### 2025-01-09 CIV 102 Lecture Notes/main.tex
Initial commit

\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf CIV 102 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Structures and Materials, Fall 2024 \hfill CIV102\\
\begin{center}
\textit{"Structural Engineering is the art and science of designing and molding structures with economy and elegance so that they can safely resist the force that they are subjected''} \\ - Prof. \textsc{Evan Bentz, 2024}
\end{center}
\vspace{10pt}
\section{Force, Moment, and Geometry} 
    \paragraph{Moment for One Force} The moment due to only one force is:
    \begin{equation}
        \mu = \mathrm{moment} = \Vec{F}\times\Vec{d_\perp}
    \end{equation}
Where:
\begin{equation*}
\begin{split}
\Vec{d_\perp} = \text{The perpendicular displacement to the center of rotation}
\end{split}
\end{equation*}
\paragraph{Centroid} The centroid of a shape with multiple geometries is calculated by: 
\begin{equation}
\bar{y} = \frac{\sum_i A_i  y_i}{\sum_i A_i}
\end{equation}
\paragraph{Parallel axis Theorem} The moment of inertia is calculated by the following
\begin{equation}
    I = \sum_i I_{i} + \sum_i A_i d_i^2
\end{equation}
Where:
\begin{equation*}
\begin{split}
d_i = \text{The displacement of $\bar{y_i}$ to $\bar{y}$}
\end{split}
\end{equation*}
\paragraph{First Moment of Area (Q)} It is expressed as.
\begin{equation}
    Q = \sum_i A_i \cdot d_i
\end{equation}
\section{Truss}
\paragraph{Area} Design area against tension/compression by:
\begin{equation}
    A \geq \frac{2F}{\sigma_y}
\end{equation}
\paragraph{Moment of Intertia} Design MOI against Euler's bucking by:
\begin{equation}
    I \geq \frac{3FL^2}{\pi^2E}
\end{equation}
\paragraph{Radius of Gyration} Design radius of gyration against slenderness ratio by:
\begin{equation}
    r \geq \frac{L}{200}
\end{equation}
\section{Beam}
\paragraph{Navier's equation} 
\begin{equation}
    \sigma = \frac{My}{I}
\end{equation}

\paragraph{Curvature Equation}
\begin{equation}
    \phi = \frac{M}{EI}
\end{equation}

\paragraph{MAT 1} The change in slope between two points is given by the first moment area theorem:
\begin{equation}
    \Delta_{AB} = \theta_B - \theta_A = \int_A^B \phi(x) \, dx
\end{equation}

\paragraph{MAT 2} The deviation of point $A$ from the tangent drawn at point $B$ is given by the second moment area theorem:
\begin{equation}
    EIt_{A/B} = \int_B^A x M(x) \, dx = \bar{x}_{AB} \int_B^A M(x) \, dx
\end{equation}
\paragraph{Shear Stress} Given by Jourawski's equation:
\begin{equation}
    \tau = \frac{VQ}{Ib}
\end{equation}
\section{Virtual Work}
\paragraph{Work} In elastic deformation, for internal energy, we have
$$
W_\text{int} = V\frac{\sigma \epsilon}{2} = \frac{P\Delta}{2}
$$
In Hookes Law, for external energy, we have:
$$
W_\text{ext} = F\Delta r
$$
\paragraph{Change of length} For a change of length, we have 
\begin{equation}
\Delta = \frac{PL}{EA}
\end{equation}
\paragraph{Deflection} to calculate bean deflection, we sum the total virtual force multiplied by extension (worked done by victual forces, virtual work):
\begin{equation}
    F^\star \Delta_{\hat{r}} = \sum_{i}P^\star_i \Delta_i
\end{equation}
\section{Vibraion}
\subsection{Free Vibration}
Since for change of length, we have:
\begin{align}
    \Delta = \frac{PL}{EA} \nonumber 
    \intertext{This could be rewritten as: }
    P = \frac{EA}{L}\Delta \nonumber
    \intertext{For stiffness $k$, this could be molded as a simple harmonic motion with:}
    k = \frac{EA}{L}
\end{align}
\paragraph{For Truss} We can use the method of virtual load to determine $\Delta_0$. 
\paragraph{For Beam} We can use the method of MAT to determine $\Delta_0$.
\paragraph{Point Load} The natural frequency is: \begin{equation}
f_n = \frac{15.76}{\sqrt{\Delta_0}}
\end{equation}
\paragraph{Uniform Load} The natural frequency is 
\begin{equation}
f_n = \frac{17.76}{\sqrt{\Delta_0}}
\end{equation}
\paragraph{Dynamic Amplification Factor} Denoted as DAF, for forced vibration at a frequency $f$, DAF is computed as:
\begin{equation}
    \text{DAF} = \frac{1}{\sqrt{(1-(\frac{f}{f_n})^2)^2+ (2\beta \frac{f}{f_n})^2}}
\end{equation}
\paragraph{Amplification} Members would be subject to loads $\text{DAF} \times P$ when the load is vibrating. 
\section{Shear and Local Buckling}
\begin{table}[ht]
\centering
\caption{Summary of plate buckling failure modes}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Failure Mode}                                           & \textbf{Failure Condition}                                                                                      & \textbf{Equation}                          \\ \hline
Buckling of the compressive flange                              & $\sigma = \dfrac{4 \pi^2 E}{12 (1 - \mu^2)} \left( \dfrac{t}{b} \right)^2$                                      & $\sigma = \dfrac{M y}{I}$                                  \\
between the webs                                                &                                                                                                                &                                                           \\ \hline
Buckling of the tips of the                                     & $\sigma = \dfrac{0.425 \pi^2 E }{12 (1 - \mu^2)} \left( \dfrac{t}{b} \right)^2$                                & $\sigma = \dfrac{M y}{I}$                                  \\
compressive flange                                              &                                                                                                                &                                                           \\ \hline
Buckling of the webs due to the                                 & $\sigma = \dfrac{6 \pi^2 E}{12 (1 - \mu^2)} \left( \dfrac{t}{b} \right)^2$                                      & $\sigma = \dfrac{M y}{I}$                                  \\
flexural stresses                                               &                                                                                                                &                                                           \\ \hline
Shear buckling of the webs                                      & $\tau = \dfrac{5 \pi^2 E}{12 (1 - \mu^2)} \left( \left( \dfrac{t}{h} \right)^2 + \left( \dfrac{t}{a} \right)^2 \right)$ & $\tau = \dfrac{V Q}{I b}$                                  \\ \hline
\end{tabular}
\end{table}
\section{Concrete}
\subsection{Material Properties}
\paragraph{Tensile Strength} The compressive strength and the tensile strength of concrete is related as follows:
\begin{equation}
    f'_t = 0.33\sqrt{f'_c}
\end{equation}
\paragraph{Young's Modulus} The compressive strength and the Young's modulus of concrete is related as follows:
\begin{equation}
    E_c = 4730\sqrt{f'_c}
\end{equation}
\begin{shaded}
\textit{Typical Values}:\\
Steel's Young's modulus is usually $E_s = 200,000$ MPa and the yield strength is $f_y = 400$ MPa.
\end{shaded}

The modular ratio $n$ is given as:
\begin{equation}
    n = \frac{E_s}{E_c}
\end{equation}

The quantity of longitudinal reinforcement $\rho$ is given as:
\begin{equation}
    \rho = \frac{A_s}{bd}
\end{equation}
where:
\begin{equation*}
\begin{split}
    A_s = \text{The area of the steel reinforcements.}\\
    b = \text{The width of the cross-sectional region of interest.}\\
    d = \text{The distance from the edge of the region of interest to the opposing reinforcements.}
\end{split}
\end{equation*}

\paragraph{$k$, Scaling Factor} of extreme compression fiber to the neutral axis is given as:
\begin{equation}
    k = \sqrt{(n\rho)^2 + 2n\rho} - n\rho
\end{equation}
\paragraph{$j$, Scaling factor } of the flexural lever is given as:
\begin{equation}
    j = 1 - \frac{1}{3}k
\end{equation}
\begin{shaded}
    \textit{Typical Values:}\\  The scaler factors are usually $k = \frac{3}{8}$ and $j = \frac{7}{8}$.
\end{shaded}

\subsection{Flexural Stress Analysis}
\paragraph{In Reinforcement} The stress is given by:
\begin{equation}
    f_s = \frac{M}{A_s j d}
\end{equation}
\paragraph{In Concrete} The stress is given by:
\begin{equation}
    f_c = \frac{k}{1-k} \cdot \frac{M}{n A_s j d}
\end{equation}

\subsection{Shear Stress Analysis}
\paragraph{Maximum Shear Stress in Cracked Concrete}
The maximum shear stress $v$ in a cracked concrete member's web is given by:
\begin{equation}
    v = \frac{V}{b_w j d}
\end{equation}
where $b_w$ is the effective web width.

\paragraph{Buckling Shear Stress}
The shear stress $v_{\text{max}}$ that causes buckling from diagonal compression is:
\begin{equation}
    v_{\text{max}} = 0.25 f'_c
\end{equation}
\paragraph{Steps for checking Shear Stress} Steps by step, if the shear strength does not pass the conditions below:
\begin{itemize}
    \item \textbf{Concrete Crushing Limit} Concrete will crush when:
    \begin{equation}
        V \geq \min(V_r, V_{\text{max}})
    \end{equation}
    where:
    \begin{equation}
    V_{\text{max}} = 0.25 f'_c b_w j d
    \end{equation}
    \item \textbf{Shear Strength of the Member}
    The shear strength $V_r$ of the member is:
    \begin{equation}
        V_r = V_c + V_s
    \end{equation}
    \item \textbf{Safety Factor for Design}
    For design purposes, select $V_r$ such that:
    \begin{equation}
        V_r = 0.5 V_c + 0.5 V_s \leq 0.5 V_{\text{max}}
    \end{equation}
\end{itemize}
Then we pass on another case  below if it involves $V_r$
\begin{itemize}
    \item \textbf{Without Reinforcement} If no shear reinforcement is present, the shear strength $V_c$ of the concrete is:
    \begin{equation}
        V_c = 230 \sqrt{f'_c} + 0.9 d b_w j d
    \end{equation}
    \item \textbf{With Minimum Reinforcement} When using shear reinforcement (stirrups), the shear strength $V_c$ of the concrete is:
    \begin{equation}
        V_c = 0.18 \sqrt{f'_c} b_w j d
    \end{equation}
    This equation is valid if:
    \begin{align}
        \frac{A_v f_y}{b_w s} 
 & \geq 0.06 \sqrt{f'_c}\\
        \Leftrightarrow s & \leq \frac{A_v f_y}{0.06 \cdot b_w \cdot \sqrt{f'_c}} \nonumber
    \end{align}
    where:
    \begin{equation*}
    \begin{split}
    A_v= \text{The effective area of stirrups; and}\\
    s = \text{Spacing between stirrups}
    \end{split}
    \end{equation*}
    \item \textbf{With Additional Reinforcement} If shear reinforcement is used, the maximum shear force $V_s$ carried is:
    \begin{equation}
        V_s = \frac{A_v f_y j d}{s} \cot(35^\circ)
    \end{equation}
\end{itemize}

\paragraph{Design Suggestions for Safety}
If a design is unsafe, consider the following:
\begin{itemize}
    \item If $V \geq 0.5 V_{\text{max}}$, resize the cross-section.
    \item If $V \geq 0.5 V_c$, add reinforcements.
    \item If $V \geq 0.5 V_c + 0.5 V_s$, adjust the spacing $s$:
    \begin{equation}
    s = \frac{0.5 A_v f_y j d \cot(35^\circ)}{V - 0.5 \times 0.18 \sqrt{f'_c} b_w j d}
    \end{equation}
\end{itemize}
\appendix
\input{appendix}
\end{document}

### 2025-01-09 ECE159 Lecture Notes/main.tex
Initial commit

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf ECE 159 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Fundamentals of Electric Circuits, Winter 2024 \hfill ECE 159\\
\vspace{10pt}
\section{Basic Concepts}
\subsection{Basic Defintions}
\begin{definition}[Voltage]
    The difference of the electric potential between two points in a circuit.
    \begin{equation}
        V \defeq \int_C \textbf{E} \cdot d\textbf{l}
    \end{equation}
    \end{definition}

    \begin{definition}[Current]
    Rate of change of charge at a particular point in the circuit.
    \begin{equation}
        I \defeq \frac{dq}{dt}
    \end{equation}
    \end{definition}

    \begin{definition}[Power]
    The rate of change of energy for an element in the circuit. Power is the product of voltage and current.
    \end{definition}

    \begin{definition}[Resistors]
    A circuit element that restricts the flow of charge.
    \end{definition}

    \begin{definition}[Ideal Sources]
    A circuit element that provides a set voltage or current regardless of what it is connected to.
    \end{definition}

    \begin{definition}[Dependent Sources]
    A circuit element that provides a voltage or current to the circuit depending on another voltage or current in the circuit.
    \end{definition}

    \begin{definition}[Passive Sign Convention]
    If a positive current flows out of the positive side of a voltage, that element is \textbf{delivering} power. Otherwise, it is \textbf{absorbing} power. All resistors absorb power.
    \end{definition}
\subsection{High School Review}
\paragraph{Electric Field} $\textbf{E}$ is created when charges are separated. It leads to a potential difference.
\paragraph{Worked done by Electric Field} Force due to the field is:
\begin{equation}
    \textbf{F} = q\textbf{E}
\end{equation}
, such that it moves charges and do work.
\paragraph{Capacitor} Energy stored in capacity is:
\begin{equation}
    U_\text{E} = \frac{1}{2}CV^2
\end{equation}
\paragraph{Energy Density} The energy [J/m$^3$] in a dielectric material is:
\begin{equation}
    u_\text{E} = \frac{1}{2}\epsilon_r \epsilon_0 E^2
\end{equation}
where:
\begin{center}
    $\epsilon_0$ = Permittivity of free space \\
    $\epsilon_r$ = Relative Permittivity
\end{center}
\paragraph{Magnetic field} \textbf{B} is created when charges move.
\paragraph{An electromotive force (emf)} can be induced when the magnetic flux
through a loop is changing with time. Induced emf does the same thing as the potential difference of a
battery. 
\paragraph{Induced EMF} An inductor can store energy in a ``magnetic form"
\begin{equation}
    U_B = \frac{1}{2} L i^2 \quad 
\end{equation}
\paragraph{Energy Density}
\begin{equation}
    u_B = \frac{1}{2 \mu_r \mu_0} B^2
\end{equation}
\subsection{Voltage and Current}
\paragraph{} Now instead of dealing with electric and magnetic vector fields, we can
use the scalar quantities of voltage and current:
\begin{align}
    \text{Voltage} &\defeq \text{Difference in Electric Potential}\nonumber \\ 
    \Delta V &= V_A - V_B \nonumber \\
    &= \int_B^A \textbf{E} \cdot dl \nonumber\\
    &= \frac{\Delta W}{q}
\end{align}

\paragraph{Electromotive Force (emf) / Voltage} Work is performed by 
the electric field to move a charge from one point to another. 
The work done per unit charge is the voltage.
\begin{equation}
    \text{Voltage} = \frac{\Delta W}{q}
\end{equation}
\subsection{Power and Energy}
\paragraph{Power} To relate power to  voltage and current, we can use the following equation:
\begin{equation}
    P \defeq \frac{dW}{dt} = \frac{dW}{dq} \cdot \frac{dq}{dt} = V \cdot I
\end{equation}
\paragraph{Sign} If $P = V \cdot I > 0$, the element is delivering power. If $P = V \cdot I < 0$, the element is absorbing power.
\paragraph{Energy}  the energy absorbed or supplied by an element
from time $t_0$ to time $t_1$ is:
\begin{equation}
    W = \int_{t_0}^{t_1} P(t) dt
\end{equation}
\subsection{Circuit Elements}
\paragraph{Active Elements} Active elements are sources of energy. They can deliver power to the circuit. While passive elements are elements that can only absorb power.
\paragraph{Ideal Sources} Ideal sources are sources that provide a set voltage or current regardless of what it is connected to. We often assume that sources are ideal.
\paragraph{Independent Sources} Independent sources are sources that provide a set voltage or current regardless of what it is connected to.
\paragraph{Dependent Sources} Dependent sources are sources that provide a voltage or current to the circuit depending on another voltage or current in the circuit.
\paragraph{Resistor} A resistor is a circuit element that restricts the flow of charge. The voltage across a resistor is proportional to the current through it.

\subsection{Resistance}
\paragraph{Ohm's Law} Ohm's Law states that the current through a resistor is proportional to the voltage across it.
\begin{equation}
    V = IR
\end{equation}
\paragraph{Resistivity} The resistance of a material is proportional to the length of the material and inversely proportional to the cross-sectional area of the material.
\begin{equation}
    R = \rho \frac{l}{A}
\end{equation}
\paragraph{Short and Open Circuits} A short circuit is a circuit with no resistance, while an open circuit is a circuit with infinite resistance.
\paragraph{Conductance} The reciprocal of resistance is conductance. The unit of conductance is mho (ohm spelled backwards), \rotatebox[origin=c]{180}{$\Omega$}.
\begin{equation}
    G = \frac{1}{R}
\end{equation}
\paragraph{Power in Resistors} The power dissipated in a resistor is given by:
\begin{equation}
    P = I^2R = \frac{V^2}{R}
\end{equation}

\section{DC Circuits}
\paragraph{}

\end{document}

### 2025-01-09 ESC195 Lecture Notes/main.tex
Initial commit

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf ESC 195 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Caculus II, Winter 2024 \hfill ESC 195\\
\section{More on Integrals}
\subsection{Riemann Sum - Non-Uniform Petition}
\begin{example}
    Given the following definite integral:
    $$\int^2_0 \sqrt{x} dx$$, we cannot evaluate its Riemann sum with uniforms partition, since the series of root cannot be easily evaluated. 
\end{example}

The definite integral of $\sqrt{x}$ from $0$ to $2$ using a Riemann sum with a non-uniform partition is given by:

$$
\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i
$$

where:
\begin{itemize}
    \item $x_0 = 0, \, x_n = 2$,
    \item $x_i = i^2 \cdot \frac{2}{n^2}$ for $i = 0, 1, 2, \dots, n$,
    \item $\Delta x_i = x_i - x_{i-1} = \frac{2}{n^2} \cdot (2i - 1)$.
\end{itemize}

The Riemann sum becomes:
$$
S_n = \sum_{i=1}^n \sqrt{i^2 \cdot \frac{2}{n^2}} \cdot \frac{2}{n^2} \cdot (2i - 1).
$$

Simplifying further:
$$
S_n = \sum_{i=1}^n \sqrt{\frac{2i^2}{n^2}} \cdot \frac{2}{n^2} \cdot (2i - 1).
$$

Taking the limit as $n \to \infty$, the sum converges to the exact value of the integral using the series of squares. 
$$
\int_{0}^{2} \sqrt{x} \, dx = \frac{4\sqrt{2}}{3}.
$$
\paragraph{Condition} $n \to \infty$ Ensures $\Delta x_i \to 0$
\paragraph{} As $n \to \infty$, the partition points $x_i$ become increasingly dense. This ensures that the partition becomes infinitely fine.

\subsection{Integration By Parts}
Using the product rule:
$$
\frac{d}{dx} \big[ u(x) v(x) \big] = u'(x)v(x) + u(x)v'(x),
$$
integrating both sides with respect to $x$ gives:
$$
u(x)v(x) = \int u'(x) v(x) \, dx + \int u(x) v'(x) \, dx.
$$
Rearranging this:
$$
\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.
$$

\paragraph{Integration by parts formula}
\begin{equation}
\int u \, dv = uv - \int v \, du.
\end{equation}
\begin{example}
We want to solve the integral $$ \int x e^{2x} \, dx $$ using integration by parts.

Let:
$$ u = x, \quad dv = e^{2x} \, dx. $$

Then, we compute the derivatives and integrals:
$$ du = dx, \quad v = \frac{e^{2x}}{2}. $$

Now, apply the integration by parts formula:
$$ \int u \, dv = uv - \int v \, du. $$

Substituting in the values:
$$ \int x e^{2x} \, dx = x \cdot \frac{e^{2x}}{2} - \int \frac{e^{2x}}{2} \, dx. $$

Next, compute the remaining integral:
$$ \int \frac{e^{2x}}{2} \, dx = \frac{e^{2x}}{4}. $$

Thus, the result is:
$$ \int x e^{2x} \, dx = \frac{x e^{2x}}{2} - \frac{e^{2x}}{4} + C. $$

\end{example}

\begin{example}
We want to solve $$ \int x^2 \sin(2x) \, dx $$ using double integration by parts.

First, let:
$$ u = x^2, \quad dv = \sin(2x) \, dx. $$

Then:
$$ du = 2x \, dx, \quad v = -\frac{1}{2} \cos(2x). $$

Using the IBP formula:
$$ \int u \, dv = uv - \int v \, du, $$

we get:
$$ \int x^2 \sin(2x) \, dx = -\frac{x^2}{2} \cos(2x) + \int x \cos(2x) \, dx. $$

Now, apply IBP again to \( \int x \cos(2x) \, dx \), let:
$$ u = x, \quad dv = \cos(2x) \, dx. $$

Then:
$$ du = dx, \quad v = \frac{1}{2} \sin(2x). $$

Using the IBP formula again:
$$ \int x \cos(2x) \, dx = \frac{x}{2} \sin(2x) - \int \frac{1}{2} \sin(2x) \, dx, $$

and solving the remaining integral:
$$ \int \frac{1}{2} \sin(2x) \, dx = -\frac{1}{4} \cos(2x). $$

Thus, the final result is:
$$ \int x^2 \sin(2x) \, dx = -\frac{x^2}{2} \cos(2x) + \frac{x}{2} \sin(2x) + \frac{1}{4} \cos(2x) + C. $$

\end{example}
\paragraph{} 
\end{document}

### 2025-01-09 MAT185 Lecture Notes/main.tex
Initial commit

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MAT 185 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Linear Algebra, Winter 2024 \hfill MAT 185\\
\section{Vector Space}
\begin{definition}[Vector Space in $\mathbb{R}$]
    A vectors space,${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
    \begin{enumerate}
        \item \textbf{Closure Under Addition} $\textbf{x},\textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$
        \item \textbf{Order of Addition} $\textbf{x},\textbf{y}, \textbf{z} \in V \implies (\textbf{x} + \textbf{y}) + \textbf{z} = \textbf{x} + (\textbf{y} + \textbf{z}) \in V$
        \item \textbf{Existence of Additional Identity} $\exists \textbf{0} \in V$ s.t. $\textbf{x} \in V \implies \textbf{x} + \textbf{0} = \textbf{x}$
        \item \textbf{Existence of Additional Inverse} $\forall \textbf{x} \in V \exists \textbf{-x} \in V$ s.t. $\textbf{x} + \textbf{-x} = \textbf{0}$
        \item \textbf{Closure under Scalar Multiplication} $\forall \textbf{x} \in V$ and $\forall \alpha \in \mathbb{R} \implies \alpha\textbf{x} \in V$
        \item \textbf{Order of Scalar Multiplication} $\forall \textbf{x} \in V$ and $\alpha, \beta \in \mathbb{R}$, $(\alpha\beta)\textbf{x} \implies \alpha(\beta\textbf{x})$
        \item \textbf{Distributive Scalar Multiplication} $\forall \textbf{x} \in V$ and $\alpha, \beta \in \mathbb{R}$, $(\alpha + \beta)\textbf{x} = \alpha\textbf{x} + \beta\textbf{x}$
        \item \textbf{Existence of Multiplicative Identity} $\forall \textbf{x} \in V$, $1\textbf{x} = \textbf{x}$
    \end{enumerate}
\paragraph{Note} It could be shown that the axiom imply the commutativity of in addition, namely $\forall \textbf{x} \implies \textbf{y} \in \mathbb{R}$, $\textbf{x} + \textbf{y} = \textbf{y} + \textbf{x}$
\end{definition}
\end{document}

### 2025-01-09 MSE160 Lecture Notes/main.tex
Initial commit

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MSE 160 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Molecules and Materials, Winter 2024 \hfill MSE 160\\
\begin{center}
\textit{"In this class we are mostly understanding solids''} \\ - Prof. \textsc{Scott Ramsay}
\end{center}
\vspace{10pt}
\section{Mechanical Behavior}
\paragraph{Classes of Materials} In this class, we look at three classes of materials (non-exhaustive):
\begin{itemize}
    \item \textbf{Metal} held together with metallic bonds, typically \textbf{ductile} and \textbf{conductive}.
    \item \textbf{Ceramics}  (often metal oxides [excp: diamond]) held together via covalent \& ionic bonds, typically \textbf{brittle} and \textbf{insulating}.
    \item \textbf{Polymers} Molecules (often hydrocarbons) typically \textbf{ductile} and \textbf{insulating}
\end{itemize}
\paragraph{Engineering Stress} We know that:
\begin{equation}
    \sigma = \frac{F}{A_0}
\end{equation}
\paragraph{Engineering Strein} Also:
\begin{equation}
    \epsilon = \frac{\Delta l}{l_0}
\end{equation}
\paragraph{Young's Moduclus} For elastic deformation, $E$, is given, by Hooke's Law, as follows:
\begin{equation}
    \sigma = E \epsilon
\end{equation}
\paragraph{Tensile Test} We apply force as to the ends of a dogbone-sample, with $l_0$ being the gauge length and $A_0$ being the area of the cross-section at the middle. 
\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}

### 2025-01-10 ECE159 Lecture Notes/main.tex
Add Force-Separation Curve and update lecture notes

@@
 \end{equation}
-
+\paragraph{Resistors in Series} If resistors are placed in series with each other, their current must be the same, The equivalent resistance of resistors in series is the sum of the resistances.
+\begin{equation}
+    R_{\text{eq}} = R_1 + R_2 + \dots + R_n
+\end{equation}
+\paragraph{Resistors in Parallel} If resistors are placed in parallel with each other, their voltage must be the same.
+The equivalent resistance of resistors in parallel is the reciprocal of the sum of the reciprocals of the resistances.
+\begin{equation}
+    \frac{1}{R_{\text{eq}}} = \frac{1}{R_1} + \frac{1}{R_2} + \dots + \frac{1}{R_n}
+\end{equation}
+\paragraph{Comination of Series and Parallel Resistors} To find the equivalent resistance of a combination of series and parallel resistors, first find the equivalent resistance of the resistors in series, then find the equivalent resistance of the resistors in parallel.
+\subsection{Node, Branch, and Loop}
+\paragraph{} Three important ideas in electric circuit analysis are:
+\begin{enumerate}
+    \item \textbf{Branch}: A branch is an element in an electric circuit, such as a source or a resistor.
+    \item \textbf{Node}: A point of connection of two or more branches within an electric circuit.
+    \item \textbf{Loop}: Any closed path in an electric circuit. To form a loop, one must start and end at the same point and never pass through the same node twice.
+\end{enumerate}
 \section{DC Circuits}

### 2025-01-10 ESC195 Lecture Notes/main.tex
Add Force-Separation Curve and update lecture notes

@@
 \end{example}
-\paragraph{} 
+\subsection{Trigonometric Integrals}
+\paragraph{Case I} This is the case I trigonometric integrals, the strategy is using the pythagorean identity.
+\paragraph{} We want to solve the class of integrals:
+\begin{equation}
+ \int \sin^n(x) \cos^m(x) \, dx
+\end{equation}
+, where $m$ or $n$ is odd.
+\begin{example}
+We want to solve the integral:
+$$ I = \int \sin^3(x) \cos^2(x) dx $$
+We use the identity: $ \sin^2(x) = 1 - \cos^2(x)$. Thus, the integral becomes:
+\begin{align*}
+    I &= \int \sin(x)(1 - \cos^2(x)) \cos^2(x) \, dx \\
+    &= \int (\cos^2(x) \sin(x) - \cos^4(x) \sin(x)) \, dx
+\end{align*} 
+which is now easily solvable with substution $u=\cos(x)$.
+\end{example}
+\paragraph{Case II} The follwoing is generally solvable via case I and case III below.
+In general, we solve the integral by reducing the power of the trigonometric functions to arrive at a solvable integral Case I or III.
+\paragraph{} We want to solve the class of integrals:
+\begin{equation} \int \sin^n(x) \cos^m(x) \, dx \end{equation}
+, where $m$ and $n$ is even.
+\begin{example}
+We want to solve the integral:
+$$ I = \int \sin^2(x) \cos^4(x) \, dx $$
+We can apply the double angle formulas:
+\begin{align}
+    \sin(x)\cos(x) &= \frac{1}{2} \sin(2x) \\
+    \sin^2(x) &= \frac{1 - \cos(2x)}{2} \\
+    \cos^2(x) &= \frac{1 + \cos(2x)}{2}
+\end{align}
+Thus, the integral becomes:
+$$ I = \frac{1}{8} \int \sin^2(2x) \, dx + \frac{1}{8} \int \cos(2x)\sin^2(2x) \, dx $$
+\end{example}
+\paragraph{Case III} This is the case III trigonometric integrals, the strategy is using the reduction formula via IBP.
+\paragraph{Reduction Formula} We can solve integrals by reducing the power of the trigonometric functions. These are done using IBP and trigonometric identities.
+\paragraph{} We want to solve the classes of integrals:
+\begin{equation} 
+    \int \sin^n(x) \, dx, \quad \int \cos^n(x) \, dx 
+\end{equation}
+, where $n$ is a positive integer. For demonstartion, We can obtain the reduction formula of $\sin^n(x)$ via IBP.
+\begin{align*}
+    I_n &= \int \sin^n(x) \, dx \\
+    &= \int \sin^{n-1}(x) \sin(x) \, dx \\
+    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \cos^2(x) \sin^{n-2}(x) \, dx \\
+    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} I_{n-2}
+\end{align*}
+\paragraph{Case IV} The following is generally solvable via simple trigonometric integrals. In general, we solve the integral by applying the angle sum formulas.
+\paragraph{} We want to solve the classes of integrals:
+\begin{equation}
+    \int \sin(mx) \cos(nx) \, dx, \quad \int \sin(mx) \sin(nx) \, dx, \quad \int \cos(mx) \cos(nx) \, dx 
+\end{equation}
+We could apply the angle sum formulas:
+\begin{align}
+    \sin(mx)\cos(nx) &= \frac{1}{2} \left[ \sin((m+n)x) + \sin((m-n)x) \right] \\
+    \sin(mx)\sin(nx) &= \frac{1}{2} \left[ \cos((m-n)x) - \cos((m+n)x) \right] \\
+    \cos(mx)\cos(nx) &= \frac{1}{2} \left[ \cos((m-n)x) + \cos((m+n)x) \right]
+\end{align}
+\paragraph{Case V} The following is generally solvable via the following trigonometric identities listed below, which convert it into a reduction formula.
+\begin{align}
+    \tan^2(x) &= \sec^2(x) - 1 \label{eq:tan2} \\
+    \cot^2(x) &= \csc^2(x) - 1 \label{eq:cot2}  \\
+    \frac{d}{dx} \tan(x) &= \sec^2(x)
+\end{align}
+\paragraph{} We want to solve the classes of integral:
+\begin{equation}
+    \int \tan^n(x) \, dx, \quad \int \cot^n(x) \, dx 
+\end{equation}
+We know that $tan^2(x) = \sec^2(x) - 1$. Thus, we can solve the integral by reducing the power of the tangent function.
+\paragraph{Case VI} The following is generally solvable via the trigonometric identities (\ref{eq:tan2}) and (\ref{eq:cot2}).
+\paragraph{} We want to solve the classes of integral:
+\begin{equation} 
+    \int \tan^m(x) \sec^n(x) \, dx, \quad \int \cot^m(x) \csc^n(x) \, dx 
+\end{equation}
+We can solve the integral by reducing the power of the converted trigonometric functions using (\ref{eq:tan2}) and (\ref{eq:cot2}).
+\paragraph{Case VII} The following is generally solvable via the trigonometric identities (\ref{eq:tan2}) and (\ref{eq:cot2}).
+\paragraph{} We want to solve the classes of integral:
+\begin{equation} \int \tan^m(x) \sec^n(x) \, dx, \quad \int \cot^m(x) \csc^n(x) \, dx \end{equation}
+We can solve the integral by converting between the tangent and secant functions using the trigonometric identities.
+\paragraph{Case VIII} This is the case VIII integrals, the strategy is using the trigonometric substitution.
+\paragraph{Trigonometric Substitution} We can solve integrals by substituting the trigonometric functions with other trigonometric functions.
+\begin{example}
+We want to solve the integral:
+$$ \int \frac{dx}{\sqrt{1 - x^2}} $$
+\end{example}
+We can substitute $x = \sin(\theta)$, then $dx = \cos(\theta) \, d\theta$. The integral becomes:
+$$ \int \frac{\cos(\theta) \, d\theta}{\sqrt{1 - \sin^2(\theta)}} = \int \frac{\cos(\theta) \, d\theta}{\cos(\theta)} = \int d\theta = \theta + C. $$
+\paragraph{} In general, we can use the following substitutions:
+\begin{itemize}
+    \item $x = a\sin(\theta)$ for $\sqrt{a^2 - x^2}$,
+    \item $x = a\tan(\theta)$ for $\sqrt{a^2 + x^2}$,
+    \item $x = a\sec(\theta)$ for $\sqrt{x^2 - a^2}$.
+\end{itemize}
+\paragraph{Summary} We can solve trigonometric integrals by using the following strategies:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|l|}
+    \hline
+    \textbf{Case} & \textbf{Strategy and General Form} \\ \hline
+    I   & Use \(\sin^2(x) + \cos^2(x) = 1\); simplify using substitution. \\
+        & General Form: \(\int \sin^n(x) \cos^m(x) dx\), where \(m\) or \(n\) is odd. \\ \hline
+    II  & Convert to Case I and Case III using double angle formulas. \\
+        & General Form: \(\int \sin^n(x) \cos^m(x) dx\), where both \(m\) and \(n\) are even. \\ \hline
+    III & Apply reduction formulas derived via integration by parts. \\
+        & General Form: \(\int \sin^n(x) dx\) or \(\int \cos^n(x) dx\). \\ \hline
+    IV  & Use angle sum formulas to simplify. \\
+        & General Form: \(\int \sin(mx) \cos(nx) dx\). \\ \hline
+    V   & Reduce tangent/cotangent powers using \(\tan^2(x) = \sec^2(x) - 1\) and substitution. \\
+        & General Form: \(\int \tan^n(x) dx\) or \(\int \cot^n(x) dx\). \\ \hline
+    VI  & Convert to Case V via Pythagorean identities. \\
+        & General Form: \(\int \tan^m(x) \sec^n(x) dx\) or \(\int \cot^m(x) \csc^n(x) dx\). \\ \hline
+    VII & Convert between tangent and secant functions for simplification. \\
+        & General Form: \(\int \tan^m(x) \sec^n(x) dx\). \\ \hline
+    VIII & Use trigonometric substitution: \(x = a \sin(\theta), a \tan(\theta), a \sec(\theta)\). \\
+         & General Form: \(\int \frac{dx}{\sqrt{a^2 - x^2}}, \int \frac{dx}{\sqrt{a^2 + x^2}}, \int \frac{dx}{\sqrt{x^2 - a^2}}\). \\ \hline
+    \end{tabular}
+    \caption{Strategies and General Forms for Case I-VIII Integrals}
+\end{table}
+
+\section{Hyperbolic Trigonometric Functions}
+\begin{definition}[Hyperbolic Sine]
+    The hyperbolic sine function is defined as:
+    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
+\end{definition}
+\begin{definition}[Hyperbolic Cosine]
+    The hyperbolic cosine function is defined as:
+    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
+\end{definition}
+\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
+\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
+\begin{align}
+    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
+    \frac{d}{dx} \cosh(x) &= \sinh(x).
+\end{align}
+\paragraph{Identities} The hyperbolic functions satisfy the following identities:
+\begin{align}
+    \cosh^2(x) - \sinh^2(x) &= 1, \\
+    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
+    \sinh(2x) &= 2\sinh(x)\cosh(x).
+\end{align}
+\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$.
+Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g.
+$x = \cosh(t)$, $y = \sinh(t)$).
+\paragraph{Area} The area of a sector of the hyperbola is given by:
+\begin{equation}
+    A = t/2,
+\end{equation} 
+where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
+\begin{figure}[h!]
+    \begin{center}
+        \begin{tikzpicture}[scale=0.8]
+            % Draw the axes
+            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
+            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};
+
+            % Draw ticks on the axes
+            \foreach \x in {-4, -2, 2, 4}
+                \draw (\x, 0.1) -- (\x, -0.1) node[below] {\x};
+            \foreach \y in {-4, -2, 2, 4}
+                \draw (0.1, \y) -- (-0.1, \y) node[left] {\y};
+
+            % Draw the hyperbola branches
+            \draw[thick, red, domain=-2:2, samples=100] plot ({cosh(\x)}, {sinh(\x)});
+            \draw[thick, red, domain=-2:2, samples=100] plot ({-cosh(\x)}, {sinh(\x)});
+
+            % Draw the asymptotes
+            \draw[red, dashed] (-5, -5) -- (5, 5);
+            \draw[red, dashed] (-5, 5) -- (5, -5);
+
+            % Add a point on the hyperbola
+            \coordinate (P) at ({cosh(1)}, {sinh(1)});
+            \draw[fill] (P) circle (0.05) node[above right] {$(\cosh(t), \sinh(t))$};
+
+            % Connect the labeled point to the origin
+            \draw[thick, blue] (0, 0) -- (P);
+
+            % Fill the sector area between hyperbola, line, and x-axis
+            \fill[blue, opacity=0.3, domain=0:1, variable=\t]
+                (0, 0) -- plot ({cosh(\t)}, {sinh(\t)}) -- cycle;
+        \end{tikzpicture}
+    \end{center}
+    \caption{Sector of the hyperbola}
+    \centering
+\end{figure}
+\paragraph{Catenary} The hyperbolic cosine function describes the shape of a hanging chain or cable. The catenary is the curve formed by a chain hanging from two points. It is given by the equation:
+\begin{equation}
+    a \cosh\left(\frac{x}{a}\right),
+\end{equation}
+\begin{definition}[Hyperbolic Tangent]
+    The hyperbolic tangent function is defined as:
+    \begin{equation}
+        \tanh(x) = \frac{\sinh(x)}{\cosh(x)} = \frac{e^x - e^{-x}}{e^x + e^{-x}}. 
+    \end{equation}
+\end{definition}
+\paragraph{Derivative} The derivative of the hyperbolic tangent function resembles the derivative of the regular tangent function:
+\begin{equation} \frac{d}{dx} \tanh(x) = \text{sech}^2(x). \end{equation}
+\paragraph{Identities} The hyperbolic tangent function satisfies the following identities:
+\begin{align}
+    \tanh(x) &= \frac{\sinh(x)}{\cosh(x)}, \\
+    \text{sech}^2(x) &= 1 - \tanh^2(x).
+\end{align}
+\paragraph{Secant, Cosecant, Cotangent} The hyperbolic secant, cosecant, and cotangent functions are defined similarly to the regular secant, cosecant, and cotangent functions.
+They are the \textbf{reciprocal} of the hyperbolic cosine, sine, and tangent functions, respectively.
+\paragraph{Inverse Hyperbolic Functions} The inverse hyperbolic functions are defined as the inverse of the hyperbolic functions.
+They are denoted by $\sinh^{-1}(x)$, $\cosh^{-1}(x)$, $\tanh^{-1}(x)$, etc.
+\begin{align}
+    \text{arsinh } x &= \ln\left(x + \sqrt{x^2 + 1}\right) \\[10pt]
+    \text{arcosh } x &= \ln\left(x + \sqrt{x^2 - 1}\right) \\[10pt]
+    \text{artanh } x &= \frac{1}{2} \ln\left(\frac{1 + x}{1 - x}\right) \\[10pt]
+    \text{arcsch } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} + 1}\right) \\[10pt]
+    \text{arsech } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} - 1}\right) \\[10pt]
+    \text{arcoth } x &= \frac{1}{2} \ln\left(\frac{x + 1}{x - 1}\right)
+\end{align}
+
 \end{document}

### 2025-01-10 MAT185 Lecture Notes/main.tex
Add Force-Separation Curve and update lecture notes

@@
     \begin{enumerate}
-        \item \textbf{Closure Under Addition} $\textbf{x},\textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$
-        \item \textbf{Order of Addition} $\textbf{x},\textbf{y}, \textbf{z} \in V \implies (\textbf{x} + \textbf{y}) + \textbf{z} = \textbf{x} + (\textbf{y} + \textbf{z}) \in V$
-        \item \textbf{Existence of Additional Identity} $\exists \textbf{0} \in V$ s.t. $\textbf{x} \in V \implies \textbf{x} + \textbf{0} = \textbf{x}$
-        \item \textbf{Existence of Additional Inverse} $\forall \textbf{x} \in V \exists \textbf{-x} \in V$ s.t. $\textbf{x} + \textbf{-x} = \textbf{0}$
-        \item \textbf{Closure under Scalar Multiplication} $\forall \textbf{x} \in V$ and $\forall \alpha \in \mathbb{R} \implies \alpha\textbf{x} \in V$
-        \item \textbf{Order of Scalar Multiplication} $\forall \textbf{x} \in V$ and $\alpha, \beta \in \mathbb{R}$, $(\alpha\beta)\textbf{x} \implies \alpha(\beta\textbf{x})$
-        \item \textbf{Distributive Scalar Multiplication} $\forall \textbf{x} \in V$ and $\alpha, \beta \in \mathbb{R}$, $(\alpha + \beta)\textbf{x} = \alpha\textbf{x} + \beta\textbf{x}$
-        \item \textbf{Existence of Multiplicative Identity} $\forall \textbf{x} \in V$, $1\textbf{x} = \textbf{x}$
+        \item \textbf{Addition Axioms}
+        \begin{enumerate}
+            \item \textbf{Closure Under Addition:} \\
+            $\textbf{x}, \textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$
+
+            \item \textbf{Associativity of Addition:} \\
+            $\textbf{x}, \textbf{y}, \textbf{z} \in V \implies (\textbf{x} + \textbf{y}) + \textbf{z} = \textbf{x} + (\textbf{y} + \textbf{z})$
+
+            \item \textbf{Existence of Additive Identity:} \\
+            $\exists \, \textbf{0} \in V$ such that $\textbf{x} \in V \implies \textbf{x} + \textbf{0} = \textbf{x}$
+
+            \item \textbf{Existence of Additive Inverse:} \\
+            $\forall \, \textbf{x} \in V, \, \exists \, \textbf{-x} \in V$ such that $\textbf{x} + \textbf{-x} = \textbf{0}$
         \end{enumerate}
+
+        \item \textbf{Scalar Multiplication Axioms}
+        \begin{enumerate}
+            \item \textbf{Closure Under Scalar Multiplication:} \\
+            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha \in \mathbb{R}, \, \alpha \textbf{x} \in V$
+
+            \item \textbf{Associativity of Scalar Multiplication:} \\
+            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha, \beta \in \mathbb{R}, \, (\alpha \beta) \textbf{x} = \alpha (\beta \textbf{x})$
+
+            \item \textbf{Distributive Property of Scalar Multiplication:} \\
+            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha, \beta \in \mathbb{R}, \, (\alpha + \beta) \textbf{x} = \alpha \textbf{x} + \beta \textbf{x}$
+
+            \item \textbf{Existence of Multiplicative Identity:} \\
+            $\forall \, \textbf{x} \in V, \, 1 \textbf{x} = \textbf{x}$
+        \end{enumerate}
+    \end{enumerate}
+
     \paragraph{Note} It could be shown that the axiom imply the commutativity of in addition, namely $\forall \textbf{x} \implies \textbf{y} \in \mathbb{R}$, $\textbf{x} + \textbf{y} = \textbf{y} + \textbf{x}$
 \end{definition}
+\paragraph{Example} $\mathbb{R}^n$ is a vector space over $\mathbb{R}$, where $\textbf{x} = (x_1, x_2, \ldots, x_n)$ and $\textbf{y} = (y_1, y_2, \ldots, y_n)$, then $\textbf{x} + \textbf{y} = (x_1 + y_1, x_2 + y_2, \ldots, x_n + y_n)$ and $\alpha\textbf{x} = (\alpha x_1, \alpha x_2, \ldots, \alpha x_n)$
+\begin{theorem}[Cancelation, Part 1]
+    Let $V$ be a vector space and $\textbf{x}, \textbf{y}, \textbf{z} \in V$, then:
+    \begin{align*}
+        \textbf{x} + \textbf{z} &= \textbf{y} + \textbf{z} \\
+        \textbf{x} &= \textbf{y}
+    \end{align*} 
+\end{theorem}
+\begin{proof}
+    \begin{align*}
+        \textbf{x} + \textbf{z} &= \textbf{y} + \textbf{z} \\
+        \intertext{From additive inverse we know that \textbf{-z} exists}
+        (\textbf{x} + \textbf{z}) + \textbf{-z} &= (\textbf{y} + \textbf{z}) + \textbf{-z} \\
+        \intertext{By order of addition we have: }
+        \textbf{x} + (\textbf{z} + \textbf{-z}) &= \textbf{y} + (\textbf{z} + \textbf{-z}) \\
+        \textbf{x} + \textbf{0} &= \textbf{y} + \textbf{0} \\
+        \intertext{By additive identity we have: }
+        \textbf{x} &= \textbf{y}
+    \end{align*}
+\end{proof}
+\begin{theorem}[Cancelation, Part 2]
+    Let $V$ be a vector space and $\textbf{x}, \textbf{y}, \textbf{z} \in V$, then:
+    \begin{align*}
+        \textbf{z} + \textbf{x} &= \textbf{z} + \textbf{y} \\
+        \textbf{x} &= \textbf{y}
+    \end{align*} 
+\end{theorem}
+To proof that, it would require the following propostion:
+\begin{lemma}
+    Let $V$ be a vector space and $\textbf{z} \in V$, then $\textbf{-z} + \textbf{z} = 0 $
+\end{lemma}
+\begin{proof}
+    \begin{align*}
+        \intertext{We know:}
+        \textbf{-z} + \textbf{z} &= (\textbf{-z} + \textbf{z}) + 0\\
+        &= (\textbf{-z} + \textbf{z}) + (\textbf{-z} + \textbf{-(-z)}) \\
+        \intertext{By the order of additon, }
+        &= \textbf{-z} + (\textbf{z} + (\textbf{-z} + \textbf{-(-z)})) \\
+        \intertext{Again, by the order of additon, }
+        &= \textbf{-z} + ((\textbf{z} + \textbf{-z}) + \textbf{-(-z)}) \\
+        &= \textbf{-z} + \textbf{-(-z)} \\
+        &= 0
+    \end{align*}
+\end{proof}
+Now, to prove the part 2 of the Cancelation Theorem:
+\begin{proof}
+    \begin{align*}
+        \textbf{z} + \textbf{x} &= \textbf{z} + \textbf{y} \\
+        \intertext{From additive inverse we know that \textbf{-z} exists}
+        \textbf{-z} + (\textbf{z} + \textbf{x}) &= \textbf{-z} + (\textbf{z} + \textbf{y})\\
+        (\textbf{-z} + \textbf{z}) + \textbf{x} &= (\textbf{-z} + \textbf{z}) + \textbf{y}\\
+        \intertext{From above, we have $\textbf{-z} + \textbf{z} = 0$}
+        0 + \textbf{x} &= 0 + \textbf{y} \\
+        \textbf{x} &= \textbf{y}
+    \end{align*}
+\end{proof}
+\begin{lemma}[Inverse of an inverse]
+    Let $V$ be a vector space and $\textbf{x} \in V$, then:
+    \begin{align*}
+        \textbf{-(-x)} &= \textbf{x}
+    \end{align*}
+\end{lemma}
+\begin{proof}
+    Assume $0$, $0^\star$ are the additive identity of $V$ and $\textbf{-x}$, $\textbf{-x}^\star$ are the additive inverse of $\textbf{x}$. We have:
+    \begin{equation*}
+        u+0 = u + 0^\star
+    \end{equation*}
+    By Cancelation Theorem, we have $0 = 0^\star$. Thus, the additive identity is unique. Now, we have:
+    \begin{align*}
+        \textbf{x} + \textbf{-x} &= 0 \\
+        \textbf{x} + \textbf{-x}^\star &= 0
+        \intertext{Thus, by the uniqueness of the additive inverse, we have:}
+        \textbf{x} + \textbf{-x} &= \textbf{x} + \textbf{-x}^\star
+        \intertext{By the Cancelation Theorem, we have:}
+        \textbf{-x} &= \textbf{-x}^\star
+    \end{align*}
+    Now, we have that the additive inverse is unique. Thus, $\textbf{-(-x)}$ must be unique and nessarily $\textbf{x}$. 
+\end{proof}
+\paragraph{Results from above}
+\begin{enumerate}
+    \item The additive identity is unique
+    \item The additive inverse is unique
+\end{enumerate}
+\begin{definition}[Subtraction]
+    Let $V$ be a vector space and $\textbf{x}, \textbf{y} \in V$, then:
+    \begin{align*}
+        \textbf{x} - \textbf{y} &\defeq \textbf{x} + \textbf{-y}
+    \end{align*}
+\end{definition}
+\begin{theorem}[Commutativity of Addition]
+    Let $V$ be a vector space and $\textbf{x}, \textbf{y} \in V$, then:
+    \begin{align*}
+        \textbf{x} + \textbf{y} &= \textbf{y} + \textbf{x}
+    \end{align*} 
+\end{theorem}
+\begin{proof}
+    \begin{align*}
+        \textbf{x} + \textbf{y} &=  
+    \end{align*}
+\end{proof}   
 \end{document}

### 2025-01-10 MSE160 Lecture Notes/main.tex
Add Force-Separation Curve and update lecture notes

@@
 \end{itemize}
-\paragraph{Engineering Stress} We know that:
+\paragraph{Engineering Stress} For normal stress, we know that:
 \begin{equation}
@@
 \paragraph{Tensile Test} We apply force as to the ends of a dogbone-sample, with $l_0$ being the gauge length and $A_0$ being the area of the cross-section at the middle. 
+\paragraph{Tensile Strein} Maximum tensile strain on the engineeging stress-strain curve.
+\subsection{Understanding Elastic Properties in terms of Atomic Configuration}
+\paragraph{Atomic Configuration} We can understand the elastic properties of a material by looking at the atomic configuration.
+Skemetically, we can represent the atomic configuration as a spring system:
+\begin{enumerate}
+    \item \textbf{Intial - Before Loading} Atoms are in equilibrium, with the interatomic forces being balanced.
+    \item \textbf{Loading} We apply a force to the material, causing the atoms to move from their equilibrium positions. The bond stretches and the atoms move further apart.
+    \item \textbf{Unloading} We remove the force, causing the atoms to return to their equilibrium positions. 
+\end{enumerate}
+\paragraph{Atom Positions} Elastic modulus is dependent on the atomic interatomic bonding force. Thus, The elastic modulus is prootionaly to the slope of the interatomic force-seperation curve.
+\paragraph{Force-Seperation Curve} The force-seperation curve is a plot of the force between two atoms as a function of the distance between them.
+The slope of the curve is proportional to the elastic modulus near the equilibrium position.
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+        \begin{axis}[
+            xlabel={Separation Distance (d)},
+            ylabel={Force (F)},
+            grid=major,
+            width=10cm,
+            height=6cm,
+            domain=-0.3:10,
+            samples=100,
+            xtick = \empty,
+            ytick = {0}
+        ]
+        \addplot[blue, thick] {exp(-x/2) - exp(-x)};
+        \node at (axis cs:0.2,0) [anchor=north] {$r_0$};
+
+    \end{axis}
+    \end{tikzpicture}
+    \caption{Force-Separation Curve (Lennard-Jones Force)}
+    \label{fig:force-separation}
+\end{figure}
+\begin{equation}
+    E \propto \frac{dF}{dr} \Bigg|_{r_0}
+\end{equation}
+\begin{definition}[Equilibrium interatomic seperation distance]
+The equilibrium interatomic seperation distance, $r_0$, is the distance between two atoms at which the interatomic force is zero.
+This is due to the interatomic forces being the sum of attractive and repulsive forces.
+\end{definition}
+\paragraph{Elastic Modulus} Thus, strongly bonded materials have a higher elastic modulus and the slope of the force-seperation curve is steeper at $r_0$.
+\subsection{Understanding Other Properties in terms of Atomic Configuration}
+\paragraph{Potential Energy-Separation Curve} The potential energy-seperation curve is a plot of the potential energy between two atoms as a function of the distance between them.
+The potential energy is the area under the force-seperation curve.
+\paragraph{Depth of the Minimum Energy Well} The depth of the minimum energy well, $E_0$, is the energy required to break the bond between two atoms.
+This is the energy required to move the atoms from the equilibrium position to infinity.
+It is proportional to the melting temperature of the material.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+        \begin{axis}[
+            xlabel={Separation Distance (d)},
+            ylabel={Potential Energy (U)},
+            grid=major,
+            width=10cm,
+            height=6cm,
+            domain=-0.3:10,
+            samples=100,
+            xtick = \empty,
+            ytick = {0}
+        ]
+        \addplot[blue, thick] {-exp(-x/2) + exp(-x)^2};
+        \node at (axis cs:1,-0.3) [anchor=north] {$E_0$};
+
+    \end{axis}
+    \end{tikzpicture}
+    \caption{Potential Energy-Separation Curve}
+    \label{fig:potential-energy}
+\end{figure}
+\paragraph{Coefficient of Thermal Expansion} The coefficient of thermal expansion, $\alpha$, is the fractional change in length per degree change in temperature. 
+\paragraph{Depth of Potential Energy Curve} The deeper the potential energy curve, the higher the melting temperature and more symetric the curve near $E_0$.
+This would give the follwoing three properties:
+\begin{enumerate}
+    \item \textbf{Higher Melting Temperature} The higher the melting temperature, the deeper the potential energy curve.
+    \item \textbf{Higher Elastic Modulus} The steeper the slope of the force-seperation curve at $r_0$, the higher the elastic modulus.
+    \item \textbf{Lower Coefficient of Thermal Expansion} The more symetric the potential energy curve near $E_0$, the lower the coefficient of thermal expansion.
+\end{enumerate}
+\subsection{Shear and Tensile Stress}
+\subsubsection{Shear}
+\paragraph{Shear Stress} Shear stress is the force per unit area acting parallel to the surface. It is given by:
+\begin{equation}
+    \tau = \frac{F}{A_0}
+\end{equation}
+\paragraph{Shear Strain} Shear strain is the change in angle between two lines originally perpendicular to each other. It is given by:
+\begin{equation}
+    \gamma = \frac{\Delta l}{l_0} \approx \tan \theta \approx \theta = \frac{\pi}{2} - \phi
+\end{equation}
+\paragraph{Shear Modulus} The shear modulus, $G$, is the ratio of shear stress to shear strain. It is given by:
+\begin{equation}
+    \tau = G \gamma
+\end{equation}
+\paragraph{Relationship between Shear and Tensile Modulus} The shear modulus is related to the tensile modulus by the following equation:
+\begin{equation}
+    G = \frac{E}{2(1+\nu)}
+\end{equation}
+where $\nu$ is the Poisson's ratio.
+\paragraph{Poisson's Ratio} Poisson's ratio, $\nu$, is the ratio of lateral strain to axial strain. It is given by:
+\begin{equation}
+    \nu = -\frac{\epsilon_{\text{lat}}}{\epsilon_{\text{axial}}}
+\end{equation}
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-01-16 ECE159 Lecture Notes/main.tex
Add Potential Energy-Separation Curve and update lecture notes

@@
 \begin{definition}[Voltage]
-    The difference of the electric potential between two points in a circuit.
+    The difference of the electric potential between two points in a circuit. The enegy required to move a charge from one point to another.
     \begin{equation}
-        V \defeq \int_C \textbf{E} \cdot d\textbf{l}
+        V \defeq \int_C \textbf{E} \cdot d\textbf{l} = \frac{dW}{dq}
     \end{equation}
     \end{definition}
-
+    \paragraph{Polarity} The polarity is fliped as illustrated when the voltage is indicated as negative.
     \begin{definition}[Current]
@@
     The rate of change of energy for an element in the circuit. Power is the product of voltage and current.
+    \begin{equation}
+        P \defeq \frac{dW}{dt} = \frac{dW}{dq} \cdot \frac{dq}{dt} = V \cdot I
+    \end{equation}
     \end{definition}
@@
 \paragraph{Comination of Series and Parallel Resistors} To find the equivalent resistance of a combination of series and parallel resistors, first find the equivalent resistance of the resistors in series, then find the equivalent resistance of the resistors in parallel.
+\paragraph{Voltage Division} The voltage across a resistor in series is proportional to the resistance of the resistor.
+\begin{equation}
+    V_i = V \frac{R_i}{R_{\text{eq}}}
+\end{equation}
+\paragraph{Current Division} The current through a resistor in parallel is proportional to the conductance of the resistor.
+\begin{equation}
+    I_i = I \frac{G_i}{G_{\text{eq}}}
+\end{equation}
 \subsection{Node, Branch, and Loop}
@@
 \end{enumerate}
+\subsection{Kirchhoff's Laws}
+\paragraph{} Kirchhoff's Laws are two laws that are used to analyze electric circuits.
+\paragraph{Kirchhoff's Current Law (KCL)} The sum of currents entering a node is equal to the sum of currents leaving a node.
+\begin{equation}
+    \sum I_{\text{in}} = \sum I_{\text{out}}
+\end{equation}
+\paragraph{Kirchhoff's Voltage Law (KVL)} The sum of voltages around a closed loop is equal to zero.
+\begin{equation}
+    \sum V_{\text{loop}} = 0
+\end{equation}
+
 \section{DC Circuits}

### 2025-01-16 ESC195 Lecture Notes/AppendixItems/trigDeri.tex
Add Potential Energy-Separation Curve and update lecture notes

\section{Trigonometric Derivatives}
$$
\setlength{\tabcolsep}{8pt} % Adjust column spacing
\renewcommand{\arraystretch}{1.5} % Adjust row spacing
\begin{array}{|l|l|}
\hline
\textbf{Trigonometric Function} & \textbf{Inverse Trigonometric Function} \\ \hline
\dfrac{d}{dx} \text{sin}(x) = \text{cos}(x) & \dfrac{d}{dx} \text{arcsin}(x) = \dfrac{1}{\sqrt{1 - x^2}} \\ \hline
\dfrac{d}{dx} \text{cos}(x) = -\text{sin}(x) & \dfrac{d}{dx} \text{arccos}(x) = \dfrac{-1}{\sqrt{1 - x^2}} \\ \hline
\dfrac{d}{dx} \text{tan}(x) = \text{sec}^2(x) & \dfrac{d}{dx} \text{arctan}(x) = \dfrac{1}{1 + x^2} \\ \hline
\dfrac{d}{dx} \text{cot}(x) = -\text{csc}^2(x) & \dfrac{d}{dx} \text{arccot}(x) = \dfrac{-1}{1 + x^2} \\ \hline
\dfrac{d}{dx} \text{sec}(x) = \text{sec}(x)\text{tan}(x) & \dfrac{d}{dx} \text{arcsec}(x) = \dfrac{1}{|x|\sqrt{x^2 - 1}} \\ \hline
\dfrac{d}{dx} \text{csc}(x) = -\text{csc}(x)\text{cot}(x) & \dfrac{d}{dx} \text{arccsc}(x) = \dfrac{-1}{|x|\sqrt{x^2 - 1}} \\ \hline
\hline
\textbf{Hyperbolic Function} & \textbf{Inverse Hyperbolic Function} \\ \hline
\dfrac{d}{dx} \text{sinh}(x) = \text{cosh}(x) & \dfrac{d}{dx} \text{arsinh}(x) = \dfrac{1}{\sqrt{x^2 + 1}} \\ \hline
\dfrac{d}{dx} \text{cosh}(x) = \text{sinh}(x) & \dfrac{d}{dx} \text{arcosh}(x) = \dfrac{1}{\sqrt{x^2 - 1}} \quad (x > 1) \\ \hline
\dfrac{d}{dx} \text{tanh}(x) = \text{sech}^2(x) & \dfrac{d}{dx} \text{artanh}(x) = \dfrac{1}{1 - x^2} \quad (|x| < 1) \\ \hline
\dfrac{d}{dx} \text{coth}(x) = -\text{csch}^2(x) & \dfrac{d}{dx} \text{arcoth}(x) = \dfrac{1}{1 - x^2} \quad (|x| > 1) \\ \hline
\dfrac{d}{dx} \text{sech}(x) = -\text{sech}(x)\text{tanh}(x) & \dfrac{d}{dx} \text{arsech}(x) = \dfrac{-1}{x\sqrt{1 - x^2}} \quad (0 < x \leq 1) \\ \hline
\dfrac{d}{dx} \text{csch}(x) = -\text{csch}(x)\text{coth}(x) & \dfrac{d}{dx} \text{arcsch}(x) = \dfrac{-1}{|x|\sqrt{x^2 + 1}} \\ \hline
\end{array}
$$

### 2025-01-16 ESC195 Lecture Notes/main.tex
Add Potential Energy-Separation Curve and update lecture notes

@@
 , where $n$ is a positive integer. For demonstartion, We can obtain the reduction formula of $\sin^n(x)$ via IBP.
-\begin{align*}
-    I_n &= \int \sin^n(x) \, dx \\
-    &= \int \sin^{n-1}(x) \sin(x) \, dx \\
-    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \cos^2(x) \sin^{n-2}(x) \, dx \\
+\begin{align}
+    I_n &= \int \sin^n(x) \, dx \nonumber \\
+    &= \int \sin^{n-1}(x) \sin(x) \, dx \nonumber \\
+    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \cos^2(x) \sin^{n-2}(x) \, dx \nonumber \\
     &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} I_{n-2}
-\end{align*}
+\end{align}
+\paragraph{Reduction for $\cos^n(x)$} We can obtain the reduction formula of $\cos^n(x)$ via IBP.
+\begin{align}
+    I_n &= \int \cos^n(x) \, dx \nonumber \\
+    &= \int \cos^{n-1}(x) \cos(x) \, dx \nonumber \\
+    &= \frac{\sin(x) \cos^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \sin^2(x) \cos^{n-2}(x) \, dx \nonumber \\
+    &= \frac{\sin(x) \cos^{n-1}(x)}{n} + \frac{(n-1)}{n} I_{n-2}
+\end{align} 
 \paragraph{Case IV} The following is generally solvable via simple trigonometric integrals. In general, we solve the integral by applying the angle sum formulas.
@@
 We know that $tan^2(x) = \sec^2(x) - 1$. Thus, we can solve the integral by reducing the power of the tangent function.
+\paragraph{Reduction for $\tan^n(x)$} We can obtain the reduction formula of $\tan^n(x)$ via the trigonometric identities.
+\begin{align}
+    I_n &= \int \tan^n(x) \, dx \nonumber \\
+    &= \int \tan^{n-1}(x) \tan(x) \, dx \nonumber \\
+    &= \frac{\tan^{n-1}(x)}{n-1} - \int \tan^{n-2}(x) \, dx \nonumber \\
+    &= \frac{\tan^{n-1}(x)}{n-1} - I_{n-2}
+\end{align}
+\paragraph{Reduction for $\cot^n(x)$} We can obtain the reduction formula of $\cot^n(x)$ via the trigonometric identities.
+\begin{align}
+    I_n &= \int \cot^n(x) \, dx \nonumber \\
+    &= \int \cot^{n-1}(x) \cot(x) \, dx \nonumber \\
+    &= \frac{\cot^{n-1}(x)}{n-1} - \int \cot^{n-2}(x) \, dx \nonumber \\
+    &= \frac{\cot^{n-1}(x)}{n-1} - I_{n-2}
+\end{align}
 \paragraph{Case VI} The following is generally solvable via the trigonometric identities (\ref{eq:tan2}) and (\ref{eq:cot2}).
@@
 \end{itemize}
+\paragraph{Weierstrass Substitution} We can solve integrals by using the Weierstrass substitution:
+\begin{equation}
+    \tan\left(\frac{x}{2}\right) = t \Leftrightarrow \cos(x) = \frac{1-t^2}{1+t^2} \quad \sin(x) = \frac{2t}{1+t^2} \quad dx = \frac{2dt}{1+t^2}
+\end{equation}
+This substitution is useful for solving integrals with trigonometric functions, by converting them into rational functions.
 \paragraph{Summary} We can solve trigonometric integrals by using the following strategies:
@@
 \end{table}
+\subsection{Partial Fractions}
+\paragraph{Partial Fraction Decomposition} We can decompose a rational function into partial fractions to simplify integration.
+\paragraph{} We want to solve the integral:
+\begin{equation}
+    \int \frac{P(x)}{Q(x)} \, dx
+\end{equation}
+, where $P(x)$ and $Q(x)$ are polynomials.
+\begin{example}
+We want to solve the integral:
+$$ \int \frac{2x^2 + 3x + 1}{x^3 + 2x^2 + x} \, dx $$
+\end{example}
+We can decompose the rational function into partial fractions:
+$$ \frac{2x^2 + 3x + 1}{x^3 + 2x^2 + x} = \frac{A}{x} + \frac{B}{x+1} + \frac{C}{(x+1)^2} $$
+\paragraph{} We can solve for the constants $A$, $B$, and $C$ by equating the coefficients of the partial fractions to the original function. We have:
+\begin{align*}
+    2x^2 + 3x + 1 &= A(x+1)^2 + Bx(x+1) + Cx \\
+    \intertext{Thus, we can solve it by setting $x = 0$, $x = -1$}
+\end{align*}
+\paragraph{Summary} We can solve integrals using partial fraction decomposition by following these steps:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|l|}
+    \hline
+    \textbf{Step} & \textbf{Description} \\ \hline
+    1 & Factorize the denominator of the rational function. \\
+    2 & Write the partial fraction decomposition. \\
+    3 & Solve for the constants by equating the coefficients. \\
+    4 & Integrate the partial fractions. \\ \hline
+    \end{tabular}
+    \caption{Steps for Partial Fraction Decomposition}
+\end{table}
+\subsection{Improper Integrals}
+\paragraph{Improper Integral} An improper integral is an integral with an infinite limit or a discontinuity in the interval of integration.
+\paragraph{} We want to solve the integral:
+\begin{equation}
+    \int_a^b f(x) \, dx
+\end{equation}
+, where $a$ or $b$ is infinite or $f(x)$ is discontinuous.
+\begin{example}
+We want to solve the integral:
+$$ \int_0^\infty e^{-x} \, dx $$
 
+Evaluating the integral:
+\begin{align*}
+    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
+    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
+    &= \lim_{b \to \infty} -e^{-b} + 1 \\
+    &= 1
+\end{align*}
+\end{example}
 \section{Hyperbolic Trigonometric Functions}
@@
 \end{align}
-
+\appendix
+\input{appendix.tex}
 \end{document}

### 2025-01-16 MAT185 Lecture Notes/main.tex
Add Potential Energy-Separation Curve and update lecture notes

@@
 \end{proof} 
+\subsection{Vector Subspace}
+\begin{definition}[Vector Subspace]
+    Let $V$ be a vector space and $W \subseteq V$, then $W$ is a vector subspace of $V$ if $W$ is a vector space.
+\end{definition}
+\begin{theorem}[Subspace Test, I]
+    Let $V$ be a vector space over $\mathbb{R}$ and $W \subseteq V$ with the usual addition and scalar multiplication, then $W$ is a subspace of $V$ iff for all $\textbf{x}, \textbf{y} \in W$ and $\alpha \in \mathbb{R}$:
+    \begin{enumerate}
+        \item $\textbf{x} + \textbf{y} \in W$
+        \item $\alpha \textbf{x} \in W$
+        \item $\textbf{0} \in W$
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    ($\Rightarrow$) Assume $W$ is a subspace of $V$, then $W$ is a vector space. Thus, the axioms of vector space are satisfied. \\
+    ($\Leftarrow$) Assume the three conditions are satisfied, then $W$ is a vector space. Thus, $W$ is a subspace of $V$.
+\end{proof}
+\subsection{Linear combinations}
+\begin{definition}[Linear Combination]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
+    \begin{align*}
+        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
+    \end{align*}
+    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
+\end{definition}
+\begin{definition}[Linear independence]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If the equation:
+    \begin{align*}
+        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
+    \end{align*}
+    has only the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.
+
+    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
+\end{definition}
+\begin{definition}[Span]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
+    \begin{align*}
+        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
+    \end{align*}
+    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
+\end{definition}
+
 \end{document}

### 2025-01-16 ESC195 Lecture Notes/main.tex
jan 16

@@
 \end{align}
+\section{Further Applications of Integrals}
+\subsection{Arc Length}
+\paragraph{Arc Length}
+    Given the curve $C: y = f(x)$, the arc length of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the arc length of the curve:
+    $$ L = \int_C ds$$
+\paragraph{Formula} We can derive the formula for the arc length of a curve $y = f(x)$ between two points $a$ and $b$ as:
+\begin{align}
+    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
+    \Delta s &\approx \sqrt{(x_{i+1} - x_i)^2 + (y_{i+1} - y_i)^2} \nonumber \\
+            &= \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
+    \intertext{We also have, by MVT:}
+    \frac{\Delta y}{\Delta x} &= y'(c) \nonumber \\
+    \Delta y &= y'(c) \Delta x \nonumber \\
+    \Delta s &\approx \sqrt{\Delta x^2 + (y'(c) \Delta x)^2} \nonumber \\
+            &= \sqrt{1 + y'(c)^2} \Delta x \nonumber \\
+    \intertext{Taking a Riemann sum:}
+    L &= \lim_{n \to \infty} \sum_{i=1}^n \sqrt{1 + y'(c_i)^2} \Delta x \nonumber \\
+    &= \int_a^b \sqrt{1 + y'(x)^2} \, dx 
+\end{align}
+\begin{example}
+    Given $f(x) = x^\frac{3}{2}$ between $x = 0$ and $x = 44$, we would like to compute the arc length of the curve:
+    \begin{align*}
+        L &= \int_0^{44} \sqrt{1 + \left(\frac{3}{2}x^{\frac{1}{2}}\right)^2} \, dx \\
+        &= \int_0^{44} \sqrt{1 + \frac{9}{4}x} \, dx \\
+        &= \frac{8}{27} \left(1 + \frac{9}{4}x\right)^{\frac{3}{2}} \Big|_0^{44} = 296
+    \end{align*}
+\end{example}
+\subsection{Surface Area of Revolution}
+\paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, the surface area of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the surface area of the curve:
+    $$ A = \int_C 2\pi y \, ds$$
+\paragraph{Formula} Deriving the formula for the surface area of a curve $y = f(x)$ between two points $a$ and $b$ is similar to the arc length formula:
+\begin{align}
+    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
+    \Delta s &\approx \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
+    \intertext{The surface area of the segment is, by continuity:}
+    A_i &= \pi (y_i + y_{i+1}) \Delta s = 2\pi y^\star \Delta s \quad 
+    \intertext{As the difference between $y^\star$ and $y_i$ diminishes as $\Delta x \rightarrow 0$, taking a Riemann sum:}
+    A &= \lim_{n \to \infty} \sum_{i=1}^n 2\pi y_i \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
+    &= \int_a^b 2\pi y \sqrt{1 + y'(x)^2} \, dx
+\end{align}
+\begin{example}
+    Given $y = \sqrt{x}$ between $x = 0$ and $x = 1$, we would like to compute the surface area of the curve:
+    \begin{align*}
+        A &= \int_0^1 2\pi \sqrt{x} \sqrt{1 + \frac{1}{4x}} \, dx \\
+        &= \pi \int^1_0 \sqrt{4x + 1} \, dx \\
+        \intertext{Let $u = 4x + 1$, then $du = 4dx$:}
+        &= \frac{\pi}{4} \int^5_1 \sqrt{u} \, du \\
+        &= \frac{\pi}{6}\left(5^\frac{3}{2} -1\right)
+    \end{align*}
+\end{example}
+\subsection{Applications to Physics and Engineering}
+\paragraph{Hydrostatic Pressure} The hydrostatic pressure and Force is given by the Archimedes' principle:
+\begin{subequations}
+    \begin{align}
+        P &= \rho g \cdot d, \\
+        F &= \rho g \cdot d \cdot A
+    \end{align}
+\end{subequations}
+\paragraph{Formula} Given a tank rectangular with width defined by $w = f(x)$, where $x$ is depth, we would like to compute the force exerted on the tank by the water:
+\begin{align*}
+    F &= \int_0^b \rho g x f(x) \cdot dx
+\end{align*}
+
 \appendix

### 2025-01-16 MAT185 Lecture Notes/main.tex
jan 16

@@
 \begin{definition}[Vector Space in $\mathbb{R}$]
-    A vectors space,${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
+    This course concerns with real vector spaces. A vectors space, ${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
     \begin{enumerate}
@@
             \item \textbf{Existence of Multiplicative Identity:} \\
-            $\forall \, \textbf{x} \in V, \, 1 \textbf{x} = \textbf{x}$
+            $\forall \, \textbf{x} \in V, \, \textbf{1} \textbf{x} = \textbf{x}$
         \end{enumerate}
@@
 \end{proof}
-\paragraph{Results from above}
+\paragraph{Additional Results from Above}
 \begin{enumerate}
@@
 \end{definition}
-\begin{theorem}[Commutativity of Addition]
+\begin{theorem}[Addition is Commutative]
     Let $V$ be a vector space and $\textbf{x}, \textbf{y} \in V$, then:
@@
     \begin{enumerate}
-        \item $\textbf{x} + \textbf{y} \in W$
-        \item $\alpha \textbf{x} \in W$
-        \item $\textbf{0} \in W$
+        \item \textbf{Closure Under Addition:} \\
+        $\textbf{x} + \textbf{y} \in W$
+        \item \textbf{Closure Under Scalar Multiplication:} \\
+        $\alpha \textbf{x} \in W$
+        \item \text{Additive Identity:} \\
+        $\textbf{0} \in W$
     \end{enumerate}
@@
     ($\Leftarrow$) Assume the three conditions are satisfied, then $W$ is a vector space. Thus, $W$ is a subspace of $V$.
+\end{proof}
+\begin{definition}[Null Space]
+    Let $V$ be a vector space and $A \in {^m\mathbb{R}^n}$\footnote{Same as $\mathbb{R}^{m\times n}$, The set of all $m \times n$ matrices with real entries}, then:
+    \begin{align*}
+        \text{null}(A) = \left\{ \textbf{x} \in {^n\mathbb{R}} \, | \, A\textbf{x} = \textbf{0} \right\}
+    \end{align*}
+    is the null space of $A$, otherwise known as the \textbf{kernel} of $A$ or the solution space of $A\textbf{x} = \textbf{0}$
+\end{definition}
+    \paragraph{} We can use the Subspace Test I to show that the null space of a matrix is a subspace of $^n\mathbb{R}$.
+        \begin{enumerate}
+            \item \textbf{Existence of Additive Identity:} \\
+            The zero vector is in the null space of $A$ as the trivial solution to the equation $A\textbf{x} = \textbf{0}$
+            \item \textbf{Closure Under Addition:} \\
+            Let $\textbf{x}, \textbf{y} \in \text{null}(A)$, then $A(\textbf{x} + \textbf{y}) = A\textbf{x} + A\textbf{y} = \textbf{0} + \textbf{0} = \textbf{0}$.
+Thus, $\textbf{x} + \textbf{y} \in \text{null}(A)$.
+This holds since $^n\mathbb{R}$ is a vector space.
+            \item \textbf{Closure Under Scalar Multiplication:} \\
+            Let $\textbf{x} \in \text{null}(A)$ and $\alpha \in \mathbb{R}$, then $A(\alpha \textbf{x}) = \alpha A\textbf{x} = \alpha \textbf{0} = \textbf{0}$.
+Thus, $\alpha \textbf{x} \in \text{null}(A)$.
+This holds since $^n\mathbb{R}^m$ is a vector space under usual addition and scalar multiplication.
+        \end{enumerate}
+\begin{theorem}[Subspace Test, II]
+    Let $V$ be a vector space over $\mathbb{R}$ and $W \subseteq V$ with the usual addition and scalar multiplication, then $W$ is a subspace of $V$ iff for all $\textbf{x}, \textbf{y} \in W$ and $\alpha, \beta \in \mathbb{R}$:
+    \begin{enumerate}
+        \item $\alpha \textbf{x} + \beta \textbf{y} \in W$
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    ($\Rightarrow$) Assume $W$ is a subspace of $V$, then $W$ is a vector space. Thus, the axioms of vector space are satisfied. \\
+    ($\Leftarrow$) Assume the condition is satisfied, then $W$ is a vector space. Thus, $W$ is a subspace of $V$.
 \end{proof}
@@
 \end{definition}
-
 \end{document}

### 2025-01-16 MSE160 Lecture Notes/main.tex
Add Potential Energy-Separation Curve and update lecture notes
jan 16

@@
 \end{equation}
+\subsection{Testing}
+\begin{definition}[Gauge Length]
+The gauge length, $l_0$, is the length of the sample over which the strain is measured.
+\end{definition}
+\begin{definition}[Reduced Section]
+The reduced section is the part of the sample where the cross-sectional area is reduced to a smaller value.
+\end{definition}
+\paragraph{} Gauge length is always no longer than the reduced section. The reduced section is where the sample will likely break.
+\paragraph{Testing Ceremics} In relation to tensile testing, ceramics have the following properties:
+\begin{itemize}
+    \item \textbf{Brittle} Ceramics are brittle and will break suddenly.
+    \item \textbf{High Strength} Ceramics have high strength and thus difficult to machine the sample.
+    \item \textbf{Sample Alignment} The sample must be aligned properly to test for pure tension. Unlike metals and polymers, which are self-aligning.
+    \item \textbf{Fracture} Ceramics will fracture while still off-axis. Hence, there would be a large shear component. 
+\end{itemize}
+\paragraph{} Thus, we often approxiate tensile behaviour with a point loading on a horizontal beam, with two point support (3 point bending test). Peak stress is given by:
+\begin{equation}
+    \sigma_{\text{peak}} = \frac{3FL}{2bd^2}
+\end{equation}
+, where:
+\begin{itemize}
+    \item $L$ (span) is the distance between the two supports.
+    \item $b$ is the width of the sample
+    \item $d$ is the thickness/depth of the sample
+\end{itemize}
+\section{Seleection of Materials}
+\begin{example}[Aircraft Wing Spar]
+    The aircraft wing spar is beam (loaded in bending) that supports the wing. The spar is made of a material with the objective of minimize mass under the following constraints:
+    \begin{itemize}
+        \item \textbf{Deflection} There is a maximum allowable deflection of the wing.
+        \item There is more..., but for this example, we will only consider the deflection.
+    \end{itemize}
+    The material selection solve for a \textbf{light stiff beam}.
+
+    \paragraph{Mass} The mass of the beam is given by:
+    $$ m = \rho V = \rho A L $$
+    \paragraph{Deflection} The deflection of the beam is given by:
+    \begin{align*} 
+        \delta &= \frac{FL^3}{48EI} 
+        \intertext{For a beam with a rectangular cross-section, we have:}
+        \delta &= \frac{FL^3}{48E} \cdot \frac{12}{bh^3} = \frac{FL^3}{4Ebh^3}  
+        \intertext{We can set $b$ proportional to $h$:}
+        \delta &= \frac{FL^3}{cE} \cdot \frac{1}{A^2},\quad \text{for some constant $c$} 
+        \intertext{We can then isolate for $A$, the free variable, and minimize the mass via the objective equation $m = \rho A L$:}
+        A &= \sqrt{\frac{FL^3}{cE\delta}} \\
+        m &= \rho L \sqrt{\frac{FL^3}{cE\delta}} = \rho L \sqrt{\frac{FL^3}{cE\delta}}
+        \intertext{Arrange into the form (functional)(geometric)(material):}
+        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\left(\frac{\rho}{E^\frac{1}{2}}\right)
+    \end{align*}
+\end{example}
+\paragraph{Material Peformance Index} The material performance index is given by:
+    \begin{equation}
+        \text{Material Performance Index (MSI)} = \frac{E^\frac{1}{2}}{\rho}
+\end{equation}
+\paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
+\appendix
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-01-17 ESC195 Lecture Notes/main.tex
jan17

@@
 \begin{align}
-    \cosh^2(x) - \sinh^2(x) &= 1, \\
+    \cosh^2(x) + \sinh^2(x) &= 1, \\
     \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
@@
 \subsection{Applications to Physics and Engineering}
+\subsubsection{Hydrostatic Pressure and Force}
 \paragraph{Hydrostatic Pressure} The hydrostatic pressure and Force is given by the Archimedes' principle:
@@
 \paragraph{Formula} Given a tank rectangular with width defined by $w = f(x)$, where $x$ is depth, we would like to compute the force exerted on the tank by the water:
+\begin{equation}
+    F = \int_0^b \rho g x f(x) \cdot dx
+\end{equation}
+, which is derived by considering $dx$ as the depth a horizontal slice of the tank.
+\subsubsection{Center of Mass and Moments of Inertia}
+\paragraph{Properties of the Center of Mass} The center of mass of a region $R$ has the following properties:
+\begin{enumerate}
+    \item \textbf{Symetry}: For all axis of symmetry, the center of mass lies on the axis.
+    \item \textbf{Additivity}: The center of mass of a region is the weighted average of the centers of mass of its parts. The weights are the areas of the parts.
+    \begin{equation}
+        \bar{x} = \sum_i \frac{A_i\bar{x_i}}{A}
+    \end{equation}
+\end{enumerate}
+\paragraph{Formula} The centroid of a uniformly-dense region $R$ bounded by $y=f(x),\,x\in[a,b],\,y\in[f(a),f(b)]$ is derived as follows:
+\begin{align}
+    \intertext{We first have $x_i^\star$ and $A_i$ as follows:}
+    x_i^\star = \frac{1}{2}(x_i + x_{i+1}) &\quad A_i = f(x_i) \Delta x \nonumber \\
+    \intertext{For $\bar{x}$, we have:}
+    A\bar{x} &\approx \sum_i x_i^\star A_i \nonumber \\
+    \intertext{Taking a Riemann sum:}
+    A\bar{x} &= \lim_{n \to \infty} \sum_i x_i^\star A_i  \nonumber \\
+    \bar{x} &= \frac{1}{A}\int_a^b x f(x) \, dx
+    \intertext{Similarly, for $\bar{y}$:}
+    A\bar{y} &= \lim_{n \to \infty} \sum_i f(x_i^\star) A_i \nonumber \\
+    \bar{y} &= \frac{1}{A}\int_a^b \frac{1}{2}(f(x))^2 \, dx
+\end{align}
+\begin{theorem}[Pappus's Theorem]
+    The volume of a solid of revolution is given by the product of the area of the region and the distance, $R$, traveled by the centroid of the region from the axis of rotation.
+    \begin{equation}
+        V = 2\pi R A
+    \end{equation}
+\end{theorem}
+\begin{example}[Eliptical Torus]
+    Given an eliptical torus (tall donut) with major axis $a$ and minor axis $b$, we would like to compute the volume of the torus:
     \begin{align*}
-    F &= \int_0^b \rho g x f(x) \cdot dx
+        V &= 2\pi R A \\
+        &= 2\pi R \cdot \pi ab = 2\pi^2 abR
     \end{align*}
+    This could be, alternatively, shown via the washer method and the shell method.
+\end{example}
+\subsection{Applications to Economics and Biology}
+``Biologists are so bad at math'' - Prof. Davis as he proceeds to erase the board and skip to the next section.
+\section{Parametric Equations and Polar Coordinates}
+\subsection{Curves Defined by Parametric Equations}
+\paragraph{Parametric Equations} For $t \in \mathbb{R}$, we have:
+\begin{equation}
+    x = x(t), \quad y = y(t)
+\end{equation}
+\begin{example}[Newton's Laws of Motion]
+    Given $\ddot{x} = 0$, $\ddot{y} = -g$, we have:
+    \begin{align*}
+        x(t) &= A_1 t + A_2, \\
+        y(t) &= -\frac{1}{2}gt^2 + B_1 t + B_2
+    \end{align*}
+    At initial conditions $t = 0$, we have $x(0) = 0$, $y(0) = 0$, $\dot x(0) = v_0 \cos(\theta)$, $\dot y(0) = v_0 \sin(\theta) - \frac{1}{2}gt^2$.
+\end{example}
+\begin{example}[Ellipse]
+    Given $x = a\cos(t)$, $y = b\sin(t)$, we have:
+    \begin{align*}
+        x^2 &= a^2\cos^2(t), \\
+        y^2 &= b^2\sin^2(t)
+    \end{align*}
+    , which satisfies $x^2/a^2 + y^2/b^2 = 1$.
+\end{example}
+\paragraph{Intersections of Parametric Curves} Given two parametric curves $C_1,\,C_2$, defined by:
+\begin{align*}
+    C_1:\, x_1 &= x_1(t), \quad y_1 = y_1(t) \\
+    C_2:\, x_2 &= x_2(t), \quad y_2 = y_2(t)
+\end{align*}
+, we can find the intersection points by solving the system of equations:
+\begin{align*}
+    x_1(t) &= x_2(t) \\
+    y_1(t) &= y_2(t)
+\end{align*}
+\begin{example}
+    Given $x_1 = 2t+t$, $y_1 = 5-4t$, $x_2 = 3 - 5\cos(\pi t)$, $y_2 = 1 + 5\sin(\pi t)$, we have:
+    \begin{align*}
+        2t + 1 &= 3 - 5\cos(\pi t) \\
+        5 - 4t &= 1 + 5\sin(\pi t)
+    \end{align*}
+\end{example}
 

### 2025-01-17 MAT185 Lecture Notes/main.tex
jan17

@@
     Let $V$ be a vector space and $A \in {^m\mathbb{R}^n}$\footnote{Same as $\mathbb{R}^{m\times n}$, The set of all $m \times n$ matrices with real entries}, then:
-    \begin{align*}
+    \begin{align}
         \text{null}(A) = \left\{ \textbf{x} \in {^n\mathbb{R}} \, | \, A\textbf{x} = \textbf{0} \right\}
-    \end{align*}
+    \end{align}
     is the null space of $A$, otherwise known as the \textbf{kernel} of $A$ or the solution space of $A\textbf{x} = \textbf{0}$
@@
 \end{proof}
+\begin{definition}[Intersection of Sets]
+    Let $A$ and $B$ be sets, then:
+    \begin{align}
+        A \cap B = \left\{ x \, | \, x \in A \, \text{and} \, x \in B \right\}
+    \end{align}
+\end{definition}
+\begin{definition}[Union of Sets]
+    Let $A$ and $B$ be sets, then:
+    \begin{align}
+        A \cup B = \left\{ x \, | \, x \in A \, \text{or} \, x \in B \right\}
+    \end{align}
+\end{definition}
+
 \subsection{Linear combinations}
@@
     Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
-    \begin{align*}
+    \begin{align}
         \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
-    \end{align*}
+    \end{align}
     is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$

### 2025-01-17 MSE160 Lecture Notes/main.tex
jan17

@@
         \intertext{Arrange into the form (functional)(geometric)(material):}
-        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\left(\frac{\rho}{E^\frac{1}{2}}\right)
+        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\cdot\left(\frac{\rho}{E^\frac{1}{2}}\right)
     \end{align*}
@@
 \paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
+\paragraph{Tempered Glass} Temperglass are made to resist tension.
+It is done by applying a compressive stress to the surface of the glass.
+This is done by cooling the surface of the glass faster than the core or chemically treating the surface.
 \appendix

### 2025-01-17 linalgA1/main.tex
jan17

\documentclass[10pt,addpoints,answers]{exam}
\input{preamble}

\begin{document}

\vspace*{-0.5cm}
\begin{center}
  \large
  \textbf{\Large \course~--~Linear Algebra}\\[0.1cm]
  \textbf{\assignmenttitle}
\end{center}
\bigskip

\textbf{\large Instructions:}\\
\normalsize

Please read the {\bf MAT185 Assignment Policies \& FAQ} document for details on
submission policies, collaboration rules and academic integrity, and general
instructions.

\begin{enumerate}

\item \textbf{Submissions are only accepted by}
  \href{https://www.gradescope.ca}{Gradescope}. Do not send anything by email.
  Late submissions are not accepted under any circumstance. Remember you can
  resubmit anytime before the deadline.

\item \textbf{Submit solutions using only this template pdf}.  Your submission
  should be a single pdf with your full written solutions for each question. If
  your solution is not written using this template pdf (scanned print or
  digital) then your submission will not be assessed. Organize your work neatly
  in the space provided.  Do not submit rough work.

\item \textbf{Show your work and justify your steps} on every question but do
  not include extraneous information.  Put your final answer in the box
  provided, if necessary.  We recommend you write draft solutions on separate
  pages and afterwards write your polished solutions here on this template.

\item \textbf{You must fill out and sign the academic integrity statement
    below}; otherwise, you will receive zero for this assignment.

\end{enumerate}

\vspace{10pt}

\textbf{\large Academic Integrity Statement:} \\

%%% Student information

% Student 1
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: Cheung, Hei Shing\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 101090782\enspace\hrulefill}\\
  \end{minipage}
}

\vspace*{0.1in}

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name:\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number:\enspace\hrulefill}\\
  \end{minipage}
}

\bigskip
\large \textbf{I confirm that:}
\normalsize

\begin{itemize} 
\item I have read and followed the policies described in the document {\bf
    MAT185 Assignment Policies \& FAQ}.
\item In particular, I have read and understand the rules for
  collaboration, and permitted resources on assignments as described in
  subsection II of the the aforementioned document. I have not violated
  these rules while completing and writing this assignment.
\item I have not used generative AI in writing this assignment.
\item I understand the consequences of violating the University's academic
  integrity policies as outlined in the
  \href{http://www.governingcouncil.utoronto.ca/policies/behaveac.htm}{Code of
    Behaviour on Academic Matters}. I have not violated them while completing
  and writing this assignment.
\end{itemize}
\bigskip

\large \textbf{By submitting this assignment to Gradescope, I agree that the
  statements above are true.}  \normalsize

\newpage

\begin{questions}
  \question In this problem, you will prove that the elementary operations you
  learnt in ESC103 do not change the set of solutions of a system of linear
  equations.  Consider the four linear systems
$$
\hspace{.8cm} \mathcal{A}: \hspace{.5cm} \left\{ \arraycolsep=1.4pt
	\begin{array}{rcrrcrcr}
          3x &+&2y &+&2z &= &9\\
          11x &+&7y &+&3z &=& 15\\
          3x &+&2y &+&z &=& 5
	\end{array} \right.
      \hspace{1.9cm}
      \mathcal{B}: \hspace{.5cm} \left\{
        \arraycolsep=1.4pt
	\begin{array}{rcrcrccr}
          3x &+&2y &+&2z &= &9\\
          11x &+&7y &+&3z &=& 15\\
          (3 + 11 \beta) x &+&(2+7 \beta) y &+&(1+3 \beta) z &=& 5 + 15 \beta
	\end{array} \right.
$$
$$
\hspace{-3.3cm} \mathcal{C}: \hspace{.5cm} \left\{ \arraycolsep=1.4pt
	\begin{array}{rcrcrccr}
          3x &+&2y &+&2z &= &9\\
          11 \alpha x &+&7 \alpha y &+&3 \alpha z &=& 15 \alpha\\
          3x &+&2y &+&z &=& 5
	\end{array} \right.
      \hspace{.9cm}
      \mathcal{D}: \hspace{.5cm} \left\{
        \arraycolsep=1.4pt
	\begin{array}{rcrcrccr}
          11x &+&7y &+&3z &=& 15\\
          3x &+&2y &+&2z &= &9\\
          3x &+&2y &+&z &=& 5
	\end{array} \right.
$$
Above, $x$, $y$, and $z$ are all \textbf{variables}.  The parameters $\alpha$
and $\beta$ are real numbers.

Let $A \subseteq \R^3$ be the \textbf{set of solutions} of the linear system
$\mathcal{A}$.  A point $(a,b,c) \in A$ means that
$$
\hspace{.8cm} \mathcal{A}: \hspace{.5cm} \left\{ \arraycolsep=1.4pt
	\begin{array}{rcrcrccrl}
          3a &+&2b &+&2c &= &9  &\hspace{.5cm} \mbox{\textbf{TRUE}}\\
          11a &+&7b &+&3c &=& 15 & \mbox{\textbf{TRUE}}\\
          3a &+&2b &+&c &=& 5 & \mbox{\textbf{TRUE}}
	\end{array} \right.
$$
Because $a$, $b$, and $c$ are all \textbf{real numbers}, equations involving
them are either true (as in $1+1=2$ is true) or false (as in $1+1=3$ is false).
And so, $(a,b,c)$ is in $A$ (is a solution of $\mathcal{A}$) if all three
equations are true when the variables $x$, $y$, and $z$ take on the values $a$,
$b$, and $c$ respectively.  If there are no points $(a,b,c)$ for which the three
equations are all true then $\mathcal{A}$ has no solutions and
$A = \varnothing$.  \textit{Note: this does not contradict the statement that
  $A \subseteq \R^3$ because $\varnothing \subseteq \R^3$.}

\underline{Pro tip:} If you have an equation where the left-hand side and
right-hand side are real numbers (or are numbers in any field) then the equation
$LHS = RHS$ is \textbf{TRUE} if and only if $LHS-RHS = 0$ and is \textbf{FALSE}
if and only if $LHS-RHS \neq 0$.

Finally, there is the question of how you could show that two sets, $A$ and $B$,
are equal.
\begin{itemize}
\item If you want to prove that $A = B$ and $A = \varnothing$ then you need to
  prove that $B = \varnothing$.  %That is, $A$ and $B$ are both the empty set.
\item If you want to prove that $A = B$ and $B = \varnothing$ then you need to
  prove that $A = \varnothing$.  %That is, $A$ and $B$ are both the empty set.
\item If you want to prove that $A = B$ and neither $A$ nor $B$ are nonempty
  then you can do this by proving
  \begin{itemize}
  \item if $a \in A$ then $a \in B$ (this proves $A \subseteq B$) \textbf{and}
  \item if $b \in B$ then $b \in A$ (this proves $B \subseteq A$).
  \end{itemize}
\end{itemize} If you have sets $A$ and $B$ and you don't know whether or not
they are empty or nonempty, you have do all three steps above.

\textbf{For the curious:} The third case is related to another common proof
technique: if you want to prove that two real numbers are equal, $a=b$, you
first use one idea/approach to prove $a \leq b$ and a different idea/approach to
prove $a \geq b$.  This technique is jarring the first few times one sees it.
This approach works for any totally-ordered field; it doesn't work for the
complex numbers, for example.

\vfil

\begin{parts}
  \part Let $A$ be the set of solutions of $\mathcal{A}$ and $B$ be the set of
  solutions of $\mathcal{B}$.  Prove that $A = B$.    \textit{Note: you don't know
    whether or not $\mathcal{A}$ or $\mathcal{B}$ have any solutions.  This
    means that you must address all three possible cases: $A = \varnothing$,
    $B = \varnothing$, and neither $A$ nor $B$ are the empty set.}

    \textbf{Please start your answer on the next page, not on this one.  The grader will not look at this page.  You can continue your answer on the top of page 4, if needed.}

  \newpage

   \textcolor{black}{\underline{case 1: $A = \varnothing$}}
  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{130pt}{
    \paragraph{Note} I denote the $k^{\text{th}}$ equation of $\mathcal{P}$ as $\mathcal{P}_k$ for linear system $\mathcal{P} \in \{\mathcal{A}, \mathcal{B}, \mathcal{C}, \mathcal{D}\}$.

    %%% Your work goes here!  
   \paragraph{} We prove that $B = \varnothing$. Assume, for the sake of contradiction, that $B \neq \varnothing$. Then there exists a point $(a,b,c) \in B$ such that
       $\mathcal{B}_1$, $\mathcal{B}_2$, and $\mathcal{B}_3$ are \textbf{TRUE}. Thus:
       \begin{itemize}
        \item For all $\beta \in \mathbb{R}$, $\beta\mathcal{B}_2: 11\beta a + 7\beta b + 3\beta c = 15\beta$ is \textbf{TRUE}; also
        \item $\mathcal{B}_3$ is \textbf{TRUE}. 
       \end{itemize}
    Hence, $\mathcal{B}_3 - \beta \mathcal{B}_2$ is \textbf{TRUE} for all $\beta \in \mathbb{R}$. This implies there exist $(a,b,c)$ s.t.: $3a + 2b + c = 5$ is \textbf{TRUE}. It follows from $\mathcal{A}_3: 3a + 2b + c = 5$ and that $\mathcal{B}_2 = \mathcal{A}_2$ and that $\mathcal{B}_1 = \mathcal{A}_1$ are all \textbf{TRUE}. Thus, $A \neq \varnothing$, a contradiction. Therefore, $B = \varnothing$.

% end of your answer to case 1 unless you need to continue the answer on page 4
}

   \textcolor{black}{ \underline{case 2: $B = \varnothing$} }
   %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{130pt}{
   %%% Your work goes here!  
   \paragraph{} We prove that $A = \varnothing$. Assume, for the sake of contradiction, that $A \neq \varnothing$. Then there exists a point $(a,b,c) \in A$ such that $\mathcal{A}_1$, $\mathcal{A}_2$, and $\mathcal{A}_3$ are \textbf{TRUE}. Thus:
       \begin{itemize}
        \item For all $\beta \in \mathbb{R}$, $\beta\mathcal{A}_2: 11\beta a + 7\beta b + 3\beta c = 15\beta$ is \textbf{TRUE}; also
        \item $\mathcal{A}_3$ is \textbf{TRUE}. 
       \end{itemize}
    Hence, $\mathcal{B}_3 = \mathcal{A}_3 + \beta \mathcal{A}_2$ is \textbf{TRUE} for all $\beta \in \mathbb{R}$. It follows from $\mathcal{A}_2 = \mathcal{B}_2$ and that $\mathcal{A}_1 = \mathcal{B}_1$ are all \textbf{TRUE}. Thus, $B \neq \varnothing$, a contradiction. Therefore, $A = \varnothing$.
% end of your answer to case 2 unless you need to continue the answer on page 4
}

   \textcolor{black}{ \underline{case 3: both $A$ and $B$ are nonempty} }
   %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{120pt}{
   %%% Your work goes here!  
   \paragraph{} ($A \subseteq B$) Take $(a_1, a_2, a_3) \in A$. Then $\mathcal{A}_1$, $\mathcal{A}_2$, and $\mathcal{A}_3$ are \textbf{TRUE}. Thus:
       \begin{itemize}
        \item For all $\beta \in \mathbb{R}$, $\beta\mathcal{A}_2: 11\beta a_1 + 7\beta a_2 + 3\beta a_3 = 15\beta$ is \textbf{TRUE}; also
        \item $\mathcal{A}_3$ is \textbf{TRUE}. 
       \end{itemize}
    Hence, $\mathcal{B}_3 = \mathcal{A}_3 + \beta \mathcal{A}_2$ is \textbf{TRUE} for all $\beta \in \mathbb{R}$. It follows from $\mathcal{A}_2 = \mathcal{B}_2$ and that $\mathcal{A}_1 = \mathcal{B}_1$ are all \textbf{TRUE}. Thus, $(a_1, a_2, a_3) \in B$. Therefore, $A \subseteq B$.

    \paragraph{} ($B \subseteq A$) Take $(b_1, b_2, b_3) \in B$. Then $\mathcal{B}_1$, $\mathcal{B}_2$, and $\mathcal{B}_3$ are \textbf{TRUE}. Thus:
       \begin{itemize}
        \item For all $\beta \in \mathbb{R}$, $\beta\mathcal{B}_2: 11\beta b_1 + 7\beta b_2 + 3\beta b_3 = 15\beta$ is \textbf{TRUE}; also
        \item $\mathcal{B}_3$ is \textbf{TRUE}. 
       \end{itemize}
    Hence, $\mathcal{A}_3 = \mathcal{B}_3 - \beta \mathcal{B}_2$ is \textbf{TRUE} for all $\beta \in \mathbb{R}$. It follows from $\mathcal{B}_2 = \mathcal{A}_2$ and that $\mathcal{B}_1 = \mathcal{A}_1$ are all \textbf{TRUE}. Thus, $(b_1, b_2, b_3) \in A$. Therefore, $B \subseteq A$.

% end of your answer to case 3 unless you need to continue the answer on page 4
}

  \newpage

  \textbf{You can continue your answer here.  If you do so, please make it clear which of the
  three cases you're continuing!}

    %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{232pt}{
    %%% Your work goes here!  Here is where you can put the continuation of your work from page 3
    %%% to page 4.  It is your responsibility to trim down your answer so that it doesn't run into 
    %%% the next question

    %%% end of your answer    
  }

 %\vspace{1cm}

 \part Let $A$ be the set of solutions of $\mathcal{A}$ and $C$ be the set of
solutions of $\mathcal{C}$.

 \begin{subparts} \subpart The system $\mathcal{C}$ has a parameter $\alpha \in
\R$.  Under what conditions on $\alpha$ is $A \subseteq C$?  \textit{No justification needed,
just one complete sentence.}

\vspace{.15in}

   %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{25pt}{ 
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    $\alpha \in \mathbb{R}$
    %%% end of your answer    
  }

  \vspace{.15in}

  \subpart Under what conditions on $\alpha$ is $C \subseteq A$? \textit{No justification needed,
just one complete sentence.} \vspace{.15in}
\makenonemptybox{25pt}{ 
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    $\alpha \neq 0$
    %%% end of your answer    
  }

  \subpart Assuming the condition of part ii. holds, prove that $C \subseteq A$.
\textit{You may assume that $C \neq \varnothing$; we'll assume that if you can
address the $C = \varnothing$ case based on your methods in part (a).}\\

%%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{25pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    \paragraph{} Given nonempty $A,\, C$, take $(a, b, c) \in C$ such that $\mathcal{C}_1$, $\mathcal{C}_2$, and $\mathcal{C}_3$ are \textbf{TRUE}.  For all $\alpha \neq 0$, we have $\frac{1}{\alpha} \in \mathbb{R}$. So $\frac{1}{\alpha}\mathcal{C}_2$ is \textbf{TRUE}. Thus $\mathcal{B}_2 = \mathcal{C}_2$ is \textbf{TRUE}. It follows that $\mathcal{C}_3 = \mathcal{B}_3$ is \textbf{TRUE} and that $\mathcal{C}_1 = \mathcal{B}_1$ is \textbf{TRUE}. Therefore, $(a, b, c) \in A$. Hence, $C \subseteq A$.

    %%% end of your answer      
  }

 \end{subparts}

 \newpage

 \begin{tikzpicture}
\path (0,0) (.5,0) (.5,.5) (0,.5) (0,0);
\end{tikzpicture}

 \part Let $D$ be the set of solutions of $\mathcal{D}$.  Prove that $A = D$.
\textit{You may assume that both $A$ and $D$ are nonempty; we'll assume you can
address the case where either of them is the empty set based on your work in
part (a).}

\textbf{This part is worth zero points.  Your proof will not be
graded.  If you got stuck, or wrote a proof
you aren't quite sure of, please post to piazza or come to office hours.}

\textbf{YOU MUST UPLOAD THIS PAGE EVEN IF YOU WROTE NOTHING ON IT.}

%%% Do not change the height of this box. Your work must fit inside it.

\makenonemptybox{500pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    \paragraph{} We have $\mathcal{A}_1 = \mathcal{D}_2$ and $\mathcal{A}_2 = \mathcal{D}_1$ and $\mathcal{A}_3 = \mathcal{D}_3$. Given nonempty $A,\, B$, we have:

    \paragraph{} ($A \subseteq D$) Take $(a_1, a_2, a_3) \in A$. Then $\mathcal{A}_1$, $\mathcal{A}_2$, and $\mathcal{A}_3$ are \textbf{TRUE}. Thus, in respective order:
       \begin{itemize}
        \item $\mathcal{D}_2$ is \textbf{TRUE}; also
        \item $\mathcal{D}_1$ is \textbf{TRUE}; also
        \item $\mathcal{D}_3$ is \textbf{TRUE},
       \end{itemize} 
       so, it follows that $(a_1, a_2, a_3) \in D$. Therefore, $A \subseteq D$.

    \paragraph{} ($D \subseteq A$) Take $(d_1, d_2, d_3) \in D$. Then $\mathcal{D}_1$, $\mathcal{D}_2$, and $\mathcal{D}_3$ are \textbf{TRUE}. Thus, in respective order:
        \begin{itemize}
          \item $\mathcal{A}_2$ is \textbf{TRUE}; also
          \item $\mathcal{A}_1$ is \textbf{TRUE}; also
          \item $\mathcal{A}_3$ is \textbf{TRUE},
        \end{itemize}
        so, it follows that $(d_1, d_2, d_3) \in A$. Therefore, $D \subseteq A$.

    %%% end of your answer
 }

 \newpage

 \part At this point, you should have proven that the three elementary
operations did not change the set of solutions of the linear system
$\mathcal{A}$.  The key insights of your proofs didn't rely on $\mathcal{A}$'s
having exactly three equations and three variables and having coefficients $3$,
$2$, $11$ and so forth.  We asked you to think about a simple, concrete example
so that you could focus on the elementary operations and not be distracted by
notation.  Now we are asking you to face the notation and prove something about
a general linear system.

 Consider a system of $m$ linear equations with $n$ variables
 $$
 \mathcal{A}: \hspace{.5cm} \left\{ \sum_{j=1}^n a_{ij} x_j = b_i \qquad 1 \leq
i \leq m \right.
 $$
 where the coefficients $a_{ij}$ and constants $b_i$ are all real numbers.  Let
$\mathcal{B}$ be the linear system where the $i_0$th equation of $\mathcal{A}$
has been multiplied by $\alpha \neq 0$.

 Assume that both $\mathcal{A}$ and $\mathcal{B}$ have solutions.  Prove that
$\mathcal{A}$ and $\mathcal{B}$ have the same set of solutions.

\textbf{This part is worth zero points.  Your proof will not be
graded.  If you got stuck, or wrote a proof
you aren't quite sure of, please post to piazza or come to office hours.}

\textbf{YOU MUST UPLOAD THIS PAGE EVEN IF YOU WROTE NOTHING ON IT.}

%%% Do not change the height of this box. Your work must fit inside it.
\makenonemptybox{200pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't

    %%% end of your answer   
 }

 \newpage

 \part Consider writing down the augmented matrix of a linear system
$\mathcal{A}$ and applying elementary row operations to the matrix until you
have a matrix in reduced row echelon form (RREF).  Let $\mathcal{R}$ be a linear
system represented by that RREF matrix.  Explain why $\mathcal{A}$ and
$\mathcal{R}$ have the same set of solutions.  \textit{There's a lot of white
space here but this isn't to indicate that you need to write at length to answer
this question.  You should be able to write your explanation w/ 100 or fewer
words (in the space above the horizontal line). }

%%% Do not change the height of this box. Your work must fit inside it.
\makenonemptybox{250pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    Consider the augmented matrix of $\mathcal{A}$ and apply elementary row operations to obtain the RREF matrix of $\mathcal{A}$. For each elementary row operation, the set of solutions of $\mathcal{A}$ remains unchanged. Thus, $\mathcal{A}$ and $\mathcal{R}$ have the same set of solutions.
    %%% end of your answer  
}

 \noindent\rule{.93\textwidth}{1pt}

\newpage

 \end{parts}

 \question One of the goals of this question is to get you into good shape for
True/False questions on exams.

 For questions (b), (c), and (d) below, please assume the ``natural'' vector
addition and scalar multiplication.  For example, for functions assume the
addition and multiplication given in section 4.2 of Medici.  The ones given in
the ``An Unusual Vector Space'' example are completely valid but are considered
``unnatural'' for this problem.

 \begin{parts}
 \part Prove the following lemma by showing that all the vector space axioms
hold.  It's very useful and once you've proven it you can use it whenever you
want.

\noindent{\bf \color{dkrgreen} Lemma}: Let $\V$ be a vector space over the
real numbers. If $\V_0 \subseteq \V$ contains ${\bf 0}$ and,
with the same vector addition and scalar multiplication of $\V$,
 is closed under vector
addition and scalar multiplication 
then $\V_0$ is a vector space over the
real numbers.

%%% Do not change the height of this box. Your work must fit inside it.
\makenonemptybox{400pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
    %%% either, please!)

    %%% end of your answer     
}

 \newpage

 \begin{EnvUplevel} \textbf{True or False:} Unsupported answers will not receive
full credit. Organize your work in a reasonably neat and coherent way.\\[0.25cm]
Indicate your final answers by \textbf{filling in exactly one circle} for each
part below (unfilled \tikz\draw[black] (0,0) circle (.8ex); \, filled
\tikz\draw[black,fill=black] (0,0) circle (.8ex);).

\textbf{Hint: if you think any of the following are vector spaces, ask yourself whether 
there's a way of 
proving that this is true without having to start at the definition and proving that all eight
axioms hold.}

 \end{EnvUplevel}

 \part The set $\V$ of all nonpositive real-valued functions on $[2,3]$ 
  with the usual addition and scalar multiplication
 is a
vector space over $\R$.  \textit{Note: a real-valued function on $[2,3]$ is a function
$f: [2,3] \to \R$.  That is, its domain is $[2,3]$ and its range is a subset of
$\R$.  Its range could be all of $\R$, of course.  A function is nonpositive if
$f(x) \leq 0$ for all $x$ in its domain.}

 % mark option as correct by using \CorrectChoice
\begin{checkboxes}
  \choice True
  \choice False
\end{checkboxes}

 %%% Do not change the height of this box. Your work must fit inside it.
 \makenonemptybox{150pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
    %%% either, please!)

    %%% end of your answer   
 }

\part The set $\V$ of all real polynomials of degree exactly $n$ 
 with the usual addition and scalar multiplication
is a vector
space.  \textit{Note: A ``real polynomial'' is a polynomial whose coefficients
are all real numbers.}

% mark option as correct by using \CorrectChoice
\begin{checkboxes}
  \choice True
  \choice False
\end{checkboxes}

 %%% Do not change the height of this box. Your work must fit inside it.
 \makenonemptybox{100pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
    %%% either, please!)

    %%% end of your answer    
 }

\newpage
 \begin{EnvUplevel} \textbf{True or False:} Unsupported answers will not receive
full credit. Organize your work in a reasonably neat and coherent way.\\[0.25cm]
Indicate your final answers by \textbf{filling in exactly one circle} for each
part below (unfilled \tikz\draw[black] (0,0) circle (.8ex); \, filled
\tikz\draw[black,fill=black] (0,0) circle (.8ex);).

\textbf{Hint: if you think the following is a vector space, ask yourself whether 
there's a way of 
proving that this is true without having to start at the definition and proving that all eight
axioms hold.}
 \end{EnvUplevel}

\part The set of $2 \times 2$ upper triangular real matrices with
the usual entry-wise addition and scalar multiplication
is a vector
space over the real numbers.

\textit{Notes: An $n \times n$ matrix ${\bf A} = [a_{ij}]$ is upper triangular if
$a_{ij}=0$ for all $1 \leq j < i \leq n$.  A matrix is real if all of its entries are
real numbers.  All integer matrices and all rational matrices are also real
matrices.  After doing this problem, ask yourself how your answer would apply
to $n \times n$ upper triangular matrices.}

% mark option as correct by using \CorrectChoice
\begin{checkboxes}
  \choice True
  \choice False
\end{checkboxes}

 %%% Do not change the height of this box. Your work must fit inside it.
 \makenonemptybox{175pt}{
    %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
    %%% either, please!)

    %%% end of your answer     
 }

 \end{parts}
\end{questions}

\end{document}

### 2025-01-20 ECE159 Lecture Notes/main.tex
jan 20

@@
 \end{equation}
+\paragraph{Node and Mesh Analysis} Node and Mesh Analysis are two methods used to analyze electric circuits.
+\paragraph{Node Analysis} Node Analysis is a method used to analyze electric circuits by writing KCL equations at each node.
+\paragraph{Ground Node} A ground node is a node in an electric circuit that is used as a reference point for voltage measurements.
+\paragraph{Mesh Analysis} Mesh Analysis is a method used to analyze electric circuits by writing KVL equations for each loop.
+\paragraph{Supermesh} A supermesh is a loop that contains a current source.
+To analyze a supermesh, write a KVL equation for the supermesh and a KCL equation for the node that the current source is connected to.
+\paragraph{Multiple Meshes} To analyze a circuit with multiple meshes, write KVL equations for each mesh and KCL equations for each node. There would be times where we add currents via KCL equations.
 

### 2025-01-21 ECE159 Lecture Notes/main.tex
added link to github
gitignore

--- 2025-01-16 ECE159 Lecture Notes/main.tex
@@
 Fundamentals of Electric Circuits, Winter 2024 \hfill ECE 159\\
+\\
+The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
 \vspace{10pt}
@@
 \end{equation}
+\section{DC Circuits}
 \subsection{Circuit Elements}
@@
 \begin{equation}
-    \sum I_{\text{in}} = \sum I_{\text{out}}
+    \sum I_{\text{in}} = \sum I_{\text{out}} \quad \text{or} \quad \sum_\text{$n$ in node} I_n = 0
 \end{equation}
+\paragraph{Node Analysis} Node Analysis is a method used to analyze electric circuits by writing KCL equations at each node.
+\paragraph{Ground Node} Ground is a reference point in a circuit. It is often used as a reference point for voltage.
 \paragraph{Kirchhoff's Voltage Law (KVL)} The sum of voltages around a closed loop is equal to zero.
 \begin{equation}
-    \sum V_{\text{loop}} = 0
+    \sum_\text{$n$ in loop} V_{n} = 0
 \end{equation}
-
-\section{DC Circuits}
+\paragraph{Mesh Analysis} Mesh Analysis is a method used to analyze electric circuits by writing KVL equations for each loop.
+\paragraph{Multiple Meshes} To analyze a circuit with multiple meshes, write KVL equations for each mesh and KCL equations for each node. There would be times where we add currents via KCL equations.
 \paragraph{}

### 2025-01-21 linalgA1/main.tex
gitignore

@@
     \makebox[\textwidth]{\large Full Name: Cheung, Hei Shing\enspace\hrulefill}\\[0.75cm]
-    \makebox[\textwidth]{\large Student number: 101090782\enspace\hrulefill}\\
+    \makebox[\textwidth]{\large Student number: 1010907823\enspace\hrulefill}\\
   \end{minipage}
@@
     \vspace{0.75cm}
-    \makebox[\textwidth]{\large Full Name:\enspace\hrulefill}\\[0.75cm]
-    \makebox[\textwidth]{\large Student number:\enspace\hrulefill}\\
+    \makebox[\textwidth]{\large Full Name: Lin, Zhiyue\enspace\hrulefill}\\[0.75cm]
+    \makebox[\textwidth]{\large Student number: 1010900759\enspace\hrulefill}\\
   \end{minipage}
@@
 It follows from $\mathcal{A}_3: 3a + 2b + c = 5$ and that $\mathcal{B}_2 = \mathcal{A}_2$ and that $\mathcal{B}_1 = \mathcal{A}_1$ are all \textbf{TRUE}.
+Hence $(a,b,c) \in A$.
 Thus, $A \neq \varnothing$, a contradiction.
@@
 It follows from $\mathcal{A}_2 = \mathcal{B}_2$ and that $\mathcal{A}_1 = \mathcal{B}_1$ are all \textbf{TRUE}.
+Hence $(a,b,c) \in B$.
 Thus, $B \neq \varnothing$, a contradiction.
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    $\alpha \in \mathbb{R}$
+    $\alpha \in \mathbb{R}$. It can be any real number.
     %%% end of your answer    
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    $\alpha \neq 0$
+    $\alpha \neq 0$. It can be any non zero real number.
     %%% end of your answer    
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
+    \paragraph{} We have $\mathcal{A}_i = \mathcal{B}_i$ for all $i \neq i_0$ and $\mathcal{B}_{i_0} = \alpha \mathcal{A}_{i_0}$. Given nonempty $A,\, B$, we have:
+    \begin{equation*}
+      \mathcal{A}_{i_0} \text{ is \textbf{TRUE}} \implies \alpha \mathcal{A}_{i_0} \text{ is \textbf{TRUE}} \implies \mathcal{B}_{i_0} \text{ is \textbf{TRUE}}.
+    \end{equation*}
+    Therefore, $\mathcal{A}$ and $\mathcal{B}$ have the same set of solutions.
 
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
-    %%% either, please!)
+    \paragraph{Additve Axioms}
+    \begin{enumerate}
+      \item By assumption, $\V_0$ is closed under vector addition. Thus, for all $\mathbf{u}, \mathbf{v} \in \V_0$, $\mathbf{u} + \mathbf{v} \in \V_0$.
+      \item By assumption, $\V_0$ has the same vector addition as $\V$ and $\V$ is a vetor space.
+Thus, for all $\mathbf{u}, \mathbf{v}, \mathbf{w} \in \V_0$, $(\mathbf{u} + \mathbf{v}) + \mathbf{w} = \mathbf{u} + (\mathbf{v} + \mathbf{w})$.
+      \item By assumption, $\V_0$ contains $\mathbf{0}$ and $\V$ is a vector space. Thus, for all $\mathbf{u} \in \V_0 \subseteq \V$, $\mathbf{u} + \mathbf{0} = \mathbf{u}$.
+      \item For all $\textbf{u} \in \V_0 \subseteq \V$, a vector space. Under the usual scalar multiplication, $0\textbf{u} = \mathbf{0}$. With $(1 + (-1)) = 0$, we have:
+      \begin{equation*}
+        \begin{aligned}
+          1\textbf{u} + (-1)\textbf{u} &= (1 + (-1))\textbf{u} \\
+          &= 0\textbf{u} \\
+          &= \mathbf{0}.
+        \end{aligned}
+      \end{equation*}
+      Since $\V_0$ is closed under scalar multiplication, $(-1)\textbf{u} \in \V_0$. Thus, $\V_0$ contains the additive inverse of $\textbf{u}$.
+      \end{enumerate}
+    \paragraph{Scalar Multiplication Axioms}
+    \begin{enumerate}
+      \item By assumption, $\V_0$ is closed under scalar multiplication. Thus, for all $c \in \mathbb{R}$ and $\textbf{u} \in \V_0$, $c\textbf{u} \in \V_0$.
+      \item By assumption, $\V_0$ has the same scalar multiplication as $\V$ and $\V$ is a vector space.
+Thus, for all $c, d \in \mathbb{R}$ and $\textbf{u} \in \V_0 \subseteq \V$, $(cd)\textbf{u} = c(d\textbf{u})$.
+      \item By assumption, $\V_0$ has the same scalar multiplication as $\V$ and $\V$ is a vector space.
+Thus, for all $c \in \mathbb{R}$ and $\textbf{u}, \textbf{v} \in \V_0 \subseteq \V$, $c(\textbf{u} + \textbf{v}) = c\textbf{u} + c\textbf{v}$.
+      \item By assumption, $\V_0$ has the same scalar multiplication as $\V$ and $\V$ is a vector space. Thus, for all $\textbf{u} \in \V_0 \subseteq \V$, $1\textbf{u} = \textbf{u}$.
+    \end{enumerate}
 
@@
   \choice True
-  \choice False
+  \CorrectChoice False
 \end{checkboxes}
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
-    %%% either, please!)
+    Pick $f \in \V$ such that there exist $x_0 \in [2,3]$, $f(x_0) \neq 0$.
+Choose real number $c<0$, then $cf(x_0) > 0$, hence $c\cdot f \notin \V$.
+Thus, $\V$ is not closed under scalar multiplication.
+Therefore, $\V$ is not a vector space over $\mathbb{R}$.
 
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
-    %%% either, please!)
+    Pick $p, q \in \V$ for which $A, -A \in \mathbb{R}$ such that $p(x) = Ax^n$ and $q(x) = -Ax^n$.
+Then $p(x) + q(x) = A_nx^n + (-A_n)x^n = 0x^n = 0$.
+However, 0 is not a polynomial of degree exactly $n$.
+Thus, $\V$ is not closed under vector addition.
+Therefore, $\V$ is not a vector space over $\mathbb{R}$.
 
@@
 \begin{checkboxes}
-  \choice True
+  \CorrectChoice True
   \choice False
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    %%% run into the next question or off the bottom of the page (and don't get too close to the bottom
-    %%% either, please!)
+    Given that $^2\mathbb{R}^2$ is a vector space, the set of $2 \times 2$ upper triangular real matrices is a subset of $^2\mathbb{R}^2$.
+Thus, we prove that $\V$ is a vector space of $^2\mathbb{R}^2$ by showing subspace test holds.
+    \begin{enumerate}
+      \item $\mathbf{0} \in \V$: $\mathbf{0}$ is an upper triangular matrix with $\textbf{0} = \begin{bmatrix}
+        0 & 0 \\
+        0 & 0
+      \end{bmatrix}$. Given usual entry-wise addition, for all $\textbf{A} \in \V$, $\textbf{A} + \textbf{0} = \begin{bmatrix}
+        a_{11} + 0 & a_{12} + 0 \\
+        0 & a_{22} + 0
+      \end{bmatrix} = \textbf{A}$. Thus, $\mathbf{0} \in \V$.
+      \item Closure under addition: Let $A, B \in \V$. Then $A + B = \begin{bmatrix}
+        a_{11} + b_{11} & a_{12} + b_{12} \\
+        0 & a_{22} + b_{22}
+      \end{bmatrix}$. So $A + B$ is upper triangular, and $A + B \in \V$.
+      \item Closure under scalar multiplication: Let $A \in \V$ and $c \in \mathbb{R}$. Then $cA = \begin{bmatrix}
+        ca_{11} & ca_{12} \\
+        0 & ca_{22}
+      \end{bmatrix}$. So $cA$ is upper triangular, and $cA \in \V$.
+    \end{enumerate}
 

### 2025-01-23 ESC195 Lecture Notes/main.tex
added link to github
jan 22

@@
 Caculus II, Winter 2024 \hfill ESC 195\\
+\\
+The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
+
 \section{More on Integrals}
@@
 \subsection{Surface Area of Revolution}
-\paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, the surface area of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the surface area of the curve:
+\paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, we would like to compute the surface area of the curve between two points $a$ and $b$ when rotated about the $x$-axis:
     $$ A = \int_C 2\pi y \, ds$$
@@
 \end{subequations}
+\paragraph{Riemann Sum} Given a tank rectangular with width defined by $w = f(x)$, we can take a horizontal slice, $x^\star$, of the tank to compute the force exerted on the tank by the water:
+\begin{equation}
+    F =\sum_{i=1}^n \rho g x_i f(x_i) \Delta x
+\end{equation}
 \paragraph{Formula} Given a tank rectangular with width defined by $w = f(x)$, where $x$ is depth, we would like to compute the force exerted on the tank by the water:
@@
 \end{example}
+\subsection{Calculus with Parametric Curves}
+\paragraph{Tangents} The tangent to a parametric curve $C$ at a point $t$ is given by:
+\begin{equation}
+    \frac{dy}{dx} = \frac{dy/dt}{dx/dt} = \frac{y'(t)}{x'(t)}
+\end{equation}
+\paragraph{Equation of the Tangent} The equation of the tangent to a parametric curve $C$ at a point $t_0$ is given by:
+\begin{equation}
+    y'(t_0) (x-x(t_0)) - x'(t_0) (y-y(t_0)) = 0
+\end{equation}
+Hence, if $x'(t_0) = 0$, we have a vertical tangent; if $y'(t_0) = 0$, we have a horizontal tangent. However, if $x'(t_0) = 0$ and $y'(t_0) = 0$, we have no information.
+\begin{example}
+    Let $x = \sin 2t$, $y = \sin t$, we have:
+    \begin{align*}
+        x'(t) &= 2\cos 2t, \\
+        y'(t) &= \cos t
+    \end{align*}
+    \textbf{Vertical Tangent}: $2\cos 2t = 0 \Rightarrow t = \frac{\pi}{4} + \frac{n\pi}{2}$.
+    \textbf{Horizontal Tangent}: $\cos t = 0 \Rightarrow t = \frac{\pi}{2} + n\pi$.
+    \textbf{At $t = 0$} we have $x' = y' = 0$. Thus, we have no information. On the graph, we see that the curve intersects itself at $t = 0$.
+\end{example}
+\paragraph{Area under Parametric Curve} The area under a parametric curve $C$ between $t_1$ and $t_2$ is given by:
+\begin{equation}
+    A = \int^{x(t_2)}_{x(t_1)} y(x) \, dx = \int_{t_1}^{t_2} y(t) x'(t) \, dt 
+\end{equation}
+\begin{definition}[Orientation]
+    The orientation of a parametric curve $C$ is given by the direction of the curve as $t$ increases.
+If the enclosed area is to the left of the trace of $t$ (counter-clockwise) the orientation is positive; otherwise, it is negative.
+\end{definition}
+\paragraph{} The sign of the area of a parametric curve $C$ is given by the orientation of the curve.
+\paragraph{Area of Closed Curves} The area of a closed parametric curve $C$ is given by:
+\begin{equation}
+    A = \oiint_C dA = \int_{t_1}^{t_2} y(t) x'(t) \, dt = \int_{t_1}^{t_2} x(t) y'(t) \, dt
+\end{equation}
+, where
+\begin{equation*}
+    x(t_1) = x(t_2), \quad y(t_1) = y(t_2) \quad \text{and} \quad \text{$t_2$ is the smallest $t>t_1$ to satisfy the condition}.
+\end{equation*}
+\paragraph{Arc Length} The arc length of a parametric curve $C$ is given by:
+\begin{equation}
+    L = \int_a^b \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
+\end{equation}
+\begin{example}
+    Given $x = t\cos t$, $y = t\sin t$, we have:
+    \begin{align*}
+        \frac{dx}{dt} &= \cos t - t\sin t, \\
+        \frac{dy}{dt} &= \sin t + t\cos t
+    \end{align*}
+    Thus, the arc length is given by:
+    \begin{align*}
+        L &= \int_0^{2\pi} \sqrt{(\cos t - t\sin t)^2 + (\sin t + t\cos t)^2} \, dt \\
+        &= \int_0^{2\pi} \sqrt{1 + t^2} \, dt
+        \intertext{Using the substitution $t = \tan \theta$, we have:}
+        &= \pi \sqrt{1 + 4\pi^2} + \frac{1}{2}\ln(2\pi + \sqrt{1 + 4\pi^2})
+    \end{align*}
+\end{example}
+\paragraph{Surface Area of Revolution} The surface area of a parametric curve $C$ rotated about the $x$-axis is given by:
+\begin{align}
+    A &= \int_a^b 2\pi y(t) ds \nonumber \\
+    \intertext{From the arc length formula, we have $ds = \sqrt{(\frac{dx}{dt})^2 + (\frac{dy}{dt})^2} \, dt$, thus:}
+    &= \int_a^b 2\pi y(t) \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
+\end{align}
+Similarly, the surface area of a parametric curve $C$ rotated about the $y$-axis is given by:
+\begin{equation}
+    A = \int_a^b 2\pi x(t) \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
+\end{equation}
+\begin{example}[Surface Area of of a Ellipse]
+    Given $x = a\sin t$, $y = b\cos t$, we have:
+    \begin{align*}
+        \frac{dx}{dt} &= a\cos t, \\
+        \frac{dy}{dt} &= -b\sin t
+    \end{align*}
+    Thus, the surface area of the ellipse is given by:
+    \begin{align*}
+        A &= \int_0^{2\pi} 2\pi b\cos t \sqrt{a^2\cos^2 t + b^2\sin^2 t} \, dt \\
+        &= 2\pi b \int_0^{2\pi} \sqrt{a^2 (1 - \sin^2 t) + b^2 \sin^2 t} \, dt \\
+        &= 2\pi b \int_0^{2\pi} \sqrt{a^2 + b^2 - (a^2 - b^2)\sin^2 t} \, dt
+    \end{align*}
+    This has no analytical solution, but can be solved numerically. Typically, we take $\epsilon = \sqrt{\frac{a^2-b^2}{a^2}}$.
+\end{example}
 

### 2025-01-23 MAT185 Lecture Notes/main.tex
added link to github
jan 22

@@
 Linear Algebra, Winter 2024 \hfill MAT 185\\
+\\
+The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
 \section{Vector Space}
+\subsection{Foundamental Properties}
 \begin{definition}[Vector Space in $\mathbb{R}$]
@@
 \begin{definition}[Linear independence]
-    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If the equation:
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If and only if the equation:
     \begin{align*}
@@
     \end{align*}
-    has only the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.
+    has \textbf{only} the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.
 
@@
 \end{definition}
+
 \begin{definition}[Span]
@@
 \end{definition}
+\begin{definition}[Basis]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$.
+If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent and $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = V$, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ is a basis of $V$.
+\end{definition}
 \end{document}

### 2025-01-23 MSE160 Lecture Notes/main.tex
added link to github
jan 22

@@
 Molecules and Materials, Winter 2024 \hfill MSE 160\\
+\\
+The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
 \begin{center}
@@
     \paragraph{Deflection} The deflection of the beam is given by:
+    \begin{equation}
+        \delta = \frac{FL^3}{48EI} 
+    \end{equation}
     \begin{align*} 
-        \delta &= \frac{FL^3}{48EI} 
         \intertext{For a beam with a rectangular cross-section, we have:}
@@
 This is done by cooling the surface of the glass faster than the core or chemically treating the surface.
-\appendix
+\subsection{Density}
+\paragraph{Density} The density of a material is given by:
+\begin{equation}
+    \rho = \frac{m}{V}
+\end{equation}
+\paragraph{Archimedes' Principle} The buoyant force on an object is equal to the weight of the fluid displaced by the object. Derive from that, we have:
+\begin{equation}
+    \rho = \frac{m}{V} = \frac{m_{\text{object}}}{m_{\text{object}} - m_{\text{object in fluid}}}
+\end{equation}
+\paragraph{Ordered structures} We have the following three orders:
+\begin{enumerate}
+    \item \textbf{Long Range Order (LRO)} Atoms are arranged in a well-defined pattern over long distances in repeating units. \\
+    (Example) Diamonds, some polymers, most metals, many ceramics, graphite, quartz, etc.
+    \item \textbf{Short Range Order (SRO)} Periodic arrangement of atoms over a few atomic or molecular spacings. \\
+    (Example) Most polymers, glasses, amorphous materials, etc.
+    \item \textbf{No Order (NO)} Atoms are randomly arranged. \\
+    (Example) Ideal gases, etc.
+\end{enumerate}
+\paragraph{Describing Crystal Structures} We can describe crystal structures by the following:
+\begin{itemize}
+    \item \textbf{Unit Cell} The smallest repeating unit of a crystal structure that could be used to represent the entire crystal.
+    \item \textbf{Lattice} The repeating arrangement of points that repersent the positions of atoms in the unit cell.\\
+    (Example) Simple Cubic, Body-Centered Cubic, Face-Centered Cubic, etc.
+    \item \textbf{Lattice Plus Basis} \textit{Not required for this course}
+\end{itemize}
+\paragraph{Theortical Density of Metals} The theoretical density of metals is given by:
+\begin{equation}
+    \rho = \frac{nM}{V_c N_A}
+\end{equation}
+, where:
+\begin{itemize}
+    \item $n$ is the number of atoms in the unit cell.
+    \item $M$ is the atomic mass of the element (amu = g/mol).
+    \item $V_c$ is the volume of the unit cell.
+    \item $N_A$ is Avogadro's number ($6.022 \times 10^{23} \text{mol}^{-1}$)
+\end{itemize}
+\paragraph{Simple Cubic} The simple cubic structure has the following properties:
+\begin{itemize}
+    \item \textbf{Centers of atoms} Located at the eight corners of a cube
+    \item \textbf{Packing Density} Rare due to low packing density (only known example: Polonium, Po)
+    \item \textbf{Close-packed directions} As cube edges.
+    \item \textbf{Coordination Number} Number of nearest neighbor
+\end{itemize}
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-01-23 linalgA1/main.tex
jan 22

@@
 For each elementary row operation, the set of solutions of $\mathcal{A}$ remains unchanged.
+This is proven in the previous parts.
     Thus, $\mathcal{A}$ and $\mathcal{R}$ have the same set of solutions.
@@
       \item By assumption, $\V_0$ has the same scalar multiplication as $\V$ and $\V$ is a vector space.
-Thus, for all $c \in \mathbb{R}$ and $\textbf{u}, \textbf{v} \in \V_0 \subseteq \V$, $c(\textbf{u} + \textbf{v}) = c\textbf{u} + c\textbf{v}$.
+Thus, for all $c \in \mathbb{R}$ and $\textbf{u}, \textbf{v} \in \V_0 \subseteq \V$, $(c + d)\textbf{u} = c\textbf{u} + d\textbf{u}$.
       \item By assumption, $\V_0$ has the same scalar multiplication as $\V$ and $\V$ is a vector space. Thus, for all $\textbf{u} \in \V_0 \subseteq \V$, $1\textbf{u} = \textbf{u}$.
@@
     %%% Your work goes here!  It is your responsibility to trim down your answer so that it doesn't 
-    Given that $^2\mathbb{R}^2$ is a vector space, the set of $2 \times 2$ upper triangular real matrices is a subset of $^2\mathbb{R}^2$.
-Thus, we prove that $\V$ is a vector space of $^2\mathbb{R}^2$ by showing subspace test holds.
+    We have the following:
+    \paragraph{Additive Axioms}
     \begin{enumerate}
-      \item $\mathbf{0} \in \V$: $\mathbf{0}$ is an upper triangular matrix with $\textbf{0} = \begin{bmatrix}
+      \item For all $A, B \in \V$, $A + B = \begin{bmatrix}
+        a_{11} + b_{11} & a_{12} + b_{12} \\
+        0 & a_{22} + b_{22}
+      \end{bmatrix}$ is upper triangular.
+      \item For all $A, B, C \in \V$, for $(A+B)+C$, we have:
+      \begin{equation*}
+      \begin{aligned}
+        (A+B)+C &= \begin{bmatrix}
+          a_{11} + b_{11} & a_{12} + b_{12} \\
+          0 & a_{22} + b_{22}
+        \end{bmatrix} + C \\
+        &= \begin{bmatrix}
+          a_{11} + b_{11} + c_{11} & a_{12} + b_{12} + c_{12} \\
+          0 & a_{22} + b_{22} + c_{22}
+        \end{bmatrix} 
+        = A + \begin{bmatrix}
+          b_{11} + c_{11} & b_{12} + c_{12} \\
+          0 & b_{22} + c_{22}
+        \end{bmatrix} \\
+        &= A + (B+C).
+      \end{aligned}
+    \end{equation*}
+      \item $\mathbf{0}$ is upper triangular. We take $\textbf{0} = \begin{bmatrix}
         0 & 0 \\
         0 & 0
-      \end{bmatrix}$. Given usual entry-wise addition, for all $\textbf{A} \in \V$, $\textbf{A} + \textbf{0} = \begin{bmatrix}
+      \end{bmatrix}$. For all $A \in \V$, $A + \textbf{0} = \begin{bmatrix}
         a_{11} + 0 & a_{12} + 0 \\
         0 & a_{22} + 0
-      \end{bmatrix} = \textbf{A}$. Thus, $\mathbf{0} \in \V$.
-      \item Closure under addition: Let $A, B \in \V$. Then $A + B = \begin{bmatrix}
-        a_{11} + b_{11} & a_{12} + b_{12} \\
-        0 & a_{22} + b_{22}
-      \end{bmatrix}$. So $A + B$ is upper triangular, and $A + B \in \V$.
-      \item Closure under scalar multiplication: Let $A \in \V$ and $c \in \mathbb{R}$. Then $cA = \begin{bmatrix}
+      \end{bmatrix} = A$.
+      \item For all $A \in \V$, We take another upper triangular matrix $-A = \begin{bmatrix}
+        -a_{11} & -a_{12} \\
+        0 & -a_{22}
+      \end{bmatrix}$. For all $A \in \V$, $A + (-A) = \begin{bmatrix}
+        a_{11} - a_{11} & a_{12} - a_{12} \\
+        0 & a_{22} - a_{22}
+      \end{bmatrix} = \textbf{0}$.
+    \end{enumerate}
+    \paragraph{Scalar Multiplication Axioms}
+    \begin{enumerate}
+      \item For all $A \in \V$, $c \in \mathbb{R}$, $cA = \begin{bmatrix}
         ca_{11} & ca_{12} \\
         0 & ca_{22}
-      \end{bmatrix}$. So $cA$ is upper triangular, and $cA \in \V$.
+      \end{bmatrix}$ is upper triangular. So $cA \in \V$.
+      \item For all $A \in \V$, $c(dA) = \begin{bmatrix}
+        cda_{11} & cda_{12} \\
+        0 & cda_{22}
+      \end{bmatrix} = (cd)A$.
+      \item For all $A \in \V$, $(c+d)A = \begin{bmatrix}
+        (c+d)a_{11} & (c+d)a_{12} \\
+        0 & (c+d)a_{22} \end{bmatrix} = \begin{bmatrix}
+          ca_{11} & ca_{12} \\
+          0 & ca_{22}
+        \end{bmatrix} + \begin{bmatrix}
+          da_{11} & da_{12} \\
+          0 & da_{22}
+        \end{bmatrix} = cA + dA$.
+      \item For all $A \in \V$, $1A = \begin{bmatrix}
+        1a_{11} & 1a_{12} \\
+        0 & 1a_{22}
+      \end{bmatrix} = A$.
     \end{enumerate}

### 2025-01-23 ESC195 Lecture Notes/main.tex
jan23

@@
 \end{example}
+\subsection{Convergence Test}
+\paragraph{Convergence} An improper integral converges if the limit of the integral exists. 
+\paragraph{Comparison Test} We can compare the integral to another integral to determine convergence or divergence.
+\paragraph{Spliting the Integral} We can split the integral into two parts to determine convergence. We split it into two limits on the left and right, for example:
+$$ \int_a^b f(x)\, dx = \lim_{t \to c^-} \int_a^t f(x) \, dx + \lim_{t \to c^+} \int_t^b f(x) \, dx $$
 \section{Hyperbolic Trigonometric Functions}
@@
 \end{example}
+\paragraph{Inverse} Also, we can take $x = f^{-1}(y)$ to compute the arc length of a curve.
+\begin{equation}
+    L = \int_{f(a)}^{f(b)} \sqrt{1 + ((f{^{-1}})'(y))^2} \, dy
+\end{equation}
 \subsection{Surface Area of Revolution}
 \paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, we would like to compute the surface area of the curve between two points $a$ and $b$ when rotated about the $x$-axis:
-    $$ A = \int_C 2\pi y \, ds$$
-\paragraph{Formula} Deriving the formula for the surface area of a curve $y = f(x)$ between two points $a$ and $b$ is similar to the arc length formula:
+    $$ A_x = \int_C 2\pi y \, ds$$
+\paragraph{Formula} Deriving the formula for the surface area of a curve $y = f(x)$ between two points $a$ and $b$ rotated about the $x$-axis as follows:
 \begin{align}
@@
     \intertext{As the difference between $y^\star$ and $y_i$ diminishes as $\Delta x \rightarrow 0$, taking a Riemann sum:}
-    A &= \lim_{n \to \infty} \sum_{i=1}^n 2\pi y_i \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
+    A_x &= \lim_{n \to \infty} \sum_{i=1}^n 2\pi y_i \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
     &= \int_a^b 2\pi y \sqrt{1 + y'(x)^2} \, dx
+    \intertext{The surface area of a curve $y = f(x)$ (i.e. $x = f^{-1}$)between two points $f(a)$ and $f(b)$ rotated about the $y$-axis is:}
+    A_y &= \int_{f(a)}^{f(b)} 2\pi x \sqrt{1 + x'(y)^2} \, dy
 \end{align}
@@
     \begin{align*}
-        A &= \int_0^1 2\pi \sqrt{x} \sqrt{1 + \frac{1}{4x}} \, dx \\
+        A_x &= \int_0^1 2\pi \sqrt{x} \sqrt{1 + \frac{1}{4x}} \, dx \\
         &= \pi \int^1_0 \sqrt{4x + 1} \, dx \\
@@
 \begin{equation}
-    A = \oiint_C dA = \int_{t_1}^{t_2} y(t) x'(t) \, dt = \int_{t_1}^{t_2} x(t) y'(t) \, dt
+    A = \oint_C dA = \int_{t_1}^{t_2} y(t) x'(t) \, dt = \int_{t_1}^{t_2} x(t) y'(t) \, dt
 \end{equation}

### 2025-01-28 ESC195 Lecture Notes/main.tex
jan 28

@@
 
-$$
-\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i
-$$
-
-where:
+$$\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i$$
+, where:
 \begin{itemize}
@@
 Rearranging this:
-$$
-\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.
-$$
+$$\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.$$
 
@@
 \end{example}
-
+\section{Polar Coordinates}
+\paragraph{Polar Coordinates} Given a point $P$ in the plane, we can define the polar coordinates of $P$ as $(r, \theta)$, where $r$ is the distance from the origin to $P$ and $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to $P$.
+\paragraph{Transformation} We can convert between polar and Cartesian coordinates as follows:
+\begin{equation}
+    x = r\cos \theta, \quad y = r\sin \theta
+\end{equation}
+\paragraph{Reverse Transformation} We can convert between Cartesian and polar coordinates as follows:
+\begin{equation}
+    r = \sqrt{x^2 + y^2}, \quad \theta = \arctan\left(\frac{y}{x}\right)
+\end{equation}
+\paragraph{Note} The angle $\theta$ is not unique, as $\theta + 2\pi n$ for $n \in \mathbb{Z}$ represents the same point, beware of this when converting between polar and Cartesian coordinates.
+\begin{example}[Lines]
+    \begin{align*}
+        y = mx + b &\implies \theta = \alpha = \arctan m \\
+        x = a &\implies r = a\sec \theta \\
+        y = a &\implies r = a\csc \theta
+    \end{align*}
+\end{example}
+\begin{example}[Cirlces]
+    \begin{align*}
+        x^2 + y^2 = a^2 &\implies r = a \\
+    \end{align*}
+    Let $r = 6\sin \theta$, we have:
+    \begin{align*}
+        r^2 &= 6r\sin \theta \\
+        x^2 + y^2 &= 6y \\
+        x^2 + (y-3)^2 &= 9
+    \end{align*}
+    We deduce that the curve is a circle with radius $3$ and center $(0,3)$.
+\end{example}
+\paragraph{Cylinrical Coordinates} Given a point $P$ in space, we can define the cylindrical coordinates of $P$ as $(r, \theta, z)$, where $r$ is the distance from the $z$-axis to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $z$ is the distance from the $xy$-plane to $P$.
+\paragraph{Spherical Coordinates} Given a point $P$ in space, we can define the spherical coordinates of $P$ as $(\rho, \theta, \phi)$, where $\rho$ is the distance from the origin to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to $P$.
+\subsection{Graphing in Polar Coordinates}
+\begin{example}
+    Let $r = \frac{1}{2} + \cos \theta$, we first figure out the origin of the curve by setting $r = 0$:
+    \begin{align*}
+        0 &= \frac{1}{2} + \cos \theta \\
+        \theta &= \frac{2\pi}{3},\, \frac{4\pi}{3}
+    \end{align*}
+    We then deduce the maximum and minimum of the curve to the origin by setting $\frac{dr}{d\theta} = 0$:
+    \begin{align*}
+        \frac{dr}{d\theta} &= -\sin \theta = 0 \\
+        \theta &= 0,\, \pi,\, 2\pi
+    \end{align*}
+    We also look for symmetry in the curve by checking $r(\theta) = r(-\theta)$:
+    \begin{align*}
+        \frac{1}{2} + \cos \theta &= \frac{1}{2} + \cos(-\theta) \\
+    \end{align*}
+    Thus, the curve is symmetric about the $x$-axis. We also look for symmetry about the $y$-axis by checking $r(\theta) = r(\pi + \theta)$:
+    \begin{align*}
+        \frac{1}{2} + \cos \theta \neq \frac{1}{2} + \cos(\pi + \theta) \\
+    \end{align*}
+    Thus, the curve is not symmetric about the $y$-axis. We also look for intervals in which the curve is increasing or decreasing by checking $r'(\theta) > 0$ or $r'(\theta) < 0$.
+\end{example}
+\paragraph{Common Polar Curves} Some common polar curves are:
+\begin{enumerate}
+    \item \textbf{Cardioid}: $r = a(1 + \cos \theta)$ (heart-shaped)
+    \item \textbf{Circle}: $r = a\cos \theta$ 
+    \item \textbf{Limascons}: $r = a + b\sin \theta$ (heart with a hole)
+    \item \textbf{Lemniscate}: $r^2 = a^2\sin 2\theta$ (infinity symbol, diagonal) \vspace{1em} $\quad r^2 = a^2\cos 2\theta$ (infinity symbol, horizontal)
+    \item \textbf{Petal Curves}: $r = a\cos n\theta$ ($mn$-petal flower) \vspace{1em} $\quad r = a\sin n\theta$ ($mn$-petal flower); $m=2$ if $n$ is even, $m=1$ if $n$ is odd.
+\end{enumerate}
+\subsection{The intersection of Polar Curves}
+\begin{example}
+    Let $r_1 = \sin \theta$, $r_2 = -\cos \theta$, we have:
+    \begin{align*}
+        \sin \theta &= -\cos \theta \\
+        \tan \theta &= -1 \\
+        \theta &= \frac{3\pi}{4},\, \frac{7\pi}{4}
+    \end{align*}
+    At these points, we can find the $x$ and $y$ coordinates of the intersection points.
+But also, we must check for the origin point $(x,y) = (0,0)$.
+Also $r_1 = r_2$ is not always reliable so we have to check it intuitively.
+\end{example}
 \appendix

### 2025-01-28 MAT185 Lecture Notes/main.tex
jan 28

@@
 
-\subsection{Linear combinations}
+\section{Linear combinations}
 \begin{definition}[Linear Combination]
@@
 \end{definition}
+\subsection{Linear Independence, Span, and Basis}
 \begin{definition}[Linear independence]
@@
 \end{definition}
-
 \begin{definition}[Span]
@@
 \end{definition}
+\begin{theorem}[Unique Representation]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}/0$, and that they span $V$.
+For any $\textbf{v}_{n+1} \in V$, if $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent, then the equation:
+    \begin{align*}
+        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{v}_{n+1}  
+    \end{align*} 
+    is the only way to express $\textbf{v}_{n+1}$ as a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$.
+\end{theorem}
+We can prove by contrapositive:
+\begin{proof}
+    Assume that there exists another way to express $\textbf{v}_{n+1}$ as a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$, then:
+    \begin{align*}
+        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \beta_1 \textbf{v}_1 + \beta_2 \textbf{v}_2 + \ldots + \beta_n \textbf{v}_n
+    \end{align*}
+    Rearranging the equation, we have:
+    \begin{align*}
+        (\alpha_1 - \beta_1) \textbf{v}_1 + (\alpha_2 - \beta_2) \textbf{v}_2 + \ldots + (\alpha_n - \beta_n) \textbf{v}_n = \textbf{0}
+    \end{align*}
+    Since ($\alpha_i - \beta_i$) for $i = 1, 2, \ldots, n$ are not all zero, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly dependent. Thus, the contrapositive is true.
+\end{proof}
+\begin{theorem}[Growing and Pruning] The following are the two theorems describe the relationship between linear independence and the span: \\ \\
+    \textbf{Growing} Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ be linearly independent.
+If $\textbf{v}_{n+1} \in V$ and $\textbf{v}_{n+1} \notin \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n)$, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n, \textbf{v}_{n+1}$ are linearly independent.
+\\ \\
+    \textbf{Pruning} Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$.
+If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly dependent, then there exists a $\textbf{v}_i$ such that $\textbf{v}_i \in \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_{i-1}, \textbf{v}_{i+1}, \ldots, \textbf{v}_n)$.
+\end{theorem}
 \begin{definition}[Basis]
@@
 \end{definition}
+\subsection{Fundamental Theorem of Linear Algebra}
+\begin{theorem}[Fundamental Theorem of Linear Algebra]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$.
+If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent and $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = V$, then:
+    \begin{enumerate}
+        \item Any basis of $V$ has the same number of elements.
+        \item Any linearly independent set of $V$ has at most $n$ elements.
+        \item Any spanning set of $V$ has at least $n$ elements.
+    \end{enumerate}
+\end{theorem}
 \end{document}

### 2025-01-28 MSE160 Lecture Notes/main.tex
jan 28

@@
 \section{Seleection of Materials}
+\subsection{Material Performance}
 \begin{example}[Aircraft Wing Spar]
@@
 \end{equation}
+\section{Atomic Structures}
+\subsection{General Definitions}
 \paragraph{Ordered structures} We have the following three orders:
@@
 \paragraph{Theortical Density of Metals} The theoretical density of metals is given by:
-\begin{equation}
+\begin{equation} \label{eq:theoretical-density}
     \rho = \frac{nM}{V_c N_A}
@@
 \end{itemize}
-\paragraph{Simple Cubic} The simple cubic structure has the following properties:
+\subsection{Describing Basic Unit Cells}
+\begin{definition}[Coordination Number (CN)]
+    The coordination number is the number of nearest neighbors an atom has in a crystal structure.
+\end{definition}
+\begin{definition}[Atomic Packing Factor (APF)]
+    The atomic packing factor is the fraction of the volume of the unit cell that is occupied by hard spheres. It is given by:
+    \begin{equation}
+        \text{APF} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell}}
+    \end{equation}
+\end{definition}
+\paragraph{Simple Cubic (SC)} The simple cubic structure has the following properties:
 \begin{itemize}
     \item \textbf{Centers of atoms} Located at the eight corners of a cube
-    \item \textbf{Packing Density} Rare due to low packing density (only known example: Polonium, Po)
+    \item \textbf{Rare Packing Density}  Due to low packing density (only known example: Polonium, Po)
     \item \textbf{Close-packed directions} As cube edges.
-    \item \textbf{Coordination Number} Number of nearest neighbor
+    \item \textbf{Coordination Number} Number of nearest neighbor: 6
 \end{itemize}
+\paragraph{Face-Centered Cubic (FCC)} The face-centered cubic structure has an atom on the centre of each face of the cube. It has the following properties:
+\begin{itemize}
+    \item \textbf{Ductile} and \textbf{Malleable} due to the close-packed planes. Common examples include: \\ $\ch{Al}\, ,\ch{Cu}\, ,\ch{Au}\, ,\ch{Ag}$.
+    \item \textbf{Coordination Number} Number of nearest neighbor: 12
+    \item \textbf{Close-packed directions} As cube diagonals.
+    \item \textbf{Theortical Density} The theoretical density of FCC is derived from Equation (\ref{eq:theoretical-density}):
+    \begin{align}
+        \rho &= \frac{nM}{V_c N_A} \nonumber \\
+        \intertext{With $n = 4$ (8 corners $\times \frac{1}{8}$ atom each + 6 faces $\times \frac{1}{2}$ atom each) and $V_c = a^3$ (where $a$ is the \textbf{lattice parameter}):}
+        \rho_\text{FCC} &= \frac{4M}{a^3 N_A} \nonumber \\
+        \intertext{We can then obtain $a$ via the relationship between the atomic radius, $r$, and the lattice parameter via the geometry on the close-packed direction (face diagonal):}
+        a &= 2\sqrt{2}r \nonumber \\
+        \intertext{Substitute $a$ back into the equation:}
+        \rho_\text{FCC} &= \frac{4M}{(2\sqrt{2}r)^3 N_A} = \frac{4M}{16\sqrt{2}r^3 N_A} \nonumber \\
+        &= \frac{M}{4\sqrt{2}r^3 N_A} 
+    \end{align}
+    \item \textbf{Atomic Packing Factor} The atomic packing factor of FCC is given by:
+    \begin{equation}
+        \text{APF}_\text{FCC} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{4 \times \frac{4}{3}\pi r^3}{a^3} = \frac{4 \times \frac{4}{3}\pi r^3}{(2\sqrt{2}r)^3} = \frac{\pi}{6\sqrt{2}} = \textbf{0.74}
+    \end{equation}
+    This is the highest possible packing factor for spheres. Thus, no other structure can have a higher packing factor.
+\end{itemize}
+\begin{example}[Density of Aluminum]
+    Given that the atomic radius of aluminum is $r = 143$ pm and the atomic mass is $M = 26.98$ g/mol, we can calculate the density of aluminum with FCC structure. We have:
+    \begin{align*}
+        \rho_\text{Al} &= \frac{M}{4\sqrt{2}r^3 N_A} = \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} \\
+        &= \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} = 2.7 \text{ g/cm}^3
+    \end{align*}
+\end{example}
+\section{Geometric Descriptions of Crystals}
+\subsection{Crystallographic Directions and Planes}
+\paragraph{Point Coordinates} A lattice positon in a unit cell is determined as fractional multiples of the unit cell edge lengths $a$, $b$, and $c$.
+(Defining the origin at one corner of the unit cell and $a=b=c=\textbf{1}$)
+\paragraph{Crystallographic Directions} A crystallographic direction is a line between two points in a crystal.
+It is denoted by square brackets, $[uvw]$.
+The direction is determined from the ``tail'' to the ``head'' of the vector.
+Also:
+\begin{enumerate}
+    \item It \textbf{shall not} include commas; and
+    \item It \textbf{shall be reduced}to the smallest integers; and
+    \item Negative signs are represented by a bar over the number.
+\end{enumerate}
+\begin{example}
+    The crystallographic direction from $(0,0,0)$ to $(-1,0,1/2)$ is $[\bar{2}01]$.
+\end{example}
+\paragraph{Family of Directions} A family of directions is a set of directions that are parallel to each other. It is denoted by $<uvw>$.
+\begin{example}
+    The family of directions $<100>$ is the set of directions $[100]$, $[010]$, $[001]$, $[\bar{1}00]$, $[0\bar{1}0]$, and $[00\bar{1}]$.
+\end{example}
+\paragraph{Angle between Directions} The angle between two directions in a \textbf{cubic} crystal is given by:
+\begin{equation}
+    \cos \theta = \frac{h_1h_2 + k_1k_2 + l_1l_2}{\sqrt{h_1^2 + k_1^2 + l_1^2}\sqrt{h_2^2 + k_2^2 + l_2^2}}
+\end{equation}
+, which is simply the normalized dot product of the two vectors.
+\paragraph{Crystallographic Planes} A crystallographic plane is a set of parallel planes in a crystal.
+It is denoted by $(hkl)$.
+The plane is determined by the intercepts on the $x$, $y$, and $z$ axes.
+\paragraph{} Generally, we find the intercepts by finding the points where the plane intersects the axes.
+We then take the reciprocals of the intercepts and multiply by the smallest integer to get the Miller Indices.
+\paragraph{Family of Planes} A family of planes is a set of planes that are parallel to each other. It is denoted by $\{hkl\}$.
+\paragraph{Distance between Planes} The distance between two planes with miller indices $(h_1k_1l_1)$ and $(h_2k_2l_2)$ is given by:
+\begin{equation}
+    d = \frac{a}{\sqrt{h^2 + k^2 + l^2}}
+\end{equation}
+, where $a$ is the lattice parameter.
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-01-28 ESC195 Lecture Notes/main.tex
jan 28

@@
 \end{example}
+\subsection{Area and Length in Polar Coordinates}
+\paragraph{Area in Polar Coordinates} The area of a region $R: r = g(\theta)$ in polar coordinates is given by:
+\begin{align}
+    A &\approx \sum_{i} \frac{1}{2} [g(\theta_i)]^2 \Delta \theta_i \nonumber \\
+    \intertext{Taking a Riemann sum:}
+    &= \int_{\alpha}^{\beta} \frac{1}{2} [g(\theta)]^2 \, d\theta
+\end{align}
+\begin{example}
+    Given $r = 1 + \cos \theta$, we have:
+    \begin{align*}
+        A &= \int_0^{2\pi} \frac{1}{2} (1 + \cos \theta)^2 \, d\theta \\
+        &= \int_0^{2\pi} \frac{1}{2} (1 + 2\cos \theta + \cos^2 \theta) \, d\theta \\
+        &= \frac{1}{2}\int_0^{2\pi} 2 + 2\cos \theta + \frac{1 + \cos 2\theta}{2} \, d\theta \\
+        &= \frac{1}{2}\left(2\pi + 2\sin \theta + \frac{\theta}{2} + \frac{\sin 2\theta}{4}\right) \Big|_0^{2\pi}= \frac{3\pi}{2}
+    \end{align*}
+\end{example}
+\paragraph{Area between Two Polar Curves} The area between two polar curves $r^2 = f(\theta)$ and $r = g(\theta)$ is given by:
+\begin{equation}
+    A = \int_{\alpha}^{\beta} \frac{1}{2} [f(\theta)^2 - g(\theta)^2] \, d\theta
+\end{equation}
+\begin{example}
+    Given lemniscate $r^2 = 4\cos 2\theta$, circle $r = 1$, to find the area between the two curves, outside the circle, and inside the lemniscate. We have:
+    \begin{align*}
+        \intertext{We first consider the intersection points}
+        4\cos 2\theta &= 1 \\
+        \cos 2\theta &= \frac{1}{4} \\
+        \intertext{We have points at $\theta = \pm 0.659$. Also, since the lemniscate is symmetric about the $y$-axis, we have:}
+        \frac{1}{2}A &= \int_{-0.659}^{0.659} 4\cos^2 2\theta - 1 \, d\theta \\
+        A &= 2.554
+    \end{align*}
+\end{example}
+\begin{example}
+    Given $r = \sin \theta$, $r = \cos \theta$, we have:
+    \begin{align*}
+        \intertext{We first consider the intersection points}
+        \sin \theta &= \cos \theta \\
+        \tan \theta &= 1 \\
+        \theta &= \frac{\pi}{4}
+        \intertext{We then consider the area between the two curves, notice that (0, 0) is also an intersection point.}
+        A &= \frac{1}{2}\int_0^{\frac{\pi}{4}}  (\sin \theta)^2 - (\cos \theta)^2 \, d\theta \\
+        &= \frac{\pi - 2}{8}
+    \end{align*}
+\end{example}
+\paragraph{Tangent Lines in Polar Coordinates} The tangent line to a polar curve $r = r(\theta)$ is derived as follows:
+\begin{align}
+    x(\theta) &= r(\theta)\cos \theta, \quad y(\theta) = r(\theta)\sin \theta \nonumber \\
+    \intertext{By chain rule, we have and product rule, we have:}
+    \frac{dy}{dx} &= \frac{r'(\theta)\sin \theta + r(\theta)\cos \theta}{r'(\theta)\cos \theta - r(\theta)\sin \theta} \\
+\end{align}
+\paragraph{Length of a Polar Curve} Derived from above, the length of a polar curve $r = r(\theta)$ between $\alpha$ and $\beta$ is given by:
+\begin{equation}
+    L = \int_{\alpha}^{\beta} \sqrt{[r(\theta)]^2 + [r'(\theta)]^2} \, d\theta
+\end{equation}
+\begin{example}
+    Given $r = a - a\cos \theta$, we have:
+    \begin{align*}
+        r' &= a\sin \theta \\
+        L &= \int_0^{2\pi} \sqrt{(a - a\cos \theta)^2 + a^2\sin^2 \theta} \, d\theta \\
+        &= a\int_0^{2\pi} \sqrt{2 - 2\cos \theta} \, d\theta \\
+        \intertext{We can use the double angle formula to simplify the integral, with $\cos 2\theta = 1 - 2\sin^2 \theta$:}
+        &= 8a
+    \end{align*}
+\end{example}
+\section{Infinite Sequences and Series}
+\subsection{Sequences}
+\paragraph{Sequences} A sequence is a list of numbers written in a specific order.
+\begin{example}
+    Let $f: A \to \mathbb{R}$ be a function, and $f(x) = \frac{1}{x}$.
+We can restrict the domain of $f$ to $A = \{1, 2, 3, \ldots\} = \mathbb{N}$, and we can define the sequence $a_n = f(n) = \frac{1}{n}$.
+\end{example}
+\paragraph{Notation} Typcially, we denote a sequence as $\{a_n\}$, where $a_n$ is the $n$th term of the sequence.
+We can also denote with curly bracket as $\{a_n\} = \{a_1, a_2, a_3, \ldots\} = \{f(n)\}$.
+\paragraph{Ratio} The ratio of two sequences $\{a_n\}$ and $\{b_n\}$ for some $n \le N$ is a sequence $\{c_n\}$ such that $c_n = \frac{a_n}{b_n}$.
+\begin{example}
+    Let $a_n = n^2$, $b_n = e^n$, we have:
+    $$\frac{a_n}{b_n} = \{\frac{1}{e}, \frac{4}{e^2}, \frac{9}{e^3}, \ldots\}$$
+\end{example}
+\paragraph{Bounded Above} A sequence $\{a_n\}$ is said to be bounded above if there exists a number $M$ such that $|a_n| \le M$ for all $n$. (Similarly for bounded below).
+\begin{definition}[Monotonic Sequences]
+    A sequence $\{a_n\}$ is said to be non decreasing if $a_{n+1} \ge a_n$ for all $n$.
+Similarly, a sequence is said to be non increasing if $a_{n+1} \le a_n$ for all $n$.
+A sequence is said to be increasing if $a_{n+1} > a_n$ for all $n$.
+Similarly, a sequence is said to be decreasing if $a_{n+1} < a_n$ for all $n$.
+\\ A sequence $\{a_n\}$ is said to be monotonic if it is either increasing, decreasing, non increasing, or non decreasing.
+\end{definition}
+\paragraph{Convergent Sequences} A sequence $\{a_n\}$ is said to be convergent if there exists a number $L$ such that for all $\epsilon > 0$, there exists an $N$ such that $|a_n - L| < \epsilon$ for all $n \ge N$.
 \appendix

### 2025-01-30 ESC195 Lecture Notes/main.tex
none

@@
 \begin{align}
-    \cosh^2(x) + \sinh^2(x) &= 1, \\
+    \cosh^2(x) - \sinh^2(x) &= 1, \\
     \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
@@
     \item \textbf{Circle}: $r = a\cos \theta$ 
-    \item \textbf{Limascons}: $r = a + b\sin \theta$ (heart with a hole)
+    \item \textbf{Limasçons}: $r = a + b\sin \theta$ (heart with a hole)
     \item \textbf{Lemniscate}: $r^2 = a^2\sin 2\theta$ (infinity symbol, diagonal) \vspace{1em} $\quad r^2 = a^2\cos 2\theta$ (infinity symbol, horizontal)
@@
 \end{enumerate}
+\paragraph{} Below are some examples of polar curves:
+\begin{figure}[h!]
+    \centering
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, red] plot ({(1+cos(\x))*cos(\x)}, {(1+cos(\x))*sin(\x)});
+            \node at (0,-3) {Cardioid: $r = 1 + \cos \theta$};
+        \end{tikzpicture}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, blue] plot ({cos(\x)*cos(\x)}, {cos(\x)*sin(\x)});
+            \node at (0,-3) {Circle: $r = \cos \theta$};
+        \end{tikzpicture}
+    \end{minipage}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, green] plot ({(1+0.5*sin(\x))*cos(\x)}, {(1+0.5*sin(\x))*sin(\x)});
+            \node at (0,-3) {Limaçon: $r = 1 + 0.5\sin \theta$};
+        \end{tikzpicture}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, orange] plot ({sqrt(abs(sin(2*\x)))*cos(\x)}, {sqrt(abs(sin(2*\x)))*sin(\x)});
+            \node at (0,-3) {Lemniscate: $r^2 = \sin 2\theta$};
+        \end{tikzpicture}
+    \end{minipage}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, purple] plot ({cos(3*\x)*cos(\x)}, {cos(3*\x)*sin(\x)});
+            \node at (0,-3) {Petal Curve: $r = \cos 3\theta$};
+        \end{tikzpicture}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.48\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=1.5]
+            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
+            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
+            \draw[domain=0:360, smooth, samples=100, brown] plot ({sin(2*\x)*cos(\x)}, {sin(2*\x)*sin(\x)});
+            \node at (0,-3) {Petal Curve: $r = \sin 2\theta$};
+        \end{tikzpicture}
+    \end{minipage}
+\end{figure}
+\newpage
+\paragraph{} \label{sec:polar} A detailed list of polar curves and their properties could be found in Appendix \ref{app:polarCurves}. 
 \subsection{The intersection of Polar Curves}

### 2025-01-30 MSE160 Lecture Notes/main.tex
jan 30

@@
     This is the highest possible packing factor for spheres. Thus, no other structure can have a higher packing factor.
+    \item \textbf{Stacking} The FCC structure can be thought of as stacking close-packed planes.
+The stacking sequence is ABCABC...
+(where A, B, and C describe the orientation hexagonal close-packed planes).
+\end{itemize}
+\paragraph{Hexagonal Close-Packed (HCP)} The hexagonal close-packed structure is \textbf{crytal structure} by stacking close-packed hexagonal planes. It has the following properties:
+\begin{itemize}
+    \item \textbf{Coordination Number} Number of nearest neighbor: 12
+    \item \textbf{Close-packed directions} As cube diagonals.
+    \item \textbf{Theortical Density} The theoretical density of HCP is derived from Equation (\ref{eq:theoretical-density}):
+    \begin{align}
+        \rho &= \frac{nM}{V_c N_A} \nonumber \\
+        \intertext{With $n = 6$ (12 corners $\times \frac{1}{6}$ atom each) and $V_c = a^3$:}
+        \rho_\text{HCP} &= \frac{6M}{a^3 N_A} 
+    \end{align}
+    \item \textbf{Atomic Packing Factor} The atomic packing factor of HCP is given by:
+    \begin{equation}
+        \text{APF}_\text{HCP} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{6 \times \frac{4}{3}\pi r^3}{a^3} = \frac{6 \times \frac{4}{3}\pi r^3}{a^3} = \textbf{0.74}
+    \end{equation}
+    \item \textbf{Stacking} The HCP structure can be thought of as stacking close-packed planes. The stacking sequence is ABAB...
 \end{itemize}
@@
 \end{example}
-\section{Geometric Descriptions of Crystals}
+\paragraph{Body-Centered Cubic (BCC)} The body-centered cubic structure has an atom at the centre of the cube. It has the following properties:
+\begin{enumerate}
+    \item \textbf{Coordination Number} Number of nearest neighbor: 8
+    \item \textbf{Atomic Packing Factor} The atomic packing factor of BCC is given by:
+    \begin{equation}
+        \text{APF}_\text{BCC} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{2 \times \frac{4}{3}\pi r^3}{a^3} = \frac{2 \times \frac{4}{3}\pi r^3}{(4r/\sqrt{3})^3} = \textbf{0.68}
+    \end{equation}
+    \item \textbf{Close Packed Directions} As cube diagonals.
+    \item \textbf{Theortical Density} The theoretical density of BCC is derived from Equation (\ref{eq:theoretical-density}):
+    \begin{align}
+        \rho &= \frac{nM}{V_c N_A} \nonumber \\
+        \intertext{With $n = 2$ (8 corners $\times \frac{1}{8}$ atom each + 1 center atom) and $V_c = a^3$:}
+        \rho_\text{BCC} &= \frac{2M}{a^3 N_A} 
+    \end{align}
+    \item \textbf{Stacking} The BCC structure can be thought of as stacking close-packed planes. The stacking sequence is ABAB...
+\end{enumerate}
+\section{Geometric Properties of Crystals}
 \subsection{Crystallographic Directions and Planes}
@@
 , where $a$ is the lattice parameter.
+\subsection{X-Ray Diffraction} The diffraction pattern is used to determine the crystal structure.
+By measuring the n$^\text{th}$ order peak angle $\theta_c$ and the wavelength of the X-ray, we can determine the distance between planes:
+\begin{align}
+    2d\sin \theta &= n\lambda \nonumber \\
+    d &= \frac{n\lambda}{2\sin \theta_c}
+    \intertext{Thus, we could determine the lattice parameter $a$ by the following equation:}
+    a &= \frac{n\lambda}{2\sin \theta_c} \sqrt{h^2 + k^2 + l^2}
+\end{align}
+
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-02-06 ESC195 Lecture Notes/main.tex
feb 6

@@
 \end{definition}
-\paragraph{Convergent Sequences} A sequence $\{a_n\}$ is said to be convergent if there exists a number $L$ such that for all $\epsilon > 0$, there exists an $N$ such that $|a_n - L| < \epsilon$ for all $n \ge N$.
+\begin{definition}[Unbounded Sequences]
+    A sequence $\{a_n\}$ is said to be unbounded if it is not bounded above or below.
+We can prove it using a epsilon-delta proof or considering the derivatives of the function that defines the sequence.
+\end{definition}
+\begin{example}
+    Proof that the sequence $\{a_n\} = 2^n$ is unbounded. Let $M$ be a number, we can find an $n$ such that $2^n > M$. We can find this $n$ by taking the logarithm of $M$ to the base $2$.
+\end{example}
+\begin{example}
+    Proof that the sequence $\{a_n\} = \frac{x^2}{e^x}$ is bounded above. We show that the derivative is negative after a certain point ($x = 2$), thus the sequence is bounded above.
+\end{example}
+\subsection{The Limit of a Sequence}
+\begin{definition}[Limit of a Sequence]
+    The limit of a sequence $\{a_n\}$ is given by:
+    \begin{equation*}
+        \lim_{n \to \infty} a_n = L
+    \end{equation*}
+    if for every $\epsilon > 0$, there exists an $N$ such that $|a_n - L| < \epsilon$ for all $n > N$. We denote this as $a_n \to L$ as $n \to \infty$.
+\end{definition}
+\paragraph{Convergence} A sequence $\{a_n\}$ is said to converge to $L$ if the limit of the sequence is $L$. And the following are true:
+\begin{enumerate}
+    \item If a sequence converges, then it is bounded.
+    \item If a sequence is unbounded, then it does not converge, i.e. it diverges.
+    \item However, if a sequence is bounded, it may or may not converge.
+\end{enumerate}
+\begin{example}
+    Prove that the sequence $\{a_n\} = \frac{n}{n+1}$ converges to $1$.
+We can show that $|a_n - 1| < \epsilon$ for all $n > N$ by taking $N = \frac{1}{\epsilon}$ by some algebraic manipulation.
+We have:\\
+    Given $\epsilon > 0$, let $N = \frac{1}{\epsilon}$, then for all $n > N$:
+    \begin{align*}
+        |a_n - 1| &= |\frac{n}{n+1} - 1| \\
+        &= |\frac{n - n - 1}{n+1}| \\
+        &= \frac{1}{n+1} < \frac{1}{n} < \epsilon
+    \end{align*}
+\end{example}
+\begin{theorem}[Uniqueness of Limits]
+    The limit of a sequence is unique. That is, of $\lim_{n \to \infty} a_n = L$ and $\lim_{n \to \infty} a_n = M$, then necessarily $L = M$.
+\end{theorem}
+\begin{theorem}[Monotone Convergence Theorem]
+    A monotonic sequence is convergent if and only if it is bounded. And necessarily the following are true if the sequence converges:
+    \begin{enumerate}
+        \item If the sequence is increasing or non decreasing, then the limit is the supremum of the sequence.
+        \item If the sequence is decreasing or non increasing, then the limit is the infimum of the sequence.
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Limit Operations]
+    Given two sequences $\{a_n\}$ and $\{b_n\}$, and $c \in \mathbb{R}$, we have:
+    \begin{enumerate}
+        \item \textbf{Sum of Limits}: $\lim_{n \to \infty} (a_n + b_n) = \lim_{n \to \infty} a_n + \lim_{n \to \infty} b_n$
+        \item \textbf{Product of Limits}: $\lim_{n \to \infty} (a_n \cdot b_n) = \lim_{n \to \infty} a_n \cdot \lim_{n \to \infty} b_n$
+        \item \textbf{Scalar Multiple of Limits}: $\lim_{n \to \infty} c \cdot a_n = c \cdot \lim_{n \to \infty} a_n$
+        \item \textbf{reciprocal of Limits}: $\lim_{n \to \infty} \frac{1}{a_n} = \frac{1}{\lim_{n \to \infty} a_n}$
+        \item \textbf{Quotient of Limits}: $\lim_{n \to \infty} \frac{a_n}{b_n} = \frac{\lim_{n \to \infty} a_n}{\lim_{n \to \infty} b_n}$, provided $\lim_{n \to \infty} b_n \neq 0$
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Squeeze Theorem\footnote{Also known as the Sandwich Theorem or the Pinching Theorem}]
+    Given three sequences $\{a_n\}$, $\{b_n\}$, and $\{c_n\}$, if $a_n \le b_n \le c_n$ for all $n$, and $\lim_{n \to \infty} a_n = \lim_{n \to \infty} c_n = L$, then necessarily $\lim_{n \to \infty} b_n = L$.
+This is useful for proving limits of sequences.
+\end{theorem}
+\begin{theorem}[Composite Function]
+    Given sequence $\{a_n\}$ covnerges to $L$ and continuous function $f$ at $L$, we have:
+    \begin{equation}
+        \lim_{n \to \infty} f(a_n) = f(\lim_{n \to \infty} a_n)
+    \end{equation}
+\end{theorem}
+\paragraph{Some Common Limits} Some common limits of sequences are:
+\begin{enumerate}
+    \item for $a > 0$, $\lim_{n \to \infty} a^{1/n} = 1$
+    \item for $|a| < 1$, $\lim_{n \to \infty} a^n = 0$
+    \item for $a > 0$, $\lim_{n \to \infty} \frac{1}{n^a} = 0$
+    \item for $a > 0$, $\lim_{n \to \infty} \frac{a^n}{n!} = 0$
+\end{enumerate}
+\subsection{Series}
+\begin{definition}[Infinite Series]
+    An infinite series is the sum of the terms of an infinite sequence. We denote the sum of the series as:
+    \begin{equation}
+        S = \sum_{n=1}^{\infty} a_n = a_1 + a_2 + a_3 + \ldots
+    \end{equation}
+    We also denote it as just $\sum a_n$.
+\end{definition}
+\begin{definition}[Partial Sum]
+    The partial sum of an infinite series is the sum of the first $n$ terms of the series. We denote the partial sum as:
+    \begin{equation}
+        S_n = \sum_{i=1}^{n} a_i = a_1 + a_2 + \ldots + a_n
+    \end{equation}
+    If the sequance $\{S_n\}$ converges to $S$, then the series is said to converge to $S$; otherwise, it is said to diverge.
+\end{definition}
+\paragraph{Geometric Series} A geometric series is a series of the form:
+\begin{equation}
+    \sum_{n=0}^{\infty} ar^n = a + ar + ar^2 + \ldots
+\end{equation}
+\begin{theorem}[Geometric Series]
+    A geometric series converges if $|r| < 1$, and the sum of the series is given by:
+    \begin{equation}
+        S = \frac{a}{1 - r}
+    \end{equation}
+    If $|r| \ge 1$, then the series diverges.
+\end{theorem}
+\begin{theorem}[Addition of Series]
+    Given two series $\sum a_n$ and $\sum b_n$, we have:
+    \begin{equation}
+        \sum a_n + \sum b_n = \sum (a_n + b_n)
+    \end{equation}
+\end{theorem}
+\begin{theorem}[Convergence]
+    $\sum a_n$ converges if and only if $\sum_{k=j}^{\infty} a_k$ converges for all $j$.
+\end{theorem}
+\begin{example}
+    Given $\sum \frac{3^{k-1}}{4^{3k+1}}$ converges. Then we have:
+    \begin{align*}
+        \sum \frac{3^{k-1}}{4^{3k+1}} = \frac{1}{12} + \frac{1}{64} + \frac{1}{4^7} + \ldots_n + \sum^{\infty}_{k=n} \frac{3^{k-1}}{4^{3k+1}}
+    \end{align*}
+    converges for all $n$.
+\end{example}
+\begin{theorem}[Zero Test, Divergence Test]
+    If $\lim_{n \to \infty} a_n \ne 0$, then $\sum a_n$ diverges. Note that the contrapositive is true, i.e. if $\sum a_n$ converges, then $\lim_{n \to \infty} a_n = 0$.
+\end{theorem}
+\subsection{The Integral Test and Estimates of Sums}
+\begin{theorem}[Integral Test]
+    Let $f$ be a continuous, positive, and decreasing function on $[1, \infty)$, and let $a_n = f(n)$.
+Then the series $\sum a_n$ converges if and only if the improper integral $\int_1^{\infty} f(x) \, dx$ converges.
+\end{theorem}
+\begin{example}[$p$=2-series]
+    Given $\sum \frac{1}{n^2}$, we can use the integral test to show that the series converges. We have:
+    \begin{align*}
+        \int_1^{\infty} \frac{1}{x^2} \, dx &= -\frac{1}{x} \Big|_1^{\infty} \\
+        &= 1
+    \end{align*}
+    Thus, the series converges.
+\end{example}
+Note that the ``iff'' statement allow us to determine the convergence of an integral.
+\begin{example}[Harmonic Series]
+    Given $\sum \frac{1}{n}$, we can use the integral test to show that the series diverges. We have:
+    \begin{align*}
+        \int_1^{\infty} \frac{1}{x} \, dx &= \ln x \Big|_1^{\infty} \\
+        &= \infty
+    \end{align*}
+    Thus, the series diverges. The harmonic series is a classic example of a divergent series.
+\end{example}
+\begin{definition}[Remainder]
+    The remainder of a series $\sum a_n$ is given by:
+    \begin{equation}
+        R_n = S - S_n = \sum_{k=n+1}^{\infty} a_k
+    \end{equation}
+    We can use the remainder to estimate the sum of a series.
+\end{definition}
+\paragraph{Estimating Sums} We can estimate the sum of a series by using the remainder of the series. We can use the remainder to estimate the sum of a series. Note that we always have:
+\begin{equation}
+    R_n \le \int_n^{\infty} f(x) \, dx /quad \text{and} \quad R_n \ge \int_{n+1}^{\infty} f(x) \, dx
+\end{equation}
+We can use it to estimate the sum of a series from a partial sum.
+\begin{equation}
+    S_n + \int_{n+1}^{\infty} f(x) \, dx \le S \le S_n + \int_n^{\infty} f(x) \, dx
+\end{equation}
+\begin{example}
+    Given $\sum \frac{1}{n^2}$, we can use the remainder to estimate the sum of the series. We have:
+    \begin{align*}
+        S_n + \int_{n+1}^{\infty} \frac{1}{x^2} \, dx &\le S \le S_n + \int_n^{\infty} \frac{1}{x^2} \, dx \\
+        \frac{1}{n} + \int_{n+1}^{\infty} \frac{1}{x^2} \, dx &\le S \le \frac{1}{n} + \int_n^{\infty} \frac{1}{x^2} \, dx
+    \end{align*}
+    We can then evaluate the integrals to obtain the value bounds.
+\end{example}
+\subsection{Comparison Tests}
+\begin{theorem}[Basic Comparison Test]
+    Let $\{a_n\}$ and $\{b_n\}$ be two sequences such that $0 \le a_n \le b_n$ for all $n$. Then:
+    \begin{enumerate}
+        \item If $\sum b_n$ converges, then $\sum a_n$ converges.
+        \item If $\sum a_n$ diverges, then $\sum b_n$ diverges.
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    Assume, WLOG, that $a_k \le b_k$ for all $k$. Define $S_n = \sum_{k=1}^{n} a_k$ and $T_n = \sum_{k=1}^{n} b_k$. Also, $T = \sum b_k$ exists. We have:
+    \begin{align*}
+        S_{n+1} &= S_n + a_{n+1} > S_n
+        \intertext{Thus, $\{S_n\}$ is increasing. Also, we have:}
+        S_n &\le T_n < T
+        \intertext{Thus, $\{S_n\}$ is bounded above. Thus, $\{S_n\}$ converges by the Monotone Convergence Theorem.}
+    \end{align*}
+\end{proof}
+\begin{example}
+    Let $S = \sum \frac{7}{17n^2 + 3n^{1/3} + 5}$ We know that $17n^2 > 17n^{1/3} + 5$ for all $n$. Thus, we have:
+    \begin{align*}
+        \frac{7}{17n^2} &> \frac{7}{17n^2 + 3n^{1/3} + 5} \\
+        \intertext{We can then use the comparison test to show that the series converges.}
+    \end{align*}
+\end{example}
 \appendix

### 2025-02-06 MAT185 Lecture Notes/main.tex
feb 6

@@
 \subsection{Foundamental Properties}
-\begin{definition}[Vector Space in $\mathbb{R}$]
+\begin{definition}[Vector Space in $\mathbb{R}$, Field]
     This course concerns with real vector spaces. A vectors space, ${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
@@
 
-    \paragraph{Note} It could be shown that the axiom imply the commutativity of in addition, namely $\forall \textbf{x} \implies \textbf{y} \in \mathbb{R}$, $\textbf{x} + \textbf{y} = \textbf{y} + \textbf{x}$
+    \paragraph{Note} It could be shown that the axiom imply the commutativity of in addition, namely $\forall \textbf{x}, \textbf{y} \in \mathbb{R}$, $\textbf{x} + \textbf{y} = \textbf{y} + \textbf{x}$
 \end{definition}
@@
 \end{theorem}
+\begin{theorem}[Span and Linear Independence]
+    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$.
+For every $\textbf{v}_k$ with $k = 1, 2, \ldots, n$, $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_{k-1}, \textbf{v}_{k+1}, \ldots, \textbf{v}_n) \subset \text{Span}(\textbf{v}_1, \ldots, \textbf{v}_n)$ if and only if $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent.
+\end{theorem}
 \begin{definition}[Basis]
@@
 \end{theorem}
+\begin{example}[Dimension of Even 4-degree polynomial]
+    Given $W \subseteq \mathbb{P}_4$ where $W = \left\{ p(x) \, | \, p(x) = p(-x) \right\}$, then $W$ is a subspace of $\mathbb{P}_4$.
+Since $x^3, x \in P_4$ but $x^3, x \notin W$,  Span$(\{W, x, x^3\}) = \mathrm{dim}(W) + 2$ So $\mathrm{dim}(W) \le 3$; We can find 3 linearly independent vectors in $W$ such as $1, x^2, x^4$.
+Thus, $\mathrm{dim}(W) \ge 3$.
+Thus, $\mathrm{dim}(W) = 3$.
+\end{example}
+
+\begin{definition}[Column Space]
+    Let $a_1, a_2, \ldots, {a_n} \in {^m\mathbb{R}}$, then:   
+    \begin{equation}
+        \text{Col}(A) = \text{Span}({a_1}, {a_2}, \ldots, {a_n})
+    \end{equation}
+    is the column space of $A$.
+\end{definition}
+Note that the span of set is a subspace of the vector space.
+Thus, the column space of a matrix is a subspace of $^m\mathbb{R}$.
+This is also call the set of images of the transformation.
+We can also note that matrix multiplication is is commutative with scalar multiplication.
+We can conduct proofs using such properties.
+\begin{definition}[Rank]
+    Let $A \in {^m\mathbb{R}^n}$, then:
+    \begin{equation}
+        \text{rank}(A) = \text{dim}(\text{Col}(A))
+    \end{equation}
+    is the rank of $A$.
+\end{definition}
 \end{document}

### 2025-02-06 MSE160 Lecture Notes/main.tex
feb 6

@@
 \end{align}
-
+\subsection{Geometrcally Ideal Crystal Structures}
+\paragraph{Coordination Number and Ratio} The coordination number increases with $\frac{r_{\text{cation}}}{r_{\text{anion}}}$ (The ionic radii ratio).  It relates to coordination number as follows:
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \textbf{Ionic Radii Ratio} & \textbf{Coord. Number} & \textbf{Name} \\
+        \hline
+        $<$ 0.155 & 2 & Linear \\
+        0.155 - 0.225 & 3 & Trigonal Planar \\
+        0.225 - 0.414 & 4 & Tetrahedral \\
+        0.414 - 0.732 & 6 & Octahedral \\
+        0.732 - 1.0 & 8 & Cubic \\
+        \hline
+    \end{tabular}
+    \caption{Coordination Number and Ionic Radii}
+    \label{tab:coordination-number}
+\end{table} 
+\begin{definition}[AX Type]
+    The AX type is a crystal structure where the cation and anion are of the same ratio. 
+\end{definition}
+\begin{example}[AX Type: $\ch{NaCl}$ rock salt]
+    The $\ch{NaCl}$ structure is an AX type structure.
+The cation and anion are of the same ratio.
+Soldium (Cation) has radius $r_{\text{Na}} = 0.102$ nm and Chlorine (Anion) has radius $r_{\text{Cl}} = 0.181$ nm.
+The ratio is:
+    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.102}{0.181} = 0.56 $$
+    Thus, the coordination number is 6.\\
+    Additionally, we could calculate its theoretical density. It has 4 cation and 4 anion in the unit cell. The volume of the unit cell is $a^3$, and it follows a FCC structure. Thus, the density is: 
+    \begin{align*}
+        \rho &= \frac{4M_{\text{Na}} + 4M_{\text{Cl}}}{a^3 N_A} = \frac{4(22.99) + 4(35.45)}{a^3 N_A} \\
+        &= \frac{4(22.99) + 4(35.45)}{(2\sqrt{2}r_{\text{Na}})^3 N_A} = \frac{4(22.99) + 4(35.45)}{(2\sqrt{2} \times 0.102)^3 N_A} \\
+        &= \frac{4(22.99) + 4(35.45)}{(2\sqrt{2} \times 0.102)^3 N_A} = 2.16 \text{ g/cm}^3
+    \end{align*}
+\end{example}
+\begin{example}[AX Type: $\ch{ZnS}$, Zinc Blende]
+    The $\ch{ZnS}$ structure is an AX type structure.
+It has a FCC structure with CN = 4.
+The radius of Zinc (Cation) is $r_{\text{Zn}} = 0.074$ nm and the radius of Sulfur (Anion) is $r_{\text{S}} = 0.184$ nm.
+The ratio is:
+    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.074}{0.184} = 0.40 $$
+\end{example}
+\begin{example}[AX$_2$ Type: $\ch{CaF2}$]
+    The $\ch{CaF2}$ structure is an AX$_2$ type structure.
+It has a FCC structure with CN = 8.
+The radius of Calcium (Cation) is $r_{\text{Ca}} = 0.099$ nm and the radius of Fluorine (Anion) is $r_{\text{F}} = 0.133$ nm.
+The ratio is:
+    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.099}{0.133} = 0.74 $$
+\end{example}
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-02-08 ESC195 Lecture Notes/main.tex
feb 8

@@
 \end{example}
-\subsection{Comparison Tests}
+\subsection{Other Comparison Tests}
 \begin{theorem}[Basic Comparison Test]
@@
 \end{example}
+\begin{theorem}[Limit Comparison Test]
+    Let $\{a_n\}$ and $\{b_n\}$ be two sequences such that $a_n, b_n > 0$ for all $n$. Then:
+    \begin{enumerate}
+        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = c > 0$, then either both series $\sum a_n$ and $\sum b_n$ converge or both diverge.
+        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = 0$ and $\sum b_n$ converges, then $\sum a_n$ converges.
+        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = \infty$ and $\sum b_n$ diverges, then $\sum a_n$ diverges.
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    We can prove the first part of the theorem by epsilon-delta proof. 
+    \\
+    Proof of part 1: Let $\lim_{n \to \infty} \frac{a_n}{b_n} = c > 0$. Then, for all $\epsilon > 0$, for all $n > N$, we have:
+    \begin{align*}
+        |\frac{a_n}{b_n} - c| &< \epsilon \\
+        \intertext{Choose $\epsilon = \frac{c}{2}$, then we have:}
+        |\frac{a_n}{b_n} - c| &< \frac{c}{2} \\
+        \frac{c}{2} < \frac{a_n}{b_n} &< \frac{3c}{2} \\
+    \end{align*}
+    We can then use the basic comparison test to show that the series converge or diverge.
+\end{proof}
+\begin{example}
+    Let $a_n = \frac{1}{n^3 - n}$ and choose $b_n = \frac{1}{n^3}$. We can use the limit comparison test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \frac{n^3}{n^3 - n} \\
+        &= 1
+    \end{align*}
+    Since $\sum b_n$ converges, then $\sum a_n$ converges.
+\end{example}
+\begin{example}
+    Let $a_n = \frac{\ln(n/1000)}{n}$ and chooise $b_n = \frac{1}{n}$. We can use the limit comparison test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \frac{\ln(n/1000)}{n} \\
+        &\to \infty
+    \end{align*}
+    Since $\sum b_n$ diverges, then $\sum a_n$ diverges.
+\end{example}
+\subsection{Alternating Series and Absolute Convergence}
+\begin{definition}[Alternating Series]
+    An alternating series is a series of the form:
+    \begin{equation}
+        \sum_{n=1}^{\infty} (-1)^{n-1} a_n = a_1 - a_2 + a_3 - a_4 + \ldots
+    \end{equation}
+    where $a_n > 0$ for all $n$.
+\end{definition}
+\begin{proof}
+    The proof for this is straight forward.
+We can show that the series converges by showing that the sequence $\{S_n\}$ is bounded above and below.
+Left as an exercise to the reader (Hint: Consider both the odd and even partial sums).
+\end{proof}
+\begin{example}
+    Let $a_n = \frac{1}{n}$, then the series $\sum (-1)^{n-1} \frac{1}{n}$ is an alternating series.
+\end{example}
+\paragraph{Note} $\sum \cos n$ is not an alternating series, though it alternates between positive and negative values, the negative and postive values are not necessarily next to each other.
+\begin{theorem}[Alternating Series Test]
+    Let $\{a_n\}$ be a sequence such that $a_n > 0$ for all $n$ and $\lim_{n \to \infty} a_n = 0$. Then the alternating series $\sum (-1)^{n-1} a_n$ converges if:
+    \begin{enumerate}
+        \item $a_{n+1} \le a_n$ for all $n$.
+        \item $\lim_{n \to \infty} a_n = 0$.
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Value of Alternating Series]
+    The value of an alternating series is given by:
+    \begin{equation}
+        S = a_1 - a_2 + a_3 - a_4 + \ldots = \lim_{n \to \infty} S_n
+    \end{equation}
+    where $S_n$ is the $n$th partial sum of the series. Additionally, the limit would always be between two consecutive partial sums.
+\end{theorem}
+\begin{example}
+    Let $S = e^{-1} = \sum (-1)^{n-1} \frac{1}{n!}$, we can use the above to estimate the value of $e^{-1}$. For $n=5$ we have:
+    \begin{align*}
+        e^{-1} &= 0.366 \pm 0.0015
+    \end{align*}
+\end{example}
+\begin{definition}[Absolute Convergence]
+    A series $\sum a_n$ is said to be absolutely convergent if the series $\sum |a_n|$ converges.
+Conversely, a series is said to be conditionally convergent if the series converges but the absolute value of the series diverges.
+\end{definition}
+\begin{example}
+    Let $a_n = \frac{(-1)^{n-1}}{n}$, then the series $\sum \frac{(-1)^{n-1}}{n}$ is conditionally convergent (Hint: 1/2 $L$).
+\end{example}
+\begin{theorem}[Absolute Convergence Test]
+    If the series $\sum |a_n|$ converges, then the series $\sum a_n$ converges.
+\end{theorem}
+\begin{proof}
+    We can prove this by using the comparison test.
+We can show that $-|a_n| \le a_n \le |a_n|$ and $0 \le a_n + |a_n| \le 2|a_n|$.
+Thus, we can use the comparison test to show that the series converges.
+Since $\sum a_n + |a_n|$ converges, then $\sum a_n$ converges.
+\end{proof}
+\subsection{Ratio and Root Tests}
+\begin{theorem}[Root Test]
+    Let $\{a_n\}$ be a sequence, and let $L = \lim_{n \to \infty} \sqrt[n]{|a_n|}$. Then:
+    \begin{enumerate}
+        \item If $L < 1$, then the series $\sum a_n$ converges.
+        \item If $L > 1$, then the series $\sum a_n$ diverges.
+        \item If $L = 1$, then the test is inconclusive.
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    Proof of part 1: Let $L < 1$.
+Then, choose $\mu$ such that $L < \mu < 1$.
+Then for some $N$, $\sqrt[n]{|a_n|} < \mu$ for all $n > N$, then $a_n < \mu^n$ for all $n > N$.
+We can then use the comparison test to show that the series converges since $\sum \mu^n$ converges for $|\mu| < 1$.
+\end{proof}
+\begin{example}
+    Let $a_n = \left ( \frac{n^2 + 1}{2n^2 + 1} \right )^n$. We can use the root test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \sqrt[n]{\left ( \frac{n^2 + 1}{2n^2 + 1} \right )^n} &= \lim_{n \to \infty} \frac{n^2 + 1}{2n^2 + 1} \\
+        &= \frac{1}{2}
+    \end{align*}
+    Thus, the series converges.
+\end{example}
+\begin{example}
+    Let $a_n = \frac{2^n}{n}$. So $\sqrt[n]{\frac{2^n}{n}} = \frac{2}{n^{1/n}}$. So the limit $2$ (denom. converges to 2), is greater than 1, thus the series diverges.
+\end{example}
+\begin{theorem}[Ratio Test]
+    Let $\{a_n\}$ be a sequence, and let $L = \lim_{n \to \infty} \left | \frac{a_{n+1}}{a_n} \right |$. Then:
+    \begin{enumerate}
+        \item If $L < 1$, then the series $\sum a_n$ converges.
+        \item If $L > 1$, then the series $\sum a_n$ diverges.
+        \item If $L = 1$, then the test is inconclusive.
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    Proof of part 1: Let $L < 1$.
+Then, choose $\mu$ such that $L < \mu < 1$.
+Then for some $N$, $ \frac{a_{n+1}}{a_n}  < \mu$ for all $n > N$, then $a_{n} < \mu a_{n+1} < \mu^2 a_{n+2} < \ldots < \mu^j a_{n+j}$ for all $n > N$.
+So we have:
+    \begin{align*}
+        \sum a_n &< \sum \frac{a_k}{\mu^k} \mu^n \\
+        &= \frac{a_k}{\mu_k} \sum \mu^n
+    \end{align*}
+    Since $\sum \mu^n$ converges, then $\sum a_n$ converges.
+\end{proof}
+\begin{example}
+    Let $a_n = \frac{k^2}{e^k}$ for some $k$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \left | \frac{a_{n+1}}{a_n} \right | &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{e^{k+1}} \cdot \frac{e^k}{k^2} \right | \\
+        &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{k^2} \cdot \frac{e^k}{e^{k+1}} \right | \\
+        &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{k^2} \cdot \frac{1}{e} \right | \\
+        &= \frac{1}{e} < 1
+    \end{align*}
+    Thus, the series converges.
+\end{example}
+\subsection{Power Series}
+\begin{definition}[Power Series]
+    A power series is a series of the form:
+    \begin{equation}
+        \sum_{n=0}^{\infty} c_n x^n = c_0 + c_1x + c_2x^2 + \ldots
+    \end{equation}
+    where $c_n$ are constants and $x$ is a variable. 
+\end{definition}
+\begin{example}[Geometric Series]
+    Let $c_n = 1$ for all $n$, then the series $\sum x^n$ is a geometric series:
+    \begin{equation*}
+        \sum x^n = 1 + x + x^2 + \ldots = \frac{1}{1-x}
+    \end{equation*}
+\end{example}
+\begin{definition}[Power Series at a Point]
+    A power series at a point $a$ is a series of the form:
+    \begin{equation}
+        \sum_{n=0}^{\infty} c_n (x-a)^n = c_0 + c_1(x-a) + c_2(x-a)^2 + \ldots
+    \end{equation}
+    where $c_n$ are constants and $x$ is a variable.
+\end{definition}
+\paragraph{} A special convergent limit for the power series when $x = a$ is called the radius of convergence. This always converges to $c_0$.
+\begin{example}
+    Let $a_n = \frac{x^n}{n^2}$ for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1} \cdot n^2}{x^n \cdot (n+1)^2} \right| \\
+        &= \lim_{n \to \infty} \left| x \cdot \frac{n^2}{(n+1)^2} \right| \\
+        &= \left| x \right| \lim_{n \to \infty} \left( \frac{n}{n+1} \right)^2 \\
+        &= \left| x \right|
+    \end{align*}
+According to the ratio test, the series $\sum a_n$ converges if $\left| x \right| < 1$.
+And if $x=1$ it is a $p=2$ series which converges.
+And for $x=-1$ we can use the alternating series test to show that the series converges since $|a_{n+1}| \le |a_n|$ and $\lim_{n \to \infty} a_n = 0$.
+\end{example}
+\begin{example}
+    Let $a_n = \frac{x^n(1+5^n)}{n!}$  for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1}(1+5^{n+1})}{(n+1)!} \cdot \frac{n!}{x^n(1+5^n)} \right| \\
+        &= \lim_{n \to \infty} \left| \frac{x}{1+n} \cdot \frac{1+5^{n+1}}{1+5^n} \right| \\
+        \intertext{The fist part goes to 5, and the second part goes to 0. Thus, the limit is 0.}
+        &= 0
+    \end{align*}
+According to the ratio test, the series $\sum a_n$ converges.
+\end{example}
+\begin{example}
+    Let $a_n = x^n n!$ for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges only when $x=0$. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1}(n+1)!}{x^n n!} \right| \\
+        &= \lim_{n \to \infty} \left| x(n+1) \right| 
+    \end{align*}
+    Thus, the series converges only when $x=0$.
+\end{example}
 \appendix

### 2025-02-08 MAT185 Lecture Notes/main.tex
feb 8

@@
 \end{definition}
+\begin{definition}[Row Space]
+    Let $a_1, a_2, \ldots, {a_n} \in {^m\mathbb{R}}$, then:   
+    \begin{equation}
+        \text{Row}(A) = \text{Span}({a_1}, {a_2}, \ldots, {a_n})
+    \end{equation}
+    is the row space of $A$.
+\end{definition}
 Note that the span of set is a subspace of the vector space.
@@
 \end{definition}
+\begin{theorem}[Rank of RREF]
+    Let $A \in {^m\mathbb{R}^n}$, then:
+    \begin{equation}
+        \text{rank}(A) = \text{rank}(\text{RREF}(A))
+    \end{equation}
+\end{theorem}
+\begin{theorem}[Dimension of Row and Column Space]
+    Let $A \in {^m\mathbb{R}^n}$, then:
+    \begin{align}
+        \text{dim}(\text{Row}(A)) &= \text{rank}(A) \\
+        \text{dim}(\text{Col}(A)) &= \text{rank}(A)
+    \end{align}
+\end{theorem}
+\begin{definition}[Matrix Multiplication]
+    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
+    \begin{equation}
+        (AB)_{ij} = \sum_{k=1}^{n} a_{ik}b_{kj}
+    \end{equation}
+\end{definition}
+\begin{definition}[Column View of Matrix Multiplication]
+    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
+    \begin{equation}
+        AB = \begin{bmatrix} A\textbf{b}_1 & A\textbf{b}_2 & \ldots & A\textbf{b}_p \end{bmatrix}
+    \end{equation}
+\end{definition}
+\begin{theorem}[Column Space of Matrix Multiplication]
+    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
+    \begin{equation}
+        \text{Col}(AB) \subseteq \text{Col}(A)
+    \end{equation}
+    If $B$ is a invertible matrix, then also $\text{Col}(A) \subseteq \text{Col}(AB)$
+\end{theorem}
+\begin{definition}[Invertible]
+    Let $A \in {^n\mathbb{R}^n}$, then $A$ is invertible if there exists a matrix $B \in {^n\mathbb{R}^n}$ such that:
+    \begin{align}
+        AB = BA = I
+    \end{align}
+    where $I$ is the identity matrix, and we denote $B = A^{-1}$, the inverse of $A$.
+\end{definition}
+\begin{proof}
+    Let $A$, $B$ be matrices, so $\text{col}(AB) = \text{Span}(\text{columns of }AB)$. In the column view, we have $AB = A(\text{columns of }B) = \begin{bmatrix}
+        A\textbf{b}_1 & A\textbf{b}_2 & \ldots & A\textbf{b}_p
+    \end{bmatrix}$
+    Each column of $AB$ is in $\text{Col}(A)$. So each column of $AB$ is a linear combination of the columns of $A$. Thus, $\text{Col}(AB) \subseteq \text{Col}(A)$.
+
+    Consider the case where $B$ is invertible.
+Let $C = AB$, Then $CB^{-1} = A$.
+Since it is proven that $\text{col}(CB^{-1}) \subseteq \text{col}(C)$, hence,  $\text{col}(A) \subseteq \text{col}(AB)$.
+\end{proof}
+\begin{definition}[Null Space of Matrix Multiplication]
+    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
+    \begin{equation}
+        \text{null}(AB) \supseteq \text{null}(B)
+    \end{equation}
+\end{definition}
 \end{document}

### 2025-02-08 MSE160 Lecture Notes/main.tex
feb 8

@@
 \end{center}
+% TODO: add a to r ratio table for different crystal structures/cubic structures
 \vspace{10pt}
@@
 \end{example}
+\section{Stress-Strain Behavior}
+\paragraph{Uniformly Distributed Elongation} The elongation of a material is uniformly distributed within the gauge length.
+The initial values, referred to as the ``elastic region," are characterized as uniformly elastic.
+\paragraph{Non-Linear Elastic} There is also a non-linear region where the material remains elastic but is not uniformly distributed.
+\paragraph{Localized Plastic Deformation} The material then enters the plastic region where the deformation is localized.
+The material is permanently deformed.
+This is characterized by the ultimate tensile strength (this is bad when it happens).
+\paragraph{Necking} The material then necks, where the cross-sectional area decreases. The material is then strained until it breaks.
+\paragraph{True Stress} The true stress is given by:
+\begin{equation}
+    \sigma = \frac{F}{A}
+\end{equation}
+, where $A$ is the instantaneous cross-sectional area.
+\paragraph{Ultimate Tensile Strength} The ultimate tensile strength is the maximum stress on the stress-strain curve.
+\paragraph{Strengthening} Plastic deformation is made more difficult by the following mechanisms:
+\begin{enumerate}
+    \item \textbf{Strain Hardening} Preloading beyond the elastic region - upon plastic deformation, the material will have a higher yield strength.
+    \item \textbf{For Metals and Alloys} The following mechanisms are used:
+\end{enumerate}
+
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}

### 2025-02-27 ECE159 Lecture Notes/main.tex
idk

--- 2025-01-16 ECE159 Lecture Notes/main.tex
@@
 \input{preamble}
+
 \begin{document}
@@
 Fundamentals of Electric Circuits, Winter 2024 \hfill ECE 159\\
+\\
+The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
 \vspace{10pt}
 \section{Basic Concepts}
+\paragraph{} One should have understood, after the course of high school physics, the basic concepts of electric circuits.
+This includes the concepts of voltage, current, resistance, and power.
+In this course, we will be building upon these concepts and applying them to more complex circuits.
 \subsection{Basic Defintions}
@@
 \end{equation}
+\paragraph{} The above concepts might be revisited in the future, but it is important to understand the basic concepts of voltage, current, resistance, power, and energy.
+\section{Basic DC Circuit Analysis with Resistors and Sources}
 \subsection{Circuit Elements}
@@
 \paragraph{Resistor} A resistor is a circuit element that restricts the flow of charge. The voltage across a resistor is proportional to the current through it.
-
+\paragraph{Passive Sign Convention} If a positive current flows out of the positive side of a voltage, that element is delivering power. Otherwise, it is absorbing power. All resistors absorb power.
 \subsection{Resistance}
@@
 \begin{equation}
-    I_i = I \frac{G_i}{G_{\text{eq}}}
+    I_i = I \frac{G_i}{G_{\text{eq}}} = I \frac{1}{R_i} \frac{1}{\sum \frac{1}{R_i}}
 \end{equation}
@@
 \begin{equation}
-    \sum I_{\text{in}} = \sum I_{\text{out}}
+    \sum I_{\text{in}} = \sum I_{\text{out}} \quad \text{or} \quad \sum_\text{$n$ in node} I_n = 0
 \end{equation}
@@
 \begin{equation}
-    \sum V_{\text{loop}} = 0
+    \sum_\text{$n$ in loop} V_{n} = 0
 \end{equation}
-
-\section{DC Circuits}
-\paragraph{}
-
+\subsection{Nodal Analysis}
+\paragraph{Nodal Analysis} Nodal analysis is a method used to analyze electric circuits. It involves writing Kirchhoff's Current Law for each node in the circuit.
+\paragraph{Steps for Nodal Analysis}
+\begin{enumerate}
+    \item Identify the nodes in the circuit.
+    \item Assign a voltage to each node.
+    \item Write Kirchhoff's Current Law for each node.
+    \item Solve the resulting system of equations.
+\end{enumerate}
+\paragraph{Supernode} A supernode is a combination of two nodes in a circuit.
+It is used to simplify the analysis of the circuit.
+A supernode is identified by a voltage source between two nodes or two nodes are shorted together.
+\begin{example}[Supernode]
+Consider the following circuit, where $V_s$ is a voltage source across two nodes.
+\end{example} 
+\begin{center}
+\begin{circuitikz}
+    \draw (0,0) to[V, v=$V_s$] (0,2) to[R, l=$R_1$] (2,2) to[R, l=$R_2$] (2,0) to[short] (0,0);
+\end{circuitikz}
+\end{center} 
+    \begin{align*}
+        V_1 - V_2 &= V_s  \\
+        \frac{V_1 - V_2}{R_1} + \frac{V_2}{R_2} &= 0
+    \end{align*}
+\paragraph{Strategies} When solving nodal analysis problems, it is important to choose a good ground node, most likely a question consist of two nodes equation with one or two addition simple equation.
+\subsection{Mesh Analysis}
+\paragraph{Mesh Analysis} Mesh analysis is a method used to analyze electric circuits. It involves writing Kirchhoff's Voltage Law for each loop in the circuit.
+\paragraph{Steps for Mesh Analysis}
+\begin{enumerate}
+    \item Identify the loops in the circuit.
+    \item Assign a current to each loop.
+    \item Write Kirchhoff's Voltage Law for each loop.
+    \item Solve the resulting system of equations.
+\end{enumerate}
+\paragraph{Supermesh} A supermesh is a combination of two loops in a circuit.
+It is used to simplify the analysis of the circuit.
+A supermesh is identified by a current source between two loops or two loops are shorted together.
+\begin{example}[Supermesh]
+Consider the following circuit, where $I_s$ is a current source between two loops.
+\end{example}
+\begin{center}
+    \begin{circuitikz}
+        \draw
+        (0,0) to[R, l=$R_1$] (0,3) -- (2,3) to[R, l=$R_2$] (2,0) -- (0,0)
+        (2,3) -- (4,3) to[R, l=$R_3$] (4,0) -- (2,0)
+        (0,3) to[I, l=$I_s$] (4,3);
+    \end{circuitikz}
+\end{center}
+\begin{align*}
+    \text{Mesh 1:} & \quad -V_1 + I_1 R_1 + I_1 R_2 - I_s R_2 = 0 \\
+    \text{Mesh 2:} & \quad -V_2 + I_2 R_3 + I_2 R_2 + I_s R_2 = 0 \\
+    \text{Supermesh:} & \quad I_1 - I_2 = I_s
+    \intertext{Addtionally, we can consider the entire circuit (or any loop that includes the current source) as a supermesh.}
+\end{align*}
+\paragraph{Strategies} When solving mesh analysis problems, it is important to choose a good loop, and to avoid current sources in the loop.
+\subsection{Superpostion, Tricks and Equivalent Circuits}
+\paragraph{Superposition} Superposition is a method used to analyze electric circuits. It involves analyzing the circuit with each source individually.
+\paragraph{Steps for Superposition}
+\begin{enumerate}
+    \item Turn off all sources except one. (voltage sources are shorted, current sources are open)
+    \item Analyze the circuit with the one source.
+    \item Repeat for each source.
+    \item Add the results together.
+\end{enumerate}
+\paragraph{Circuit of Dependent Sources} If a cicuit is composed of entirely dependent sources, the circuit has no current or voltage.
+However, we might be asked to find the equivalent resistance of the circuit, which can be done by applying a test voltage or current.
+\paragraph{Nodal Analysis by Inspection} We can create a matrix of conductances and voltages to solve for the nodal voltages.
+The diagonal elements are the sum of the conductances connected to the node, while the off-diagonal elements are the negative conductance between the nodes.
+\begin{theorem}[Thevenin's Theorem]
+    Thevenin's Theorem states that any linear circuit can be replaced by an equivalent circuit consisting of a voltage source and a resistor.
+    \begin{equation}
+        V_{\text{th}} = V_{\text{oc}} \quad R_{\text{th}} = R_{\text{eq}}
+    \end{equation}
+\end{theorem}
+\begin{theorem}[Norton's Theorem]
+    Norton's Theorem states that any linear circuit can be replaced by an equivalent circuit consisting of a current source and a resistor.
+    \begin{equation}
+        I_{\text{N}} = I_{\text{sc}} \quad R_{\text{N}} = R_{\text{eq}}
+    \end{equation}
+\end{theorem} 
+\paragraph{Thevenin Resistance} The Thevenin resistance is the resistance seen by the load resistor when the load resistor is removed (V = 0 and I open). There are cases we could consider:
+\begin{enumerate}
+    \item If there is only independent sources, then we simply remove the sources and find the equivalent resistance.
+    \item If there is only dependent sources, then we could apply a test voltage or current to find the equivalent resistance.
+    \item If there is a combination of independent and dependent sources, we could apply a test voltage or current to find the equivalent resistance after removing the independent sources.
+\end{enumerate}
+The above theorems are direct results of each other.
+\begin{example}[Thevenin's Theorem]
+    Consider the following circuit, where we are asked to find the Thevenin equivalent circuit.
+    \end{example}
+    \begin{center}
+\begin{circuitikz}
+    \draw
+    (0,0) to[V, v=$V_s$] (0,3) -- (2,3) to[R, l=$R_1$] (2,0) -- (0,0)
+    (2,3) -- (4,3) to[R, l=$R_2$] (4,0) -- (2,0)
+    (4,3) -- (6,3) to[R, l=$R_L$] (6,0) -- (4,0)
+    (2,0) -- (2,-2) to[I, l=$I_s$] (4,-2) -- (4,0);
+\end{circuitikz}
+\end{center}
+We remove the voltage and current sources and find the equivalent resistance.
+\begin{center}
+\begin{circuitikz}
+    \draw
+    (0,0) -- (0, 3) -- (2,3) to[R, l=$R_1$] (2,0) -- (0,0)
+    (2,3) -- (4,3) to[R, l=$R_2$] (4,0) -- (2,0)
+    (4,3) -- (6,3) to[R, l=$R_L$] (6,0) -- (4,0)
+    (2,0) -- (2,-2) to[open, i^=$I_s$] (4,-2) -- (4,0);
+\end{circuitikz}
+\end{center}
+\begin{theorem}[Maximum Power Transfer Theorem]
+    The maximum power transfer theorem states that the maximum power is transferred to the load resistor when the load resistor is equal to the Thevenin resistance.
+    \begin{equation}
+        R_L = R_{\text{th}}
+    \end{equation}
+\end{theorem}
+\section{Operational Amplifiers}
+\begin{definition}[Operational Amplifier (Op-Amp)]
+    An operational amplifier is a high-gain electronic voltage amplifier with a differential input and, usually, a single-ended output.
+\end{definition}
+\subsection{Ideal Op-Amp}
+\paragraph{Ideal Op-Amp} An ideal op-amp has the following characteristics:
+\begin{enumerate}
+    \item $V_+ = V_-$
+    \item $I_+ = I_- = 0$
+\end{enumerate}
+\paragraph{Common Op-Amp Circuits} There are several common op-amp circuits that are used in electric circuits. They could be found in the appendix \ref{app:opampcircuits}.
+\section{First Order Circuits}
+\begin{definition}[First Order Circuits]
+    A first order circuit is a circuit that can be described by a first order differential equation.
+\end{definition}
+\appendix
+\input{appendix}
 \end{document}

### 2025-02-27 ESC195 Lecture Notes/main.tex
idk

@@
 \begin{example}
-    Let $a_n = \frac{\ln(n/1000)}{n}$ and chooise $b_n = \frac{1}{n}$. We can use the limit comparison test to show that the series $\sum a_n$ converges. We have:
-    \begin{align*}
-        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \frac{\ln(n/1000)}{n} \\
+    Let $a_n = \frac{\ln(n/1000)}{n}$ and chooise $b_n = \frac{1}{n}$. We can use the limit comparison test to show that the series $\sum a_n$ diverges. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \ln(n/1000) \\
         &\to \infty
@@
 \end{example}
+\begin{theorem}
+    For a power series $\sum c_n (x-a)^n$, there are three possibilities in respect to convergence:
+    \begin{enumerate}
+        \item The series converges only when $x=a$.
+        \item The series converges for all $x$.
+        \item There exists a radius of convergence $R$ such that the series converges for $|x-a| < R$ and diverges for $|x-a| > R$.
+    \end{enumerate}
+\end{theorem}
+\paragraph{Interval of Convergence} The interval of convergence could also have differnt types of convergence at the endpoints. It could be either open or closed on either side.
+\begin{example}
+    Let $a_n = \frac{(-2)^n (x-1)^n}{n+2}$, we can use the ratio test to show that the series $\sum a_n$ converges for $|x-1| < 2$. We have:
+    \begin{align*}
+        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{(-2)^{n+1} (x-1)^{n+1}}{n+3} \cdot \frac{n+2}{(-2)^n (x-1)^n} \right| \\
+        &= \lim_{n \to \infty} \left| \frac{2(x-1)}{n+3} \right| \\
+        \intertext{It would reqiroes $|x-1| < 2$ for the series to converge, so $R=1/2$ or $x \in (1/2, 3/2)$. However, we also need to check the endpoints. For $x=1/2$, we have:}
+        \sum \frac{(-2)^n (1/2-1)^n}{n+2} &= \sum \frac{(-2)^n (-1/2)^n}{n+2} \\
+        &= \sum \frac{1}{n+2}
+        \intertext{This is a $p=1$ series, thus it diverges. For $x=3/2$, we have:}
+        \sum \frac{(-2)^n (3/2-1)^n}{n+2} &= \sum \frac{(-2)^n (1/2)^n}{n+2} \\
+        &= \sum \frac{(-1)^n}{n+2}
+        \intertext{This is an alternating series, thus it converges. Thus, the interval of convergence is $(1/2, 3/2]$.}
+    \end{align*}
+\end{example}
+\subsection{Representing Functions as Power Series}
+\paragraph{Representing Functions} We can represent functions as power series. We have, for some analytic function $f(x)$:
+\begin{equation}
+    f(x) = \sum_{n=0}^{\infty} c_n x^n = c_0 + c_1x + c_2x^2 + \ldots 
+\end{equation}
+\begin{example}
+    Let $f(x) = \frac{1}{1-x}$, we can represent the function as a power series. We have:
+    \begin{align*}
+        \frac{1}{1-x} &= 1 + x + x^2 + x^3 + \ldots \\
+        &= \sum x^n
+    \end{align*}
+\end{example}
+\begin{example}
+    Let $f(x) = \frac{x}{x-3}$, we can represent the function as a power series. We have:
+    \begin{align*}
+        \frac{x}{x-3} &= x \cdot \frac{1}{x-3} = -x \cdot \frac{1}{3-x} \\
+        &= -x \cdot \frac{1}{-3(1-x/3)} \\
+        \intertext{Using the geometric series, we have:}
+        &= -x \cdot \sum \frac{1}{3} \left( \frac{x}{3} \right)^n \\
+        &= -\sum \frac{x^{n+1}}{3^{n+1}},\quad |x/3| < 1
+    \end{align*}
+\end{example}
+\begin{theorem}[Term-by-term Differentiation and Integration]
+    Consider the power series $\sum c_n x^n$ with radius of convergence $R$. Then:
+    \begin{enumerate}
+        \item The power series $\sum n c_n x^{n-1}$ has the same radius of convergence $R$.
+        \item The power series $\sum \frac{c_n}{n+1} x^{n+1}$ has the same radius of convergence $R$.
+    \end{enumerate}
+    However, the \textbf{interval endpoints may change}.
+\end{theorem}
+\paragraph{Order of Operation of Differentiation, Integration and Summation} The following are true:
+\begin{enumerate}
+    \item $\frac{d}{dx} \sum c_n x^n = \sum \frac{d}{dx} c_n x^n$
+    \item $\int \sum c_n x^n \, dx = \sum \int c_n x^n \, dx$
+\end{enumerate}
+However, note that the constnat of integration is important in the integration case.
+\begin{example}
+    Let $f(x) = \frac{1}{(1+x)^2}$, we can represent the function as a power series. Since:
+    \begin{align*}
+        \frac{d}{dx} \frac{-1}{1+x} &= \frac{1}{(1+x)^2} \\
+        \intertext{We can then use the power series of $\frac{-1}{1+x}$ to find the power series of $\frac{1}{(1+x)^2}$.}
+        \frac{d}{dx} \sum (-1)^n x^n &= \sum \frac{d}{dx} (-1)^n x^n \\
+        &= \sum (-1)^n n x^{n-1} \\
+    \end{align*}
+    Thus, the power series of $\frac{1}{(1+x)^2}$ is $\sum (-1)^n n x^{n-1}$. The radius of convergence is the same as the original power series (1).
+\end{example}
+\begin{example}
+    Let $f(x) = \ln(1-x)$, we can represent the function as a power series. Since:
+    \begin{align*}
+        \int \frac{1}{1-x} \, dx &= -\ln(1-x) \\
+        \ln(1-x) &= -\int \sum x^n \, dx \\
+        &= C - \sum \frac{x^{n+1}}{n+1} \\
+    \end{align*}
+    Let $x=0$, then $C = 0$.
+Thus, the power series of $\ln(1-x)$ is $\sum \frac{x^{n+1}}{n+1}$.
+The radius of convergence is the same as the original power series (1).
+Also, we can show that $x=1$ diverges and $x=-1$ conditionally converges.
+\end{example}
+\begin{example}[Estimation]
+    We would like to estimate the following definite integral:
+    $$
+    I = \int_0^0.1 \frac{dx}{1+x^4}
+    $$
+    We can use the power series of $\frac{1}{1-x}$ to estimate the integral. Take $-x= -(-x)^4$, then we have:
+    \begin{align*}
+        I = \int_0^0.1 \frac{dx}{1+x^4} &= \int_0^0.1 \sum (-x)^{4n} \, dx \\
+        &= \sum \int_0^0.1 (-x)^{4n} \, dx \\
+        &= \sum \frac{(-0.1)^{4n+1}}{4n+1}
+    \end{align*}
+    By alternate series estimation, we can show that that $I \approx 0.099998 \pm 1.1 \times 10^{-10}$.
+\end{example}
+\begin{example}\label{ex:arctan}
+    Let $f(x) = \arctan(x)$, we can represent the function as a power series. Since:
+    \begin{align*}
+        \frac{d}{dx} \arctan(x) &= \frac{1}{1+x^2} \\
+        \int \frac{1}{1+x^2} \, dx &= \arctan(x) \\
+        \arctan(x) &= \int \sum (-1)^n x^{2n} \, dx \\
+        &= C + \sum \frac{(-1)^n x^{2n+1}}{2n+1} \\
+    \end{align*}
+    Let $x=0$, then $C = 0$. Thus, the power series of $\arctan(x)$ is $\sum \frac{(-1)^n x^{2n+1}}{2n+1}$. The radius of convergence is the same as the original power series (1). 
+\end{example}
+\begin{example}[Leibniz Formula]
+    The result from Example \ref{ex:arctan} can be used estimate the value of $\pi$. Since $\arctan(1) = \frac{\pi}{4}$, we have:
+    \begin{align*}
+        \arctan(1) &= \sum \frac{(-1)^n}{2n+1} \\
+        &= 1 - \frac{1}{3} + \frac{1}{5} - \frac{1}{7} + \ldots \\
+        &= \frac{\pi}{4}
+    \end{align*}
+\end{example}
+\subsection{Taylor and Maclaurin Series}
+\begin{center}
+    ``\textit{The amazing thing about Taylor series is that they swiftly converges}" - Prof. J. Davis
+\end{center}
+\paragraph{} We begin by assuming all function discussed in this section has a power series representation (i.e. it is infinitely differentiable), and that there is a non zero radius of convergence.
+\begin{example}[n-th derivatives of power series]
+    Let $f(x) = \sum c_n x^n$, we can find the n-th derivative of the function. We have:
+    \begin{align*}
+        f(x) &= \sum c_n x^n \\
+        f'(x) &= \sum c_n n x^{n-1} \\
+        f''(x) &= \sum c_n n(n-1) x^{n-2} \\
+        &\vdots \\
+        f^{(n)}(x) &= \sum c_n n(n-1) \ldots (n-n+1) x^{n-n} \\
+        &= \sum c_n n!
+    \end{align*}
+    Thus, by taking $f^{(n)}(0)$, we can find the n-th coefficient of the power series, i.e. $c_n = \frac{f^{(n)}(0)}{n!}$.
+\end{example}
+\begin{theorem}
+    If $f$ has a power series representation about $a$, then the coefficients of the power series are given by:
+    \begin{equation}
+        c_n = \frac{f^{(n)}(a)}{n!}
+    \end{equation}
+\end{theorem}
+\begin{definition}[Taylor Series]
+    The Taylor series of a function $f$ about $a$ is given by:
+    \begin{equation}
+        f(x) = \sum \frac{f^{(n)}(a)}{n!} (x-a)^n 
+    \end{equation}
+\end{definition}
+\begin{definition}[Maclaurin Series]
+    The Maclaurin series of a function $f$ is the Taylor series of $f$ about $a=0$.
+    \begin{equation}
+        f(x) = \sum \frac{f^{(n)}(0)}{n!} x^n
+    \end{equation}
+\end{definition}
+\begin{definition}[Analytic Function]
+    A function $f$ is said to be analytic if can be represented as a power series about 0. In turn, it is infinitely differentiable and has a power series representation about any point.
+\end{definition}
+\begin{example}
+    Let $f(x) = e^x$, we can find the Maclaurin series of the function. We have:
+    \begin{align*}
+        f(x) &= e^x \\
+        f'(x) &= e^x \\
+        f''(x) &= e^x \\
+        &\vdots \\
+        f^{(n)}(x) &= e^x \\
+        f^{(n)}(0) &= 1
+    \end{align*}
+    Thus, the Maclaurin series of $e^x$ is $\sum \frac{x^n}{n!}$. Also, by the ratio test, we can show that the series converges for all $x$ (i.e. the radius of convergence is $\infty$).
+\end{example}
+\begin{center}
+    ``\textit{When is it that $f(x) = \sum f^{(n)}(a) (x-a)^n/n!$?}''
+\end{center}
+\begin{definition}[n-th Order Taylor Polynomial]
+    The n-th order Taylor polynomial of a function $f$ about $a$ is given by:
+    \begin{equation}
+        P_n(x) = \sum_{k=0}^{n} \frac{f^{(k)}(a)}{k!} (x-a)^k
+    \end{equation}
+\end{definition}
+\begin{definition}[Remainder]
+    The remainder of the n-th order Taylor polynomial is given by:
+    \begin{equation}
+        R_n(x) = f(x) - P_n(x) = \frac{f^{(n+1)}(c)}{(n+1)!} (x-a)^{n+1}
+    \end{equation}
+    for some $c$ between $a$ and $x$.
+\end{definition}
+\begin{theorem}[Remainder Convergence]
+    If $f(a) = P_n(x) + R_n(x)$ and $\lim_{n \to \infty} R_n(x) = 0$, then $f(x) = \sum \frac{f^{(n)}(a)}{n!} (x-a)^n$.
+\end{theorem}
+\begin{theorem}[Taylor Theorem]
+    Let $f$ be a function that is $n+1$ times differentiable on an interval $I$ containing $a$. If the $n$-th derivative of $f$ is continuous on $I$, then for all $x$ in $I$:
+    \begin{equation}
+        f(x) = f(a) + f'(a)(x-a) + \ldots + \frac{f^{(n)}(a)}{n!} (x-a)^n + R_n(x)
+    \end{equation}
+    where $R_n(x) = \int \frac{f^{(n+1)}(c)}{n!} (x-a)^{n+1} \, dx = \frac{f^{(n+1)}(c)}{(n+1)!} (x-a)^{n+1}$.
+\end{theorem}
+\paragraph{} And for bounded deriviites, i.e. $|f^{(n+1)}(x)| \le M$ for all $x$ in $I$, we have:
+\begin{equation}
+    |R_n(x)| \le \frac{M}{(n+1)!} |x-a|^{n+1}
+\end{equation}
+We can quickly obtain an error bound for the Taylor polynomial by using the above inequality.
+\paragraph{Taylor Series of Common Functions}
+\begin{align}
+    e^x &= \sum \frac{x^n}{n!} \\
+    \sin(x) &= \sum \frac{(-1)^n x^{2n+1}}{(2n+1)!} \\
+    \cos(x) &= \sum \frac{(-1)^n x^{2n}}{(2n)!} \\
+    \ln(1+x) &= \sum \frac{(-1)^{n+1} x^n}{n} \\
+    (1+x)^p &= \sum \binom{p}{n} x^n
+\end{align}
+\begin{example}
+    Let $f(x) = \cos(x)$, we want ro find the taylor series of the function about $a=17\frac{\pi}{4}$ The derivatives and their respective value at $a$ are:
+    \begin{align*}
+        f(x=a) &= \cos(a) = \frac{1}{\sqrt{2}} \\
+        f'(x=a) &= -\sin(a) = -\frac{1}{\sqrt{2}} \\    
+        f''(x=a) &= -\cos(a) = -\frac{1}{\sqrt{2}} \\
+        f'''(x=a) &= \sin(a) = \frac{1}{\sqrt{2}} \\
+    \end{align*}
+    We can see that the derivatives repeat every 4th derivative, and that we could seperate it into two alternating series, we have:
+    \begin{align*}
+        \cos(x) &= \sum \frac{1}{\sqrt{2}} (-1)^n \frac{(x-17\pi/4)^{2n}}{(2n)!} + \sum \frac{1}{\sqrt{2}} (-1)^n \frac{(x-17\pi/4)^{2n+1}}{(2n+1)!} \\
+    \end{align*}
+\end{example}
+\paragraph{Multiplication and Division of Power Series} We can multiply and divide power series as we would with polynomials.
+We can also represent functions as products of known power series.
+We have some examples below.
+\begin{example}
+    Let $f(x) = \frac{e^x}{1+x}$, we can represent the function as product of two known power series. We have:
+    \begin{align*}
+        f(x) &= e^x \cdot \frac{1}{1+x} \\
+        &= \sum \frac{x^n}{n!} \cdot \sum (-1)^n x^n \\
+        &= (1 + x + \frac{x^2}{2} + \ldots) \cdot (1 - x + x^2 - \ldots) \\
+        &= \sum \sum \frac{(-1)^n x^{2n}}{n!}
+    \end{align*}
+\end{example}
+\paragraph{} We shall just deduce it with powers of terms.
+\begin{example}
+    Let $f(x) = \tan(x) = \frac{\sin(x)}{\cos(x)}$, we can represent the function as division of two known power series. We have:
+    \begin{align*}
+        f(x) &= \frac{\sin(x)}{\cos(x)} \\
+        &= \frac{x - \frac{x^3}{3!} + \frac{x^5}{5!} - \ldots}{1 - \frac{x^2}{2!} + \frac{x^4}{4!} - \ldots} \\
+        &= x - \frac{x^3}{3} + \frac{2x^5}{15} - \ldots
+    \end{align*}
+\end{example}
+\paragraph{} We shall just deduce it with long division, note that the radius of covnergence is not the same as the original power series.
+\subsection{Applications of Taylor Series}
+\paragraph{Approximation} We can use the Taylor series to approximate functions.
+We can use the first few terms of the Taylor series to approximate the function.
+We can also use the remainder to find an error bound for the approximation.
+\paragraph{Error Estimation} We can summerize two ways:
+\begin{enumerate}
+    \item If we know that it is an alternating series, then $|R_n(x)| \le a_{n+1}$.
+    \item Otherwise, we can use the remainder formula to find an error bound. $|R_n(x)| \le \frac{M}{(n+1)!} |x-a|^{n+1}$.
+\end{enumerate}
+\begin{example}
+    Let $f(x) = \sqrt{x}$, we can use the Taylor series to approximate $\sqrt{1.25}$.
+    \begin{align*}
+        f(x) &= \sqrt{x} = 1\\
+        f'(x) &= \frac{1}{2\sqrt{x}} = \frac{1}{2} \\
+        f''(x) &= -\frac{1}{4x^{3/2}} = -\frac{1}{4} \\
+        f'''(x) &= \frac{3}{8x^{5/2}} = \frac{3}{8} \\
+        \intertext{We can then use the Taylor series to approximate $\sqrt{1.25}$.}
+        f(1.25) &\approx 1 + \frac{1}{2}(1.25-1) - \frac{1}{8}(1.25-1)^2 + \frac{1}{16}(1.25-1)^3 \\
+        &= 1 + \frac{1}{2}(0.25) - \frac{1}{8}(0.25)^2 + \frac{1}{16}(0.25)^3 \\
+        &= 1 + \frac{1}{8} - \frac{1}{32} + \frac{1}{128} \\
+        &= 1.1182
+    \end{align*}
+    Since it is a alternating series, we also have that the error is less than $\frac{3}{8} \cdot \frac{1}{16} (0.25)^4 = 0.00015$. Thus, the error is less than 0.00015.
+\end{example}
+\begin{example}
+    Let $f(x) = \ln(1-x)$, we would like to find a term such that the estimate of $\ln(1.4)$ is less than $10^{-3}$.
+Since the Maclaurin series of $\ln(1-x)$ is $\sum \frac{x^n}{n}$, we can use the remainder for alternating series to find the error bound.
+We have:
+    \begin{align*}
+        |R_1(1.4)| &\le \frac{1}{2} (0.4)^2 = 0.08 \\
+        \vdots \\
+        |R_4(1.4)| &\le \frac{1}{5} (0.4)^5 = 0.002\\
+        |R_5(1.4)| &\le \frac{1}{6} (0.4)^6 = 0.0007 \\
+    \end{align*}
+    Thus, we can use the first 5 terms of the Maclaurin series to approximate $\ln(1.4)$ to less than $10^{-3}$.
+\end{example}
+\subsection{Binomial Series}
+\begin{definition}[Binomial Series]
+    The binomial series is a power series of the form:
+    \begin{equation}
+        (1+x)^p = \sum \binom{p}{n} x^n
+    \end{equation}
+    , where $\binom{p}{n} = \frac{p(p-1)\ldots(p-n+1)}{n!}$ and $\binom{k}{0} = 1$.
+\end{definition}
+\begin{example}[$(1+x)^k$]
+    Let $f(x) = (1+x)^k$, we can find the binomial series of the function. We have:
+    $$
+    (1+x)^k = \sum \binom{k}{n} x^n
+    $$
+
+    This could be derived, also, via Maclaurin series, we have:
+    \begin{align*}
+        f(x) = (1+x)^k &\quad f(0) = 1 \\
+        f'(x) = k(1+x)^{k-1} &\quad f'(0) = k \\
+        f''(x) = k(k-1)(1+x)^{k-2} &\quad f''(0) = k(k-1) \\
+        \vdots \\
+        f^{(n)}(x) &= k(k-1)\ldots(k-n+1)(1+x)^{k-n} \\
+        f^{(n)}(0) &= k(k-1)\ldots(k-n+1) \\
+        \intertext{Thus, the binomial series of $(1+x)^k$ is $\sum \binom{k}{n} x^n$.}
+    \end{align*}
+
+    We can also show that the binomial series converges for $|x| < 1$. With the ratio test, we have:
+    \begin{align*}
+        \lim_{n \to \infty} \left| \frac{\binom{k}{n+1} x^{n+1}}{\binom{k}{n} x^n} \right| &= \lim_{n \to \infty} \left| \frac{k-n}{n+1} x \right| \\
+        &= |x|
+    \end{align*}
+    Thus, the series converges for $|x| < 1$. Also for the endpoints, Prof. J. Davis didn't show the process, but we have:
+    $$
+    I = \begin{cases}
+        (-1, 1) & k \le -1 \\
+        (-1, 1] & k \in (-1, 0) \\
+        [-1, 1] & k \ge 0
+    \end{cases}
+    $$
+\end{example}
+\subsection{Fourier Series}
+\begin{definition}[Periodic Function]
+    A function $f$ is said to be periodic if there exists a positive number $P$ such that $f(x+P) = f(x)$ for all $x$.
+\end{definition}
+\begin{example}
+    Let $f(x) = \cos(nt) + \frac{1}2{\sin(2\pi t)}$ for some $n$. We can show that the function is periodic. We have:
+    \begin{align*}
+        f(x+P) &= \cos(n(t+P)) + \frac{1}{2}\sin(2\pi(t+P)) \\
+        &= \cos(nt + nP) + \frac{1}{2}\sin(2\pi t + 2\pi P) \\
+        &= \cos(nt) + \frac{1}{2}\sin(2\pi t) \\
+        &= f(x)
+    \end{align*}
+    Thus, the function is periodic.
+\end{example}
+\begin{theorem}
+    For periodic function $f$, with period $P$, continuous and piecewise differentiable, then the Fourier series of the function converges to the function.
+\end{theorem}
+\begin{definition}[Fourier Series]
+    The Fourier series of a function $f$ is a series of the form:
+    \begin{equation}
+        f(x) = \frac{a_0}{2} + \sum_{n=1}^{\infty} a_n \cos(n\omega x) + b_n \sin(n \omega x)
+    \end{equation}
+    , where $a_0, a_n, b_n$ are the Fourier coefficients of the function, and $\omega = \frac{2\pi}{P}$.
+\end{definition}
+\paragraph{Integral Properties of Fourier Series} First, we have these integral properites of $\cos$ and $\sin$:
+\begin{align}
+    \int_{-P/2}^{P/2} \cos(n\omega x) \, dx &= \begin{cases}
+        0 & n \ne 0 \\
+        P & n = 0
+    \end{cases} \\
+    \int_{-P/2}^{P/2} \sin(n\omega x) \, dx &= 0
+\end{align}
+Also, we have the product of $\cos$ and $\sin$:
+\begin{align}
+    \int_{-P/2}^{P/2} \sin(m\omega x) \sin(n\omega x) \, dx &= \begin{cases}
+        0 & m \ne n \\
+        \frac{P}{2} & m = n
+    \end{cases} \\ 
+    \int_{-P/2}^{P/2} \cos(m\omega x) \cos(n\omega x) \, dx &= \begin{cases}
+        0 & m \ne n \\
+        \frac{P}{2} & m = n
+    \end{cases} 
+\end{align}
+\begin{theorem}[Fourier Coefficients]
+    The Fourier coefficients of a function $f$ are given by:
+    \begin{align}
+        a_0 &= \frac{1}{P} \int_{-P/2}^{P/2} f(x) \, dx \\
+        a_n &= \frac{2}{P} \int_{-P/2}^{P/2} f(x) \cos(n\omega x) \, dx \\
+        b_n &= \frac{2}{P} \int_{-P/2}^{P/2} f(x) \sin(n\omega x) \, dx
+    \end{align}
+\end{theorem}
 \appendix

### 2025-02-27 MAT185 Lecture Notes/main.tex
idk

@@
 \end{definition}
+\begin{definition}[Rank-Nullity Theorem]
+    Let $A \in \mathbb{R}^{m \times n}$. Then:
+    \begin{equation}
+        \dim(\text{null}(A)) = n - \text{rank}(A)
+    \end{equation}
+    where $n$ is the number of columns of $A$.
+\end{definition}
+
+\begin{proof}
+    Consider $A \in {^m\mathbb{R}^n}$, Let $S = \{x_1, x_2, \ldots, x_k\}$ be a basis of $\text{null}(A)$, then $\{x_1, x_2, \ldots, x_k\}$ are linearly independent and that $\text{dim}(\text{null}(A)) = k$.
+We can extend $S$ to a basis of $^n\mathbb{R}$ by adding $n-k$ vectors, $\{x_{k+1}, x_{k+2}, \ldots, x_n\}$, then $\{x_1, x_2, \ldots, x_n\}$ is a basis of $^n\mathbb{R}$.
+Since $\{x_1, x_2, \ldots, x_n\}$ is a basis of $^n\mathbb{R}$, then $\text{rank}(A) = n - k$.
+Assert hat $n-k \ge 0$.
+If we can show that $\{Ax_1, Ax_2, \ldots, Ax_n\}$ is a basis of $\text{Col}(A)$, then we can show that $\text{rank}(A) = n - k$.
+We can show that $\{Ax_1, Ax_2, \ldots, Ax_n\}$ is a basis of $\text{Col}(A)$ by showing that $\{Ax_1, Ax_2, \ldots, Ax_n\}$ is linearly independent and that $\text{Span}(\{Ax_1, Ax_2, \ldots, Ax_n\}) = \text{Col}(A)$.
+
+\end{proof}
+\begin{theorem}[Properties of Row and Column Space w.r.t RREF]
+    Let $A \in {^m\mathbb{R}^n}$ and $\tilde{A} = \text{RREF}(A)$, then:
+    \begin{align}
+        \text{dim}(\text{Row}(\tilde{A})) &= \text{dim}(\text{Row}(A)) \\
+        \text{dim}(\text{Col}(\tilde{A})) &= \text{dim}(\text{Col}(A)) \\
+        \text{Row}(\tilde{A}) &= \text{Row}(A) \\
+        \text{Col}(\tilde{A}) &\neq \text{Col}(A)
+    \end{align}
+\end{theorem}
+\section{Linear Transformations}
+\begin{definition}[Linear Transformation]
+    Let $V$ and $W$ be vector spaces, then $T: V \rightarrow W$ is a linear transformation if:
+    \begin{enumerate}
+        \item $T(\textbf{x} + \textbf{y}) = T(\textbf{x}) + T(\textbf{y})$
+        \item $T(\alpha \textbf{x}) = \alpha T(\textbf{x})$
+    \end{enumerate}
+\end{definition}
+\begin{example}[$Ax + b$ is not LT]
+    Let $A \in {^m\mathbb{R}^n}$ and $b \in \mathbb{R}^m$, then $T: \mathbb{R}^n \rightarrow \mathbb{R}^m$ defined by $T(\textbf{x}) = A\textbf{x} + b$ is not a linear transformation.
+It violates the first property of linear transformation:
+    \begin{align*}
+        T(\textbf{x} + \textbf{y}) &= A(\textbf{x} + \textbf{y}) + b \\
+        &= A\textbf{x} + A\textbf{y} + b \\
+        &= T(\textbf{x}) + T(\textbf{y}) + b
+    \end{align*}
+\end{example}
+\begin{example}[Rotate by $90^\circ$]
+    Let $T: \mathbb{R}^2 \rightarrow \mathbb{R}^2$ defined by $T(\textbf{x}) = \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}\textbf{x}$, then $T$ is a linear transformation.
+It satisfies the properties of linear transformation.
+    \begin{align*}
+        T(\textbf{x} + \textbf{y}) &= \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}(\textbf{x} + \textbf{y}) \\
+        &= \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}\textbf{x} + \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}\textbf{y} \\
+        &= T(\textbf{x}) + T(\textbf{y})
+    \end{align*}
+    \begin{align*}
+        T(\alpha \textbf{x}) &= \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}(\alpha \textbf{x}) \\
+        &= \alpha \begin{bmatrix} 0 & -1 \\ 1 & 0 \end{bmatrix}\textbf{x} \\
+        &= \alpha T(\textbf{x})
+    \end{align*}
+
+\end{example}
+\begin{theorem}[Properties of Linear Transformation]
+    Let $T: V \rightarrow W$ be a linear transformation, then:
+    \begin{enumerate}
+        \item $T(\textbf{0}) = \textbf{0}$
+        \item $T(\textbf{x} - \textbf{y}) = T(\textbf{x}) - T(\textbf{y})$
+    \end{enumerate}
+\end{theorem}
+\begin{proof}
+    We have $T(\textbf{0} + X) = T(X) = T(\textbf{0}) + T(X)$.
+Thus, $T(\textbf{0}) = \textbf{0}$, by the Cancelation Theorem.
+We also have $T(\textbf{x} - \textbf{y}) = T(\textbf{x} + (-\textbf{y})) = T(\textbf{x}) + T(-\textbf{y}) = T(\textbf{x}) - T(\textbf{y})$.
+\end{proof}
 \end{document}

### 2025-02-27 MSE160 Lecture Notes/main.tex
idk

@@
 \begin{equation}
-    \sigma = \frac{F}{A}
+    \sigma_T = \frac{F}{A_i}
 \end{equation}
-, where $A$ is the instantaneous cross-sectional area.
+, where $A_i$ is the instantaneous cross-sectional area.
 \paragraph{Ultimate Tensile Strength} The ultimate tensile strength is the maximum stress on the stress-strain curve.
@@
 \end{enumerate}
-
+\paragraph{Strain Hardening Equation} The true stress-strain curve can be emprically described by the following equation:
+\begin{equation}
+    \sigma_T = K \epsilon^n_T
+\end{equation}
+, where $K$ is the Strain Hardening coefficient and $n$ is the strain hardening exponent. Both of these values are material properties.
+\paragraph{True Strain} Since $\epsilon_i = \frac{\Delta l_i}{l_0 + \Delta l_{i-1}}$ we can compute a Rieman sum to get the true strain:
+\begin{equation}
+    \epsilon_T = \int^{l_f}_{l_0} \frac{dl}{l} = \ln \left(\frac{l_f}{l_0}\right)
+\end{equation}
+\paragraph{Strain Rate} The strain rate is the rate at which the material is deformed. It is given by:
+\paragraph{} But, before necking, we can assume $V_0 = V_i$ and $A_i = \frac{A_0 l_0}{l_i}$. Thus, we have:
+\begin{align}
+    \sigma_T &= \frac{F}{A_i} = \frac{F}{\frac{A_0 l_0}{l_i}} = \frac{F l_i}{A_0 l_0} = \frac{F}{A_0} \cdot \frac{l_i}{l_0} = \sigma \cdot \frac{l_i}{l_0}  \nonumber \\
+    &= \sigma \cdot (1 + \epsilon)
+    \intertext{and,}
+    \epsilon_T &= \ln \left(1 + \epsilon\right)
+\end{align}
+\subsection{Imperfections}
+\paragraph{Dislocation} Recal that plastic deformation occurs by step-wise breaking of atomic bonds.
+This is done by the movement of dislocations.
+\textbf{Strengthening mechanisms in Metals} rely on inhibiting the movement of dislocations.
+\begin{definition}[Dislocation]
+    A dislocation is the breaking and reforming of bonds, one row at a time. It is a line defect in the crystal structure.
+\end{definition}
+\paragraph{Crystalline Imperfections} We use crystalline imperfections to inhibit the movement of dislocations. These imperfections include:
+\begin{itemize}
+    \item \textbf{Point Imperfections (Defects)} Vacancies, interstitials, and substitutional atoms.
+    (Example) Solid solution strengthening.
+    \item \textbf{Line Imperfections} Dislocations.
+    \item \textbf{Area Imperfections} Grain boundaries of the interface between two grains and free surfaces.
+    \item \textbf{Volume Imperfections} Second phase particles.
+\end{itemize}
+\paragraph{Solid Solution Strengthening} [Alloying] The addition of a solute to a solvent to inhibit the movement of dislocations. There is two types:
+\begin{enumerate}
+    \item \textbf{Interstitial Impurities} The solute atoms are smaller than the solvent atoms. This the impurites to be in the interstitial sites.
+    \item \textbf{Substitutional Impurities} The solute atoms are larger than or as large as the solvent atoms ($<10 \% \Delta \varnothing$, similar electornegativity.).
+This causes the impurities substitute the lattice sites.
+\end{enumerate}
+It slightly nudges nearby atoms, creating \textbf{lattice strain} that make it harder for dislocations to move.
+\paragraph{Machaism - Diffusion of Impurities} In positve edge dislcation, it has compession above disvlation line, and tension below.
+The impurities will `pin' the dislocation, making it harder to move.
+\paragraph{Plastic Deformation} We can also strengthen the material by increasing the number of dislocations. This is done by plastic deformation methods:
+\begin{enumerate}
+    \item Cold Work
+    \item Forging/Strain Hardening
+\end{enumerate}
+\paragraph{Mechanisms of Strengthening} Plastic deformation increases number of dislocations, which inhibits one another due to the strain fields.
+\begin{definition}[Grain Boundaries]
+    Grain boundaries are the interface between two grains. Grains are regions of the same crystal structure but with different orientations.
+\end{definition}
+\paragraph{Grain Size Reduction} Dislcations tend to move in the direction of the grain boundry (GB).
+Thus, reducing the grain size reduces the distance the dislocation can move.
+In addition, there are atomic strain due to the difference of bond lenght between grains, contributing to the strain field.
+It is given by:
+\begin{equation}
+    \sigma = \sigma_0 + \frac{K_y}{d}
+\end{equation}
+\paragraph{Second Phase Particles} Second phase particles are particles of a different phase in the material.
+They can be used to inhibit the movement of dislocations.
+They are often hard, brittke phases.
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}
+\paragraph{Dislcations (Slips)} Occurs on the plane with the highest planar density.
+The slip direction is the direction with the highest linear density.
+The slip system is the combination of the slip plane and slip direction.
+\paragraph{} FCC has 4 unique planes, each with 3 unique directions. Thus, there are 12 unique slip systems. Compare this to HCP, which has only base plane.
+\paragraph{Resolved Stear Stress} Consider single crystal with a certain plane.
+Let $\phi$ be the angle between the normal of the plane and the direction of the applied stress, and $\lambda$ be the angle between the plane and the orthonal of the applied stress.
+The resolved shear stress is given by:
+\begin{equation}
+    \tau = \sigma \cos \phi \cos \lambda
+\end{equation}
+\paragraph{Critcal Resolved Shear Stress} The critical resolved shear stress is the minimum shear stress required to move a dislocation.
+When the resolved shear stress equal the critical resolved shear stress, the dislocation will move, and:
+\begin{equation}
+    \sigma = \sigma_{y}
+\end{equation}
+So,
+\begin{equation}
+   \tau_{\text{CRSS}} = \sigma_{y} \cos \phi \cos \lambda
+\end{equation}
+\paragraph{Note} Most crystals are polycrystalline, so dislocation must be able to move to the neibouring crystal.
+\paragraph{Burgers Vector} The Burgers vector is the vector that describes the dislocation. It is the vector that closes the loop of the dislocation. It is given by:
+\begin{equation}
+    \vec{b} = \vec{b}_1 + \vec{b}_2
+\end{equation}
 \end{document}

### 2025-03-09 ESC195 Lecture Notes/main.tex
march 9

@@
     $$
-    I = \int_0^0.1 \frac{dx}{1+x^4}
+    I = \int_0^{0.1} \frac{dx}{1+x^4}
     $$
@@
     \begin{align*}
-        I = \int_0^0.1 \frac{dx}{1+x^4} &= \int_0^0.1 \sum (-x)^{4n} \, dx \\
-        &= \sum \int_0^0.1 (-x)^{4n} \, dx \\
+        I = \int_0^{0.1} \frac{dx}{1+x^4} &= \int_0^{0.1} \sum (-x)^{4n} \, dx \\
+        &= \sum \int_0^{0.1} (-x)^{4n} \, dx \\
         &= \sum \frac{(-0.1)^{4n+1}}{4n+1}
@@
 \end{theorem}
+\section{Vector and the Geometry of Space}
+\subsection{Equations of Lines and Planes}
+\begin{definition}[Vector in $\mathbb{R}^3$]
+    A vector in $\mathbb{R}^3$ is an ordered triple of real numbers $(a, b, c)$. We also write $\vec{v} = ai + bj + ck$.
+\end{definition}
+\paragraph{Radius Vector} The radius vector of a point $P(x, y, z)$ is the vector $\vec{r} = xi + yj + zk$, it starts at the origin and ends at the point $P$.
+\begin{definition}[Planes]
+    A plane in $\mathbb{R}^3$ is a flat surface that extends infinitely in all directions.
+A plane is determined by a point $P(x_0, y_0, z_0)$ and a normal vector $\vec{n} = ai + bj + ck$.
+The equation of the plane is given by:
+    \begin{equation}
+        a(x-x_0) + b(y-y_0) + c(z-z_0) = 0
+    \end{equation}
+\end{definition}
+\begin{example}[Vector Form of a Plane]
+    Given the normal vector and two points on the plane, we can find the vector form of the plane.
+Let $P_1(x_1, y_1, z_1)$ and $P_2(x_2, y_2, z_2)$ be two points on the plane, and $\vec{n} = ai + bj + ck$ be the normal vector.
+We have:
+    \begin{align*}
+        \vec{r} \cdot \vec{n} &= \vec{r_1} \cdot \vec{n} \\
+        (x-x_1)a + (y-y_1)b + (z-z_1)c &= 0 \\
+        \vec{r} \cdot \vec{n} &= \vec{r_2} \cdot \vec{n} \\
+        (x-x_2)a + (y-y_2)b + (z-z_2)c &= 0
+    \end{align*}
+\end{example}
+Thus, the vector form of the plane is given by:
+\begin{equation}
+    \vec{r} \cdot \vec{n} = \vec{r_1} \cdot \vec{n}
+\end{equation}
+\begin{definition}[Lines]
+    A line in $\mathbb{R}^3$ is a straight path that extends infinitely in two directions.
+A line is determined by a point $P(x_0, y_0, z_0)$ and a direction vector $\vec{d} = ai + bj + ck$.
+The equation of the line is given by:
+    \begin{equation}
+        \vec{r} = \vec{r_0} + t\vec{d}
+    \end{equation}
+\end{definition}
+\begin{example}[Vector Form of a Line]
+    Given two points on the line, we can find the vector form of the line. Let $P_1(x_1, y_1, z_1)$ and $P_2(x_2, y_2, z_2)$ be two points on the line. We have:
+    \begin{align*}
+        \vec{r} &= \vec{r_1} + t(\vec{r_2} - \vec{r_1}) \\
+        &= \vec{r_1} + t(\vec{d})
+    \end{align*}
+\end{example}
+    Thus, the vector form of the line is given by:
+    \begin{equation}
+        \vec{r} = \vec{r_0} + t\vec{d}
+    \end{equation}
+    \paragraph{Parametric Form of a Line} We can also write the line in parametric form. Let $P(x, y, z)$ be a point on the line, we have:
+    \begin{align}
+        x &= x_0 + at \\
+        y &= y_0 + bt \\
+        z &= z_0 + ct
+    \end{align}
+    We can get $t$ by solving the equations:
+    \begin{align}
+        \frac{x-x_0}{a} &= \frac{y-y_0}{b} = \frac{z-z_0}{c} = t
+    \end{align}
+\subsection{Cyliner and Quadric Surfaces}
+\begin{definition}[Cylinder]
+    A cylinder is a surface formed by the straight lines parallel to a given line and lying on a given surface. The equation of a cylinder is given by:
+    \begin{equation}
+        F(x, y) = 0
+    \end{equation}
+\end{definition}
+\begin{definition}[Quadric Surface]
+    A quadric surface is a surface that can be represented by a second degree equation in $x, y, z$. The general form of a quadric surface is given by:
+    \begin{equation}
+        Ax^2 + By^2 + Cz^2 + Dxy + Exz + Fyz + Gx + Hy + Iz + J = 0
+    \end{equation}
+\end{definition}
+\paragraph{Sketching Surfaces} We can sketch surfaces by:
+\begin{enumerate}
+    \item Find the domain and range of the surface.
+    \item Find the intercepts of the surface.
+    \item Find the traces with the coordinate planes (i.e. $x=0$, $y=0$, $z=0$).
+    \item Find sections of the surface.
+    \item Find the centre of the surface.
+    \item Find the symmetry of the surface.
+    \item Find the boundries and boundedness of the surface.
+\end{enumerate}
+\begin{example}[Hyperboloid]
+    We can sketch the hyperboloid $\frac{x^2}{a^2} + \frac{y^2}{b^2} - \frac{z^2}{c^2} = 1$. We could rearragne it for $z$:
+    \begin{align*}
+        \frac{z^2}{c^2} &= \frac{x^2}{a^2} + \frac{y^2}{b^2} - 1 \\
+        z &= \pm c \sqrt{\frac{x^2}{a^2} + \frac{y^2}{b^2} - 1}
+    \end{align*}
+    Clearly, we have intercepts at $z = \pm c$. It has no trace with the $xy$ plane. On other, it has traces of hyperbola. Taking the section for some $z_0 > c$, we have a ellipse.\\
+    For symmetry, it is symmetric about the $z$ axis. It is unbounded for all directions.
+\end{example}
+\paragraph{Projections - Curve of Intersection} We can find the curve of intersection of two surfaces by setting the two equations equal to each other.
+We can then project the curve of intersection onto the coordinate planes.
+\begin{enumerate}
+    \item We have $C: (x, y, z)$ such that $z = f(x, y)$ and $z = g(x, y)$.
+    \item We can obtain a vertical cylinder by $f(x, y) = g(x, y)$.
+    \item We can obtain a projection by setting $z = 0$.
+\end{enumerate}
+\section{Vector Functions}
+\subsection{Vector Functions and Space Curves}
+\begin{definition}[Vector Function]
+    A vector function is a function that maps a real number to a vector. A vector function in $\mathbb{R}^3$ is given by:
+    \begin{equation}
+        \vec{r}(t) = f(t)i + g(t)j + h(t)k
+    \end{equation}
+\end{definition}
+\begin{definition}[Space Curve]
+    A space curve is a curve in $\mathbb{R}^3$ that is determined by a vector function. The space curve is given by:
+    \begin{equation}
+        \vec{r}(t) = f(t)i + g(t)j + h(t)k
+    \end{equation}
+\end{definition}
+\begin{definition}[Limit]
+    Let the vector function $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ be defined on an open interval $I$ containing $t = a$, and let $\vec{L}$ be a vector.
+We say that the limit of $\vec{r}(t)$ as $t$ approaches $a$ is $\vec{L}$ if $\lim_{t \to t_0} \text{norm} (\vec{r}(t) - \vec{L}) = 0$.
+We write:
+    \begin{equation}
+        \lim_{t \to a} \vec{r}(t) = \vec{L}
+    \end{equation}
+\end{definition}
+\begin{theorem}[Vector Limit Rules] The following limit rules apply to vector functions:
+    \begin{enumerate}
+        \item \textbf{Norm Limit Rule}: $\lim_{t \to a} \text{norm} (\vec{r}(t)) = \text{norm} (\lim_{t \to a} \vec{r}(t))$.
+        \item \textbf{Sum Limit Rule}: $\lim_{t \to a} (\vec{r}(t) + \vec{s}(t)) = \lim_{t \to a} \vec{r}(t) + \lim_{t \to a} \vec{s}(t)$.
+        \item \textbf{Scalar Multiple Limit Rule}: $\lim_{t \to a} c\vec{r}(t) = c\lim_{t \to a} \vec{r}(t)$.
+        \item \textbf{Product Limit Rule}: $\lim_{t \to a} (\vec{r}(t) \cdot \vec{s}(t)) = \lim_{t \to a} \vec{r}(t) \cdot \lim_{t \to a} \vec{s}(t)$.
+        \item \textbf{Cross Product Limit Rule}: $\lim_{t \to a} (\vec{r}(t) \times \vec{s}(t)) = \lim_{t \to a} \vec{r}(t) \times \lim_{t \to a} \vec{s}(t)$.
+    \end{enumerate} 
+\end{theorem}
+\subsection{Derivatives and Integrals of Vector Functions}
+\begin{definition}[Derivative]
+    Let the vector function $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ be defined on an open interval $I$ containing $t = a$. The derivative of $\vec{r}(t)$ is given by:
+    \begin{equation}
+        \vec{r'}(t) = \lim_{h \to 0} \frac{\vec{r}(t+h) - \vec{r}(t)}{h}
+    \end{equation}
+\end{definition}
+\begin{definition}[Integration]
+    Let the vector function $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ be defined on an open interval $I$ containing $t = a$. The integral of $\vec{r}(t)$ is given by:
+    \begin{equation}
+        \int \vec{r}(t) \, dt = \int f(t) \, dt i + \int g(t) \, dt j + \int h(t) \, dt k
+    \end{equation}
+\end{definition}
+\begin{definition}[Triangle Inequality for Integrals]
+    Let the vector function $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ be defined on an open interval $I$ containing $t = a$. The triangle inequality for integrals is given by:
+    \begin{equation}
+        \left| \int \vec{r}(t) \, dt \right| \le \int \left| \vec{r}(t) \right| \, dt
+    \end{equation}
+\end{definition}
+\begin{definition}[Composite Functions]
+    Let $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ be a vector function and $u: \mathbb{R} \to \mathbb{R}$ be a scalar function. The composite function of $\vec{r}(t)$ and $u$ is given by:
+    \begin{equation}
+        (\vec{r}\circ u)(t) \vec{r}(u(t)) = f(u(t))i + g(u(t))j + h(u(t))k
+    \end{equation}
+\end{definition}
+\begin{theorem}[Differentiation Rules]
+    Below are the differentiation rules for vector functions:
+    \begin{enumerate}
+        \item \textbf{Sum Rule}: $(\vec{r} + \vec{s})' = \vec{r'} + \vec{s}'$.
+        \item \textbf{Scalar Multiple Rule}: $(c\vec{r})' = c\vec{r'}$.
+        \item \textbf{Product Rule}: $(\vec{r} \cdot \vec{s})' = \vec{r'} \cdot \vec{s} + \vec{r} \cdot \vec{s}'$.
+        \item \textbf{Cross Product Rule}: $(\vec{r} \times \vec{s})' = \vec{r'} \times \vec{s} + \vec{r} \times \vec{s}'$. (order matters)
+        \item \textbf{Chain Rule}: $(\vec{r}(u(t)))' = \vec{r'}(u(t)) \cdot u'(t)$.
+        \item \textbf{Quotient Rule}: $\left( \frac{\vec{r}}{\vec{s}} \right)' = \frac{\vec{r'} \vec{s} - \vec{r} \vec{s}'}{\vec{s}^2}$.
+    \end{enumerate}
+\end{theorem}
+\begin{example}[Dot Product]
+    Let $\vec{r}(t) = f(t)i + g(t)j + h(t)k$. Since we have $\vec{r} \dot \vec{r} = r^2$, differencating both sides, we have:
+    \begin{align*}
+        \frac{d\vec{r}}{dt} \cdot \vec{r} + \vec{r} \cdot \frac{d\vec{r}}{dt} &= 2r \frac{dr}{dt} \\
+        2\vec{r} \cdot \frac{d\vec{r}}{dt} &= 2r \frac{dr}{dt} \\
+        \vec{r} \cdot \frac{d\vec{r}}{dt} &= r \frac{dr}{dt}
+    \end{align*}
+\end{example}
+\begin{example}
+    Let $\vec{r}(t) = \frac{\vec{r}}{r}$ this would be a unit vector, so its derivative would be perpendicular to the vector. We have:
+    \begin{align*}
+       \frac{d}{dt}\frac{\vec{r}}{r} &= \frac{1}{r^3}\left[ r^2\frac{d\vec{r}}{dt} - \vec{r}\frac{dr}{dt} \right] \\
+         &= \frac{1}{r^3}\left[ (\vec{r} \cdot \vec{r})\frac{d\vec{r}}{dt} - \vec{r}(\vec{r} \cdot \frac{d\vec{r}}{dt}) \right] \\
+         \intertext{Since $\vec{a} \times \vec{b} \times \vec{c} = \vec{a}(\vec{b} \cdot \vec{c}) - \vec{b}(\vec{a} \cdot \vec{c})$, we have:}
+         &= \frac{1}{r^3}((\vec{r} \times \frac{d\vec{r}}{dt}) \times \vec{r}) 
+    \end{align*}
+\end{example}
+\begin{definition}[Tangent Vector]
+    Let $C$ be parameterized by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ and a differentiable curve.
+Then the nonzero vector $\vec{r'}$ is tangent to the curve at the point $\vec{r}(t)$ at the point $(f(t), g(t), h(t))$.
+\end{definition}
+\begin{definition}[Unit Tangent Vector]
+    The unit tangent vector is given by:
+    \begin{equation}
+        \vec{T} = \frac{\vec{r'}}{\text{norm}({\vec{r'}})}
+    \end{equation}
+\end{definition}
+\begin{theorem}[Properties of the Unit Tangent Vector]
+    The unit tangent vector $\vec{T}$ has the following properties:
+    \begin{enumerate}
+        \item $\vec{T} \cdot \vec{T} = 1$.
+        \item $\vec{T} \cdot \vec{r'} = \text{norm}({\vec{r'}})$.
+        \item $\vec{T} \cdot \vec{T'} = 0 = \vec{T} \cdot \vec{r''}$. (Since we are only changing direction)
+    \end{enumerate}
+\end{theorem}
+\subsection{Arc Length and Curvature}
+\begin{definition}[Arc Length]
+    The arc length of a curve $C$ parameterized by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ is given by:
+    \begin{align}
+        s &= \int_a^b \text{norm}(\vec{r'}(t)) \, dt \\
+        &= \int_a^b \sqrt{f'(t)^2 + g'(t)^2 + h'(t)^2} \, dt
+    \end{align}
+\end{definition}
+\begin{example}[Helix]
+    Consider the helix $C$ perametized by $r(t) = 3\cos(t)i - 3\sin(t)j + 4t$. We can find the arc length of the helix. We have:
+    \begin{align*}
+        s &= \int_0^{2\pi} \sqrt{9\cos ^2 t + 9\sin ^2 t + 16} \, dt \\
+        &= \int_0^{2\pi} \sqrt{9 + 16} \, dt \\
+        &= \int_0^{2\pi} 5 \, dt \\
+        &= 10\pi
+    \end{align*}
+\end{example}
+\begin{definition}[Speed]
+    The speed of a curve $C$ parameterized by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ is given by:
+    \begin{equation}
+        v = \text{norm}(\vec{r'}(t)) = \sqrt{f'(t)^2 + g'(t)^2 + h'(t)^2}
+    \end{equation}
+\end{definition}
+\begin{definition}[Curvature]
+    The curvature of a curve $C$ parameterized by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$ is given by:
+    \begin{equation}
+        \kappa = \left| \frac{d\phi}{ds} \right|
+    \end{equation}
+    , where $\phi$ is the angle between the tangent vector and the $x$ axis.
+\end{definition}
+\begin{example}
+    Let $C$ be a curve defined by $y = y(x)$, we can find the curvature of the curve. We have:
+    \begin{align*}
+        \frac{dy}{dx} = y' &= \tan \phi \\
+        \phi &= \arctan(y') \\
+        \frac{d\phi}{dx} &= \frac{y''}{1 + y'^2} \\
+        \intertext{Also, for $ds$, we have:}
+        ds &= \sqrt{1 + y'^2} \, dx \\
+        \intertext{By chain rule, we have:}
+        \frac{d\phi}{dx} &= \frac{d\phi}{ds} \frac{ds}{dx} \\
+        \intertext{Solving for $\kappa$, we have:}
+        \kappa &= \frac{d\phi}{ds} = \frac{y''}{(1 + y'^2)^{3/2}}
+    \end{align*}
+\end{example}
+\begin{theorem}
+    For a curve $C$ parameterized by $y = y(x)$, the curvature of the curve is given by:
+    \begin{equation}
+        \kappa = \frac{y''}{(1 + y'^2)^{3/2}}
+    \end{equation}
+\end{theorem}
+\begin{example}[Straight Line]
+    Let $C$ be a straight line defined by $y = mx + b$, we can find the curvature of the curve. We have:
+    \begin{align*}
+        y' &= m \\
+        y'' &= 0 \\
+        \kappa &= \frac{0}{(1 + m^2)^{3/2}} = 0
+    \end{align*}
+\end{example}
+\paragraph{} For curves perametized by $\vec{r}(t) = x(t)i + y(t)j$, we have:
+\begin{equation}
+    \frac{dy}{dx} = \frac{dy/dt}{dx/dt} = \frac{y'(t)}{x'(t)}
+\end{equation}
+For $d^2y/dx^2$, we have, by quotient rule:
+\begin{align}
+    \frac{d^2y}{dx^2} &= \frac{d}{dx} \left( \frac{dy}{dx} \right) \nonumber  \\
+    &= \frac{dx'/dx \dot y' - dy'/dx \dot x'}{(dx/dt)^2} \nonumber  \\
+    \intertext{By chain rule, we have:}
+    \frac{dx'}{dx} &= \frac{dx'/dt}{dx/dt} = \frac{x''}{x'} \nonumber  \\
+    \frac{dy'}{dx} &= \frac{dy'/dt}{dx/dt} = \frac{y''}{x'} \nonumber 
+    \intertext{Substituting back, we have:} 
+    \frac{d^2y}{dx^2} &= \frac{x''y' - y''x'}{x'^3}
+\end{align}
+\begin{theorem}
+    For a curve $C$ parameterized by $\vec{r}(t) = x(t)i + y(t)j$, the curvature of the curve is given by:
+    \begin{equation}
+        \kappa = \frac{x''y' - y''x'}{(x'^2 + y'^2)^{3/2}}
+    \end{equation}
+\end{theorem}
+\begin{example}[Circle]
+    Let $C$ be a circle defined by $r\cos(t)i + r\sin(t)j$, we can find the curvature of the curve. We have:
+    \begin{align*}
+        x &= r\cos(t) \\
+        y &= r\sin(t) \\
+        x' &= -r\sin(t) \\
+        y' &= r\cos(t) \\
+        x'' &= -r\cos(t) \\
+        y'' &= -r\sin(t) \\
+        \kappa &= \frac{(-r\cos(t))(r\cos(t)) - (-r\sin(t))(r\sin(t))}{(r^2)^{3/2}} \\
+        &= \frac{r^2}{r^3} = \frac{1}{r}
+    \end{align*}
+\end{example}
+\begin{definition}[Radius of Curvature]
+    The radius of curvature of a curve $C$ is given by:
+    \begin{equation}
+        \rho = \frac{1}{\kappa}
+    \end{equation}
+\end{definition}
+\begin{example}[Ellipse]
+    Let $C$ be an ellipse defined by $a\cos(t)i + b\sin(t)j$, we can find the radius of curvature of the curve. We have:
+    \begin{align*}
+        x &= a\cos(t) \\
+        y &= b\sin(t) \\
+        x' &= -a\sin(t) \\
+        y' &= b\cos(t) \\
+        x'' &= -a\cos(t) \\
+        y'' &= -b\sin(t) \\
+        \kappa &= \frac{(-a\cos(t))(b\cos(t)) - (-b\sin(t))(-a\sin(t))}{(a^2\sin^2(t) + b^2\cos^2(t))^{3/2}} \\
+        &= \frac{ab}{(a^2\sin^2(t) + b^2\cos^2(t))^{3/2}}
+    \end{align*}
+\end{example}
+\paragraph{Relationship to $\vec{T}$} For an angle $\phi$ between the $x$ axis and the tangent vector $\vec{T}$, we have:
+\begin{align}
+    \vec{T} &= \cos(\phi)i + \sin(\phi)j \\
+    \intertext{Also, for $d\vec{T}/ds$, we have:}
+    \frac{d\vec{T}}{ds} &= -\sin(\phi)\frac{d\phi}{ds}i + \cos(\phi)\frac{d\phi}{ds}j \nonumber \\
+    \left|\frac{d T}{ds}\right| &= \frac{d\phi }{ds} \sqrt{\sin^2 (\phi) + \cos^2 (\phi)} = \kappa \\
+\end{align}
+\begin{definition}[Curvature for $n$-D curves]
+    For a curve $C$ parameterized by $\vec{r}(t) = f_1(t)i + f_2(t)j + \ldots + f_n(t)k$, the curvature of the curve is given by:
+    \begin{equation}
+        \kappa = \left| \frac{d\vec{T}}{ds} \frac{dt}{ds} \right| = \frac{|\vec{T'}|}{|\vec{r'}|} 
+    \end{equation}
+\end{definition}
+\begin{example}[Helix]
+    Let $C$ be a helix defined by $3\cos(t)i + 3\sin(t)j + 4tk$, we can find the curvature of the curve. We have:
+    \begin{align*}
+        \frac{d\vec{r}}{dt} &= -3\sin(t)i + 3\cos(t)j + 4k \\
+        |\frac{d\vec{r}}{dt}| &= \sqrt{9\sin^2(t) + 9\cos^2(t) + 16} = 5 \\
+        \vec{T} &= \frac{1}{5}(-3\sin(t)i + 3\cos(t)j + 4k) \\
+        \frac{d\vec{T}}{dt} &= \frac{1}{5}(-3\cos(t)i - 3\sin(t)j) \\
+        |\frac{d\vec{T}}{dt}| &= \frac{3}{5} \\
+        \kappa &= \frac{3}{5} \cdot \frac{1}{5} = \frac{3}{25} \\
+        \rho &= \frac{25}{3}
+    \end{align*}
+\end{example}
+\begin{definition}[Cross Product Defintion of Curvature]
+    For a curve $C$ parameterized by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$, the curvature of the curve is given by:
+    \begin{equation}
+        \kappa = \frac{|\vec{r'} \times \vec{r''}|}{|\vec{r'}|^3}
+    \end{equation}
+\end{definition}
+\paragraph{The Normal and Binormal Vectors} We can use the tangent vector to find the normal and binormal vectors. We can then constuct the osculating plane.
+\begin{definition}[Normal Vector]
+    The normal vector of a curve $C$ is a vector that is perpendicular to the tangent vector $\vec{T}$ of the curve. The normal vector is given by:
+    \begin{equation}
+        \vec{N} = \frac{d\vec{T}}{ds}
+    \end{equation}
+\end{definition}
+\begin{definition}[Binormal Vector]
+    The binormal vector of a curve $C$ is a vector that is perpendicular to both the tangent vector $\vec{T}$ and the normal vector $\vec{N}$ of the curve. The binormal vector is given by:
+    \begin{equation}
+        \vec{B} = \vec{T} \times \vec{N} = \frac{\vec{T'}}{|\vec{T'}|}
+    \end{equation}
+\end{definition}
+\begin{definition}[Osculating Plane]
+    The osculating plane of a curve $C$ is a plane that is tangent to the curve at a point and contains the tangent vector $\vec{T}$ and the normal vector $\vec{N}$ of the curve.
+\end{definition}
+\subsection{Motion in Space: Velcity and Acceleration}
+\begin{definition}[Location, Velocity and Acceleration]
+    Let the position of an object be given by $\vec{r}(t) = f(t)i + g(t)j + h(t)k$. The velocity of the object is given by:
+    \begin{equation}
+        \vec{v} = \vec{r'}(t) = f'(t)i + g'(t)j + h'(t)k
+    \end{equation}
+    The acceleration of the object is given by:
+    \begin{equation}
+        \vec{a} = \vec{v'}(t) = f''(t)i + g''(t)j + h''(t)k
+    \end{equation}
+\end{definition}
+\paragraph{Circular Motion} For circular motion, we have:
+\begin{align}
+    \vec{r}(t) &= r\cos(\omega t)i + r\sin(\omega t)j \\
+    \vec{v}(t) &= -r\omega\sin(\omega t)i + r\omega\cos(\omega t)j \\
+    \vec{a}(t) &= -r\omega^2\cos(\omega t)i - r\omega^2\sin(\omega t)j
+\end{align}
+\paragraph{Clockwise Convention} Note that for $\theta$ increasing, the vector rotates counter-clockwise. For $\theta$ decreasing, the vector rotates clockwise.
+\begin{definition}[Angluar Velocity and Speed]
+    The angular velocity of a curve $C$ is given by:
+    \begin{equation}
+        \omega = \frac{d\theta}{dt}
+    \end{equation}
+    For angular speed, we have:
+    \begin{equation}
+        \omega = \left|\frac{d\theta}{dt}\right|
+    \end{equation}
+\end{definition}
+\subsection{Introduction to Vector Mechanics}
+\begin{definition}[Newton's Second Law]
+    Newton's second law states that the acceleration of an object is directly proportional to the net force acting on the object and inversely proportional to the mass of the object.
+The law is given by:
+    \begin{equation}
+        \vec{F} = m\vec{a}
+    \end{equation}
+\end{definition}
+\begin{definition}[Momentum]
+    The momentum of an object is the product of the mass of the object and the velocity of the object. The momentum is given by:
+    \begin{equation}
+        \vec{p} = m\vec{v}
+    \end{equation}
+\end{definition}
+\begin{definition}[Conservation of Momentum]
+    The conservation of momentum states that the total momentum of a system of objects is conserved if the net external force acting on the system is zero. The law is given by:
+    \begin{equation}
+        \vec{p}_{\text{initial}} = \vec{p}_{\text{final}}, \quad \vec{p^\prime} = \vec{F} = 0
+    \end{equation}
+\end{definition}
+\begin{definition}[Angualr Momentum]
+    The angular momentum of an object is the product of the moment of inertia of the object and the angular velocity of the object. The angular momentum is given by:
+    \begin{equation}
+        \vec{L} = \vec{r} \times \vec{p}
+    \end{equation}
+\end{definition}
+\begin{example}[Angular Momentum]
+    Let $C$ be a curve defined by $r\cos(t)i + r\sin(t)j$, we can find the angular momentum of the curve. We have:
+    \begin{align*}
+        \vec{r} &= r\cos(t)i + r\sin(t)j \\
+        \vec{v} &= -r\sin(t)i + r\cos(t)j \\
+        \vec{p} &= m(-r\sin(t)i + r\cos(t)j) \\
+        \vec{L} &= \vec{r} \times \vec{p} = mr^2\omega k \\
+        |\vec{L}| &= mr^2\omega
+    \end{align*}
+\end{example}
+\begin{definition}[Torque]
+    The torque of an object is the product of the moment arm of the object and the force acting on the object. The torque is given by:
+    \begin{equation}
+        \vec{\tau} = \vec{r} \times \vec{F}
+    \end{equation}
+\end{definition}
+For $\vec{\tau} = 0$, $\vec{L}$ is conserved.
+\begin{definition}[Acceleration]
+    For velocity, we can use the tagental vector to define a velociy vector. For acceleration, differenciate it and get a ccaletion with respect to the normal vector:
+    \begin{align}
+        \vec{v} &= \frac{ds}{dt}\vec{T} \\
+        \vec{a} &= \frac{d\vec{v}}{dt} = \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\vec{T}}{dt} \nonumber \\
+        \intertext{Recall that $\vec{T} = \left | \frac{d\vec{T}}{dt} | \right | \vec{N}$, we have:}
+        \vec{a} &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\vec{T}}{dt} \nonumber \\
+        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\phi}{dt}\vec{N} \nonumber \\
+        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\kappa\vec{N}
+    \end{align}
+\end{definition}
+
 \appendix

### 2025-03-09 MAT185 Lecture Notes/main.tex
march 9

@@
 \end{proof}
+\begin{definition}[Kernel and Image]
+    Let $T: V \rightarrow W$ be a linear transformation, then: 
+    \begin{enumerate}
+        \item The \textbf{kernel} of $T$, denoted as $\ker(T)$, is the set of all vectors in $V$ that map to the zero vector in $W$. Formally,
+        \[
+        \ker(T) = \{ \textbf{v} \in V \mid T(\textbf{v}) = \textbf{0} \}
+        \]
+        \item The \textbf{image} of $T$, denoted as $\text{Im}(T)$, is the set of all vectors in $W$ that are images of vectors in $V$. Formally,
+        \[
+        \text{Im}(T) = \{ T(\textbf{v}) \mid \textbf{v} \in V \}
+        \]
+    \end{enumerate}
+\end{definition}
+\begin{theorem}[Rank-Nullity for LT]
+    Let $T: V \rightarrow W$ be a linear transformation, then:
+    \begin{equation}
+        \dim(\ker(T)) + \dim(\text{Im}(T)) = \dim(V)
+    \end{equation}
+\end{theorem}
+\begin{definition}[Injective LT]
+    Let $T: V \rightarrow W$ be a linear transformation, then $T$ is injective if $x_1 \neq x_2 \implies T(x_1) \implies T(x_2)$
+\end{definition}
+\begin{theorem}
+    If the nullity of $T$, $\text{dim ker} T$ is greater than zero, then $T$ is not injective.
+\end{theorem}
+\begin{definition}[Surjective]
+    Let $T: V \rightarrow W$ be a linear transformation, then $T$ is surjective if for every $y \in W$, there exist an $x \in V$ so that $T(x) = y$.
+\end{definition}
+\begin{definition}[Bijective]
+    Let $T: V \rightarrow W$ be a linear transformation, then $T$ is bijective if $T$ is both injective and surjective.
+\end{definition}
+\begin{theorem}[Basis function is LT]
+    Let $\alpha = \{\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n\}$ be a basis for vector space $V$. The function $T: V \to \mathbb{R}^n$ defined by $T(v) = [v]_\alpha$ is a linear transformation.
+\end{theorem}
+\subsection{Change of Basis}
+\begin{definition}[Change of Basis]
+    Let $\alpha = \{\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n\}$ and $\beta = \{\textbf{w}_1, \textbf{w}_2, \ldots, \textbf{w}_n\}$ be two bases for vector space $V$.
+The change of basis matrix $P$ is defined as:
+    \begin{equation}
+        P = [\textbf{w}_1, \textbf{w}_2, \ldots, \textbf{w}_n]_\alpha
+    \end{equation}
+\end{definition}
+\begin{theorem}[Change of Basis Matrix]
+    Let $\alpha = \{\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n\}$ and $\beta = \{\textbf{w}_1, \textbf{w}_2, \ldots, \textbf{w}_n\}$ be two bases for vector space $V$.
+The change of basis matrix $P$ is defined as:
+    \begin{equation}
+        P = [\textbf{w}_1, \textbf{w}_2, \ldots, \textbf{w}_n]_\alpha
+    \end{equation}
+\end{theorem}
+\paragraph{Stacking Chnage of Basis Matrix} We can stack the change of basis matrix to form a matrix $P$, and also incoperate linear transformations in other bases to a vector.
+\section{Determinants}
+\begin{definition}[Determinant]
+    Let $A \in \mathbb{R}^{n \times n}$, then the determinant of $A$ is denoted as $\det(A)$.
+\end{definition}
+\begin{example}
+    Consider the following $2\times 2$ system:
+    \begin{align*}
+        ax + by &= 0 \\
+        cx + dy &= 0
+    \end{align*}
+    The the system has the general solution:
+    \begin{align*}
+        x &= \frac{dy}{ad - bc} \\
+        y &= \frac{-cx}{ad - bc}
+    \end{align*}
+    The denominator $ad - bc$ is the determinant of the matrix. Notice that if the determinant is zero, then the system has no unique solution.
+\end{example}
+\begin{theorem}[Propeties of $2 \times 2$ Determinant]
+    Let $A = \begin{bmatrix} a & b \\ c & d \end{bmatrix}$, then:
+    \begin{align}
+        \det(A) &= ad - bc \\
+        \det\left(\begin{bmatrix}
+            br_1 + cr_1^\prime \\ r_2
+        \end{bmatrix}\right) &= b\det\left(\begin{bmatrix}
+            r_1 \\ r_2
+        \end{bmatrix}\right) + c\det\left(\begin{bmatrix}
+            r_1^\prime \\ r_2 
+        \end{bmatrix}\right) \\
+        \det\left(\begin{bmatrix}
+            r_1 \\ r_2
+        \end{bmatrix}\right) &= -\det\left(\begin{bmatrix}
+            r_2 \\ r_1
+        \end{bmatrix}\right) \\ 
+        \det I &= 1
+\end{align}
+\begin{definition}[Determinant Functions]
+    Any function $f$ that satisfies the properties:
+    \begin{enumerate}
+        \item $f\left(\begin{bmatrix}
+            br_1 + cr_1^\prime \\ r_2 \\ \vdots \\ r_n
+        \end{bmatrix}\right) = bf\left(\begin{bmatrix}
+            r_1 \\ r_2 \\ \vdots \\ r_n
+        \end{bmatrix}\right) + cf\left(\begin{bmatrix}
+            r_1^\prime \\ r_2 \\ \vdots \\ r_n
+        \end{bmatrix}\right)$ (Multilinearity in rows)
+
+        \item $f\left(\begin{bmatrix}
+            r_1 \\ r_2 \\ \vdots \\ r_i \\ \vdots \\ r_j \\ \vdots \\ r_n
+        \end{bmatrix}\right) = -f\left(\begin{bmatrix}
+            r_1 \\ r_2 \\ \vdots \\ r_j \\ \vdots \\ r_i \\ \vdots \\ r_n
+        \end{bmatrix}\right)$ (Alternating property)
+
+        \item $f(A) = (\det A )f(I)$
+    \end{enumerate}
+    is called a determinant function.
+\end{definition}
+\end{theorem}
 \end{document}

### 2025-03-09 MSE160 Lecture Notes/main.tex
march 9

@@
 \end{enumerate}
+\paragraph{Vacancies} Vacancies are missing atoms in the lattice. They are created by thermal vibrations. They are used to inhibit the movement of dislocations.
 It slightly nudges nearby atoms, creating \textbf{lattice strain} that make it harder for dislocations to move.
@@
 They are often hard, brittke phases.
-\includepdf[pages=-, angle=90]{EquationSheet.pdf}
 \paragraph{Dislcations (Slips)} Occurs on the plane with the highest planar density.
@@
 \end{equation}
+\section{Polymers} 
+\paragraph{Polymer} Polymers could be britte or elastic (elastomers). They are made of long chains of monomers. Interestingly, polymers have necking points lower than its ultimate tensile strength.
+\begin{center}
+    \textit{``Why does polymers continue to stretch after necking?''}
+\end{center}
+\paragraph{Polymer Structure} Polymers are made of long chains of monomers. The chains are held together by weak van der Waals forces. The chains are entangled.
+\\ (Example) Polyethylene are made of long chains of carbon atoms with two hydrogen atoms attached to each carbon atom.
+\paragraph{Ans: (Chain-Oriented)} The chains are originally entangled.
+When the polymer is stretched, the chains are straightened to align with the load.
+More of the load is taken by the covalent bonds, which are stronger than the van der Waals forces.
+This serves as an important strengthening mechanism.
+\paragraph{Note} In a polymer, the elastic response is from numerrous bond types.
+\paragraph{Crylistallinity} Polymers cannot be 100\% crystalline.
+A polymer could be semi-crystalline, where the polymer has both crystalline and amorphous regions.
+Since the crystalline regions are stronger, increasing the crystallinity is a important strengthening mechanism.
+\begin{enumerate}
+    \item \textbf{Tmerpeature} Increasing the crtystallinity permits the polymer have higher service temperature. 
+    \item \textbf{Chemical Resistance} Crystalline regions are more resistant to chemcial dissolution.
+\end{enumerate}
+\paragraph{Different Mer Units} Different mer units have different properties. Listed below are some common mer units:
+\begin{itemize}
+    \item \textbf{Polyethylene} (PE) - Made of long chains of carbon atoms with two hydrogen atoms attached to each carbon atom.
+    \item \textbf{Polypropylene} (PP) - Similar to PE, but with a methyl group attached to the carbon atom.
+    \item \textbf{Polyvinyl Chloride} (PVC) - Made of long chains of carbon atoms with a chlorine atom attached to each carbon atom.
+Chlorine are more electronegative than carbon, so the bond is polar and form stornger van der Waals forces.
+    \item \textbf{Polystyrene} (PS) - Made of long chains of carbon atoms with a phenyl group attached to each carbon atom.
+    \item \textbf{Polytetrafluoroethylene} (PTFE) - Made of long chains of carbon atoms with two fluorine atoms attached to each carbon atom.
+The florine is to protect the carbon backbone.
+Hoowever, it is mechanically weak as the dipole moment is low.
+    \item \textbf{Polymethyl Methacrylate} (PMMA) - Made of long chains of carbon atoms with a methyl group and a methacrylate ($-\ch{OC(=O)}$) group attached to each carbon atom.
+It is opticall tranparnet the methacrylate group prevent crtystallinity, hence amorphous and no crystal boundaries.
+\end{itemize}
+\subsection{Viscoelasitcity and Thermal Properties}
+\paragraph{Viscoelasticity} Polymers are viscoelastic.
+They have both viscous and elastic properties.
+The viscous properties are due to the entanglement of the chains.
+The elastic properties are due to the covalent bonds.
+\paragraph{Time-Dependent Behavior} For a given stress/strain, the strain/stress will decrease with time. This is due to the viscous properties of the polymer.
+\paragraph{Relaxation Modulus} The relaxation modulus is the modulus of the polymer as a function of time. It is given by:
+\begin{equation}
+    E(t) = \frac{\sigma(t)}{\epsilon_0}
+\end{equation}
+The relaxation modulus is also relate to the temperature of the polymer. The relaxation modulus decreases with temperature.
+\paragraph{Glassy Transition} The glassy transition is the temperature at which the polymer changes from a glassy state to a rubbery state.
+The glassy state is characterized by a high modulus and low strain.
+The rubbery state is characterized by a low modulus and high strain.
+The temperature is denoted as $T_g$.
+This is due to the temperature overcoming the intermolecular forces of the amorphous reigions, but the crystalline regions are still intact as the van der Waals foce is stronger.
+\paragraph{Melting} THe melting temperature is the temperature at which the polyme the intermolecular forces of the crystalline regions are overcome.
+The polymer then melts as chains slide past each other.
+\paragraph{Thermosat} Thermosat are polymers that do not melt.
+They are made of cross-linked chains or network polymers.
+The cross-links prevent the chains from sliding past each other.
+Elastomers are typically thermosets.
+\subsection{Optical Properties}
+\paragraph{PMMA} PMMA is a transparent polymer. It is used in optical applications. The transparency is due to the lack of \textbf{scattering events}.
+\begin{definition}[Scatter Event]
+    A scatter event is an event that causes light to be scattered. It is due to the difference in refractive index between the polymer and the air.
+\end{definition}
+\includepdf[pages=-, angle=90]{EquationSheet.pdf}
+
 \end{document}

### 2025-03-18 ESC195 Lecture Notes/main.tex
mar 18

@@
         \vec{a} &= \frac{d\vec{v}}{dt} = \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\vec{T}}{dt} \nonumber \\
-        \intertext{Recall that $\vec{T} = \left | \frac{d\vec{T}}{dt} | \right | \vec{N}$, we have:}
+        \intertext{Recall that $\vec{T} = \left | \frac{d\vec{T}}{dt} \right | \vec{N}$, we have:}
         \vec{a} &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\vec{T}}{dt} \nonumber \\
-        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\phi}{dt}\vec{N} \nonumber \\
-        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\kappa\vec{N}
-    \end{align}
-\end{definition}
-
+        &= \frac{d^2s}{dt^2}\vec{T} + \frac{d^2s}{dt^2}\kappa\vec{N}
+    \end{align}
+\end{definition}
+\section{Partial Derivatives}
+\subsection{Functions of Several Variables}
+\begin{definition}[Functions of Several Variables]
+    A function of several variables is a function that maps a point in $\mathbb{R}^n$ to a real number. A function of two variables is given by:
+    \begin{equation}
+        f(x, y)
+    \end{equation}
+\end{definition}
+\begin{example}[Temperature]
+    We have been duelign with funciton that maps a line (like a rod) to a real number (temperature). We can extend this to a function that maps a plane to a real number (temperature of a surface):
+    $$
+    T(x, y)
+    $$
+    Similarly, we can extend this to a function that maps a volume to a real number (temperature of a room):
+    $$
+    T(x, y, z)
+    $$
+
+\end{example}
+\begin{example}[Confinement of Plasma in Magnetic Field in Fusion Reactors]
+    Plasma are matter so hot that it is ionized. Hence, they consist in a soup of particles. It is used in fusion reactors. The plasma is confined in a magnetic field. The magnetic force is given by:
+    \begin{align}
+        \vec{F} &= q\vec{E}
+        \vec{F} &= q\vec{v} \times \vec{B} 
+    \end{align}
+    Let $\vec{B} = B_0\vec{k}$, the trajectory of the particle is given by:
+    \begin{align}
+        \vec{F} &= m\vec{r''} \nonumber \\
+        \intertext{Let $\omega_L = \frac{qB_0}{m}$ (gyrofrequency), we have:} 
+        \vec{v'} = w_L\vec{v} \times \vec{k} \nonumber
+        \intertext{Thus, we have the following 3 coupled differential equations:}
+        v_x' &= -\omega_Lv_y \nonumber \\
+        v_y' &= \omega_Lv_x \nonumber \\
+        v_z' &= 0 \nonumber
+        \intertext{Differentiating the first equation, we have:}
+        v_x'' &= -w_Lv_y' = -w_L^2v_x \nonumber \\
+        v_x = A\sin(\omega _Lt) \hspace{0.4in} v_y &= A\cos(\omega _Lt) \hspace{0.4in} v_z = C \nonumber \\
+        \vec{v} &= A\sin(\omega_Lt)i + A\cos(\omega_Lt)j + Ck \\ 
+        \vec{r} &= \int \vec{v} \, dt = \frac{A}{\omega_L}\cos(\omega_Lt)i - \frac{A}{\omega_L}\sin(\omega_Lt)j + Ct\vec{k}
+    \end{align}
+    Which, we can see is a helix.
+The current state-of-the art uses a torus to confine the plasma.
+But this leads to a cicular magnetic field; the gradient of the magnetic field is not constant, this leads to and electric field that can cause the plasma to drift:
+\end{example}
+\subsection{Partial Derivatives}
+\begin{example}[Upper Hemishere]
+    Let $f(x, y) = \sqrt{25 - x^2 - y^2}$, we cut the shphere along the surface $y=2$. We can find the slope of the curve at $x=3$. We have:
+    \begin{align*}
+        f(x, 2) &= \sqrt{25 - x^2 - 4} \\
+        f'(x, 2) &= \frac{-x}{\sqrt{25 - x^2 - 4}} \\
+        f'(3, 2) &= \frac{-3}{\sqrt{25 - 9 - 4}} = -\frac{3}{\sqrt{12}}
+    \end{align*}
+\end{example}
+\begin{definition}
+    In general, we treat the other variable as a constant. We can find the partial derivative of $f(x, y)$ with respect to $x$ as:
+    \begin{subequations}
+        \begin{equation}
+        f_x = \frac{\partial f}{\partial x} = \lim_{h \to 0} \frac{f(x+h, y) - f(x, y)}{h}
+    \end{equation}
+    Similarly, we can find the partial derivative of $f(x, y)$ with respect to $y$ as:
+    \begin{equation}
+        f_y = \frac{\partial f}{\partial y} = \lim_{h \to 0} \frac{f(x, y+h) - f(x, y)}{h}
+    \end{equation}
+    \end{subequations}
+\end{definition}
+\paragraph{Notation} We can also write the partial derivative of $f(x, y)$ with respect to $x$ as:
+\begin{equation*}
+    f_x = \frac{\partial f}{\partial x} = \frac{\partial f}{\partial x} = \frac{\partial}{\partial x}f = D_x f
+\end{equation*}
+\paragraph{Note} \textit{We prnounce $\partial$ as `die'}
+\begin{example}
+    Let $f(x, y, z) = e^{xy}\sin(xz)$, we can find the partial derivative of $f(x, y, z)$ with respect to $x$. We have:
+    \begin{align*}
+        f_x &= \frac{\partial f}{\partial x} = y e^{xy}\sin(xz) + ze^{xy}\cos(xz) \\
+        f_y &= \frac{\partial f}{\partial y} = x e^{xy}\sin(xz) \\
+        f_z &= \frac{\partial f}{\partial z} = x e^{xy}\cos(xz)
+    \end{align*}
+\end{example}
+\paragraph{Mixed Partial Derivatives} We can also find the mixed partial derivatives of $f(x, y)$ with respect to $x$ and $y$ as:
+\begin{subequations}
+    \begin{equation}
+    f_{xy} = \frac{\partial^2 f}{\partial y \partial x} = \frac{\partial}{\partial y} \left( \frac{\partial f}{\partial x} \right)
+    \end{equation}
+    \begin{equation}
+    f_{yx} = \frac{\partial^2 f}{\partial x \partial y} = \frac{\partial}{\partial x} \left( \frac{\partial f}{\partial y} \right)
+    \end{equation}
+\end{subequations}
+\begin{theorem}[Clairaut's Theorem]
+    If $f(x, y)$ is defined on an open disk $D$ containing the point $(a, b)$ and if the partial derivatives $f_{xy}$ and $f_{yx}$ are continuous on $D$, then $f_{xy}(a, b) = f_{yx}(a, b)$.
+\end{theorem}
+\subsection{Directional Derivatives and the Gradient}
+\paragraph{Differentiability and Gradients} Using the first principle, we can find the directional derivative of $f(a_1, a_2, \dots, a_n)$ as $\vec{h} \to 0$:
+\begin{equation}
+    \lim_{\vec{h} \to 0} \frac{f(\vec{a} + \vec{h}) - f(\vec{a})}{\vec{h}}
+\end{equation}
+\paragraph{A note on $\Pi$ Notation} We can write the product of a series using $\Pi$ notation:
+\begin{equation}
+    \prod_{i=1}^n a_i = a_1 \cdot a_2 \cdot \ldots \cdot a_n
+\end{equation}
 \appendix

### 2025-03-18 MAT185 Lecture Notes/main.tex
mar 18

@@
 
-\section{Linear combinations}
+\section{Linear combinations and Bases}
 \begin{definition}[Linear Combination]
@@
 \paragraph{Stacking Chnage of Basis Matrix} We can stack the change of basis matrix to form a matrix $P$, and also incoperate linear transformations in other bases to a vector.
-\section{Determinants}
+\section{Determinants and Inverses}
+\subsection{Determinants}
 \begin{definition}[Determinant]
@@
 \end{align}
+\end{theorem}
 \begin{definition}[Determinant Functions]
@@
 \end{definition}
+\begin{theorem}
+    If $f$ is alternating on rows and $A$ is a matrix that has two identical rows, then $f(A) = 0$.
+\end{theorem}
+\subsection{Constructing Determinants and Inverses}
+\begin{definition}[Minor Matrix]
+    Let $A \in \mathbb{R}^{n \times n}$, then the minor of $A$ is denoted as $A_{ij}$, which the matrix obtained by removing the $i$-th row and $j$-th column of $A$.
+\end{definition}
+\begin{example}
+    Let $A = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \end{bmatrix}$, then the minor of $A$ is:
+    $$ A_{11} = \begin{bmatrix} 5 & 6 \\ 8 & 9 \end{bmatrix} \quad A_{12} = \begin{bmatrix} 4 & 6 \\ 7 & 9 \end{bmatrix} \quad A_{13} = \begin{bmatrix} 4 & 5 \\ 7 & 8 \end{bmatrix} $$
+\end{example}
+\paragraph{Find the Determinant from a Row} We can find the determinant of a matrix by expanding along a row. We can expand along any row or column, but it is easier to expand along a row with zeros.
+\begin{definition}[Cofactor]
+    Let $A \in \mathbb{R}^{n \times n}$, then the cofactor of $A$ is denoted as $C_{ij}$, which is defined as:
+    \begin{equation}
+        C_{ij} = (-1)^{i+j}\det(A_{ij})
+    \end{equation}
+\end{definition}
+\paragraph{Checkerboard Pattern} We can use the checkerboard pattern to find the determinant of a matrix. We can expand along any row or column, but it is easier to expand along a row with zeros.
+\begin{definition}[Determinant]
+    Let $A \in \mathbb{R}^{n \times n}$, then the determinant of $A$ is denoted as $\det(A)$, which is defined as:
+    \begin{equation}
+        \det(A) = \sum_{j=1}^{n} a_{ij}C_{ij}
+    \end{equation}
+\end{definition}
+\begin{example}
+    Let $A = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \end{bmatrix}$, then the cofactor of $A$ is:
+    $$ C_{11} = \begin{bmatrix} 5 & 6 \\ 8 & 9 \end{bmatrix} \quad C_{12} = \begin{bmatrix} 4 & 6 \\ 7 & 9 \end{bmatrix} \quad C_{13} = \begin{bmatrix} 4 & 5 \\ 7 & 8 \end{bmatrix} $$
+\end{example}
+\paragraph{Find the Determinant from a Column} We can find the determinant of a matrix by expanding along a column.
+We can expand along any row or column, but it is easier to expand along a column with zeros.
+\paragraph{Using Elementary Row Operations} We can use elementary row operations to simplify the matrix on the row or column we are expanding along (by having zeros).
+\begin{theorem}[Determinant w.r.t Elementary Row Operations]
+    Consider the following elementary row operations:
+    \begin{enumerate}
+        \item \textbf{Swapping two rows}: The determinant of the matrix changes sign due to the alternating property.
+        \item \textbf{Multiplying a row by a scalar}: The determinant of the matrix is multiplied by the scalar. This is obvious.
+        \item \textbf{Adding a multiple of one row to another row}: The determinant of the matrix remains the same. This is due to the multilinearity property.
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Determinant of a Transpose]
+    Let $A \in \mathbb{R}^{n \times n}$, then:
+    \begin{equation}
+        \det(A) = \det(A^T)
+    \end{equation}
+\end{theorem}
+\begin{example}[Determinant of a $5\times5$ Matrix]
+    Let $A = \begin{bmatrix} 3 & 2 & 1 & 4 & -1 \\ 5 & 4 & 3 & 2 & 1 \\ 2 & 1 & 2 & 3 & 1 \\ -3 & 4 & 1 & 6 & 7 \\ 1 & 2 & 3 & 4 & 5 \end{bmatrix}$. We have:
+    \begin{align*}
+        \intertext{Step 1: Swap row 1 and row 5}
+        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 5 & 4 & 3 & 2 & 1 \\ 2 & 1 & 2 & 3 & 1 \\ -3 & 4 & 1 & 6 & 7 \\ 3 & 2 & 1 & 4 & -1 \end{vmatrix} \\
+        \intertext{Step 2: Get row 2 - 5 start with 0}
+        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 0 & -6 &-12 & -18 & -24 \\ 0 & -3 & -4 & -5 & -9 \\ 0 & 10 & 10 & 18 & 22 \\ 0 & -4 & -8 & -8 & -16 \end{vmatrix} \\
+        \intertext{Step 3: Get it to a UTM}
+        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 0 & -6 &-12 & -18 & -24 \\ 0 & 0 & 2 & 4 & 3 \\ 0 & 0 & 0 & 4 & 0 \\ 0 & 0 & 0 & 0 & -3 \end{vmatrix} \\
+        \intertext{We can see that each row expansion is just a multiple of the minor below it. Thus, the determinant is the product of the diagonal elements.}
+        \det A &= -1 \times -6 \times 2 \times 4 \times -3 = -144
+    \end{align*}
+\end{example}
+\begin{theorem}[The theorem of Determinants and Inverse]
+    Let $A \in \mathbb{R}^{n \times n}$, then the following properties hold:
+    \begin{enumerate}
+        \item $A$ is invertible.
+        \item $\det(A) \neq 0$.
+        \item $\text{RREF}(A) = I$.
+        \item $A$ is a product of elementary matrices.
+        \item $\det(A^{-1}) = \frac{1}{\det(A)}$.
+        \item $\text{rank}(A) = n$.
+    \end{enumerate}
+\end{theorem}
+\section{Eigen and Diagonalization}
+\subsection{Eigenvalues and Eigenvectors}
+\begin{definition}[Eigenvalues and Eigenvectors]
+    Let $A \in {^n\mathbb{R}^n}$, then $\lambda$ is an eigenvalue of $A$ if there exists a \textbf{non-zero} vector $\textbf{v}$ such that:
+    \begin{equation}
+        A\textbf{v} = \lambda \textbf{v}
+    \end{equation}
+    The vector $\textbf{v}$ is called an eigenvector of $A$.
+\end{definition}
+\begin{theorem}[Properties of Eigenvalues and Eigenvectors]
+    Let $A \in {^n\mathbb{R}^n}$, then the following properties hold:
+    \begin{enumerate}
+        \item The eigenvalues of $A$ are the roots of the characteristic equation:
+        \begin{equation}
+            \det(A - \lambda I) = 0
+        \end{equation}
+        \item The eigenvectors of $A$ are the solutions to the system of equations:
+        \begin{equation}
+            (A - \lambda I)\textbf{v} = \textbf{0}
+        \end{equation}
+    \end{enumerate}
+\end{theorem}
+\paragraph{Note} If $A\textbf{v} = \textbf{0}$, then $\textbf{v}$ is an eigenvector of $A$ with eigenvalue $\lambda = 0$.
+\begin{example}[Eigenvalues and Eigenvectors]
+    Let $A = \begin{bmatrix} 1 & 2 \\ 2 & 1 \end{bmatrix}$, then the characteristic equation is:
+    \begin{align*}
+        \det(A - \lambda I) &= \begin{vmatrix} 1-\lambda & 2 \\ 2 & 1-\lambda \end{vmatrix} \\
+        &= (1-\lambda)(1-\lambda) - 4 \\
+        &= \lambda^2 - 2\lambda - 3 = 0
+    \end{align*}
+    The roots of the characteristic equation are $\lambda = 3, -1$. The eigenvectors are:
+    \begin{align*}
+        \text{For } \lambda = 3: \begin{bmatrix} -2 & 2 \\ 2 & -2 \end{bmatrix}\textbf{v} &= \textbf{0} \implies \textbf{v} = \begin{bmatrix} 1 \\ 1 \end{bmatrix} \\
+        \text{For } \lambda = -1: \begin{bmatrix} 2 & 2 \\ 2 & 2 \end{bmatrix}\textbf{v} &= \textbf{0} \implies \textbf{v} = \begin{bmatrix} 1 \\ -1 \end{bmatrix}
+    \end{align*}
+\end{example}
+\begin{example}
+    Let $A \in {^n\mathbb{R}^n}$ and has eigen value 2 with multiplicity 2, and eigen value 3 with multiplicity 1.
+Let $x, y, z$ be eigenvectors of $A$ corresponding to the eigenvalues 2, 2, 3 respectively.
+Then:
+    \begin{enumerate}
+        \item \textbf{2x is an eigenvector of A}: Since $A(2x) = 2Ax = 2(2x)$, then $2x$ is an eigenvector of $A$ with eigenvalue 2.
+        \item \textbf{x + y is an eigenvector of A}: Since $A(x + y) = Ax + Ay = 2x + 2y = 2(x + y)$, then $x + y$ is an eigenvector of $A$ with eigenvalue 2.
+        \item \textbf{x + z is an not an eigenvector of A}: Since $A(x + z) = Ax + Az = 2x + 3z \neq \lambda(x + z)$, then $x + z$ is not an eigenvector of $A$.
+    \end{enumerate}
+\end{example}
+\subsection{Eigenspace and Multiplicity}
+\begin{definition}[Eigenspace]
+    Let $A \in {^n\mathbb{R}^n}$, then the eigenspace of $A$ corresponding to the eigenvalue $\lambda$ is denoted as $E_\lambda$, which is defined as:
+    \begin{equation}
+        E_\lambda = \{ \textbf{v} \in \mathbb{R}^n \mid A\textbf{v} = \lambda \textbf{v} \}
+    \end{equation}
+    , which is the union of all eigenvectors of $A$ corresponding to the eigenvalue $\lambda$ and ${\textbf{0}}$.
+\end{definition}
+\begin{theorem}[Eigenspace is a Subspace]
+    Let $A \in {^n\mathbb{R}^n}$, then the eigenspace of $A$ corresponding to the eigenvalue $\lambda$ is a subspace of $\mathbb{R}^n$ with dimension equal to the multiplicity of $\lambda$ ($\ge 1$).
+\end{theorem}
+\begin{proof}
+    This is obvious as eigenspace is a span of eigenvectors, and the span of eigenvectors is a subspace. To proof it, we use the subspace test:
+    \begin{enumerate}
+        \item \textbf{Non-emptiness}: Since $\textbf{0}$ satisfies $A\textbf{0} = \lambda \textbf{0}$, then $\textbf{0} \in E_\lambda$.
+        \item \textbf{Closure under addition}: Let $\textbf{v}_1, \textbf{v}_2 \in E_\lambda$, then $A(\textbf{v}_1 + \textbf{v}_2) = A\textbf{v}_1 + A\textbf{v}_2 = \lambda \textbf{v}_1 + \lambda \textbf{v}_2 = \lambda(\textbf{v}_1 + \textbf{v}_2)$, then $\textbf{v}_1 + \textbf{v}_2 \in E_\lambda$.
+        \item \textbf{Closure under scalar multiplication}: Let $\textbf{v} \in E_\lambda$, then $A(c\textbf{v}) = cA\textbf{v} = c\lambda \textbf{v} = \lambda(c\textbf{v})$, then $c\textbf{v} \in E_\lambda$.
+    \end{enumerate}
+\end{proof}
+\begin{theorem}[Intersection of Eigenspaces]
+    Let $A \in {^n\mathbb{R}^n}$, then the intersection of eigenspaces of $A$ corresponding to distinct eigenvalues is $\{\textbf{0}\}$, that is, for eigenvalues $\lambda_1, \lambda_2$, if $\lambda_1 \neq \lambda_2$, then $E_{\lambda_1} \cap E_{\lambda_2} = \{\textbf{0}\}$.
+\end{theorem}
+\begin{definition}[Algebraic Multiplicity (AM)]
+    Let $A \in {^n\mathbb{R}^n}$, then the algebraic multiplicity of an eigenvalue $\lambda$ is the number of times $\lambda$ appears as a root of the characteristic equation.
+\end{definition}
+\begin{definition}[Geometric Multiplicity (GM)]
+    Let $A \in {^n\mathbb{R}^n}$, then the geometric multiplicity of an eigenvalue $\lambda$ is the dimension of the eigenspace of $\lambda$.
+\end{definition}
+\begin{theorem}[Multiplicity of Eigenvalues]
+    Let $A \in {^n\mathbb{R}^n}$, then the algebraic multiplicity of an eigenvalue is greater than or equal to the geometric multiplicity of the eigenvalue, that is:
+    \begin{equation}
+        \text{AM}(\lambda) \ge \text{GM}(\lambda) \ge 1
+    \end{equation}
+\end{theorem}
+\begin{theorem}[Product of Eigenvalues]
+    Let $A \in \mathbb{R}^{n \times n}$, then the product of the eigenvalues of $A$ is equal to the determinant of $A$:
+    \begin{equation}
+        \prod_{i=1}^{n} \lambda_i = \det(A)
+    \end{equation}
 \end{theorem}

### 2025-03-18 MSE160 Lecture Notes/main.tex
mar 18

@@
 \includepdf[pages=-, angle=90]{EquationSheet.pdf}
+\subsection{Matter-Energy Interactions}
+\paragraph{Photo Wavelength-Frequency} The wavelength of a photon is inversely proportional to its frequency. The relationship is given by:
+\begin{equation}
+    c = f\lambda
+\end{equation}
+\paragraph{Energy of a Photon} The energy of a photon is given by:
+\begin{equation}
+    E = hf = \frac{hc}{\lambda}
+\end{equation}
+\paragraph{Visible Light} Visible light is the range of wavelengths that the human eye can see. It is between 400 nm and 700 nm (3eV to 2eV).
+\paragraph{Bohr's Model} Bohr's model is a model of the atom that describes the electron as a particle in a circular orbit around the nucleus.
+It discribes electrons in well-defined qunatized orbits.
+Thus, electrion promotes and demotes between orbitals, creating discrete spectra.
+\paragraph{Lmitation of Bohr's Model} Bohrs model cannot captures the energy levels and orbitals of atoms with more than one electron. It also cannot explain the fine structure of the spectra.
+\paragraph{Wave-Mechanical Model} The wave-mechanical model describes the electron as a wave.
+It describes the electron as a wavefunction, $\psi$, that describes the probability of finding the electron at a certain position.
+\paragraph{Quantum Numbers} The wave-mechanical model uses quantum numbers to describe the electron. The quantum numbers are:   
+\begin{enumerate}
+    \item \textbf{Principal Quantum Number} $n = 1,2,3,\dots$ - Describes the energy level of the electron (size of the orbital). \\
+    (note) $n$ is sometimes given by $n = K, L, M, \dots$
+    \item \textbf{Angular Momentum Quantum Number} $l = 0,1,2,\dots,n-1$ - Describes the shape of the orbital.
+    (note) Those are the $s,p,d,f$ orbitals.
+    \begin{enumerate}
+        \item \textbf{$l=0$ or $s$} Spherical
+        \item \textbf{$l=1$ or $p$} Dumbbell
+        \item \textbf{$l=2$ or $d$} Four-leaf clover
+        \item \textbf{$l=3$ or $f$} Complex
+    \end{enumerate}
+    \item \textbf{Magnetic Quantum Number} $m_l = -l, -l+1, \dots, 0, \dots, l-1, l$ - Describes the orientation of the orbital.
+    \item \textbf{Spin Quantum Number} $m_s = \pm \frac{1}{2}$ - Describes the spin of the electron.
+\end{enumerate}
+\paragraph{Relative Energy Levels} The energy levels of the orbitals are given by:
+\begin{equation}
+    E = -\frac{Z^2}{n^2} \text{Ry}
+\end{equation}
+\paragraph{Electrion Configuration} The electron configuration is the distribution of electrons in the orbitals. It is given by:
+\begin{equation}
+    \text{Element} = [n]l^m
+\end{equation}
+(Example) The electron configuration of Carbon is $1s^2 2s^2 2p^2$. That for iron is $1s^2 2s^2 2p^6 3s^2 3p^6 4s^2 3d^6$.
+\paragraph{Stable Octet} The $ns^2 np^6$ is a stable configuration that forms the basis for the primary bonds.
+Atoms tend to gain, lose or share electrons to achieve this configuration.
+As we know, if we share electrons, we form covalent bonds; if we lose or gain electrons, we form ionic bonds.
+\paragraph{Half filled and Fully Filled Orbitals} Half filled and fully filled orbitals are more stable than partially filled orbitals.
+Thus at $d^5$ and $d^{10}$, the $d$ orbitals are more stable and it tends to lend electron from the $s$ orbital.\\
+(Example) Chromium is $[Ar] 3d^5 4s^1$ and Copper is $[Ar] 3d^{10} 4s^1$.
+\paragraph{Pauli Exclusion Principle} The Pauli Exclusion Principle states that no two electrons can have the same set of quantum numbers. 
+\paragraph{Tightly Packed Atoms} When atoms are packed, due to the pauli exclusion principle, it creates a bend of energy levels of the same orbital.
+In ground state, the electrons will fill the lower energy levels first.
+\subsection{Band Theory}
+\paragraph{Conductivity} At room temperature, the thermal promotion within the bend makse it possible for conduction in conductors.
+\paragraph{Overlaping Energy Level} In some atoms, the bend of energy level gets larger and overlap with the energy level of the next orbital. 
+\paragraph{Conductive Band and Valence Band} The conductive band is the band of energy levels that are above the valence band.
+The valence band is the band of energy levels that are filled with electrons.
+There is a band gap between the two bands, the size of the band gap determines the conductivity of the material.
+\paragraph{Conductors} Conductors have a no band gap. The valence band and conductive band overlap. Thus, electrons can easily move between the two bands.
+\paragraph{Insulators} Insulators have a large band gap. The valence band and conductive band are far apart. Thus, electrons cannot easily move between the two bands.
+\paragraph{Semiconductors} Semiconductors have a small/moderate band gap.
+The valence band and conductive band are close together.
+Thus, electrons can move between the two bands with the addition of energy.
+\textbf{For This Coruse, we take the band gap for semiconductors to be $\le 4$ eV.}
+\paragraph{Optically transparency} A material would be optically transparent if the band gap is larger than the energy of the visible light (3 eV).
+\subsection{Silicon}
+\paragraph{Intrinsic Silicon} Intrinsic silicon is pure silicon.
+When electron is promoted to the conduction band, it leaves a hole in the valence band.
+The hole can move through the lattice as a postive charge.
+The electron can move through the lattice as a negative charge.
+The electron-hole pair is called an exciton.
+For intrinsic silicon, the number of holes is equal to the number of electrons.
+Also, the charge carrier contcentration is given by:
+\begin{equation}
+    \frac{n=p}{a^3}
+\end{equation}
+For silicon, at room temperature, the charge carrier concentration is $10^{16}$ m$^{-3}$.
+\paragraph{Doping} Doping is the addition of impurities to a semiconductor.
+They are point defects in the lattice.
+Doping is used to increase the conductivity of the semiconductor.
+There are two types of doping:
+\begin{enumerate}
+    \item \textbf{N-Type Doping} N-type doping is the addition of impurities that increase the number of electrons.
+The extra electron is in the higher end of the band gap.
+The impurities are called donor impurities.
+The donor impurities are typically group 15 elements.
+The donor impurities are negatively charged.
+    \item \textbf{P-Type Doping} P-type doping is the addition of impurities that increase the number of holes.
+The impurities are called acceptor impurities.
+The acceptor impurities are typically group 13 elements.
+The acceptor impurities are positively charge the the hole is on the higher end of the valence band and to be promoted to the lower end of the band gap.
+\end{enumerate}
 

### 2025-04-14 ESC195 Lecture Notes/main.tex
exam

@@
 \end{equation}
+\begin{definition}[Differentiability]
+    We say that $f$ is differentiable at $\vec{x}$ if there exists a vector $\nabla f$ such that:
+    \begin{equation}
+        \lim_{\vec{h} \to 0} \frac{f(\vec{x} + \vec{h}) - f(\vec{x}) - \nabla f \cdot \vec{h}}{\text{norm}(\vec{h})} = 0
+    \end{equation}
+    Thus,
+    \begin{equation}
+        f(\vec{x} + \vec{h}) - f(\vec{x}) = \nabla f \cdot \vec{h} + o(\vec{h})
+    \end{equation}
+    , where if $\lim_{\vec{h} \to 0} \frac{g(\vec{h})}{\text{norm}(\vec{h})} = 0$, then $g(\vec{h}) = o(\vec{h})$.
+\end{definition}
+\begin{definition}[Gradient]
+    The gradient of a function $f(x, y)$ is a vector that points in the direction of the greatest increase of the function. The gradient is given by:
+    \begin{equation}
+        \nabla f = \frac{\partial f}{\partial x}i + \frac{\partial f}{\partial y}j
+    \end{equation}
+\end{definition}
+\begin{example}
+    Let $f(x, y) = xyz$, we can find the gradient of the function from first principles with $h = h_1i + h_2j + h_3k$. We have:
+    \begin{align*}
+        f(x+h_1, y+h_2, z+h_3) - f(x, y, z) &= (x+h_1)(y+h_2)(z+h_3) - xyz \\
+        &= yzh_1 + xzh_2 + xyh_3 + h_1h_2z + h_1h_3y + h_2h_3x + h_1h_2h_3 - xyz \\
+        &= \vec{h} \cdot (yz i + xz j + xy k) + xh_1h_2 + yh_1h_3 + zh_2h_3 + h_1h_2h_3 \\
+    \end{align*}
+    Now, we consider the term $o(\vec{h})$, WLOG, consider $g(x) =xh_1x_3 = xh_2k\cdot \vec{h}$. We have:
+    \begin{align*}
+        \lim_{\vec{h} \to 0} \frac{|g(\vec{h})|}{\text{norm}(\vec{h})} &= \lim_{\vec{h} \to 0} \frac{|xh_2|\text{norm}(\vec{h})|\cos \theta|}{\text{norm}(\vec{h})} \le |xh_2| \\ 
+        \intertext{Since we have:}
+        \lim_{\vec{h} \to 0} |xh_2| &= 0\\
+        \intertext{So $o(\vec{h}) = 0$, and we have:}
+        \nabla f &= yz i + xz j + xy k
+    \end{align*}
+\end{example}
+\begin{theorem}[Gradient is Partial Derivative]
+    The gradient of a function $f(x, y)$ is given by:
+    \begin{equation}
+        \nabla f = \frac{\partial f}{\partial x}i + \frac{\partial f}{\partial y}j
+    \end{equation}
+    And in general, the gradient of a function $f(x_1, x_2, \dots, x_n)$ is given by:
+    \begin{equation}
+        \nabla f = \frac{\partial f}{\partial x_1}b_1 + \frac{\partial f}{\partial x_2}b_2 + \dots + \frac{\partial f}{\partial x_n}b_n
+    \end{equation}
+\end{theorem}
+\begin{example}
+    Given $T_x = 3$ and $T_y = 4$ at some point $(x_0, y_0)$, we can find the gradient of $T$ at that point. We have:
+    \begin{align*}
+        \nabla T &= 3i + 4j
+    \end{align*}
+    Also, the magnitude of the gradient is 5.
+\end{example}
+\begin{example}
+    Let $f(x, y, z) = xy^2z^3$, we can find the gradient of the function. We have:
+    \begin{align*}
+        f_x &= y^2z^3 \\
+        f_y &= 2xyz^3 \\
+        f_z &= 3xy^2z^2 \\
+        \intertext{Thus, we have:}
+        \nabla f &= y^2z^3i + 2xyz^3j + 3xy^2z^2k
+    \end{align*}
+\end{example}
+\begin{example}
+    Let $\vec{r} = x^2i + y^2j + z^2k$, we can find the gradient of the function radius funciton $r = \sqrt{x^2 + y^2 + z^2}$. We have:
+    \begin{align*}
+        r &= \sqrt{x^2 + y^2 + z^2} \\
+        r_x &= \frac{2x}{2r} = \frac{x}{r} \\
+        r_y &= \frac{y}{r} \\
+        r_z &= \frac{z}{r} \\
+        \intertext{Thus, we have:}
+        \nabla r &= \frac{x}{r}i + \frac{y}{r}j + \frac{z}{r}k = \frac{\vec{r}}{r}
+    \end{align*}
+\end{example}
+\begin{definition}[Directional Derivative]
+    The directional derivative of a function $f(x_1, x_2, \dots, x_n)$ in the direction of a unit vector $\hat{u}$ is given by:
+    \begin{equation}
+        D_{\hat{u}} = f_{\hat{u}} = \lim_{h \to 0} \frac{f(\vec{x} + h\hat{u}) - f(\vec{x})}{h}
+    \end{equation}
+\end{definition}
+\paragraph{Dot Product} We can write the directional derivative of $f(x, y)$ in the direction of a unit vector $\hat{u}$ as:
+\begin{equation}
+    f_{\hat{u}} = \nabla f \cdot \hat{u}
+\end{equation}
+\begin{example}
+    Let $f(x, y) = x^2 + y^2$, we can find the directional derivative of the function in the direction of the unit vector $\hat{u} = \frac{1}{\sqrt{2}}i + \frac{1}{\sqrt{2}}j$. We have:
+    \begin{align*}
+        \nabla f &= 2xi + 2yj \\
+        \hat{u} &= \frac{1}{\sqrt{2}}i + \frac{1}{\sqrt{2}}j \\
+        f_{\hat{u}} &= \nabla f \cdot \hat{u} = 2x\frac{1}{\sqrt{2}} + 2y\frac{1}{\sqrt{2}} = \frac{2x + 2y}{\sqrt{2}} = \sqrt{2}(x + y)
+    \end{align*}
+\end{example}
+\begin{example}[Path of Steepest Descent]
+    Let $f(x, y) = A + x + 2y - x^2 - 3y^2$, we can find the direction of the path of steepest descent from $(0, 0 , A)$. We have:
+    \begin{align*}
+        \nabla f &= (1-2x)i + (2-6y)j \\
+        \intertext{Let $C: \vec{r}(t) = x(t)i + y(t)j$ be the path of steepest descent, we have:}
+        \vec{r'}(t) &= -\nabla f \\
+        x'(t) &= 2x - 1 \\
+        y'(t) &= 3y - 2 \\
+        \intertext{Thus, we have:}
+        \frac{dy}{dx} &= \frac{dy/dt}{dx/dt} = \frac{3y - 2}{2x - 1}
+        \intertext{Solving for $y$ in terms of $x$, we have:}
+        3y &= (2s - 1)^3 + 1
+    \end{align*}
+\end{example}
+\begin{example}[Quadric Surface]
+    Let $z(x, y) = 20 - x^2 - y^2$, we can find the gradient of the function. We have:
+    \begin{align*}
+        z_x &= -2x \\
+        z_y &= -2y \\
+        \intertext{Thus, we have:}
+        \nabla z &= -2xi - 2yj
+    \end{align*}
+    Now, we can find the directional derivative of the function at (1, 2) along the direction of (-1, 1). We have:
+    \begin{align*}
+        \intertext{We first find the unit vector in the direction of (-1, 1):}
+        \hat{u} &= \frac{-1}{\sqrt{2}}i + \frac{1}{\sqrt{2}}j \\
+        z_{\hat{u}}(1, 2) &= \nabla z \cdot \hat{u} = \frac{2(1) + 2(2)}{\sqrt{2}} = 3\sqrt{2}
+    \end{align*}
+\end{example}
+\begin{definition}[Level Curves]
+    The level curves of a function $f(x, y)$ are the curves in the $xy$ plane where $f(x, y) = c$ for some constant $c$.
+\end{definition}
+\begin{example}
+    Let $f(x, y) = x^2 + y^2$, we can find the level curves of the function. We have:
+    \begin{align*}
+        x^2 + y^2 &= c \\
+        \intertext{Thus, we have:}
+        y &= \pm \sqrt{c - x^2}
+    \end{align*}
+\end{example}
+\subsection{The Chain Rule}
+\begin{theorem}[Chain Rule along a Curve]
+    Let $f(x, y, z)$ be a function of three variables and let $\vec{r}(t) = x(t)i + y(t)j + z(t)k$ be a curve in the $xy$ plane.
+The chain rule states that the derivative of $f$ with respect to $t$ is given by:
+    \begin{align}
+        \frac{d}{dt}[f(\vec{r}(t))] &= \nabla f (\vec{r}(t)) \cdot \vec{r'}(t)
+        &= \frac{\partial f}{\partial x} \frac{dx}{dt} + \frac{\partial f}{\partial y} \frac{dy}{dt} + \frac{\partial f}{\partial z} \frac{dz}{dt} 
+        &= ()\nabla f \cdot T)\frac{ds}{dt}
+    \end{align}
+\end{theorem}
+\begin{example}
+    Let $\vec{r}(t) = t^3 i + \cos t j$ and $T(x, y) = xy^2$, we can find the derivative of $T$ with respect to $t$. We have:
+    \begin{align*}
+        T(x, y) &= xy^2 \\
+        x(t) &= t^3 \\
+        y(t) &= \cos t \\
+        \frac{dT}{dt} &= \frac{\partial T}{\partial x} \frac{dx}{dt} + \frac{\partial T}{\partial y} \frac{dy}{dt} \\
+        &= y^2 \cdot 3t^2 + 2xy \cdot -\sin t \\
+        &= 3t^2\cos^2 t - 2t^3\cos t \sin t
+    \end{align*}
+\end{example}
+\begin{theorem}[Implicit Differentiation]
+    Let $F(x, y) = 0$ be an equation that defines $y$ as a function of $x$. The derivative of $y$ with respect to $x$ is given by:
+    \begin{equation}
+        \frac{dy}{dx} = -\frac{f_x}{f_y}
+    \end{equation}
+\end{theorem}
+\paragraph{Parametrization} We can also parametrize the curve $F(x, y) = 0$ as $x=t$, $y=g(t)$, and then use the chain rule to find the derivative of $y$ with respect to $x$.
+\subsection{Tangent Planes and Linear Approximations}
+\begin{example}[Gradient is Normal to Level Curves]
+    Let $C: 20 - x^2 - y^2 = z$.
+At $\vec{r} = (1,2)$, we have $z = 15$ and a level curve of $x^2 + y^2 = 5$.
+Also, let $\vec{t} = t_1i + t_2j$ be a tangent vector to the level curve at (1, 2).
+We have:
+    \begin{align*}
+        \vec{t}\cdot \vec{r} &= 0 \\
+        t_1(1) + t_2(2) &= 0 \\
+        \intertext{We choose $t_1 = 2$, then:}
+        2(1) + t_2(2) &= 0 \\
+        t_2 &= -1 \\
+        \intertext{Thus, the tangent vector is $2i - j$. We can find the gradient of the function at (1, 2):}
+        2(1) + 2(2) &= 0 \\
+        \nabla z = -2xi - 2yj \\
+        \nabla z \cdot \vec{t} &= -4x + 2y = -4 + 4 = 0
+    \end{align*}
+\end{example}
+\begin{definition}[Tangent Plane]
+    The tangent plane to a surface $z = f(x, y)$ at a point $(x_0, y_0, z_0)$ is a plane that is tangent to the surface at that point.
+\end{definition}
+\begin{definition}[Linear Approximation]
+    The linear approximation of a function $f(x, y)$ at a point $(x_0, y_0)$ is a linear function that approximates the function at that point.
+\end{definition}
+\begin{theorem}[Tangent Plane]
+    The tangent plane to a surface $z = f(x, y)$ at a point $(x_0, y_0, z_0)$ is given by:
+    \begin{equation}
+        z = f(x_0, y_0) + f_x(x_0, y_0)(x - x_0) + f_y(x_0, y_0)(y - y_0)
+    \end{equation}
+\end{theorem}
+\subsection{Maximum and Minimum Values}
+\begin{definition}[Local Maximum and Minimum]
+    A function $f(x, y)$ has a local maximum at a point $(x_0, y_0)$ if $f(x, y) \le f(x_0, y_0)$ for all points $(x, y)$ in a small disk around $(x_0, y_0)$.
+A function $f(x, y)$ has a local minimum at a point $(x_0, y_0)$ if $f(x, y) \ge f(x_0, y_0)$ for all points $(x, y)$ in a small disk around $(x_0, y_0)$.
+\end{definition}
+\begin{theorem}
+    If $f(x, y)$ has a local maximum or minimum at a point $(x_0, y_0)$ and the partial derivatives $f_x(x_0, y_0)$ and $f_y(x_0, y_0)$ exist, then $f_x(x_0, y_0) = 0$ and $f_y(x_0, y_0) = 0$.
+Otherwise, the gradient DNE.
+\end{theorem}
+\begin{definition}[Critical Point]
+    A point $(x_0, y_0)$ is a critical point of a function $f(x, y)$ if $\nabla f(x_0, y_0) = 0$ or the gradient DNE.
+\end{definition}
+\begin{definition}[Stationary Point]
+    A point $(x_0, y_0)$ is a stationary point of a function $f(x, y)$ if $f_x(x_0, y_0) = 0$ and $f_y(x_0, y_0) = 0$.
+\end{definition}
+\begin{definition}[Saddle Point]
+    A point $(x_0, y_0)$ is a saddle point of a function if it is neither a local maximum nor a local minimum, but the gradient is zero.
+\end{definition}
+\begin{theorem}[Second Derivative Test]
+    Let $f(x, y)$ be a function of two variables and let $(x_0, y_0)$ be a critical point of $f$.
+We have the following cases depending on $\det{J} = f_{xx}f_{yy} - f_{xy}^2$, where $J$ is the Hessian matrix of $f$:
+    \begin{enumerate}
+        \item If $\det{J} > 0$ and $f_{xx} > 0$, then $f$ has a local minimum at $(x_0, y_0)$.
+        \item If $\det{J} > 0$ and $f_{xx} < 0$, then $f$ has a local maximum at $(x_0, y_0)$.
+        \item If $\det{J} < 0$, then $f$ has a saddle point at $(x_0, y_0)$.
+        \item If $\det{J} = 0$, then the test is inconclusive.
+    \end{enumerate} 
+\end{theorem}
+\paragraph{Tips/Tricks}
+\begin{enumerate}
+    \item Check if the gradient is DNE. If it is, then the point is a critical point.
+    \item Check if the gradient is zero. If it is, then the point is a stationary point.
+    \item Check the boundry if we are dealing with a closed interval.
+    \item Check the endpoints of the boundry if we are dealing with a closed interval.
+\end{enumerate}
+\subsection{Lagrange Multipliers}
+\begin{theorem}[Lagrange Multipliers]
+    Let $f(x, y)$ be a function of two variables and let $g(x, y) = 0$ be a constraint.
+The method of Lagrange multipliers states that the maximum or minimum of $f(x, y)$ subject to the constraint $g(x, y) = 0$ occurs at a point $(x_0, y_0)$ where:
+    \begin{equation}
+        \nabla f(x_0, y_0) = \lambda \nabla g(x_0, y_0)
+    \end{equation}
+    where $\lambda$ is a Lagrange multiplier. Thus we can set up a system of equations:
+    \begin{align}
+        f_x(x_0, y_0) &= \lambda g_x(x_0, y_0) \\
+        f_y(x_0, y_0) &= \lambda g_y(x_0, y_0) \\
+        g(x_0, y_0) &= k
+    \end{align}
+
+\end{theorem}
+\subsection{Reconstructing a Function from its Gradient}
+\begin{theorem}[Reconstructing a Function from its Gradient]
+    Let $f(x, y)$ be a function of two variables and let $\nabla f(x, y) = g(x, y)$. The function $f(x, y)$ can be reconstructed from its gradient by integrating the gradient:
+    \begin{equation}
+        f(x, y) = \int g_x(x, y) \, dx + \int g_y(x, y) \, dy
+    \end{equation}
+    where the integrals are taken with respect to $x$ and $y$ respectively.
+    The function $f(x, y)$ is unique up to a constant of integration.
+    \begin{equation}
+        f(x, y) = \int g_x(x, y) \, dx + \int g_y(x, y) \, dy + C
+    \end{equation}
+    where $C$ is a constant of integration.
+\end{theorem}
+\subsection{Rocket Science}
+\paragraph{Tsiovsky Rocket Equation} The Tsiovsky rocket equation is given by:
+\begin{equation}
+    \Delta v = I_{sp} \ln \left( \frac{m_0}{m_f} \right)
+\end{equation}
+, where $\Delta v$ is the change in velocity, $I_{sp}$ is the specific impulse of the fuel, $m_0$ is the initial mass of the rocket, and $m_f$ is the final mass of the rocket.
+\paragraph{Escape Velocity} The escape velocity of a rocket is given by:
+\begin{equation}
+    v_e = \sqrt{2g_0R}
+\end{equation}
+, on earth, $v_e = 11.2 km/s$.
+\paragraph{Specific Impulse} The specific impulse of a rocket is given by:
+\begin{equation}
+    I_{sp} = \frac{F}{\dot{m}g_0}
+\end{equation}
+\begin{example}[Liquuid Oxygen Kerosene Rocket]
+    The specific impulse of a liquid oxygen kerosene rocket is given by:
+    $$
+        I_{sp} = 3.2 km/s
+    $$
+    Thus, we have:
+    $$
+    \frac{m_i}{m_f} = e^{\frac{v_e}{I_{sp}}} = 31.1
+    $$
+    The mass of the paylod have to be 1/31.1 of the mass of the rocket. Which means that the rocket is $97\%$ fuel. Which is not good. Let's try to optimize that. First, we have the initial mass:
+    $$
+    m_i = A + m_{fuel} + m_s
+    $$
+    , where $A$ is the mass of the payload, $m_{fuel}$ is the mass of the fuel, and $m_s$ is the mass of the structure.\\
+    We define $S$, the structure mass fraction as:
+    $$
+    S = \frac{m_s}{m_i + m_s}
+    $$
+    And define rocket mass as:
+    $$
+    m_r = m_s + m_{fuel}
+    $$
+    So we have:
+    $$
+        \frac{m_i}{m_f} = \frac{A + m_r}{A + Sm_r} = \frac{1 + \frac{m_r}{A}}{1 + S}
+    $$
+    We then have the $\Delta V$ equation:
+    $$
+    \Delta V = I_{sp} \ln(1/0.2) = 5.15 km/s
+    $$
+    which is not enough to get to orbit. We could stage the rocket:
+    \begin{align*}
+        \Delta V_{total} &= \Delta V_1 + \Delta V_2 + \Delta V_3 \\
+        &= I_{sp} \ln \left( \frac{A + m_1 + m_2 + m_3}{A + Sm_1 + m_2 + m_3} \right) + I_{sp} \ln \left( \frac{A + m_2 + m_3}{A + Sm_2 + m_3} \right) + I_{sp} \ln \left( \frac{A + m_3}{A + Sm_3} \right) \\
+        &= I_{sp} (\ln \left( \frac{A + m_1 + m_2 + m_3}{A + Sm_1 + m_2 + m_3} \right) + \ln \left( \frac{A + m_2 + m_3}{A + Sm_2 + m_3} \right) + \ln \left( \frac{A + m_3}{A + Sm_3} \right) ) \\
+    \end{align*}
+    Thus we have the constraint $\Delta V_{total} = 11.2 / 3.2 = 3.5$. And we minimize:
+    $$
+    f(m_1, m_2, m_3) = m_1 + m_2 + m_3
+    $$
+    So $\nabla f = (1, 1, 1)$ and we have the following system of equations be setting $\lambda \nabla f = \nabla g$:
+    \begin{align*}
+       \lambda &= \frac{1}{A + m_1 + m_2 + m_3} - \frac{S}{A + Sm_1 + m_2 + m_3} \\
+        \lambda &= \frac{1}{A + m_1 + m_2 + m_3} - \frac{1}{A + Sm_1 + m_2 + m_3} + \frac{1}{A + m_3} - \frac{S}{A + Sm_3} \\ 
+        \lambda &= \frac{1}{A + m_1 + m_2 + m_3} - \frac{1}{A + Sm_1 + m_2 + m_3} + \frac{1}{A + m_2 + m_3} - \frac{1}{A + Sm_2 + m_3} + \frac{1}{A + m_3} - \frac{S}{A + Sm_3} \\
+        \end{align*}
+    We can also, with rearranging:
+    $$
+    0 = \frac{1}{A + Sm_1 + m_2 + m_3} - \frac{S}{A + Sm_2 + m_3} + \frac{S}{A + Sm_3} - \frac{1}{A + Sm_3}
+    $$ 
+    We evantullaly find that the arguments in the logarithm are equal. Thus, we have:
+    $$
+    \frac{A + m_1 + m_2 + m_3}{A + Sm_1 + m_2 + m_3} = \frac{A + m_2 + m_3}{A + Sm_2 + m_3} = \frac{A + m_3}{A + Sm_3}
+    $$
+    Thus, we can use the following equation:
+    \begin{equation}
+        N = e^{\Delta V / (kI_{sp})}
+    \end{equation}
+    , where $k=3$ is the number of stage and $N$ is the terms in the logarithm. we can now solve for the masses of each stage:
+    \begin{align*}
+        m_3 &= A\left(\frac{N-1}{1-SN}\right)
+        m_2 &= (A + m_3)\left(\frac{N-1}{1-SN}\right)
+        m_1 &= (A + m_2 + m_3)\left(\frac{N-1}{1-SN}\right)
+    \end{align*}
+    Solving for $m_1$, $m_2$, and $m_3$ to $A=10000$kg into orbit, and $S=0.2$:
+    \begin{align*}
+        m_3 &= 24,100kg \\
+        m_2 &= 82,100kg \\
+        m_1 &= 280,000kg \\
+        m_{total} &= 386,200kg \\
+        \Delta V &= 11.2km/s
+    \end{align*}
+\end{example}
+\subsection{Differentiability with respect to a parameter}
+\begin{theorem}
+    Let $f(x, y)$ be a function of two variables in close rectangle $\{(x,y) \in \mathbb{R}^2: a \leq x \leq b, c \leq y \leq d\}$, then $f(x, y)$ have contious deritivcece with respect to $x$, then:
+    $$ \frac{dF}{dx} = \frac{d}{dx}\int^d_c f(x, y) dy = \int^d_c \frac{\partial f}{\partial x} dy$$
+    and:
+    $$ \frac{dF}{dy} = \frac{d}{dy}\int^b_a f(x, y) dx = \int^b_a \frac{\partial f}{\partial y} dx$$
+    where $F(x,y) = \int^d_c f(x, y) dy$ and $F(x,y) = \int^b_a f(x, y) dx$.
+\end{theorem}
+\begin{example}
+    Let $A(t) = \int_{x_1(t)}^{x_2(t)} f(x) dx$. To obtain $A'(t)$, we first have $\Delta A$ for some $\Delta t$:
+    \begin{align*}
+        A(t + \Delta t) - A(t) &= \int_{x_1(t + \Delta t)}^{x_2(t) + \Delta t} f(x) dx - \int_{x_1(t)}^{x_2(t+ \Delta t)} f(x) dx \\
+    \end{align*}
+    Recall, the MVT for integrals:
+    \begin{align*}
+        \int b_ a f(x) dx &= f(c)(b-a) \\
+    \end{align*}
+    So we would be arriving at the chauin rule:
+    \begin{align*}
+        A'(t) &= \frac{d}{dt} \int_{x_1(t)}^{x_2(t)} f(x) dx \\
+        &= f(x_2(t))\frac{dx_2}{dt} - f(x_1(t))\frac{dx_1}{dt} + \int_{x_1(t)}^{x_2(t)} f'(x) dx
+        &= f(x_2(t))\frac{dx_2}{dt} - f(x_1(t))\frac{dx_1}{dt} + \int_{x_1(t)}^{x_2(t)} \frac{\partial f}{\partial x} dx
+    \end{align*}
+\end{example}
+\begin{theorem}[Leibniz's Rule]
+    Given a reigion $R$ in the $xy$ plane, and a functions $\phi_1(x)$ and $\phi_2(x)$ have continuous derivatives with $x$ in $R$, and that we have continuous differentiable functions $f(x, y)$ in $R$, if:
+    \begin{align*}
+       F(x) &= \int_{\phi_1(x)}^{\phi_2(x)} f(x, y) dy \\
+    \end{align*}
+    Then:
+    \begin{align*}
+        F'(x) &= f(x, \phi_1(x))\frac{d\phi_1}{dx} - f(x, \phi_2(x))\frac{d\phi_2}{dx} + \int_{\phi_1(x)}^{\phi_2(x)} \frac{\partial f}{\partial x} dy
+    \end{align*}
+\end{theorem}
+\begin{example}[Integral Equations]
+    Let:
+    $$ f(x) = x - 2\int_0^x e^{2(x-t)} f(t) dt $$
+    We can solve this by turing it into a differential equation:
+    \begin{align*}
+        f'(x) &= 1 - 2e^{2(x-x)}f(x) + 2\int_0^x e^{2(x-t)}f'(t) dt \\
+        &= 1 - 2\left[2\int_{x}^{0} e^{2(x-t)}f(t) dt\right] - 2f(x) \\ 
+        &= 1 - 2x + 2f(x) - 2f(x) \\
+        &= 1 - 2x \\
+        \intertext{Thus, we have:}
+        f(x) &= \int_0^x (1 - 2t) dt \\
+        &= x - x^2 \\
+        \intertext{Thus, we have:}
+        f(x) &= \frac{1}{2}x^2 - \frac{1}{3}x^3 + C \\
+        \intertext{We can find the constant $C$ by substituting $x=a$ into the original equation:}
+        f(a) &= a - 2\int_0aa e^{2(a-t)} f(t) dt = a
+        So we have:
+        c &= a^2
+        \end{align*}
+\end{example}
+\begin{example}[Feynman's Trick]
+    Let:
+    $$ F(x) = \int_0^1 \frac{y^x -1}{\ln y} dy $$
+    We can find the derivative of $F(x)$ with respect to $x$:
+    \begin{align*}
+        F'(x) &= \int_0^1 \frac{\partial}{\partial x} \left( \frac{y^x - 1}{\ln y} \right) dy \\
+        &= \int_0^1 \frac{y^x \ln y}{\ln y} dy \\
+        &= \int_0^1 y^x dy \\
+        &= \left[ \frac{y^{x+1}}{x+1} \right]_0^1 \\
+        &= \frac{1}{x+1} - 0 \\
+        &= \frac{1}{x+1} \\
+        \intertext{Thus, we have:}
+        F'(x) &= \frac{1}{x+1} \\
+        F(x) &= \ln(x+1) + C \\
+        \intertext{We can find the constant $C$ by substituting $x=0$ into the original equation:}
+        F(0) &= \int_0^1 \frac{y^0 - 1}{\ln y} dy = 0
+        \intertext{Thus, we have:}
+        C &= 0 \\
+        \intertext{Thus, we have:}
+        F(x) &= \ln(x+1) \\
+        \end{align*}
+\end{example}
 \appendix

### 2025-04-14 MAT185 Lecture Notes/main.tex
exam

@@
 \begin{definition}[Minor Matrix]
-    Let $A \in \mathbb{R}^{n \times n}$, then the minor of $A$ is denoted as $A_{ij}$, which the matrix obtained by removing the $i$-th row and $j$-th column of $A$.
+    Let $A \in {^n\mathbb{R}^n}$, then the minor of $A$ is denoted as $A_{ij}$, which the matrix obtained by removing the $i$-th row and $j$-th column of $A$.
 \end{definition}
@@
 \begin{definition}[Cofactor]
-    Let $A \in \mathbb{R}^{n \times n}$, then the cofactor of $A$ is denoted as $C_{ij}$, which is defined as:
+    Let $A \in  {^n\mathbb{R}^n}$, then the cofactor of $A$ is denoted as $C_{ij}$, which is defined as:
     \begin{equation}
@@
 \begin{definition}[Determinant]
-    Let $A \in \mathbb{R}^{n \times n}$, then the determinant of $A$ is denoted as $\det(A)$, which is defined as:
+    Let $A \in  {^n\mathbb{R}^n}$, then the determinant of $A$ is denoted as $\det(A)$, which is defined as:
     \begin{equation}
@@
 \begin{theorem}[Determinant of a Transpose]
-    Let $A \in \mathbb{R}^{n \times n}$, then:
+    Let $A \in  {^n\mathbb{R}^n}$, then:
     \begin{equation}
@@
 \end{example}
+\begin{definition}[Singular Matrix]
+    Let $A \in  {^n\mathbb{R}^n}$, then $A$ is singular if $\det(A) = 0$.
+\end{definition}
+\begin{definition}[Adjugate Matrix]
+    Let $A \in  {^n\mathbb{R}^n}$, then the adjugate of $A$ is denoted as $\text{adj}(A)$, which is defined as:
+    \begin{equation}
+        \text{adj}(A) = \begin{bmatrix} C_{11} & -C_{12} & C_{13} & \ldots \\ -C_{21} & C_{22} & -C_{23} & \ldots \\ C_{31} & -C_{32} & C_{33} & \ldots \\ \vdots & \vdots & \vdots & \ddots \end{bmatrix}^T
+    \end{equation}
+\end{definition}
+\begin{theorem}[Determinant and Inverse]
+    Let $A \in  {^n\mathbb{R}^n}$, then the following properties hold:
+    \begin{enumerate}
+        \item $A$ is invertible if and only if $\det(A) \neq 0$.
+        \item $A^{-1} = \frac{1}{\det(A)}\text{adj}(A)$.
+    \end{enumerate}
+\end{theorem}
+\subsection{Determinants and Inverses of Elementary Matrices}
+\begin{definition}[Elementary Matrices]
+    Let $E \in  {^n\mathbb{R}^n}$, then $E$ is an elementary matrix if $E$ is obtained by performing a single elementary row operation on $I$.
+\end{definition}
+\begin{theorem}[Determinants and Inverses of Elementary Matrices]
+    Let $E \in  {^n\mathbb{R}^n}$ be an elementary matrix, then the following properties hold:
+    \begin{enumerate}
+        \item $\det(E) = 1$.
+        \item $E^{-1} = E^T$.
+    \end{enumerate}
+\end{theorem}
+\begin{example}[Determinants and Inverses of Elementary Matrices]
+    Let $E_1 = \begin{bmatrix} 1 & 0 \\ 0 & 2 \end{bmatrix}$, then:
+    \begin{align*}
+        \det(E_1) &= 1 \times 2 = 2 \\
+        E_1^{-1} &= E_1^T = \begin{bmatrix} 1 & 0 \\ 0 & \frac{1}{2} \end{bmatrix}
+    \end{align*}
+\end{example}
+\begin{theorem}[Determinants and Inverses of Product of Matrices]
+    Let $A, B \in  {^n\mathbb{R}^n}$, then the following properties hold:
+    \begin{enumerate}
+        \item $\det(AB) = \det(A)\det(B)$.
+        \item $(AB)^{-1} = B^{-1}A^{-1}$.
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Derminants and Inverse w.r.t. RREF]
+    Let $A \in  {^n\mathbb{R}^n}$, then the following properties hold:
+    \begin{enumerate}
+        \item $\det(A) = \det(\text{RREF}(A))$.
+        \item $A^{-1} = \text{RREF}(A)^{-1}$.
+    \end{enumerate}
+    However:
+    \begin{enumerate}
+        \item $\det(\text{RREF}(A)) = \det(A)$.
+        \item $\text{RREF}(A)^{-1} \neq A^{-1}$.
+    \end{enumerate}
+
+\end{theorem}
 \begin{theorem}[The theorem of Determinants and Inverse]
-    Let $A \in \mathbb{R}^{n \times n}$, then the following properties hold:
+    Let $A \in  {^n\mathbb{R}^n}$, then the following properties hold:
     \begin{enumerate}
@@
 \end{example}
+\paragraph{0 can be an Eigenvalue} If there exist $\textbf{x} \neq 0$ such that $A\textbf{x} = 0$, then 0 is an eigenvalue of $A$. This require the nullity of $A$ to be greater than 0.
 \subsection{Eigenspace and Multiplicity}
@@
 \end{theorem}
+\begin{theorem}[Theorem of Eigenvalues]
+    The following statements are equivalent:
+    \begin{enumerate}
+        \item $\lambda$ is an eigenvalue of $A$.
+        \item $Ax = \lambda x$ for some non-zero vector $x$.
+        \item $\det(A - \lambda I) = 0$.
+        \item $A - \lambda I$ is singular.
+        \item $A - \lambda I$ has a non-trivial null space.
+    \end{enumerate}
+
+\end{theorem}
 \begin{definition}[Algebraic Multiplicity (AM)]
@@
 \end{theorem}
+\begin{theorem}
+    Given $A\in {^n\mathbb{R}^n}$, we can find $P {^n\mathbb{R}^n}$ and $\Lambda \in {^n\mathbb{R}^n}$ where $\Lambda$ is a diagonal matrix so that $A = P\Lambda P^{-1}$ if and only if $A$ has $n$ linearly independent eigenvectors.
+\end{theorem}
+\begin{theorem}
+    $A \in {^n\mathbb{R}^n}$ is diagonalizable if and only if there is a basis for the eigenvectors of $A$ in $\mathbb{R}^n$.
+\end{theorem}
+
+\begin{theorem}
+    If $\Lambda$ is a diagonal matrix with the eigenvalues of $A$ on the diagonal, then $A$ is similar to $\Lambda$.
+\end{theorem}
+\begin{lemma}[Linear Independence of Eigenvectors]
+    Let $A \in {^n\mathbb{R}^n}$ and $v_1, v_2, \dots v_j$ be eigenvector assocated with distinct eigenvalues, then they are linearly independent.
+\end{lemma}
+\begin{theorem}
+    Let $A \in {^n\mathbb{R}^n}$ and $A$ has $n$ distict eigenvalues, then $A$ is diagonalizable.
+\end{theorem}
+\begin{theorem}[AM, GM, and Diagonalization]
+    Let $A \in {^n\mathbb{R}^n}$, then the following statements are equivalent:
+    \begin{enumerate}
+        \item $A$ is diagonalizable.
+        \item The algebraic multiplicity of each eigenvalue is equal to its geometric multiplicity.
+        \item The number of linearly independent eigenvectors of $A$ is equal to the dimension of the eigenspace of $A$.
+    \end{enumerate}
+\end{theorem}
+\paragraph{Similar Matrices} Two matrices $A$ and $B$ are similar if there exists an invertible matrix $P$ such that $A = PBP^{-1}$.
+This means that $A$ and $B$ have the same eigenvalues and the same characteristic polynomial.
+\begin{theorem}[Similar Matrices]
+    Let $A, B \in {^n\mathbb{R}^n}$, then the following statements are equivalent:
+    \begin{enumerate}
+        \item $A$ and $B$ are similar.
+        \item $A$ and $B$ have the same eigenvalues.
+        \item $A$ and $B$ have the same characteristic polynomial.
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Similarity of Charistic Polynomial]
+    Let $A, B \in {^n\mathbb{R}^n}$ be similar matrices, then:
+    \begin{enumerate}
+        \item $\lambda I - A$ and $\lambda I - B$ are similar matrices.
+        \item $\det(\lambda I - A) = \det(\lambda I - B)$ (by product of determinants).
+        \item $\dim \null (\lambda I - A) = \dim \null (\lambda I - B)$ (by rank-nullity), but not necessarily the same null space. 
+    \end{enumerate}
+    , thus the characteristic polynomial of $A$ and $B$ are the same and they have the same eigenvalues, which also means that they have the same AM.
+\end{theorem}
+\begin{theorem}
+    If $A$ is a diagonalizable matrix, then $A$ is has equal AM and GM.
+\end{theorem}
+\begin{proof}
+    $A$ is similar to the diagonal matrix $D$, so $\dim null D = \dim null A$. Thus, the AM and GM of $A$ are equal since AM and GM are equal for diagonal matrices.
+\end{proof}
+\begin{theorem}[Distinct Eigenvalues]
+    Let $A \in {^n\mathbb{R}^n}$, then the following statements are equivalent:
+    \begin{enumerate}
+        \item $A$ has $n$ distinct eigenvalues.
+        \item $A$ is diagonalizable.
+        \item The algebraic multiplicity of each eigenvalue is equal to its geometric multiplicity.
+        \item The number of linearly independent eigenvectors of $A$ is equal to the dimension of the eigenspace of $A$.
+    \end{enumerate}
+\end{theorem}
+\section{Differential Equations}
+\begin{definition}
+    A system of the form:
+    \begin{align*}
+        x_1^\prime &= a_{11}x_1 + a_{12}x_2 + \ldots + a_{1n}x_n \\
+        x_2^\prime &= a_{21}x_1 + a_{22}x_2 + \ldots + a_{2n}x_n \\
+        \vdots \\
+        x_n^\prime &= a_{n1}x_1 + a_{n2}x_2 + \ldots + a_{nn}x_n
+    \end{align*}
+    , wherer $a_{ij} \in \mathbb{R}$ is a $n \times n$ system of homogeneous linear differential equations with constant coefficients. For $A = \{a_{ij}\}$, we can write the system as:
+    \begin{equation}
+       X^\prime = AX
+    \end{equation}
+\end{definition}
+\begin{example}[Two Wafers in Breeze]
+    Let $\alpha = ha/mc$, $\beta = ka/mcd$, $T = [\hat{T}_1, \hat{T}_2]^T$, wher $h, a, m, c, k, d$ are constants. The system of equations is:
+    \begin{align*}
+        \frac{d\hat{T}_1}{dt} &= -(\alpha+\beta)\hat{T}_1 + \beta\hat{T}_2 \\
+        \frac{d\hat{T}_2}{dt} &= \beta\hat{T}_1 - (\alpha+\beta)\hat{T}_2
+    \end{align*}
+    Then the coefficients matrix is:
+    $$
+        A = \begin{bmatrix}
+            -(\alpha+\beta) & \beta \\
+            \beta & -(\alpha+\beta)
+        \end{bmatrix} = \begin{bmatrix} 
+            -(ha/mc + ka/mcd) & ka/mcd \\
+            ka/mcd & -(ha/mc + ka/mcd)
+        \end{bmatrix}
+    $$
+\end{example}
+\begin{example}
+    Let:
+    $$
+    A = \begin{bmatrix}
+        17 & -30 \\
+        10 & -18 
+    \end{bmatrix}
+    $$
+    , then particular solutions are
+    \begin{align*}
+        y_1 &= \begin{bmatrix}
+            2 \\ 1
+        \end{bmatrix}e^{2 t} \\
+        y_2 &= \begin{bmatrix}
+            3 \\ 2
+        \end{bmatrix}e^{3 t}
+    \end{align*}
+    Our initial conditions for two solutions are:
+    \begin{align*}
+        y_1(0) &= \begin{bmatrix}
+            2 \\ 1
+        \end{bmatrix} \\
+        y_2(0) &= \begin{bmatrix}
+            3 \\ 2
+        \end{bmatrix}
+    \end{align*}
+    , which are the eigenvectors of $A$ corresponding to the eigenvalues $2$ and $3$ respectively. Thus, the intial solution of the system:
+    \begin{align*}
+        y(0) &= c_1\begin{bmatrix}
+            2 \\ 1
+        \end{bmatrix} + c_2\begin{bmatrix}
+            3 \\ 2
+        \end{bmatrix} 
+    \end{align*}
+    is the span of the eigenvectors of $A$ corresponding to the eigenvalues $2$ and $3$ respectively.
+\end{example}
+\begin{theorem}
+    If $A$ is diagonalizable, then the system of DEs.
+    \begin{equation}
+        X^\prime = AX
+    \end{equation}
+    has a basis of solution and we can solve any IVP w.r.t. the system of DEs.
+\end{theorem}
+\subs
 \end{document}

### 2025-04-14 MAT185_Assignment_4/main.tex
exam

\documentclass[10pt,answers]{exam}

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{graphicx}
\usepackage[top=2.0cm,bottom=2.0cm,left=2.5cm,right=2.5cm]{geometry}
\usepackage{tikz}
\usetikzlibrary{decorations.pathmorphing,patterns}
\usetikzlibrary {arrows.meta} 
\usepackage{float}
\usepackage{multicol}
\usepackage{pgfplots}
\usepackage{lastpage}
\usepackage{soul}
\usepackage{libertine}
\usepackage{lipsum}
\usepackage{siunitx}
\usepackage{xspace}
\usepackage[labelfont=bf]{caption}
\usepackage[hidelinks, urlcolor=blue, linkcolor=blue, citecolor=blue, colorlinks=true]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage[absolute]{textpos}

% code below allows to put more spaces between rows in bmatrix
\makeatletter
\renewcommand*\env@matrix[1][\arraystretch]{%
  \edef\arraystretch{#1}%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{*\c@MaxMatrixCols c}}
\makeatother

%========================================================================
% These are commands needed for the Markov chain example
%
\usepackage{tikz-network} % Required for drawing network diagrams for Markov chains
\usepackage{nicematrix} % Allows highlighting rows and columns of matrices
\usepackage[customcolors]{hf-tikz}

\usepackage{colortbl}

\tikzset{style blue/.style={
    set fill color=tolBlue!30,
    set border color = tolBlue!30
  },
  style orange/.style={
    set fill color=tolOrange!40,
    set border color = tolOrange!40
  },
  style red/.style={
    ,set fill color=tolRed!40,
    set border color = tolRed!40
  },
  style teal/.style={
    set fill color=tolTeal!40,
    set border color = tolTeal!40
  },
  style purple/.style={
    set fill color=tolPurple!40,
    set border color = tolPurple!40
  },
  style grey/.style={
    set fill color=tolGrey!40,
    set border color = tolGrey!40
  },
  hor/.style={
    above left offset={-0.1,0.31},
    below right offset={0.1,-0.125},
    #1
  },
  horshort/.style={
    above left offset={0.1,0.31},
    below right offset={-0.1,-0.125},
    #1
  },
  ver/.style={
    above left offset={-0.1,0.3},
    below right offset={0.1,-0.15},
    #1
  }
}
%========================================================================

\newcommand{\R}{ \mathbb{R}}
\newcommand{\C}{ \mathbb{C}}
\newcommand{\bA}{ \mathbf{A}}
\newcommand{\bp}{ \mathbf{p}}
\newcommand{\bx}{ \mathbf{x}}
\newcommand{\bz}{ \mathbf{0}}
% this gives me an open circle
\newcommand{\choice}{\begin{tikzpicture}\draw (1,0) circle (1.0ex);\end{tikzpicture} \hspace{.1cm}}

% %correct choices in solution
\CorrectChoiceEmphasis{\rm}
\checkedchar{\tikz\draw[blue,fill=blue] (0,0) circle (1ex);}
\newcommand{\emptychoice}{\begin{tikzpicture}\draw (1,0) circle (1.0ex);\end{tikzpicture}~~}
\newcommand{\fullchoice}{\begin{tikzpicture}\draw[blue,fill=blue] (0,0) circle (1ex);\end{tikzpicture}~~}

\newcommand{\makeframednonemptybox}[3]{%
\setlength{\fboxrule}{0.5pt}
\fbox{%
\parbox[c][#1][t]{#2}{%
  \hrule width \hsize height 0pt%
  \vspace{\dimexpr(#1/2-0.1cm)}%
  \color{SolutionColor}#3\color{black}
 }%
}%
\setlength{\fboxrule}{1pt}
}

%% define course title
\newcommand{\course}{MAT185}
\newcommand{\assignmenttitle}{Assignment 4}

%% header and footer
\firstpageheader{}{}{\textbf{{\color{red} Due}: 10:00pm, Sunday Apr. 6, 2025}}
\firstpageheadrule
\runningheader{}{Page~\thepage~of~\numpages}{\course~--~\assignmenttitle}
\footer{}{}{}

\setlength\parindent{0pt} % no indentation in document

%% formats exam class
\qformat{\textbf{Question \thequestiontitle:}\hfill} % title of question 
\boxedpoints
\pointpoints{mark}{marks}
\pointsinrightmargin
\hpword{Marks:}
\hsword{Your score:}
\unframedsolutions
\totalformat{\boxed{\textnormal{\totalpoints~\if\totalpoints1 mark\else marks\fi}}}
\definecolor{SolutionColor}{rgb}{0,0,1}
\renewcommand{\solutiontitle}{}
\AtBeginEnvironment{solution}{\color{blue}}

% %correct choices in solution
\CorrectChoiceEmphasis{\rm}
\checkedchar{\tikz\draw[blue,fill=blue] (0,0) circle (1ex);}

% % increase distance between checkbox items
\renewcommand{\checkboxeshook}{\setlength{\itemsep}{6pt}}

%% distance between questions and parts
\renewcommand{\questionshook}{\setlength{\parsep}{10pt}}
\renewcommand{\partshook}{\setlength{\parsep}{15pt}}

%% define arrows in text
\newcommand{\arrow}{$\rightarrow$\xspace}
\newcommand{\Arrow}{$\Rightarrow$\xspace}

% % math notation:
%\veccol{1}{2}{3}{4}
\newcommand{\veccol}[4]{
    \begin{bmatrix}
      #1\\
      #2\\
      #3\\
      #4\\
    \end{bmatrix}}

\newcommand{\veccolTwo}[2]{
    \begin{bmatrix}
      #1\\
      #2\\
    \end{bmatrix}}

%%% This command makes a framed box of a chosen height.
\newcommand{\makenonemptybox}[2]{%
\par\nobreak\vspace{\ht\strutbox}\noindent
\setlength{\fboxrule}{0pt} % set this to 0pt to make invisible
\fbox{%
\parbox[c][#1][t]{\dimexpr\linewidth-2\fboxsep}{
  \hrule width \hsize height 0pt
  \vspace{-0.6cm}
  \color{SolutionColor}#2\color{black}
 }%
}%
\setlength{\fboxrule}{1pt}
}

%\vecrow{1}{2}{3}
\newcommand{\vecrow}[3]{\left[#1~#2~ #3\right]}

%\matrixTwo{1}{2}{3}{4}
\newcommand{\matrixTwo}[4]{\left[\begin{array}{cc}#1&#2\\#3&#4\end{array}\right]}

% \matrixThree{1}{2}{3}{4}{5}{6}{7}{8}{9}
\newcommand{\matrixThree}[9]{\left[\begin{array}{ccc}#1&#2&#3\\#4&#5&#6\\#7&#8&#9\end{array}\right]}

%\matrixCorner{1}{2}{3}{4}
\newcommand{\matrixCorner}[4]{\left[\begin{array}{ccc}#1& \cdots&#2\\ \vdots & \ddots & \vdots\\#3&
      \cdots&#4\end{array}\right]}

% \nR
\newcommand{\nR}{{}^{n}\mathbb{R}}
% \Rn
\newcommand{\Rn}{\mathbb{R}^{n}}
% \nRn
\newcommand{\nRn}{{}^{n}\mathbb{R}^{n}}
% \nRm
\newcommand{\nRm}{{}^{n}\mathbb{R}^{m}}
% \nRm
\newcommand{\mRn}{{}^{m}\mathbb{R}^{n}}
% \mRm
\newcommand{\mRm}{{}^{m}\mathbb{R}^{m}}

%% define abbreviations
\newcommand{\row}{\operatorname{row}\,}
\newcommand{\col}{\operatorname{col}\,}
\renewcommand{\dim}{\operatorname{dim}\,}
\renewcommand{\span}{\operatorname{span}\,}
\newcommand{\rank}{\operatorname{rank}\,}
\renewcommand{\ker}{\operatorname{ker}\,}
\newcommand{\nul}{\operatorname{null}\,}
\renewcommand{\det}{\operatorname{det}\,}
\newcommand{\adj}{\operatorname{adj}\,}

\begin{document}

\vspace*{-0.5cm}
\begin{center}
  \large
  \textbf{\Large \course~--~Linear Algebra}\\[0.1cm]
  \textbf{\assignmenttitle}
\end{center}
\bigskip

\textbf{\large Instructions:}\\
\normalsize

Please read the {\bf MAT185 Assignment Policies \& FAQ} document for
details on submission policies, collaboration rules and academic integrity, and
general instructions.

\begin{enumerate}

\item \textbf{Submissions are only accepted by}
  \href{https://www.gradescope.ca}{Gradescope}. Do not send anything by email.
  Late submissions are not accepted under any circumstance. Remember you can
  resubmit anytime before the deadline.

\item \textbf{Submit solutions using only this template pdf}.  Your submission
  should be a single pdf with your full written solutions for each question. If
  your solution is not written using this template pdf (scanned print or
  digital) then your submission will not be assessed. Organize your work neatly
  in the space provided.  Do not submit rough work.

\item \textbf{Show your work and justify your steps} on every question but do
  not include extraneous information.  Put your final answer in the box
  provided, if necessary.  We recommend you write draft solutions on separate
  pages and afterwards write your polished solutions here on this template.

\item \textbf{You must fill out and sign the academic integrity statement
    below}; otherwise, you will receive zero for this assignment.

\end{enumerate}

\vspace{10pt}

\textbf{\large Academic Integrity Statement:} \\

%%% Student information

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: Siu, Nelson\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010940608\enspace\hrulefill}\\
  \end{minipage}
}
\vspace*{0.1in}

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: Cheung, Hei Shing\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010907823\enspace\hrulefill}\\
  \end{minipage}
}
\bigskip
\large \textbf{I confirm that:}
\normalsize

\begin{itemize} 
\item I have read and followed the policies described in the document {\bf
    MAT185 Assignment Policies \& FAQ}.
\item In particular, I have read and understand the rules for
  collaboration, and permitted resources on assignments as described in
  subsection II of the the aforementioned document. I have not violated
  these rules while completing and writing this assignment.
\item I have not used generative AI in writing this assignment.
\item I understand the consequences of violating the University's academic
  integrity policies as outlined in the
  \href{http://www.governingcouncil.utoronto.ca/policies/behaveac.htm}{Code of
    Behaviour on Academic Matters}. I have not violated them while completing
  and writing this assignment.
\end{itemize}
\bigskip

\large \textbf{By submitting this assignment to Gradescope, I agree that the
  statements above are true.}  \normalsize

\newpage

\section*{Preamble}

This assignment explores two applications for eigenvectors and eigenvalues:
(i) the PageRank algorithm and (ii) solving linear constant-coefficient second-order ordinary differential
equations.

\vspace{.2cm}

In Questions~\ref{question:PageRank} and \ref{question:PageRank2}, you will
explore using eigenvectors to calculate the ranking of nodes in a directed
graph, which is useful for ranking websites by how important they are, e.g., to
show a sorted list of websites based on a search criterion. The main problem
with ranking websites is that not every website is connected to every other
website.\footnote{A similar problem would be if you consider a tournament
  (e.g. a chess tournament), in which not every player is playing against every
  other player. How can we now calculate a ranking of the best players if we
  only have insufficient data?}  We are introducing a technique which is the
basis for the PageRank-algorithm, which is the main algorithm used by Google to
sort its search results in the 90s and is still in use today. Bryan and
Leise~\cite{Bryan2006} describe this algorithm in more detail and it is a great
insight if you are interested in more technical details. Reading this
publication is not necessary to solve the problems in this assignment. However,
we have uploaded
\href{https://q.utoronto.ca/courses/375036/modules/items/6183944}{the
  publication to Quercus} for the curious.

\vspace{.2cm}

In Question~\ref{question:ODE} we are using eigenvalues and eigenvectors to
solve an ordinary differential equation (ODE). In that case, we are using a
mass-spring system, representing other mechanics ODEs you will solve in the
following terms.

\subsection*{The PageRank Algorithm}
The PageRank algorithm is an important part of the internal functionality of web
search engines. It was established by Google and its founders Sergey Brin and
Larry Page in the 1990s. Its main functionality is based on their publication in
Reference~\cite{Brin1998}. We present here a key idea behind the PageRank
algorithm, which involves eigenvectors, eigenvalues, and directed graphs.  You
will work through a small example.  In reality, one needs to apply efficient
numerical algorithms to estimate the PageRank eigenvector.

\vspace{.2cm}

As an example, the directed graph in \cref{fig:graph1} represents a small
network of four websites. This graph has 4 \textit{nodes}, each representing a
website. If a website links to another website, this is shown by a directed
edge.  For example, website \#1 has two links: one to website \#2 and one to
website \#3.  Website \#2 also has two links: one to website \#3 and one to
website \#4.  We will only consider graphs for which every website links to some
other website and no website links to itself. (Although we encourage the curious
to explore such situations once they've worked through Question 1 and 2.)

\begin{figure}[H]
  \begin{center}
    \resizebox{0.25\textwidth}{0.25\textwidth}{
      \begin{tikzpicture}[scale=0.75]
        \tikzstyle{every node}=[font=\small]
        \tikzset{vertex/.style = {shape=circle,draw,minimum size=0.5em}}
        \tikzset{edge/.style = {->,> = Stealth}}
        % vertices
        \node[vertex,fill={black!20}] (1) at  (0,1) {$1$};
        \node[vertex,fill={black!20}] (2) at  (4,0) {$2$};
        \node[vertex,fill={black!20}] (3) at  (0,-2) {$3$};
        \node[vertex,fill={black!20}] (4) at  (2,-3) {$4$};
        % edges
        \draw[edge] (1) to (2);
        \draw[edge] (2.185) to (3.50);
        \draw[edge] (1.260) to (3.100);
        \draw[edge] (3) to (2);
        \draw[edge] (2) to (4);
        \draw[edge] (3.80) to (1.280);
        \draw[edge] (3) to (4);
        \draw[edge] (4.37) to (2.255);
      \end{tikzpicture}}
  \end{center}
  \caption{Directed graph, representing a small network of linked websites.}
  \label{fig:graph1}
\end{figure}

One ranking strategy would be to use the number of incoming hyperlinks.  If lots
of other websites are linking to one website, surely that one website must be
very useful and reliable, right?  This strategy is flawed because it can be
easily manipulated by someone who creates lots of spurious websites all with
hyperlinks pointing to the one website they want to boost up in the ranking.
When creating a ranking, not all hyperlinks should carry the same weight!

\vspace{.2cm}

A hyperlink from a low-ranked website should have less of a contribution
to a website's ranking than a hyperlink from a high-ranked website.  Separately, there should
be some sort of dilution effect: if a website has two
hyperlinks pointing to other websites, those two links should count for more (on the receiving
ends) than if the website had ten hyperlinks pointing to other websites.

\newpage

The PageRank approach encapsulates these three ideas:
\begin{itemize}
\item The rank of a website is the sum of the weights of all incoming
  hyperlinks.
\item The higher the ranking of a website, the higher the weights are of its
  outgoing hyperlinks.
\item The more outgoing hyperlinks a website has, the lower the weight each of
  those hyperlinks has.
\end{itemize}
With this weighing strategy, we calculate the PageRank for each site as follows:
The PageRank $p_i$ of website $i$ is a positive scalar and is the weighted sum
of the PageRanks of all the websites that have hyperlinks pointing to it:
\begin{equation}
  \label{eq:1}
  p_i = \sum_{j\rightarrow i} \frac{p_j}{N_j}
\end{equation}
where $j \rightarrow i$ denotes there is a hyperlink from site $j$ to site $i$,
$p_j$ is the PageRank of website $j$, and $N_j$ is the total number of outgoing
hyperlinks from site $j$.  For example, to calculate the PageRank of website 2
in \cref{fig:graph1}, there is one incoming hyperlink from site 1 (which has 2
outgoing links), one from site 3 (which has 3 outgoing links), and one from site
4 (which has only one outgoing link). And so, the PageRank of $p_2$ is defined
as
\begin{equation}
  \label{eq:2}
  p_2 = \frac{1}{2} p_1 + \frac{1}{3} p_3 + p_4.
\end{equation}

In Question 1, you will find the PageRank of every website shown in
\cref{fig:graph1}.  In Question 2, you will continue studying the PageRank model
for this network and develop tools that will work for much, much larger
networks.

\vspace{.2cm}

\begin{thebibliography}{3}
\bibitem{Bryan2006} K. Bryan and T. Leise, \textit{The \$25,000,000,000
    Eigenvector: The Linear Algebra behind Google}, SIAM Review, 48:3, 569-581,
  2006,
  \href{https://q.utoronto.ca/courses/375036/modules/items/6183944}{[online]}.
\bibitem{Brin1998} S. Brin and L. Page, \textit{The anatomy of a large-scale
    hypertextual Web search engine}, Computer Networks and ISDN Systems, 30:1-7,
  107-117, 1998.
\end{thebibliography}

\newpage

\begin{questions}
  \question \label{question:PageRank}
  \begin{parts}
    \part \Cref{eq:2} determines the PageRank of node~2. Give the system of
    PageRank equations for the four nodes in \cref{fig:graph1}, as defined by
    \cref{eq:1}. \textit{No justification is required.}

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{12em}{
      %%% Add your explanations here!      
      $$\begin{cases}
        p_1 &= \frac{1}{3}p_3 \\
        p_2 &= \frac{1}{2}p_1 + \frac{1}{3}p_3 + p_4 \\
        p_3 &= \frac{1}{2}p_1 + \frac{1}{2}p_2 \\
        p_4 &= \frac{1}{2}p_2 + \frac{1}{3}p_3 
      \end{cases}$$
      In matrix form, this is:
        $$\begin{bmatrix}
            0 & 0 & 1/3 & 0 \\
            1/2 & 0 & 1/3 & 1 \\
            1/2 & 1/2 & 0 & 0 \\
            0 & 1/2 & 1/3 & 0
            \end{bmatrix}
            \begin{bmatrix}
              p_1 \\ p_2 \\ p_3 \\ p_4
            \end{bmatrix}
            =
            \begin{bmatrix}
                p_1 \\ p_2 \\ p_3 \\ p_4
            \end{bmatrix}$$
      %%% end of your answer    
    }

    \part Formulate the linear system you found in part (a) as an
    eigenproblem. Identify the matrix $\mathbf{A}$, the eigenvalue and the
    eigenvector.  \label{part:q1_eigen}
        \textit{Note: you'll solve for the eigenvectors in the next part of this question.}

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{15em}{
      %%% Add your explanations here!      
        For some eigenvector of $A$: $\mathbf{p} = \begin{bmatrix}
          p_1 & p_2 & p_3 & p_4
        \end{bmatrix}^T$, we can write the system of equations as
        $$A\mathbf{p} = \mathbf{p} = \begin{bmatrix}
            0 & 0 & 1/3 & 0 \\
            1/2 & 0 & 1/3 & 1 \\
            1/2 & 1/2 & 0 & 0 \\
            0 & 1/2 & 1/3 & 0
            \end{bmatrix}
            \begin{bmatrix}
              p_1 \\ p_2 \\ p_3 \\ p_4
            \end{bmatrix}
            =
            \lambda
            \begin{bmatrix}
              p_1 \\ p_2 \\ p_3 \\ p_4
            \end{bmatrix}$$
        where $\lambda = 1$. The matrix $\mathbf{A}$ is
        $$\mathbf{A} = \begin{bmatrix}
            0 & 0 & 1/3 & 0 \\
            1/2 & 0 & 1/3 & 1 \\
            1/2 & 1/2 & 0 & 0 \\
            0 & 1/2 & 1/3 & 0
            \end{bmatrix}$$
      %%% end of your answer    
    }

    \part \label{part:solve_eigen} Solve the eigenproblem you formulated in part
    (b).  Calculate the PageRank vector $\mathbf{p}$ for the graph in
    \cref{fig:graph1}.  You will need to use that the components of a PageRank
    vector must add up to 1.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{18em}{
      %%% Add your explanations here!
    Identifying $A\mathbf{p} = \mathbf{p}$ as an eigenvalue problem, we can
    solve for the eigenvector $\mathbf{p}$ by solving the system of equations $A\textbf{p} - \textbf{p} = 0$.
    This is equivalent to solving the system of equations $(A - I)\textbf{p} = 0$. So, we have:
    $$\begin{bmatrix}
        -1 & 0 & 1/3 & 0 \\
        1/2 & -1 & 1/3 & 1 \\
        1/2 & 1/2 & -1 & 0 \\
        0 & 1/2 & 1/3 & -1
        \end{bmatrix}
        \begin{bmatrix}
          p_1 \\ p_2 \\ p_3 \\ p_4
        \end{bmatrix}
        =
        \begin{bmatrix}
          0 \\ 0 \\ 0 \\ 0
        \end{bmatrix}$$
    We can solve this system of equations using computers method. I used the scipy.nullspace function in Python to find the null space of the matrix. The result is:
    $$\mathbf{p} = \text{span}\left(\begin{bmatrix}
        0.14547859 & 0.72739297 & 0.43643578 & 0.50917508       \end{bmatrix}^T\right)$$
    We can normalize this vector by dividing each component by the sum of all components:
    $$\sum_{i=1}^4 p_i = 0.14547859 + 0.72739297 + 0.43643578 + 0.50917508 = 1.81848242$$
    So, the PageRank vector is: 
    $$\mathbf{p} = \begin{bmatrix}
        0.14547859 & 0.72739297 & 0.43643578 & 0.50917508
      \end{bmatrix}^T/1.81848242 = \begin{bmatrix}
        0.08 & 
        0.4 & 
        0.24 & 
        0.28
        \end{bmatrix}^T$$
    }

    \newpage

    \part Which website of the graph in \cref{fig:graph1} has the highest and
    which has the lowest PageRank?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{10em}{
      %%% Add your explanations here!
      \begin{itemize}
        \item Website 2 has the highest PageRank of 0.4
        \item Website 1 has the lowest PageRank of 0.08
      \end{itemize}
      %%% end of your answer    
    }

    \part   
    Think about other possible networks.  What would be a network that would
    lead to all four webpages having the same PageRank?  Add the hyperlinks to
    the figure below on the left.  What would be a \textit{different} network
    would lead to all four webpages having the same PageRank? Add the hyperlinks
    to the figure on the right. \textit{No justification is required.}

    \textit{There are more than two possible answers to this question.  Please
      make sure that your hyperlinks are clearly legible in your scan, including
      the direction they're pointing.}

    \vspace{1cm}

    \begin{minipage}{.5\textwidth}
      \begin{center}
        \vspace{.3cm}
        \begin{tikzpicture}[scale=.55]
          \Vertex[size = .5]{4} \Vertex[x=4,y=0, size = .5]{3}
          \Vertex[x=0,y=4,size = .5]{1} \Vertex[x=4,y=4, size = .5]{2}

          \draw (4) node {4};
          \draw (3) node {3};
          \draw (1) node {1};
          \draw (2) node {2};
           % % YOU CAN DRAW ARROWS WITH THE FOLLOWING COMMANDS
          \Edge[Direct,bend=12](1)(2)
            \Edge[Direct,bend=12](2)(3)
            \Edge[Direct,bend=12](3)(4)
            \Edge[Direct,bend=12](4)(1)

          % \Edge[Direct,loopshape=45,loopposition=0](2)(3)                        
        \end{tikzpicture}
        \end{center}
    \end{minipage}    
    %
 \begin{minipage}{.5\textwidth}
      \begin{center}
        \vspace{.3cm}
        \begin{tikzpicture}[scale=.55]
          \Vertex[size = .5]{4} \Vertex[x=4,y=0, size = .5]{3}
          \Vertex[x=0,y=4,size = .5]{1} \Vertex[x=4,y=4, size = .5]{2}

          \draw (4) node {4};
          \draw (3) node {3};
          \draw (1) node {1};
          \draw (2) node {2};
           % % YOU CAN DRAW ARROWS WITH THE FOLLOWING COMMANDS
          % \Edge[Direct,bend=12](1)(2)
          \Edge[Direct,bend=12](1)(3)
            \Edge[Direct,bend=12](2)(4)
            \Edge[Direct,bend=12](3)(1)
            \Edge[Direct,bend=12](4)(2)
          % \Edge[Direct,loopshape=45,loopposition=0](2)(3)                        
        \end{tikzpicture}
        \end{center}
    \end{minipage}

    \vspace{1.5cm}

  \end{parts}

\newpage

\question \label{question:PageRank2} Now that you are more familiar with the
eigenproblem from Question 1, you're going to study the eigenvalues and
eigenvectors of the matrix $\bA$ from Question 1.

\begin{parts}

  \part Compute $\bA \bx$ where $\bx \in {^4}\R$.  How is
  $\displaystyle\sum_{i=1}^4 (\bA \bx)_i$ related to
  $\displaystyle\sum_{i=1}^4 x_i$?\\[-1.0cm]

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{11em}{
    \vspace{0.7cm}
    %%% Add your explanations here!
    %Since $A\mathbf{x} = \lambda \mathbf{x}$, we have $\sum_{i=1}^4 (\bA \bx)_i = \sum_{i=1}^4 \lambda x_i$. Which means:
    %$$\sum_{i=1}^4 (\bA \bx)_i = \sum_{i=1}^4 x_i$$
    %, since $\lambda = 1$.
    We consider the sum of the components of $\bA \bx$. First, we compute $\bA \bx$:
    $$\bA \bx = \begin{bmatrix}
        0 & 0 & 1/3 & 0 \\
        1/2 & 0 & 1/3 & 1 \\
        1/2 & 1/2 & 0 & 0 \\
        0 & 1/2 & 1/3 & 0
      \end{bmatrix}
      \begin{bmatrix}
        x_1 \\ x_2 \\ x_3 \\ x_4
      \end{bmatrix} =
      \begin{bmatrix}
        (1/3)x_3 \\ (1/2)x_1 + (1/3)x_3 + x_4 \\ (1/2)x_1 + (1/2)x_2 \\ (1/2)x_2 + (1/3)x_3
      \end{bmatrix}$$
    $$\sum_{i=1}^4 (\bA \bx)_i = (1/3)x_3 + (1/2)x_1 + (1/3)x_3 + x_4 + (1/2)x_1 + (1/2)x_2 + (1/2)x_2 + (1/3)x_3 = x_1 + x_2 + x_3 + x_4$$
      Thus, we have  $\sum_{i=1}^4 (\bA \bx)_i = \sum_{i=1}^4 x_i$

    %%% end of your answer    
  }

  Assume $\bx$ is an eigenvector of $\bA$ with eigenvalue $\lambda$.  What can
  you say about $\lambda$ and/or $\bx$?

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{8em}{
    %%% Add your explanations here!
    Since $\bA \bx = \lambda \bx$, we can write: $\sum_{i=1}^4 (\bA \bx)_i = \sum_{i=1}^4 \lambda x_i$. 
    We know from the previous part that $\sum_{i=1}^4 (\bA \bx)_i = \sum_{i=1}^4 x_i$. So, we have $\sum_{i=1}^4 \lambda x_i = \sum_{i=1}^4 x_i$. 
    So, one of the following must be true:
    \begin{itemize}
      \item $\lambda = 1$ and $\bx$ is an eigenvector of $\bA$.
      \item $\sum_{i=1}^4 x_i = 0$ and $\lambda$ can be any value.
    \end{itemize}
    So, we can conclude that if $\bx$ is an eigenvector of $\bA$, then either
    $\lambda = 1$ or $\sum_{i=1}^4 x_i = 0$.
    %%% end of your answer    
  }

  \part Using your computation of $\bA \bx$ where $\bx \in {^4}\R$, how is
  $\displaystyle\sum_{i=1}^4 | (\bA \bx)_i |$ related to
  $\displaystyle\sum_{i=1}^4 |x_i|$?  \textit{You will want to use the triangle
    inequality: if $a, b \in \mathbb{R}$ then $|a+b| \leq |a| + |b|$.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{11em}{
    %%% Add your explanations here!
    For all not necessarily eigenvectors $\bx \in {^4}\R$, we have:
    $
    (Ax)_i = \sum_{j=1}^4 A_{ij} x_j
    $
    Using the triangle inequality, we have:
    $$
    |(Ax)_i| = |\sum_{j=1}^4 A_{ij} x_j| \leq \sum_{j=1}^4 |A_{ij}| |x_j|
    $$
     \\Taking the sum over $i$ gives:
    $
    \sum_{i=1}^4 |(Ax)_i| \leq \sum_{i=1}^4 \sum_{j=1}^4 |A_{ij}| |x_j| = \sum_{j=1}^4 |x_j| \sum_{i=1}^4 |A_{ij}|
    $
    \\ Since $\sum_{i=1}^4 |A_{ij}| = 1$ for all $j$, we have:
    $$
    \sum_{i=1}^4 |(Ax)_i| \leq \sum_{j=1}^4 |x_j|
    $$

    %%% end of your answer    
  }

  If $\bx$ is an eigenvector of $\bA$, what can you say about its eigenvalue $\lambda$?

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{8em}{
    %%% Add your explanations here!
    From the previous part, we know that $\sum_{i=1}^4 |(Ax)_i| \leq \sum_{j=1}^4 |x_j|$.
    Since $\bA \bx = \lambda \bx$, we have:
    $$
    \sum_{i=1}^4 |(Ax)_i| = |\lambda| \sum_{i=1}^4 |x_i| \leq \sum_{i=1}^4 |x_i|
    $$
    If $\bx$ is an eigenvector of $\bA$, be defintion we have $\sum_{i=1}^4 |x_i| \neq 0$. So, we can divide both sides by $\sum_{i=1}^4 |x_i|$:
    $$
    |\lambda| \leq 1
    $$

    %%% end of your answer    
  }

  \newpage

  \part Combining your answer to (a) and (b) what can you say about $\bA$'s
  eigenvalues and eigenvectors?

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{3em}{
    %%% Add your explanations here!
    We have the following cases:
  \begin{itemize}
      \item If $\sum_{i=1}^4 x_i \neq 0$, then $|\lambda| = 1$.
      \item If $\sum_{i=1}^4 x_i = 0$, then $|\lambda| < 1$.
  \end{itemize}
  So, we can conclude that the eigenvalues of $\bA$ are all less than or equal to 1.
  %%% end of your answer    
  }

  \part Using matlab or python or wolframalpha or whatever, what are the
  eigenvalues of $\bA$?  Please give the eigenvalues in order of decreasing
  magnitude, largest to smallest.  Remember, $| a + i b | = \sqrt{a^2 + b^2}$.
  Round to two decimal places. \textit{No justification is required.}

  \begin{minipage}{.2\textwidth}
    $\lambda_1 \approx  $\makeframednonemptybox{0.5cm}{2cm}{1.00
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.2\textwidth}
    $\lambda_2 \approx  $\makeframednonemptybox{0.5cm}{2cm}{-0.6 + 0.22i
    }
    [$|\lambda_2| \approx 0.64$]
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.2\textwidth}
    $\lambda_3 \approx  $\makeframednonemptybox{0.5cm}{2cm}{-0.6 - 0.22i
    }
    [$|\lambda_3| \approx 0.64$]
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.2\textwidth}
    $\lambda_4 \approx  $\makeframednonemptybox{0.5cm}{2cm}{0.20
    }
  \end{minipage}

  %% REPLACE \choice with \fullchoice FOR YOUR ANSWER
  Is $\bA \in {^4}\R^4$ diagonalizable? \hspace{.2cm}  \choice Yes \hspace{.4cm}  \fullchoice No

  %% REPLACE \choice with \fullchoice FOR YOUR ANSWER
  Is $\bA \in {^4}\C^4$ diagonalizable? \hspace{.2cm}  \fullchoice Yes \hspace{.4cm}  \choice No

  \part Let $\bx \in {^4}\C$ be a linear combination of the eigenvectors of $\bA$;
  that is, $\bx = c_1 \bx_1 + c_2 \bx_2 + c_3 \bx_3 + c_4 \bx_4$ where
  $c_i \in \C$.  What is $\bA \bx$?  What is $\bA^2 \,\bx$?  What is
  $\bA^{100} \,\bx$?  \textit{Your answers should include the symbols $\lambda_i$
    when referring to eigenvalues, rather than the approximate values you found
    for eigenvalues.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{7em}{
    %%% Add your explanations here!
    Since $x_i$ are eigenvectors of $A$, we have:
    $$\bA \bx = c_1 \lambda_1 \bx_1 + c_2 \lambda_2 \bx_2 + c_3 \lambda_3 \bx_3 + c_4 \lambda_4 \bx_4$$
    Since $\bA$ is a linear operator, $\bA(c_1 \bx_1 + c_2 \bx_2 + c_3 \bx_3 + c_4 \bx_4) = c_1 \bA \bx_1 + c_2 \bA \bx_2 + c_3 \bA \bx_3 + c_4 \bA \bx_4$ and we can write:
    $$\bA^2 \bx = c_1 \lambda_1^2 \bx_1 + c_2 \lambda_2^2 \bx_2 + c_3 \lambda_3^2 \bx_3 + c_4 \lambda_4^2 \bx_4$$
    $$\bA^{100} \bx = c_1 \lambda_1^{100} \bx_1 + c_2 \lambda_2^{100} \bx_2 + c_3 \lambda_3^{100} \bx_3 + c_4 \lambda_4^{100} \bx_4$$
    %%% end of your answer    
  }

  As $n \to \infty$, what does $\bA^n \, \bx$ do, and why?

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{3em}{
    %%% Add your explanations here!
    Since $|\lambda_1| = 1$ and $|\lambda_2|, |\lambda_3|, |\lambda_4| < 1$, we have:
    $$\bA^n \bx = c_1 \lambda_1^n \bx_1 + c_2 \lambda_2^n \bx_2 + c_3 \lambda_3^n \bx_3 + c_4 \lambda_4^n \bx_4$$
    As $n \to \infty$, the terms with $\lambda_2$, $\lambda_3$ and $\lambda_4$ will go to 0, and we will be left with $\bA^n \bx \to c_1 \bx_1$.
    %%% end of your answer    
  }

  \part You have the matrix $\bA$ and you would like to find an approximation of
  an eigenvector for the eigenvalue $1$.  You have access to matlab and python but
  not to subroutines that compute eigenvalues and eigenvectors.
  Playing around, you discover that if you choose a random vector $\bx \in {^4}\R$
  and you compute $\bA^n \, \bx$ for larger and larger values of $n$ then
  $\bA^n \, \bx$ seems to stop changing (so far as the computer can see).  If you
  call that computational limit $\bx_\infty$, is $\bx_\infty$ (approximately) an
  eigenvector for $\bA$ with eigenvalue $1$?

  %% REPLACE \choice with \fullchoice FOR YOUR ANSWER
  \hspace{.2cm}  \choice Yes \hspace{.4cm}  \fullchoice No

  If yes, why?  If no, why not?
  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{4em}{
    %%% Add your explanations here!
    No, since $x_\infty$ is the image of the random vector $\bx$ under the
    transformation $\bA^n$, it is not guaranteed that $\bx$ not be spanned only by $\bx_2, \bx_3$ and $\bx_4$. 
    If $c_1 = 0$, then $\bA^n \bx_\infty$ will not converge to an eigenvector of $\bA$ and converge to the zero vector.
    For instance, we pick $\bx = \bx_3$, then $\bA^n \bx_\infty = c_3 \lambda_3^n \bx_3$ and as $n \to \infty$, $\bA^n \bx_\infty$ will converge to 0.
    So, we can conclude that $\bx_\infty$ is not guaranteed to converge to the eigenvector of $\bA$ assocated with the eigenvalue 1.
    %%% end of your answer    
  }

\end{parts}

\begin{EnvUplevel}  
  \textit{\textbf{For the curious:}} The linear system in question 1(a) is a
  small linear system: four equations in four unknowns.  Your immediate reaction
  upon seeing the system should have been to set up an augmented matrix, do some
  elementary row operations, and find all solutions.  Finding solutions by
  viewing it as an eigenproblem seems pretty artificial!  And even if you do
  decide to take the eigenproblem view of things; you can find the null space of
  $ \mathbf{I} - A$ via row reduction.  So row reduction must be the right
  approach, right?

  For a internet search engine, how big is the matrix $\mathbf{A}$?  Its size
  depends on the number of considered webpages, which is roughly on the order of
  tens of billions for a search engine.  We do know that the $\textbf{A}$ is
  very sparse because, on average, there are fewer than 100 links per webpage
  pointing to different page. So, $\bA$ is a sparse matrix of size 10 billion
  $\times$ 10 billion. The number of operations needed to transform an
  $n \times n$ matrix to reduced row echelon form is $\mathcal{O}(n^3)$.  But
  the number of operations needed to compute $\bA \bx$ once is
  $\mathcal{O}(n^2)$.  When you have $\bA \bx$, computing $\bA^2 \bx$ is a
  matter of computing $\bA (\bA \bx)$ --- this will be another
  $\mathcal{O}(n^2)$.  Computing $\bA^k \, \bx$ is $\mathcal{O}(k n^2)$ and so,
  if $k$ isn't large, this will be faster than computing the reduced row echelon
  form.  Using this approach to approximate an eigenvector for the largest
  eigenvalue is called \textit{power method}.  It works well as long as the
  geometric multiplicity of the largest eigenvalue is 1.  The speed with which
  it converges has to do with how big the gap is between $|\lambda_1|$ and
  $|\lambda_2|$ where $\lambda_2$ is an eigenvalue with ``next largest" size.
  The bigger $|\lambda_1| - |\lambda_2|$ is, the faster the power method
  converges.  You can find Python implementations of the power method to
  calculate the dominant eigenvalue and eigenvector
  \href{https://www.geeksforgeeks.org/power-method-determine-largest-eigenvalue-and-eigenvector-in-python/}{online}.

  Numerical methods for computing eigenvalues have been an area of active
  research and continue to be so.  The power method is a snowflake on top of an
  iceberg when it comes to the topic!

  A separate question is: what properties of a network ensure that $1$ will be
  an eigenvalue of the matrix $\bA$?  What properties ensure that its eigenspace
  will be one dimensional?

\end{EnvUplevel}

\newpage

\question \label{question:ODE} The typical ODEs in mechanics are second-order
ODEs, because the acceleration is the second derivative of the position in
respect to time. If the ODEs are linear, they can be reformulated as two linear
first-order ODEs.  If the ODEs have constant coefficients and if the forcing is
constant, we can solve the system using the methods from this course.

In this problem, you will will explore solving a second-order ODE which
describes a mass-spring system using the techniques discussed in this course.
\textit{For the curious: the methods used in this problem work equally well if
  there is damping in the system and these methods apply for other oscillatory
  systems such as RL circuits and RLC circuits.}

Consider the following mass-spring system with one mass and two springs:
\bigskip

\begin{figure}[H]
  \begin{center}
    \begin{tikzpicture}
      \draw[decoration={aspect=0.3, pre length=0.5cm,post length=1.0cm, segment
        length=2.5mm, amplitude=3mm,coil},decorate] (-5,0) -- (0,0);
      \draw[decoration={aspect=0.3, pre length=1.0cm,post length=0.5cm, segment
        length=2.5mm,amplitude=3mm,coil},decorate] (0,0) -- (5,0);
      \node[rectangle,fill=gray,inner sep=6mm] (a) at (-0.1,0) {};
      \fill [pattern = north east lines] (-5.5,-0.6) rectangle (-5,0.6);
      \draw[very thick]
      (-5,-0.6) -- (-5,0.6);
      \fill [pattern = north east lines] (5.5,-0.6) rectangle (5,0.6); \draw[very thick]
      (5,-0.6) -- (5,0.6);
      \fill [pattern = north east lines] (-5.5,-0.6) rectangle (5.5,-1.0);
      \draw[very thick] (-5,-0.6) -- (5,-0.6);
      \node[] at (-0.1,0.9) {\Large $m$};
      \node[] at (-2.8,0.7) {\Large $k_1$};
      \node[] at (2.8,0.7) {\Large $k_2$};
      \draw[-{Stealth[length=3mm]},thick] (-0.1,-1.5) -- (0.6,-1.5);
      \draw[thick] (-0.1,-1.75) -- (-0.1,-1.25);
      \node[] at (0.25,-2.0) {\Large $q$};
    \end{tikzpicture}
  \end{center}
  \caption{Mass-spring system with the mass $m$ and two springs of
    stiffness $k_1$ and $k_2$.}
  \label{fig:mass_spring}
\end{figure}
We will ignore the effect of gravity and of friction and we will assume the
displacements are small enough that Hooke's law applies.  Let $m$ be the mass,
and $k_1$ and $k_2$ be the spring constants (i.e. the stiffnesses) of the
springs. The equation of motion for the system in \cref{fig:mass_spring} is
\begin{equation}
  \label{eq:ode}
  m \ddot{q}(t) = -k_1 q(t) - k_2 q(t)  = -(k_1 + k_2)q(t).
\end{equation}
Here, $q$ is the displacement from rest; if the system is at rest then
$q(t)=0$ for all time $t$.  To understand the force: a displacement of the
mass $m$ in positive $q$ direction result in the left spring to lengthen,
which leads to pulling in $-q$ direction, and the right spring to compress,
which leads to pushing in $-q$ direction.

  \begin{parts}

    \part Using the substitution of $x_1(t) = q(t)$ and $x_2(t) = \dot q(t)$, show
 that  
    \cref{eq:ode}  can
    be expressed as the following system of two first-order ODEs:
    \begin{equation}
      \label{eq:subst}
      \dot{\mathbf{x}}(t)=\veccolTwo{\dot{x}_1(t)}{\dot{x}_2(t)}=\matrixTwo{0}{1}{-\frac{k_1+k_2}{m}}{0}\veccolTwo{x_1(t)}{x_2(t)}=\mathbf{A}\mathbf{x}(t).
    \end{equation}

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{10em}{
      %%% Add your explanations here!
      From \cref{eq:ode}, we have:
       $$m \dot{x}(t) = m \ddot{q}(t) = -k_1 q(t) - k_2 q(t)  = -(k_1 + k_2)q(t) = -k q(t)$$
      So, we have:
        $$\dot{x}(t) = -\frac{k_1 + k_2}{m} x(t) \cdot (1)$$
      Also, we have the definition of $x_1$ and $x_2$: $x_1(t) = q(t)$ and $x_2(t) = \dot{q}(t)$, so we can write:
    $$\dot{x}_1(t) = \dot{q}(t) = x_2(t)$$ 
    Constructing the system of equations, we have:
    $$\begin{bmatrix}
        0 & 1 \\
        -\frac{k_1 + k_2}{m} & 0
        \end{bmatrix} \begin{bmatrix}
        x_1(t) \\
        x_2(t)
        \end{bmatrix} = \begin{bmatrix}
        \dot{x}_1(t) \\
        \dot{x}_2(t)
        \end{bmatrix}$$

       %%% end of your answer    
    }
  \newpage

  \begin{EnvUplevel}
    You will now solve \cref{eq:ode}, using \cref{eq:subst} and
    the methods introduced in this course.
  \end{EnvUplevel}

  \part \label{part:eigenvalue} View $\bA$ as a complex matrix, rather than a
  real one, $\bA \in {^2}\C^2$, and compute the eigenvalues of $\mathbf{A}$ from
  \cref{eq:subst}.  To keep your computations simple, replace $k_1+k_2$ with $k$
  (the effective spring constant).  Your eigenvalues should be of the form
  $\pm i \omega$ where $\omega$ is a function of $k$ and $m$.  The expression you will find for $\omega$ is called the
  ``natural angular frequency'' or the ``natural frequency''.  For the rest of
  Question 3, refer to $\omega$ rather than to expressions involving $k$ and
  $m$.

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{25em}{
    %%% Add your explanations here!
    The eigenvalues of $A$ are the solutions of the characteristic polynomial:
    $$\det(A - \lambda I) = 0$$
    We have:
    $$A - \lambda I = \begin{bmatrix}
        0 & 1 \\
        -\frac{k}{m} & 0
      \end{bmatrix} - \begin{bmatrix}
        \lambda & 0 \\
        0 & \lambda
      \end{bmatrix} = \begin{bmatrix}
        -\lambda & 1 \\
        -\frac{k}{m} & -\lambda
      \end{bmatrix}$$
    So, we have:
    $$\det(A - \lambda I) = \det \begin{bmatrix}
        -\lambda & 1 \\
        -\frac{k}{m} & -\lambda
      \end{bmatrix} = (-\lambda)(-\lambda) - (1)(-\frac{k}{m}) = \lambda^2 + \frac{k_1 + k_2}{m}$$
    Setting the determinant to 0, we have:
    $$\lambda^2 + \frac{k}{m} = 0$$
    $$\lambda^2 = -\frac{k}{m}$$
    $$\lambda = \pm i \sqrt{\frac{k}{m}}$$
    Setting $\omega = \sqrt{\frac{k}{m}}$, we have:
    $$\lambda = \pm i \omega$$
    %%% end of your answer    
  }

  \part \label{part:e_vec} Compute the eigenvectors for the two complex
  eigenvalues calculated in
  Question~\ref{question:ODE}\,(\ref{part:eigenvalue}). \textit{\textbf{Hint:}}
  The eigenvectors will be complex as well.  \textit{You will need to rewrite
    $\bA$ so that it no longer involves $k$ and $m$, only $\omega$.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{20em}{
    %%% Add your explanations here!
    We consider the matrix:
    $$A \pm i \omega I = \begin{bmatrix}
        0 & 1 \\
        -\omega^2 & 0
      \end{bmatrix} \pm i \begin{bmatrix}
        \omega & 0 \\
        0 & \omega
        \end{bmatrix} = \begin{bmatrix}
        \pm i \omega & 1 \\
        -\omega^2 & \pm i \omega
      \end{bmatrix}$$
    %%% end of your answer  
    Considering the first eigenvalue $\lambda_1 = i \omega$, and consider any $c_1, c_2 \in \C$, we have:
    $$\begin{bmatrix}
        i \omega & 1 \\
        -\omega^2 & i \omega
      \end{bmatrix} \begin{bmatrix}
        x_1 \\
        x_2
      \end{bmatrix} = 0$$
    So the first eigenvector is $c_2[1, i \omega]^T$. Considering the second eigenvalue $\lambda_2 = -i \omega$, we have:
    $$\begin{bmatrix}
        -i \omega & 1 \\
        -\omega^2 & -i \omega
      \end{bmatrix} \begin{bmatrix}
        x_1 \\
        x_2
      \end{bmatrix} = 0$$
    So the second eigenvector is $c_2[1, -i \omega]^T$.  So the eigenvectors are:
    $$c_1\begin{bmatrix}
        1 \\
        i \omega
      \end{bmatrix} \text{ and } c_2\begin{bmatrix}
        1 \\
        -i \omega
      \end{bmatrix}$$
  }

  \newpage

  \part Using the eigenvalues and eigenvectors you found, what is the general
  solution, $\bx(t)$, for the system in \cref{eq:subst}?  What is each component, $x_1(t)$ and
  $x_2(t)$?  \textit{Please keep working with $\omega$ rather than $k$ and $m$.
    Also, you want to check your work by checking on a piece of scratch paper that $\dot{x}_1(t) = x_2(t)$.  There's no need to present that verification here.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{6em}{
       Given $\lambda_1 = i \omega$ and $\lambda_2 = -i \omega$, we have:
    $$\begin{bmatrix}
        x_1(t) \\
        x_2(t)
      \end{bmatrix} = c_1 e^{i \omega t} \begin{bmatrix}
        1 \\
        i \omega
        \end{bmatrix} + c_2 e^{-i \omega t} \begin{bmatrix}
        1 \\
        -i \omega
        \end{bmatrix}$$
    So, we have:
    $$x_1(t) = c_1 e^{i \omega t} + c_2 e^{-i \omega t}$$
    $$x_2(t) = c_1 i \omega e^{i \omega t} - c_2 i \omega e^{-i \omega t} = i \omega (c_1 e^{i \omega t} - c_2 e^{-i \omega t})$$

  }

  \part Assume that the mass is initially at rest, $q(0)=q_0$ and
  $\dot{q}(0) = 0$.  This determines $\bx(0)$.  What is $\bx(t)$?
 \textit{At this
    point, we haven't discussed what it means to have a complex number in the
    exponent but you do still know that $e^0=1$.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{8em}{
    %%% Add your explanations here!
    First, we solve for $c_1$ and $c_2$ using the given initial conditions:
    $$x_1(0) = c_1 = c_2 = q_0 \quad x_2(0) = i \omega (c_1 - c_2) = \dot{q}(0) = 0$$
    Thus, by sovling the system of equations, we have $c_1 + c_2 = q_0/2$. Therefore, we have:
    $$\bx = \begin{bmatrix}
        x_1(t) \\
        x_2(t)
      \end{bmatrix} = \begin{bmatrix}
        c_1 e^{i \omega t} + c_2 e^{-i \omega t} \\
        i \omega (c_1 e^{i \omega t} - c_2 e^{-i \omega t})
      \end{bmatrix} = \begin{bmatrix}
        \frac{q_0}{2} e^{i \omega t} + \frac{q_0}{2} e^{-i \omega t} \\
        i \omega (\frac{q_0}{2} e^{i \omega t} - \frac{q_0}{2} e^{-i \omega t})
      \end{bmatrix}$$
    %%% end of your answer    
  }

  \part At this point, you have an expression for $q(t)$ that involves $i$.  Use
  Euler's formula, $e^{i\theta} =\cos \theta + i \sin \theta$, to express $q(t)$
  in terms of real quantities only.  \textit{\textbf{For the curious:} you can
    use these ideas to rewrite the general solution from part (d) as something
    involving $\cos$ and $\sin$ if you assume the initial conditions
    $q(0) = q_0$ and $\dot{q}(0) = v_0$.}

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{18em}{
    %%% Add your explanations here!
    Using the results of the initial conditions, using Euler's formula, we can rewrite $q(t)$ as:
    $$q(t) = \frac{q_0}{2} e^{i \omega t} + \frac{q_0}{2} e^{-i \omega t} = \frac{q_0}{2} (\cos(\omega t) + i \sin(\omega t)) + \frac{q_0}{2} (\cos(-\omega t) + i \sin(-\omega t))$$
    $$= \frac{q_0}{2} (\cos(\omega t) + i \sin(\omega t)) + \frac{q_0}{2} (\cos(\omega t) - i \sin(\omega t))$$
    The imaginary parts cancel out, so we have:
    $$q(t) = \frac{q_0}{2} (\cos(\omega t) + \cos(\omega t)) = q_0 \cos(\omega t)$$
    So, we have:
    $$q(t) = q_0 \cos(\omega t)$$
    %%% end of your answer    
  }

  \part Check if your found solution for $q(t)$ satisfies the ODE in
  \cref{eq:ode}.

  %%% Do not change the height of this box. Your work must fit inside it.
  \makenonemptybox{3em}{
    %%% Add your explanations here!
    $$ \ddot{q}(t) = -\omega^2 q(t)$$
    $$m \ddot{q}(t) = -k_1 q(t) - k_2 q(t) = -(k_1 + k_2)q(t)$$
    %%% end of your answer    
  }

\end{parts}

\end{questions}

\end{document}

### 2025-04-14 MSE160 Lecture Notes/main.tex
exam

@@
 \end{enumerate}
+\section{Thermaldynamics}
+\begin{center}
+    ``Free Time, Free Money, Free Energy?''
+\end{center}
+\paragraph{Thermaldynamics} Thermaldynamics is the study of heat and work. It is the study of the relationship between heat, work, and energy.
+\subsection{System, Surroundings, Universe and Boundary}
+\paragraph{System} The system is the part of the universe that we are interested in. It is the part of the universe that we are studying.
+\paragraph{Surroundings} The surroundings are the part of the universe that is not the system. It is the part of the universe that interacts with the system.   
+\paragraph{Universe} The universe is the system and the surroundings. It is the entire system that we are studying.
+\paragraph{Boundary} The boundary is the interface between the system and the surroundings. It is the interface that allows the system and the surroundings to interact.
+\paragraph{Open System} An open system is a system that can exchange both matter and energy with the surroundings.
+\paragraph{Closed System} A closed system is a system that can exchange energy but not matter with the surroundings.
+\paragraph{Isolated System} An isolated system is a system that cannot exchange matter or energy with the surroundings.
+\subsection{Thermaldynamics Variables}
+\paragraph{State Function} A state function is a function that depends only on the state of the system. It does not depend on the path taken to reach the state. \\
+(Example) Any energy is a state function: enthalpy, internal energy, Gibbs free energy, Helmholtz free energy, temperature, pressure, volume, entropy, etc.
+\paragraph{Path Function (Process Variable)} A path function is a function that depends on the path taken to reach the state. Examples of path functions include heat and work.
+(Example) Heat and work are path functions.
+\paragraph{Heat} Heat is the transfer of energy between a system and its surroundings due to a temperature difference. It is denoted by $q$.
+\paragraph{Work} Work is the transfer of energy between a system and its surroundings due to a force acting through a distance. It is denoted by $w$.
+\section{First Law of Thermaldynamics}
+\paragraph{Internal Energy} The internal energy of a system is the sum of the kinetic and potential energies of the particles in the system. It is denoted by $U$.
+\paragraph{First Law of Thermaldynamics} The first law of thermaldynamics is the law of conservation of energy.
+It states that the change in internal energy of a system is equal to the heat added to the system plus the work done on the system.
+It is given by:
+\begin{equation}
+    \Delta U = Q + W + W_{\text{other}}
+\end{equation}
+, where $Q$ is the heat added to the system, $W$ is the work done on the system, and $W_{\text{other}}$ are all other kinds of work done on the system. $W$ is given by:
+\begin{equation}
+    W = -P\Delta V
+\end{equation}
+\paragraph{} Thus, the energy in the universe is conserved, and the energy of a isolated system is constant. 
+\paragraph{Infinitesimal Change} For an infinitesimal change in the system, the first law of thermaldynamics is given by:
+\begin{equation}
+    dU = \delta Q + \delta W + \delta W_{\text{other}} 
+\end{equation}
+\paragraph{Sign Convention} The sign convention for heat and work is as follows:
+\begin{itemize}
+    \item \textbf{Heat} Heat added to the system is positive. Heat removed from the system is negative.
+    \item \textbf{Work} Work done on the system is positive. Work done by the system is negative.
+\end{itemize}
+\subsection{Second Law of Thermaldynamics}
+\paragraph{Second Law of Thermaldynamics} The entropy of an isolated system always increases in the course of a spontaneous process. It is given by:
+\begin{equation}
+    \Delta S \ge 0
+\end{equation}3
+\paragraph{Spontaneous Process} A spontaneous process is a process that occurs without any external intervention. It is a process that occurs naturally.
+\paragraph{Entropy} Entropy is defined by the change in heat transfer divided by the temperature. It is given by:
+\begin{equation}
+    dS = \frac{\delta Q}{T}
+\end{equation}
+\paragraph{Change in Entropy} The change in entropy of a system is given by:
+\begin{equation}
+    \Delta S_{\text{system}} = \Delta S_t + \Delta S_{p}
+\end{equation}
+\begin{example}[Cooling Coffee]
+    When you put a cup of hot coffee in a room, the coffee cools down and the room heats up. The entropy of the coffee decreases, and the entropy of the room increases:
+    \begin{align*}
+        \Delta S_{\text{coffee}} &= -\frac{\delta Q}{T_{\text{coffee}}} \\
+        \Delta S_{\text{room}} &= \frac{\delta Q}{T_{\text{room}}}
+    \end{align*}
+    The entropy of the universe increases:
+    $$
+    \Delta S_{\text{universe}} = \Delta S_{\text{coffee}} + \Delta S_{\text{room}} = \delta Q(\frac{1}{T_{\text{room}}} - \frac{1}{T_{\text{coffee}}}) > 0
+    $$
+\end{example}
+\subsection{Revrsibility and Irreversibility}
+\paragraph{Note} This concerns with the process variable, not the state variable.
+\paragraph{Reversible Process} A reversible process is a process that can be reversed without any change in the system or the surroundings. It is a process that occurs without any loss of energy.
+\paragraph{Irreversible Process} An irreversible process is a process that cannot be reversed without any change in the system or the surroundings.
+When reverese tue external influence on the system, the complciated sequence of events will not be the same as the original.
+Reversible is only an idealization.
+\paragraph{Spontaneous Process} A spontaneous process is an irreversible process. It is a process that occurs without any external intervention.
+\subsection{Enthaly and Gibb's Energy}
+\paragraph{Enthalpy} Enthalpy is the heat content of a system. It is given by:
+\begin{equation}
+    H = U + PV
+\end{equation}
+\paragraph{Infinitesimal Change in Enthalpy} The infinitesimal change in enthalpy is given by:
+\begin{equation}
+    dH = TdS + VdP + \delta W'
+\end{equation}
+\begin{Gibb's Energy} Gibb's energy is the energy of a system that is available to do work. It is given by:
+\begin{equation}
+    \Delta G = \Delta H - T\Delta S
+\end{equation}
+\subsection{Thermaldynamics in Standard Chemical Reactions}
+\paragraph{Standard State} The standard state is the state of a material at 1 atm and 25$^\circ$C. It is denoted by a circle above the variable.
+\paragraph{Standard Enthalpy of Formation} The standard enthalpy of formation is the enthalpy change when one mole of a compound is formed from its elements in their standard states.
+It is denoted by $\Delta H_f^\circ$.\\
+\begin{example}
+    For element $\ch{A}$ and $\ch{B}$, the standard enthalpy of formation for $\ch{A}_a\ch{B}_b$ is denoted as:
+    \begin{equation}
+        \Delta H_f^\circ(\ch{A}_a\ch{B}_b) = a\Delta H_f^\circ(\ch{A}) + b\Delta H_f^\circ(\ch{B})
+    \end{equation}
+\end{example}
+\paragraph{Enthalpy of Reaction} The enthalpy of reaction is the enthalpy change when a reaction occurs. It is denoted by $\Delta H_r$. It is given by:
+\begin{equation}
+    \Delta H_r = \sum \Delta H_f^\circ(\text{products}) - \sum \Delta H_f^\circ(\text{reactants})
+\end{equation}
+\begin{example}[Carbon Dionixe]
+    We have the reacittion, in standard state:
+    \begin{align*}
+        \ch{CO} + \frac{1}{2}\ch{O2} &\rightarrow \ch{CO2} \\
+        \Delta H_f^\circ(\ch{CO2}) &= \Delta H_f^\circ(\ch{CO}) + \frac{1}{2}\Delta H_f^\circ(\ch{O2})
+        \intertext{And for $\ch{CO}$:}
+        \Delta H_f^\circ(\ch{CO}) &= -110.5 \text{ kJ/mol}
+        \intertext{And for $\ch{O2}$:}
 
+    \end{align*}
+\end{example}
+\subsection{Change of Entropy for Chemical Reactions}
+\paragraph{Single Phase} Recall the chagne of entropy as a function of pressure and temperature:
+\begin{equation}
+    dS = \frac{\delta Q}{T} = \frac{C_p}{T}dT
+\end{equation}
+, wehre $C_p$ is the heat capacity at constant pressure, which is energy for a mole of substance to raise the temperature by 1 K.\\
+Like enthalpy, the change of entropy is a state function. Thus, we can integrate it:
+\begin{equation}
+    \Delta S = \int_{T_1}^{T_2} \frac{C_p}{T}dT
+\end{equation}
+\paragraph{Entropy of Melting and Vaporization} The entropy of melting and vaporization is the change in entropy when a substance melts or vaporizes. It is given by:
+\begin{equation}
+    \Delta S_{melt} = \frac{\Delta H_{melt}}{T_{melt}}
+\end{equation}
+\begin{equation}
+    \Delta S_{vap} = \frac{\Delta H_{vap}}{T_{vap}}
+\end{equation}
+\paragraph{Entropy of Reaction} The entropy of reaction is the change in entropy when a reaction occurs. It is given by:    
+\paragraph{Hess's Law} Hess's law states that the change in enthalpy for a reaction is the same regardless of the number of steps in the reaction. It is given by:
+\section{Phase Diagram}
+\paragraph{Phase Diagram} A phase diagram is a graph that shows the relationship between temperature on the vertical axis, and compositon of the right component on the horizontal axis.
+The phase diagram shows the different phases of a material at different temperatures and compositions.
+\paragraph{Isotherm} An isotherm is a line on the phase diagram that is a constant temperature. 
+\paragraph{Solutbiliy Limit} The solubility limit is the maximum amount of solute that can be dissolved in a solvent at a given temperature.
+It is the line on the phase diagram that separates the single phase region from the two phase region.
+\paragraph{Eutetic Point} The eutectic point is the point on the phase diagram where the solubility limit is reached and some temperture, in which a liquid solution form two solid phases. 
+\paragraph{Eutetic Reaction} The eutectic reaction is the reaction that occurs at the eutectic point. It is given by:
+\begin{equation}
+    L \rightarrow S_1 + S_2
+\end{equation}
+, for example, Syrup to Ice and Sugar Crystal.
+\paragraph{Single Phase Region} The single phase region is the region on the phase diagram that is a single phase
+\paragraph{Two Phase Region} The two phase region is the region on the phase diagram that is a two phase region. 
+\paragraph{Tie Line} A tie line is a horizontal line drawn in two phase region. It is used to determine the composition of the two phases in the two phase region.
+\paragraph{After the Eutectic Point} After the eutectic point, the liquid phase is no longer stable.
+The liquid phase will solidify into two solid phases.
+The composition of the two solid phases can be determined by the tie line.
+\paragraph{Liquidus Line} The liquidus line is the line on the phase diagram that separates the single phase region from the two phase region.
+It is the line that indicates the temperature at which a liquid solution forms two solid phases.
 \end{document}

### 2025-09-01 YEAR1/CIV 102 Lecture Notes/AppendixItems/old.tex
pre year 2

\input{preamble}
\begin{document}

\thispagestyle{empty}
\begin{center}
{\LARGE \bf CIV 102 Notes}\\
{\large Hayson Cheung}\\
Structures and Materials, Fall 2024\\
\vspace{10pt}
\textit{"Structural Engineering is the art and science of designing and molding structures with economy and elegance so that they can safely resist the force that they are subjected''}
\end{center}
\section{Force, Moment, and Load} 
\subsection{Moment}
\begin{definition}[Moment]
    The moment is an object's tendency to rotate, measured in Newton-Meters (\textbf{Nm}).
\end{definition}
\begin{shaded}
    \paragraph{Force-Coupled Moment} Two forces act in \textbf{opposite directions} but the \textbf{same magnitude} creates force-coupled moment.
    \begin{equation}
        \mu = \mathrm{moment} = \Vec{F}\times\Vec{d}
    \end{equation}
    Where:
    \begin{equation*}
    \begin{split}
    \Vec{F} = \text{Coupled forces (N)} \\
    \Vec{d} = \text{Displacement between the coupled forces (m)} \\
    \end{split}
    \end{equation*}
    \\Note that the moment is calculated in the cross-products, and the direction of that vector is the axis of rotation,
    \end{shaded}
\paragraph{Signs} If the moments return a positive value, it is a counter-clockwise rotation, and vice versa. 
\paragraph{Static Equilibrium} The state in which the sum of all component forces, including moments, equals zero. (In civ, things are most likely not supposed to move, so this is used often)
\begin{equation}
    \sum F_x=0,\sum F_y=0, \sum \mu = 0
\end{equation}
\begin{shaded}
    \paragraph{Moment for One Force} The moment due to only one force is:
    \begin{equation}
        \mu = \mathrm{moment} = \Vec{F}\times\Vec{d_\perp}
    \end{equation}
Where:
\begin{equation*}
\begin{split}
\Vec{d_\perp} = \text{The perpendicular displacement to the center of rotation}
\end{split}
\end{equation*}
\end{shaded}
\subsection{Moment of Inertia}
\paragraph{} For Linear Force, we have:
$$
F=ma
$$
\begin{shaded}
    \paragraph{Mass-moment Inertia} Its counterpart in moment is the following
    \begin{equation}
        \mu = \mathrm{moment} = \omega\alpha
    \end{equation}
    Where:
    \begin{equation*}
    \begin{split}
    \omega = \text{Mass-Moment Inertia} \\
    \Vec{d} = \text{Rotational Acceleration} \\
    \end{split}
    \end{equation*}
\end{shaded}
\paragraph{}
\paragraph{} and for position, acceleration, and velocity:
\begin{equation}
    x = \theta y
\end{equation}
\begin{equation}
    \dot{x} = \dot{\theta} y
\end{equation}
\begin{equation}
    \ddot{x} = \ddot{\theta} y
\end{equation}
\paragraph{Multiple point Masses} 
\begin{equation}
    I_m = \Sigma_i y_i^2 m_i
\end{equation}
\paragraph{Distributed Mass} 
\begin{equation}
    I_m = \int y^2 d m
\end{equation}
\paragraph{} Let $\rho$ be the mass density, assume constant thickness, so $dm = \rho t dA$
\subsection{Hanging Cables}
\begin{equation}
    I_m = \rho t\iint y^2 dA
\end{equation}
$I = \iint y^2 dA$ is called the moment of inertia. For $I$ in a rectangle with the axis of rotation in its middle, $R$, we have.
\begin{equation}
    I = bh^3/12
\end{equation}
\begin{minipage}{0.55\textwidth}
\begin{definition} [Type I Hanging Cables]
    Cables under its weight, follow the shape of a \textbf{catenery}. It has \textbf{non-uniform} horizontal weight distribution. 
\end{definition}
\paragraph{} This is not a practical case as there certainly would always be loads other than the cable.
\end{minipage}
\begin{minipage}{0.05\textwidth}
    \hspace{\textwidth}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\noindent
\resizebox{0.75\textwidth}{1in}{
\begin{tikzpicture}
    % Define the catenary curve (sinh-based approximation)
    \draw[thick, domain=-3.5:3.5, samples=100, smooth, variable=\x] 
    plot ({\x}, {cosh(\x)-1});

    \foreach \x in {-3.5,-3,-2.25,-1.5,-0.75,0,0.75,1.5,2.25,3,3.5} {
        \pgfmathsetmacro{\y}{cosh(\x)-1}
        \fill[red] (\x, \y) circle (0.08); % Circles representing loads
        \draw[dashed] (\x, \y) -- (\x, 0);
    }

\end{tikzpicture}
}
\end{minipage}
\begin{definition} [Type II Hanging Cables]
    Cables loaded with uniform weight horizontally. It follows the shape of a \textbf{parabola}. It has \textbf{uniform} horizontal weight distribution. (This is because $\ddot{y}$ is constant.)
\end{definition}

\subsection{Load and its Distribution}
\paragraph{Axial Force} Forces that align on their member axis. 
\paragraph{Load Path} The paths in which the load follows.
\begin{example}[load path of a suspension bridge]
    car-deck-hanger cable-main cable-tower-earth
\end{example}
\subsubsection{Distributed Load/Force}
\begin{definition}[load]
    Load is force distributed across a surface, measured in Nm$^{-1}$ 
\end{definition}
\paragraph{} There are two methods to describe load distribution. Multiple Point Load and Distributed Load (UDL).
\paragraph{Multiple Point Load} This could be useful for specific points acting as the load. However, for large structures, the points tend to infinity and we shall use Distributed Load.
\begin{center}
\resizebox{0.7\textwidth}{!}{ % Adjust the width to your preference
\begin{tikzpicture}

    % Draw the surface (flat beam)
    \draw[thick] (-5,0) -- (5,0); % Length L = 10 units

    % Label the surface as length L
    \node[below] at (0,-0.2) {$L$};

    % Define the locations of point loads (can be modified as needed)
    \def\pointLoadLocations{-4,-2,0,2,4} % x-coordinates of point loads
    \def\pointLoadLabels{} % Labels for point loads

    % Draw point loads
    \foreach \x [count=\i from 1] in \pointLoadLocations {
        % Draw the arrow for the point load
        \draw[->, thick] (\x, 1) -- (\x, 0);
        % Label the point load
        \node[above] at (\x, 1) {$P_\i$};
    }

\end{tikzpicture}
}
\end{center}
\paragraph{Distributed load} Load is distributed across $L$. it models the case where each load is infinitesimally small. 
\begin{center}
\resizebox{0.7\textwidth}{!}{ % Adjust the width to your preference
\begin{tikzpicture}

    % Draw the surface (flat beam)
    \draw[thick] (-5,0) -- (5,0); % Length L = 10 units
    \draw[thick] (-5,1.5) -- (5,1.5);

    % Label the surface as length L
    \node[below] at (0,-0.2) {$L$};

    % Define the locations of point loads (can be modified as needed)
    \def\pointLoadLocations{-4.5,-4,-4,4,4.5} % x-coordinates of point loads
    \def\pointLoadLabels{} % Labels for point loads

    % Draw point loads
    \foreach \x [count=\i from 1] in \pointLoadLocations {
        % Draw the arrow for the point load
        \draw[->, thick] (\x, 1.5) -- (\x, 0);
        % Label the point load
    }

    \foreach \x [count=\i from 1] in {-3,-2,-1,0,1,2,3} {
        % Draw the arrow for the point load
        \filldraw (\x,0.75) circle (1.5pt);;
        % Label the point load
    }
\end{tikzpicture}
}
\end{center}
When the load ($w$) is uniform across (\textbf{U}niformly \textbf{D}istributed \textbf{L}oad), the load  is calculated by:
\begin{equation}
    w = \frac{F}{L}
\end{equation}
\paragraph{Total Force due to Load} The load could be non-uniform. The total force is, in general:
\begin{equation}
    F=\int w dl
\end{equation}
the location of the force would be on the centroid (center of mass). Which is in the middle of \textbf{UDL}. In addition, for UDL, the total force is simply the area so:
\begin{equation}
    F= wL
\end{equation}
moreover, for non-UDL that increase/decrease linearly, it is the case of a \textbf{triangle} so:
\begin{equation}
    F= \frac{w_\text{max}L}{2}
\end{equation}
\section{Property of Materials}
\subsection{Stress and Strain}
\begin{center}
    \textit{"Stress and Strain are orthogonal - they are independent of each other''}
\end{center}
\begin{shaded}
\paragraph{Stress} ($\sigma$) is the distribution of applied force normalized on the plane. It is in the unit of MPa in CIV. Given by:
\begin{equation}\label{stress}
    \sigma = \frac{\Vec{F}}{A_\perp}
\end{equation}
Where:
    \begin{equation*}
    \begin{split}
    \Vec{F} = \text{Applied Force (MN)} \\
    A = \text{Cross section area normal to $\Vec{F}$ (m$^2$)} \\
    \end{split}
    \end{equation*}
\begin{note}
It is usually scaler considering uni-axial problems.
\end{note}
\end{shaded}
\begin{shaded}
    \paragraph{Strain} ($\epsilon$) is the unitless ratio between the material's actual length ($L$) and its change of length ($\Delta L$). Given by:
    \begin{equation}\label{strain}
        \epsilon = \frac{\Delta L}{L_0}
    \end{equation}
\begin{note}
Units of mm/m could also be used. $0.001=$1 mm/m is often an acceptable strain value in structures.
\end{note}
\end{shaded}
\paragraph{$\sigma$-$\epsilon$ Diagram} \textbf{Tensional stress} involves forces in opposite directions, which results in a strain. The relationship between stress and stress obeys \textbf{Hooke's law} at low stress and strain.
    \begin{equation}\label{sigmaepsilon}
        \sigma = E\epsilon
    \end{equation}
    Where:
    \begin{equation*}
    \begin{split}
    E = \text{\textbf{Young's Modulus}, the ratio between stress and stain} \\
    \end{split}
    \end{equation*}
so $\sigma$-$\epsilon$ diagram is linear when these two qualities are small and there are no effects of deformation. 
\begin{example}[Combined Stress-Stain Equation]
    Combining (\ref{sigmaepsilon}), (\ref{strain}) and (\ref{stress}), we have:
    \begin{equation}
        F = \frac{EA}{L}\Delta L
    \end{equation}
    And by Hooke's law, the spring constant, $k$, is 
    \begin{equation}
        k = \frac{EA}{L}
    \end{equation}
\end{example}
\subsection{Mild Steel}
\paragraph{Ductility} Ductile material could resist stress. It follows Hooke's law before it goes to a smaller strain-stress ratio. 
\paragraph{Brittle} Brittle material breaks at a certain lower level of stress-strain. 
\paragraph{Mild Steel} It follows a special stress-strain relationship. 
\begin{center}
\resizebox{0.7\textwidth}{!}{
\begin{tikzpicture}[scale=1.5]

    % Axis
    \draw[->] (0,0) -- (7,0) node[below] {Strain};
    \draw[->] (0,0) -- (0,5) node[left] {Stress};

    % Elastic region
    \draw[thick] (0,0) -- (2,3.5) node[midway, above left, font=\small] {Elastic Region};

    % Yield point
    \node at (2,3.5) [circle,fill=black,inner sep=1.5pt]{};
    \node[font=\small] at (-1,3.5) {Yield Stress ($\sigma_y$)};
    \draw[thick,dashed] (0,3.5) -- (2,3.5);

    % Yield plateau
    \draw[thick] (2,3.5) -- (4,3.5) node[midway, above, font=\small] {Yield Plateau};

     \draw[thick] (4,3.5) -- (5.5,4.5) node[midway, above] {Strain Hardening};

    % Necking and fracture
    \draw[dashed, thick] (5.5,4.5) -- (6.5,3.5) node[midway, right, font=\small] {Necking};

    % Labels
    \node[font=\small] at (5.7,4.7) {Ultimate Strength ($\sigma_u$)};

    % Ultimate stress point
    \node at (5.5,4.5) [circle,fill=black,inner sep=1.5pt]{};

    % Fracture point
    \node at (6.5,3.5) [circle,fill=black,inner sep=1.5pt]{};
    \node[font=\small] at (6.5,3.3) {Fracture};

\end{tikzpicture}
}
\end{center}
The stress-strain diagram has the following regions and properties.
\begin{enumerate}
    \item \textbf{Elastic Region} 
    \begin{itemize}
        \item Strains recoverable
        \item Linear, obeys Hooke's Law
    \end{itemize}
    \item \textbf{Yield Plateau Region}
    \begin{itemize}
        \item Non-recoverable strain
        \item Permanent damage
    \end{itemize}
    \item \textbf{Strain Hardening Region} Often ignored in design.
\end{enumerate}
\begin{note}
    There are two types of stress. The stress we would only discuss considers the original cross-section area, but true stress considers the current cross-section area. Ruptures occur when true stress is at a maximum, and that explains the decrease in stress in the necking region. 
\end{note}
\paragraph{Strain Energy} Is the area under the curve of a $\sigma-\epsilon$ diagram. It is in the unit of (MPa=MJm$^{-3}$).
\begin{equation}
    \int \sigma d\epsilon
\end{equation}
\paragraph{} For mild steel, the recoverable, elastic region is said to be the elastic strain energy. The non-revocable region is said to be the plastic strain energy.
\section{Oscillations}
\subsection{Dynamics in Single Degree of Freedom Systems}
\paragraph{Single Degree of Freedom Systems} Could be Subsets of the oscillation system
\begin{itemize}
    \item \textbf{Electrical}
    \item \textbf{Mechanical} $a\neq0$
    \item \textbf{Electromagnetic system}
\end{itemize}
\paragraph{Dynamic Equilibrium} D'Alembert's principle: An accelerating system may be placed in dynamic equilibrium by introducing and inertia for $F_i$ acting in the opposite direction to the acceleration. The have the following ODE:
\begin{align}
    0&=kx+m\ddot{x}\\\nonumber
    \Leftrightarrow x(t)&=A\sin(\omega_n+\phi)
\end{align} 
To verify, we have:
\begin{equation*}
    0 = A\sin(\omega_n+\phi)- mA\omega_n^2\sin(\omega_n+\phi)
\end{equation*}
That is true as per the natural frequency of vibration
\begin{equation}
    \omega_n=\sqrt{\frac{k}{m}}
\end{equation}
hence,
\begin{equation}
    f_n = \frac{1}{2\pi}\sqrt{\frac{k}{m}}
\end{equation}
\paragraph{Free Vibrations with Gravity} We have the rest displacement $\Delta_0=\frac{mg}{k}$
\begin{equation}
    x(t)=A\sin(\omega_n+\phi)+\Delta_0
\end{equation}
$k$ is not usually easy to find, so we can find frequency by  $\Delta_0=\frac{mg}{k}$
\begin{equation}
    f_n=\frac{1}{2\pi}\sqrt{\frac{g}{\Delta_0}}
\end{equation}
\section{Safety}
\begin{center}
    \textit{``The concept of safety is to avoid people getting hurt"}
\end{center}
\paragraph{Uncertainty Quantification} The measure of the significance of uncertainty, these could stem from:
    \begin{enumerate}
        \item Strength of Material and Geometry.
        \item Applied Load: Dead Load (unmovable) of Life Load (movable).
    \end{enumerate}
% APPENDIX
\paragraph{Evaluation of Safety} There are two main methods of determining the safety level:
\begin{enumerate}
    \item \textbf{Reliability} Risk, the probability of failure, multiplied by the consequences of failure.
    (Example) \textbf{Histograms of strength and load, risk related to overlapping region}
    \item \textbf{Working Stress/Allowable Stress Method} For factor of safety, we have
    \begin{equation}
        \text{Factor of safety (F.O.S)}=\frac{\text{Average Strength}}{\text{Average Applied Load}}
    \end{equation}
    In practice, to calculate allowable stress, we have:
    \begin{equation}
        \sigma_\text{allowable}=\frac{\sigma_y}{\text{F.O.S.}}
    \end{equation}
    Historically, F.O.S. is 10$\times$ during the 19$^\text{th}$ century. For example,
    \begin{itemize}
        \item \textbf{Brooklyn Bridge} (1800s): F.O.S. = 5.0$\times$.
        \item \textbf{Golden Gate Bridge} (1930s): F.O.S. = 2.6$\times$.
        \item \textbf{Akashi Kaikyo Bridge} (1998s): F.O.S. = 2.25$\times$.
    \end{itemize}
    F.O.S. decreases as we are more confident in our calculation. 
\end{enumerate}
\subsection{Structural Analysis}
\paragraph{Beam and Colume} Beam are horizontal and volume is vertical usually carrying axial load. 
\section{Design of Truss Bride}
\begin{center}
    \textit{``Strong, light, and stiff"}
\end{center}
\paragraph{Anatomy of Truss Bridge} See figure. MEmeber connects with joints (pin joints, hinge joints). 
\paragraph{}
\subsection{Moment of Inertia}
\begin{definition}[Curvature]
    Curvature $\phi$ [m$^{-1}$] is the measure of bentness between two ends $A$ and $B$, defined by:
    \begin{equation}
        \phi = \frac{\theta_A - \theta B}{L_0} = \frac{d \theta}{dx}
    \end{equation}
\end{definition}
    On the strain diagram plotted against depth, the angle is $\phi$, by the small angle approximation:
    \begin{equation}
        \phi = \frac{\epsilon_\text{top} - \epsilon_\text{bottom}}{h}
    \end{equation}
    Or we can use the relationship between $\phi$ and moment:
    \begin{equation}
        \phi =  \frac{M}{EI}
    \end{equation}
We have:
\begin{align*}
    \epsilon_i &= \phi y_i\\
    \sigma_i &= E\phi y_i\\
    F_i = E\phi y_i \Delta A
    \sum_i F_i = \sum_i E\phi y_i \Delta A
\end{align*}
\paragraph{First Moeenot of area} Since total force is zero, So we have:
\begin{equation}
    \phi E \int y dA = 0 
\end{equation}
The integral is known as the first moment of area. 
\begin{equation}
     \int y dA = 0 
\end{equation}
\paragraph{Stress and Bending Moment} The relation is given by the Navier's Equation. Since $\phi = \frac{\sigma}{Ey} = \frac{M}{EI}$, we have:
\begin{equation}
    \sigma = \frac{My}{I}
\end{equation}
\begin{shaded}
    \begin{equation}
        I = \iint y^2 dA
    \end{equation}
    Strain at a depth of $y$:
    \begin{equation}
        \epsilon_y = \phi y
    \end{equation}
    Moment:
    \begin{equation}
        M - EI\phi
    \end{equation}
    \begin{equation}
        I = \frac{bh^3}{12}
    \end{equation}
\end{shaded}
\subsection{Euler Buckling}
\paragraph{Buckling} Buckling is only possible in compression. The transition between axial shortening and buckling is said to be $P_\text{crit}$. It is important to prevent buckling as the structure fails without warning.
\begin{definition}[Elastic Buckling, Euler (1957)]
    Assume constant $E$ along the length, and $I$ is constant along the length. We have a pin at the top and bottom, that is, two ends can rotate. We also assume that the length is perfectly straight before buckling and we would use small angle approximation (since we only look at $P_\text{crit}$). \\
    \par Let $\Delta y$ be the horizontal distance of the bent. For small angle, we have
    $$
    \theta = \frac{\delta y}{\delta x} = \frac{y(L/2)}{L/2} 
    $$
    $$
        P_n = \frac{\pi^2 n^2 EI}{L^2}
    $$
    $$
        P_\text{crit} = \frac{\pi^2 EI}{L^2}
    $$
    $n$ is the count of the node in between the lenght
\end{definition}
\subsection{Curvature}
\paragraph{Curvature by 3 Method} Given $R$, the radius of the arc, the curvature
\begin{equation}
    \phi = 1/R
\end{equation}
\section{Wind}
\paragraph{Wind Bracing} 
\section{Virtual Work}
\paragraph{Work} In elastic deformation, for internal energy, we have
$$
W_\text{int} = V\frac{\sigma \epsilon}{2} = \frac{P\Delta}{2}
$$
In Hookes Law, for external energy, we have:
$$
W_\text{ext} = F\Delta r
$$
\paragraph{} For change of length, we have 
\begin{equation}
\Delta = \frac{PL}{EA}
\end{equation}
\paragraph{Deflection} To calculate the extension of the truss we can use a combined equation incorporating a virtual unit load $F^\star = 1$, and the equation $W_\text{ext} = W_\text{int}$
\begin{subequations}
    \begin{equation}
        \frac{1}{2}F_x\Delta x + \frac{1}{2}F_y\Delta y = \frac{1}{2}P_x\Delta x + \frac{1}{2}P_y\Delta y
    \end{equation}
    \begin{equation}
        \frac{1}{2}F^\star_x\Delta^\star x + \frac{1}{2}F^\star_y\Delta^\star y = \frac{1}{2}P^\star_x\Delta^\star x + \frac{1}{2}P^\star_y\Delta^\star y
    \end{equation}
    \begin{equation}
        (F^\star_x+F_x)(\Delta^\star x +\Delta x) + (F^\star_y+F_y)(\Delta^\star y +\Delta y) = (P^\star_x+P_x)(\Delta^\star x +\Delta x) + (P^\star_y+P_y)(\Delta^\star y +\Delta y)
    \end{equation}
\end{subequations}
Combining all three, after some algebra, we have:
\begin{equation}
    F^\star \Delta_{\hat{r}} = \sum^{i}P^\star_i \Delta_i
\end{equation}
\section{Vibraion}
\subsection{Free Vibration}
Since for change of length, we have:
\begin{align}
    \Delta = \frac{PL}{EA} \nonumber 
    \intertext{This could be rewritten as}
    P = \frac{EA}{L}\Delta \nonumber
    \intertext{For stiffness $k$, this could be molded as a simple harmonic motion with:}
    k = \frac{EA}{L}
\end{align}
\paragraph{} The solution to the differential equation $m\ddot{x} + k\dot{x} = mg$ is:
\begin{equation}
    x= A\sin{(\omega_nt + \phi)} + \Delta_0
\end{equation}
Where:
\begin{equation*}
\begin{split}
\omega_n = \sqrt{\frac{k}{m}}\text{, [rad s$^{-1}$] The natural frequency of vibration}\\
\Delta_0 =\text{Initial Displacement of the joint}
\end{split}
\end{equation*}
\paragraph{For Truss} We can use the method of virtual load to determine $\Delta_0$. 
\paragraph{Point Load} The natural frequency is \begin{equation}
\frac{15.76}{\sqrt{\Delta_0}}
\end{equation}
\paragraph{Uniform Load} The natural frequency is 
\begin{equation}
\frac{17.76}{\sqrt{\Delta_0}}
\end{equation}
\paragraph{Damping} Energy absorbed by friction, wind, etc.
\paragraph{} The differential equation for this is $m\ddot{x} + c\dot{x} + kx = mg$, where for some critical damping usually $\beta \in [0, 1 \text{ or } 2], c = 2\beta\sqrt{mk}$, for which the solution would be:
\begin{equation}
    x= A\sin{(\omega_n\sqrt{1-\beta}t + \phi)} + \Delta_0
\end{equation}
\subsection{Forced Vibration - Dynamic Amplification Factor} For some frequency of vibration $\omega$, the differential equation would be $m\ddot{x} + c\dot{x} + kx = mg + F_0\sin(\omega t)$
\begin{equation}
    x = \text{DAF} \frac{F_0}{k}\sin{(\omega t + \phi)}
\end{equation}
\paragraph{Dynamic Amplification Factor} Denoted as DAF, defined as
\begin{equation}
    \text{DAF} = \frac{1}{\sqrt{(1-(\frac{f}{f_n})^2)^2+ (2\beta \frac{f}{f_n})^2}}
\end{equation}
\section{Shear and Bending Moment} Consider a member AB with weight $w$, length $L$ and uniform load $P$ with a roller support on the left and a pin support on the right side. The member also has a constant axial force $N$. Let $N, V, M$ be the functions of axial force, sheer force and bending moment respectively w.r.t. $x$, the distance from left and right. We have, by equilibrium equations
\begin{subequations}
    \begin{equation}
        F_x: N(x) = N
    \end{equation}
    \begin{equation}
        F_y: V(x) = P + wx - Ay
    \end{equation}
    \begin{equation}
        M: M(x) = -\frac{wx^2}{2} - P(x-\frac{L}{2}) + xA_y
    \end{equation}
\end{subequations}
\paragraph{Relationship between $V$ and $M$}
\begin{subequations}
\begin{equation}
    V(x) = \frac{dM}{dX}
\end{equation}
\begin{equation}
    M(x) = \int_0^x V(x) dx
\end{equation}
\paragraph{Drawing Sheer Force Diagram} Imagine walking along and accumulating sheer force.
\paragraph{Drawing Bending Moment Diagram}  Imagine walking along and accumulating the area under the sheer force diagram. 
\paragraph{} Both graph at the end should add to zero. 
\end{subequations}
\subsection{Bending Movement and Stress}
\appendix
\def\thesection{\Alph{section}}
\input{appendix}
\end{document}

### 2025-09-01 YEAR1/MAT185_Assignments/MAT185_Assignment_2/MAT185_Assignment_2.tex
pre year 2

\documentclass[10pt,addpoints]{exam} %% no solution

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[top=2.0cm,bottom=2.0cm,left=2.5cm,right=2.5cm]{geometry}
\usepackage{tikz}
\usepackage{float}
\usepackage{multicol}
\usepackage{pgfplots}
\usepackage{lastpage}
\usepackage{soul}
\usepackage{libertine}
\usepackage{lipsum}
\usepackage{siunitx}
\usepackage{xspace}
\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}
\usepackage[labelfont=bf]{caption}
\usepackage[hidelinks, urlcolor=blue, linkcolor=black, colorlinks=true]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage[absolute]{textpos}

\usepackage[normalem]{ulem}

\newcommand{\todo}[1]{\hl{\textbf{ToDo:}~#1}}

%% define course title
\newcommand{\course}{MAT185}
\newcommand{\assignmenttitle}{Assignment 2}

%% header and footer
\firstpageheader{}{}{\textbf{{\color{red} Due}: 10:00pm, Tuesday Feb. 4, 2025}}
\firstpageheadrule
\runningheader{}{Page~\thepage~of~\numpages}{\course~--~\assignmenttitle}
\footer{}{}{}

\setlength\parindent{0pt} % no indentation in document

%% formats exam class
\qformat{\textbf{Question \thequestiontitle:}\hfill} % title of question 
\boxedpoints
\pointpoints{mark}{marks}
\pointsinrightmargin
\hpword{Marks:}
\hsword{Your score:}
\unframedsolutions
\totalformat{\boxed{\textnormal{\totalpoints~\if\totalpoints1 mark\else marks\fi}}}
\definecolor{SolutionColor}{rgb}{0,0,1}
\renewcommand{\solutiontitle}{}
\AtBeginEnvironment{solution}{\color{blue}}

% %correct choices in solution
\CorrectChoiceEmphasis{\rm}
\checkedchar{\tikz\draw[blue,fill=blue] (0,0) circle (1ex);}

% %
\newcommand{\mblue}[1]{\color{blue}\textit{#1}\color{black}}

% % increase distance between checkbox items
\renewcommand{\checkboxeshook}{\setlength{\itemsep}{6pt}}

%% distance between questions and parts
\renewcommand{\questionshook}{\setlength{\parsep}{10pt}}
\renewcommand{\partshook}{\setlength{\parsep}{15pt}}

%% define arrows in text
\newcommand{\arrow}{$\rightarrow$\xspace}
\newcommand{\Arrow}{$\Rightarrow$\xspace}

% % math notation:
%\veccol{1}{2}{3}
\newcommand{\veccol}[3]{
    \begin{bmatrix}
        #1\\
        #2\\
        #3\\
    \end{bmatrix}}

%\vecrow{1}{2}{3}
\newcommand{\vecrow}[3]{\left[#1~#2~#3\right]}

%\matrixTwo{1}{2}{3}{4}
\newcommand{\matrixTwo}[4]{\left[\begin{array}{cc}#1&#2\\#3&#4\end{array}\right]}

% \matrixThree{1}{2}{3}{4}{5}{6}{7}{8}{9}
\newcommand{\matrixThree}[9]{\left[\begin{array}{ccc}#1&#2&#3\\#4&#5&#6\\#7&#8&#9\end{array}\right]}

%\matrixCorner{1}{2}{3}{4}
\newcommand{\matrixCorner}[4]{\left[\begin{array}{ccc}#1& \cdots&#2\\ \vdots & \ddots & \vdots\\#3&
      \cdots&#4\end{array}\right]}

% real numbers
\newcommand{\R}{\mathbb{R}}

% \nR
\newcommand{\nR}{{}^{n}\mathbb{R}}
% \mR
\newcommand{\mR}{{}^{m}\mathbb{R}}
% \Rn
\newcommand{\Rn}{\mathbb{R}^{n}}
% \nRn
\newcommand{\nRn}{{}^{n}\mathbb{R}^{n}}
% \nRm
\newcommand{\nRm}{{}^{n}\mathbb{R}^{m}}
% \nRm
\newcommand{\mRn}{{}^{m}\mathbb{R}^{n}}
% \mRm
\newcommand{\mRm}{{}^{m}\mathbb{R}^{m}}

%% define abbreviations
\newcommand{\row}{\operatorname{row}\,}
\newcommand{\col}{\operatorname{col}\,}
\renewcommand{\dim}{\operatorname{dim}\,}
\renewcommand{\span}{\operatorname{span}\,}
\newcommand{\rank}{\operatorname{rank}\,}
\renewcommand{\ker}{\operatorname{ker}\,}
\newcommand{\nul}{\operatorname{null}\,}
\renewcommand{\det}{\operatorname{det}\,}
\newcommand{\adj}{\operatorname{adj}\,}

%%% This command makes a framed box of a chosen height.
\newcommand{\makenonemptybox}[2]{%
\par\nobreak\vspace{\ht\strutbox}\noindent
\setlength{\fboxrule}{0pt} % set this to 0pt to make invisible
\fbox{%
\parbox[c][#1][t]{\dimexpr\linewidth-2\fboxsep}{
  \hrule width \hsize height 0pt
  \vspace{-0.6cm}
  \color{SolutionColor}#2\color{black}
 }%
}%
}

\begin{document}

\vspace*{-0.5cm}
\begin{center}
  \large
  \textbf{\Large \course~--~Linear Algebra}\\[0.1cm]
  \textbf{\assignmenttitle}
\end{center}
\bigskip

\textbf{\large Instructions:}\\
\normalsize

Please read the {\bf MAT185 Assignment Policies \& FAQ} document for
details on submission policies, collaboration rules and academic integrity, and
general instructions.

\begin{enumerate}

\item \textbf{Submissions are only accepted by}
  \href{https://www.gradescope.ca}{Gradescope}. Do not send anything by email.
  Late submissions are not accepted under any circumstance. Remember you can
  resubmit anytime before the deadline.

\item \textbf{Submit solutions using only this template pdf}.  Your submission
  should be a single pdf with your full written solutions for each question. If
  your solution is not written using this template pdf (scanned print or
  digital) then your submission will not be assessed. Organize your work neatly
  in the space provided.  Do not submit rough work.

\item \textbf{Show your work and justify your steps} on every question but do
  not include extraneous information.  Put your final answer in the box
  provided, if necessary.  We recommend you write draft solutions on separate
  pages and afterwards write your polished solutions here on this template.

\item \textbf{You must fill out and sign the academic integrity statement
    below}; otherwise, you will receive zero for this assignment.

\end{enumerate}

\vspace{10pt}

\textbf{\large Academic Integrity Statement:} \\

%%% Student information

% Student 1
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: CHEUNG, Hei Shing (Hayson)\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010907823\enspace\hrulefill}\\
  \end{minipage}
}

\vspace*{0.1in}

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: SIU, Nelson\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010940608 \enspace\hrulefill}\\
  \end{minipage}
}

\bigskip
\large \textbf{I confirm that:}
\normalsize

\begin{itemize} 
\item I have read and followed the policies described in the document {\bf
    MAT185 Assignment Policies \& FAQ}.
\item In particular, I have read and understand the rules for
  collaboration, and permitted resources on assignments as described in
  subsection II of the the aforementioned document. I have not violated
  these rules while completing and writing this assignment.
\item I have not used generative AI in writing this assignment.
\item I understand the consequences of violating the University's academic
  integrity policies as outlined in the
  \href{http://www.governingcouncil.utoronto.ca/policies/behaveac.htm}{Code of
    Behaviour on Academic Matters}. I have not violated them while completing
  and writing this assignment.
\end{itemize}
\bigskip

\large \textbf{By submitting this assignment to Gradescope, I agree that the
  statements above are true.}  \normalsize

\newpage

\section*{Preamble:}

Image processing is the manipulation of digital images by applying mathematical tools and
algorithms. A wide range of applications based on digital image processing are, for example, medical
imaging, image optimization in consumer cameras, computer vision, and satellite imagery. Linear
algebra, and the techniques you will learn during this course, plays a crucial role in that field by
providing a mathematical foundation.\\

A typical digital image can be considered as a 2-dimensional matrix of \textit{pixels} (abbreviation
for \textit{picture element}). Methods of digital image processing are manipulating this
2D-matrix. A pixel is the smallest element of a digitally acquired raster image, and can be
considered as a colour sample at each point of an image. Typically, each pixel is represented by 3
positive integer values from 0 to 255 for each colour red, green, and blue: 0 is representing black
and 255 ($= 2^8-1$) either red, green, or blue. In that case, 24 Bits are used to code 16,777,216
distinct colours for each pixel. This is called a 24\,bpp (24\,Bits-per-pixel) colour depth. If a
grey-scale image is sampled, each pixel samples the light intensity. In that case, only 8\,Bits
(single integers from 0 to 255) are typically necessary, as shown in \cref{fig:imagematrix}, to
store grey-scale images.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\textwidth]{imagematrix.pdf}
  \caption{Grey-scale image represented by 8 bpp
    (Bits-per-pixel). [\textit{source: stanford.edu}]}
  \label{fig:imagematrix}
\end{figure}

One application of linear algebra in the context of digital image processing is for a linear
transformation of the images / 2D-matrices. We will discuss linear transformations in more detail
later this term. Common transformations include scaling, rotation, and translation of the image. All
of these linear transformations can be represented by a matrix
multiplication.\\

Another important application of linear algebra in the context of digital image processing is
filtering, which will be explored in Question~\ref{q:filtering} of this assignment. Filtering
includes, for example, methods for noise reduction, blurring/sharpening, edge
detection, white balancing, colour correction, and many more.\\

Lastly, images have to be stored in efficient ways. For a colour image of size $n\times m$ with a
24\,bpp colour depth, $m \times n \times 24$\,Bits of memory are necessary. An image of
$8000\times 6000$ Pixels (4:3 aspect ratio and 48MP resolution of modern smartphone cameras) with
24\,Bit colour depth will lead to an uncompressed image filesize of 144\,MB. To reduce the image
filesize, lossy (permanently removing \textit{unnecessary} information of the original image) and
lossless compression methods are applied. For lossy compression algorithms, it is vital to identify
unnecessary information of the image, which are not perceived by the viewer and can be removed. One
approach for a lossy compression is based on the Singular Value Decomposition (SVD), which will be
explored in more detail in Question~\ref{q:sdv}.

\newpage

\begin{questions}
  \question\label{q:filtering} For simplicity, we assume that each
    pixel of a grey-scale image is considered as a real value in
    Question~\ref{q:filtering}(\ref{part:first}) to (\ref{part:second_last}).
  Let $G \in \nRm$ be a grey-scale image of $n \times m$ pixels. A collection of
  image filters $ F_1, F_2, \dots, F_k \in \nRm$ can be used to process this
  image, resulting in filtered images $A_l$ represented as:
  \begin{equation}
    \label{eq:filter}
    A_l = F_l \circ G, \quad l = 1, 2, \dots, k,
  \end{equation}
  where $ \circ $ denotes the entry-wise product. The entry-wise product (also called
  \textit{Hadamard} product) of two matrices of the same size is defined as the product, where each
  entry of the resulting matrix is the product of the corresponding entries of the original
  matrices.
  For example, if $P = \matrixTwo{p_{11}}{p_{12}}{p_{21}}{p_{22}}$ and
  $Q= \matrixTwo{q_{11}}{q_{12}}{q_{21}}{q_{22}}$ are $2\times 2$ matrices, the Hadamard product is
  defined as
  \begin{equation}
  P\circ Q = \matrixTwo{p_{11}q_{11}}{p_{12}q_{12}}{p_{21}q_{21}}{p_{22}q_{22}}.
\end{equation}

  \begin{parts}

    \part\label{part:first} Can the filter $F_1$ in \cref{eq:filter} be used to blur or smooth an image
      $A_1$? Unsupported answers will not receive full credit.

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{3cm}{
        %%% Your work goes here!
        %        
        % \textbf{YES}, since $F_1$ can be anything, assume that we can compute the filter $F_1$ based on the image $A_1$.
        % We can apply a bluring algorithm to the image $G = \{g_{ij}\}$ to get the filter $F_1 = \{f_{ij}\}$. One naive way to do this is to average the pixel values in a $3\times 3$ window around each pixel:
        \textbf{YES}. Define \textit{blurring} as an operation so that the average gradient between pixels decreases. 
        Apply a blurring algorithm to the image $G = \{g_{ij}\}$ to get the filter $F_1 = \{f_{ij}\}$. 
        One naive way to do this is by averaging the each pixel values in a $3\times 3$ window, and devide by the pixel value to get the filter $F^\prime_1$, hence $F^\prime_1 \circ G$ is \textit{blurred}. 
        You specified for the same filter $F_1$ to blur the filterd image $A_1 = F_1 \circ G$ when applied as a Hadamard Product. 
        We can modify the same said averaging filter to blur both $A_1$ and $G$. We can do it by taking the root of the filter values, which allows us to gradually reuse the same filter as Haddamard product: $A_2 = F_1 \circ A_1 = F_1 \circ F_1 \circ G$. Further explanation and results: \href{https://github.com/HaysonC/svdAssignment}{https://github.com/HaysonC/svdAssignment}
        \begin{equation*}
           f_{ij} = \sqrt[3]{\frac{1}{9g_{ij}} \sum_{0\le s,t\le 2} g_{i+s-1,j+t-1}}  = \sqrt[3]{\frac{\operatorname{conv}(\{1/9\}_{3\times 3}, G)_{ij}}{g_{ij}}}
        \end{equation*}
        % Then, we can apply the filter $F_1$ to the image $G$ to get the blurred image $A_1= F_1 \circ G$ \\
        % \tiny{\\ \textbf{Note for 1(a)} Assume we can compute $F_1$ based on the entire image. Otherwise, the answer to 1(a) is \textbf{NO}, because smoothing and bluring must consider nearby pixels, element-wise multiplication cannot achieve that since each operation is independent of other pixels in the matrix. For instance, the example as shown in my solution is that of a convolutional multiplication with a $3\times3$ kernel $K = \frac{1}{9}\{1\}_{3\times3}$, which is a simple blurring kernel, with $A_1 = K \ast G$. Also, if you consider a zero matrix (black iamge) as a blured one, then for all image we can apply the the zero matrix filter.}  The filter $F_1$ would be a $3\times 3$ matrix with entries $\frac{1}{9}$.
        }
    \part Name at least \textbf{three} applications in the context of digital
    image processing for a filter as defined in \cref{eq:filter}. Additionally,
    give details how $F_i$ would look like for your applications.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{10cm}{
      \begin{enumerate}
        \item \textbf{Adusting white balance/color temperature} The filter $F_i$ can be used to adjust the white balance of an image. By setting a filter to enhance the red or blue channel, we can adjust the color temperature of the image by mulitplying the filter to the specific channel. We assume that a colored image is defined as $\{R, G, B\}$, then $\{F_i\circ R, G,B\}$ will be the new image with adjusted white balance. In this case $F_i$ would be a uniform matrix with entries greater than one if we want to enhance the color, and less than one if we want to decrease the color.
        \item \textbf{Cinematic Effects} Certain filters can be used to create cinematic effects. For example, a filter that dims the edges of image (vignette) can be applied by multiplying the filter to the image. The filter will have a ones in the center and approaches zero as it goes to the edges.
        \item \textbf{Artifcially Increasing Exposure} A filter can be used to artificially increase the exposure of an image. By multiplying the filter to the image, the pixel values will be increased by a certain factor. The filter will have values greater than one, otherwise, we can use the filter to decrease the exposure. The increased exposure image will be $\{F_i\circ G\}$, and it would look like a brighter version of the original image.
      \end{enumerate}

      %%% end of your answer    
    }

\newpage

\part \label{part:linind} Assume that the entries of the image $G$ are all nonzero.  Prove that the
  set of filtered images $\{F_1\circ G, F_2\circ G, \dots, F_k\circ G\}$ is linearly independent if
  and only if the set of filters $\{F_1, F_2, \dots, F_k\}$ is linearly independent.

  %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{17cm}{
      %%% Your work goes here!      
      (``$\Rightarrow$") Assume, for the sake of contrapositive, that the set of filters $\{F_1, F_2, \dots, F_k\}$ is linearly dependent. Then, we proof that exists a non-trivial linear combination of the filtered images that equals the zero vector. We first consider the linear combination of the filtered images, by the Distributive Property, as:
      $$ \lambda_1(F_1\circ G) + \lambda_2(F_2\circ G) + \dots + \lambda_k(F_k\circ G)$$
      Since multiplicati on in the Hadamard product is commutative, we can consider $\lambda_j(F_j\circ G) = G\circ(\lambda_j F_j)$. We can then rewrite the linear combination as:
      $$ G\circ(\lambda_1F_1 + \lambda_2F_2 + \dots + \lambda_kF_k)$$
      Since there exist a non-trivial linear combination of the filters that equals the zero vector, we take ${\mu_1, \mu_2, \dots, \mu_k} \neq \textbf{0}_k$ such that $\mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k = 0$. We coule then pick $\lambda_j = \mu_j$ and rewrite the linear combination as:
      $$ G\circ(\mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k) = G\circ\textbf{0}$$
      Thus, since $G \neq \textbf{0}$, there also exist a non-trivial linear combination of the filtered images that equals the zero vector, which implies that the set of filtered images is linearly dependent.\\ \\
      (``$\Leftarrow$") Assume, for the sake of contrapositive, that the set of filtered images is linearly dependent. Then, we proof that there exist a non-trivial linear combination of the filters that equals the zero vector. We first consider the linear combination of the filters:
      $$ \mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k$$
      Since the images are of any real matrix of a specific size, we can take the Hadamard product of the linear combination with the image $G$, again, by the Distributive Property, we have:
      $$ G\circ(\mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k) = G\circ\textbf{0} = \textbf{0}$$
      Again, by communativity of the Hadamard product, we can rewrite the linear combination as:
      $$ \mu_1(F_1\circ G) + \mu_2(F_2\circ G) + \dots + \mu_k(F_k\circ G) = \textbf{0}$$
      Since $\{\mu_1, \mu_2, \dots, \mu_k\} \neq \textbf{0}_k$, there exist a non-trivial linear combination of the filtered images that equals the zero vector, which implies that the set of filters is linearly dependent.\\
    }

   \newpage

  \part Assume that the entries of the image $G$ are all nonzero.  Let $\mathcal{W}$ be the set of
    the filtered images $\{F_1 \circ G, F_2 \circ G, \dots, F_k \circ G\}$. Suppose a new image
    filter $ F_{k+1} $ is introduced. Prove that the filtered image $ F_{k+1} \circ G $ lies in
    $\operatorname{span} \mathcal{W}$ if and only if $ F_{k+1} $ is a linear combination of
    $ \{F_1, F_2, \dots, F_k\} $.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{17cm}{
      %%% Your work goes here!      
      (``$\Rightarrow$") Assume that $F_{k+1}\circ G$ lies in $\operatorname{span} \mathcal{W}$. Then, we proof that $F_{k+1}$ is a linear combination of $\{F_1, F_2, \dots, F_k\}$. Since $F_{k+1}\circ G$ lies in $\operatorname{span} \mathcal{W}$, we can write $F_{k+1}\circ G$ as a linear combination of the filtered images:
      $$ F_{k+1}\circ G = \lambda_1(F_1\circ G) + \lambda_2(F_2\circ G) + \dots + \lambda_k(F_k\circ G)$$
      By the Distributive Property, we can rewrite the linear combination as:
      $$ F_{k+1}\circ G = (\lambda_1F_1 + \lambda_2F_2 + \dots + \lambda_kF_k) \circ G$$
      Since $G = \{g_{ij}\}$ is a non-zero matrix, we can multiply both sides by $G^{-1} = \{1/g_{ij}\}$ (The inverse such that $G\circ G^{-1} = \textbf{1}$):
      $$ F_{k+1} \circ G \circ G^{-1} = (\lambda_1F_1 + \lambda_2F_2 + \dots + \lambda_kF_k) \circ G \circ G^{-1}$$
      Since $G\circ G^{-1} = \textbf{1}$, we have:
      $$ F_{k+1} \circ \textbf{1} = (\lambda_1F_1 + \lambda_2F_2 + \dots + \lambda_kF_k) \circ \textbf{1}$$
      So we have:
      $$ F_{k+1} = \lambda_1F_1 + \lambda_2F_2 + \dots + \lambda_kF_k$$
      Thus, $F_{k+1}$ is a linear combination of $\{F_1, F_2, \dots, F_k\}$.\\ \\
      (``$\Leftarrow$") Assume that $F_{k+1}$ is a linear combination of $\{F_1, F_2, \dots, F_k\}$. Then, we proof that $F_{k+1}\circ G$ lies in $\operatorname{span} \mathcal{W}$. Since $F_{k+1}$ is a linear combination of $\{F_1, F_2, \dots, F_k\}$, we can write $F_{k+1}$ as:
      $$ F_{k+1} = \mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k$$
      We can then multiply both sides by $G$:
      $$ F_{k+1}\circ G = (\mu_1F_1 + \mu_2F_2 + \dots + \mu_kF_k) \circ G$$
      By the Distributive Property, we can rewrite the linear combination as:
      $$ F_{k+1}\circ G = \mu_1(F_1\circ G) + \mu_2(F_2\circ G) + \dots + \mu_k(F_k\circ G)$$
      Thus, $F_{k+1}\circ G$ lies in $\operatorname{span} \mathcal{W}$.\\
      %%% end of your answer    
    }

    \newpage

    \part \label{part:second_last} Assume $ F_1 = \matrixTwo{1}{0}{0}{1} $,
    $ F_2 = \matrixTwo{0}{1}{1}{0} $, and $ F_3 = \matrixTwo{1}{1}{0}{0} $ are
    filters for a grey-scale image $G \in {}^2\mathbb{R}^2$. Verify whether the
    filters $ F_1, F_2, F_3 $ are linearly independent, and determine whether
    $ F_4 = \matrixTwo{2}{1}{1}{4} $ lies in the span of $ \{F_1, F_2, F_3\} $.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{14cm}{
      %%% Your work goes here!      
      Consider the linear combination of the filters, with $a,b,c \in \mathbb{R}$:
      $$ aF_1 + bF_2 + cF_3 = \matrixTwo{a}{0}{0}{a} + \matrixTwo{0}{b}{b}{0} + \matrixTwo{c}{c}{0}{0} = \matrixTwo{a+c}{b+c}{b}{a} = \matrixTwo{0}{0}{0}{0}$$
      So we have the system of equations:
      $$
      S = \begin{cases}
        a+c = 0\\
        b+c = 0\\
        b = 0\\
        a = 0
      \end{cases}
      $$
      The system of equations has only the trivial solution $a=b=c=0$, so the filters $F_1, F_2, F_3$ \textbf{are linearly independent}.\\

      To determine whether $F_4$ lies in the span of $\{F_1, F_2, F_3\}$, we can write $F_4$ as a linear combination of $\{F_1, F_2, F_3\}$, we consider the system of equations:
      $$
      S = \begin{cases}
        a+c = 2\\
        b+c = 1\\
        b = 1\\
        a = 4\\
      \end{cases}
      $$
      Substitute $b=1$ and $a=4$ into the first two equations, we have $c=-2$ and $c=0$, which is a contradiction. Thus, $F_4$ \textbf{does not} lie in the span of $\{F_1, F_2, F_3\}$.
      %%% end of your answer    
    }

  \part For saving grey-scale images $A$ of size $n\times m$ pixels in an efficient way, each pixel
    is stored as an 8\,Bit integer, where 0 represents black and 255 represents white (see
    preamble). Let $\mathcal{V}$ be the set of these grey-scale images. Is $\mathcal{V}$ a vector
    space if one uses entry-wise vector addition and scalar multiplication?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{3cm}{
      %%% Your work goes here!      
      \textbf{NO}, $\mathcal{V}$ is not a vector space. The set of grey-scale images $\mathcal{V}$ is not closed under addition. For example, we take two images $A$ and $B$ where they are both full white images, we have:
      $$ A = \{255\}_{n\times m}, B = \{255\}_{n\times m} \quad A+B = \{510\}_{n\times m} \notin \mathcal{V}$$      
      %%% end of your answer    
    }
  \end{parts}

\newpage

\question \label{q:sdv} Let $A \in \mRn$ be a grey-scale image of size $m \times n$, where each
  entry is representing a grey pixel. As you learnt in ESC103, the rank of an $n\times m$ matrix is
  the number of linear independent rows, which is the same as the number of the linear independent
  columns.

  The Singular Value Decomposition\footnote{In this writing assignment, we are introducing you to
    the compact (or reduced) SVD, which is slightly different from the full SVD. To save space, we
    refer to it as the ``SVD'' throughout.} of a matrix $A$ is a way of writing $A$ as the product of three
  matrices:
\begin{equation}
  \label{eq:1}
  A=U \Sigma V^T
\end{equation}
with the
matrices $U$, $\Sigma$, and $V^T$ defined as

\begin{equation}
  \begin{aligned}
    & A=\left[\begin{array}{lllll}
                \mathbf{u}_{\mathbf{1}} & \mathbf{u}_{\mathbf{2}} & \cdots & \mathbf{u}_{\mathbf{n}-\mathbf{1}} & \mathbf{u}_{\mathbf{r}}
              \end{array}\right]\left[\begin{array}{lllll}
                                        \sigma_1 & & \cdots& & 0\\
                                                 & \sigma_2 & & & \\
                                                 \vdots& & \ddots & & \vdots\\
                                                 & & & \sigma_{r-1} & \\
                                                 0& & \cdots& & \sigma_r
                                      \end{array}\right]\left[\begin{array}{c}
                                                                \mathbf{v}_{\mathbf{1}}^T \\
                                                                \mathbf{v}_{\mathbf{2}}{ }^T \\
                                                                \vdots \\
                                                                \mathbf{v}_{\mathbf{r-1}}{ }^T \\
                                                                \mathbf{v}_{\mathbf{r}}{ }^T
                                                              \end{array}\right]
  \end{aligned}
  \label{eq:sdv_1}
\end{equation}

where $\mathbf{u}_i$ are the columns of $U$ and $\mathbf{v}^T_i$ are the rows of $V^T$. The matrix
$\Sigma$ is a diagonal matrix with the entries $\sigma_i$ on the diagonal (the diagonal entries of
$\Sigma$ are nonnegative and all other entries of $\Sigma$ are zero). Because $\Sigma$ is diagonal,
\cref{eq:sdv_1} can be written as:

    \begin{equation}
      \label{eq:sumA}
      A=\mathbf{u}_{\mathbf{1}} \sigma_1 \mathbf{v}_{\mathbf{1}}{
      }^T+\mathbf{u}_{\mathbf{2}} \sigma_2 \mathbf{v}_{\mathbf{2}}{
      }^T+\cdots+\mathbf{u}_{\mathbf{r}-\mathbf{1}} \sigma_{r-1}
      \mathbf{v}_{\mathbf{r}-\mathbf{1}}{ }^T+\mathbf{u}_{\mathbf{r}} \sigma_r
      \mathbf{v}_{\mathbf{r}}{ }^T.
    \end{equation}

    As you can see, $A$ is separated into the sum of (rank=1) matrices of the
    form $\sigma_i \mathbf{u}_{\mathbf{i}}\mathbf{v}_{\mathbf{i}}{ }^T$. With
    that, the sum in \cref{eq:sumA} can be expressed as
    \begin{equation}
      \label{eq:2}
      A=\sum_{i=1}^r \sigma_i \mathbf{u}_{\mathbf{i}} \mathbf{v}_{\mathbf{i}}^T
    \end{equation}
    with $r$ so called singular values
    $\left\{\sigma_i \in \mathbb{R}~|~\sigma_i > 0\right\}$ and
    both $\{ \mathbf{u}_1, \dots \mathbf{u}_r \} \subseteq \mR$ and 
    $\{ \mathbf{v}_1, \dots \mathbf{v}_r \} \subseteq \nR$ are orthonormal sets of vectors.

    What's an ``orthonormal set of vectors''?  In simple language, each vector has length one and
    any two vectors are perpendicular (also called ``orthogonal'').  More specifically, two vectors,
    $\mathbf{x}, \mathbf{y} \in \nR$, are called \mblue{orthogonal}\ if and only if their
    dot-product is equal to zero: $\mathbf{x} \cdot \mathbf{y} = \mathbf{x}^T\mathbf{y} = 0$.  The
    set $\{\mathbf{x}_1,\mathbf{x}_2,\dots \mathbf{x}_k\} \subseteq \nR$ is an \mblue{orthogonal set
      of vectors}\ if and only if every pair of vectors is orthogonal:
    $\mathbf{x}_i \cdot \mathbf{x}_j = 0 $ if $i \neq j$.  The set
    $\{\mathbf{x}_1,\mathbf{x}_2,\dots \mathbf{x}_k\} \subseteq \nR$ is an \mblue{orthonormal set of
      vectors}\ if and only if every pair of vectors is orthogonal and each vector has length $1$:
    $\mathbf{x}_i \cdot \mathbf{x}_j = 0 $ if $i \neq j$ and $\mathbf{x}_i \cdot \mathbf{x}_i = 1$
    for every $1 \leq i, j \leq k$.

    At the end of this course, you will have the tools you need to be able to find the SVD of a
    general matrix.\footnote{\textit{For the curious:} $\mathbf{u}_i$ are eigenvectors of the matrix
      $AA^T$ and $\mathbf{v}_i$ are eigenvectors of the matrix $A^TA$.}  We can use \cref{eq:sumA}
    to compress the image $A$ by keeping the ``most relevant'' terms in the sum: As you can see from
    \cref{eq:sumA}, some matrices $\sigma_i \mathbf{u}_{\mathbf{i}} \mathbf{v}_{\mathbf{i}}^T$ could
    be neglected if $\sigma_i$ is very small. Therefore, a typical strategy for compressing images
    is putting the singular values $\sigma_i$ in nonincreasing order,
    $\sigma_1 \geq \sigma_2 \geq \dots \sigma_r > 0$ (as well as sorting the corresponding
    $\mathbf{u}_i$ and $\mathbf{v}_i$ in $U$ and $V$, respectively) and only storing $k < r$ of
    these components in an image file ($k$ is an integer of at least 1). The compressed image $A_k$
    is constructed as
  \begin{equation}
    \label{eq:3}
    A_k=\sum_{i=1}^k \sigma_i \mathbf{u}_{\mathbf{i}} \mathbf{v}_{\mathbf{i}}^T
  \end{equation}

  This will reduce the image size dramatically, however, the image quality will
  be degraded if $k$ was picked too small, see for example
  \cref{fig:svd} where 30 components were used ($k = 30$) for the compression of
  the initial grey-scale image.

  \begin{figure}[H]
    \centering
    \includegraphics[width=1.0\textwidth]{svd}
    \caption{Lossy compression of a $700\times 700$ pixel grey-scale image via
      Single Value Decomposition (SVD).}
    \label{fig:svd}
  \end{figure}

  \begin{parts}

    \part For what values of $k$ would $A_k$ corresponds to a lossy
      compressed image? No additional explanation is necessary.

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{0.5cm}{
        %%% Your work oes here!    
        For all $0 \le k < \operatorname{rank}(A)$

        %%% end of your answer    
      }

    \part Let $A= \sigma\mathbf{x}\mathbf{y}^T$ where
      $\mathbf{x},\mathbf{y} \in {}^n\mathbb{R}$ and $\sigma \in \R$ is nonzero. Prove that
      $\operatorname{rank}(A)=1$. \color{red}\textit{Hint:~}\color{black}
    Try constructing a $3 \times 3$ example, $A$, by choosing some
    nonzero $\sigma\in\mathbb{R}$ and
      $\mathbf{x}, \mathbf{y} \in {}^3\mathbb{R}$. This should give you the insight you 
      need to answer this question for
      general $\mathbf{x},\mathbf{y} \in {}^n\mathbb{R}$ below.

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{9cm}{
        %%% Your work goes here!      
        Let $A = \sigma\mathbf{x}\mathbf{y}^T$ where $\mathbf{x},\mathbf{y} \in {}^3\mathbb{R}$ and $\sigma \in \R$ is nonzero. Then, WLOG, we use a $3 \times 3$ example:
        $$ A = \sigma\mathbf{x}\mathbf{y}^T = \sigma\begin{bmatrix} x_1 \\ x_2 \\ x_3 \end{bmatrix} \begin{bmatrix} y_1 & y_2 & y_3 \end{bmatrix} = \begin{bmatrix} \sigma x_1y_1 & \sigma x_1y_2 & \sigma x_1y_3 \\ \sigma x_2y_1 & \sigma x_2y_2 & \sigma x_2y_3 \\ \sigma x_3y_1 & \sigma x_3y_2 & \sigma x_3y_3 \end{bmatrix}$$
        This extends to a general $n\times n$ matrix $A$:
        $$ A = \sigma\mathbf{x}\mathbf{y}^T = \sigma\begin{bmatrix} x_1 \\ x_2 \\ \vdots \\ x_n \end{bmatrix} \begin{bmatrix} y_1 & y_2 & \cdots & y_n \end{bmatrix} = \begin{bmatrix} \sigma x_1y_1 & \sigma x_1y_2 & \cdots & \sigma x_1y_n \\ \sigma x_2y_1 & \sigma x_2y_2 & \cdots & \sigma x_2y_n \\ \vdots & \vdots & \ddots & \vdots \\ \sigma x_ny_1 & \sigma x_ny_2 & \cdots & \sigma x_ny_n \end{bmatrix}$$
        We can see that the matrix $A$ is a rank-1 matrix, since all the columns are scalar multiples of each other. Thus, $\operatorname{rank}(A) = 1$.

        %%% end of your answer    
      }

\newpage

\part Let
$A = \begin{bmatrix} 36 & 9 & 12 \\ -48 & -12 & -16 \\ 144 & 36 &
  48 \end{bmatrix}$. Determine the singular value decomposition
$A = \sigma {\bf x} {\bf y}^T$.

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{3cm}{
        %%% Your work goes here!      
        $\operatorname{Col}(A) = \begin{bmatrix}
          3 & -4 & 6
        \end{bmatrix}^T$ so $\operatorname{Rank}{A} = 1$. 
        We normalize the column vector to get ${\bf x} = \begin{bmatrix} 3/\sqrt{61} \\ -4/\sqrt{61} \\ 6/\sqrt{61} \end{bmatrix}$.
        Also, we normalize $y = \begin{bmatrix} 12/13 & 3/13 & 4/13 \end{bmatrix}^T$.
        \\Thus, $A = \sigma {\bf x} {\bf y}^T = 13\sqrt{61}\begin{bmatrix} 3/\sqrt{61} \\ -4/\sqrt{61} \\ 6/\sqrt{61} \end{bmatrix} \begin{bmatrix} 12/13 & 3/13 & 4/13 \end{bmatrix}$

        %%% end of your answer    
      }

    \part\label{part:ortho} Assume  $\{\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3\}$ is an
      orthogonal set of nonzero vectors.  Prove that the set is linearly
      independent. \label{part:orthogonal}

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{10cm}{
        %%% Your work goes here!      
        We consider the linear combinations of the set of vectors that equals the zero vector:
        $$ \lambda_1\mathbf{v}_1 + \lambda_2\mathbf{v}_2 + \lambda_3\mathbf{v}_3 = \textbf{0}$$
        Since the set is orthogonal, we have:
        $$ 
        \begin{cases}
          \mathbf{v}_1 \cdot \mathbf{v}_2 = 0\\
          \mathbf{v}_1 \cdot \mathbf{v}_3 = 0\\
          \mathbf{v}_2 \cdot \mathbf{v}_3 = 0
        \end{cases}
        $$
        We can then take the dot product of the linear combination with $\mathbf{v}_1$, $\mathbf{v}_2$, and $\mathbf{v}_3$, also since $0\cdot\mathbf{v}_i = 0$, we have:
        $$ \mathbf{v}_1 \cdot (\lambda_1\mathbf{v}_1 + \lambda_2\mathbf{v}_2 + \lambda_3\mathbf{v}_3) = 0 \implies \lambda_1\mathbf{v}_1 \cdot \mathbf{v}_1 + \lambda_2\mathbf{v}_1 \cdot \mathbf{v}_2 + \lambda_3\mathbf{v}_1 \cdot \mathbf{v}_3 = \lambda_1\|\mathbf{v}_1\|^2 = 0$$
        Similarly, we will have:
        $$ \lambda_2\|\mathbf{v}_2\|^2 = 0 \quad \lambda_3\|\mathbf{v}_3\|^2 = 0$$
        Since $\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3$ are non-zero vectors, we have $\|\mathbf{v}_1\|^2, \|\mathbf{v}_2\|^2, \|\mathbf{v}_3\|^2 > 0$. Thus, $\lambda_1 = \lambda_2 = \lambda_3 = 0$, which implies that the set of vectors is linearly independent.
        %%% end of your answer    
      }

    \part Consider the set $\{\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3 \}$ from
      part~(\ref{part:ortho}).
    What property of a 4th vector $\mathbf{v}_4 \in\ ^4\R$ 
       would ensure that the extended set
      $\{\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3, \mathbf{v}_4 \}$ spans
      ${}^4\mathbb{R}$? Give a short explanation.  Note: your explanation does not
      have to include how you would find such a $\mathbf{v}_4$.

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{3cm}{
        %%% Your work goes here!      
        $v_4$ shall be orthogonal to all vectors in the set $\{\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3\}$. This ensures that the extended set $\{\mathbf{v}_1, \mathbf{v}_2, \mathbf{v}_3, \mathbf{v}_4\}$ spans ${}^4\mathbb{R}$, since the set would remain linearly independent. By the Fundamental Theorem of Linear Algebra we learnt in MAT185, the size 4 set of linear independent vectors would span ${}^4\mathbb{R}$, a dim 4 vecotor space. This is beacuse the spanning set of $^4\mathbb{R}$ is at least 4, so a set of 4 linearly independent vectors would span $^4\mathbb{R}$ as it constitutes the basis thereof.

        %%% end of your answer    
      }

\newpage

    \part Download the Jupyter notebook (see Quercus page of \textit{Assignment
      2}), which includes the python code of an SVD analysis.  Please also
    download the three example grey-scale images \textit{image1.png},
    \textit{image2.png}, and \textit{image3.png} from the same Quercus page.
    You can run the code on the
    \href{https://jupyter.utoronto.ca/hub/user-redirect/tree}{U\,of\,T Jupyter
      server}, on your own local Jupyter installation, or on the
    \href{https://undergrad.engineering.utoronto.ca/campus-facilities/engineering-computing-facility-ecf/remote-access/}{ECF
      lab PCs}. We recommend using the U\,of\,T Jupyter server, since all
    necessary packages are already installed:
    \href{https://jupyter.utoronto.ca/hub/user-redirect/tree}{https://jupyter.utoronto.ca/hub/user-redirect/tree}. For
    that, please login with your U\,of\,T credentials and click
    '\textit{upload}' to upload the IPYNB file and images. After that, you can
    open the uploaded notebook and run the Python code.

    Use the Jupyter Python code to analyze the provided three images. Plot the
    sorted singular values $\sigma_i$ over the index $i$ in one graph and
    discuss the results for all three provided images. Do you see a pattern of
    the result depending on each input image?

    Insert your plot below.  In addition, briefly discuss which of the 3 images
    would allow for the highest and which for the lowest compression ratio by
    singular value
    decomposition. \textcolor{red}{\textit{Hint:}}\color{black}~The compression
    ratio is the uncompressed file size over the compressed file size. You don't
    need to calculate this ratio. However, think about which image would need
    the largest $k$ and which the lowest $k$ in \cref{eq:3} for an accurate
    representation of the uncompressed image, \cref{eq:2}.  For plotting
    $\sigma_i$, see the comments and plotting commands in the Jupyter
    file. Also, keep the logarithmic Y-axis for your graph.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{9cm}{
      %%% Add your graph and explanations here!      
      \begin{center}

        \includegraphics[width=0.6\textwidth]{graphall.png}
      \end{center}
      The graph shows the sorted singular values $\sigma_i$ over the index $i$ for the three provided images. We can see that the singular values for \textit{image1.png} decay the slowest, while the singular values for \textit{image3.png} decay the fastest. This implies that \textit{image1.png} would allow for the highest compression ratio, since it requires less $k$ for an accurate representation of the uncompressed image. On the other hand, \textit{image3.png} would allow for the lowest compression ratio, since it requires more $k$ for an accurate representation of the uncompressed image. If it has a faster decay, it means that the there would be less small singular values.
      %%% end of your answer    
    }

    \part Calculate the SVD of \textit{image1.png} with the given Jupyter code.
    For what choice of $k \in \mathbb{Z}$ would you expect a reduction in file
    size? To answer this question, estimate the memory needed to store the
    SVD-compressed image and compare it to the memory needed for the
    uncompressed 8\,bpp version of the image. \color{red}\textit{Hint:}
    \color{black} Estimate the memory size of the uncompressed image by using
    its size and given colour depth. Also note that each floating-point number
    uses 32\,Bit memory when stored.

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{2.5cm}{
      %%% Your work goes here!      
      Size of \textit{image1.png} = $512 \times 512$ pixels, 8\,bpp.\\
      Memory needed for uncompressed image = $512 \times 512 \times 8 = 2^{21} = 2\,097\,152$ bits.\\

      \textbf{Assumption} We assume that the fullt \textit{image1.png} is rank 512. In reality it might not, but 512 is the maximum rank possible, so any $k$ as a result of this will garantee a reduction in file size.\\

      Memory needed for $\Sigma = k \times 32$ bits.\\
      Memory needed for each of $U, V = 2 \times k \times 512 \times 32$ bits.\\
      Total memory needed for SVD-compressed image = $k \times 32 + 2 \times  k \times 512 \times 32 = 32k + 32768k = 32800k$ bits.\\

      To expect a reduction in file size, we would need $32800k < 2\,097\,152 \implies k < 64$.

      %%% end of your answer    
    }

  \end{parts}

\end{questions}

\end{document}

### 2025-09-01 YEAR1/MAT185_Assignments/MAT185_Assignment_3/MAT185_Assignment_3.tex
pre year 2

\documentclass[10pt,answers]{exam}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[top=2.0cm,bottom=2.0cm,left=2.5cm,right=2.5cm]{geometry}
\usepackage{tikz}
\usepackage{float}
\usepackage{multicol}
\usepackage{pgfplots}
\usepackage{lastpage}
\usepackage{siunitx}
\usepackage{xspace}
\usepackage[labelfont=bf]{caption}
\usepackage[hidelinks, urlcolor=blue, linkcolor=blue, colorlinks=true]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage[absolute]{textpos}
\usepackage{systeme}

\newcommand{\mblue}[1]{\color{blue}\textit{#1}\color{black}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\bz}{\mathbf{0}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\ba}{\mathbf{a}}
\newcommand{\bb}{\mathbf{b}}
\newcommand{\bc}{\mathbf{c}}
\newcommand{\TT}{ \mathbf{{\tiny T}}}
\newcommand{\T}{ \mathbf{{T}}}

\newcommand{\ds}{\displaystyle}

%% define course title
\newcommand{\course}{MAT185}
\newcommand{\assignmenttitle}{Assignment 3}

%% header and footer
\firstpageheader{}{}{\textbf{{\color{red} Due:} 10:00pm, Tuesday March 18, 2025}}
\firstpageheadrule
\runningheader{}{Page~\thepage~of~\numpages}{\course~--~\assignmenttitle}
\footer{}{}{}

\setlength\parindent{0pt} % no indentation in document

%% formats exam class
\qformat{\textbf{Question \thequestiontitle:}\hfill} % title of question 
\boxedpoints
\pointpoints{mark}{marks}
\pointsinrightmargin
\hpword{Marks:}
\hsword{Your score:}
\unframedsolutions
\totalformat{\boxed{\textnormal{\totalpoints~\if\totalpoints1 mark\else marks\fi}}}
\definecolor{SolutionColor}{rgb}{0,0,1}
\renewcommand{\solutiontitle}{}
\AtBeginEnvironment{solution}{\color{blue}}

% %correct choices in solution
\CorrectChoiceEmphasis{\rm}
\checkedchar{\tikz\draw[blue,fill=blue] (0,0) circle (1ex);}
\newcommand{\emptychoice}{\begin{tikzpicture}\draw (1,0) circle (1.0ex);\end{tikzpicture}}
\newcommand{\fullchoice}{\begin{tikzpicture}\draw[blue,fill=blue] (0,0) circle (1ex);\end{tikzpicture}}

% % checkbox which can be used outside the checkbox environment
\newcommand{\TrueChoice}{\ifprintanswers \fullchoice \else \emptychoice \fi}

% % increase distance between checkbox items
\renewcommand{\checkboxeshook}{\setlength{\itemsep}{6pt}}

%% distance between questions and parts
\renewcommand{\questionshook}{\setlength{\parsep}{10pt}}
\renewcommand{\partshook}{\setlength{\parsep}{15pt}}

%% define arrows in text
\newcommand{\arrow}{$\rightarrow$\xspace}
\newcommand{\Arrow}{$\Rightarrow$\xspace}

% % math notation:
%\veccol{1}{2}{3}
\newcommand{\veccol}[3]{
    \begin{bmatrix}
        #1\\
        #2\\
        #3\\
    \end{bmatrix}}

%\vecrow{1}{2}{3}
\newcommand{\vecrow}[3]{\left[#1~#2~ #3\right]}

%\matrixTwo{1}{2}{3}{4}
\newcommand{\matrixTwo}[4]{\left[\begin{array}{cc}#1&#2\\#3&#4\end{array}\right]}

% \matrixThree{1}{2}{3}{4}{5}{6}{7}{8}{9}
\newcommand{\matrixThree}[9]{\left[\begin{array}{ccc}#1&#2&#3\\#4&#5&#6\\#7&#8&#9\end{array}\right]}

%\matrixCorner{1}{2}{3}{4}
\newcommand{\matrixCorner}[4]{\left[\begin{array}{ccc}#1& \cdots&#2\\ \vdots & \ddots & \vdots\\#3&
      \cdots&#4\end{array}\right]}

% \nR
\newcommand{\nR}{{}^{n}\mathbb{R}}
% \mR
\newcommand{\mR}{{}^{m}\mathbb{R}}
% \Rn
\newcommand{\Rn}{\mathbb{R}^{n}}
% \nRn
\newcommand{\nRn}{{}^{n}\mathbb{R}^{n}}
% \nRm
\newcommand{\nRm}{{}^{n}\mathbb{R}^{m}}
% \nRm
\newcommand{\mRn}{{}^{m}\mathbb{R}^{n}}
% \mRm
\newcommand{\mRm}{{}^{m}\mathbb{R}^{m}}

% \u
\renewcommand{\u}{{\bf u}}      
% \v
\renewcommand{\v}{{\bf v}}      
% \w
\newcommand{\w}{{\bf w}}    
% \V
\newcommand{\V}{{\bf V}}

%% define abbreviations
\newcommand{\row}{\operatorname{row}\,}
\newcommand{\col}{\operatorname{col}\,}
\renewcommand{\dim}{\operatorname{dim}\,}
\renewcommand{\span}{\operatorname{span}\,}
\newcommand{\rank}{\operatorname{rank}\,}
\renewcommand{\ker}{\operatorname{ker}\,}
\newcommand{\nul}{\operatorname{null}\,}
\renewcommand{\det}{\operatorname{det}\,}
\newcommand{\adj}{\operatorname{adj}\,}

\usepackage{xcolor}
\definecolor{dkrgreen}{HTML}{009988} % this is the color-blind friendly teal from below
\definecolor{dkred}{HTML}{EE3377}  % this is the colour-blind friendly magenta from below
\definecolor{blue}{HTML}{1965B0} % this is the colour-blind friendly blue from below
%
% colour-blind-friendly colours from https://personal.sron.nl/~pault/
\definecolor{tolBlue}{HTML}{1965B0}
\definecolor{tolMedBlue}{HTML}{5289C7}
\definecolor{tolLightBlue}{HTML}{7BAFDE} 
\definecolor{tolRed}{HTML}{E8601C} 
\definecolor{tolYellow}{HTML}{F6C141}
\definecolor{tolTeal}{HTML}{009988}
\definecolor{tolCyan}{HTML}{33BBEE}
\definecolor{tolTeal}{HTML}{009988} 
\definecolor{tolOrange}{HTML}{EE7733} 
\definecolor{tolMagenta}{HTML}{EE3377} 
\definecolor{tolGrey}{HTML}{BBBBBB}

%%% This command makes a framed box of a chosen height.
\newcommand{\makenonemptybox}[2]{%
\par\nobreak\vspace{\ht\strutbox}\noindent
\setlength{\fboxrule}{0pt} % set this to 0pt to make invisible
\fbox{%
\parbox[c][#1][t]{\dimexpr\linewidth-2\fboxsep}{
  \hrule width \hsize height 0pt
  \vspace{-0.6cm}
  \color{SolutionColor}#2\color{black}
 }%
}%
\setlength{\fboxrule}{1pt}
}

\newcommand{\makeframednonemptybox}[3]{%
\setlength{\fboxrule}{0.5pt}
\fbox{%
\parbox[c][#1][t]{#2}{%
  \hrule width \hsize height 0pt%
  \vspace{\dimexpr(#1/2-0.1cm)}%
  \color{SolutionColor}#3\color{black}
 }%
}%
\setlength{\fboxrule}{1pt}
}

\begin{document}

\vspace*{-0.5cm}
\begin{center}
  \large
  \textbf{\Large \course~--~Linear Algebra}\\[0.1cm]
  \textbf{\assignmenttitle}
\end{center}
\bigskip

\textbf{\large Instructions:}\\
\normalsize

Please read the {\bf MAT185 Assignment Policies \& FAQ} document for details on
submission policies, collaboration rules and academic integrity, and general
instructions.

\begin{enumerate}

\item \textbf{Submissions are only accepted by}
  \href{https://www.gradescope.ca}{Gradescope}. Do not send anything by email.
  Late submissions are not accepted under any circumstance. Remember you can
  resubmit anytime before the deadline.

\item \textbf{Submit solutions using only this template pdf}.  Your submission
  should be a single pdf with your full written solutions for each question. If
  your solution is not written using this template pdf (scanned print or
  digital) then your submission will not be assessed. Organize your work neatly
  in the space provided.  Do not submit rough work.

\item \textbf{Show your work and justify your steps} on every question but do
  not include extraneous information.  Put your final answer in the box
  provided, if necessary.  We recommend you write draft solutions on separate
  pages and afterwards write your polished solutions here on this template.

\item \textbf{You must fill out and sign the academic integrity statement
    below}; otherwise, you will receive zero for this assignment.

\end{enumerate}

\vspace{10pt}

\textbf{\large Academic Integrity Statement:} \\

%%% Student information

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: Siu, Nelson\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010940608\enspace\hrulefill}\\
  \end{minipage}
}
\vspace*{0.1in}

% Student 2
\fbox{
  \begin{minipage}{\textwidth}
    \vspace{0.75cm}
    \makebox[\textwidth]{\large Full Name: Cheung, Hei Shing\enspace\hrulefill}\\[0.75cm]
    \makebox[\textwidth]{\large Student number: 1010907823\enspace\hrulefill}\\
  \end{minipage}
}

\bigskip
\large \textbf{I confirm that:}
\normalsize

\begin{itemize} 
\item I have read and followed the policies described in the document {\bf
    MAT185 Assignment Policies \& FAQ}.
\item In particular, I have read and understand the rules for
  collaboration, and permitted resources on assignments as described in
  subsection II of the the aforementioned document. I have not violated
  these rules while completing and writing this assignment.
\item I have not used generative AI in writing this assignment.
\item I understand the consequences of violating the University's academic
  integrity policies as outlined in the
  \href{http://www.governingcouncil.utoronto.ca/policies/behaveac.htm}{Code of
    Behaviour on Academic Matters}. I have not violated them while completing
  and writing this assignment.
\end{itemize}
\bigskip

\large \textbf{By submitting this assignment to Gradescope, I agree that the
  statements above are true.}  \normalsize

\newpage

\begin{questions}

   \question

   Consider the real vector space $V = \text{span}\{x^2 e^x, x e^x, e^x\}$.  \textit{No justification needed for parts (a)--(d); just fill in the boxes. :)}

    \begin{parts}
     \part Let $T: V \to V$ be the linear transformation corresponding to
     differentiation: $T(v) = v'$.  Using the basis
     $\alpha = \{x^2 e^x, x e^x, e^x\}$ for the domain and the codomain, what
     matrix $[T]_{\alpha \alpha}$ represents this linear transformation?

     %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
     %% node[midway,blue]{YOUR ANSWER}
     %%
     %% for example the following has 1, 2, 3 in the three boxes.
     %% \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1} ;& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{3}; \\
     %%
     $$[T]_{\alpha \alpha} = \left[\begin{array}{ccc}
                                     \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
                                     \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
                                     \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
          \end{array}\right]$$

        \part Demonstrate that your matrix is correct by computing
        $ [T]_{\alpha \alpha} [v]_\alpha$ for $v =  3 x^2 e^x - 2 x e^x + 6 e^x$ to find 
        $[T(v)]_\alpha =[ 3 x^2 e^x + 4 x e^x + 4 e^x ]_\alpha$.

        %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
        %% node[midway,blue]{YOUR ANSWER}
        %%

        $$[T]_{\alpha \alpha} [v]_\alpha = \left[\begin{array}{ccc}
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
\end{array}\right]  \left[\begin{array}{c}
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{3}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-2}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{6}; \\
          \end{array}\right] = 
           \left[\begin{array}{c}
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{3}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{4}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{4}; \\
          \end{array}\right] = [T(v)]_\alpha$$

        \part What is $[T]_{\alpha \alpha}^{-1}$ for your matrix?  
%        In the
%        language of Calculus, what linear transformation should this matrix
%        represent?

        %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
        %% node[midway,blue]{YOUR ANSWER}
        %% 
        $$[T]_{\alpha \alpha}^{-1} = \left[\begin{array}{ccc}
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
         \end{array}\right]$$

        %Do not change the height of this box. Your work must fit inside it.
        \makenonemptybox{1.0cm}{
          %Add your explanations here!

          %end of your answer    
        }
        \vspace{-1cm}

        \part Using your $[T]_{\alpha \alpha}^{-1}$, find $v \in V$ such that $v'(x) = x^2 e^x$.

        %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
        %% node[midway,blue]{YOUR ANSWER}
        %% 
        $$[T]_{\alpha \alpha}^{-1} [v]_\alpha = \left[\begin{array}{ccc}
           \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
           \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
           \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-1};&\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
          \end{array}\right]  \left[\begin{array}{c}
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \end{array}\right] = 
           \left[\begin{array}{c}
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-2}; \\
\tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2}; \\
          \end{array}\right] = [T^{-1}(v)]_\alpha$$
       $v(x)$= \makeframednonemptybox{1cm}{6cm}{$x^2e^x - 2xe^x + 2e^x$}

    \part You've learnt that a function has infinitely many antiderivatives.
    Why is it that this method has found only one of them?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{2.5cm}{
      %%% Add your explanations here!      
      In general, we have:
      $$\int f(x) dx = F(x) + c$$ 
      ,where $F(x)$ is the antiderivative of $f(x)$. \\
      The linear transformation $T$ maps to the special case $c=0$. The constant function is not spanned by the bases of vector space $V$, so the antiderivative found is unique with $c=0$. In general, linear transformation that represent the antiderivative cannot have non-zero constant such that the additivity property of the linear transformation is preserved. (i. e. $T(f+g) = F + G + c = T(f) + T(g) = F + G + 2c$)
      %%% end of your answer    
    }

  \end{parts}

  \question

  \begin{parts}
    \part What is a parametrized curve that represents the line segment that
    connects $(x_0,y_0)$ to $(x_1,y_1)$?  That is, what is an $\ell(t)$ such
    that $\ell(0) = (x_0,y_0)$, $\ell(1) = (x_1,y_1)$, and
    $\{ \ell(t) \mid 0 \leq t \leq 1 \}$ is a line segment?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{2cm}{
      %%% Add your explanations here!      
      $$\ell(t) = (x(t), y(t)) = (1-t)(x_0, y_0) + t(x_1, y_1),\, 0 \leq t \leq 1$$
      Putting into standard form, we have:
      $$\ell(t) = (x_0, y_0) + t(x_1 - x_0, y_1 - y_0)$$
    }

    \part Consider the linear transformation $T: \R^2 \to \R^2$ where
    $T((x,y)) = (2x+3y,-6x+y)$.  What is the image of the above line segment
    $\{ \ell(t) \mid 0 \leq t \leq 1 \}$ under this linear transformation?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{3cm}{
      %%% Add your explanations here!      
      For $0 \leq t \leq 1$, we have:
      $$T(\ell(t)) = (2x(t) + 3y(t), -6x(t) + y(t)) = (2(1-t)x_0 + 2tx_1 + 3(1-t)y_0 + 3ty_1, -6(1-t)x_0 - 6tx_1 + (1-t)y_0 + ty_1)$$
      Simplifying, we have the line segment connecting $(2x_0 + 3y_0, -6x_0 + y_0)$ to $(2x_1 + 3y_1, -6x_1 + y_1)$:
      $$T(\ell(t)) = ((1-t)(2x_0 + 3y_0) + t(2x_1 + 3y_1), (1-t)(-6x_0 + y_0) + t(-6x_1 + y_1))$$
      Putting into standard form, we have:
      $$T(\ell(t)) = (2x_0 + 3y_0, -6x_0 + y_0) + t(2(x_1 - x_0) + 3(y_1 - y_0), -6(x_1 - x_0) + (y_1 - y_0))$$
      %%% end of your answer    
    }

    \part Consider a general linear transformation $T: \R^2 \to \R^2$;  that is, $T((x,y)) = (a_{11} x + a_{12} y, a_{21} x + a_{22} y )$ where $a_{11}, a_{12}, a_{21}, a_{22} \in \R$.  Let 
 $\ell \subseteq \R^2$ be a line in the plane.  In ten words or less, what is the image of $\ell$?  That is, what is $T(\ell)$?

    %%% Do not change the height of this box. Your work must fit inside it.
    \makenonemptybox{1cm}{
      %%% Add your explanations here!      
      $T(\ell)$ is a line or a point in $\R^2$.
      %%% end of your answer    
    }
    \vspace{-1cm}

   \part  \textit{No justification is needed. You can select more than one answer.} Let $P \subseteq \R^2$ be a polygon\footnote{Here are the \href{https://www.geeksforgeeks.org/polygon/}{\underline{types of polygons}} we mean here.  Think about both regular and irregular ones, please.} and consider a general linear transformation as in part (c).  The image of $P$ under
   $T$ could be

   \vspace{-.5cm}

    \begin{checkboxes}
      \choice a circle
      \correctchoice a point
      \correctchoice a line segment
      \correctchoice a polygon
      \choice something else
      %REPLACE \choice with \correctchoice TO MARK YOUR ANSWER 
    \end{checkboxes}

  \end{parts}

  \newpage

  \question

  In the following, you will consider a sequence of linear transformations.  For
  each linear transformation you will asked to draw the image of a square under
  the linear transformation.
    \textit{No justification needed for parts (a)--(f); just fill in the boxes. :)}

  \begin{parts}

  \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
  $T((x,y)) = (x,x/2+y)$.  In the domain, draw the unit square
  $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
  codomain draw the image of the unit square $T(\mathcal{S})$.

  \begin{center}

    \begin{minipage}{.15\textwidth}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[axis x line=center,axis line style = thick,axis y
            line=center,axis
            equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \end{minipage}
    \hspace{2cm}
    \begin{minipage}{.3\textwidth}
      \begin{tikz}
        \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
      \end{tikz}

      \vspace{2cm}
    \end{minipage}
    \hspace{-4.5cm}
    \begin{minipage}{.35\textwidth}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[axis x line=center,axis line style = thick,axis y
            line=center,axis
            equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0.5) (1,1.5) (0,1)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \end{minipage}
  \end{center}
  What is the area of $\mathcal{S}$?  What is the area of its image
  $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is the
  matrix $[T]_{\alpha \alpha}$ that represents the linear transformation $T$?
  What is $\mbox{det} [T]_{\alpha \alpha}$?

  \begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{$\frac{1}{2}$};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{1
    }
   \end{minipage}

   \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
   $T((x,y)) = (2x,2y)$.  In the domain, draw the unit square
   $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
   codomain draw the image of the unit square $T(\mathcal{S})$.

   \begin{center}

     \begin{minipage}{.15\textwidth}
       \begin{center}
         \begin{tikzpicture}
           \begin{axis}[axis x line=center,axis line style = thick,axis y
             line=center,axis
             equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
           \end{axis}
         \end{tikzpicture}
       \end{center}
     \end{minipage}
     \hspace{2cm}
     \begin{minipage}{.3\textwidth}
       \begin{tikz}
         \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
       \end{tikz}

       \vspace{2cm}
     \end{minipage}
     \hspace{-4.5cm}
     \begin{minipage}{.35\textwidth}
       \begin{center}
         \begin{tikzpicture}
           \begin{axis}[axis x line=center,axis line style = thick,axis y
             line=center,axis
             equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (2,0) (2,2) (0,2)};
           \end{axis}
         \end{tikzpicture}
       \end{center}
     \end{minipage}
   \end{center}
   What is the area of $\mathcal{S}$?  What is the area of its image
   $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is the
   matrix $[T]_{\alpha \alpha}$ that represents the linear transformation $T$?
   What is $\mbox{det} [T]_{\alpha \alpha}$?

\begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{4
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{4
    }
   \end{minipage}

   \newpage

   \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
   $T((x,y)) = (2x,y/2)$.  In the domain, draw the unit square
   $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
   codomain draw the image of the unit square $T(\mathcal{S})$.

   \begin{center}

     \begin{minipage}{.15\textwidth}
       \begin{center}
         \begin{tikzpicture}
           \begin{axis}[axis x line=center,axis line style = thick,axis y
             line=center,axis
             equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
           \end{axis}
         \end{tikzpicture}
       \end{center}
     \end{minipage}
     \hspace{2cm}
     \begin{minipage}{.3\textwidth}
       \begin{tikz}
         \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
       \end{tikz}

       \vspace{2cm}
     \end{minipage}
     \hspace{-4.5cm}
     \begin{minipage}{.35\textwidth}
       \begin{center}
         \begin{tikzpicture}
           \begin{axis}[axis x line=center,axis line style = thick,axis y
             line=center,axis
             equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
            \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (2,0) (2,0.5) (0,0.5)};
           \end{axis}
         \end{tikzpicture}
       \end{center}
     \end{minipage}
   \end{center}
   What is the area of $\mathcal{S}$?  What is the area of its image
   $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is the
   matrix $[T]_{\alpha \alpha}$ that represents the linear transformation $T$?
   What is $\mbox{det} [T]_{\alpha \alpha}$?

\begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{2};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{$\frac{1}{2}$}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{1
    }
   \end{minipage}

      \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
      $T((x,y)) = (-y,x)$.  In the domain, draw the unit square
      $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
      codomain draw the image of the unit square $T(\mathcal{S})$.

      \begin{center}

        \begin{minipage}{.15\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
              \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
        \hspace{2cm}
        \begin{minipage}{.3\textwidth}
          \begin{tikz}
            \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
          \end{tikz}

          \vspace{2cm}
        \end{minipage}
        \hspace{-4.5cm}
        \begin{minipage}{.35\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
                \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (-1,0) (-1,1) (0,1)};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
      \end{center}
      What is the area of $\mathcal{S}$?  What is the area of its image
      $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is
      the matrix $[T]_{\alpha \alpha}$ that represents the linear transformation
      $T$?  What is $\mbox{det} [T]_{\alpha \alpha}$?

\begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-1}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{0}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{1}

   \end{minipage}

      \newpage

      \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
      $T((x,y)) = (x+y,x-2y)$.  In the domain, draw the unit square
      $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
      codomain draw the image of the unit square $T(\mathcal{S})$.

      \begin{center}

        \begin{minipage}{.15\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
              \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
        \hspace{2cm}
        \begin{minipage}{.3\textwidth}
          \begin{tikz}
            \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
          \end{tikz}

          \vspace{2cm}
        \end{minipage}
        \hspace{-4.5cm}
        \begin{minipage}{.35\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
              \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,-2) (2, -1) (1,1)};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
      \end{center}
      What is the area of $\mathcal{S}$?  What is the area of its image
      $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is
      the matrix $[T]_{\alpha \alpha}$ that represents the linear transformation
      $T$?  What is $\mbox{det} [T]_{\alpha \alpha}$?

\begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{3
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-2}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{-3
    }
   \end{minipage}

      \part Consider the linear transformation $T: \R^2 \to \R^2$ given by
      $T((x,y)) = (-x-y,-x/2-y/2)$.  In the domain, draw the unit square
      $\mathcal{S} = \{ (x,y) \mid 0 \leq x \leq 1, 0 \leq y \leq 1 \}$.  In the
      codomain draw the image of the unit square $T(\mathcal{S})$.

      \begin{center}

        \begin{minipage}{.15\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
              \addplot[fill=blue,opacity=0.7] coordinates {(0,0) (1,0) (1,1) (0,1)};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
        \hspace{2cm}
        \begin{minipage}{.3\textwidth}
          \begin{tikz}
            \draw[-latex,line width=2pt] (1,0) to[out=25,in=155] node[above]{$T$} (2,0);
          \end{tikz}

          \vspace{2cm}
        \end{minipage}
        \hspace{-4.5cm}
        \begin{minipage}{.35\textwidth}
          \begin{center}
            \begin{tikzpicture}
              \begin{axis}[axis x line=center,axis line style = thick,axis y
                line=center,axis
                equal,xmin=-2.5,ymin=-2.5,xmax=2.5,ymax=2.5,xtick={-2,-1,...,2},ytick={-2,-1,...,2},grid=both,xlabel={$x$},ylabel={$y$},width=6cm,height=6cm]
                %\addplot [mark=none,very thick, blue, domain=-2.5:2.5] {1.5*x}; %% USE THIS TO ADD A LINE
                \addplot [mark=none,very thick, blue, domain=-2:0] {0.5*x};
              \end{axis}
            \end{tikzpicture}
          \end{center}
        \end{minipage}
      \end{center}
      What is the area of $\mathcal{S}$?  What is the area of its image
      $T(\mathcal{S})$?  If $\alpha$ is the standard basis for $\R^2$, what is
      the matrix $[T]_{\alpha \alpha}$ that represents the linear transformation
      $T$?  What is $\mbox{det} [T]_{\alpha \alpha}$?

\begin{minipage}{.2\textwidth}
    Area of $\mathcal{S}=$ \makeframednonemptybox{0.5cm}{0.5cm}{1
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    Area of $T(\mathcal{S})=$ \makeframednonemptybox{0.5cm}{0.5cm}{0
    }
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}{.25\textwidth}
    %% TO ADD YOUR SOLUTION TO THE RECTANGLE, ADD THE NEXT LINE BEFORE EACH ';'
    %% node[midway,blue]{YOUR ANSWER}
    %% 
  $[T]_{\alpha \alpha} = \left[\begin{array}{cc}
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-1};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-1}; \\
            \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{$-\frac{1}{2}$};& \tikz\draw[black] (0,0) rectangle (5ex,5ex) node[midway,blue]{-$\frac{1}{2}$}; \\
          \end{array}\right]$
   \end{minipage}
   \hspace{.8cm}
   \begin{minipage}{.2\textwidth}
     $\mbox{det} [T]_{\alpha \alpha}=$\makeframednonemptybox{0.5cm}{0.5cm}{0
    }
   \end{minipage}

      \newpage

\part Continuing with the unit square $\mathcal{S}$, as shown in parts (a)--(f), the area of $T(\mathcal{S})$ and
$\mbox{det} [T]_{\alpha \alpha}$ are related. What is the relationship you observed?  Prove this relationship holds for a general linear transformation $T$.  \textit{See question 2(c) for what we mean by a general linear transformation.  Your answer should not need to refer to the entries of $T$; there should be no $a_{12}$ or the like in your answer.} \label{part:area-det}

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{14cm}{
        %%% Add your explanations here!      
        Area of $T(\mathcal{S}) = \text{Area of }\mathcal{S} \times \mbox{det} [T]_{\alpha \alpha} = \mbox{det} [T]_{\alpha \alpha}$
        Let $V$ be a 2-dimensional vector space and $T: V \to V$ be a general linear transformation for standard basis $\alpha$. We have $\alpha = \{\alpha_1, \alpha_2\} = \{(1,0), (0,1)\}$ hence $\alpha_1$ and $\alpha_2$ encloses the unit square $\mathcal{S}$. Let $v_1, v_2$ be the images of $\alpha_1, \alpha_2$ under $T$ respectively. Then $v_1, v_2$ encloses the image of the unit square $T(\mathcal{S})$ which is a parallelogram. Let $\theta$ be the angle between $v_1$ and $v_2$, we have:
        $$
        P = \text{Area of parallelogram $T(\mathcal{S})$} = |v_1||v_2|\sin\theta
        $$
        Hence, we have:
        $$
        P = |v_1||v_2|\sqrt{1 - \left(\frac{v_1 \cdot v_2}{|v_1||v_2|}\right)^2} = \sqrt{|v_1|^2|v_2|^2 - (v_1 \cdot v_2)^2}
        $$
        Expanding, we have:
        $$
        P = \sqrt{(v_{11}v_{21})^2 + (v_{12}v_{22})^2 + (v_{11}v_{22})^2 + (v_{12}v_{21})^2 - (v_{11}v_{21} + v_{12}v_{22})^2} 
        $$ 
        $$
        = \sqrt{v_{11}^2v_{22}^2 + v_{12}^2v_{21}^2 - 2v_{11}v_{21}v_{12}v_{22}}
        $$
        Simplifying, we have:
        $$
        P = \sqrt{(v_{11}v_{21})^2 + (v_{12}v_{22})^2 - 2v_{11}v_{21}v_{12}v_{22}} = \sqrt{(v_{11}v_{22} - v_{12}v_{21})^2} = |v_{11}v_{22} - v_{12}v_{21}|
        $$
        Let $V = \begin{bmatrix} v_1^\T & v_2^\T \end{bmatrix}$ and $\mathcal{A} = \begin{bmatrix} \alpha_1^\T & \alpha_2^\T \end{bmatrix}$, then $V = \mathcal{A}T$. We have:
        $$
        P = |v_{11}v_{22} - v_{12}v_{21}| = |\text{det}(V)| = |\text{det}(\mathcal{A}T) = |\text{det}(\mathcal{A})\text{det}(T)|.
        $$
        Since $A = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix} = I$, we have $\text{det}(\mathcal{A}) = 1$. Therefore, we have:
        $$
        P = \text{det}(T) = \text{det}[T]_{\alpha \alpha}
        $$
        Hence, the area of $T(\mathcal{S})$ is equal to $|\text{det}[T]_{\alpha \alpha}|$.

        %%% end of your answer    
      }

      \part If you started with a square of area 4,
      $\mathcal{S}_1 = \{ (x,y) \mid 0 \leq x \leq 2, 0 \leq y \leq 2 \}$ how
      would you use the determinant to compute the area of $T(\mathcal{S}_1)$
      for a general linear transformation?

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{4cm}{
        I would use the determinant to compute the area of $T(\mathcal{S}_1)$ for a general linear transformation by:
        $$\text{Area of }T(\mathcal{S}_1) = \text{Area of }\mathcal{S}_1 \times \mbox{det} [T]_{\alpha \alpha} = 4\mbox{det} [T]_{\alpha \alpha}$$
        For all parallelogram $\mathcal{S}_r$ enclosed by two vectors in $V$ with area $A_r$, we can find a linear transformation $Q_r: V \to V$ such that $Q_r(\mathcal{S}_0) = \mathcal{S}_r$ where $\mathcal{S}_0$ is the unit square. Therefore, the area of $\mathcal{S}_r$ is $A_r = \mbox{det} [Q]_{\alpha \alpha}$. Hence $T(\mathcal{S}_r) = T(Q(\mathcal{S}_0))$. From part \ref{part:area-det}, we have $T(\mathcal{S}_r) = T(Q(\mathcal{S}_0)) = \mbox{det} [TQ]_{\alpha \alpha} = \mbox{det} [T]_{\alpha \alpha}\mbox{det} [Q]_{\alpha \alpha} = A_r\mbox{det} [T]_{\alpha \alpha}$. Also, since $Q_r(\mathcal{S}_0) = \mathcal{S}_r$ from part \ref{part:area-det}, we have $\mbox{det} [Q]_{\alpha \alpha} = A_r$. Therefore, the area of $T(\mathcal{S}_r)$ is $A_r\mbox{det} [T]_{\alpha \alpha}$.\\
        In this case where $\mathcal{S}_1$ is a square of area 4, the area of $T(\mathcal{S}_1)$ is $4\mbox{det} [T]_{\alpha \alpha}$.

      }

    \end{parts}

    \newpage

    \begin{EnvUplevel}

      \textbf{The following question 4 is worth zero points.  Your work will not be
        graded.  The question generalizes the previous question and is within your
        reach if you are comfortable with dot products and cross products.  It is
        material that you may be assumed to already know when you're in AER210.  We
        encourage you to think about and work on this problem and we are happy to
        talk about it with you, but it is provided ``for the curious'' in that you
        will not be tested on it.}

      \textbf{YOU MUST UPLOAD THIS PAGE EVEN IF YOU WROTE NOTHING ON IT.}

    \end{EnvUplevel}

    \question What about volumes of parallelepipeds?

    \begin{minipage}{.6\textwidth}
      Each face of a parallellepiped is parallel to the opposite face; the
      parallelepiped is determined by the three edges coming out from any one
      corner.  In the image, we have a corner at $(0,0,0)$ and the three edges are
      represented using the vectors in $\ba, \bb, \bc \in {^3}\R$, for example the
      vectors based at the origin $(0,0,0)$.
    \end{minipage}
    \begin{minipage}{.35\textwidth}
      \flushright\begin{tikzpicture}
        \draw[-stealth,thick] (0,0) -- (0.7,2) node[pos=0.7,above,rotate=70.7] {\small$\textbf{a}$};
        \draw[-stealth,thick] (0,0) -- (3.5,0) node[pos=0.7,below] {\small$\textbf{b}$};
        \draw[-stealth,thick] (0,0) -- (1.2,0.5) node[pos=0.7,above,rotate=22.6] {\small$\textbf{c}$};
        \draw[dashed,thick] (1.2,0.5) -- (4.7,0.5) -- (3.5,0) -- (4.2,2) -- (0.7,2) -- (1.9,2.5) -- (1.2,0.5);
        \draw[dashed,thick] (1.9,2.5) -- (5.3,2.5) -- (4.7,0.5);
        \draw[dashed,thick] (5.3,2.5) -- (4.2,2);
        \node at (-0.1,-0.1) {\small$\mathbf{0}$};
      \end{tikzpicture}
    \end{minipage}

    To find the volume of a parallelepiped\footnote{A parallelepiped is a solid
      object with corners and edges and an interior.  We can understand it using
      Cartesian coordinates the moment we say where the origin is and where the
      $x$, $y$, and $z$ axes are.  In the picture we're implicitly taking the
      view that the positive $x$ axis is in the same direction as $\bb$, the
      positive $y$ axis is in the same direction as $\bc$, and the positive $z$
      axis is in the same direction as $\ba$.  But we need to be careful to
      remember that edges are collections of points like $(0,0,0)$ and $(1,0,0)$
      and so forth while the vectors $\ba$, $\bb$, and $\bc$ are in $^3\R$.},
    you need to find the area of one face and the distance of the opposite face
    to that one face.  For example, if you found the area of the bottom face
    (the one determined by $\bb$ and $\bc$) you would need to find the height of
    the parallelepiped.  You would then multiply the two areas.

    \begin{parts}
      \part Using the above figure, explain why the volume of the parallelipiped
      is $| \ \ba \cdot ( \bb \times \bc ) \ |$.  Explain how this is related to
      the determinant of the matrix whose rows are given by $\ba$, $\bb$, and
      $\bc$:
      $$
      \det \begin{bmatrix} \ba^\T \\ \bb^\T \\ \bc^\T \end{bmatrix}.
      $$

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{8cm}{
        %%% Add your explanations here!

        %%% end of your answer    
      }

      \newpage

          \begin{EnvUplevel}

      \textbf{YOU MUST UPLOAD THIS PAGE EVEN IF YOU WROTE NOTHING ON IT.}

    \end{EnvUplevel}

      \part Explain why you would have gotten the same volume if you had based your
      calculation on the front of the parallelepiped (determined by $\ba$ and $\bb$)
      or if you had based your calculation on the side of the parallelepiped
      (determined by $\ba$ and $\bc$).
      $$
      \mbox{volume of parallelepiped} = \det \begin{bmatrix} \bc^\T \\ \ba^\T \\
        \bb^\T \end{bmatrix} = \det \begin{bmatrix} \bb^\T \\ \bc^\T \\
        \ba^\T \end{bmatrix}.
      $$

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{18cm}{
        %%% Add your explanations here!

        %%% end of your answer    
      }

      \newpage

                \begin{EnvUplevel}

      \textbf{YOU MUST UPLOAD THIS PAGE EVEN IF YOU WROTE NOTHING ON IT.}

    \end{EnvUplevel}

      \part Consider the unit cube
      $\mathcal{C} = \{ (x,y,z) \mid 0 \leq x \leq 1, 0 \leq y \leq 1, 0 \leq z \leq
      1\}$.  Let $T: \R^3 \to \R^3$ be a linear transformation.  What is the image
      of $\mathcal{C}$?  (What is $T(\mathcal{C})$ geometrically?)  If $\alpha$ is
      the standard basis for $\R^3$ and $[T]_{\alpha \alpha}$ is the matrix that
      represents the linear transformation, what is $\det [T]_{\alpha \alpha}$ and
      how does it relate to the volume of $T(\mathcal{C})$?

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{14cm}{
        %%% Add your explanations here!

        %%% end of your answer    
      }

      \part If you started with a cube of volume 8,
      $\mathcal{C}_1 = \{ (x,y,z) \mid 0 \leq x \leq 2, 0 \leq y \leq 2, 0 \leq z
      \leq 2\}$, how would you use the determinant to compute the volume of
      $T(\mathcal{C}_1)$?

      %%% Do not change the height of this box. Your work must fit inside it.
      \makenonemptybox{3.0cm}{
        %%% Add your explanations here!

        %%% end of your answer    
      }

    \end{parts}

  \end{questions}

\end{document}

### 2025-09-01 YEAR2/AER210_LectureNotes/main.tex
pre year2

--- 2025-03-09 MSE160 Lecture Notes/main.tex
@@
-\documentclass[11pt]{article}
+\documentclass[11pt]{report}
 \input{preamble}
@@
 \thispagestyle{empty}
-{\LARGE \bf MSE 160 Lecture Notes}\\
+{\LARGE \bf AER 210 Lecture Notes}\\
 {\large Hei Shing Cheung}\\
-Molecules and Materials, Winter 2024 \hfill MSE 160\\
+Vector Calculus \& Fluid Mechanics, Fall 2025 \hfill AER210\\
 \\
 The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
-\begin{center}
-\textit{"In this class we are mostly understanding solids''} \\ - Prof. \textsc{Scott Ramsay}
-\end{center}
-% TODO: add a to r ratio table for different crystal structures/cubic structures
-\vspace{10pt}
-\section{Mechanical Behavior}
-\paragraph{Classes of Materials} In this class, we look at three classes of materials (non-exhaustive):
-\begin{itemize}
-    \item \textbf{Metal} held together with metallic bonds, typically \textbf{ductile} and \textbf{conductive}.
-    \item \textbf{Ceramics}  (often metal oxides [excp: diamond]) held together via covalent \& ionic bonds, typically \textbf{brittle} and \textbf{insulating}.
-    \item \textbf{Polymers} Molecules (often hydrocarbons) typically \textbf{ductile} and \textbf{insulating}
-\end{itemize}
-\paragraph{Engineering Stress} For normal stress, we know that:
-\begin{equation}
-    \sigma = \frac{F}{A_0}
-\end{equation}
-\paragraph{Engineering Strein} Also:
-\begin{equation}
-    \epsilon = \frac{\Delta l}{l_0}
-\end{equation}
-\paragraph{Young's Moduclus} For elastic deformation, $E$, is given, by Hooke's Law, as follows:
-\begin{equation}
-    \sigma = E \epsilon
-\end{equation}
-\paragraph{Tensile Test} We apply force as to the ends of a dogbone-sample, with $l_0$ being the gauge length and $A_0$ being the area of the cross-section at the middle. 
-\paragraph{Tensile Strein} Maximum tensile strain on the engineeging stress-strain curve.
-\subsection{Understanding Elastic Properties in terms of Atomic Configuration}
-\paragraph{Atomic Configuration} We can understand the elastic properties of a material by looking at the atomic configuration.
-Skemetically, we can represent the atomic configuration as a spring system:
-\begin{enumerate}
-    \item \textbf{Intial - Before Loading} Atoms are in equilibrium, with the interatomic forces being balanced.
-    \item \textbf{Loading} We apply a force to the material, causing the atoms to move from their equilibrium positions. The bond stretches and the atoms move further apart.
-    \item \textbf{Unloading} We remove the force, causing the atoms to return to their equilibrium positions. 
-\end{enumerate}
-\paragraph{Atom Positions} Elastic modulus is dependent on the atomic interatomic bonding force. Thus, The elastic modulus is prootionaly to the slope of the interatomic force-seperation curve.
-\paragraph{Force-Seperation Curve} The force-seperation curve is a plot of the force between two atoms as a function of the distance between them.
-The slope of the curve is proportional to the elastic modulus near the equilibrium position.
 
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}
-        \begin{axis}[
-            xlabel={Separation Distance (d)},
-            ylabel={Force (F)},
-            grid=major,
-            width=10cm,
-            height=6cm,
-            domain=-0.3:10,
-            samples=100,
-            xtick = \empty,
-            ytick = {0}
-        ]
-        \addplot[blue, thick] {exp(-x/2) - exp(-x)};
-        \node at (axis cs:0.2,0) [anchor=north] {$r_0$};
+\chapter{Vector Calculus}
 
-    \end{axis}
-    \end{tikzpicture}
-    \caption{Force-Separation Curve (Lennard-Jones Force)}
-    \label{fig:force-separation}
-\end{figure}
-\begin{equation}
-    E \propto \frac{dF}{dr} \Bigg|_{r_0}
-\end{equation}
-\begin{definition}[Equilibrium interatomic seperation distance]
-The equilibrium interatomic seperation distance, $r_0$, is the distance between two atoms at which the interatomic force is zero.
-This is due to the interatomic forces being the sum of attractive and repulsive forces.
-\end{definition}
-\paragraph{Elastic Modulus} Thus, strongly bonded materials have a higher elastic modulus and the slope of the force-seperation curve is steeper at $r_0$.
-\subsection{Understanding Other Properties in terms of Atomic Configuration}
-\paragraph{Potential Energy-Separation Curve} The potential energy-seperation curve is a plot of the potential energy between two atoms as a function of the distance between them.
-The potential energy is the area under the force-seperation curve.
-\paragraph{Depth of the Minimum Energy Well} The depth of the minimum energy well, $E_0$, is the energy required to break the bond between two atoms.
-This is the energy required to move the atoms from the equilibrium position to infinity.
-It is proportional to the melting temperature of the material.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}
-        \begin{axis}[
-            xlabel={Separation Distance (d)},
-            ylabel={Potential Energy (U)},
-            grid=major,
-            width=10cm,
-            height=6cm,
-            domain=-0.3:10,
-            samples=100,
-            xtick = \empty,
-            ytick = {0}
-        ]
-        \addplot[blue, thick] {-exp(-x/2) + exp(-x)^2};
-        \node at (axis cs:1,-0.3) [anchor=north] {$E_0$};
-
-    \end{axis}
-    \end{tikzpicture}
-    \caption{Potential Energy-Separation Curve}
-    \label{fig:potential-energy}
-\end{figure}
-\paragraph{Coefficient of Thermal Expansion} The coefficient of thermal expansion, $\alpha$, is the fractional change in length per degree change in temperature. 
-\paragraph{Depth of Potential Energy Curve} The deeper the potential energy curve, the higher the melting temperature and more symetric the curve near $E_0$.
-This would give the follwoing three properties:
-\begin{enumerate}
-    \item \textbf{Higher Melting Temperature} The higher the melting temperature, the deeper the potential energy curve.
-    \item \textbf{Higher Elastic Modulus} The steeper the slope of the force-seperation curve at $r_0$, the higher the elastic modulus.
-    \item \textbf{Lower Coefficient of Thermal Expansion} The more symetric the potential energy curve near $E_0$, the lower the coefficient of thermal expansion.
-\end{enumerate}
-\subsection{Shear and Tensile Stress}
-\subsubsection{Shear}
-\paragraph{Shear Stress} Shear stress is the force per unit area acting parallel to the surface. It is given by:
-\begin{equation}
-    \tau = \frac{F}{A_0}
-\end{equation}
-\paragraph{Shear Strain} Shear strain is the change in angle between two lines originally perpendicular to each other. It is given by:
-\begin{equation}
-    \gamma = \frac{\Delta l}{l_0} \approx \tan \theta \approx \theta = \frac{\pi}{2} - \phi
-\end{equation}
-\paragraph{Shear Modulus} The shear modulus, $G$, is the ratio of shear stress to shear strain. It is given by:
-\begin{equation}
-    \tau = G \gamma
-\end{equation}
-\paragraph{Relationship between Shear and Tensile Modulus} The shear modulus is related to the tensile modulus by the following equation:
-\begin{equation}
-    G = \frac{E}{2(1+\nu)}
-\end{equation}
-where $\nu$ is the Poisson's ratio.
-\paragraph{Poisson's Ratio} Poisson's ratio, $\nu$, is the ratio of lateral strain to axial strain. It is given by:
-\begin{equation}
-    \nu = -\frac{\epsilon_{\text{lat}}}{\epsilon_{\text{axial}}}
-\end{equation}
-\subsection{Testing}
-\begin{definition}[Gauge Length]
-The gauge length, $l_0$, is the length of the sample over which the strain is measured.
-\end{definition}
-\begin{definition}[Reduced Section]
-The reduced section is the part of the sample where the cross-sectional area is reduced to a smaller value.
-\end{definition}
-\paragraph{} Gauge length is always no longer than the reduced section. The reduced section is where the sample will likely break.
-\paragraph{Testing Ceremics} In relation to tensile testing, ceramics have the following properties:
-\begin{itemize}
-    \item \textbf{Brittle} Ceramics are brittle and will break suddenly.
-    \item \textbf{High Strength} Ceramics have high strength and thus difficult to machine the sample.
-    \item \textbf{Sample Alignment} The sample must be aligned properly to test for pure tension. Unlike metals and polymers, which are self-aligning.
-    \item \textbf{Fracture} Ceramics will fracture while still off-axis. Hence, there would be a large shear component. 
-\end{itemize}
-\paragraph{} Thus, we often approxiate tensile behaviour with a point loading on a horizontal beam, with two point support (3 point bending test). Peak stress is given by:
-\begin{equation}
-    \sigma_{\text{peak}} = \frac{3FL}{2bd^2}
-\end{equation}
-, where:
-\begin{itemize}
-    \item $L$ (span) is the distance between the two supports.
-    \item $b$ is the width of the sample
-    \item $d$ is the thickness/depth of the sample
-\end{itemize}
-\section{Seleection of Materials}
-\subsection{Material Performance}
-\begin{example}[Aircraft Wing Spar]
-    The aircraft wing spar is beam (loaded in bending) that supports the wing. The spar is made of a material with the objective of minimize mass under the following constraints:
-    \begin{itemize}
-        \item \textbf{Deflection} There is a maximum allowable deflection of the wing.
-        \item There is more..., but for this example, we will only consider the deflection.
-    \end{itemize}
-    The material selection solve for a \textbf{light stiff beam}.
-
-    \paragraph{Mass} The mass of the beam is given by:
-    $$ m = \rho V = \rho A L $$
-    \paragraph{Deflection} The deflection of the beam is given by:
-    \begin{equation}
-        \delta = \frac{FL^3}{48EI} 
-    \end{equation}
-    \begin{align*} 
-        \intertext{For a beam with a rectangular cross-section, we have:}
-        \delta &= \frac{FL^3}{48E} \cdot \frac{12}{bh^3} = \frac{FL^3}{4Ebh^3}  
-        \intertext{We can set $b$ proportional to $h$:}
-        \delta &= \frac{FL^3}{cE} \cdot \frac{1}{A^2},\quad \text{for some constant $c$} 
-        \intertext{We can then isolate for $A$, the free variable, and minimize the mass via the objective equation $m = \rho A L$:}
-        A &= \sqrt{\frac{FL^3}{cE\delta}} \\
-        m &= \rho L \sqrt{\frac{FL^3}{cE\delta}} = \rho L \sqrt{\frac{FL^3}{cE\delta}}
-        \intertext{Arrange into the form (functional)(geometric)(material):}
-        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\cdot\left(\frac{\rho}{E^\frac{1}{2}}\right)
-    \end{align*}
-\end{example}
-\paragraph{Material Peformance Index} The material performance index is given by:
-    \begin{equation}
-        \text{Material Performance Index (MSI)} = \frac{E^\frac{1}{2}}{\rho}
-\end{equation}
-\paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
-\paragraph{Tempered Glass} Temperglass are made to resist tension.
-It is done by applying a compressive stress to the surface of the glass.
-This is done by cooling the surface of the glass faster than the core or chemically treating the surface.
-\subsection{Density}
-\paragraph{Density} The density of a material is given by:
-\begin{equation}
-    \rho = \frac{m}{V}
-\end{equation}
-\paragraph{Archimedes' Principle} The buoyant force on an object is equal to the weight of the fluid displaced by the object. Derive from that, we have:
-\begin{equation}
-    \rho = \frac{m}{V} = \frac{m_{\text{object}}}{m_{\text{object}} - m_{\text{object in fluid}}}
-\end{equation}
-\section{Atomic Structures}
-\subsection{General Definitions}
-\paragraph{Ordered structures} We have the following three orders:
-\begin{enumerate}
-    \item \textbf{Long Range Order (LRO)} Atoms are arranged in a well-defined pattern over long distances in repeating units. \\
-    (Example) Diamonds, some polymers, most metals, many ceramics, graphite, quartz, etc.
-    \item \textbf{Short Range Order (SRO)} Periodic arrangement of atoms over a few atomic or molecular spacings. \\
-    (Example) Most polymers, glasses, amorphous materials, etc.
-    \item \textbf{No Order (NO)} Atoms are randomly arranged. \\
-    (Example) Ideal gases, etc.
-\end{enumerate}
-\paragraph{Describing Crystal Structures} We can describe crystal structures by the following:
-\begin{itemize}
-    \item \textbf{Unit Cell} The smallest repeating unit of a crystal structure that could be used to represent the entire crystal.
-    \item \textbf{Lattice} The repeating arrangement of points that repersent the positions of atoms in the unit cell.\\
-    (Example) Simple Cubic, Body-Centered Cubic, Face-Centered Cubic, etc.
-    \item \textbf{Lattice Plus Basis} \textit{Not required for this course}
-\end{itemize}
-\paragraph{Theortical Density of Metals} The theoretical density of metals is given by:
-\begin{equation} \label{eq:theoretical-density}
-    \rho = \frac{nM}{V_c N_A}
-\end{equation}
-, where:
-\begin{itemize}
-    \item $n$ is the number of atoms in the unit cell.
-    \item $M$ is the atomic mass of the element (amu = g/mol).
-    \item $V_c$ is the volume of the unit cell.
-    \item $N_A$ is Avogadro's number ($6.022 \times 10^{23} \text{mol}^{-1}$)
-\end{itemize}
-\subsection{Describing Basic Unit Cells}
-\begin{definition}[Coordination Number (CN)]
-    The coordination number is the number of nearest neighbors an atom has in a crystal structure.
-\end{definition}
-\begin{definition}[Atomic Packing Factor (APF)]
-    The atomic packing factor is the fraction of the volume of the unit cell that is occupied by hard spheres. It is given by:
-    \begin{equation}
-        \text{APF} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell}}
-    \end{equation}
-\end{definition}
-\paragraph{Simple Cubic (SC)} The simple cubic structure has the following properties:
-\begin{itemize}
-    \item \textbf{Centers of atoms} Located at the eight corners of a cube
-    \item \textbf{Rare Packing Density}  Due to low packing density (only known example: Polonium, Po)
-    \item \textbf{Close-packed directions} As cube edges.
-    \item \textbf{Coordination Number} Number of nearest neighbor: 6
-\end{itemize}
-\paragraph{Face-Centered Cubic (FCC)} The face-centered cubic structure has an atom on the centre of each face of the cube. It has the following properties:
-\begin{itemize}
-    \item \textbf{Ductile} and \textbf{Malleable} due to the close-packed planes. Common examples include: \\ $\ch{Al}\, ,\ch{Cu}\, ,\ch{Au}\, ,\ch{Ag}$.
-    \item \textbf{Coordination Number} Number of nearest neighbor: 12
-    \item \textbf{Close-packed directions} As cube diagonals.
-    \item \textbf{Theortical Density} The theoretical density of FCC is derived from Equation (\ref{eq:theoretical-density}):
-    \begin{align}
-        \rho &= \frac{nM}{V_c N_A} \nonumber \\
-        \intertext{With $n = 4$ (8 corners $\times \frac{1}{8}$ atom each + 6 faces $\times \frac{1}{2}$ atom each) and $V_c = a^3$ (where $a$ is the \textbf{lattice parameter}):}
-        \rho_\text{FCC} &= \frac{4M}{a^3 N_A} \nonumber \\
-        \intertext{We can then obtain $a$ via the relationship between the atomic radius, $r$, and the lattice parameter via the geometry on the close-packed direction (face diagonal):}
-        a &= 2\sqrt{2}r \nonumber \\
-        \intertext{Substitute $a$ back into the equation:}
-        \rho_\text{FCC} &= \frac{4M}{(2\sqrt{2}r)^3 N_A} = \frac{4M}{16\sqrt{2}r^3 N_A} \nonumber \\
-        &= \frac{M}{4\sqrt{2}r^3 N_A} 
-    \end{align}
-    \item \textbf{Atomic Packing Factor} The atomic packing factor of FCC is given by:
-    \begin{equation}
-        \text{APF}_\text{FCC} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{4 \times \frac{4}{3}\pi r^3}{a^3} = \frac{4 \times \frac{4}{3}\pi r^3}{(2\sqrt{2}r)^3} = \frac{\pi}{6\sqrt{2}} = \textbf{0.74}
-    \end{equation}
-    This is the highest possible packing factor for spheres. Thus, no other structure can have a higher packing factor.
-    \item \textbf{Stacking} The FCC structure can be thought of as stacking close-packed planes.
-The stacking sequence is ABCABC...
-(where A, B, and C describe the orientation hexagonal close-packed planes).
-\end{itemize}
-\paragraph{Hexagonal Close-Packed (HCP)} The hexagonal close-packed structure is \textbf{crytal structure} by stacking close-packed hexagonal planes. It has the following properties:
-\begin{itemize}
-    \item \textbf{Coordination Number} Number of nearest neighbor: 12
-    \item \textbf{Close-packed directions} As cube diagonals.
-    \item \textbf{Theortical Density} The theoretical density of HCP is derived from Equation (\ref{eq:theoretical-density}):
-    \begin{align}
-        \rho &= \frac{nM}{V_c N_A} \nonumber \\
-        \intertext{With $n = 6$ (12 corners $\times \frac{1}{6}$ atom each) and $V_c = a^3$:}
-        \rho_\text{HCP} &= \frac{6M}{a^3 N_A} 
-    \end{align}
-    \item \textbf{Atomic Packing Factor} The atomic packing factor of HCP is given by:
-    \begin{equation}
-        \text{APF}_\text{HCP} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{6 \times \frac{4}{3}\pi r^3}{a^3} = \frac{6 \times \frac{4}{3}\pi r^3}{a^3} = \textbf{0.74}
-    \end{equation}
-    \item \textbf{Stacking} The HCP structure can be thought of as stacking close-packed planes. The stacking sequence is ABAB...
-\end{itemize}
-\begin{example}[Density of Aluminum]
-    Given that the atomic radius of aluminum is $r = 143$ pm and the atomic mass is $M = 26.98$ g/mol, we can calculate the density of aluminum with FCC structure. We have:
-    \begin{align*}
-        \rho_\text{Al} &= \frac{M}{4\sqrt{2}r^3 N_A} = \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} \\
-        &= \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} = 2.7 \text{ g/cm}^3
-    \end{align*}
-\end{example}
-\paragraph{Body-Centered Cubic (BCC)} The body-centered cubic structure has an atom at the centre of the cube. It has the following properties:
-\begin{enumerate}
-    \item \textbf{Coordination Number} Number of nearest neighbor: 8
-    \item \textbf{Atomic Packing Factor} The atomic packing factor of BCC is given by:
-    \begin{equation}
-        \text{APF}_\text{BCC} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{2 \times \frac{4}{3}\pi r^3}{a^3} = \frac{2 \times \frac{4}{3}\pi r^3}{(4r/\sqrt{3})^3} = \textbf{0.68}
-    \end{equation}
-    \item \textbf{Close Packed Directions} As cube diagonals.
-    \item \textbf{Theortical Density} The theoretical density of BCC is derived from Equation (\ref{eq:theoretical-density}):
-    \begin{align}
-        \rho &= \frac{nM}{V_c N_A} \nonumber \\
-        \intertext{With $n = 2$ (8 corners $\times \frac{1}{8}$ atom each + 1 center atom) and $V_c = a^3$:}
-        \rho_\text{BCC} &= \frac{2M}{a^3 N_A} 
-    \end{align}
-    \item \textbf{Stacking} The BCC structure can be thought of as stacking close-packed planes. The stacking sequence is ABAB...
-\end{enumerate}
-\section{Geometric Properties of Crystals}
-\subsection{Crystallographic Directions and Planes}
-\paragraph{Point Coordinates} A lattice positon in a unit cell is determined as fractional multiples of the unit cell edge lengths $a$, $b$, and $c$.
-(Defining the origin at one corner of the unit cell and $a=b=c=\textbf{1}$)
-\paragraph{Crystallographic Directions} A crystallographic direction is a line between two points in a crystal.
-It is denoted by square brackets, $[uvw]$.
-The direction is determined from the ``tail'' to the ``head'' of the vector.
-Also:
-\begin{enumerate}
-    \item It \textbf{shall not} include commas; and
-    \item It \textbf{shall be reduced}to the smallest integers; and
-    \item Negative signs are represented by a bar over the number.
-\end{enumerate}
-\begin{example}
-    The crystallographic direction from $(0,0,0)$ to $(-1,0,1/2)$ is $[\bar{2}01]$.
-\end{example}
-\paragraph{Family of Directions} A family of directions is a set of directions that are parallel to each other. It is denoted by $<uvw>$.
-\begin{example}
-    The family of directions $<100>$ is the set of directions $[100]$, $[010]$, $[001]$, $[\bar{1}00]$, $[0\bar{1}0]$, and $[00\bar{1}]$.
-\end{example}
-\paragraph{Angle between Directions} The angle between two directions in a \textbf{cubic} crystal is given by:
-\begin{equation}
-    \cos \theta = \frac{h_1h_2 + k_1k_2 + l_1l_2}{\sqrt{h_1^2 + k_1^2 + l_1^2}\sqrt{h_2^2 + k_2^2 + l_2^2}}
-\end{equation}
-, which is simply the normalized dot product of the two vectors.
-\paragraph{Crystallographic Planes} A crystallographic plane is a set of parallel planes in a crystal.
-It is denoted by $(hkl)$.
-The plane is determined by the intercepts on the $x$, $y$, and $z$ axes.
-\paragraph{} Generally, we find the intercepts by finding the points where the plane intersects the axes.
-We then take the reciprocals of the intercepts and multiply by the smallest integer to get the Miller Indices.
-\paragraph{Family of Planes} A family of planes is a set of planes that are parallel to each other. It is denoted by $\{hkl\}$.
-\paragraph{Distance between Planes} The distance between two planes with miller indices $(h_1k_1l_1)$ and $(h_2k_2l_2)$ is given by:
-\begin{equation}
-    d = \frac{a}{\sqrt{h^2 + k^2 + l^2}}
-\end{equation}
-, where $a$ is the lattice parameter.
-\subsection{X-Ray Diffraction} The diffraction pattern is used to determine the crystal structure.
-By measuring the n$^\text{th}$ order peak angle $\theta_c$ and the wavelength of the X-ray, we can determine the distance between planes:
-\begin{align}
-    2d\sin \theta &= n\lambda \nonumber \\
-    d &= \frac{n\lambda}{2\sin \theta_c}
-    \intertext{Thus, we could determine the lattice parameter $a$ by the following equation:}
-    a &= \frac{n\lambda}{2\sin \theta_c} \sqrt{h^2 + k^2 + l^2}
-\end{align}
-\subsection{Geometrcally Ideal Crystal Structures}
-\paragraph{Coordination Number and Ratio} The coordination number increases with $\frac{r_{\text{cation}}}{r_{\text{anion}}}$ (The ionic radii ratio).  It relates to coordination number as follows:
-\begin{table}[h]
-    \centering
-    \begin{tabular}{|c|c|c|}
-        \hline
-        \textbf{Ionic Radii Ratio} & \textbf{Coord. Number} & \textbf{Name} \\
-        \hline
-        $<$ 0.155 & 2 & Linear \\
-        0.155 - 0.225 & 3 & Trigonal Planar \\
-        0.225 - 0.414 & 4 & Tetrahedral \\
-        0.414 - 0.732 & 6 & Octahedral \\
-        0.732 - 1.0 & 8 & Cubic \\
-        \hline
-    \end{tabular}
-    \caption{Coordination Number and Ionic Radii}
-    \label{tab:coordination-number}
-\end{table} 
-\begin{definition}[AX Type]
-    The AX type is a crystal structure where the cation and anion are of the same ratio. 
-\end{definition}
-\begin{example}[AX Type: $\ch{NaCl}$ rock salt]
-    The $\ch{NaCl}$ structure is an AX type structure.
-The cation and anion are of the same ratio.
-Soldium (Cation) has radius $r_{\text{Na}} = 0.102$ nm and Chlorine (Anion) has radius $r_{\text{Cl}} = 0.181$ nm.
-The ratio is:
-    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.102}{0.181} = 0.56 $$
-    Thus, the coordination number is 6.\\
-    Additionally, we could calculate its theoretical density. It has 4 cation and 4 anion in the unit cell. The volume of the unit cell is $a^3$, and it follows a FCC structure. Thus, the density is: 
-    \begin{align*}
-        \rho &= \frac{4M_{\text{Na}} + 4M_{\text{Cl}}}{a^3 N_A} = \frac{4(22.99) + 4(35.45)}{a^3 N_A} \\
-        &= \frac{4(22.99) + 4(35.45)}{(2\sqrt{2}r_{\text{Na}})^3 N_A} = \frac{4(22.99) + 4(35.45)}{(2\sqrt{2} \times 0.102)^3 N_A} \\
-        &= \frac{4(22.99) + 4(35.45)}{(2\sqrt{2} \times 0.102)^3 N_A} = 2.16 \text{ g/cm}^3
-    \end{align*}
-\end{example}
-\begin{example}[AX Type: $\ch{ZnS}$, Zinc Blende]
-    The $\ch{ZnS}$ structure is an AX type structure.
-It has a FCC structure with CN = 4.
-The radius of Zinc (Cation) is $r_{\text{Zn}} = 0.074$ nm and the radius of Sulfur (Anion) is $r_{\text{S}} = 0.184$ nm.
-The ratio is:
-    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.074}{0.184} = 0.40 $$
-\end{example}
-\begin{example}[AX$_2$ Type: $\ch{CaF2}$]
-    The $\ch{CaF2}$ structure is an AX$_2$ type structure.
-It has a FCC structure with CN = 8.
-The radius of Calcium (Cation) is $r_{\text{Ca}} = 0.099$ nm and the radius of Fluorine (Anion) is $r_{\text{F}} = 0.133$ nm.
-The ratio is:
-    $$ \frac{r_{\text{cation}}}{r_{\text{anion}}} = \frac{0.099}{0.133} = 0.74 $$
-\end{example}
-\section{Stress-Strain Behavior}
-\paragraph{Uniformly Distributed Elongation} The elongation of a material is uniformly distributed within the gauge length.
-The initial values, referred to as the ``elastic region," are characterized as uniformly elastic.
-\paragraph{Non-Linear Elastic} There is also a non-linear region where the material remains elastic but is not uniformly distributed.
-\paragraph{Localized Plastic Deformation} The material then enters the plastic region where the deformation is localized.
-The material is permanently deformed.
-This is characterized by the ultimate tensile strength (this is bad when it happens).
-\paragraph{Necking} The material then necks, where the cross-sectional area decreases. The material is then strained until it breaks.
-\paragraph{True Stress} The true stress is given by:
-\begin{equation}
-    \sigma_T = \frac{F}{A_i}
-\end{equation}
-, where $A_i$ is the instantaneous cross-sectional area.
-\paragraph{Ultimate Tensile Strength} The ultimate tensile strength is the maximum stress on the stress-strain curve.
-\paragraph{Strengthening} Plastic deformation is made more difficult by the following mechanisms:
-\begin{enumerate}
-    \item \textbf{Strain Hardening} Preloading beyond the elastic region - upon plastic deformation, the material will have a higher yield strength.
-    \item \textbf{For Metals and Alloys} The following mechanisms are used:
-\end{enumerate}
-\paragraph{Strain Hardening Equation} The true stress-strain curve can be emprically described by the following equation:
-\begin{equation}
-    \sigma_T = K \epsilon^n_T
-\end{equation}
-, where $K$ is the Strain Hardening coefficient and $n$ is the strain hardening exponent. Both of these values are material properties.
-\paragraph{True Strain} Since $\epsilon_i = \frac{\Delta l_i}{l_0 + \Delta l_{i-1}}$ we can compute a Rieman sum to get the true strain:
-\begin{equation}
-    \epsilon_T = \int^{l_f}_{l_0} \frac{dl}{l} = \ln \left(\frac{l_f}{l_0}\right)
-\end{equation}
-\paragraph{Strain Rate} The strain rate is the rate at which the material is deformed. It is given by:
-\paragraph{} But, before necking, we can assume $V_0 = V_i$ and $A_i = \frac{A_0 l_0}{l_i}$. Thus, we have:
-\begin{align}
-    \sigma_T &= \frac{F}{A_i} = \frac{F}{\frac{A_0 l_0}{l_i}} = \frac{F l_i}{A_0 l_0} = \frac{F}{A_0} \cdot \frac{l_i}{l_0} = \sigma \cdot \frac{l_i}{l_0}  \nonumber \\
-    &= \sigma \cdot (1 + \epsilon)
-    \intertext{and,}
-    \epsilon_T &= \ln \left(1 + \epsilon\right)
-\end{align}
-\subsection{Imperfections}
-\paragraph{Dislocation} Recal that plastic deformation occurs by step-wise breaking of atomic bonds.
-This is done by the movement of dislocations.
-\textbf{Strengthening mechanisms in Metals} rely on inhibiting the movement of dislocations.
-\begin{definition}[Dislocation]
-    A dislocation is the breaking and reforming of bonds, one row at a time. It is a line defect in the crystal structure.
-\end{definition}
-\paragraph{Crystalline Imperfections} We use crystalline imperfections to inhibit the movement of dislocations. These imperfections include:
-\begin{itemize}
-    \item \textbf{Point Imperfections (Defects)} Vacancies, interstitials, and substitutional atoms.
-    (Example) Solid solution strengthening.
-    \item \textbf{Line Imperfections} Dislocations.
-    \item \textbf{Area Imperfections} Grain boundaries of the interface between two grains and free surfaces.
-    \item \textbf{Volume Imperfections} Second phase particles.
-\end{itemize}
-\paragraph{Solid Solution Strengthening} [Alloying] The addition of a solute to a solvent to inhibit the movement of dislocations. There is two types:
-\begin{enumerate}
-    \item \textbf{Interstitial Impurities} The solute atoms are smaller than the solvent atoms. This the impurites to be in the interstitial sites.
-    \item \textbf{Substitutional Impurities} The solute atoms are larger than or as large as the solvent atoms ($<10 \% \Delta \varnothing$, similar electornegativity.).
-This causes the impurities substitute the lattice sites.
-\end{enumerate}
-\paragraph{Vacancies} Vacancies are missing atoms in the lattice. They are created by thermal vibrations. They are used to inhibit the movement of dislocations.
-It slightly nudges nearby atoms, creating \textbf{lattice strain} that make it harder for dislocations to move.
-\paragraph{Machaism - Diffusion of Impurities} In positve edge dislcation, it has compession above disvlation line, and tension below.
-The impurities will `pin' the dislocation, making it harder to move.
-\paragraph{Plastic Deformation} We can also strengthen the material by increasing the number of dislocations. This is done by plastic deformation methods:
-\begin{enumerate}
-    \item Cold Work
-    \item Forging/Strain Hardening
-\end{enumerate}
-\paragraph{Mechanisms of Strengthening} Plastic deformation increases number of dislocations, which inhibits one another due to the strain fields.
-\begin{definition}[Grain Boundaries]
-    Grain boundaries are the interface between two grains. Grains are regions of the same crystal structure but with different orientations.
-\end{definition}
-\paragraph{Grain Size Reduction} Dislcations tend to move in the direction of the grain boundry (GB).
-Thus, reducing the grain size reduces the distance the dislocation can move.
-In addition, there are atomic strain due to the difference of bond lenght between grains, contributing to the strain field.
-It is given by:
-\begin{equation}
-    \sigma = \sigma_0 + \frac{K_y}{d}
-\end{equation}
-\paragraph{Second Phase Particles} Second phase particles are particles of a different phase in the material.
-They can be used to inhibit the movement of dislocations.
-They are often hard, brittke phases.
-\paragraph{Dislcations (Slips)} Occurs on the plane with the highest planar density.
-The slip direction is the direction with the highest linear density.
-The slip system is the combination of the slip plane and slip direction.
-\paragraph{} FCC has 4 unique planes, each with 3 unique directions. Thus, there are 12 unique slip systems. Compare this to HCP, which has only base plane.
-\paragraph{Resolved Stear Stress} Consider single crystal with a certain plane.
-Let $\phi$ be the angle between the normal of the plane and the direction of the applied stress, and $\lambda$ be the angle between the plane and the orthonal of the applied stress.
-The resolved shear stress is given by:
-\begin{equation}
-    \tau = \sigma \cos \phi \cos \lambda
-\end{equation}
-\paragraph{Critcal Resolved Shear Stress} The critical resolved shear stress is the minimum shear stress required to move a dislocation.
-When the resolved shear stress equal the critical resolved shear stress, the dislocation will move, and:
-\begin{equation}
-    \sigma = \sigma_{y}
-\end{equation}
-So,
-\begin{equation}
-   \tau_{\text{CRSS}} = \sigma_{y} \cos \phi \cos \lambda
-\end{equation}
-\paragraph{Note} Most crystals are polycrystalline, so dislocation must be able to move to the neibouring crystal.
-\paragraph{Burgers Vector} The Burgers vector is the vector that describes the dislocation. It is the vector that closes the loop of the dislocation. It is given by:
-\begin{equation}
-    \vec{b} = \vec{b}_1 + \vec{b}_2
-\end{equation}
-\section{Polymers} 
-\paragraph{Polymer} Polymers could be britte or elastic (elastomers). They are made of long chains of monomers. Interestingly, polymers have necking points lower than its ultimate tensile strength.
-\begin{center}
-    \textit{``Why does polymers continue to stretch after necking?''}
-\end{center}
-\paragraph{Polymer Structure} Polymers are made of long chains of monomers. The chains are held together by weak van der Waals forces. The chains are entangled.
-\\ (Example) Polyethylene are made of long chains of carbon atoms with two hydrogen atoms attached to each carbon atom.
-\paragraph{Ans: (Chain-Oriented)} The chains are originally entangled.
-When the polymer is stretched, the chains are straightened to align with the load.
-More of the load is taken by the covalent bonds, which are stronger than the van der Waals forces.
-This serves as an important strengthening mechanism.
-\paragraph{Note} In a polymer, the elastic response is from numerrous bond types.
-\paragraph{Crylistallinity} Polymers cannot be 100\% crystalline.
-A polymer could be semi-crystalline, where the polymer has both crystalline and amorphous regions.
-Since the crystalline regions are stronger, increasing the crystallinity is a important strengthening mechanism.
-\begin{enumerate}
-    \item \textbf{Tmerpeature} Increasing the crtystallinity permits the polymer have higher service temperature. 
-    \item \textbf{Chemical Resistance} Crystalline regions are more resistant to chemcial dissolution.
-\end{enumerate}
-\paragraph{Different Mer Units} Different mer units have different properties. Listed below are some common mer units:
-\begin{itemize}
-    \item \textbf{Polyethylene} (PE) - Made of long chains of carbon atoms with two hydrogen atoms attached to each carbon atom.
-    \item \textbf{Polypropylene} (PP) - Similar to PE, but with a methyl group attached to the carbon atom.
-    \item \textbf{Polyvinyl Chloride} (PVC) - Made of long chains of carbon atoms with a chlorine atom attached to each carbon atom.
-Chlorine are more electronegative than carbon, so the bond is polar and form stornger van der Waals forces.
-    \item \textbf{Polystyrene} (PS) - Made of long chains of carbon atoms with a phenyl group attached to each carbon atom.
-    \item \textbf{Polytetrafluoroethylene} (PTFE) - Made of long chains of carbon atoms with two fluorine atoms attached to each carbon atom.
-The florine is to protect the carbon backbone.
-Hoowever, it is mechanically weak as the dipole moment is low.
-    \item \textbf{Polymethyl Methacrylate} (PMMA) - Made of long chains of carbon atoms with a methyl group and a methacrylate ($-\ch{OC(=O)}$) group attached to each carbon atom.
-It is opticall tranparnet the methacrylate group prevent crtystallinity, hence amorphous and no crystal boundaries.
-\end{itemize}
-\subsection{Viscoelasitcity and Thermal Properties}
-\paragraph{Viscoelasticity} Polymers are viscoelastic.
-They have both viscous and elastic properties.
-The viscous properties are due to the entanglement of the chains.
-The elastic properties are due to the covalent bonds.
-\paragraph{Time-Dependent Behavior} For a given stress/strain, the strain/stress will decrease with time. This is due to the viscous properties of the polymer.
-\paragraph{Relaxation Modulus} The relaxation modulus is the modulus of the polymer as a function of time. It is given by:
-\begin{equation}
-    E(t) = \frac{\sigma(t)}{\epsilon_0}
-\end{equation}
-The relaxation modulus is also relate to the temperature of the polymer. The relaxation modulus decreases with temperature.
-\paragraph{Glassy Transition} The glassy transition is the temperature at which the polymer changes from a glassy state to a rubbery state.
-The glassy state is characterized by a high modulus and low strain.
-The rubbery state is characterized by a low modulus and high strain.
-The temperature is denoted as $T_g$.
-This is due to the temperature overcoming the intermolecular forces of the amorphous reigions, but the crystalline regions are still intact as the van der Waals foce is stronger.
-\paragraph{Melting} THe melting temperature is the temperature at which the polyme the intermolecular forces of the crystalline regions are overcome.
-The polymer then melts as chains slide past each other.
-\paragraph{Thermosat} Thermosat are polymers that do not melt.
-They are made of cross-linked chains or network polymers.
-The cross-links prevent the chains from sliding past each other.
-Elastomers are typically thermosets.
-\subsection{Optical Properties}
-\paragraph{PMMA} PMMA is a transparent polymer. It is used in optical applications. The transparency is due to the lack of \textbf{scattering events}.
-\begin{definition}[Scatter Event]
-    A scatter event is an event that causes light to be scattered. It is due to the difference in refractive index between the polymer and the air.
-\end{definition}
-\includepdf[pages=-, angle=90]{EquationSheet.pdf}
-
+\chapter{Fluid Mechanics}
 \end{document}

### 2025-09-01 YEAR2/ECE253_LectureNotes/main.tex
pre year2

\documentclass[11pt]{report}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf ECE 253 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Digital and Computer Systems, Fall 2025 \hfill ECE253\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\

\section{Introduction to Digital Logic}

\section{Digital Storage Elements}

\section{Finite State Machines (FSM)}

\section{Computer Organization}

\section{Assembly Language}
\end{document}

### 2025-09-03 YEAR2/CHE260_LectureNotes/main.tex
pre year2
2 sept

--- 2025-09-01 YEAR2/AER210_LectureNotes/main.tex
@@
 \thispagestyle{empty}
-{\LARGE \bf AER 210 Lecture Notes}\\
+{\LARGE \bf CHE 260 Lecture Notes}\\
 {\large Hei Shing Cheung}\\
-Vector Calculus \& Fluid Mechanics, Fall 2025 \hfill AER210\\
+Thermodynamics and Heat Transfer, Fall 2025 \hfill CHE260\\
 \\
@@
 
-\chapter{Vector Calculus}
+\begin{center}
+    \textit{``If there is one word that describes this class it would be \textbf{energy}.''}
+\end{center}
+\chapter{Thermodynamics}
 
-\chapter{Fluid Mechanics}
+\section{Systems and Properties in Thermodynamics}
+\begin{definition}[Energy]
+    You should have learned that energy is the ability to do work.
+\end{definition}
+
+\begin{definition}[Work]
+    You also learned that work is the transfer of energy.
+\end{definition}
+
+We have a problem. The above two definitions are in terms of each other. This touches on the theory of fundamental concepts:
+
+\paragraph{Fundamental Concepts} For example, the following are fundamental concepts in physics:
+\begin{itemize}
+    \item Time
+    \item \textbf{Mass} Interestingly, we don't measure mass directly; instead, we measure weight, which is the force exerted by gravity on an object, and from that we deduce mass.
+    \item Space
+\end{itemize}
+In this course, we are going to explore two:
+\begin{itemize}
+    \item \textbf{Energy} Energy is pretty familiar to us. 
+    \item \textbf{Entropy} Entropy is what gives students the most trouble; you can't show someone a picture of entropy, it is abstract.
+\end{itemize}
+
+These fundamental concepts are in arbitrary units and form the foundation `axioms' in science.
+
+\subsection{Energy}
+\paragraph{Energy} An understanding of energy is its ability to \textbf{lift weights}. This is a way we could test if there is energy.
+\begin{example}[Potential Energy]
+    A 1 kg mass lifted 1 meter has a potential energy of about 10 J. Imagine that it is attached to a balance, with a small mass on the other side, then, it is able to lift that weight.
+\end{example}
+
+\begin{example}[Kinetic Energy]
+    Given a flying ball that hits a lever with a mass rested on it, given the appropriate angle, the ball can transfer its kinetic energy to the mass, causing it to lift.
+\end{example}
+
+\begin{example}[Heat]
+    Given a gas in a container, if we heat the gas, it expands and lifts objects on top of it.
+\end{example}
+\begin{shaded}
+\begin{example}[Newcomen Engine]
+    The above example is actually a simplified representation of how the Newcomen engine operates.
+In the Newcomen engine, steam is used to create a vacuum that lifts a piston, demonstrating the conversion of thermal energy into mechanical work.
+\end{example}
+\end{shaded}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{newcomen_engine.png}
+    \caption{A Newcomen Engine}
+\end{figure}
+When Valve 1 opens, steam fills the chamber and lifts up the piston. When Valve 2 opens (with valve 1 closes),
+water gets sprayed, evaporates, and condenses, causing the piston to lower.
+
+However, the Newcomen engine is not the most efficient steam engine, most heat is lost in the heating and cooling process.
+Solution?
+A external condenser was added to improve efficiency by cooling the steam and creating a vacuum, allowing the piston to be pulled down more effectively.
+This is the \textbf{Watt Engine}.
+
+\paragraph{Steam Cycle} The steam cycle is a thermodynamic cycle that converts heat energy into mechanical work. It consists of four main processes: 
+\begin{enumerate}
+    \item \textbf{Heating} - The working fluid (steam) is heated in the boiler, converting water into steam and increasing its energy.
+    \item \textbf{Expansion} - The high-pressure steam expands in the piston, doing work on the piston and converting thermal energy into mechanical work.
+    \item \textbf{Cooling} - The steam is cooled in the condenser, releasing heat to the surroundings and condensing back into water.
+    \item \textbf{Compression} - The water is pumped back into the boiler, completing the cycle.
+\end{enumerate}
+
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+        \node (boiler) [rectangle, draw] {1. Boiler};
+        \node (turbine) [rectangle, draw, right=of boiler] {2. Turbine};
+        \node (condenser) [rectangle, draw, below=of turbine] {3. Condenser};
+        \node (pump) [rectangle, draw, below=of boiler] {4. Pump};
+
+        \draw[->] (boiler) -- (turbine);
+        \draw[->] (turbine) -- (condenser);
+        \draw[->] (condenser) -- (pump);
+        \draw[->] (pump) -- (boiler);
+    \end{tikzpicture}
+    \caption{Steam Cycle Diagram}
+\end{figure}
+
+\begin{definition}[Heat Engine]
+    A heat engine is a device that converts thermal energy into mechanical work.
+It operates by taking in heat from a high-temperature source, performing work using that heat, and then releasing some waste heat to a low-temperature sink.
+The steam engine is a classical instance of a heat engine.
+\end{definition}
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+        \node (Engine) [circle, draw] {Heat Engine};
+        \node (Source) [rectangle, draw, left=of Engine] {High-Temp Source ($Q_H$)};
+        \node (Sink) [rectangle, draw, right=of Engine] {Low-Temp Sink ($Q_C$)};
+        \draw[->] (Source) -- (Engine);
+        \draw[->] (Engine) -- (Sink);
+
+        \node (Work) [rectangle, draw, below=of Engine] {Work Output ($W$)};
+        \draw[->] (Engine) -- (Work);
+    \end{tikzpicture}
+    \caption{A Heat Engine}
+\end{figure}
+
+\paragraph{Efficiency of a Heat Engine} The efficiency of a heat engine is defined as the ratio of the work output to the heat input:
+
+\begin{equation}
+    \eta = \frac{W}{Q_H}
+\end{equation}
+
+Obviously, the best engine is when $\eta$ is maximized ($W$ is maximized, and $Q_H$ is minimized).
+For example, a Newcomen engine has a low efficiency ($\eta \approx 0.34\%$), while a classic watt engine has a higher efficiency ($\eta \approx 4\%$).
+The modern powerplant is optimized such that it has $\eta \approx 30\%$.
+
+This begs the question, can we create a heat engine with 100\% efficiency (i.e., $Q_C = 0$)?
+The answer is no, due to the second law of thermodynamics [Sadi Carnot (1830)].
+For now, consider the wasted energy $Q_c$.
+
+\paragraph{Irreversibility and Spontaneity} The process of losing heat is spontaneous and irreversible.
+Once heat is lost to the cooler surroundings, it cannot be completely recovered and converted back into work, however, the process of bringing heat from a cooler body to a hotter body is non-spontaneous and requires external work (e.g., a refrigerator).
+These give rise to the concept of irreversibility in thermodynamic processes, from the laws of thermodynamics.
+
+The first law of thermodynamics is simple and easy to understand:
+\begin{definition}[First Law of Thermodynamics]
+    The first law of thermodynamics states that energy cannot be created or destroyed, only transformed from one form to another.
+In the context of a heat engine, this means that the work output ($W$) is equal to the heat input ($Q_H$) minus the heat rejected ($Q_C$):
+    \begin{equation}
+        W = Q_H - Q_C
+    \end{equation}
+\end{definition}
+The second law formalizes the concept of the directionality of thermodynamic processes, stating that heat cannot spontaneously flow from a colder body to a hotter body.
+This law introduces the idea of entropy, a measure of the disorder or randomness in a system, which tends to increase over time in isolated systems.
+
+\begin{definition}[Second Law of Thermodynamics]
+    The second law of thermodynamics states that the total entropy of an isolated system can never decrease over time. We consdier entropy as:
+    \begin{equation}
+        S = \frac{Q}{T}
+    \end{equation}
+    In the above example, we can see that entropy increases as heat is transferred from the hot reservoir to the cold reservoir:
+
+    $$
+        \Delta S = S_{\text{final}} - S_{\text{initial}} = \frac{Q_C}{T_C} - \frac{Q_H}{T_H} > 0
+    $$
+
+    Since $Q_C = Q_H$ in a closed system and $T_H > T_C$, we have $\Delta S > 0$.
+\end{definition}
+
+\subsection{Modeling Ideal Gas}
+\begin{definition}[Ideal Gas]
+    We consider ideal gas conditions as those where the gas particles do not interact with each other except during elastic collisions, and the volume of the gas particles themselves is negligible compared to the volume of the container.
+In this course, we consider 1 atm as the threshold for ideal gas behavior.
+Of course, ideal gas obey the ideal gas law:
+    \begin{equation}
+        PV = nRT
+    \end{equation}
+\end{definition}
+
+\chapter{Heat Transfer}
+
 \end{document}

### 2025-09-03 YEAR2/MAT292_LectureNotes/main.tex
pre year2
2 sept

--- 2025-09-01 YEAR2/ECE253_LectureNotes/main.tex
@@
-\documentclass[11pt]{report}
+\documentclass[11pt]{article}
 \input{preamble}
@@
 \thispagestyle{empty}
-{\LARGE \bf ECE 253 Lecture Notes}\\
+{\LARGE \bf MAT 292 Lecture Notes}\\
 {\large Hei Shing Cheung}\\
-Digital and Computer Systems, Fall 2025 \hfill ECE253\\
+Ordinary Differential Equations, Fall 2025 \hfill MAT292\\
 \\
@@
 
-\section{Introduction to Digital Logic}
+\begin{center}
+    ``\textit{ODEs are the bread and butter of engineering. }''
+\end{center}
+\begin{shaded}
+    \textbf{Key Concepts:}
+    \begin{itemize}
+        \item \textbf{Conflicting Definitions. } This course has multiple definitions for key terms, which may vary between different texts and contexts.
+        \item \textbf{Practice Derivatives. } Regular practice with derivatives is essential for mastering ODEs.
+        \begin{example}
+            The following is a good example of good intuition:
 
-\section{Digital Storage Elements}
+            \textit{What is the antiderivative of $f(x) = \frac{\ln x}{x}$?}
 
-\section{Finite State Machines (FSM)}
+            We know that this is of the form $g \cdot g'$, where $g(x) = \ln x$ and $g'(x) = \frac{1}{x}$. Thus, by the rule that:
 
-\section{Computer Organization}
+            $$
+            \frac{1}{2} \frac{d}{dx} [g(x)]^2 = g(x) g'(x)
+            $$ 
+            we can deduce that"
+            $$
+            \frac{1}{2} \frac{d}{dx} [(\ln x)^2] = \ln x \cdot \frac{1}{x}
+            $$
+        \end{example}
+        \item \textbf{Practice Linear Algebra. } Familiarity with linear algebra concepts is crucial for understanding ODEs.
+    \end{itemize}
+\end{shaded}
+\section{Examples and Review}
+\subsection{What is a Differential Equation?}
+\begin{definition}[Differential Equation]
+     Any relationship between a variable and its derivatives is called a differential equation. 
+\end{definition}
 
-\section{Assembly Language}
+\begin{example}[Newton Second Law]
+    Newton's second law states that the force acting on an object is equal to the mass of the object multiplied by its acceleration. Mathematically, this can be expressed as:
+    $$
+    F = m \frac{d^2x}{dt^2}
+    $$
+    where \( F \) is the force, \( m \) is the mass, and \( \frac{d^2x}{dt^2} \) is the acceleration (the second derivative of position with respect to time).
+This is a second-order ordinary differential equation.
+
+    If force is constant, this is a simple form of a differential equation, which we simply rearrange and integrate it (twice), which gives, simply:
+    $$
+    x(t) = \frac{F}{2m} t^2 + C_1 t + C_0
+    $$
+    where \( C_1 \) and \( C_0 \) are constants determined by initial conditions.
+\end{example}
+
+\begin{example}
+    Consider the following ODE:
+    $$
+    x' = f(t)
+    $$
+    So we have:
+    $$
+    \frac{dx}{dt} = f(t)
+    $$
+    Integrating both sides with respect to \( t \) gives:
+    $$
+    \int_0^t dx = \int_0^t f(t) \, dt
+    $$
+    So we have:
+    $$
+    x(t) = \int_0^t f(t) \, dt + C
+    $$
+    where \( C \) is a constant of integration. In which initial conditions can be used to determine the value of \( C \).  
+\end{example}
+
+\begin{example}[Standard Trick: Turning Higher Order into a System]
+    Consider Hooke's Law:
+    $$
+    F = m\ddot{x} = -kx
+    $$
+    Then, we let:
+    $$
+    \begin{cases}
+        x_1 = x \\
+        x_2 = \dot{x} \\
+        \dot{x}_1 = x_2 \\
+        \dot{x}_2 = -\frac{k}{m} x_1
+    \end{cases}
+    $$
+    This system can be solved using the techniques for first-order ODEs, as a system by its eigenvalues and eigenvectors:
+    $$
+    \begin{pmatrix}
+        \dot{x}_1 \\
+        \dot{x}_2
+    \end{pmatrix}
+    =
+    \begin{pmatrix}
+        0 & 1 \\
+        -\frac{k}{m} & 0
+    \end{pmatrix}
+    \begin{pmatrix}
+        x_1 \\
+        x_2
+    \end{pmatrix}
+    $$
+\end{example}
+
+The above idea scales:
+
+\begin{example}
+    Take $F(t, x, \dot{x}, \ddot{x}, \ldots, x^{(n)}) = 0$. Then, we can let:
+   \begin{equation}
+    \begin{cases}
+        x_1 = x \\
+        x_2 = \dot{x} \\
+        x_3 = \ddot{x} \\
+        \vdots \\
+        x_n = x^{(n)}
+    \end{cases}
+    $$
+    This allows us to rewrite the original equation as a system of first-order equations:
+    $$ 
+    \begin{bmatrix}
+        \dot{x}_1 \\
+        \dot{x}_2 \\
+        \vdots \\
+        \dot{x}_n
+    \end{bmatrix}
+    =
+    A
+    \begin{bmatrix}
+        x_1 \\
+        x_2 \\
+        \vdots \\
+        x_n
+    \end{bmatrix}
+\end{equation}
+
+Where we take $A$ to be the appropriate matrix.
+\end{example}
+
+\begin{example}[Exponential Growth]
+    You should have already know, that:
+    $$
+    \frac{dx}{dt} = kx
+    $$
+    is a first-order linear ODE. The solution to this equation is given by:
+    $$
+    x(t) = Ce^{kt}
+    $$
+    where \( C \) is a constant determined by initial conditions.
+
+We could understand the eigenproblems associated with systems of ODEs as an extension of the above, where we look for solutions of the form:
+\begin{equation}
+\begin{bmatrix}
+x_1(t) \\
+x_2(t) \\
+\vdots \\
+x_n(t)
+\end{bmatrix}
+=
+e^{At}
+\begin{bmatrix}
+x_1(0) \\
+x_2(0) \\
+\vdots \\
+x_n(0)
+\end{bmatrix}
+\end{equation}
+where \( A \) is the matrix associated with the system of ODEs.
+$\exp(At)$ is the matrix exponential of \( At \), which can be computed using various methods, including power series or diagonalization.
+\end{example}
+
+\begin{example}[Superposition]
+    You should also have already known that the solution to:
+    $$ \ddot{x} = -x $$
+    is given by:
+    $$ x(t) = A \cos(t) + B \sin(t) $$
+    where \( A \) and \( B \) are constants determined by initial conditions.
+The solutions $x_1 = A \cos(t)$ and $x_2 = B \sin(t)$ can be combined to form the general solution.
+Thus, additional conditions must be satisfied to determine the values of \( A \) and \( B \).
+\end{example}
+You might have observed that:
+\begin{theorem}[Number of Initial Conditions]
+    You need as many initial conditions as the order of the ODE to uniquely determine a solution.
+\end{theorem}
+
+\begin{example}[Newton's Law of Cooling]
+    Let $u(t)$ be the temperature of the object at time $t$. Then, according to Newton's Law of Cooling, we have:
+    $$
+    \frac{du}{dt} = -k(u - T_a)
+    $$
+    where \( T_a \) is the ambient temperature, and \( k \) is a positive constant (the transmission coefficient).
+This is a first-order linear ODE that can be solved using the techniques discussed earlier.
+
+    \textbf{Fixed-Point Solution} We denote the trivial case where \( u(0) = T_a \) as the fixed-point solution, where the temperature of the object is equal to the ambient temperature at time \( t = 0 \).
+It is easy to see that \( u' = 0 \) in this case so \( u(t) = T_a \) for all \( t \).
+
+    One way of solving the general solution is:
+
+    \begin{align*}
+        \frac{du}{dt} &= -k(u - T_a) \\
+        \int_0^t \frac{du}{u - T_a} &= -k \int_0^t dt \\
+        \ln|u - T_a| &= -kt + C \\
+        u - T_a &= e^C e^{-kt} \\
+        u &= Be^{-kt} + T_a
+    \end{align*}
+
+\end{example}
+
+\begin{definition}[Phase Portraits]
+    A phase portrait is a graphical representation of the trajectories of a dynamical system in the phase plane.
+Each point in the phase plane corresponds to a unique state of the system, and the trajectories represent the evolution of the system over time.
+Phase portraits are useful for visualizing the behavior of systems of ODEs, particularly in understanding stability and equilibrium points.
+\end{definition}
+
 \end{document}

### 2025-09-04 YEAR2/ECE253_LectureNotes/main.tex
2 sept
4 sept

@@
 
+\begin{shaded}
+\section*{Introduction}
+\paragraph{Layers of Computation} In hardware, we have the following layers of abstraction:
+\begin{itemize}
+    \item Computation
+    \item Adders 
+    \item Logic Gates
+    \item Transistors
+    \item Silicon
+\end{itemize}
+In this course, we will focus on the first three layers, on top of the logic gate level.
+
+\paragraph{Layer of abstraction} At this course, for the digital systems part, we would start from understanding logic gates, all the way to understanding computer architecture, with each level of abstraction hiding the details of the lower level.
+\end{shaded}
+
 \section{Introduction to Digital Logic}
+
+\subsection{Hierarchy, Modularity, and Regularity}
+
+\begin{definition}[Hierarchy]
+    The division of system into a set of modules, then further subdividing each module into smaller modules, and so on, until pieces are \textit{easy} to understand.    
+\end{definition}
+
+\begin{definition}[Modularity]
+    The design principle that modules have well-defined functions and interfaces so they connect easily without unintended side effects.
+\end{definition}
+
+\begin{definition}[Regularity]
+    The uniformality of modules, such that the reusability of common modules reduces the number of distinct modules to be designed.
+\end{definition}
+
+\subsection{Digital Logic Gates}
+Logic gates are made out of transistors:
+\begin{definition}[Transistor]
+    A transistor is a 3-terminal device behaving as a switch. When the voltage on the terminal is HI, the switch is closed, and when the voltage is LO, the switch is open.
+\end{definition}
 

### 2025-09-04 YEAR2/PHY293_LectureNotes/main.tex
pre year2
4 sept

--- 2025-09-01 YEAR2/ECE253_LectureNotes/main.tex
@@
 \thispagestyle{empty}
-{\LARGE \bf ECE 253 Lecture Notes}\\
+{\LARGE \bf PHY 293 Lecture Notes}\\
 {\large Hei Shing Cheung}\\
-Digital and Computer Systems, Fall 2025 \hfill ECE253\\
+Waves and Modern Physics, Fall 2025 \hfill PHY293\\
 \\
@@
 
-\section{Introduction to Digital Logic}
+\chapter{Waves}
+\section{Harmonic Oscillators}
 
-\section{Digital Storage Elements}
+\subsection{Governing Equations of Harmonic Oscillators}
 
-\section{Finite State Machines (FSM)}
+\paragraph{Types of Harmonic Oscillators} There are three types of harmonic oscillators: simple, damped, and driven harmonic oscillators.
+Consider a simple one dimensional harmonic oscillator, they are defined by the following differential equations:
 
-\section{Computer Organization}
+\begin{definition}[Simple Harmonic Oscillator]
+    A simple harmonic oscillator is described by Hooke's law:
 
-\section{Assembly Language}
+    \begin{equation}
+        m \frac{d^2 x}{dt^2} + kx = 0
+    \end{equation}
+    where \( k \) is the spring constant, \( m \) is the mass, and \( x \) is the displacement from equilibrium.
+\end{definition}
+
+\begin{definition}[Damped Harmonic Oscillator]
+    A damped harmonic oscillator is described by the following differential equation, by adding a damping term proportional to $\dot{x}$ to the simple harmonic oscillator equation:
+    \begin{equation}
+        m \frac{d^2 x}{dt^2} + \gamma \frac{dx}{dt} + kx = 0   
+    \end{equation}
+    where \( \gamma \) is the damping coefficient.
+\end{definition}
+
+\begin{definition}[Driven Harmonic Oscillator]
+    A driven harmonic oscillator is described by the following differential equation, which includes an external driving force $F(t)$:
+    \begin{equation}
+        m \frac{d^2 x}{dt^2} + \gamma \frac{dx}{dt} + kx = F(t)   
+    \end{equation}
+\end{definition}
+\subsection{The Wave Equation}
+
+\begin{definition}[The Wave Equation]
+    The wave equation is a second-order linear partial differential equation that describes the propagation of waves, such as sound waves, light waves, and water waves, through a medium.
+In one dimension, it is given by the following PDE:
+    \begin{equation}
+        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
+    \end{equation}
+    where \( u(x,t) \) is the wave function, \( c \) is the speed of the wave in the medium, \( x \) is the spatial coordinate, and \( t \) is time.
+\end{definition}
+
+\chapter{Modern Physics}
 \end{document}

### 2025-09-05 YEAR2/AER210_LectureNotes/main.tex
sept 5

@@
 
+% set section counter from 15 
+\setcounter{section}{14}
+\paragraph{Note} the section numbering is based on Stewart's book.
+\section{Double and Triple Integrals}
+\begin{definition}[Double Integral]
+    Let $f(x,y)$ be a function defined on a closed and bounded region $R$ in the $xy$-plane. The double integral of $f$ over $R$ is denoted by
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \iint_R f(x,y) \, dA
+    \end{equation}
+    where $dA$ represents an infinitesimal area element in the region $R$. The double integral can be interpreted as the volume under the surface defined by $z = f(x,y)$ over the region $R$.
+\end{definition}
+
+\subsection{Double Integrals in a Rectangular Region}
+
+ By the point of seeing this note, you should be familiar with the simple case of rectangular, simple cases are provided as examples:
+
+\begin{example}
+    Find the area under the quadric surface $z = 16 - x^2 - y^2$ over the square region $R = \{ (x,y) \mid 0 \le x \le 2, 0 \le y \le 2\}$.
+
+    \textbf{Note} We would have to ensure that the surface is above the $xy$-plane in the region of interest, which is true in this case.
+
+    We can set up the double integral as follows:
+    $$
+        \iint_R (16 - x^2 - y^2) \, dA = \int_0^2 \int_0^2 (16 - x^2
+        - y^2) \, dy \, dx
+    $$
+    First, we integrate with respect to $y$:
+    $$
+        \int_0^2 (16 - x^2 - y^2) \, dy = \left[ 16y - x^2y - \frac{y^3}{3} \right]_0^2 = 32 - 2x^2 - \frac{8}{3} = \frac{88}{3} - 2x^2
+    $$
+    Next, we integrate with respect to $x$:
+    $$
+        \int_0^2 \left( \frac{88}{3} - 2x^2 \right) \, dx = \left[ \frac{88}{3}x - \frac{2x^3}{3} \right]_0^2 = \frac{176}{3} - \frac{16}{3} = \frac{160}{3}
+    $$
+    Therefore, the area under the surface is $\frac{160}{3}$.
+
+\end{example}
+
+\begin{example}
+    Evaluate the double integral of $f(x,y) = x - 3y^2$ over the rectangular region $R = \{ (x,y) \mid 0 \le x \le 2, 1 \le y \le 2\}$.
+
+    We can set up the double integral as follows:
+    $$
+        \iint_R (x - 3y^2) \, dA = \int_0^2 \int_1^2 (x - 3y^2) \, dy \, dx
+    $$
+    First, we integrate with respect to $y$:
+    $$
+        \int_1^2 (x - 3y^2) \, dy = \left[ xy - y^3 \right]_1^2 = 2x - 8 - (x - 1) = x - 7
+    $$
+    Next, we integrate with respect to $x$:
+    $$
+        \int_0^2 (x - 7) \, dx = \left[ \frac{x^2}{2} - 7x \right]_0^2 = \left( 2 - 14 \right) - 0 = -12
+    $$
+    Therefore, the value of the double integral is $-12$.
+\end{example}
+
+\begin{theorem}
+    If the integrand function $f(x,y)$ is seperable, i.e., $f(x,y) = g(x)h(y)$, then the double integral can be computed as follows:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \left( \int_a^b g(x) \, dx \right) \left( \int_c^d h(y) \, dy \right)
+    \end{equation}
+    where $R = \{ (x,y) \mid a \le x \le b, c \le y \le d \}$.
+\end{theorem}
+\begin{proof}
+    \textbf{Sketch:} $h(y)$ is a constant when integrating with respect to $x$, and vice versa.
+\end{proof}
+
+\begin{example}
+    Let $f(x,y) = \sin{x} \cos{y}$ and $R = \{ (x,y) \mid 0 \le x \le \frac{\pi}{2}, 0 \le y \le \frac{\pi}{2} \}$. Evaluate the double integral $\iint_R f(x,y) \, dA$.
+
+    Since $f(x,y)$ is seperable, we can write:
+    $$
+        \iint_R f(x,y) \, dA = \left( \int_0^{\frac{\pi}{2}} \sin{x} \, dx \right) \left( \int_0^{\frac{\pi}{2}} \cos{y} \, dy \right)
+    $$
+    Evaluating each integral separately gives 1 for both, so the final result is: $1 \times 1 = 1$.
+\end{example}
+\subsection{Double Integrals in General Regions}
+\paragraph{Types of Regions} When the region $R$ is not rectangular, we can still compute the double integral by expressing the region in terms of inequalities.
+There are three common types of regions:
+\begin{definition}[Type I Region]
+    A region $R$ is called a Type I region if it can be described by the inequalities:
+    $$
+        a \le x \le b, \quad g_1(x) \le y \le g_2(x)
+    $$
+    where $g_1(x)$ and $g_2(x)$ are continuous functions on the interval $[a,b]$.
+Then, to evaluate the double integral over a Type I region for a continuous function $f(x,y)$, we set up the integral as follows:
+
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \int_a^b \int_{g_1(x)}^{g_2(x)} f(x,y) \, dy \, dx
+    \end{equation}
+
+    \textbf{Integral Order:} Integrate with respect to $y$ first, then $x$.
+
+    \textbf{Intuition:} As we traverse the outer part ($x$), we are summing up vertical slices (in $y$), and the bounds of those slices depend on $x$ and changes.
+\end{definition}
+
+\begin{definition}[Type II Region]
+    Type II reigion is similar to Type I, but the roles of $x$ and $y$ are swapped. A region $R$ is called a Type II region if it can be described by the inequalities:
+    $$
+        c \le y \le d, \quad h_1(y) \le x \le h_2(y)
+    $$
+    where $h_1(y)$ and $h_2(y)$ are continuous functions on the interval $[c,d]$.
+Then, to evaluate the double integral over a Type II region for a continuous function $f(x,y)$, we set up the integral as follows:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \int_c^d \int_{h_1(y)}^{h_2(y)} f(x,y) \, dx \, dy
+    \end{equation}
+
+    The integral order and intuition is mirrored from Type I, but we are summing up horizontal slices (in $x$), and the bounds of those slices depend on $y$ and changes.
+\end{definition}
+
+\begin{definition}[Type III Region]
+    A region $R$ is called a Type III region if it can be described as the union of a finite number of Type I and Type II regions.
+To evaluate the double integral over a Type III region for a continuous function $f(x,y)$, we can break down the integral into separate integrals over each Type I or Type II subregion and sum them up:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \sum_{i=1}^n \iint_{R_i} f(x,y) \, dA
+    \end{equation}
+    where each $R_i$ is either a Type I or Type II region. And that:
+    $$
+        \bigcup_{i=1}^n R_i = R \quad \text{and} \quad R_i \cap R_j = \emptyset \text{ for } i \neq j
+    $$
+    This approach allows us to handle more complex regions by breaking them down into simpler parts.
+\end{definition}
+
+\begin{example}
+    Find the volume of the solid that lies under the paraboloid $z = f(x,y) = x^2 + y^2$ and above the region $R$ bounded
+     by $y = 2x$ and $y = x^2$.
+
+     First, you would sketch the region to understand its shape and boundaries at Figure \ref{fig:region}.
+
+    \begin{figure}[h]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                axis lines = middle,
+                xlabel = $x$,
+                ylabel = $y$,
+                xmin = -1, xmax = 3,
+                ymin = -1, ymax = 5,
+                xtick = {0,1,2,3},
+                ytick = {0,1,2,3,4},
+                grid = both,
+                minor tick num = 1,
+                domain = -1:3,
+            ]
+            \addplot[blue, thick, name path=twox, name path global=twox] {2*x};
+            \addplot[red, thick, name path=xsq, name path global=xsq] {x^2};
+            \addplot[fill=gray, opacity=0.5] fill between[of=twox and xsq, soft clip={domain=0:2}];          \end{axis}
+        \end{tikzpicture}
+        \caption{Region bounded by $y = 2x$ and $y = x^2$} \label{fig:region}
+    \end{figure}
+
+    We can tell that this is a Type I regionwhere $0 \le x \le 2$, and $x^2 \le y \le 2x$. Thus, we can set up the double integral as follows:
+    $$
+        \iint_R (x^2 + y^2) \, dA = \int_0^2 \int_{x^2}^{2x} (x^2 + y^2) \, dy \, dx
+    $$
+    First, we integrate with respect to $y$:
+    $$
+        \int_{x^2}^{2x} (x^2 + y^2) \, dy = \left[ x^2y + \frac{y^3}{3} \right]_{y=x^2}^{y=2x} = 2x^3 + \frac{8x^3}{3} - x^4 - \frac{x^6}{3} = \frac{14x^3}{3} - x^4 - \frac{x^6}{3}
+    $$
+    Next, we integrate with respect to $x$:
+    $$
+        \int_0^2 \left( \frac{14x^3}{3} - x^4 - \frac{x^6}{3} \right) \, dx = \left[ \frac{14x^4}{12} - \frac{x^5}{5} - \frac{x^7}{21} \right]_0^2 = \frac{216}{35}
+    $$
+    Therefore, the volume of the solid is $\frac{216}{35}$.
+
+\end{example}
+
+\begin{example}
+    Consider the above example, but we want to set it up as a Type II region.
+The region $R$ can be described by $0 \le y \le 4$, and $\frac{y}{2} \le x \le \sqrt{y}$.
+Thus, we can set up the double integral as follows:
+    $$
+        \iint_R (x^2 + y^2) \, dA = \int_0^4 \int_{\frac{y}{2}}^{\sqrt{y}} (x^2 + y^2) \, dx \, dy
+    $$
+    First, we integrate with respect to $x$:
+    $$
+        \int_{\frac{y}{2}}^{\sqrt{y}} (x^2 + y^2) \, dx = \left[ \frac{x^3}{3} + y^2x \right]_{x=\frac{y}{2}}^{x=\sqrt{y}} = \frac{y^{\frac{3}{2}}}{3
+        } + y^{\frac{5}{2}} - \frac{y^3}{24} - \frac{y^3}{2} = \frac{y^{\frac{3}{2}}}{3} + y^{\frac{5}{2}} - \frac{13y^3}{24}
+    $$
+    Next, we integrate with respect to $y$:
+    $$
+        \int_0^4 \left( \frac{y^{\frac{3}{2}}}{3} + y^{\frac{5}{2}} - \frac{13y^3}{24} \right) \, dy = \left[ \frac{2y^{\frac{5}{2}}}{15} + \frac{2y^{\frac{7}{2}}}{7} - \frac{13y^4}{96} \right]_0^4 = \frac{216}{35}
+    $$
+    Therefore, the volume of the solid is $\frac{216}{35}$, which is consistent with the previous result. This is also consistent with Fubini's Theorem.
+\end{example}
+
+\begin{example}
+    Integrate the surface given by $z = e^{x^2}$ over the triangular region with vertices at $(0,0)$, $(1,0)$, and $(1,1)$. We can describe the region as either a Type I or Type II region:
+
+    (Incorrect) Here, we will describe it as a Type II regionwhere $0 \le y \le 1$, and $y \le x \le 1$. Thus, we can set up the double integral as follows:
+    $$
+        \iint_R e^{x^2} \, dA = \int_0^1 \int_y^1 e^{x^2} \, dx \, dy
+    $$
+    We can tell that $e^{x^2}$ does not have an elementary antiderivative, so we cannot integrate with respect to $x$ directly.
+
+    (Correct) However, we can change the order of integration to make it a Type I regionwhere $0 \le x \le 1$, and $0 \le y \le x$. Thus, we can set up the double integral as follows:
+    $$
+        \iint_R e^{x^2} \, dA = \int_0^1 \int_0^x e^{x^2} \, dy \, dx
+    $$
+    First, we integrate with respect to $y$:
+    $$
+        \int_0^x e^{x^2} \, dy = \left[ y e^{x^2} \right]_0^x = xe^{x^2}
+    $$
+    Next, we integrate with respect to $x$:
+    $$
+        \int_0^1 xe^{x^2} \, dx
+    $$
+    This is now obvious, a simple $u$-substitution with $u = x^2$, $du = 2x \, dx$:
+    $$
+        \int_0^1 xe^{x^2} \, dx = \frac{1}{2} \int_0^1 e^u \, du = \frac{1}{2} \left[ e^u \right]_0^1 = \frac{e - 1}{2}
+    $$
+    Therefore, the value of the double integral is $\frac{e - 1}{2}$.
+
+    \textbf{Intuition} When the integrand is difficult to integrate with respect to one variable, consider changing the order of integration.
+You should be able to tell that $e^{x^2}$ has no elementary antiderivative, so you would have ruled out integrating with respect to $x$ first.
+\end{example}
+
+\subsection{Formal Definition of Double Integrals}
+There is two definitions of double integrals in this course, due to the discrepancy between Stewart's book and the lectures.
+
+\begin{shaded}
+    \subsubsection*{\textbf{Review.  } Formal Definition of Definite Integral (Single Variable)}
+
+    Consider $y = f(x) \ge 0$ on the interval $x \in [a,b]$.
+We divide the interval into $n$ subintervals of equal width $\Delta x = \frac{b-a}{n}$, and let $x_i^*$ be a sample point in the $i$-th subinterval.
+The Riemann sum is given by:
+    $$
+        S_n = \sum_{i=1}^n f(x_i^*) \Delta x
+    $$
+    Now, for any $x_i^*$, we consider the minimum and maximum values of $f(x_i^*)$ in the $i$-th subinterval, denoted as $m_i$ and $M_i$ respectively.
+We can then define the lower sum $L_n$ and upper sum $U_n$ as follows:
+    $$
+        L_n = \sum_{i=1}^n m_i \Delta x \quad \text{and} \quad U_n = \sum_{i=1}^n M_i \Delta x
+    $$
+    To satisfy the squeeze theorem, for all $i$, we would need:
+    $$
+        \lim_{n \to \infty} M_i - m_i = \lim_{\delta x \to 0} M_i - m_i = 0
+    $$
+    If $f(x)$ is continuous on $[a,b]$. Then, we have:
+    $$
+        \lim_{n \to \infty} L_n = \lim_{n \to \infty} U_n = \int_a^b f(x) \, dx
+    $$
+
+    For the case of discontinuous functions, if the set of discontinuities has measure zero, then the function is still integrable.
+\end{shaded}
+
+\begin{definition}[Definition of Double Integral]
+    Let $R$ be a rectangular region in the $xy$-plane given by $R = [a,b] \times [c,d]$. The double integral of a function $f(x,y)$ over the region $R$ is defined as:
+    \begin{subequations}
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \lim_{n \to \infty} \sum_{i=1}^n f(x_i^*,y_i^*) \Delta A_i \quad (\text{Riemann Definition})
+    \end{equation}
+    where $\Delta A_i$ is the area of the $i$-th subrectangle, and $(x_i^*,y_i^*)$ is a sample point in it. The limit is taken as the maximum diameter of the subrectangles approaches zero.
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \lim_{n,m \to \infty} \sum_{i=1}^n \sum_{j=1}^m f(x_i^*,y_j^*) \Delta A_{ij} \quad (\text{Grid Formulation})
+    \end{equation}
+    where $\Delta A_{ij}$ is the area of the $ij$-th subrectangle, and $(x_i^*,y_j^*)$ is a sample point in it.
+Note that the $\Delta A_{ij}$ may be non-uniform.
+The limit is taken as the maximum diameter of the subrectangles approaches zero.
+    \end{subequations}
+
+    Similarly, the lower and upper sums for double integrals are:
+    \begin{subequations}
+    \begin{equation}
+        L_n = \sum_{i=1}^n m_i \Delta A_i \quad \text{and} \quad U_n = \sum_{i=1}^n M_i \Delta A_i \quad (\text{Riemann Definition})
+    \end{equation}
+    \begin{equation}
+        L_{n,m} = \sum_{i=1}^n \sum_{j=1}^m m_{ij} \Delta A_{ij} \quad \text{and} \quad U_{n,m} = \sum_{i=1}^n \sum_{j=1}^m M_{ij} \Delta A_{ij} \quad (\text{Grid Formulation})
+    \end{equation}
+    \end{subequations}
+    Here, $m_{ij}$ and $M_{ij}$ are the minimum and maximum values of $f(x,y)$ in the $ij$-th subrectangle.
+Define $||P|| = \max ||(\Delta x_i, \Delta y_j)||$ as the maximum diameter of the subrectangles.
+For the squeeze theorem, we require:
+    $$
+        \lim_{n,m \to \infty} (M_{ij} - m_{ij}) = \lim_{||P|| \to 0} (M_{ij} - m_{ij}) = 0
+    $$
+    If $f(x,y)$ is continuous on $R$, then:
+    $$
+        \lim_{n,m \to \infty} L_{n,m} = \lim_{n,m \to \infty} U_{n,m} = \iint_R f(x,y) \, dA
+    $$
+
+    The Riemann definition and grid formulation are similar.
+\end{definition}
+
 \chapter{Fluid Mechanics}

### 2025-09-05 YEAR2/ECE253_LectureNotes/main.tex
4 sept
sept 5

@@
 
+\chapter{Digital Circuits that Compute, Store, and Control}
 \begin{shaded}
@@
 \paragraph{Layer of abstraction} At this course, for the digital systems part, we would start from understanding logic gates, all the way to understanding computer architecture, with each level of abstraction hiding the details of the lower level.
-\end{shaded}
-
-\section{Introduction to Digital Logic}
 
@@
 
+\paragraph{Factors Affecting Speed of Digital Circuits}
+\begin{itemize}
+    \item \textbf{Transistors and Electrons take time to switch.} A transistor (State of the Art) takes 2-3 picoseconds to switch. Gates takes 40 ps and an 8-bit adder takes 300 ps.
+    \item \textbf{Wires take time to propagate signals.} Signals travel at approximately 2/3 the speed of light in a vacuum, which is about 200,000 kilometers per second in a typical silicon wire.
+    \item \textbf{Capacitance} There would be RCL circuits formed by the wires and transistors, which would cause delay.
+\end{itemize}
+
+\end{shaded}
+
+\section{Information Representation}
+
+\subsection{Number Systems}
+\begin{definition}[Number System]
+    A number system is a way of representing numbers using a set of symbols (digits) and a base (radix). The base determines the number of unique digits that can be used in the number system.
+\end{definition}
+
+\paragraph{Common Number Systems} You should be familiar with the following number systems:
+\begin{itemize}
+    \item Decimal (Base 10): Digits 0-9
+    \item Binary (Base 2): Digits 0-1
+    \item Hexadecimal (Base 16): Digits 0-9, A-F
+\end{itemize}
+
+In computer systems, we use binary to represent information, and we would often use hexadecimal to represent binary numbers in a more compact way - a group of 4 bits (a nibble) can be represented by a single hexadecimal digit.
+
+\begin{example}[Binary, Decimal, and Hexadecimal Numbers]
+    Below is a table showing the conversion of binary numbers to decimal numbers, along with their hexadecimal representation.
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|}
+            \hline
+            Binary & Decimal & Hexadecimal \\
+            \hline
+            0000 & 0 & 0 \\
+            0001 & 1 & 1 \\
+            0010 & 2 & 2 \\
+            0011 & 3 & 3 \\
+            0100 & 4 & 4 \\
+            0101 & 5 & 5 \\
+            0110 & 6 & 6 \\
+            0111 & 7 & 7 \\
+            1000 & 8 & 8 \\
+            1001 & 9 & 9 \\
+            1010 & 10 & A \\
+            1011 & 11 & B \\
+            1100 & 12 & C \\
+            1101 & 13 & D \\
+            1110 & 14 & E \\
+            1111 & 15 & F \\
+            \hline
+        \end{tabular}
+        \caption{Binary to Decimal and Hexadecimal Conversion}
+    \end{table}
+
+\end{example}
+\begin{example}[Decimal to Binary Conversion]
+    To convert a decimal number to binary, we can use the method of successive division by 2. For example, to convert the decimal number 437 to binary:
+    \begin{align*}
+        437 \div 2 &= 218 & \text{remainder 1} \\
+        218 \div 2 &= 109 & \text{remainder 0} \\
+        109 \div 2 &= 54 & \text{remainder 1} \\
+        54 \div 2 &= 27 & \text{remainder 0} \\
+        27 \div 2 &= 13 & \text{remainder 1} \\
+        13 \div 2 &= 6 & \text{remainder 1} \\
+        6 \div 2 &= 3 & \text{remainder 0} \\
+        3 \div 2 &= 1 & \text{remainder 1} \\
+        1 \div 2 &= 0 & \text{remainder 1}
+    \end{align*}
+    Reading the remainders from bottom to top, we get the binary representation of 437
+\end{example}
+
+\begin{example}
+    To convert (512000)$_{10}$ to binary, we recognize that $512000 = 2^9 \times 1000$.
+We know that $2^9 = 512$ and $1000_{10} = 1111101000_2$ (by method outlined above).
+Therefore, we can shift the binary representation of 1000 left by 9 bits to get the binary representation of 512000:
+    \[(512000)_{10} = (1111101000000000000)_2\]
+\end{example}
+
+\paragraph{Note} An alternative method is to devide by powers of 2.
+
+\paragraph{Binary Arithmetic} Binary arithmetic is similar to decimal arithmetic, but it only uses two digits (0 and 1). The 0 and 1 could be associated with a xor logic.
 \section{Digital Storage Elements}
@@
 
+\chapter{Computer Organization and Assembly Language}
+\begin{shaded}
+\textbf{What is Assembly Language?} We know that we can run C/C++ on any computer (Machine Agnostic), but how does the computer understand C/C++?
+The answer is the compiler that parse it to assembly through:
+
+\begin{enumerate}
+    \item \textbf{Front-end Parser:} The front-end parser would parse the C/C++ code into an intermediate representation (IR), which is a low-level representation of the code that is easier to optimize.
+The front-end parser would also perform optimizations on the IR, such as loop unrolling, inlining, and dead code elimination.
+    \item \textbf{Back-end Parser:} The back-end parser would take the optimized IR and generate assembly code for a specific architectures.
+\end{enumerate}
+
+The assembly code is then assembled into machine code, which is a series of 0s and 1s that the computer can understand.
+The assembly would be specific to the architecture of the computer (machine dependent), which is why we have different assembly languages for different architectures (e.g., x86, RISC-V, ARM).
+\end{shaded}
 \section{Computer Organization}

### 2025-09-05 YEAR2/MAT292_LectureNotes/main.tex
4 sept
sept 5

@@
 
-            We know that this is of the form $g \cdot g'$, where $g(x) = \ln x$ and $g'(x) = \frac{1}{x}$. Thus, by the rule that:
+            We know that this is of the form $g \cdot g'$where $g(x) = \ln x$ and $g'(x) = \frac{1}{x}$. Thus, by the rule that:
 
-            $$
+            \begin{equation}
             \frac{1}{2} \frac{d}{dx} [g(x)]^2 = g(x) g'(x)
-            $$ 
-            we can deduce that"
+            \end{equation}
+            we can deduce that:
             $$
@@
 
-We could understand the eigenproblems associated with systems of ODEs as an extension of the above, where we look for solutions of the form:
+We could understand the eigenproblems associated with systems of ODEs as an extension of the abovewhere we look for solutions of the form:
 \begin{equation}
@@
 
-    \textbf{Fixed-Point Solution} We denote the trivial case where \( u(0) = T_a \) as the fixed-point solution, where the temperature of the object is equal to the ambient temperature at time \( t = 0 \).
+    \textbf{Fixed-Point Solution} We denote the trivial case where \( u(0) = T_a \) as the fixed-point solutionwhere the temperature of the object is equal to the ambient temperature at time \( t = 0 \).
 It is easy to see that \( u' = 0 \) in this case so \( u(t) = T_a \) for all \( t \).
@@
         \frac{du}{dt} &= -k(u - T_a) \\
-        \int_0^t \frac{du}{u - T_a} &= -k \int_0^t dt \\
+        \int \frac{du}{u - T_a} &= -k \int dt \\
         \ln|u - T_a| &= -kt + C \\
@@
 
+\subsection{Fixed Points}
+
+We first consider the autonomous system:
+\begin{equation}
+    \frac{dx}{dt} = f(x)
+\end{equation}
+
+\begin{definition}[Autonomous System]
+    An autonomous system is a system of ordinary differential equations (ODEs) in which the independent variable (usually time) does not explicitly appear in the equations.
+In other words, the rate of change of the dependent variable(s) depends only on the current state of the system and not on time itself.
+Autonomous systems can be expressed in the form:
+    $$
+    \frac{dx}{dt} = f(x)
+    $$
+    where \( x \) is the state vector and \( f(x) \) is a function that describes how the state changes over time.
+\end{definition}
+
+\begin{definition}[Fixed Point]
+    A fixed point (or equilibrium point) of a dynamical system is a point in the phase space where the system remains unchanged over time.
+Mathematically, for a system described by the differential equation \( \frac{dx}{dt} = f(x) \), a fixed point \( x^* \) satisfies:
+    $$
+    f(x^*) = 0
+    $$
+    This means that if the system starts at \( x^* \), it will stay at \( x^* \) for all future times.
+\end{definition}
+
+\begin{example}
+    Consider $x' = \sin(x)$. The fixed points are given by:
+    $$
+    \sin(x) = 0 \implies x = n\pi, \quad n \in \mathbb{Z}
+    $$
+    Thus, the fixed points are \( x = 0, \pm \pi, \pm 2\pi, \ldots \)
+\end{example}
+
+\subsubsection{Stability of Fixed Points}
+\paragraph{Intuition} Consider a simple pendulum parametized by the angle \( \theta \) from the vertical.
+The fixed points of this system occur when the pendulum is at rest, which happens at \( \theta = 0 \) (hanging straight down) and \( \theta = \pi \) (inverted position).
+The case where \textbf{\( \theta = 0 \) is stable}, as small perturbations will cause the pendulum to oscillate around this point.
+In contrast, \textbf{\( \theta = \pi \) is unstable,} as any small perturbation will cause the pendulum to fall away from this position.
+
+\paragraph{What determines stability?} A fixed point \( x^* \) of a dynamical system \( \frac{dx}{dt} = f(x) \) is classified as:
+
+\begin{subequations}
+    stable if:
+    \begin{equation}
+        \frac{df}{dx} \bigg|_{x=x^*} < 0
+    \end{equation}
+    unstable if:
+    \begin{equation}
+        \frac{df}{dx} \bigg|_{x=x^*} > 0
+    \end{equation}
+    semi-stable if:
+    \begin{equation}
+        \frac{df}{dx} \bigg|_{x=x^*} = 0
+    \end{equation}
+\end{subequations}
+
+\begin{proof}
+    A demonstration of this will be given in the next section on Linear Stability Analysis \ref{def:linear-stability-analysis}.
+\end{proof}
+
+\begin{example}
+    Consider the system $x' = \sin(x)$.
+The fixed points are at \( x = n\pi \)where \( n \in \mathbb{Z} \).
+To determine the stability of these fixed points, we compute the derivative of \( f(x) = \sin(x) \):
+    $$
+    \frac{df}{dx} = \cos(x)
+    $$
+    Evaluating this derivative at the fixed points $x = n\pi$ gives:
+    $$
+    \cos(n\pi) =\begin{cases}
+        1 & \text{if } n \text{ is even (unstable)} \\
+        -1 & \text{if } n \text{ is odd (stable)}
+    \end{cases}
+    $$
+    Thus, the fixed points at \( x = 0, \pm 2\pi, \pm 4\pi, \ldots \) are unstable, while the fixed points at \( x = \pm \pi, \pm 3\pi, \pm 5\pi, \ldots \) are stable. 
+\end{example}
+
+\paragraph{Drawing Fixed Point Diagrams and Phase Portraits} By visualizing the direction of $x$ due to $x' = f(x)$ on a $x$, $f(x)$ plane (fixed points diagram), we can draw a phase portrait.
+For the phase portrait, we can demonstrate the fixed points as horizontal lines, and draw points of $x$ that converge to or diverge from these fixed points according to the direction of $x'$.
+
+We draw the fixed point diagram with filled circles for stable fixed points and open circles for unstable fixed points. Arrows indicate the direction of flow towards or away from the fixed points.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.6\textwidth]{fixpointdiagram.jpeg}
+    \caption{Fixed Point Diagram and Phase Portrait for \( x' = \sin(x) \)}
+    \label{fig:fixpointdiagram}
+\end{figure}
+
+\begin{theorem}[Non-Intersection of Solutions]
+    In a dynamical system described by an ordinary differential equation (ODE) with a unique solution for given initial conditions, the trajectories of different solutions cannot intersect in the phase space.
+This means that if two solutions start from different initial conditions, they will never cross each other at any point in time.
+\end{theorem}
+\begin{proof}
+    \textbf{Sketch:} If the solutions intersect, then at the point of intersection, the solution is not unique, which contradicts the assumption of uniqueness.
+\end{proof}
+
+\paragraph{What Happens When We Consider $t \to -\infty$?} 
+\begin{theorem}[Stability Reversal]
+    A fixed point that is stable as \( t \to \infty \) becomes unstable as \( t \to -\infty \), and vice versa.
+\end{theorem}
+
+\begin{proof}
+    This can be understood by considering the time-reversed system.
+If we denote the time-reversed variable as \( t' = -t \), then the original system \( \frac{dx}{dt} = f(x) \) transforms, by the chain rule, into:
+    $$
+    \frac{dx}{dt'} = -f(x) 
+    $$
+    In this new system, the direction of flow is reversed.
+Therefore, if a fixed point \( x^* \) is stable in the original system (i.e., trajectories approach \( x^* \) as \( t \to \infty \)), it will be unstable in the time-reversed system (i.e., trajectories move away from \( x^* \) as \( t' \to \infty \)).
+Conversely, an unstable fixed point in the original system becomes stable in the time-reversed system.
+This demonstrates the reversal of stability when considering \( t \to -\infty \).
+\end{proof}
+
+\subsubsection{Linear Stability Analysis}
+
+\begin{definition}[Linear Stability Analysis] \label{def:linear-stability-analysis}
+    Consider the system $x' = f(x)$. Let $\epsilon$ be any perturbation around a fixed point $x^*$, such that:
+    $$x = x^* + \epsilon
+    $$
+    To analyze the stability of the fixed point, we analyze the behavior of the perturbation $\epsilon(t)$ over time, since $x^*$ is fixed, we have:
+    \begin{align*}
+        \epsilon' = x' = f(x) &= f(x^* + \epsilon) \\
+        \intertext{Ignoring higher order terms in Taylor expansion, we have:}
+        &\approx f(x^*) + \epsilon f'(x^*) + \cancel{\frac{1}{2} \epsilon^2 f''(x^*) + \ldots} \\
+        &= \epsilon f'(x^*) \quad \text{(since } f(x^*) = 0 \text{)} \\
+        \Rightarrow \epsilon' &= f'(x^*) \epsilon
+    \end{align*}
+
+    This is a linear ODE in $\epsilon$, which can be solved as:
+    $$\epsilon(t) = \epsilon e^{f'(x^*) t}
+    $$
+    The behavior of $\epsilon(t)z$ depends on the sign of $f'(x^*)$:
+    \begin{itemize}
+        \item If $f'(x^*) < 0$, then $\epsilon(t) \to 0$ as $t \to \infty$, indicating that the fixed point is stable.
+        \item If $f'(x^*) > 0$, then $\epsilon(t) \to \infty$ as $t \to \infty$, indicating that the fixed point is unstable.
+        \item If $f'(x^*) = 0$, the linear analysis is inconclusive, and higher-order terms must be considered.
+    \end{itemize}
+    Thus, the stability of the fixed point can be determined by the sign of $f'(x^*)$.
+\end{definition}
+
+\begin{example}
+    Consider the system $x' =x^2$. The only fix point is at \( x = 0 \). We have $f'(x) = 2x$, and at the fixed point:
+    $$
+    f'(0) = 0
+    $$
+    Thus, the linear stability analysis is inconclusive.
+We draw the fixed point half filled on the left, and half empty on the right.
+This is becuase for \( x < 0 \), \( f(x) = x^2 > 0 \), so points to the left of the fixed point move away from it (unstable).
+For \( x > 0 \), \( f(x) = x^2 > 0 \), so points to the right of the fixed point also move away from it (unstable).
+Therefore, the fixed point at \( x = 0 \) is unstable.
+\end{example}
+
+\subsection{Classification of ODEs}
+
+\begin{shaded}
+\paragraph{ODE vs PDE} An ordinary differential equation (ODE) contains functions of a single variable and their derivatives.
+A partial differential equation (PDE) contains functions of multiple variables and their partial derivatives.
+\begin{example}[Heat Equation]
+    The heat equation is a PDE that describes how heat diffuses through a given region over time. It is given by:
+    $$
+    \frac{\partial u}{\partial t} = \alpha \nabla^2 u 
+    $$
+    where \( u(x, t) \) is the temperature distribution function, \( \alpha \) is the thermal diffusivity constant, and \( \nabla^2 \) is the Laplacian operator.
+\end{example}
+\begin{example}[Laplace's Equation]
+    Laplace's equation is a PDE that describes the behavior of scalar fields such as electric potential and fluid flow. It is given by:
+    $$
+    \nabla^2 \phi = 0
+    $$
+    where \( \phi(x, y, z) \) is the scalar potential function, and \( \nabla^2 \) is the Laplacian operator.
+
+\end{example}
+\end{shaded}
+\begin{definition}[Order of an ODE]
+    The order of an ordinary differential equation (ODE) is determined by the highest derivative present in the equation.
+\end{definition}
+\subsubsection{Linearity and Homogeneity of ODEs}
+\begin{definition}[Linearity of an ODE]
+    An ordinary differential equation (ODE) is considered linear if it can be expressed in the form:
+    $$
+    a_n(x) \frac{d^n y}{dx^n} + a_{n-1}(x) \frac{d^{n-1} y}{dx^{n-1}} + \ldots + a_1(x) \frac{dy}{dx} + a_0(x) y = g(x)
+    $$
+    where \( y \) is the dependent variable, \( x \) is the independent variable, \( a_i(x) \) are functions of \( x \), and \( g(x) \) is a known function.
+In a linear ODE, the dependent variable \( y \) and its derivatives appear to the first power and are not multiplied together.
+Otherwise, the ODE is considered nonlinear.
+\end{definition}
+
+\begin{definition}[Homogenity of Linear ODEs]
+    A linear ordinary differential equation (ODE) is said to be homogeneous if the function \( g(x) \) on the right-hand side of the equation is equal to zero for all values of \( x \).
+In other words, a homogeneous linear ODE has the form:
+    $$
+    a_n(x) \frac{d^n y}{dx^n} + a_{n-1}(x) \frac{d^{n-1} y}{dx^{n-1}} + \ldots + a_1(x) \frac{dy}{dx} + a_0(x) y = 0
+    $$
+    If \( g(x) \) is not identically zero, then the ODE is considered non-homogeneous.
+Homogeneous linear ODEs have special properties and solution methods that differ from those of non-homogeneous linear ODEs.
+\end{definition}
+
+\begin{theorem}[Principle of Superposition]
+    Let:
+    $$
+    a_0(t)y + \dots + a_n(t) \frac{d^n y}{dt^n} =0
+    $$
+    Assuming that $y_1(t)$ and $y_2(t)$ are two solutions to the above equation, then any linear combination of these solutions $A y_1(t) + B y_2(t)$, where \( A \) and \( B \) are constants, is also a solution.
+\end{theorem}
+\begin{proof}
+    This could be demonstrated by the following derivation:
+    \begin{align*}
+        & a_0(t)(A y_1 + B y_2) + \dots + a_n(t) \frac{d^n}{dt^n}(A y_1 + B y_2) \\ 
+        &= A \left( a_0(t)y_1 + \dots + a_n(t) \frac{d^n y_1}{dt^n} \right) + B \left( a_0(t)y_2 + \dots + a_n(t) \frac{d^n y_2}{dt^n} \right) \\
+        &= A \cdot 0 + B \cdot 0 \\
+        &= 0
+    \end{align*}
+    Thus, \( A y_1(t) + B y_2(t) \) is also a solution to the original equation. WLOG, this can be extended to any finite number of solutions.
+\end{proof}
+
+Conisder the non-homogeneous case:
+\begin{theorem}[General Solution of Non-Homogeneous Linear ODEs]
+    Let:
+    $$
+    a_0(t)y + \dots + a_n(t) \frac{d^n y}{dt^n} =g(t)
+    $$
+    If \( y_p(t) \) is a particular solution to the non-homogeneous equation, and \( y_h(t) \) is any solution to the corresponding homogeneous equation, then:
+    $$
+    y(t) = A y_h(t) + y_p(t)
+    $$
+    is the general solution to the non-homogeneous equation, where \( A \) is an arbitrary constant.
+\end{theorem}
+\begin{proof}
+    \textbf{Sketch:} The solution for the homogeneous part would simply be 0, so the particular solution would be the only solution to the non-homogeneous equation.
+The general solution is then the sum of the homogeneous and particular solutions.
+\end{proof}
+
+\subsubsection{Separable ODEs}
+\begin{definition}[Separable ODE]
+    An ordinary differential equation (ODE) is said to be separable if it can be expressed in the form:
+    $$
+    \frac{dy}{dx} = g(x) h(y) = f(x, y)
+    $$
+    where \( g(x) \) is a function of the independent variable \( x \) only, and \( h(y) \) is a function of the dependent variable \( y \) only.
+This allows the variables to be separated on opposite sides of the equation, enabling integration with respect to each variable independently:
+
+    \begin{align*}
+        \frac{dy}{dx} &= g(x) h(y) \\
+        \int \frac{1}{h(y)} dy &= \int g(x) dx \\
+        \intertext{Let \( H(y) \) be the antiderivative of \( \frac{1}{h(y)} \) and \( G(x) \) be the antiderivative of \( g(x) \), we have:}
+        H(y) &= G(x) + C \\
+        \Rightarrow y &= H^{-1}(G(x) + C)
+    \end{align*}
+    where \( C \) is the constant of integration.
+\end{definition}
+
+\paragraph{Remarks} If $g$ and $h$ are continuous, then there need not exist solution in a neighbourhood (open $\epsilon$-ball) of any point $(x_0, y_0)$ so long as $h(y_0) \neq 0$.
+
+\begin{example}
+    Consider the IVP:
+    $$
+    \begin{cases}
+        \frac{dx}{dt} = ax \\
+        x(0) = x_0
+    \end{cases}
+    $$
+    then the solution is given by:
+    $$
+    x(t) = x_0e^{at}
+    $$
+    Is the solution unique? Suppose $x$ is any solution, then we check if $y = e^{-at}x$ is constant:
+    \begin{align*}
+        \frac{dy}{dt} &= \frac{d}{dt}(e^{-at}x) \\
+        &= -ae^{-at}x + e^{-at}\frac{dx}{dt} \\
+        &= -ae^{-at}x + e^{-at}(ax) \\
+        &= 0
+    \end{align*}
+    Thus, $y$ is constant, and since $y(0) = x_0$, we have $y(t) = x_0$ for all $t$. Thus,
+    $$x(t) = x_0 e^{at}
+    $$
+    which is the unique solution.
+\end{example}
+
+\begin{example}[Logistic Equation]
+    The logistic equation is a first-order nonlinear separable autonomous ODE that models population growth with a carrying capacity. It is given by:
+    $$
+    \frac{dx}{dt} = \alpha x \left(1 - x \right)
+    $$
+    where \( \alpha \) is the growth rate.
+
+    To solve this equation, we can separate the variables:
+    \begin{align*}
+        \frac{dx}{\alpha x(1 - x)} &= 1 dt \\
+        \int \frac{1}{\alpha x(1 - x)} dx &= \int 1 dt \\
+        \intertext{Using partial fraction decomposition, we have:}
+        \frac{1}{\alpha x(1 - x)} &= \frac{A}{x} + \frac{B}{1 - x} \\
+        1 &= A(1 - x) + Bx \\
+        1 &= A + (B - A)x \\
+        \Rightarrow A &= 1, \quad B - A = 0 \\
+        \Rightarrow B &= 1
+        \intertext{Thus, we have:}
+        \int \left( \frac{1}{\alpha x} + \frac{1}{\alpha (1 - x)} \right) dx &= \int 1 dt \\
+        \frac{1}{\alpha} \left( \ln|x| - \ln|1 - x| \right) &= t + C \\
+        \ln \left| \frac{x}{1 - x} \right| &= \alpha t + C' \\
+        \frac{x}{1 - x} &= e^{\alpha t + C'} = Ce^{\alpha t} \\
+        \Rightarrow x(t) &= \frac{Ce^{\alpha t}}{1 + Ce^{\alpha t}}
+    \end{align*}
+
+    To analyze the stability of the fixed points, we first find the fixed points by setting \( \frac{dx}{dt} = 0 \):
+    $$\alpha x(1 - x) = 0 \implies x = 0 \text{ or } x = 1
+    $$
+    Next, we compute the derivative of \( f(x) = \alpha x(1 - x) \):
+    $$\frac{df}{dx} = \alpha (1 - 2x)
+    $$
+    Evaluating this derivative at the fixed points, assumeing \( \alpha > 0 \):
+    \begin{itemize}
+        \item At \( x = 0 \):
+        $$\frac{df}{dx} \bigg|_{x=0} = \alpha > 0 \quad \text{(unstable)}
+        $$
+        \item At \( x = 1 \):
+        $$\frac{df}{dx} \bigg|_{x=1} = -\alpha < 0 \quad \text{(stable)}
+        $$
+    \end{itemize}
+    Thus, the fixed point at \( x = 0 \) is unstable, while the fixed point at \( x = 1 \) is stable.
+\end{example}
 \end{document}

### 2025-09-05 YEAR2/PHY293_LectureNotes/main.tex
sept 5

@@
 
-    \begin{equation}
+    \begin{equation} \label{eq:hooke}
         m \frac{d^2 x}{dt^2} + kx = 0
@@
 
+\subsection{Simple Harmonic Motion}
+\begin{definition}[Simple Harmonic Motion]
+    You should have leaned the Hooke's law and Newton's second law, which gives us the equation of motion for a simple harmonic oscillator.
+The same with the equation (\ref{eq:hooke}), which can be rewritten as:
+    \begin{equation}
+        F = m\ddot{x} = -kx
+    \end{equation}
+    By setting $\omega^2 = \frac{k}{m}$, ageneral solution can be written as:
+    \begin{equation}
+        x(t) = x_0 + A_1 \cos(\omega t) + A_2 \sin(\omega t)
+    \end{equation}
+    where \( A \) are the constants determined by the IVP, \( \omega \) is the angular frequency, and \( \phi \) is the phase constant.
+$x_0$ is the equilibrium position where we generally set it to be 0.
+The unknown constant can be determined by knowing $x, \dot{x}$ at specific times.
+\end{definition}
+
+\begin{definition}[Period, Frequency, and Angular Frequency]
+    The period \( T \) is the time it takes for one complete cycle of the motion, given by:
+    \begin{equation}
+        T = 2\pi \sqrt{\frac{m}{k}}
+    \end{equation}
+    The frequency \( f \) is the number of cycles per unit time, given by:
+    \begin{equation}
+        f = \frac{1}{T} = \frac{1}{2\pi} \sqrt{\frac{k}{m}}
+    \end{equation}
+    The angular frequency \( \omega \) is related to the frequency by:
+    \begin{equation}
+        \omega = 2\pi f = \sqrt{\frac{k}{m}}
+    \end{equation}
+
+\end{definition}
+
+\begin{example}
+    A simple harmonic oscillator consisting of mass \(m = 11.0\ \mathrm{kg}\) attached to a spring with spring constant \(k = 201\ \mathrm{N\,m^{-1}}\).
+At time \(t=0\ \mathrm{s}\) the oscillator is at position \(x(0) = -0.207\ \mathrm{m}\) and has velocity \(v(0) = -1.33\ \mathrm{m\,s^{-1}}\).
+Determine all coefficients of the equation describing the position \(x(t)\) of the oscillator as a function of time, assuming the offset is zero.
+
+    To solve for $A_1$ and $A_2$, while we assume $x_0 = 0$, we can use the initial conditions:
+    \begin{align*}
+        x(0) &= A_1 \cos(0) + A_2 \sin(0) = A_1 = -0.207\ \mathrm{m} \\
+        v(0) &= -A_1 \omega \sin(0) + A_2 \omega \cos(0) = A_2 \omega = -1.33\ \mathrm{m\,s^{-1}}
+    \end{align*}
+
+    We can find $\omega$ from the given $m$ and $k$:
+    $$
+    \omega = \sqrt{\frac{k}{m}} = \sqrt{\frac{201\ \mathrm{N\,m^{-1}}}{11.0\ \mathrm{kg}}} \approx 4.28\ \mathrm{rad\,s^{-1}}
+    $$
+    Therefore, we can solve for $A_2$:
+    $$
+    A_2 = \frac{v(0)}{\omega} = \frac{-1.33\ \mathrm{m\,s^{-1}}}{4.28\ \mathrm{rad\,s^{-1}}} \approx -0.311\ \mathrm{m}
+    $$
+    Thus, the equation describing the position \(x(t)\) of the oscillator as a function of time is:
+    $$
+    x(t) = -0.207 \cos(4.28 t) - 0.311 \sin(4.28 t)
+    $$
+\end{example}
+
+\begin{theorem}[A Trigonometric Identity] We can also express the solution in a more compact form using a single cosine function with a phase shift:
+\begin{equation}
+    x(t) = A \cos(\omega t + \phi)
+\end{equation}
+where:
+\begin{subequations}
+    \begin{equation}
+        A = \sqrt{A_1^2 + A_2^2}
+    \end{equation}
+    \begin{equation}
+        \phi = \tan^{-1}\left(\frac{A_2}{A_1}\right) = \tan^{-1}\left(\frac{-v(0)/\omega}{x(0)}\right)
+    \end{equation}
+\end{subequations}
+\end{theorem}
+\begin{proof}
+    Let $A = \sqrt{A_1^2 + A_2^2}$ and some $\phi$ such that $\cos(\phi) = \frac{A_1}{A}$ and $\sin(\phi) = \frac{A_2}{A}$. Then, we can rewrite our original solution as:
+    \begin{align*}
+        x(t) &= A_1 \cos(\omega t) + A_2 \sin(\omega t) \\
+        &= A \cos(\phi) \cos(\omega t) + A \sin(\phi) \sin(\omega t) \\
+        &= A [\cos(\phi) \cos(\omega t) + \sin(\phi) \sin(\omega t)] \\ 
+        &= A \cos(\omega t - (-\phi)) \qquad \text{(by the cosine addition formula)} \\
+        &= A \cos(\omega t + \phi)
+    \end{align*}
+\end{proof}
+\begin{example}
+    To determine the amplitude \(A\) and phase constant \(\phi\) for the oscillator in the previous example, we can use the values of \(A_1\) and \(A_2\) we found:
+    \begin{align*}
+        A &= \sqrt{(-0.207)^2 + (-0.311)^2} \approx 0.374\ \mathrm{m} \\
+        \phi &= \tan^{-1}\left(\frac{-0.311}{-0.207}\right) \approx 0.982 \pm n\pi \approx 4.12\ \mathrm{rad} \quad (n=1, \text{since both } A_1, A_2 < 0)
+    \end{align*}
+    Therefore, the equation describing the position \(x(t)\) of the oscillator as a function of time can also be written as:
+    $$
+    x(t) = 0.374 \cos(4.28 t + 4.12)
+    $$
+\end{example}
+
 \chapter{Modern Physics}

### 2025-09-05 YEAR1/ESC195 Lecture Notes/main.tex
pre year2
sept 5

--- 2025-04-14 ESC195 Lecture Notes/main.tex
@@
 \begin{example}
-    Let $A(t) = \int_{x_1(t)}^{x_2(t)} f(x) dx$. To obtain $A'(t)$, we first have $\Delta A$ for some $\Delta t$:
-    \begin{align*}
-        A(t + \Delta t) - A(t) &= \int_{x_1(t + \Delta t)}^{x_2(t) + \Delta t} f(x) dx - \int_{x_1(t)}^{x_2(t+ \Delta t)} f(x) dx \\
+    Let $A(t) = \int_{x_1(t)}^{x_2(t)} f(x) \, dx$. To obtain $A'(t)$, we first have $\Delta A$ for some $\Delta t$:
+    \begin{align*}
+        A(t + \Delta t) - A(t) &= \int_{x_1(t + \Delta t)}^{x_2(t + \Delta t)} f(x) \, dx - \int_{x_1(t)}^{x_2(t + \Delta t)} f(x) \, dx
     \end{align*}
@@
     \begin{align*}
-        \int b_ a f(x) dx &= f(c)(b-a) \\
-    \end{align*}
-    So we would be arriving at the chauin rule:
-    \begin{align*}
-        A'(t) &= \frac{d}{dt} \int_{x_1(t)}^{x_2(t)} f(x) dx \\
-        &= f(x_2(t))\frac{dx_2}{dt} - f(x_1(t))\frac{dx_1}{dt} + \int_{x_1(t)}^{x_2(t)} f'(x) dx
-        &= f(x_2(t))\frac{dx_2}{dt} - f(x_1(t))\frac{dx_1}{dt} + \int_{x_1(t)}^{x_2(t)} \frac{\partial f}{\partial x} dx
+        \int_a^b f(x) \, dx &= f(c)(b - a)
+    \end{align*}
+    So we would be arriving at the chain rule:
+    \begin{align*}
+        A'(t) &= \frac{d}{dt} \int_{x_1(t)}^{x_2(t)} f(x) \, dx \\
+        &= f(x_2(t)) \frac{d x_2}{dt} - f(x_1(t)) \frac{d x_1}{dt} + \int_{x_1(t)}^{x_2(t)} f'(x) \, dx \\
+        &= f(x_2(t)) \frac{d x_2}{dt} - f(x_1(t)) \frac{d x_1}{dt} + \int_{x_1(t)}^{x_2(t)} \frac{\partial f}{\partial x} \, dx
     \end{align*}
@@
 \begin{theorem}[Leibniz's Rule]
-    Given a reigion $R$ in the $xy$ plane, and a functions $\phi_1(x)$ and $\phi_2(x)$ have continuous derivatives with $x$ in $R$, and that we have continuous differentiable functions $f(x, y)$ in $R$, if:
-    \begin{align*}
-       F(x) &= \int_{\phi_1(x)}^{\phi_2(x)} f(x, y) dy \\
+    Given a region $R$ in the $xy$-plane, and functions $\phi_1(x)$ and $\phi_2(x)$ have continuous derivatives with $x$ in $R$, and that we have continuous differentiable functions $f(x, y)$ in $R$, if:
+    \begin{align*}
+       F(x) &= \int_{\phi_1(x)}^{\phi_2(x)} f(x, y) \, dy
     \end{align*}
@@
     \begin{align*}
-        F'(x) &= f(x, \phi_1(x))\frac{d\phi_1}{dx} - f(x, \phi_2(x))\frac{d\phi_2}{dx} + \int_{\phi_1(x)}^{\phi_2(x)} \frac{\partial f}{\partial x} dy
+        F'(x) &= f(x, \phi_1(x)) \frac{d \phi_1}{dx} - f(x, \phi_2(x)) \frac{d \phi_2}{dx} + \int_{\phi_1(x)}^{\phi_2(x)} \frac{\partial f}{\partial x} \, dy
     \end{align*}
@@
     Let:
-    $$ f(x) = x - 2\int_0^x e^{2(x-t)} f(t) dt $$
-    We can solve this by turing it into a differential equation:
-    \begin{align*}
-        f'(x) &= 1 - 2e^{2(x-x)}f(x) + 2\int_0^x e^{2(x-t)}f'(t) dt \\
-        &= 1 - 2\left[2\int_{x}^{0} e^{2(x-t)}f(t) dt\right] - 2f(x) \\ 
-        &= 1 - 2x + 2f(x) - 2f(x) \\
-        &= 1 - 2x \\
-        \intertext{Thus, we have:}
-        f(x) &= \int_0^x (1 - 2t) dt \\
-        &= x - x^2 \\
-        \intertext{Thus, we have:}
-        f(x) &= \frac{1}{2}x^2 - \frac{1}{3}x^3 + C \\
-        \intertext{We can find the constant $C$ by substituting $x=a$ into the original equation:}
-        f(a) &= a - 2\int_0aa e^{2(a-t)} f(t) dt = a
-        So we have:
-        c &= a^2
+    $$ f(x) = x - 2 \int_0^x e^{2(x - t)} f(t) \, dt $$
+    We can solve this by turning it into a differential equation:
+    \begin{align*}
+        f'(x) &= 1 - 2 e^{2(x - x)} f(x) + 2 \int_0^x e^{2(x - t)} f'(t) \, dt \\
+        &= 1 - 2 f(x) + 2 \int_0^x e^{2(x - t)} f'(t) \, dt \\
+        &= 1 - 2 f(x) + 2 \int_x^0 e^{2(x - t)} f'(t) \, dt \\
+        &= 1 - 2 f(x) - 2 f(x) \\
+        &= 1 - 4 f(x)
+    \end{align*}
+    Thus, we have:
+    \begin{align*}
+        f'(x) + 4 f(x) &= 1 \\
+        \intertext{Integrating factor: $e^{4x}$, so:}
+        \frac{d}{dx} (e^{4x} f) &= e^{4x} \\
+        e^{4x} f &= \frac{1}{4} e^{4x} + C \\
+        f(x) &= \frac{1}{4} + C e^{-4x} \\
+        \intertext{At $x=0$, $f(0) = 0$:}
+        0 &= \frac{1}{4} + C \\
+        C &= -\frac{1}{4} \\
+        f(x) &= \frac{1}{4} - \frac{1}{4} e^{-4x}
     \end{align*}

### 2025-09-05 YEAR2/AER210_LectureNotes/main.tex
sept 5

@@
 
-    (Incorrect) Here, we will describe it as a Type II regionwhere $0 \le y \le 1$, and $y \le x \le 1$. Thus, we can set up the double integral as follows:
+    (\xmark) Here, we will describe it as a Type II regionwhere $0 \le y \le 1$, and $y \le x \le 1$. Thus, we can set up the double integral as follows:
     $$
@@
 
-    (Correct) However, we can change the order of integration to make it a Type I regionwhere $0 \le x \le 1$, and $0 \le y \le x$. Thus, we can set up the double integral as follows:
+    (\cmark) However, we can change the order of integration to make it a Type I regionwhere $0 \le x \le 1$, and $0 \le y \le x$. Thus, we can set up the double integral as follows:
     $$

### 2025-09-05 YEAR2/CHE260_LectureNotes/main.tex
sept 5

@@
 
+\subsection{Systems and States}
+\begin{definition}[System]
+    Any piece of matter or a region of space.
+\end{definition}
+
+\begin{definition}[Boundary]
+    The boundary is the surface that separates the system from its surroundings. It can be real or imaginary, fixed or movable, and can allow or prevent the transfer of mass and energy.
+\end{definition}
+
+\textbf{Types of Systems} There are three main types of systems: closed, open, and isolated. Defined as the following:
+\begin{definition}[Closed System/Control Mass]
+    A closed system is one where mass cannot cross the boundary, but energy can. An example of a closed system is a sealed container of liquid, or a bag of gas, where we can do work on it.
+\end{definition}
+\begin{definition}[Open System/Control Volume]
+    An open system is one where both mass and energy can cross the boundary.
+An example of an open system is a flowing pipe, or a car engine, where mass (water, air) and energy (heat, work) can enter and leave the system.
+\end{definition}
+\begin{definition}[Isolated System]
+    An isolated system is one where neither mass nor energy can cross the boundary. An example of an isolated system is a thermos bottle that is perfectly insulated from its surroundings.
+\end{definition}
+\begin{definition}[Property]
+    A property is any attribute of a system that can be measured without knowing the history of the system. \\
+    (Example) Position, Temperature, Mass, Energy \\
+    (Non-example) Work, Transfer of energy (heat), Mass outside the system (denote $\delta m$)
+    \begin{equation}
+        W = \int dW \quad (\text{depends on path})
+    \end{equation}
+    Hence, to remind ourselves, we denote infinitesimal work as $\delta W$ instead of $dW$. Then, we can compute the work done on a system as:
+    \begin{equation}
+        \delta W = \int_C F \, dx 
+    \end{equation}
+\end{definition}
+
+\paragraph{Intensive and Extensive Properties} Properties can be classified into two categories: intensive and extensive.
+\begin{definition}[Extensive Property]
+    An extensive property is one that depends on the size or extent of the system. \\
+    (Example) Mass, Volume, Energy
+\end{definition}    
+\begin{definition}[Intensive Property]
+    An intensive property is one that does not depend on the size or extent of the system. \\ 
+    (Example) Temperature, Pressure, Density
+
+    We can convert any extensive property to an intensive property by dividing it by mass or volume. For example, specific volume $v$ or specific energy $u$:
+    $$
+        v = \frac{V}{m}, \quad u = \frac{U}{m}
+    $$
+\end{definition}
+
+\begin{definition}[State]
+    The state of a system is defined by its properties at a specific moment in time.
+A state is a condition of the system that can be described by a set of properties.
+For example, the state of an ideal gas can be described by its pressure, volume, and temperature.
+
+\end{definition}
+
+\begin{definition}[Steady State]
+    A system is in steady state if its properties do not change with time. In other words, the system is in a state of equilibrium, and there are no net changes in mass or energy within the system. \\
+    (Example) An overflowing bucket of water, in which case the system is unchanging but interacting with its surroundings.
+\end{definition}
+
+\begin{definition}[Equilibrium]
+    An \textbf{ISOLATED} system is in equilibrium if its properties are uniform throughout the system and do not change with time (constant). \\
+    (Example) A gas in a closed container that has been left undisturbed for a long time.
+\end{definition}
+\begin{example}
+    Consider a gas in a cylinder with a movable piston.
+If the piston is held in place, the gas is in a steady state, as its properties (pressure, volume, temperature) do not change with time.
+However, if the piston is allowed to move, the gas will expand or compress, and its properties will change with time, indicating that it is not in a steady state.
+\end{example}
+\begin{definition}[Process]
+    A process is a transformation from one state to another.
+A process can be described by the path taken by the system in the property space.
+For example, a gas can undergo a process of compression or expansion, which changes its pressure, volume, and temperature.
+\end{definition}
 \chapter{Heat Transfer}

### 2025-09-07 YEAR2/CHE260_LectureNotes/main.tex
sept 7

@@
 \chapter{Thermodynamics}
+\begin{shaded}
+    \section*{Systematic Approach to Problem Solving in Thermodynamics}
+    You should adopt a systematic approach when tackling problems to avoid getting confused by a lot
+    of information. The following steps are helpful:
 
+    \begin{enumerate}
+        \item \textbf{Find.} Read the problem and state what you are asked to find in your own words. This step
+        may seem obvious, but it is surprising how often people misunderstand the question or
+        miss essential information.
+        \item \textbf{Known.} List all the information and properties provided, as well as information about properties that remain constant (e.g.
+the volume in a isochoric process or the temperature in an isothermal process).
+        \item \textbf{Diagrams.} Draw a schematic diagram of the system.
+Mark the system boundary and decide whether the system is a control mass or volume.
+Show energy and mass transfers between the system and surroundings by arrows.
+Draw a process diagram if necessary.
+        \item \textbf{Assumptions.} Decide how you are going to model the system. List all assumptions that you make.
+        \item \textbf{Governing Equations.} Which conservation law are you going to apply?
+Depending on what you are trying to find, you may use principles of conservation of mass or energy.
+Write down the governing equations.
+        \item \textbf{Properties.} List all property values that are not given in the problem statement. This includes information extracted from tables or other sources.
+        \item \textbf{Solution.} Substitute known values of variables in the governing equations and solve them to find unknowns.
+        \item \textbf{Answer.} State the answer and confirm that it is what was asked for.
+        \item \textbf{Discussion.} Are your results reasonable? Can you draw any conclusions from them?
+    \end{enumerate}
+
+\end{shaded}
 \section{Systems and Properties in Thermodynamics}
@@
 
-    $$
+    \begin{equation}
         \Delta S = S_{\text{final}} - S_{\text{initial}} = \frac{Q_C}{T_C} - \frac{Q_H}{T_H} > 0
-    $$
+    \end{equation}
 
@@
     \end{equation}
-    Hence, to remind ourselves, we denote infinitesimal work as $\delta W$ instead of $dW$. Then, we can compute the work done on a system as:
+    Hence, to remind ourselves, we denote infinitesimal work as $\delta W$ instead of $dW$. Then, we can compute the work done by the force field $\textbf{F}$ along a curve $C$ as:
     \begin{equation}
-        \delta W = \int_C F \, dx 
+        \delta W = \int_C \textbf{F} \cdot ds
     \end{equation}
 \end{definition}
+\begin{shaded}
+Here's a quick refresher on line integrals and basic math:
+\begin{example}[Computing work]
+    \textit{The force acting on a 5 kg object varies with time as F = $(50 + 10t)\,N$, where t is the time in seconds. If the body starts from rest, how much work is done on it in 10 s?}
 
+    The object is travelling for a straight line, so we can simply parametrize the curve as $ds = v dt = r' dt$.
+We can compute the velocity $v$ by integrating the acceleration $a = F/m$.
+Firstly the work is given by:
+    $$
+        W = \int_{C} F \cdot ds = \int_0^{10} F \cdot v dt
+    $$
+    Since they are in the same direction, we can treat them as scalars. Now, we compute $v$:
+    $$
+        v = \int_0^t a dt = \int_0^t \frac{F}{m} dt = \int_0^t \frac{50 + 10t}{5} dt = 10t + t^2 + (\text{$C$ = 0 $\because$ object starts from rest})
+    $$
+    Hence, we have:
+    $$
+        W = \int_0^{10} (50 + 10t)(10t + t^2) dt = 100\,000 \text{ J}
+    $$
+\end{example}
+
+\end{shaded}
 \paragraph{Intensive and Extensive Properties} Properties can be classified into two categories: intensive and extensive.

### 2025-09-07 YEAR2/MAT292_LectureNotes/main.tex
sept 7

@@
 \paragraph{Remarks} If $g$ and $h$ are continuous, then there need not exist solution in a neighbourhood (open $\epsilon$-ball) of any point $(x_0, y_0)$ so long as $h(y_0) \neq 0$.
+\begin{example}[Proof of Uniqueness by Integrating Factor Method]
+Consider the IVP
+\[
+\frac{dx}{dt} = ax, \qquad x(0) = x_0.
+\]
+Although we know the solution is \( x(t) = x_0 e^{at} \), we will prove that this solution is unique.
 
-\begin{example}
-    Consider the IVP:
+Suppose $w(t)$ is any solution of this IVP. Define a change of variables:
 $$
-    \begin{cases}
-        \frac{dx}{dt} = ax \\
-        x(0) = x_0
-    \end{cases}
+y(t) = e^{-at} w(t).
 $$
-    then the solution is given by:
+Then, by the product rule,
 $$
-    x(t) = x_0e^{at}
+\frac{dy}{dt} = \frac{d}{dt}\!\left( e^{-at} w(t) \right) 
+= -a e^{-at} w(t) + e^{-at} \frac{dw}{dt}.
 $$
-    Is the solution unique? Suppose $x$ is any solution, then we check if $y = e^{-at}x$ is constant:
-    \begin{align*}
-        \frac{dy}{dt} &= \frac{d}{dt}(e^{-at}x) \\
-        &= -ae^{-at}x + e^{-at}\frac{dx}{dt} \\
-        &= -ae^{-at}x + e^{-at}(ax) \\
-        &= 0
-    \end{align*}
-    Thus, $y$ is constant, and since $y(0) = x_0$, we have $y(t) = x_0$ for all $t$. Thus,
-    $$x(t) = x_0 e^{at}
+Since $\tfrac{dw}{dt} = aw(t)$, this becomes
 $$
-    which is the unique solution.
+\frac{dy}{dt} = -a e^{-at} w(t) + e^{-at} (a w(t)) = 0.
+$$
+Thus, $y(t)$ is constant for all $t$. Evaluating at $t=0$ gives
+$$
+y(0) = e^{-a \cdot 0} w(0) = x_0,
+$$
+so $y(t) \equiv x_0$. Therefore,
+$$
+e^{-at} w(t) = x_0 \quad \implies \quad w(t) = x_0 e^{at}.
+$$
+Hence, the only possible solution of the IVP is
+$$
+x(t) = x_0 e^{at}.
+$$
+This shows the solution is unique.
 \end{example}
+
+\paragraph{Remark} 
+The change of variable 
+\[
+y(t) = e^{-at} x(t)
+\]
+allows us to prove the uniqueness of the solution. 
+This is essentially the same idea as the integrating factor method.
 
@@
 \end{example}
+
 \end{document}

### 2025-09-08 YEAR2/ECE253_LectureNotes/main.tex
sept 8

@@
 \paragraph{Note} An alternative method is to devide by powers of 2.
+\begin{shaded}
+    \textbf{Fractional Numbers.
+} To represent fractional numbers in binary, we can use the method of successive multiplication by 2 (fixed point representation).
+Or we can use floating point representation, which is similar to scientific notation in decimal.
+\end{shaded}
+\subsection{Binary Arithmetic and Logic}
+\paragraph{Binary Arithmetic} Binary arithmetic is similar to decimal arithmetic, but it only uses two digits (0 and 1). Addition is associated with a sum and carry.
 
-\paragraph{Binary Arithmetic} Binary arithmetic is similar to decimal arithmetic, but it only uses two digits (0 and 1). The 0 and 1 could be associated with a xor logic.
+\paragraph{Binary Addition} The rules for binary addition are as follows: \\
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        A & B & Sum, Carry \\
+        \hline
+        0 & 0 & 0, 0 \\
+        0 & 1 & 1, 0 \\
+        1 & 0 & 1, 0 \\
+        1 & 1 & 0, 1 \\
+        \hline
+    \end{tabular}
+    \caption{Binary Addition}
+\end{table}
+
+This could be summerize as the following logic:
+\begin{equation}
+    \text{Sum} = A \oplus B, \quad \text{Carry} = A \cdot B
+\end{equation}
+\paragraph{Binary Subtraction} The rules for binary subtraction are defined using the addition of negative numbers (2's complement):
+
+\begin{definition}[Least Significant Bit (LSB) and Most Significant Bit (MSB)]
+    The least significant bit (LSB) is the rightmost bit in a binary number, while the most significant bit (MSB) is the leftmost bit.`'
+
+\end{definition}
+\begin{definition}[2's Complement]
+    The 2's complement of a binary number is obtained by inverting all the bits (1's complement) and adding 1 to the least significant bit (LSB).
+\end{definition}
+
+\begin{example}[Number Inversion]
+    To find the 2's complement of the binary number (10110010)$_2$:
+    \begin{enumerate}
+        \item Invert all the bits: (01001101)$_2$
+        \item Add 1 to the LSB: 
+        \[
+        \begin{array}{c}
+        \quad 01001101 \\
+        +\, 00000001 \\
+        \hline
+        \quad 01001110 \\
+        \end{array}
+        \]
+    \end{enumerate}
+    Therefore, the 2's complement of (10110010)$_2$ is (01001110)$_2$.
+\end{example}
+
+\section{Logic Circuits}
+\begin{definition}[Logic Function]
+    A logic function $L: \{0,1\}^n \rightarrow \{0,1\}$ is a mathematical function that takes $n$ binary inputs and produces a single binary output based on a set of rules.
+\end{definition}
+
+\begin{definition}[Truth Table]
+    A truth table is a tabular representation of a logic function that lists all possible combinations of input values and their corresponding output values.
+
+\end{definition}
+
+\begin{definition}[Boolean Algebra]
+    Boolean algebra is a branch of algebra that deals with binary variables and logical operations.
+It provides a set of rules and properties for manipulating and simplifying logic functions.
+The specific rules and properties would be covered in later lectures.
+\end{definition}
+\subsection{Transistors as Switches}
+\begin{definition}[Transistor]
+    Transistor operates as a switch.
+The switch is open only when the gate is high.
+We denote the state of the gate as $x \in \{0,1\}$, where 0 is LO and 1 is HI.
+If input end of the swtich is HI, the output end could be modeled by the logic function:
+    $$
+        L(x) = x
+    $$
+\end{definition}
+
+\begin{example}[Serial Transistors]
+    Consdier two transistors connected in series, with the input end of the first transistor connected to HI. The output end of the second transistor can be modeled by the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|}
+            \hline
+            $x_1$ & $x_2$ & $L(x_1, x_2)$ \\
+            \hline
+            0 & 0 & 0 \\
+            0 & 1 & 0 \\
+            1 & 0 & 0 \\
+            1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Two Transistors in Series}
+        \label{tab:and_gate}
+    \end{table}
+    \\
+    The logic function can be expressed as:
+    $$
+        L(x_1, x_2) = x_1 \cdot x_2
+    $$
+    where $\cdot$ denotes the AND operation.
+\end{example}
+
+\begin{example}[Parallel Transistors]
+    Consider two transistors connected in parallel, with the input end of both transistors connected to HI. The output end can be modeled by the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|}
+            \hline
+            $x_1$ & $x_2$ & $L(x_1, x_2)$ \\
+            \hline  
+            0 & 0 & 0 \\
+            0 & 1 & 1 \\
+            1 & 0 & 1 \\
+            1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Two Transistors in Parallel}
+        \label{tab:or_gate}
+    \end{table}
+    \\
+    The logic function can be expressed as:
+    $$
+        L(x_1, x_2) = x_1 + x_2
+    $$
+    where $+$ denotes the OR operation.
+\end{example}
+
+\begin{example}
+    Consider a circuit with a transistor connected to LO and the output end connected to LO, The other ends of the output and the transistor are connected together to a HI (and a resistor).
+The output end of the circuit can be modeled by the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|}
+            \hline
+            $x$ & $L(x)$ \\
+            \hline
+            0 & 1 \\
+            1 & 0 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for a Transistor Connected to LO}
+        \label{tab:not_gate}
+    \end{table}
+    \\
+    The logic function can be expressed as:
+    $$
+        L(x) = \overline{x}
+    $$
+    where $\overline{x}$ denotes the NOT operation.
+\end{example}
+
+\subsection{Basic Logic Gates}
+
+\begin{definition}[AND Gate]
+    An AND gate outputs 1 only if all inputs are 1. The truth table for a 2-input AND gate is shown in Table \ref{tab:and_gate}.
+    \\
+    The logic function for an AND gate with inputs $A$ and $B$ can be expressed as:
+    $$
+        L(A, B) = A \cdot B = AB 
+    $$
+    \textbf{Note} when no operator is present, it is assumed to be AND.
+
+    The digital logic symbol for an AND gate is shown below:
+    \begin{center}
+        \begin{tikzpicture}[scale=1]
+            % Simple custom AND gate to avoid reliance on gate shape anchors
+            % Input nodes
+            \node (A) at (-1,0.3) {$A$};
+            \node (B) at (-1,-0.3) {$B$};
+            % Wires to gate
+            \draw (A.east) -- (-0.2,0.3);
+            \draw (B.east) -- (-0.2,-0.3);
+            % Gate body: rectangle on the left and semicircle on the right
+            \draw (-0.2,0.6) -- (-0.2,-0.6) -- (0.2,-0.6) arc (-90:90:0.6) -- cycle;
+            % Output
+            \node (Y) at (0.9,0) {$Y$};
+            \draw (0.6,0) -- (Y.west);
+            \draw (0.2,0) -- (0.6,0);
+        \end{tikzpicture}
+    \end{center}    
+\end{definition}
+
+\begin{definition}[OR Gate]
+    An OR gate outputs 1 if at least one input is 1. The truth table for a 2-input OR gate is shown in Table \ref{tab:or_gate}.
+    \\
+    The logic function for an OR gate with inputs $A$ and $B$ can be expressed as:
+    $$
+        L(A, B) = A + B
+    $$
+    The digital logic symbol for an OR gate is shown below:
+    \begin{center}
+        \begin{tikzpicture}[scale=1]
+            % Simple custom OR gate to avoid reliance on gate shape anchors
+            % Input nodes
+            \node (A) at (-1,0.3) {$A$};
+            \node (B) at (-1,-0.3) {$B$};
+            % Wires to gate
+            \draw (A.east) -- (-0.2,0.3);
+            \draw (B.east) -- (-0.2,-0.3);
+            % Gate body: curved shape
+            \draw (-0.2,0.6) .. controls (0,0.6) and (0.4,0.3) .. (0.6,0) .. controls (0.4,-0.3) and (0,-0.6) .. (-0.2,-0.6) -- cycle;
+            % Output
+            \node (Y) at (0.9,0) {$Y$};
+            \draw (0.6,0) -- (Y.west);
+        \end{tikzpicture}
+    \end{center}
+\end{definition}
+
+\begin{definition}[NOT Gate]
+    A NOT gate outputs the inverse of the input. The truth table for a NOT gate is shown in Table \ref{tab:not_gate}.
+    \\
+    The logic function for a NOT gate with input $A$ can be expressed as:
+    $$
+        L(A) = \overline{A}
+    $$
+    The digital logic symbol for a NOT gate is shown below:
+    \begin{center}
+        \begin{tikzpicture}[scale=1]
+            % Input node
+            \node (A) at (-1,0) {$A$};
+            % Wire to gate
+            \draw (A.east) -- (-0.2,0);
+            % Gate body: triangle with circle at output
+            \draw (-0.2,0.5) -- (0.2,0) -- (-0.2,-0.5) -- cycle;
+            \draw (0.2,0) circle (0.1);
+            % Output
+            \node (Y) at (0.6,0) {$Y$};
+            \draw (0.3,0) -- (Y.west);
+        \end{tikzpicture}
+    \end{center}
+\end{definition}
+
+\subsection{Additional Logic Gates}
+\begin{example}[XOR Operation]
+    We have two swtiches, when both switches are in the same state (both open or both closed), the output is 0.
+When the switches are in different states (one open and one closed), the output is 1.
+The truth table for this operation is shown below:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|}      
+            \hline
+            $x_1$ & $x_2$ & $L(x_1, x_2)$ \\
+            \hline
+            0 & 0 & 0 \\
+            0 & 1 & 1 \\
+            1 & 0 & 1 \\
+            1 & 1 & 0 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for XOR Operation}
+        \label{tab:xor_gate}
+    \end{table}
+    \\
+    The logic function can be expressed as:
+    $$
+        L(x_1, x_2) = x_1 \oplus x_2 = \overline{x_1}x_2 + x_1\overline{x_2}
+    $$
+    where $\oplus$ denotes the XOR operation.
+\end{example}
+
+\begin{definition}[XOR Gate]
+    An XOR gate outputs 1 if the inputs are different. The truth table for a 2-input XOR gate is shown in Table \ref{tab:xor_gate}.
+    \\
+    The logic function for an XOR gate with inputs $A$ and $B$ can be expressed as:
+    $$
+        L(A, B) = A \oplus B = \overline{A}B + A\overline{B}
+    $$
+    where $\oplus$ denotes the XOR operation.
+    The digital logic symbol for an XOR gate is shown below:
+    \begin{center}
+        \begin{tikzpicture}[scale=1]
+            % Input nodes
+            \node (A) at (-1,0.3) {$A$};
+            \node (B) at (-1,-0.3) {$B$};
+            % Wires to gate
+            \draw (A.east) -- (-0.2,0.3);
+            \draw (B.east) -- (-0.2,-0.3);
+            % Gate body: curved shape with extra curve on input side
+            \draw (-0.4,0.6) .. controls (-0.2,0.6) and (0.2,0.3) .. (0.4,0) .. controls (0.2,-0.3) and (-0.2,-0.6) .. (-0.4,-0.6) -- cycle;
+            \draw (-0.5,0.6) .. controls (-0.3,0.6) and (0.1,0.3) .. (0.3,0) .. controls (0.1,-0.3) and (-0.3,-0.6) .. (-0.5,-0.6);
+            % Output
+            \node (Y) at (0.9,0) {$Y$};
+            \draw (0.4,0) -- (Y.west);
+        \end{tikzpicture}
+    \end{center}
+\end{definition}
+
+In addition, we have the following gates:
+\begin{definition}[NAND Gate]
+    A NAND gate outputs 0 only if all inputs are 1. The truth table for a 2-input NAND gate is shown in Table \ref{tab:nand_nor_gates}.
+    \\
+    The logic function for a NAND gate with inputs $A$ and $B$ can be expressed as:
+    $$
+        L(A, B) = \overline{A \cdot B} = \overline{A} + \overline{B}
+    $$
+    The digital logic symbol for a NAND gate is shown below:
+    \begin{center}
+        \begin{tikzpicture}[scale=1]
+            % Input nodes
+            \node (A) at (-1,0.3) {$A$};
+            \node (B) at (-1,-0.3) {$B$};
+            % Wires to gate
+            \draw (A.east) -- (-0.2,0.3);
+            \draw (B.east) -- (-0.2,-0.3);
+            % Gate body: rectangle on the left and semicircle on the right with circle at output
+            \draw (-0.2,0.6) -- (-0.2,-0.6) -- (0.2,-0.6) arc (-90:90:0.6) -- cycle;
+            \draw (0.2,0) circle (0.1);
+            % Output
+            \node (Y) at (0.9,0) {$Y$};
+            \draw (0.6,0) -- (Y.west);
+            \draw (0.2,0) -- (0.6,0);
+        \end{tikzpicture}
+    \end{center}
+\end{definition}
+
+\begin{definition}
+
+\end{definition}
+
+\begin{shaded}
+    \paragraph{NAND and NOR Gates are Cheaper} NAND gates and NOR gates are cheaper than AND and OR gates because they require fewer transistors to implement.
+A 2-input NAND gate can be implemented using 4 transistors, while a 2-input AND gate requires 6 transistors (4 for the NAND gate and 2 for the NOT gate).
+The same applies to NOR and OR gates.
+
+    \paragraph{NAND and NOR Gates are Universal (Functionally Complete)} Additionally, NAND and NOR gates are universal gates, meaning that any logic function can be implemented using only NAND or NOR gates.
+
+    This makes them more versatile and cost-effective for building complex digital circuits.
+\end{shaded}
+
+\paragraph{Commonly Used Logic Operators} Below is a table summarizing the commonly used logic operators:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        Operator & Symbol & Description \\
+        \hline
+        AND & $\cdot$ or no symbol & Outputs 1 if all inputs are 1 \\
+        OR & $+$ & Outputs 1 if at least one input is 1 \\
+        NOT & $\overline{x}$ & Outputs the inverse of the input \\
+        inverse & $'$ or $\sim$ & Outputs the 2's complement of the input \\
+        XOR & $\oplus$ & Outputs 1 if inputs are different \\
+        NAND & $\overline{\cdot}$ & Outputs 0 if all inputs are 1 \\
+        NOR & $\overline{+}$ & Outputs 0 if at least one input is 1 \\
+        XNOR & $\overline{\oplus}$ & Outputs 1 if inputs are the same \\
+        \hline
+    \end{tabular}
+    \caption{Commonly Used Logic Operators}
+\end{table}
+
 \section{Digital Storage Elements}
@@
 \end{enumerate}
-
+\end{shaded}
 The assembly code is then assembled into machine code, which is a series of 0s and 1s that the computer can understand.
 The assembly would be specific to the architecture of the computer (machine dependent), which is why we have different assembly languages for different architectures (e.g., x86, RISC-V, ARM).
-\end{shaded}
 \section{Computer Organization}

### 2025-09-09 YEAR2/AER210_LectureNotes/main.tex
sept 9

@@
 
-\subsection{Formal Definition of Double Integrals}
+\subsubsection{Formal Definition of Double Integrals}
 There is two definitions of double integrals in this course, due to the discrepancy between Stewart's book and the lectures.
@@
 \begin{shaded}
-    \subsubsection*{\textbf{Review.  } Formal Definition of Definite Integral (Single Variable)}
+    \textbf{Review.  Formal Definition of Definite Integral (Single Variable)}
 
@@
 
+\begin{shaded}
+The following is the analogue of the squeeze theorem for double integrals:
+\begin{definition}[Squeeze Theorem for Double Integrals]
+    For the first defintion Consider region $R$ subdivided into $N$ subregions $R_1, R_2, \ldots, R_N$, such that all subregions $\bigcup_{i=1}^N R_i \subset R$ (They are all inside).
+For both cases, we require that $R_i \cap R_j = \emptyset$ for $i \neq j$, and then some of the area would be omited and the following would be garanteed:
+    $$
+        \sum_{i=1}^N \Delta A \le \text{Area}(R), \quad \sum_{i=1}^N m_i \Delta A_i \le \iint_R f(x,y) \, dA 
+    $$ 
+    where $m_i$ and $M_i$ are the minimum and maximum values of $f(x,y)$ in the $i$-th subregion.
+Similarly, if $\bigcup_{i=1}^N R_i \supset R$ (They all cover $R$), and that we garantee that $R_i \cap R \neq \emptyset$ for all $i$.
+Then, some of the area would be double counted and the following would be garanteed:
+    $$
+        \sum_{i=1}^N \Delta A \ge \text{Area}(R), \quad \sum_{i=1}^N M_i \Delta A_i \ge \iint_R f(x,y) \, dA
+    $$ 
+    For the second definition, the same logic applies, but we consider subrectangles that creates grid that is either inside or covering $R$.
+\end{definition}
+
+\end{shaded}
+\begin{example}
+    Estimate the volume that lies above the square $R = [0,2] \times [0,2]$ and below the surface $z = f(x,y) = 16 - x^2 - 2y^2$ by dividing the $R$ into four subrectangles of equal area and using the value of the function at the upper right corner of each subrectangle to form a Riemann sum.
+Choose the upper right corner of each subrectangle as the sample point.
+
+    We divide the square $R$ into four subrectangles, each with an area of 1. We obtain the sum:
+    $$
+        V \approx \sum_{i=1}^2 \sum_{j=1}^2 f(x_i^*,y_j^*) \Delta A = \sum_{i=1}^2 \sum_{j=1}^2 f(i,j) \cdot 1
+    $$
+    where the sample points are $(1,1)$, $(1,2)$, $(2,1)$, and $(2,2)$. Evaluating the function at these points gives:
+    $$
+        V \approx f(1,1) + f(1,2) + f(2,1) + f(2,2) = 34
+    $$
+    Therefore, the estimated volume is approximately 34.
+\end{example}
+
+\subsection{Double Integrals in Non-Rectangular Regions}
+
+\begin{theorem}[Change of Variable to Polar Coordinates]
+    Consider the double integral of a function $f(x,y)$ over a region $R$ in the $xy$-plane.
+If we change the variables from Cartesian coordinates $(x,y)$ to polar coordinates $(r,\theta)$ using the transformations:
+    $$
+        x = r \cos{\theta}, \quad y = r \sin{\theta} \implies r = \sqrt{x^2 + y^2}
+    $$
+    then the double integral can be expressed in polar coordinates as follows:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \iint_{R'} f(r \cos{\theta}, r \sin{\theta}) \, r \, dr \, d\theta
+    \end{equation}
+    where $R'$ is the corresponding region in the $r\theta$-plane, and the term $r$ arises from the Jacobian determinant of the transformation from Cartesian to polar coordinates.
+\end{theorem}
+\begin{proof}
+   \textbf{Note.
+} \textit{This change of variable can be derived using the Jacobian determinant of the transformation from Cartesian to polar coordinates, which will be covered in Section \ref{sec:change_of_variables}, which can fully prove this theorem in the case where $g < 0$ for some input.}
+
+   \textbf{Geometric Sketch: } Assume $f(r \cos{\theta}, r \sin{\theta}) = g(r, \theta) \ge 0$.
+Consider a small rectangle $\Delta A_i$ in the $xy$-plane with dimensions $\Delta x_i$ and $\Delta y_i$.
+When we transform this rectangle into polar coordinates, it becomes a small sector of a circle with radius $r_i$ and angle $\Delta \theta_i$.
+The area of this sector is given by:
+    $$
+          \Delta A_i = r_i \Delta r_i \Delta \theta_i  \left( 1 + \frac{\Delta r_i}{2r_i} \right)
+    $$
+    this is derived from the geormetric formula of the area of a sector of a circle. Then, as $\Delta r_i \to 0$, the term $\frac{\Delta r_i}{2r_i} \to 0$, and we have:
+    $$
+        \Delta A_i \approx r_i \Delta r_i \Delta \theta_i
+    $$
+    Therefore, the double integral in polar coordinates can be approximated as:
+    $$
+        \iint_R f(x,y) \, dA \approx \sum_{i=1}^n f(r_i \cos{\theta_i}, r_i \sin{\theta_i}) r_i \Delta r_i \Delta \theta_i
+    $$
+    Taking the limit as the maximum diameter of the subrectangles approaches zero, we obtain the exact double integral in polar coordinates.
+\end{proof}
+
+\begin{definition}[Reigion Defined by Varying $r$ with $\theta$]
+    Consider a region $R$ in the $xy$-plane that can be described in polar coordinates by the inequalities:
+    $$
+        \alpha \le \theta \le \beta, \quad g_1(\theta) \le r \le g_2(\theta)
+    $$
+    where $g_1(\theta)$ and $g_2(\theta)$ are continuous functions on the interval $[\alpha, \beta]$.
+Then, to evaluate the double integral over this region for a continuous function $f(x,y)$, we set up the integral as follows:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \int_\alpha^\beta \int_{g_1(\theta)}^{g_2(\theta)} f(r \cos{\theta}, r \sin{\theta}) \, r \, dr \, d\theta
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Reigion Defined by Varying $\theta$ with $r$]
+    Similarly, consider a region $R$ in the $xy$-plane that can be described in polar coordinates by the inequalities:
+    $$
+        a \le r \le b, \quad h_1(r) \le \theta \le h_2(r)
+    $$
+    where $h_1(r)$ and $h_2(r)$ are continuous functions on the interval $[a, b]$.
+Then, to evaluate the double integral over this region for a continuous function $f(x,y)$, we set up the integral as follows:
+    \begin{equation}
+        \iint_R f(x,y) \, dA = \int_a^b \int_{h_1(r)}^{h_2(r)} f(r \cos{\theta}, r \sin{\theta}) \, r \, d\theta \, dr
+    \end{equation}
+\end{definition}
+
+\begin{shaded}
+    \textbf{When to Use Polar Coordinates}
+
+    Polar coordinates are particularly useful for regions with circular or radial symmetry, as they simplify integration by transforming variables into a more natural form.
+They are also advantageous for integrands that are difficult in Cartesian coordinates, especially those involving terms like $x^2 + y^2$.
+\end{shaded}
+
+\begin{example}
+    Evaluate $\iint_R (3x + 4y^2) \, dA$, where $R$ is the region in the upper half-plane bounded by the circles $x^2 + y^2 = 4$ and $x^2 + y^2 = 1$ (Donut region).
+    We can describe the region $R$ in polar coordinates as $1 \le r \le 2$ and $0 \le \theta \le \pi$. Thus, we can set up the double integral as follows:
+    \begin{align*}
+        \iint_R (3x + 4y^2) \, dA &= \int_0^\pi \int_1^2 (3(r \cos{\theta}) + 4(r \sin{\theta})^2) \, r \, dr \, d\theta \\
+        &= \int_0^\pi \int_1^2 (3r^2 \cos{\theta} + 4r^3 \sin^2{\theta}) \, dr \, d\theta \\
+        &= \int_0^\pi (r^3\cos \theta + r^4 \sin^2 \theta) \Big|_{r=1}^{r=2} \, d\theta \\
+        &= \int_0^\pi (7\cos \theta + 15 \sin^2 \theta) \, d\theta = \frac{15}{2} \pi \quad \left(\text{Using } \sin^2 \theta = \frac{1 - \cos 2\theta}{2}\right)
+    \end{align*}
+\end{example}
+
+\begin{example}
+    Find the volume of the solid bounded by the $z=0$ plane and the paraboloid $z = 1 - x^2 - y^2$.
+We first consider the projection of the paraboloid onto the $xy$-plane, which is the circle $1 - x^2 - y^2 = 0$.
+We can describe the region $R$ in polar coordinates as $0 \le r \le 1$ and $0 \le \theta \le 2\pi$.
+Thus, we can set up the double integral as follows:
+    \begin{align*}
+        V &= \iint_R (1 - x^2 - y^2) \, dA \\
+        &= \int_0^{2\pi} \int_0^1 (1 - r^2) \, r \, dr \, d\theta \\
+        &= \int_0^{2\pi} \left( \int_0^1 (r - r^3) \, dr \right) d\theta \\
+        &= \int_0^{2\pi} \left( \frac{1}{2} - \frac{1}{4} \right) d\theta  = \frac{\pi}{2}
+    \end{align*}
+\end{example}
+
+\subsection{Applications of Double Integrals}
+\subsection{Surface Area}
+
+\subsection{Triple Integrals in Rectangular Coordinates}
+\subsection{Triple Integrals in Cylindrical Coordinates}
+\subsection{Triple Integrals in Spherical Coordinates}
+\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
 \chapter{Fluid Mechanics}

### 2025-09-09 YEAR2/CHE260_LectureNotes/main.tex
sept 9

@@
 The first law of thermodynamics is simple and easy to understand:
-\begin{definition}[First Law of Thermodynamics]
+\begin{definition}[First Law of Thermodynamics] \label{def:firstlaw}
     The first law of thermodynamics states that energy cannot be created or destroyed, only transformed from one form to another.
@@
     Since $Q_C = Q_H$ in a closed system and $T_H > T_C$, we have $\Delta S > 0$.
-\end{definition}
-
-\subsection{Modeling Ideal Gas}
-\begin{definition}[Ideal Gas]
-    We consider ideal gas conditions as those where the gas particles do not interact with each other except during elastic collisions, and the volume of the gas particles themselves is negligible compared to the volume of the container.
-In this course, we consider 1 atm as the threshold for ideal gas behavior.
-Of course, ideal gas obey the ideal gas law:
-    \begin{equation}
-        PV = nRT
-    \end{equation}
 \end{definition}
@@
 \end{definition}
+
+\subsection{Modeling Ideal Gas}
+\begin{definition}[Quasi-Equilibrium Process (Reversible Path)]
+    A quasi-equilibrium process is a process that occurs slowly enough that the system remains in a state of equilibrium at all times.
+In other words, the properties of the system change so (infinitely) slowly that they can be considered constant during the process.
+In this case, this path is reversible.
+\\
+    (Example) A gas in a cylinder with a movable piston that is compressed or expanded very slowly, allowing the gas to remain in equilibrium throughout the process.
+\end{definition}
+
+\begin{definition}[Real Process (Irreversible Path)]
+    A real process is a process that occurs at a finite rate, and the system may not remain in a state of equilibrium at all times.
+In other words, the properties of the system may change rapidly during the process, and the system may not be able to adjust to these changes quickly enough to maintain equilibrium.
+In this case, this path is irreversible.
+\\
+    (Example) A gas in a cylinder with a movable piston that is compressed or expanded quickly, causing the gas to deviate from equilibrium during the process.
+In that process, the work done is more than that of a quasi-equilibrium process.
+\end{definition}
+
+\begin{shaded}
+\paragraph{Motivation} We care about this because this is the process in which we ignite an engine, allowing gas to expand rapidly, pushing the piston down and doing work.
+\end{shaded}
+
+\begin{definition}[Ideal Gas Model]
+    Consider an isolated system with molecules that:
+    \begin{itemize}
+        \item Have negligible volume compared to the volume of the container. (point mass)
+        \item Do not interact with each other except during elastic collisions.
+        \item Are in constant random motion without spin or vibration.
+    \end{itemize}
+    The best example of this is a noble gas (e.g., helium, neon, argon). Since they are inert and monoatomic. For all we care about in this course all gas are ideal unless:
+    \begin{itemize}
+        \item They are at very high pressure (Condier more then 1 atm).
+        \item They are at very low temperature (Consider close to 0 K).
+    \end{itemize}
+\end{definition}
+
+Such a system is called an ideal gas.
+The macroscopic observation of pressure is the average force per unit area exerted by the gas molecules colliding with the walls of the container, so if we increase the number of molecules or decrease the volume, the pressure increases proportionally, this is Boyle's law:
+
+\begin{definition}[Boyle's Law]
+    For a given amount of gas at constant temperature, the pressure of the gas is inversely proportional to its volume:
+    \begin{equation}
+        P \propto \frac{1}{V} \quad \Rightarrow \quad PV = \text{constant}
+    \end{equation}
+    where $N$ is the number of molecules.
+We can proof the $PV$ inverse relationship a momentum conservation argument.
+Consider linear path with lenght $L$, since collisions are elastic, the change in momentum for each collision is $\Delta p = 2mv_x$.
+The time between collisions is $\Delta t = \frac{2L}{v_x}$, so the force exerted on the wall is:
+    $$
+        F = \frac{\Delta p}{\Delta t} = \frac{2mv_x}{\frac{2L}{v_x}} = \frac{mv_x^2}{L}
+    $$
+    The pressure is then given by:
+    $$
+        P = \frac{F}{A} = \frac{mv_x^2}{AL} = \frac{mv_x^2}{V}
+    $$
+    Thus, we have $PV \propto mv_x^2$.
+Now, define the RMS velocity as $v_{rms} = \sqrt{\frac{1}{N} \sum_{i=1}^N v_i^2}$, and consider random motion, we have $\overline{v_x^2} = \overline{v_y^2} = \overline{v_z^2} = \frac{1}{3} v_{rms}^2$.
+Hence, we have:
+    $$
+        PV \propto Nm \overline{v_x^2} = \frac{1}{3} Nm v_{rms}^2
+    $$
+    where $Nm$ is the total mass of the gas. Then, we complete the arguement of the gas moving in all direction.
+\end{definition}
+
+\begin{example}[Boyle's Law]
+    Consider a gas in a cylinder with a movable piston.
+If we compress the gas by pushing the piston down, the volume of the gas decreases, and the pressure increases.
+This relationship is described by Boyle's law, which states that for a given amount of gas at constant temperature, the pressure of the gas is inversely proportional to its volume:
+\end{example}
+
+We can also observe that increasing the temperature increases the average kinetic energy of the molecules, which increases the frequency and force of collisions with the walls of the container, leading to an increase in pressure, this is Charles's law:
+\begin{definition}[Charles's Law, Total Internal Energy]
+    For a given amount of gas at constant volume, the pressure of the gas is directly proportional to its absolute temperature:
+    \begin{equation}
+        P \propto T
+    \end{equation}
+    And the total internal energy of an ideal gas is the sum of the kinetic energies of all the gas molecules, which is also directly proportional to its absolute temperature:
+    \begin{equation}
+        U = \frac{3}{2} N k_B T = \frac{3}{2} n R_u T
+    \end{equation}
+    $U$ is a extensive property, so it is proportional to the number of molecules $N$ (or moles $n$). The intensive property is the specific internal energy:
+    \begin{equation}
+        u = \frac{U}{m} = \frac{3}{2} \frac{R_u}{M} T = \frac{3}{2} R T
+    \end{equation}
+    where $M$ is the molar mass of the gas, and
+    \begin{equation}
+        R = \frac{R_u}{M}
+    \end{equation}
+    is the specific gas constant.
+
+    This is a consequence of the kinetic theory of gases, which model temperature as the average kinetic energy of the gas molecules, so we have:
+    \begin{align*}
+        \frac{1}{2} mv_{rms}^2 &= \frac{3}{2} R_u T \\
+        \intertext{Dividing by $N_A$ (Avogadro's number), and define the boltzmann constant $k_B = \frac{R_u}{N_A}$, we have:}
+        \frac{1}{2} m_e v_{rms}^2 &= \frac{3}{2} k_B T 
+        \intertext{We multiply both sides by $N$ (number of molecules), and substitute into the previous equation, we have:}
+        \frac{1}{2} m_e v_{rms}^2 N &= \frac{3}{2} k_B T N \\
+    \end{align*}
+    The above is the internal energy of the gas. Thus, we can deduce the following Ideal Gas Law, which matches macroscopic observations:
+\end{definition}
+
+\begin{definition}[Ideal Gas Law]
+    Consider the combination of Boyle's and Charles's law, we have the ideal gas law:
+    \begin{equation}
+        PV = nR_uT = mRT
+    \end{equation}
+    where $R_u = 8.314\, \text{J/(mol K)}$ (or, equivalently, $R_u = 8 \times 10^3 \, \text{J/(kmol K)}$) is the universal gas constant, and $n$ is the number of moles of gas.
+$R = \frac{R_u}{M}$ is the specific gas constant, where $M$ is the molar mass of the gas.
+\end{definition}
+
+\begin{example}
+    \textit{A spherical balloon with $d = 6\, m$ is filled with helium gas at $P = 200$ kPa and $T = 20^\circ C$.
+Calculate the mass and number of moles of helium in the balloon.
+(Assume ideal gas behavior).}
+
+    We can use the ideal gas law to solve this problem. First, we need to calculate the volume of the balloon:
+    $$        
+    V = \frac{4}{3} \pi r^3 = \frac{4}{3} \pi (3)^3 = 113.1\, \text{m}^3
+    $$
+    Next, we can use the ideal gas law to calculate the number of moles of helium in the balloon:
+    $$
+        n = \frac{PV}{R_uT} = \frac{200 \times 10^3 \times 113.1}{8.314 \times (20 + 273)} = 9\,280\, \text{mol}
+    $$
+    Finally, we can calculate the mass of helium in the balloon using the molar mass of helium ($M = 4\, \text{g/mol}$):
+    $$
+        m = nM = 9\,280 \times 4  = 37\,120\, \text{g} = 37.12\, \text{kg}
+    $$
+\end{example}
+
+\begin{example}[Adiabatic Compression]
+    Consider an ideal gas undergoing adiabatic compression\footnote{This differs from \textbf{isothermal} compression, where the temperature remains constant.}. 
+    Since no heat is exchanged with the surroundings ($Q=0$), the work done on the gas increases its internal energy, which raises the temperature. 
+    Using the adiabatic relations,
+    $$
+        P V^{\gamma} = \text{constant}, 
+        \quad T V^{\gamma - 1} = \text{constant},
+    $$
+    we see that a decrease in $V$ leads to an increase in both $P$ and $T$.
+\end{example}
+
+\begin{example}[Isothermal Expansion]
+    Consider an ideal gas undergoing isothermal expansion\footnote{This differs from \textbf{adiabatic} expansion, where no heat is exchanged.}. 
+    Since the temperature remains constant ($\Delta T = 0$), the internal energy of the gas does not change ($\Delta U = 0$). 
+    The work done by the gas during expansion is equal to the heat absorbed from the surroundings ($Q = W$). 
+    Using the ideal gas law, we can see that as $V$ increases, $P$ decreases, while $T$ remains constant.
+
+\end{example}
+
+\paragraph{Note} For all ideal gas processes, the internal energy $U$ depends only on temperature $T$, not on pressure $P$ or volume $V$.
+This is a key property of ideal gases and simplifies the analysis of thermodynamic processes involving them.
+
+\begin{definition}[Change in Internal Energy]
+    The change in internal energy of an ideal gas can be calculated using the specific heat capacity at constant volume ($c_v$):
+    \begin{equation}
+        \Delta U = \frac{3}{2} mk_B \Delta T = m c_v \Delta T
+    \end{equation}
+    where $c_v$ [J kg$^{-1}$ K$^{-1}$] is the specific heat capacity at constant volume. For a monoatomic ideal gas, we have:
+    \begin{equation}
+        c_v = \frac{3}{2} R
+    \end{equation}
+    where $R$ is the specific gas constant.
+\end{definition}
+
+\begin{example}
+    \textit{How much energy is required to heat 10kg of air with 20$^\circ$C at 1 bar to 120$^\circ$ at constant volume?
+What is the final pressure?
+(Assume ideal gas behavior, and $c_{v, \text{air}} = 0.717$ kJ/kg K)}
+
+    Consider the change of internal energy:
+    $$
+        \Delta U = m c_v \Delta T = 10 \times 100 \times c_{v, \text{air}} = 1\,000 \times 0.717 = 717\, \text{kJ}
+    $$
+
+    Now consider the change of pressure, we first isolate for the control:
+    $$
+        \frac{mR}{V} = \frac{P}{T}
+    $$
+    So we can compute the final pressure:
+    $$
+        \frac{P_1}{T_1} = \frac{P_2}{T_2} \Rightarrow P_2 = P_1 \frac{T_2}{T_1} = 1 \times \frac{393}{293} = 1.34\, \text{bar}
+    $$
+\end{example}
+
+\section{First Law of Thermodynamics}
+\begin{shaded}
+\paragraph{Motivation} We want to understand how energy is conserved in a system, and how energy is transferred between a system and its surroundings.
+This is important for understanding how engines work, how to design efficient systems, and how to analyze thermodynamic processes.
+\end{shaded}  
+The same definition as before \ref{def:firstlaw}.
+
+\begin{definition}[First Law of Thermodynamics for a Closed or Isolated System]
+    For a closed system, the energy balance can be expressed as:
+    \begin{equation}
+        \Delta E_{in} - \Delta E_{out} = \Delta E_{system}
+    \end{equation}
+    where $\Delta E_{in}$ is the energy entering the system, $\Delta E_{out}$ is the energy leaving the system, and $\Delta E_{system}$ is the change in energy of the system.
+The total energy of the system can be expressed as:
+    \begin{equation}
+        E_{system} = U + KE + PE
+    \end{equation}
+    where $U$ is the internal energy, $KE$ is the kinetic energy, and $PE$ is the potential energy.
+\end{definition}
+
+\paragraph{Sign Convention} Relative to a system, we adopt the following sign convention:
+\begin{itemize}
+    \item Energy transfer to the system is positive.
+    \item Energy transfer from the system is negative.
+\end{itemize}
+
+\begin{definition}[Energy Transfer from a System]
+    The energy transfer from a system can occur in two ways: heat transfer ($Q$) and work transfer ($W$). The first law of thermodynamics can be expressed as:
+    \begin{equation}
+        dE = \delta Q + \delta W
+    \end{equation}
+\end{definition}
+
+Dividing by $dt$, we have the rate form of the first law:
+\begin{definition}[Rate Equation of the First Law of Thermodynamics]
+    \begin{equation}
+        \frac{dE}{dt} = \dot{Q} + \dot{W}
+    \end{equation}
+    where $\dot{Q}$ is the rate of heat transfer, and $\dot{W}$ is the rate of work transfer.
+\end{definition}
+
+\subsection{Work Modes}
+\begin{definition}[Boundry Work]
+    Force acts on the boundries of a systems and deformes them. \\
+    (Example) Expansion and compression of a gas
+
+    Assuming for a Quaso-equilibrium process, and no friction, we can compute the boundary work as:
+    \begin{equation}
+        \delta W = F dx  = P A dx = - P dV
+    \end{equation}
+    where $P$ is the pressure of the gas, $A$ is the area of the piston, and $dx$ is the displacement of the piston.
+The negative sign is because when the gas expands, it does work on the surroundings, so the work done by the system is negative.
+Also, we can consider the total work by:
+    \begin{equation}
+        w_{12} = -\int_{V_1}^{V_2} P dV
+    \end{equation}
+    \text{Note} Connecting that to the $PV$ diagram, the work done by the system is the area under the curve in the $PV$ diagram.
+\end{definition}
+
 \chapter{Heat Transfer}

### 2025-09-09 YEAR2/MAT292_LectureNotes/main.tex
sept 9

@@
 \end{example}
+\subsubsection{Integrating Factor Method for First-Order Linear ODEs}
+\begin{definition}[First-Order Linear ODE]
+    A first-order linear ordinary differential equation (ODE) is an equation of the form:
+    \begin{equation}
+        \frac{dy}{dx} + P(x)y = Q(x)
+    \end{equation}
+    where \( P(x) \) and \( Q(x) \) are functions of the independent variable \( x \).
+\end{definition}
+
+\begin{definition}[Integrating Factor Method]
+    To solve a first-order linear ODE of the form:
+    $$
+        \frac{dy}{dx} + P(x)y = Q(x)
+    $$
+    we can use the integrating factor method. The steps are as follows:
+    \begin{enumerate}
+        \begin{subequations}
+
+        \item Compute the integrating factor \( \mu(x) \):
+        \begin{equation}
+            \mu(x) = \exp\left( \int P(x) \, dx \right)
+        \end{equation}
+        \begin{shaded}
+            This is from the fact that:
+        \begin{align*}
+            \frac{d\mu}{dx} &= P(x) \mu(x) \quad \text{(by requirement of step \ref{step:integrating-factor-multiplication})} \\
+            \Rightarrow \frac{1}{\mu} \frac{d\mu}{dx} &= P(x) \\
+            \Rightarrow \int \frac{1}{\mu} \frac{d\mu}{dx} \, dx &= \int P(x) \, dx \\
+            \Rightarrow \ln|\mu| &= \int P(x) \, dx \\
+            \Rightarrow \mu(x) &= \exp\left( \int P(x) \, dx \right)
+        \end{align*}
+        \end{shaded}
+        \item Multiply both sides of the original ODE by the integrating factor: 
+        \begin{equation}
+            \mu(x) \frac{dy}{dx} + \mu(x) P(x)y = \mu(x) Q(x)
+        \end{equation}
+        \item Recognize that the left-hand side is the derivative of \( \mu(x)y \) \label{step:integrating-factor-multiplication}:
+        \begin{equation}
+            \frac{d}{dx}[\mu(x)y] = \mu(x) Q(x)
+        \end{equation}
+        \item Integrate both sides with respect to \( x \):
+        \begin{equation}
+            \mu(x)y = \int \mu(x) Q(x) \, dx + C
+        \end{equation}
+        where \( C \) is the constant of integration.
+        \item Finally, solve for \( y(x) \):
+        \begin{equation}
+            y(x) = \frac{1}{\mu(x)}\left( \int \mu(x) Q(x) \, dx + C \right)
+        \end{equation}
+        \end{subequations}
+    \end{enumerate}
+\end{definition}
+
+\begin{example}
+    Consider the IVP involving a first-order linear ODE:
+    $$
+    y' + \frac{2}{t} y = \frac{\sin t}{t^2}, \quad y(\pi) = 1
+    $$
+
+    To solve this IVP using the integrating factor method, we first identify \( P(t) = \frac{2}{t} \) and \( Q(t) = \frac{\sin t}{t^2} \).
+
+    Next, we compute the integrating factor:
+    \begin{align*}
+        \mu(t) &= \exp\left( \int \frac{2}{t} \, dt \right) \\
+        &= \exp\left( 2 \ln |t| \right) \\
+        &= |t|^2 = t^2
+    \end{align*}
+
+    Multiplying both sides of the original ODE by the integrating factor:
+    $$
+        t^2 y' + 2ty = \sin t
+    $$
+
+    Recognizing the left-hand side as the derivative of \( |t|^2 y \):
+    \begin{align*}
+        \frac{d}{dt}[t^2 y] &= \sin t \\
+        t^2 y &= -\cos t + C \\
+        y &= \frac{-\cos t + C}{t^2}
+    \end{align*}
+
+    To determine the constant \( C \), we use the initial condition \( y(\pi) = 1 \):
+    \begin{align*}
+        1 &= \frac{-\cos(\pi) + C}{\pi^2} \\
+        \Rightarrow C &= \pi^2 - 1
+    \end{align*}
+
+    Thus, the solution to the IVP is:
+    \begin{align*}
+        y(t) &= \frac{-\cos t + \pi^2 - 1}{t^2}
+    \end{align*}
+\end{example}
+
+\subsubsection{Exact ODEs}
+\begin{definition}[Exact ODE]
+    An ordinary differential equation (ODE) of the form:
+    \begin{subequations}
+    \begin{equation}
+        M(x, y) + N(x, y) \frac{dy}{dx} = 0 \quad \text{or equivalently} \quad M(x, y) dx + N(x, y) dy = 0
+    \end{equation}
+    is said to be exact iff there exists a function \( \psi(x, y) \) such that:
+    \begin{equation}
+        M_y = \frac{\partial M}{\partial y} = N_x = \frac{\partial N}{\partial x}
+    \end{equation}
+
+    This implies a potential function \( \psi(x, y) \) exists such that:
+    \begin{equation}
+        M_y = \psi_{xy} = \psi_{yx} = N_x
+    \end{equation}
+    The general solution to the exact ODE is given by:
+    \begin{equation}
+        d\psi(x, y) = 0 \implies \psi(x, y) = C \quad (\text{By chain rule of $\psi$})
+    \end{equation}
+    \end{subequations}
+    In this case, we can solve the ODE by finding the potential function \( \psi(x, y) \):
+    \begin{equation}
+        \psi(x, y) = \int M(x, y) \, dx + g(y)
+    \end{equation}
+    to find \( g(y) \), we can use the fact that:
+    \begin{equation}
+        \frac{\partial \psi}{\partial y} = \frac{\partial}{\partial y} \left( \int M(x, y) \, dx \right) = \int M_y \, dx = \int N_x \, dx = N(x, y)
+    \end{equation}
+    We can also solve this from the other direction. Or verify your answer this way.
+\end{definition}
+
+\begin{example}
+    Consider the exact ODE:
+    $$
+    (3x^2y^2 + x + \cos{y}) \frac{dy}{dx} = -(2xy^3 + y)
+    $$
+    We identify \( N(x, y) = 3x^2y^2 + x + \cos{y} \) and \( M(x, y) = 2xy^3 + y \). So \( M_y = N_x = 6x^2y + 1 \), thus the ODE is exact. Then we compute the potential function:
+    \begin{align*}
+        \psi(x, y) &= \int M(x, y) \, dx \\
+        &= \int (2xy^3 + y) \, dx \\
+        &= x^2 y^3 + xy + g(y)
+        \intertext{To find \( g(y) \), we compute:}
+        \frac{\partial \psi}{\partial y} &= 3x^2 y^2 + x + g'(y) \\
+        &= N(x, y) = 3x^2 y^2 + x + \cos{y} \\
+        \Rightarrow g'(y) &= \cos{y} \\
+        \Rightarrow g(y) &= \sin{y}
+        \intertext{Thus, the potential function is:}
+        \psi(x, y) &= x^2 y^3 + xy + \sin{y}
+        \intertext{The general solution to the ODE is then:}
+        \psi(x, y) &= C \\
+        \Rightarrow x^2 y^3 + xy + \sin{y} &= C
+    \end{align*}
+
+    We can verify our answer by computing from the other direction:
+    \begin{align*}
+        \psi(x, y) &= \int N(x, y) \, dy \\
+        &= \int (3x^2 y^2 + x + \cos{y}) \, dy \\
+        &= x^2 y^3 + xy + \sin{y} + h(x)
+        \intertext{To find \( h(x) \), we compute:}
+        \frac{\partial \psi}{\partial x} &= 2x y^3 + y + h'(x) \\
+        &= M(x, y) = 2xy^3 + y \\
+        \Rightarrow h'(x) &= 0 \\
+        \Rightarrow h(x) &= C'
+        \intertext{Thus, we obtain the same potential function:}
+        \psi(x, y) &= x^2 y^3 + xy + \sin{y} + C'
+    \end{align*}
+\end{example}
 

### 2025-09-09 YEAR2/PHY293_LectureNotes/main.tex
sept 9

@@
     A damped harmonic oscillator is described by the following differential equation, by adding a damping term proportional to $\dot{x}$ to the simple harmonic oscillator equation:
-    \begin{equation}
-        m \frac{d^2 x}{dt^2} + \gamma \frac{dx}{dt} + kx = 0   
+    \begin{equation} \label{eq:damped_ho}
+        m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = 0   
     \end{equation}
-    where \( \gamma \) is the damping coefficient.
+    where \( b \) is the damping coefficient.
 \end{definition}
@@
     \begin{equation}
-        m \frac{d^2 x}{dt^2} + \gamma \frac{dx}{dt} + kx = F(t)   
+        m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = F(t)   
     \end{equation}
@@
 \end{example}
-
-\begin{theorem}[A Trigonometric Identity] We can also express the solution in a more compact form using a single cosine function with a phase shift:
+\begin{theorem}[A Trigonometric Identity]
+We can also express the solution in a more compact form using a single cosine function with a phase shift:
 \begin{equation}
@@
 \end{equation}
-where:
+where
 \begin{subequations}
-    \begin{equation}
-        A = \sqrt{A_1^2 + A_2^2}
-    \end{equation}
-    \begin{equation}
-        \phi = \tan^{-1}\left(\frac{A_2}{A_1}\right) = \tan^{-1}\left(\frac{-v(0)/\omega}{x(0)}\right)
-    \end{equation}
+    \begin{align}
+        A &= \sqrt{A_1^2 + A_2^2}, \\
+        \phi &= \arctan\!\left(\frac{-A_2}{A_1}\right) 
+              = \arctan\!\left(\frac{-v(0)/\omega}{x(0)}\right).
+    \end{align}
 \end{subequations}
 \end{theorem}
+
 \begin{proof}
-    Let $A = \sqrt{A_1^2 + A_2^2}$ and some $\phi$ such that $\cos(\phi) = \frac{A_1}{A}$ and $\sin(\phi) = \frac{A_2}{A}$. Then, we can rewrite our original solution as:
+Let $A = \sqrt{A_1^2 + A_2^2}$ and choose $\phi$ such that 
+\[
+\cos(\phi) = \frac{A_1}{A}, \qquad \sin(\phi) = -\frac{A_2}{A}.
+\]
+Then, we can rewrite our original solution as
 \begin{align*}
-        x(t) &= A_1 \cos(\omega t) + A_2 \sin(\omega t) \\
-        &= A \cos(\phi) \cos(\omega t) + A \sin(\phi) \sin(\omega t) \\
-        &= A [\cos(\phi) \cos(\omega t) + \sin(\phi) \sin(\omega t)] \\ 
-        &= A \cos(\omega t - (-\phi)) \qquad \text{(by the cosine addition formula)} \\
-        &= A \cos(\omega t + \phi)
+    x(t) &= A_1 \cos(\omega t) - A_2 \sin(\omega t) \\
+         &= A \cos(\phi) \cos(\omega t) - A \sin(\phi) \sin(\omega t) \\
+         &= A \big[\cos(\phi)\cos(\omega t) - \sin(\phi)\sin(\omega t)\big] \\ 
+         &= A \cos(\omega t + \phi),
 \end{align*}
+by the cosine addition formula.
 \end{proof}
+
 \begin{example}
@@
         A &= \sqrt{(-0.207)^2 + (-0.311)^2} \approx 0.374\ \mathrm{m} \\
-        \phi &= \tan^{-1}\left(\frac{-0.311}{-0.207}\right) \approx 0.982 \pm n\pi \approx 4.12\ \mathrm{rad} \quad (n=1, \text{since both } A_1, A_2 < 0)
+        \phi &= \arctan\!\left(\frac{-(-0.311)}{-0.207}\right) \approx 4.12\ \mathrm{rad} \quad (\text{since } A_1 < 0 \text{ and } A_2 < 0)
     \end{align*}
@@
 
+\begin{definition}[The Energy of a Simple Harmonic Oscillator]
+    The total mechanical energy \(E\) of a simple harmonic oscillator is the sum of its kinetic energy \(K\) and potential energy \(U\). 
+    \begin{equation}
+        E = K + U
+    \end{equation}
+    First we consider the change of potential energy from a position $x_i$ to $x_f$, assuming the path is along the spring or the curve $C$ of the oscillator.
+The force exerted by the spring is given by Hooke's law, \( F = -kx \).
+The change in potential energy can be simply parametized and calculated as follows:
+    \begin{equation}
+        \Delta U = \int_C F\cdot ds = -\int_{x_i}^{x_f} F\,dx = \int_{x_i}^{x_f} kx\,dx = \left[\frac{1}{2}kx^2\right]_{x_i}^{x_f} = \frac{1}{2}k(x_f^2 - x_i^2)
+    \end{equation}
+
+    Therefore, the potential energy \(U\) at a position \(x\) (taking the reference point at \(x=0\)) is given by:
+    \begin{equation}
+        U(x) = \frac{1}{2}kx^2
+    \end{equation}
+
+    The kinetic energy \(K\) of the oscillator is given by:
+    \begin{equation}
+        K = \frac{1}{2}m\dot{x}^2
+    \end{equation}
+    Therefore, the total mechanical energy \(E\) of the simple harmonic oscillator is:
+    \begin{equation}
+        E = K + U = \frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2
+    \end{equation}
+    The total mechanical energy \(E\) remains constant over time, as energy is conserved in the absence of non-conservative forces (like friction or air resistance).
+
+\end{definition}
+
+\subsection{Damped Harmonic Motion}
+\begin{definition}[Damped Harmonic Motion]
+    For small velocities, the drag force is approximately proportional to the velocity and acts in the opposite direction.
+This drag force can be modeled as \( F_d = -\gamma \dot{x} \), where \( \gamma \) is the damping coefficient.
+Including this drag force in the equation of motion for a harmonic oscillator leads to the damped harmonic oscillator equation (\ref{eq:damped_ho}).
+Which could be rewritten as:
+    \begin{equation}
+        \ddot{x} + \gamma \dot{x} + \omega_0^2 x = 0
+    \end{equation}
+    where \( \omega_0 = \sqrt{\frac{k}{m}} \) is the natural angular frequency of the undamped oscillator, and \( \gamma = \frac{b}{m} \) is the damping coefficient per unit mass.
+
+    To skip the math, lets assume a solution of the form \( x(t) = e^{i\omega t} \), substituting into the differential equation gives us a formulation for \( \omega \):
+    \begin{align}
+        -\omega^2 - i\gamma\omega + \omega_0^2 &= 0 \nonumber \\
+        \omega &= -i\frac{\gamma}{2} \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} \label{eq:omega_damped}
+    \end{align}
+    Also, we can characterize the real and imaginary parts of \( \omega \) as:
+    \begin{subequations}
+        \begin{equation}
+            \omega_r = \mathrm{Re}(\omega) = \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}}
+        \end{equation}
+        \begin{equation}
+            \omega_i = \mathrm{Im}(\omega) = -\frac{\gamma}{2}
+        \end{equation}
+    \end{subequations}
+    The general solution for the damped harmonic oscillator can be written as:
+    \begin{equation}
+        x(t) = \exp(\omega_i t) \exp(-i \omega_r t) = \exp\left(-\frac{\gamma}{2}t\right) \exp(\mp i \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} t)
+    \end{equation}
+    \begin{subequations}
+    \begin{itemize}
+        \item \textbf{No Damping} (\( \gamma = 0 \)): The system behaves like a simple harmonic oscillator with angular frequency \( \omega_0 \). Given by:
+        \begin{equation}
+            z = \exp(-i\omega_0 t)
+        \end{equation}
+        \item \textbf{Underdamping} (\( 0 < \gamma < 2\omega_0 \)): The system oscillates with a gradually decreasing amplitude.
+The angular frequency of oscillation is given by \( \omega_r = \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} \).
+Given by:
+        \begin{equation}
+            z = \exp\left(-\frac{\gamma}{2}t\right) \exp(-i \omega_r t)
+        \end{equation}
+        The trigonometric form of the solution is:
+        \begin{equation}
+            x(t) = A_0 \exp\left(-\frac{\gamma}{2}t\right) \cos(\omega_r t + \phi)
+        \end{equation}
+        where \( A_0 \) and \( \phi \) are constants determined by the initial conditions
+        From this, we can derive the following cases:
+        \item \textbf{Critical Damping} (\( \gamma = 2\omega_0 \)): The system returns to equilibrium as quickly as possible without oscillating. The solution is given by:
+        \begin{equation}
+            z = \exp(-\omega_0 t)
+        \end{equation}\
+        \item \textbf{Overdamping} (\( \gamma > 2\omega_0 \)): The system returns to equilibrium without oscillating, but more slowly than in the critically damped case. The solution is given by:
+        \begin{equation}
+            z = \exp\left(-\frac{\gamma}{2}t\right) \exp\left(\sqrt{\frac{\gamma^2}{4} - \omega_0^2} t\right)
+        \end{equation}
+        So the general solution is (the solution is via a substitution of $x(t) = e^{-\gamma t/2} f(t)$ into the differential equation, which resolves the ODE to a simple form):
+        \begin{equation}
+            x(t) = A_1 \exp\left[\left(-\frac{\gamma}{2} + \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right] + A_2 \exp\left[\left(-\frac{\gamma}{2} - \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right]
+        \end{equation}
+        where \( A_1 \) and \( A_2 \) are constants determined by the initial conditions.
+    \end{itemize}
+    \end{subequations}
+\end{definition}
 \chapter{Modern Physics}

### 2025-09-11 YEAR2/ECE253_LectureNotes/main.tex
911

@@
 
-\section{Information Representation}
-
+\section{Logic Circuits}
 \subsection{Number Systems}
@@
 
-\section{Logic Circuits}
 \begin{definition}[Logic Function]
@@
 
+\subsection{Combinational Logic Circuits}
+
+\subsection{Sum of Products (SOP) Form}
+\begin{definition}[Literal]
+    A literal is a variable or its negation. For example, $A$ and $\overline{A}$ are literals.  A literal can be either true or false, and it represents a single value in a logical expression.
+
+\end{definition}
+
+\begin{definition}[Product Term]
+    A product term is a logical synonym for AND.
+\end{definition}
+
+\begin{definition}[Sum Term]
+    A sum term is a logical synonym for OR.
+\end{definition}
+
+\begin{definition}[Sum of Products (SOP) Form]
+    A logical expression is in sum of products (SOP) form if it is a sum of product terms. For example, the expression $AB + \overline{A}C + BC$ is in SOP form.
+\end{definition}
+
+\begin{definition}[Minterm]
+    A product term that evaluates to one for exactly one row of the truth table is called a minterm.
+\end{definition}
+\begin{example}[Minterm]
+    For a given truth table for $x_1, x_2, x_3$, the minterms are:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|c|}
+            \hline
+            $x_1$ & $x_2$ & $x_3$ & Minterm \\
+            \hline
+            0 & 0 & 0 & $m_0 = \overline{x_1}\overline{x_2}\overline{x_3}$ \\
+            0 & 0 & 1 & $m_1 = \overline{x_1}\overline{x_2}x_3$ \\
+            0 & 1 & 0 & $m_2 = \overline{x_1}x_2\overline{x_3}$ \\
+            0 & 1 & 1 & $m_3 = \overline{x_1}x_2x_3$ \\
+            1 & 0 & 0 & $m_4 = x_1\overline{x_2}\overline{x_3}$ \\
+            1 & 0 & 1 & $m_5 = x_1\overline{x_2}x_3$ \\
+            1 & 1 & 0 & $m_6 = x_1x_2\overline{x_3}$ \\
+            1 & 1 & 1 & $m_7 = x_1x_2x_3$ \\
+            \hline
+        \end{tabular}
+        \caption{Minterms for 3 Variables}
+        \label{tab:minterms}
+
+    \end{table}
+    \\
+    Note that each minterm corresponds to a unique combination of input values that produces an output of 1. To create the minterm, you would try to make every literal one.
+
+\end{example}
+
+\begin{definition}[Canonical SOP Form]
+    A logical expression is in canonical SOP form if it is a sum of minterms. 
+\end{definition}
+
+\subsection{Product of Sums (POS) Form}
+
+\begin{definition}[Product of Sums (POS) Form]
+    A logical expression is in product of sums (POS) form if it is a product of sum terms. For example, the expression $(A + B)(\overline{A} + C)(B + C)$ is in POS form.
+\end{definition}
+
+\begin{definition}[Maxterm]
+    A sum term that evaluates to zero for exactly one row of the truth table is called a maxterm.
+\end{definition}
+
+\begin{example}[Maxterm]
+    For a given truth table for $x_1, x_2, x_3$, the maxterms are:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|c|}
+            \hline
+            $x_1$ & $x_2$ & $x_3$ & Maxterm \\
+            \hline
+            0 & 0 & 0 & $M_0 = (x_1 + x_2 + x_3)$ \\
+            0 & 0 & 1 & $M_1 = (x_1 + x_2 + \overline{x_3})$ \\
+            0 & 1 & 0 & $M_2 = (x_1 + \overline{x_2} + x_3)$ \\
+            0 & 1 & 1 & $M_3 = (x_1 + \overline{x_2} + \overline{x_3})$ \\
+            1 & 0 & 0 & $M_4 = (\overline{x_1} + x_2 + x_3)$ \\
+            1 & 0 & 1 & $M_5 = (\overline{x_1} + x_2 + \overline{x_3})$ \\
+            1 & 1 & 0 & $M_6 = (\overline{x_1} + \overline{x_2} + x_3)$ \\
+            1 & 1 & 1 & $M_7 = (\overline{x_1} + \overline{x_2} + \overline{x_3})$ \\
+            \hline
+        \end{tabular}
+        \caption{Maxterms for 3 Variables}
+        \label{tab:maxterms}
+    \end{table}
+    \\
+    Note that each maxterm corresponds to a unique combination of input values that produces an output of 0. To create the maxterm, you would try to make every literal zero.
+\end{example}
+
+\begin{definition}[Canonical POS Form]
+    A logical expression is in canonical POS form if it is a product of maxterms.
+\end{definition}
+
+\begin{theorem}[Converting between Canonical Forms]
+    Any logical expression can be converted from canonical SOP form to canonical POS form and vice versa. For $i \in \{0, 1, \ldots, 2^n - 1\}$ and $S \subseteq \{0, 1, \ldots, 2^n - 1\}$, we have:
+    \[
+        f(x_1, x_2, \ldots, x_n) = \sum_{i \in S} m_i = \prod_{i \notin S} M_i
+    \]
+\end{theorem}
+\begin{example}
+    We have the following converison:
+    $$
+        f(x_1, x_2, x_3) = m_1 + m_3 + m_5 + m_7 = M_0 M_2 M_4 M_6
+    $$
+\end{example}
+
 \section{Digital Storage Elements}

### 2025-09-11 YEAR2/PHY293_LectureNotes/main.tex
911

@@
 \begin{align*}
-    x(t) &= A_1 \cos(\omega t) - A_2 \sin(\omega t) \\
+    x(t) &= A_1 \cos(\omega t) + A_2 \sin(\omega t) \\
          &= A \cos(\phi) \cos(\omega t) - A \sin(\phi) \sin(\omega t) \\
@@
         From this, we can derive the following cases:
-        \item \textbf{Critical Damping} (\( \gamma = 2\omega_0 \)): The system returns to equilibrium as quickly as possible without oscillating. The solution is given by:
+        \item \textbf{Critical Damping} (\( \gamma = 2\omega_0 \)): The system returns to equilibrium as quickly as possible without oscillating. Consider:
+        $$
+            x(t) = e^{-\frac{\gamma}{2}t}f(t)
+        $$
+        Inserting into the differential equation, we get:
+        $$
+            \ddot{f} + \left(\omega_0^2 - \frac{\gamma^2}{4}\right)f = 0
+        $$
+        Since \( \gamma = 2\omega_0 \), we have \( \omega_0^2 - \frac{\gamma^2}{4} = 0 \), leading to:
+        $$
+            \ddot{f} = 0 \implies f(t) = A_1 t + A_2
+        $$
+        Therefore, the general solution for the critically damped case is:
         \begin{equation}
-            z = \exp(-\omega_0 t)
-        \end{equation}\
+            x(t) = (A_1 t + A_2) \exp\left(-\frac{\gamma}{2}t\right)
+        \end{equation}
+        where \( A_1 \) and \( A_2 \) are constants determined by
         \item \textbf{Overdamping} (\( \gamma > 2\omega_0 \)): The system returns to equilibrium without oscillating, but more slowly than in the critically damped case. The solution is given by:
@@
 \end{definition}
+
+\subsection{Energy and Quality Factor}
+\begin{definition}[Energy of a Very Light Damping]
+    Consider a very lightly damped harmonic oscillator, where \( \gamma \ll \omega_0 \). In this case, the angular frequency of oscillation \( \omega_r \) can be approximated as:
+    $$
+    \omega_r \approx \omega_0 \left(1 - \frac{\gamma^2}{8\omega_0^2}\right) \approx \omega_0
+    $$
+    So the motion of the lightly damped oscillator can be approximated as:
+    $$
+    x(t) \approx A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi)
+    $$
+    Then, we can calculate the velocity of the oscillator:
+    \begin{align*}
+        \dot{x}(t) &= -\frac{\gamma}{2} A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi) - A_0 \omega_0 e^{-\frac{\gamma}{2}t} \sin(\omega_0 t + \phi) \\
+        &= A_0 \omega_0 e^{-\frac{\gamma}{2}t} \left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)
+    \end{align*}
+    The total mechanical energy \( E(t) \) of the lightly damped oscillator is given by:
+    \begin{align*}
+        E(t) &= \frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2 \\
+             &= \frac{1}{2}m \left[A_0 \omega_0 e^{-\frac{\gamma}{2}t} \left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)\right]^2 + \frac{1}{2}k \left[A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi)\right]^2 \\
+             &= \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[\left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)^2 + \cos^2(\omega_0 t + \phi)\right] \\
+             &= \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[\sin^2(\omega_0 t + \phi) + \cos^2(\omega_0 t + \phi) + \frac{\gamma^2}{4\omega_0^2} \cos^2(\omega_0 t + \phi) + \frac{\gamma}{\omega_0} \sin(\omega_0 t + \phi) \cos(\omega_0 t + \phi)\right] \\
+             &\approx \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[1 + \frac{\gamma^2}{4\omega_0^2} \cos^2(\omega_0 t + \phi)\right] \quad (\text{neglecting the small term } \frac{\gamma}{\omega_0} \sin(\omega_0 t + \phi) \cos(\omega_0 t + \phi)) \\
+             &\approx \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \quad (\text{since } \frac{\gamma^2}{4\omega_0^2} \text{ is very small}) \\
+             &= E_0 e^{-\gamma t} \quad \text{where } E_0 = \frac{1}{2}m A_0^2 \omega_0^2 \text{ is the initial energy at } t=0
+    \end{align*}
+    We can also define the time constant \( \tau \) as the time it takes for the energy to decrease to \( \frac{1}{e} \) of its initial value:
+    \begin{equation}
+        \tau = \frac{1}{\gamma}
+    \end{equation}
+    So we have, for very light damping:
+    \begin{equation}
+        E(t) = E_0 e^{-\frac{t}{\tau}}
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Rate of Energy Loss]
+    Taking the time derivative of the total mechanical energy \( E(t) \):
+    \begin{align*}
+        \frac{dE}{dt} &= \frac{d}{dt} \left(\frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2 \right) \\
+                      &= (ma + kx)\dot{x} \\
+    \end{align*}
+    For a undamped harmonic oscillator, \( ma + kx = 0 \), so \( \frac{dE}{dt} = 0 \), indicating that the total mechanical energy is conserved and obeys Hooke's law completely.
+However, for a damped harmonic oscillator, \( ma + kx = -b\dot{x} \), leading to:
+    \begin{equation}
+        \frac{dE}{dt} = -b\dot{x}^2
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Quality Factor (Q-Factor)]
+    The quality factor \( Q \) is a dimensionless parameter that characterizes the damping of a harmonic oscillator. It is defined as:
+    \begin{subequations}
+    \begin{equation}
+        Q = \frac{\omega}{\gamma} = \omega \tau
+    \end{equation}
+    And for very light damping, we can approximate \( \omega \approx \omega_0 \), leading to:
+    \begin{equation}
+        Q \approx \frac{\omega_0}{\gamma} = \omega_0 \tau
+    \end{equation}
+    \end{subequations}
+    This allows us to rewrite the equation of a damped harmonic oscillator as:
+    \begin{equation}
+        \ddot{x} + \frac{\omega_0}{Q} \dot{x} + \omega_0^2 x = 0
+    \end{equation}
+    and:
+    \begin{equation}
+        \omega = \omega_0 \sqrt{1 - \frac{1}{4Q^2}}
+    \end{equation}
+\end{definition}
 \chapter{Modern Physics}

### 2025-09-12 YEAR2/MAT292_LectureNotes/main.tex
911

@@
 
+\section{Qualitative Analysis of ODEs}
 \subsection{Fixed Points}
@@
 \end{example}
-\subsubsection{Integrating Factor Method for First-Order Linear ODEs}
+\subsubsection{First-Order Linear ODEs - Integrating Factor Method}
 \begin{definition}[First-Order Linear ODE]
@@
 
+\subsection{Bifurcations}
+\begin{definition}[Bifurcation]
+    A bifurcation in a dynamical system occurs when a small change in the system's parameters causes a sudden qualitative change in its behavior.
+
+    This can lead to the emergence or disappearance of fixed points, changes in stability, or the onset of periodic or chaotic behavior.
+\end{definition}
+
+\begin{definition}[Bifurcation Point]
+    A bifurcation point is a specific value of a parameter in a dynamical system at which the qualitative behavior of the system changes.
+At this point, the system may undergo a bifurcation, leading to changes in the number or stability of fixed points, periodic orbits, or other dynamical features.
+
+\end{definition}
+
+\paragraph{Types of Bifurcations} There are several common types of bifurcations in dynamical systems, including:
+    \begin{itemize}
+        \item Saddle-Node Bifurcation: Two fixed points (one stable and one unstable) collide and annihilate each other as a parameter is varied.
+        \item Transcritical Bifurcation: Two fixed points exchange their stability as a parameter is varied.
+        \item Pitchfork Bifurcation: A single fixed point splits into three fixed points (one stable and two unstable, or vice versa) as a parameter is varied.
+        \item Hopf Bifurcation: A fixed point loses stability and a small amplitude limit cycle (periodic orbit) emerges as a parameter is varied.
+    \end{itemize}
+
+\begin{example}
+    Consider the ODE:
+    $$
+    x' = ax(1 - x)
+    $$
+    This is a classic example of a bifurcation scenario, where the behavior of the system changes qualitatively as the parameter \(a\) is varied. Consider the fix points of the system:
+    $$
+    x^* = 0, \quad x^* = 1
+    $$
+    So, we have:
+    \begin{itemize}
+        \item For \( a < 0 \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is stable (since \( f'(0) = a < 0 \))
+            \item \( x^* = 1 \) is unstable (since \( f'(1) = -a > 0 \))
+        \end{itemize}
+        \item For \( a > 0 \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is unstable (since \( f'(0) = a > 0 \))
+            \item \( x^* = 1 \) is stable (since \( f'(1) = -a < 0 \))
+        \end{itemize}
+        \item For \( a = 0 \):
+        \begin{itemize}
+            \item Both \( x^* = 0 \) and \( x^* = 1 \) are semi-stable (since \( f'(0) = 0 \) and \( f'(1) = 0 \))
+        \end{itemize}
+        This indicates a bifurcation at \( a = 0 \), where the stability of the fixed points changes.
+    \end{itemize}
+\end{example}
+
+\begin{example}[Saddle-Node Bifurcation] \label{ex:bifurcation-saddle-node}
+    Consider the ODE:
+    $$
+    x' = x^2 + a
+    $$
+    This equation has a family of fixed points given by:
+    $$
+    x^* = \pm\sqrt{-a}
+    $$
+    For \( a < 0 \), we have two real fixed points, while for \( a > 0 \), there are no real fixed points.
+At \( a = 0 \), the fixed points collide and disappear, indicating a bifurcation point at \( a = 0 \).
+\end{example}
+
+\begin{definition}[Bifurcation Diagram]
+    A bifurcation diagram is a visual representation that illustrates how the fixed points of a dynamical system change as a parameter is varied.
+
+    It is a plot of the fixed points against the parameter, showing regions of stability and instability, and we denote stable fixed points with solid lines and unstable fixed points with dashed lines.
+A bifurcation diagram of Example \ref{ex:bifurcation-saddle-node} is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                axis lines = middle,
+                xlabel = \( a \),
+                ylabel = \( x^* \),
+                ymin = -2, ymax = 2,
+                xmin = -2, xmax = 2,
+                domain=-2:2,
+                samples=100,
+                width=10cm,
+                height=6cm,
+                grid=both,
+                minor tick num=1,
+                major grid style={line width=.2pt,draw=gray!50},
+                minor grid style={line width=.1pt,draw=gray!20},
+            ]
+            % Stable fixed points
+            \addplot[blue, thick] {-sqrt(-x)}; % Stable branch
+            \addplot[red, dashed, thick] {sqrt(-x)}; % Stable branch
+
+            % Bifurcation point
+            \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt]{};
+            \node at (axis cs:0.3,0.3) {\( (0,0) \)};
+
+            % Labels
+            \node at (axis cs:-1,1) [anchor=south west] {unstable};
+            \node at (axis cs:-1,-1) [anchor=north west] {stable};
+            \end{axis}
+        \end{tikzpicture}
+        \caption{Bifurcation Diagram for \( x' = x^2 + a \)}
+        \label{fig:bifurcation-diagram}
+    \end{figure}
+
+\end{definition}
+
+\begin{example}[Pitchfork Bifurcation]
+    Consider the ODE:
+    $$
+    x' = ax - x^3
+    $$
+    This equation has a family of fixed points given by:
+    $$
+    x^* = 0, \quad x^* = \pm\sqrt{a}
+    $$
+    For \( a \le 0 \), there is one real fixed point at \( x^* = 0 \).
+For \( a > 0 \), there are three real fixed points: \( x^* = 0 \) (unstable) and \( x^* = \pm\sqrt{a} \) (stable).
+At \( a = 0 \), the fixed point at \( x^* = 0 \) changes stability, indicating a bifurcation point at \( a = 0 \).
+
+    We can plot the bifurcation diagram for this system as follows:
+
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                axis lines = middle,
+                xlabel = \( a \),
+                ylabel = \( x^* \),
+                ymin = -2, ymax = 2,
+                xmin = -2, xmax = 2,
+                domain=-2:2,
+                samples=100,
+                width=10cm,
+                height=6cm,
+                grid=both,
+                minor tick num=1,
+                major grid style={line width=.2pt,draw=gray!50},
+                minor grid style={line width=.1pt,draw=gray!20},
+            ]
+            % Stable fixed point for a <= 0
+            \addplot[blue, thick, domain=-2:0] {0}; % Stable branch for a <= 0
+            % Unstable fixed point for a > 0
+            \addplot[red, dashed, thick, domain=0:2] {0}; % Unstable branch for a > 0
+            % Stable fixed points for a > 0
+            \addplot[blue, thick] {sqrt(x)}; % Stable branch for a > 0
+            \addplot[blue, thick] {-sqrt(x)}; % Stable
+            % Bifurcation point
+            \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt]{};
+            \node at (axis cs:0.3,0.3) {\( (0,0) \)};
+            % Labels
+            \node at (axis cs:-1,0.2) [anchor=south west] {stable};
+            \node at (axis cs:0.9,1.1) [anchor=south west] {stable};
+            \node at (axis cs:0.9,-1.1) [anchor=north west] {stable};
+            \node at (axis cs:0.5,0.1) [anchor=south west] {unstable};
+            \end{axis}
+        \end{tikzpicture}
+        \caption{Bifurcation Diagram for \( x' = ax - x^3 \)}
+        \label{fig:pitchfork-bifurcation-diagram}
+    \end{figure}
+\end{example}
+
+\begin{example}[Transcritical Bifurcation] \label{ex:bifurcation-transcritical}
+    Consider the ODE:
+    $$
+    x' = x(r - x)
+    $$
+    This equation has a family of fixed points given by:
+    $$
+    x^* = 0, \quad x^* = r
+    $$
+    For $r < 0$, \( x^* = 0 \) is stable.
+For $r > 0$, \( x^* = 0 \) is unstable and \( x^* = r \) is stable.
+At \( r = 0 \), the stability of the fixed point at \( x^* = 0 \) changes, indicating a bifurcation point at \( r = 0 \).
+The bifurcation diagram is as follows:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                axis lines = middle,
+                xlabel = \( r \),
+                ylabel = \( x^* \),
+                ymin = -2, ymax = 2,
+                xmin = -2, xmax = 2,
+                domain=-2:2,
+                samples=100,
+                width=10cm,
+                height=6cm,
+                grid=both,
+                minor tick num=1,
+                major grid style={line width=.2pt,draw=gray!50},
+                minor grid style={line width=.1pt,draw=gray!20},
+            ]
+            % Stable fixed point for r < 0
+            \addplot[blue, thick, domain=-2:0] {0}; % Stable branch for r < 0
+            % Unstable fixed point for r < 0
+            \addplot[red, dashed, thick, domain=-2:0] {x}; % Unstable branch for r < 0
+            % Unstable fixed point for r > 0
+            \addplot[red, dashed, thick, domain=0:2] {0}; % Unstable branch for r > 0
+            % Stable fixed point for r > 0
+            \addplot[blue, thick, domain=0:2] {x}; % Stable branch for r > 0
+            % Bifurcation point
+            \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt]{};
+            \node at (axis cs:0.3,0.3) {\( (0,0) \)};
+            % Labels
+            \node at (axis cs:-1,0.2) [anchor=south west] {stable};
+            \node at (axis cs:1,1) [anchor=south west] {stable};
+            \node at (axis cs:0.5,0.1) [anchor=south west] {unstable};
+            \end{axis}
+        \end{tikzpicture}
+        \caption{Bifurcation Diagram for \( x' = x(r - x) \)}
+        \label{fig:bifurcation-diagram-2}
+    \end{figure}
+\end{example}
+
+\paragraph{Linear Stability Analysis and Normal Forms} Consider the ODE in the above example \ref{ex:bifurcation-transcritical}:
+We can verify the stability of the fixed points by performing linear stability analysis.
+Consider $x \sim x^*$ and $r \sim r^*$, such that $|x - x^*| = \epsilon$ and $|r - r^*| = \delta$ for small $\epsilon, \delta > 0$.
+Then,
+\begin{align*}
+    f(x, r) &= f(x^*, r^*) + (x-x^*) f_x(x^*, r^*) + (r-r^*) f_r(x^*, r^*) + \\
+    &\quad \frac{(x-x^*)^2}{2} f_{xx}(x^*, r^*) + O(\epsilon^3, \delta^2, \epsilon\delta) \\
+    &\approx a(x - x^*)^2 + b(r - r^*)
+\end{align*}
+which we refer to as the normal form of the system. Here, \( a = \frac{1}{2} f_{xx}(x^*, r^*) \) and \( b = f_r(x^*, r^*) \). And that $f_x(x^*, r^*) = 0$ is called a saddle node.
+
+\begin{example}
+    Consider the ODE:
+    $$
+    x' = rx + x^3 - x^5
+    $$
+    We can find the fixed points by setting \( x' = 0 \), and we have
+    $$
+    x^\star = 0 \quad \text{or} \quad x^\star = \pm\sqrt{\frac{1 \pm \sqrt{1 + 4r}}{2}}
+    $$
+    Now, we see the bifurcation points are at \( r = -\frac{1}{4} \) and \( r = 0 \). We have:
+    \begin{itemize}
+        \item For \( r < -\frac{1}{4} \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is stable (since \( f'(0) = r < 0 \))
+        \end{itemize}
+        \item For \( r = -\frac{1}{4} \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is stable (since \( f'(0) = r < 0 \))
+            \item \( x^* = \pm\frac{1}{2} \) are semi-stable (since \( f'(\pm\frac{1}{2}) = 0 \))
+        \end{itemize}
+        \item For \( -\frac{1}{4} < r < 0 \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is stable (since \( f'(0) = r < 0 \))
+            \item \( x^* = -\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}} \) is stable (since \( f'(-\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}}) < 0 \))
+            \item \( x^* = \sqrt{\frac{1 + \sqrt{1 + 4r}}{2}} \) is unstable (since \( f'(\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}}) > 0 \))
+        \end{itemize}
+        \item For \( r = 0 \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is semi-stable (since \( f'(0) = 0 \))
+            \item \( x^* = -1 \) is stable (since \( f'(-1) < 0 \))
+            \item \( x^* = 1 \) is stable (since \( f'(1) < 0 \))
+        \end{itemize}
+        \item For \( r > 0 \):
+        \begin{itemize}
+            \item \( x^* = 0 \) is unstable (since \( f'(0) = r > 0 \))
+            \item \( x^* = -\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}} \) is stable (since \( f'(-\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}}) < 0 \))
+            \item \( x^* = \sqrt{\frac{1 + \sqrt{1 + 4r}}{2}} \) is stable (since \( f'(\sqrt{\frac{1 + \sqrt{1 + 4r}}{2}}) < 0 \))
+        \end{itemize}
+    \end{itemize}
+    The bifurcation diagram is as follows:
+    \begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+        \begin{axis}[
+            axis lines = middle,
+            xlabel = {$r$},
+            ylabel = {$x^\star$},
+            ymin = -2, ymax = 2,
+            xmin = -1, xmax = 1,
+            samples=200,
+            domain=-0.25:1,
+            legend style={at={(1.05,1)}, anchor=north west}
+        ]
+        % Stable and unstable branches from formula
+        \addplot[blue, thick, domain=-0.25:1] {sqrt((1 + sqrt(1+4*x))/2)};
+        \addplot[red, dashed, domain=-0.25:1] {sqrt((1 - sqrt(1+4*x))/2)};
+        \addplot[blue, thick, domain=-0.25:1] {-sqrt((1 + sqrt(1+4*x))/2)};
+        \addplot[red, dashed, domain=-0.25:1] {-sqrt((1 - sqrt(1+4*x))/2)};
+
+        % vertical black dashed line at r = -1/4
+        \draw[dashed] (axis cs:-0.25,-2) -- (axis cs:-0.25,2);
+        \node at (axis cs:-0.25,2.2) {\( r = -\frac{1}{4} \)};  
+        % The trivial equilibrium at x=0
+        \addplot[blue, thick, domain=-1:-0.0001] {0}; % stable for r<0
+        \addplot[red, dashed, domain=0:1] {0};        % unstable for r>0
+
+        % Mark bifurcation points
+        \node at (axis cs:-0.25,0.75) [circle,fill,inner sep=1.5pt]{};
+        \node at (axis cs:-0.25,-0.75) [circle,fill,inner sep=1.5pt]{};
+        \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt]{};
+
+        \legend{stable, unstable}
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Bifurcation diagram for $x' = rx + x^3 - x^5$. 
+    Blue = stable equilibria, red dashed = unstable equilibria.}
+\end{figure}
+
+\end{example}
+
 \end{document}

### 2025-09-14 YEAR2/AER210_LectureNotes/main.tex
push

@@
 
+\begin{example}
+    \textit{Find the area of the region $R$ enclosed by one petal of the rose given by $r = \cos{3\theta}$.}
+
+    We know that the petal has upper bound at $\theta = \pm \frac{\pi}{6}$.
+We can describe the region $R$ in polar coordinates as $0 \le r \le \cos{3\theta}$ and $-\frac{\pi}{6} \le \theta \le \frac{\pi}{6}$.
+Thus, we can set up the double integral as follows:
+    \begin{align*}
+        \text{Area}(R) &= \iint_R 1 \, dA \\
+        &= \int_{-\frac{\pi}{6}}^{\frac{\pi}{6}} \int_0^{\cos{3\theta}} r \, dr \, d\theta \\
+        &= \int_{-\frac{\pi}{6}}^{\frac{\pi}{6}} \left( \frac{r^2}{2} \Big|_{r=0}^{r=\cos{3\theta}} \right) d\theta \\
+        &= \int_{-\frac{\pi}{6}}^{\frac{\pi}{6}} \frac{\cos^2{3\theta}}{2} \, d\theta = \frac{\pi}{12} \quad \left(\text{Using } \cos^2{x} = \frac{1 + \cos{2x}}{2}\right)
+    \end{align*}
+\end{example}
+
+\begin{example}
+    \textit{Find the volume trapped between the cone $z = \sqrt{x^2 + y^2}$ and the sphere $x^2 + y^2 + z^2 = 1$.}
+
+    First, we find the intersection of the cone and the sphere:
+    $$
+        z = \sqrt{x^2 + y^2} \implies z^2 = x^2 + y^2
+    $$
+    Substituting into the sphere equation:
+    \begin{align*}
+        x^2 + y^2 + z^2 = 1 &\implies z^2 + z^2 = 1 \\ 
+        &\implies 2z^2 = 1 \implies z = \frac{1}{\sqrt{2}} \quad x^2 + y^2 = \frac{1}{2}
+    \end{align*}
+    Thus, we have the region $R = \{(r, \theta) \mid 0 \le r \le \frac{1}{\sqrt{2}}, 0 \le \theta \le 2\pi\}$. We can set up the double integral as follows:
+    \begin{align*}
+        V &= \iint_R (\sqrt{1 - x^2 - y^2} - \sqrt{x^2 + y^2} ) \, dA \\ % the shpere - cone
+        &= \int_0^{2\pi} \int_0^{\frac{1}{\sqrt{2}}} (\sqrt{1 - r^2} - r) \, r \, dr \, d\theta \\
+        &= 2\pi \left[ \int_0^{\frac{1}{\sqrt{2}}} r\sqrt{1 - r^2} \, dr - \int_0^{\frac{1}{\sqrt{2}}} r^2 dr \right] \\
+        &= 2\pi \left[ -\frac{1}{3} (1 - r^2)^{\frac{3}{2}} \Big|_0^{\frac{1}{\sqrt{2}}} - \frac{r^3}{3} \Big|_0^{\frac{1}{\sqrt{2}}} \right] = \frac{2\pi}{3} (1- \frac{1}{\sqrt{2}})
+    \end{align*}
+\end{example}
+
 \subsection{Applications of Double Integrals}
+\begin{shaded}
+    \subsubsection*{Review: Moment of Inertia using Single Integral}
+
+    Consider a thin rod of length $L$ with a linear mass density $\rho(x)$, where $x$ is the distance from one end of the rod.
+We can use the following formula to find the moment of inertia $I$ of the rod about an axis perpendicular to the rod and passing through one end:
+
+    \begin{equation}
+        I = \int_0^L x^2 \rho(x) \, dx
+    \end{equation}
+    where $x^2$ is the square of the distance from the axis of rotation, and $\rho(x) \, dx$ represents the mass element of the rod at position $x$.
+This comes from the definition of moment of inertia, which is the following sum with point masses:
+    $$        
+    I = \sum m_i r_i^2 \quad \text{s.t. } \quad \text{KE}= \frac{1}{2} I \omega^2
+    $$
+\end{shaded}
+
+\begin{definition}[Mass of a Lamina]
+    Consider a lamina occupying the region $R$ in the $xy$-plane with a surface mass density $\sigma(x,y)$, where $\sigma(x,y)$ is the mass per unit area at the point $(x,y)$.
+The mass $M$ of the lamina can be found using the following double integral:
+    \begin{equation}
+        M = \iint_R \sigma(x,y) \, dA
+    \end{equation}
+    where $dA$ represents an infinitesimal area element in the region $R$.
+
+\end{definition}
+
+\begin{definition}[Moment of a Lamina]
+    The moment of the lamina about the $x$-axis ($M_x$) and $y$-axis ($M_y$) can be found using the following formulas:
+    \begin{equation}
+        M_x = \iint_R y \sigma(x,y) \, dA, \quad M_y = \iint_R x \sigma(x,y) \, dA
+    \end{equation}
+    where $y$ and $x$ are the distances from the respective axes of rotation, and $\sigma(x,y) \, dA$ represents the mass element of the lamina at the point $(x,y)$.
+\end{definition}
+
+\begin{definition}[Center of Mass of a Lamina]
+    The center of mass $(\bar{x}, \bar{y})$ of the lamina can be found using the following formulas:
+    \begin{equation}
+        \bar{x} = \frac{1}{M} M_y = \frac{1}{M} \iint_R x \sigma(x,y) \, dA, \quad \bar{y} = \frac{1}{M} M_x = \frac{1}{M} \iint_R y \sigma(x,y) \, dA
+    \end{equation}
+    where $M$ is the total mass of the lamina as calculated in the previous example.
+\end{definition}
+
+\begin{example}
+    \textit{Find the cnetre of mass of the following plate with density funciton $\sigma(x,y) = x + y$ over the reigion bounded by the axis and $y = \sqrt{x}$ and $x = 1$.}
+
+    First, we find the mass of the lamina:
+    \begin{align*}
+        M &= \iint_R (x + y) \, dA \\
+        &= \int_0^1 \int_0^{\sqrt{x}} (x + y) \, dy \, dx \\
+        &= \int_0^1 \left[ xy + \frac{y^2}{2} \right]_{y=0}^{y=\sqrt{x}} \, dx \\
+        &= \int_0^1 \left( x\sqrt{x} + \frac{x}{2} \right) dx = \left[ \frac{2}{5} x^{\frac{5}{2}} + \frac{1}{4} x^2 \right]_{0}^{1} = \frac{13}{20}
+    \end{align*}
+    Now, we find the moments about the $x$-axis and $y$-axis:
+    \begin{align*}
+        M_x &= \iint_R y(x + y) \, dA \\
+        &= \int_0^1 \int_0^{\sqrt{x}} y(x + y) \, dy \, dx \\
+        &= \int_0^1 \left[ \frac{xy^2}{2} + \frac{y^3}{3} \right]_{y=0}^{y=\sqrt{x}} \, dx \\
+        &= \int_0^1 \left( \frac{x^2}{2} + \frac{x^{\frac{3}{2}}}{3} \right) dx = \left[ \frac{x^3}{6} + \frac{2}{15} x^{\frac{5}{2}} \right]_{0}^{1} = \frac{3}{10}
+    \end{align*}
+    and,
+    \begin{align*}
+        M_y &= \iint_R x(x + y) \, dA \\
+        &= \int_0^1 \int_0^{\sqrt{x}} x(x + y) \, dy \, dx \\
+        &= \int_0^1 \left[ x^2y + \frac{xy^2}{2} \right]_{y=0}^{y=\sqrt{x}} \, dx \\
+        &= \int_0^1 \left( x^2\sqrt{x} + \frac{x^{\frac{3}{2}}}{2} \right) dx = \left[ \frac{2}{7} x^{\frac{7}{2}} + \frac{1}{6} x^3 \right]_{0}^{1} = \frac{19}{42}
+    \end{align*}
+    Finally, we can find the center of mass:
+    $$        
+    \bar{x} = \frac{M_y}{M} = \frac{190}{273}, \quad \bar{y} = \frac{M_x}{M} = \frac{6}{13}
+    $$
+\end{example}
+
+\begin{definition}[Geometric Center of a Lamina]
+    If the surface mass density $\sigma(x,y)$ is constant, then the center of mass is also known as the geometric center (or centroid) of the lamina. The formulas for the geometric center are:
+    \begin{equation}
+        x_c = \frac{1}{\text{Area}(R)} \iint_R x \, dA, \quad y_c = \frac{1}{\text{Area}(R)} \iint_R y \, dA
+    \end{equation}
+    where $\text{Area}(R)$ is the area of the region $R$.
+
+\end{definition}
+
+\begin{definition}[Moment of Inertia of a Lamina]
+    The moment of inertia of the lamina about the $x$-axis ($I_x$) and $y$-axis ($I_y$) can be found using the following formulas:
+    \begin{equation}
+        I_x = \iint_R y^2 \sigma(x,y) \, dA, \quad I_y = \iint_R x^2 \sigma(x,y) \, dA
+    \end{equation}
+    where $y^2$ and $x^2$ are the squares of the distances from the respective axes of rotation, and $\sigma(x,y) \, dA$ represents the mass element of the lamina at the point $(x,y)$.
+Also for the moment of inertia about the origin ($I_o$):
+    \begin{equation}
+        I_o = \iint_R (x^2 + y^2) \sigma(x,y) \, dA
+    \end{equation}
+    In general, for an axis defined by a line $ax + by + c = 0$, the moment of inertia about that axis ($I_l$) can be found using the following formula:
+    \begin{equation}
+        I_l = \iint_R \left( \frac{ax + by + c}{\sqrt{a^2 + b^2}} \right)^2 \sigma(x,y) \, dA
+    \end{equation}
+\end{definition}
+
+\begin{example}
+    A rectangular plate of mass $m$, length $L$ and width $W$ is rotated about a vertical line on its left side with width $W$. Find the moment of inertia of the plate about this line in two cases:
+    \begin{enumerate}
+        \item The plate has uniform density $\sigma(x,y) = \frac{m}{LW}$.
+        \item The density varies at a point proportional to the square of the distance from the right most side.
+        \item It has uniform density, but rotated its center.
+    \end{enumerate}
+
+   \textbf{Solution.
+1.
+Uniform Density} We can describe the region $R$ in Cartesian coordinates as $0 \le x \le L$ and $0 \le y \le W$.
+The surface mass density is $\sigma(x,y) = \frac{m}{LW}$.
+Thus, we can set up the double integral as it rotates around the $y$-axis:
+    \begin{align*}
+        I_y &= \iint_R x^2 \sigma(x,y) \, dA \\
+        &= \int_0^W \int_0^L x^2 \cdot \frac{m}{LW} \, dx \, dy \\
+        &= \int_0^W \left[ \frac{m}{LW} \cdot \frac{x^3}{3} \right]_{x=0}^{x=L} dy = \int_0^W \frac{mL^2}{3W} dy = \left[ \frac{mL^2}{3W} y \right]_{y=0}^{y=W} = \frac{mL^2}{3}
+    \end{align*}
+    \textbf{Solution.
+2.
+Varying Density} We can describe the region $R$ in Cartesian coordinates as $0 \le x \le L$ and $0 \le y \le W$.
+The surface mass density is $\sigma(x,y) = k(L - x)^2$.
+To find the constant $k$, we have:
+    \begin{align*}
+        I_y &= \iint_R x^2 \sigma(x,y) \, dA  \\
+        &= \int_0^W \int_0^L x^2 \cdot k(L - x)^2 \, dx \, dy \\
+        &= \int_0^W \left[ k \int_0^L (x^2L^2 - 2Lx^3 + x^4) \, dx \right] dy \\
+        &= \int_0^W \left[ k \left( \frac{L^5}{3} - \frac{L^5}{2} + \frac{L^5}{5} \right) \right] dy = \int_0^W \left[ k \cdot \frac{L^5}{30} \right] dy = k \cdot \frac{L^5}{30} W
+    \end{align*}
+
+    \textbf{Solution.
+3.
+Rotated Center} We can describe the region $R$ in Cartesian coordinates as $-\frac{L}{2} \le x \le \frac{L}{2}$ and $-W \le y \le W$.
+The surface mass density is $\sigma(x,y) = \frac{m}{LW}$.
+Thus, we can set up the double integral as it rotates around the origin:
+    \begin{equation*}
+    \begin{split}
+        I_o &= \iint_R (x^2 + y^2) \sigma(x,y) \, dA \\
+        &= \int_{-W}^W \int_{-L/2}^{L/2} (x^2 + y^2) \cdot \frac{m}{LW} \, dx \, dy \\
+        &= \int_{-W}^W \left[ \frac{m}{LW} \left( \frac{x^3}{3} + y^2 x \right)_{x=-L/2}^{x=L/2} \right] dy \\
+        &= \int_{-W}^W \left[ \frac{mL^2}{12W} + \frac{my^2}{W} \right] dy \\
+        &= \left[ \frac{mL^2}{12W} y + \frac{my^3}{3W} \right]_{y=-W}^{y=W} = \frac{mL^2}{6} + \frac{2mW^2}{3}
+    \end{split}
+    \end{equation*}
+\end{example}
+
 \subsection{Surface Area}
+\begin{theorem}[Surface Area]
+    Given $z=f(x,y)$, where $f$ is a differentiable function over the region $R$ in the $xy$-plane, the surface area $S$ of the surface above the region $R$ is given by:
+    \begin{equation}
+        S = \iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
+    \end{equation}
+\end{theorem}
+\begin{proof}
+    Consider a small rectangle $\Delta A_i$ in the $xy$-plane with dimensions $\Delta x_i$ and $\Delta y_i$.
+When we project this rectangle onto the surface $z = f(x,y)$, it becomes a parallelogram $T_i$ tangent to the surface.
+Using the partial derivatives of the surface, we can deduce the two vectors that define the parallelogram:
+    $$
+        \vec{u} = \left( \Delta x_i, 0, f_x(x_i^*, y_i^*) \Delta x_i \right), \quad \vec{v} = \left( 0, \Delta y_i, f_y(x_i^*, y_i^*) \Delta y_i \right)
+    $$
+    The area of this parallelogram is given by the magnitude of the cross product of these two vectors:
+    \begin{align*}
+        \Delta T_i = ||\vec{u} \times \vec{v}|| &= \left|\begin{vmatrix}
+        \hat{i} & \hat{j} & \hat{k} \\
+        \Delta x_i & 0 & f_x(x_i^*, y_i^*) \Delta x_i \\
+        0 & \Delta y_i & f_y(x_i^*, y_i^*) \Delta y_i
+        \end{vmatrix}\right| \\
+        &= \sqrt{( -f_x(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + ( -f_y(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + (\Delta x_i \Delta y_i)^2} \\
+        &= \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta x_i \Delta y_i
+    \end{align*}
+    Therefore, the surface area can be approximated as:
+    $$
+        S \approx S_n = \sum_{i=1}^n \Delta T_i = \sum_{i=1}^n \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta A_i
+    $$
+    Taking the limit as the maximum diameter of the subrectangles approaches zero, we obtain the exact surface area:
+    $$
+        S = \lim_{|P| \to 0} S_n =\iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
+    $$
+    which completes our derivation.
+\end{proof}
+
+\begin{example}
+    \textit{Find the surface of the sphere $x^2 + y^2 + z^2 = a^2$. }
+
+    We first consider the first octant, where $x, y, z \ge 0$. Then the total volumn would be eight times the volumn of the first octant. We first express it in polar coordinates:
+    $$
+        x^2 + y^2 = r^2 \implies z = \sqrt{a^2 - r^2}
+    $$
+    taking the partial derivatives:
+    $$
+        \frac{\partial z}{\partial r} = \frac{-r}{\sqrt{a^2 - r^2}}, \quad \frac{\partial z}{\partial \theta} = 0
+    $$
+    Thus, we can set up the double integral as follows:
+    \begin{align*}
+        S_{\text{1st octant}} &= \iint_R \sqrt{1 + \left( \frac{\partial z}{\partial r} \right)^2 + \left( \frac{\partial z}{\partial \theta} \right)^2} \, dA \\
+        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \left( \frac{-r}{\sqrt{a^2 - r^2}} \right)^2 + 0^2} \, r \, dr \, d\theta \\
+        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \frac{r^2}{a^2 - r^2}} \, r \, dr \, d\theta \\
+        &= \int_0^{\frac{\pi}{2}} \int_0^a \frac{a}{\sqrt{a^2 - r^2}} \, r \, dr \, d\theta \\
+        &= \int_0^{\frac{\pi}{2}} \left[ -a \sqrt{a^2 - r^2} \right]_{r=0}^{r=a} d\theta = \int_0^{\frac{\pi}{2}} a^2 \, d\theta = \frac{\pi a^2}{2}
+    \end{align*}
+    Therefore, the total surface area of the sphere is:
+    $$        S = 8 \cdot S_{\text{1st octant}} = 4\pi a^2 $$
+
+\end{example}
 

### 2025-09-14 YEAR2/CHE260_LectureNotes/main.tex
push

@@
     \end{equation}
-    \text{Note} Connecting that to the $PV$ diagram, the work done by the system is the area under the curve in the $PV$ diagram.
+    \text{Note} Connecting that to the $PV$ diagram, the work done by the system is the area under the curve in the $PV$ diagram:
+
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+        \draw[->] (0,0) -- (5,0) node[right] {Volume (V)};
+        \draw[->] (0,0) -- (0,5) node[above] {Pressure (P)};
+        \draw[thick, domain=1:4, samples=100] plot (\x, {4/\x}) node[right] {$P V = \text{constant}$};
+        \fill[blue!20, opacity=0.5] (1,4) -- plot[domain=1:3] (\x, {4/\x}) -- (3, 0) -- (1, 0) -- cycle;
+        \node at (2,1) {$w_{12}$};
+        \node at (1,4.5) {State 1};
+        \node at (3,1.5) {State 2};
+    \end{tikzpicture}
+    \caption{Work done by the system during expansion from State 1 to State 2 is the area under the curve in the $PV$ diagram.}
+\end{figure}
+
+\end{definition}
+\newpage
+\begin{example}[Isochoric Process]
+    Consider a gas in a cylinder with a fixed piston.
+If we heat the gas, its temperature and pressure increase, but its volume remains constant.
+In this case, there is no boundary work done by the system, as the volume does not change ($dV = 0$).
+However, there may be other forms of work done on or by the system, such as electrical work if the gas is heated using an electric heater.
+We can demosrate by the below diagram:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \draw[->] (0,0) -- (5,0) node[right] {Volume (V)};
+            \draw[->] (0,0) -- (0,5) node[above] {Pressure (P)};
+            \draw[thick] (2,0.5) -- (2,4) node[above] {};    
+            \node at (2.25,2) {$w_{12} = 0$};
+            \node at (2,4.5) {State 2};
+            \node at (2,0.5) {State 1};
+        \end{tikzpicture}
+        \caption{Isochoric Process in a $PV$ Diagram}
+    \end{figure}
+
+\end{example}
+
+\begin{example}[Isobaric Process]
+    Consider a gas in a cylinder with a movable piston with mass $m$ on top.
+If we heat the gas, its temperature and volume increase, while its pressure remains constant (equal to the atmospheric pressure plus the pressure exerted by the mass on top of the piston).
+In this case, there is boundary work done by the system as the volume changes ($dV \neq 0$).
+The work done by the system can be calculated using the formula:
+    \begin{equation}
+        w_{12} = -\int_{V_1}^{V_2} P dV = -P (V_2 - V_1)
+    \end{equation}
+    where $P$ is the constant pressure of the gas, and $V_1$ and $V_2$ are the initial and final volumes of the gas, respectively. We can demonstrate by the below diagram:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \draw[->] (0,0) -- (5,0) node[right] {Volume (V)};
+            \draw[->] (0,0) -- (0,5) node[above] {Pressure (P)};
+            \draw[thick] (1,2) -- (4,2) node[right] {};    
+            \fill[blue!20, opacity=0.5] (1,0) -- (4,0) -- (4,2) -- (1,2) -- cycle;
+            \node at (3,1) {$w_{12} = -P (V_2 - V_1)$};
+            \node at (4,2.5) {State 2};
+            \node at (1,1.5) {State 1};
+        \end{tikzpicture}
+        \caption{Isobaric Process in a $PV$ Diagram}
+    \end{figure}
+\end{example}
+
+\begin{example}[Isothermal Process]
+    Consider a gas in a cylinder with a movable piston.
+If we compress the gas very slowly while keeping it in contact with a thermal reservoir, its temperature remains constant (isothermal process).
+In this case, there is boundary work done by the system as the volume changes ($dV \neq 0$).
+The work done by the system can be calculated using the formula:
+    \begin{equation}
+        w_{12} = -\int_{V_1}^{V_2} P dV = -nR_u T \ln\left(\frac{V_2}{V_1}\right)
+    \end{equation}
+    where $n$ is the number of moles of gas, $R_u$ is the universal gas constant, $T$ is the constant temperature of the gas, and $V_1$ and $V_2$ are the initial and final volumes of the gas, respectively.
+We can demonstrate by the below diagram:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \draw[->] (0,0) -- (5,0) node[right] {Volume (V)};
+            \draw[->] (0,0) -- (0,5) node[above] {Pressure (P)};
+            \draw[thick, domain=1:4, samples=100] plot (\x, {4/\x}) node[right] {$P V = \text{constant}$};
+            \fill[blue!20, opacity=0.5] (1,4) -- plot[domain=1:3] (\x, {4/\x}) -- (3, 0) -- (1, 0) -- cycle;
+            \node at (2,1) {$w_{12}$};
+            \node at (1,4.5) {State 1};
+            \node at (3,1.5) {State 2};
+        \end{tikzpicture}
+        \caption{Isothermal Process in a $PV$ Diagram}
+    \end{figure}
+
+    We can use the above equation to calculate the work done by the gas during the isothermal compression process.
+
+\end{example}
+
+\paragraph{Curve Fitting in the PV Diagram} In general, we can fit a curve in the $PV$ diagram using the polytropic process equation:
+\begin{equation}
+    P V^n = \text{constant}
+\end{equation}
+where $n$ is the polytropic index. Different values of $n$ correspond to different specific types of processes:
+\begin{itemize}
+    \item $n = 0$: Isobaric process (constant pressure)
+    \item $n = 1$: Isothermal process (constant temperature)
+    \item $n = \gamma$: Adiabatic process (no heat transfer, where $\gamma = \frac{c_p}{c_v}$ is the heat capacity ratio)
+    \item $n \to \infty$: Isochoric process (constant volume)
+\end{itemize}
+We can compute the work done by the system during a polytropic process as:
+\begin{equation}
+    w_{12} = -\int_{V_1}^{V_2} P dV = \frac{P_2 V_2 - P_1 V_1}{n - 1} \quad (n \neq 1)
+\end{equation}
+
+\begin{definition}[Flow Work]
+    Flow work is the work required to push mass into or out of a control volume. It is given by:
+    \begin{equation}
+        W_{flow} = \frac{PB}{m} = Pv
+    \end{equation}
+    where $P$ is the pressure, $B$ is the volume flow rate, and $m$ is the mass flow rate. $v$ is the specific volume (volume per unit mass).
+    \textbf{Energy flowing a system per unit mass of a fluid} is called \textbf{flow energy} or \textbf{flow work}.
+It is the energy required to push mass into or out of a control volume.
+The flow work per unit mass is given by:
+    \begin{equation}
+       H = U + PV
+    \end{equation}
+    which is called the \textbf{enthalpy} of the fluid. The enthalpy is an extensive property, so we can define the specific enthalpy as:
+    \begin{equation}
+        h = u + Pv
+    \end{equation}
 \end{definition}

### 2025-09-14 YEAR2/ECE253_LectureNotes/main.tex
push

@@
 \begin{shaded}
-    \paragraph{NAND and NOR Gates are Cheaper} NAND gates and NOR gates are cheaper than AND and OR gates because they require fewer transistors to implement.
+    \textbf{NAND and NOR Gates are Cheaper} NAND gates and NOR gates are cheaper than AND and OR gates because they require fewer transistors to implement.
 A 2-input NAND gate can be implemented using 4 transistors, while a 2-input AND gate requires 6 transistors (4 for the NAND gate and 2 for the NOT gate).
@@
 
-    \paragraph{NAND and NOR Gates are Universal (Functionally Complete)} Additionally, NAND and NOR gates are universal gates, meaning that any logic function can be implemented using only NAND or NOR gates.
+    \textbf{NAND and NOR Gates are Universal (Functionally Complete)} Additionally, NAND and NOR gates are universal gates, meaning that any logic function can be implemented using only NAND or NOR gates.
 
@@
 \end{table}
-
-\subsection{Combinational Logic Circuits}
 
@@
 
+\subsection{Boolean Algebra and Logic Minimization}
+\begin{definition}[Boolean Algebra]
+    Boolean algebra is a branch of algebra that deals with binary variables and logical operations.
+It is a effective means to describe logic circuits with a set of rules derived from the axioms of Boolean algebra.
+\end{definition}
+
+\begin{definition}[Axioms of Boolean Algebra]
+    The axioms of Boolean algebra are a set of fundamental rules that govern the behavior of binary variables and logical operations.
+The number stems consist only of the set $\{0, 1\}$, with the following axioms:
+    \begin{itemize}
+        \item $0 \cdot 0 = 0$
+        \item $1 \cdot 1 = 1$
+        \item $0 \cdot A = 0 \cdot 1 = 1 \cdot 0 = 0$ for any $A$
+        \item if $x=0$ then $\overline{x} = 1$
+    \end{itemize}
+
+    \textbf{Dual Form} We can also derive the following logical equivalences from the axioms:
+    \begin{itemize}
+        \item $A + 0 = A$
+        \item $A + 1 = 1$
+        \item $0 + 1 = 1 + 0 = 1$
+        \item $A + \overline{A} = 1$
+    \end{itemize}
+
+    where $1$ is the multiplicative identity and $0$ is the additive identity.
+\end{definition}
+
+\paragraph{Rules derived from the Axioms of Boolean Algebra} The following rules can be derived from the axioms of Boolean algebra:
+\begin{theorem}
+\begin{itemize}
+    \item $x \cdot 0 = 0$ (Annihilation)
+    \item $x \cdot 1 = 1 \cdot x = x$ (Identity)
+    \item $x \cdot \overline{x} = 0$ (Complementation)
+    \item $x \cdot x = x$ (Idempotent)
+    \item $x + 0 = 0 + x = x$ (Identity)
+    \item $x + 1 = 1 + x = 1$ (Annihilation)
+    \item $x + \overline{x} = 1$ (Complementation)
+\end{itemize}
+\end{theorem}
+
+\begin{theorem}
+The following identities can be derived from the axioms of Boolean algebra:
+\begin{itemize}
+    \item Commutative Laws:
+    \begin{itemize}
+        \item $A + B = B + A$
+        \item $A \cdot B = B \cdot A$
+    \end{itemize}
+    \item Associative Laws:
+    \begin{itemize}
+        \item $A + (B + C) = (A + B) + C$
+        \item $A \cdot (B \cdot C) = (A \cdot B) \cdot C$
+    \end{itemize}
+    \item Distributive Laws:
+    \begin{itemize}
+        \item $A \cdot (B + C) = A \cdot B + A \cdot C$
+        \item $A + (B \cdot C) = (A + B) \cdot (A + C)$
+    \end{itemize}
+\end{itemize}
+\end{theorem}
+\begin{proof}
+    By perfect induction. We can exhaustively check all possible values of $A$, $B$, and $C$ (0 or 1) to verify that both sides of each identity yield the same result.
+\end{proof}
+
+\begin{theorem}[Covering Theorem]
+     The following is true:
+    $$    
+    x + xy = x
+    $$
+    and its dual:
+    $$
+    x(x + y) = x
+    $$
+\end{theorem}
+
+\begin{theorem}[Combining Theorem]
+    The following is true:
+    $$
+    xy + x\overline{y} = x
+    $$
+    and its dual:
+    $$
+    (x + y)(x + \overline{y}) = x
+    $$
+
+\end{theorem}
+
+\begin{theorem}[De Morgan's Theorem]
+    The following is true:
+    $$
+    \overline{xy} = \overline{x} + \overline{y}
+    $$
+    and its dual:
+    $$
+    \overline{x + y} = \overline{x} \cdot \overline{y}
+    $$
+\end{theorem}
+\begin{proof}
+    By direct proof. We have:
+    \begin{align*}
+        \overline{xy} &= \overline{x}\overline{y} + \overline{x}y + x\overline{y} \quad (\text{In Canonical SOP Form}) \\
+        &= \overline{x}\overline{y} + \overline{x}y + x\overline{y} + x\overline{y} \quad (\text{Adding } x\overline{y} \text{ using $x + x = x$}) \\
+        &= \overline{x}(\overline{y} + y) + \overline{y}(x + \overline{x}) \quad (\text{Using Distributive Law}) \\
+        &= \overline{x} \cdot 1 + \overline{y} \cdot 1 \quad (\text{Using Complementation}) \\
+        &= \overline{x} + \overline{y} \quad (\text{Using Identity})
+    \end{align*}
+\end{proof}
+
+\begin{theorem}[Absorption / Redundancy Theorem]
+    The following is true:
+    $$
+    x + \overline{x}y = x + y
+    $$
+    and its dual:
+    $$
+    x(\overline{x} + y) = xy
+    $$
+\end{theorem}
+\begin{proof}
+    By direct proof. We have:
+    \begin{align*}
+        x + \overline{x}y &= x + \overline{x}y + xy \quad (\text{Adding } xy \text{ using } x + xy = x) \\
+        &= x(1 + y) + \overline{x}y \quad (\text{Using Distributive Law}) \\
+        &= x \cdot 1 + \overline{x}y \quad (\text{Using Identity}) \\
+        &= x + y \quad (\text{Using Combining Theorem})
+    \end{align*}
+\end{proof}
+
+\begin{shaded}
+    \textbf{Logic Minimization} The goal of logic minimization is to reduce the number of logic gates and inputs in a digital circuit while maintaining its functionality.
+This is important because it can lead to cost savings, improved performance, and reduced power consumption.
+Logic minimization can be achieved through various techniques, including Boolean algebra simplification, Karnaugh maps, and the Quine-McCluskey algorithm.
+\end{shaded}
+
+\begin{theorem}[Nand as SOP]
+    And SOP circuit can be implemented using only NAND gates.
+\end{theorem}
+
+\begin{theorem}[Nor as POS]
+    A POS circuit can be implemented using only NOR gates.
+\end{theorem}
+
+\begin{example}[Gumball Fact]
+    Consider three sensors $s_0, s_1, s_2$ that detect defects in Gumballs. Those sensors are normally 0, but would be 1 if a defect is detected as follows:
+    $$
+        \begin{cases}
+            s_0 = 1 & \text{if the Gumball is too small} \\
+            s_1 = 1 & \text{if the Gumball is too big} \\
+            s_2 = 1 & \text{if the Gumball is too light}
+        \end{cases}
+    $$
+    We are to design a circuit that would output 1 if the Gumball is either too large or too small and too light. We can express canonical SOP form as:
+    \begin{align*}
+        L(s_0, s_1, s_2) &= m_3 + m_4 + m_5 + m_6 + m_7 = \overline{s_0}s_1s_2 + s_2\overline{s_0}\overline{s_2} + s_0\overline{s_1}s_2 + s_0s_1\overline{s_2} + s_0s_1s_2 \\
+        \intertext{Using the Combining Theorem}
+        &= \overline{s_0}s_1s_2 + s_2\overline{s_0}\overline{s_2} + s_0s_2 + s_0s_1 \\
+        \intertext{Using the Absorption Theorem}
+        &= s_2\overline{s_0} + s_0s_2 + s_0s_1 \\
+        \intertext{Using the Covering Theorem}
+        &= s_2 + s_0s_1 
+    \end{align*}
+\end{example}
 \section{Digital Storage Elements}

### 2025-09-14 YEAR2/MAT292_LectureNotes/main.tex
push

@@
             % Stable and unstable branches from formula
-        \addplot[blue, thick, domain=-0.25:1] {sqrt((1 + sqrt(1+4*x))/2)};
-        \addplot[red, dashed, domain=-0.25:1] {sqrt((1 - sqrt(1+4*x))/2)};
-        \addplot[blue, thick, domain=-0.25:1] {-sqrt((1 + sqrt(1+4*x))/2)};
-        \addplot[red, dashed, domain=-0.25:1] {-sqrt((1 - sqrt(1+4*x))/2)};
+            \addplot[blue, thick, domain=-0.25:1] {
+                sqrt((1 + sqrt(1+4*x))/2)
+            };
+            \addplot[red, dashed, domain=-0.25:1] {
+                sqrt((1 - sqrt(1+4*x))/2)
+            };
+            \addplot[blue, thick, domain=-0.25:1] {
+                -sqrt((1 + sqrt(1+4*x))/2)
+            };
+            \addplot[red, dashed, domain=-0.25:1] {
+                -sqrt((1 - sqrt(1+4*x))/2)
+            };
 
@@
         \end{tikzpicture}
-    \caption{Bifurcation diagram for $x' = rx + x^3 - x^5$. 
-    Blue = stable equilibria, red dashed = unstable equilibria.}
+        \caption{Bifurcation diagram for $x' = rx + x^3 - x^5$.}
     \end{figure}
-
+\end{example}
+
+\paragraph{Conditions for Bifurcations} Consider $x' = f(x, r)$. Suppose ($x^c, r^c$) is a candidate for a bifurcation point. Then, the following conditions must hold:
+\begin{itemize}
+    \item \textbf{Equilibrium Condition} $f(x^c, r^c) = 0$ (i.e., $x^c$ is a fixed point for parameter value $r^c$)
+    \item \textbf{Non Hyperbolicity Condition} $f_x(x^c, r^c) = 0$ (i.e., the Jacobian at the fixed point has a zero eigenvalue)    
+    \item \textbf{Transversality Condition} $f_r(x^c, r^c) \neq 0$ (i.e., the fixed point changes with respect to the parameter)
+    \item \textbf{Types of Bifurcations} For different types of bifurcations, we have the following additional conditions:
+    \begin{itemize}
+        \item \textbf{Saddle-Node Bifurcation} $f_{xx}(x^c, r^c) \neq 0$        
+        \item \textbf{Pitchfork Bifurcation} $f_{xx}(x^c, r^c) = 0$ and $f_{xxx}(x^c, r^c) \neq 0$
+    \end{itemize}
+\end{itemize}
+\subsection{Existence and Uniqueness of Solutions}
+\begin{example}
+    Consider the following IVP:
+    $$
+    y' = y^{\frac{1}{3}}, \quad y(0) = 0
+    $$
+    Apart from the trivial solution \( y(t) = 0 \), we can also solve this ODE by separating variables:
+    \begin{align*}
+        \frac{dy}{y^{1/3}} &= dt \\
+        \int y^{-1/3} \, dy &= \int 1 \, dt \\
+        \frac{3}{2} y^{2/3} &= t + C \\
+        \intertext{Using the initial condition \( y(0) = 0 \), we find \( C = 0 \). Thus, the solution is:}
+        y(t) &= \left( \frac{2}{3}(t ) \right)^{3/2}
+    \end{align*}
+    Thus, we have two solutions to the IVP. Existence is guaranteed, but uniqueness is not.
+\end{example}
+
+\begin{example}
+    Consider the following IVP:
+    $$
+    y' = y^2 , \quad y(0) = 1
+    $$
+    We can solve this ODE by separating variables:
+    \begin{align*}
+        \frac{dy}{y^2} &= dt \\
+        \int y^{-2} \, dy &= \int 1 \, dt \\
+        -y^{-1} &= t + C \\
+        \intertext{Using the initial condition \( y(0) = 1 \), we find \( C = -1 \). Thus, the solution is:}
+        y(t) &= \frac{1}{1 - t}
+    \end{align*}
+    The solution exists but not globally, as it blows up at \( t = 1 \). Due to the nature of the initial condition, the solution is defined only for \( t < 1 \) to guarantee continuousness.
+\end{example}
+
+\begin{theorem}[Existence and Uniqueness Theorem for Linear ODEs]
+    Consider the IVP:
+    $$
+    y' + p(t)y = g(t), \quad y(t_0) = y_0
+    $$
+    where \( p(t) \) and \( g(t) \) are continuous on an open interval \( I = (\alpha, \beta) \) containing \( t_0 \).
+Then, there exists a unique solution \( y(t) \) defined on the entire interval \( I \).
+\end{theorem}
+\begin{proof}
+    (\textbf{Existence}) We can use the integrating factor method to find a solution. The integrating factor is given by:
+    $$\mu(t) = \exp\left( \int p(t) \, dt \right)
+    $$
+    Multiplying both sides of the ODE by the integrating factor, we have:
+    $$\mu(t) y' + \mu(t) p(t) y = \mu(t) g(t)
+    $$
+    Integrating both sides with respect to \( t \):
+    $$\mu(t)y = \int \mu(s) g(s) \, ds + C
+    $$
+    We can determine \( C \) using the initial condition \( y(t_0) = y_0 \):
+    $$y_0 = \frac{1}{\mu(t_0)}\left( \int_{t_0}^{t_0} \mu(s) g(s) \, ds + y_0 \right)
+    $$
+    Thus, we have a well defined solution \( y(t) \) defined on the interval \( I \), as long as $p$ and $g$ are continuous on $I$.
+
+    (\textbf{Uniqueness}) Suppose there are two solutions \( y_1(t) \) and \( y_2(t) \) to the IVP, then they must satisfy $y_1(t_0) = y_2(t_0) = y_0$.
+Define \( z(t) = y_1(t) - y_2(t) \), then \( z(t) \) satisfies:
+    $$
+    z' + p(t)z = g - g = 0, \quad z(t_0) = y_1(t_0) - y_2(t_0) = 0
+    $$
+    Apply the same integrating factor method, we have:
+    $$\mu(t) z' + \mu(t) p(t) z = 0
+    $$
+    Integrating both sides with respect to \( t \):
+    $$\mu(t)z = \int_{t_0}^{t} 0 \, ds + C = C
+    $$
+    Using the initial condition \( z(t_0) = 0 \), we find \( C = 0 \).
+Then, since $\mu$ is a exponential, $\mu > 0$, \( z(t) = 0 \) for all \( t \in I \), which implies \( y_1(t) = y_2(t) \) for all \( t \in I \).
+Therefore, the solution to the IVP is unique.
+\end{proof}
+
+\begin{theorem}[Cauchy-Lipschitz Theorem]
+    Consider the IVP:
+    $$
+    y' = f(t, y), \quad y(t_0) = y_0
+    $$
+    where \( f(t, y) \) and \( f_y(t, y) \) are continuous on a rectangle \( R = [\alpha, \beta] \times [\gamma, \delta] \).
+Then, for some point \((t_0, y_0) \in \mathrm{int}(R)\), there exists an interval \( I = (t_0 - h, t_0 + h) \subseteq [\alpha, \beta] \) for some \( 0 < h \le \min(t_0 - \alpha, \beta - t_0) \), such that there exists a unique solution \( y(t) \) defined on the interval \( I \).
+
+\end{theorem}
+
+\begin{proof}
+    The proof involves the method of successive approximations (Picard iterations). Which we will not cover here.
+\end{proof}
+
+\begin{example}
+    To demonstrate that the Cauchy-Lipschitz theorem applies in the linear case, consider the IVP:
+    $$
+    y' + p(t)y = g(t), \quad y(t_0) = y_0
+    $$
+    where \( p(t) \) and \( g(t) \) are continuous.
+We also have \( f_y = -p(t)y + g(t) = -p(t) \).
+So both \( f \) and \( f_y \) are continuous, thus the Cauchy-Lipschitz theorem guarantees the existence and uniqueness of the solution on some interval around \( t_0 \).
+\end{example}
+
+\begin{example}
+    Consider the IVP:
+    $$
+    y' = y^{1/3}, \quad y(0) = 0
+    $$
+    Here, \( f(t, y) = y^{1/3} \) and \( f_y(t, y) = \frac{1}{3}y^{-2/3} \).
+The function \( f(t, y) \) is continuous everywhere, but \( f_y(t, y) \) is not continuous at \( y = 0 \).
+Therefore, the Cauchy-Lipschitz theorem does not apply at the initial condition \( (0, 0) \), which explains why we have multiple solutions to this IVP.
 \end{example}

### 2025-09-14 YEAR2/PHY293_LectureNotes/main.tex
push

@@
     \begin{equation}
-        E(t) = E_0 e^{-\frac{t}{\tau}}
+        E(t) = E_0 e^{-\gamma t} = E_0 e^{-\frac{t}{\tau}}
     \end{equation}
@@
     \end{equation}
+
+We can also consider the ratio between the energy at one time and the energy one period later:
+\begin{align*}
+    \frac{E(t + T)}{E(t)} &= \frac{E_0 e^{-\gamma (t + T)}}{E_0 e^{-\gamma t}} \\
+    &= e^{-\gamma T} \approx 1 - \gamma T \\
+    \frac{E(t + T) - E(t)}{E(t)} & \approx -\gamma T = -\frac{2\pi}{\omega_0} \gamma = -\frac{2\pi}{Q}
+\end{align*}
 \end{definition}
+
+\begin{example}
+    \textit{What is the number of radians through which the damped system oscillates as its energy decreases to $1/e$ of its initial value?}
+
+    We have
+    $$
+    \frac{E}{E_0} = e^{-\gamma t} = \frac{1}{e} \implies \gamma t = 1 \implies t = \frac{1}{\gamma} = \tau
+    $$
+    So the number of radians is
+    $$
+    \theta = \omega \tau = \frac{\omega}{\gamma} = Q
+    $$
+
+\end{example}
+
+\subsection{Undamped Forced Oscillations}
+\begin{definition}[Undamped Forced Oscillations]
+    Consider a driver porves on the left side of the sping, we have the following differential equation:
+    \begin{equation}
+        m\ddot{x} + kx = \eta
+    \end{equation}
+    where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The general solution to this non-homogeneous differential equation given by:
+    \begin{equation}
+        x(t) = A(\omega)\cos(\omega t - \delta) 
+    \end{equation}
+    to derive $A(\omega)$ and $\delta$, we can use the equation of motion:
+    \begin{align*}
+        A(\omega)(-\omega^2 + \omega_0^2) \cos(\delta) = \omega_0^2 a
+    \end{align*}
+    where \( a = \frac{F_0}{k} \) is the static displacement of the mass when the driving force is constant. We also haev:
+    \begin{align*}
+        A(\omega)(-\omega^2 + \omega_0^2) \sin(\delta) = 0
+    \end{align*}
+    Now, consider the case when $\delta = 0$, we have:
+    \begin{equation}
+        A(\omega) = \frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
+    \end{equation}
+    and the case where \( \delta = \pi \), we have:
+    \begin{equation}
+        A(\omega) = -\frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
+    \end{equation}
+    Observe, when $\omega \approx \omega_0$, the amplitude \( A(\omega) \) becomes very large, indicating resonance.
+At resonance, the system oscillates with maximum amplitude, which can lead to significant energy transfer from the driving force to the oscillator.
+\end{definition}
+
+\subsection{Damped Forced Oscillations}
+\begin{definition}[Damped Forced Oscillations]
+    Consider a damped harmonic oscillator subjected to an external driving force. It is described by the following differential equation:
+    \begin{equation}
+        m\ddot{x} + b\dot{x} + kx = F_0 \cos(\omega t)
+    \end{equation}
+    where \( F_0 \) is the amplitude of the driving force, \( \omega \) is the angular frequency of the driving force, \( b \) is the damping coefficient, \( m \) is the mass, and \( k \) is the spring constant.
+The general solution to this non-homogeneous differential equation is given by:
+    \begin{equation}
+        x(t) = A(\omega)\cos(\omega t - \delta)
+    \end{equation}
+    and the amplitude \( A(\omega) \) is given by:
+    \begin{equation}
+        A(\omega) = \frac{\omega_0^2 a}{\sqrt{\omega^2 \gamma^2 + (\omega^2 - \omega_0^2)^2}}
+    \end{equation}
+    We can derive the following 4 cases:
+    \begin{enumerate}
+        \item \textbf{No Damping} (\( \gamma = 0 \)): In this case, the amplitude \( A(\omega) \) simplifies to the undamped case we discussed earlier.
+        \item \textbf{Low-Frequency Limit} (\( \omega \ll \omega_0 \)): In this limit, the amplitude \( A(\omega) \) approaches the static displacement \( a = \frac{F_0}{k} \).
+This means that at very low frequencies, the system behaves like a static spring, and the mass is displaced by an amount proportional to the applied force.
+        \item \textbf{High-Frequency Limit} (\( \omega \gg \omega_0 \)): In this limit, the amplitude \( A(\omega) \) decreases with increasing frequency, following the relation \( A(\omega) \approx \frac{\omega_0^2 a}{\omega^2} \).
+This indicates that at very high frequencies, the mass cannot respond quickly enough to the rapidly oscillating driving force, resulting in a smaller amplitude of oscillation.
+        \item \textbf{Resonance} (\( \omega \approx \omega_0 \)): At resonance, the amplitude \( A(\omega) \) reaches its maximum value but it does not become infinite due to the presence of damping: 
+        \begin{equation}
+            A_{\text{max}} = \frac{\omega_0^2 a}{\omega_0 \gamma} = \frac{\omega_0 a}{\gamma} = Q a
+        \end{equation}
+    \end{enumerate}
+    We also have the phase shift \( \delta \) given by:
+    \begin{equation}
+        \tan(\delta) = \frac{\omega \gamma}{\omega_0^2 - \omega^2}
+    \end{equation}
+    The phase shift \( \delta \) indicates how much the oscillation of the mass lags behind the driving force. The behavior of \( \delta \) can be summarized as follows:
+    \begin{itemize}
+        \item At low frequencies (\( \omega \ll \omega_0 \)), \( \delta \) approaches 0, meaning the mass oscillates in phase with the driving force.
+        \item At resonance (\( \omega = \omega_0 \)), \( \delta = \frac{\pi}{2} \), indicating that the mass oscillates a quarter cycle behind the driving force.
+        \item At high frequencies (\( \omega \gg \omega_0 \)), \( \delta \) approaches \( \pi \), meaning the mass oscillates out of phase with the driving force.
+    \end{itemize}
+\end{definition}
+
 \chapter{Modern Physics}

### 2025-09-16 YEAR2/AER210_LectureNotes/main.tex
sept 15

@@
 
-\begin{definition}[Definition of Double Integral]
+\begin{definition}[Definition of Double Integral] \label{def:double_integral}
     Let $R$ be a rectangular region in the $xy$-plane given by $R = [a,b] \times [c,d]$. The double integral of a function $f(x,y)$ over the region $R$ is defined as:
@@
     Therefore, the total surface area of the sphere is:
-    $$        S = 8 \cdot S_{\text{1st octant}} = 4\pi a^2 $$
-
+    $$        
+        S = 8 \cdot S_{\text{1st octant}} = 4\pi a^2 
+    $$   
 \end{example}
 
+\begin{example}
+    \textit{Let $R$ be the triangular region with vertices at $(0,0,0)$, $(1,0,0)$, and $(0,1,0)$. Find the surface area of the portion of $z = 3x + y^2$ that lies above the region $R$.}
+
+    We first express the function and take the partial derivatives:
+    $$
+        z = 3x + y^2, \quad \frac{\partial z}{\partial x} = 3, \quad \frac{\partial z}{\partial y} = 2y
+    $$
+    Thus, we can set up the double integral as follows:
+    \begin{align*}
+        S &= \iint_R \sqrt{1 + \left( \frac{\partial z}{\partial x} \right)^2 + \left( \frac{\partial z}{\partial y} \right)^2} \, dA \\
+        &= \int_0^1 \int_0^{y} \sqrt{1 + 3^2 + (2y)^2} \, dx \, dy \\
+        &= \int_0^1 \int_0^{y} \sqrt{10 + 4y^2} \, dx \, dy \\
+        &= \int_0^1 \sqrt{10 + 4y^2} \cdot y \, dy \\ 
+        \intertext{The extra $y$ is useful for substitution. Let $u = 10 + 4y^2$, then $du = 8y \, dy$ (that's why we integrate w.r.t $x$ first). Thus, we have:}
+        S &= \int_{u=10}^{u=14} \sqrt{u} \cdot \frac{du}{8} = \frac{1}{8} \cdot \frac{2}{3} u^{\frac{3}{2}} \Big|_{u=10}^{u=14} \\ 
+        &= \frac{1}{12} (14^{\frac{3}{2}} - 10^{\frac{3}{2}}) \approx 1.7
+    \end{align*}
+\end{example}
+
+\subsection{Triple Integrals}
+The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
+Definition \ref{def:double_integral}. We can define a triple integral as follows:
+\begin{definition}
+    Consider a function $f(x,y,z)$ that is continuous a 
+    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
+    \begin{equation}
+        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
+    \end{equation}
+    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions.
+This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
+    $$       
+     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
+    $$
+    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.
+
+    In the retangluar case, we can express $\Delta V_i$ as:
+    $$        
+    \Delta V_i = \Delta x_i \Delta y_i \Delta z_i
+    $$
+    Thus we have:
+    $$
+    dV = dx \, dy \, dz
+    $$
+\end{definition}
+
 \subsection{Triple Integrals in Rectangular Coordinates}
-\subsection{Triple Integrals in Cylindrical Coordinates}
+\begin{definition}[Triple Integral in Rectangular Region]
+    Consider a function $f(x,y,z)$ that is continuous over a rectangular box $B$ defined by the inequalities:
+    $$
+        a \le x \le b, \quad c \le y \le d, \quad r \le z \le s
+    $$
+    Then, the triple integral of $f$ over the box $B$ can be computed as an iterated integral in any order of integration:
+    \begin{equation}
+        \iiint_B f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx
+    \end{equation}
+    or any other permutation of the order of integration.
+
+\end{definition}
+
+\begin{example}
+    Consider $f(x,y,z)$ over a box $Q = \{(x,y,z) \mid a \le x \le b , c \le y \le d, r \le z \le s\}$. We would form the triple integral as follows:
+    $$
+        \iiint_Q f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx \\
+    $$
+    Of course, the order of integration can be change (there are $_3P_3= 6$ ways).
+\end{example}
+
+\begin{definition}[Triple Integrals in General Reigion]
+    Consider a function $f(x,y,z)$ that is continuous over a general reigion $V$ in the 3-D space. We can describe the reigion $V$ using the following inequalities:
+    $$
+        a \le x \le b, \quad g_1(x) \le y \le g_2(x), \quad h_1(x,y) \le z \le h_2(x,y)
+    $$
+    where $g_1(x)$, $g_2(x)$, $h_1(x,y)$, and $h_2(x,y)$ are continuous functions. Then, the triple integral of $f$ over the reigion $V$ can be computed as an iterated integral:
+    \begin{equation}
+        \iiint_V f(x,y,z) \, dV = \int_a^b \int_{g_1(x)}^{g_2(x)} \int_{h_1(x,y)}^{h_2(x,y)} f(x,y,z) \, dz \, dy \, dx = \iint_R \int_{h_1(x,y)}^{h_2(x,y)} f(x,y,z) \, dz \, dA
+    \end{equation}
+    where $R = \{(x,y) \mid a \le x \le b, g_1(x) \le y \le g_2(x)\}$ is the projection of the reigion $V$ onto the $xy$-plane.
+
+    The order of integration and their associated bounds can be changed based on the description of the region.
+\end{definition}
+
+\begin{example} \label{ex:triple_integral_tetrahedron}
+    \textit{Evaluate $\iiint_Q 6xy dV$, where $Q$ is the tetrahedron bounded by the coordinate planes $x=0$, $y=0$, $z=0$ and the plane $2x + y + z = 4$.}
+
+    Let $R$ be the projection of the tetrahedron onto the $xy$-plane. Then, we consider how the tetrahedron change in the $z$ direction over $R$:
+    $$
+        z = 4 - 2x - y \implies 0 \le z \le 4 - 2x - y
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_Q 6xy \, dV &= \iint_R \int_0^{4 - 2x - y} 6xy \, dz \, dA \\
+        &= \iint_R 6xy(4 - 2x - y) \, dA
+    \end{align*}
+    Now, we consider how the projection $R$ change in the $y$ direction over $x$:
+    $$
+        2x + y = 4 \implies 0 \le y \le 4 - 2x, \quad 0 \le x \le 2
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_Q 6xy \, dV &= \int_0^2 \int_0^{4 - 2x} 6xy(4 - 2x - y) \, dy \, dx \\
+        &= \int_0^2 \left[ 12x y^2 - 6x(4 - 2x) y^2 - 2y^3 \right]_{y=0}^{y=4 - 2x} \, dx \\
+        &= \int_0^2 (192x - 144x^2 + 24x^3 - 32x^4 + 8x^5) \, dx = \frac{64}{5}
+    \end{align*}
+\end{example}
+
+\begin{example}
+    \textit{Evaluate the integral given in the previous Example \ref{ex:triple_integral_tetrahedron} by intergrating w.r.t. $x$ first.}
+
+    We consider how the tetrahedron change in the $x$ direction over $R$, which is now the projection of the 
+    tetrahedron onto the $yz$-plane:
+    $$
+        2x + y + z = 4 \implies 0 \le x \le \frac{4 - y - z}{2}
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_Q 6xy \, dV &= \iint_R \int_0^{\frac{4 - y - z}{2}} 6xy \, dx \, dA \\
+        &= \frac{3}{4}\iint_R y(4 - y - z)^2 \, dA
+    \end{align*}
+    Now, we consider how the projection $R$ change in the $y$ direction over $z$:
+    $$
+        y + z = 4 \implies 0 \le y \le 4 - z, \quad 0 \le z \le 4
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_Q 6xy \, dV &= \frac{3}{4} \int_0^4 \int_0^{4 - z} y(4 - y - z)^2 \, dy \, dz \\
+        &= \frac{64}{5}
+    \end{align*}
+\end{example}
+
+\begin{example}
+    \textit{Find the volume of the solid bounded by the surface $z = 4 - y^2$ and planes given by $x+z = 4$, $x=0$, and $z=0$.}
+
+    First, we select the $z = 4 - y^2$ surface as our base reigion $R$. Then, we consider how the solid change in the $x$ direction over $R$:
+    $$
+        x + z = 4 \implies 0 \le x \le 4 - z
+    $$
+    So we can set up:
+    \begin{align*}
+        V &= \iint_R \int_0^{4 - z} 1 \, dx \, dA \\
+        &= \iint_R (4 - z) \, dA
+    \end{align*}
+    Now, we consider how the projection $R$ change in the $z$ direction over $y$:
+    $$        z = 4 - y^2 \implies 0 \le z \le 4 - y^2, \quad -2 \le y \le 2
+    $$
+    So we can set up:
+    \begin{align*}
+        V &= \int_{-2}^2 \int_0^{4 - y^2} (4 - z) \, dz \, dy \\
+        &= \int_{-2}^2 \left[ 4z - \frac{z^2}{2} \right]_{z=0}^{z=4 - y^2} \, dy \\
+        &= \int_{-2}^2 \left( 8 - \frac{y^2}{2} \right) \, dy \\
+        &= \frac{128}{5}
+    \end{align*}
+\end{example}
 \subsection{Triple Integrals in Spherical Coordinates}

### 2025-09-16 YEAR2/CHE260_LectureNotes/main.tex
sept 15

@@
 \end{shaded}
-\section{Systems and Properties in Thermodynamics}
+% ===================== PART 1: FOUNDATIONS =====================
+\section{Foundations and Fundamental Concepts}
 \begin{definition}[Energy]
@@
 These fundamental concepts are in arbitrary units and form the foundation `axioms' in science.
+We first build intuitive physical meaning (ability to lift, cause motion, create change) before formal laws.
 
-\subsection{Energy}
+\subsection{Mechanical Intuition for Energy Forms}
 \paragraph{Energy} An understanding of energy is its ability to \textbf{lift weights}. This is a way we could test if there is energy.
@@
 
-\paragraph{Steam Cycle} The steam cycle is a thermodynamic cycle that converts heat energy into mechanical work. It consists of four main processes: 
+\paragraph{Steam Cycle (Preview)} The steam cycle is a thermodynamic cycle that converts heat energy into mechanical work.
+It consists of four main processes (we will formalize cycles later under engines):
 \begin{enumerate}
@@
 
-The first law of thermodynamics is simple and easy to understand:
+% ===================== PART 2: LAWS OF THERMODYNAMICS (PREVIEW) =====================
+\section{Preview of Thermodynamic Laws and Directionality}
+The first intuitive principle (later formalized) is conservation of energy in transformations between forms:
 \begin{definition}[First Law of Thermodynamics] \label{def:firstlaw}
@@
 \end{definition}
-The second law formalizes the concept of the directionality of thermodynamic processes, stating that heat cannot spontaneously flow from a colder body to a hotter body.
-This law introduces the idea of entropy, a measure of the disorder or randomness in a system, which tends to increase over time in isolated systems.
+The second law formalizes directionality: heat does not spontaneously flow from cold to hot.
+This introduces \emph{entropy} as a bookkeeping tool for irreversibility.
+(A rigorous treatment of entropy will follow after we develop state / property concepts.)
 
@@
 
-\subsection{Systems and States}
+% ===================== PART 3: SYSTEM DESCRIPTORS =====================
+\section{Systems, States, and Properties}
 \begin{definition}[System]
@@
 
-\subsection{Modeling Ideal Gas}
+% ===================== PART 4: IDEAL GAS MODEL =====================
+\section{Ideal Gas Model and Molecular Basis}
 \begin{definition}[Quasi-Equilibrium Process (Reversible Path)]
@@
 
-\section{First Law of Thermodynamics}
+% ===================== PART 5: FIRST LAW FORMAL DEVELOPMENT =====================
+\section{First Law of Thermodynamics (Formal Development)}
 \begin{shaded}
@@
 
-\subsection{Work Modes}
+% ===================== PART 6: WORK AND ENERGY MODES =====================
+\section{Work and Energy Transfer Modes}
 \begin{definition}[Boundry Work]
@@
 
-\paragraph{Curve Fitting in the PV Diagram} In general, we can fit a curve in the $PV$ diagram using the polytropic process equation:
+\paragraph{General Polytropic Representation} In general, we can fit a curve in the $PV$ diagram using the polytropic process equation:
 \begin{equation}
@@
     where $P$ is the pressure, $B$ is the volume flow rate, and $m$ is the mass flow rate. $v$ is the specific volume (volume per unit mass).
+
     \textbf{Energy flowing a system per unit mass of a fluid} is called \textbf{flow energy} or \textbf{flow work}.
 It is the energy required to push mass into or out of a control volume.
-The flow work per unit mass is given by:
+The flow work per unit mass is given by enthalpy, which we will define next.
+\end{definition}
+
+\begin{definition}[Enthalpy]
+    Enthalpy is a thermodynamic property that represents the total energy of a system, including both its internal energy and the energy associated with its pressure and volume. It is defined as:
     \begin{equation}
@@
     \end{equation}
-    which is called the \textbf{enthalpy} of the fluid. The enthalpy is an extensive property, so we can define the specific enthalpy as:
+    where $H$ is the enthalpy, $U$ is the internal energy, $P$ is the pressure, and $V$ is the volume. Enthalpy is an extensive property, so we can define the specific enthalpy as:
     \begin{equation}
@@
     \end{equation}
+    where $h$ is the specific enthalpy, $u$ is the specific internal energy, and $v$ is the specific volume.
+
 \end{definition}
 
+\begin{example}[Constant Pressure Porcess]
+    Consider a gas in a cylinder with a movable piston. If we heat the gas at constant pressure, we would haveL
+    $$
+        P_{\text{gas}} = P_{\text{atm}}
+    $$
+    The work done by the gas during expansion is given by:
+    $$
+        w_{12} = -\int_{V_1}^{V_2} P dV = -P (V_2 - V_1)
+    $$
+    The energy balance for the system can be expressed as:
+    $$
+        Q_{\text{in}} - W_{\text{out}} = \Delta U_{\text{system}}
+    $$
+    where $Q_{\text{in}}$ is the heat added to the system, $W_{\text{out}}$ is the work done by the system, and $\Delta U_{\text{system}}$ is the change in internal energy of the system.
+Rearranging, we have:
+    $$
+        Q_{\text{in}} = (U_2 - U_1) + (P_2 V_2 - P_1 V_1) = (U_2 + P_2 V_2) - (U_1 + P_1 V_1) = H_2 - H_1
+    $$
+\end{example}
+
+\begin{theorem}[Enthalpy Change in a Constant Pressure Process]
+    For a constant pressure process, the heat added to the system is equal to the change in enthalpy of the system:
+    \begin{equation}
+        Q_{\text{in}} = H_2 - H_1
+    \end{equation}
+
+    For the instantaneous change, we have:
+    $$
+        \delta Q = du + Pdu
+    $$
+    from the definition of enthalpy, we have:
+    $$
+        h = u + Pv \quad \Rightarrow \quad dh = du + PdV + VdP
+    $$
+    Since the pressure is constant ($dP = 0$), we have:
+    $$
+        \delta Q = dh
+    $$
+\end{theorem}
+
+\begin{example}[Constant Volume Process]
+    Consider a gas in a cylinder with a fixed piston.
+If we heat the gas at constant volume, there is no boundary work done by the system, as the volume does not change ($dV = 0$).
+However, there may be other forms of work done on or by the system, such as electrical work if the gas is heated using an electric heater.
+The energy balance for the system can be expressed as:
+    $$
+        Q_{\text{in}} - W_{\text{out}} = \Delta U_{\text{system}}
+    $$
+    where $Q_{\text{in}}$ is the heat added to the system, $W_{\text{out}}$ is the work done by the system (which is zero in this case), and $\Delta U_{system}$ is the change in internal energy of the system.
+Rearranging, we have:
+    $$
+        Q_{\text{in}} = U_2 - U_1
+    $$
+\end{example}
+
+\begin{theorem}
+[Internal Energy Change in a Constant Volume Process]
+    For a constant volume process, the heat added to the system is equal to the change in internal energy of the system:
+    \begin{equation}
+        Q_{\text{in}} = U_2 - U_1
+    \end{equation}
+    For the instantaneous change, we have:
+    $$
+        \delta Q = du
+    $$  
+\end{theorem}
+
+\paragraph{How do we evaluate $\Delta U$ and $\Delta H$?} For an ideal gas, we can use the specific heat capacities at constant volume ($c_v$) and constant pressure ($c_p$) to evaluate the changes in internal energy and enthalpy:
+\begin{definition}[Specific Heat]
+    The specific heat is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius). It is given by:
+    \begin{equation}
+        c_{\text{average}} = \frac{Q}{\Delta T}
+    \end{equation}
+    where $Q$ is the heat added to the system and $\Delta T$ is the change in temperature. For the specific heat at each instanentious temperature, we have:
+    \begin{equation}
+        c(T) = \lim_{\Delta T \to 0} \frac{Q}{\Delta T} = \frac{\delta Q}{dT}
+    \end{equation}
+    note that $c$ could change with temperature.
+\end{definition}
+
+\begin{definition}[Specifc Heat at Constant Volume]
+    The specific heat at constant volume ($c_v$) is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius) while keeping the volume constant.
+It is given by:
+    \begin{equation}
+        c_v(T) = \left(\frac{\partial u}{\partial T}\right)_v 
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Specific Heat at Constant Pressure]
+    The specific heat at constant pressure ($c_p$) is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius) while keeping the pressure constant.
+It is given by:
+    \begin{equation}
+        c_p(T) = \left(\frac{\partial h}{\partial T}\right)_p 
+    \end{equation}
+\end{definition}
+
+\paragraph{Ideal Gas} For an ideal gas, the $h$ and $u$ depend only on temperature, so we can write:
+\begin{subequations}
+\begin{align}
+    du &= c_v(T) dT \\
+    dh &= c_p(T) dT
+\end{align}
+\end{subequations}
+Also, we have:
+$$
+    h = u + Pv = u + RT
+$$
+Differentiating, we have:
+$$
+    dh = du + R dT
+$$
+Substituting, we have:
+\begin{equation}
+    R = c_p - c_v
+\end{equation}
+
+\begin{definition}[Specifc Heat Ratio]
+    The specific heat ratio ($\gamma$) is the ratio of the specific heat at constant pressure to the specific heat at constant volume:
+    \begin{equation}
+        \gamma = \frac{c_p}{c_v}
+    \end{equation}
+    For a monoatomic ideal gas, we have:
+    \begin{equation}
+        \gamma = \frac{5}{3} \approx 1.67
+    \end{equation}
+    For a diatomic ideal gas (e.g., nitrogen, oxygen), we have:
+    \begin{equation}
+        \gamma = \frac{7}{5} = 1.4
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Change in Internal Energy and Enthalpy]
+    The change in internal energy and enthalpy of an ideal gas can be calculated using the specific heat capacities at constant volume ($c_v$) and constant pressure ($c_p$):
+    \begin{subequations}
+    \begin{align}
+        \Delta u &= \int_{T_1}^{T_2} c_v(T) dT \\
+        \Delta h &= \int_{T_1}^{T_2} c_p(T) dT
+    \end{align}
+    \end{subequations}
+    where $T_1$ and $T_2$ are the initial and final temperatures, respectively.
+
+    For small temperature changes, since $c_v$ and $c_p$ are relatively constant, we can approximate the changes as:
+    \begin{subequations}
+        \begin{align}
+            \Delta u &\approx c_v (T_2 - T_1) \\
+            \Delta h &\approx c_p (T_2 - T_1)
+        \end{align}
+    \end{subequations}
+    To choose the value of $c_v$ and $c_p$, could just use the average value over the temperature range, or use the value at the midpoint temperature.
+
+\end{definition}
+
+\begin{example}
+    Consider air heated from 300 K to 400 K.
+For air, we have $c_v = 0.721$ kJ/kg K and $c_p = 1.008$ kJ/kg K.
+The change in internal energy and enthalpy can be approximated (this is because we assume $c_v$ and $c_p$ are constant over the temperature range, which is a reasonable assumption for small temperature changes) as:
+
+    \begin{align*}
+        \Delta u &\approx \int_{T_1}^{T_2} c_v(T) dT = c_v (T_2 - T_1) = 0.721 \times (400 - 300) = 72.1\, \text{kJ/kg} \\
+        \Delta h &\approx \int_{T_1}^{T_2} c_p(T) dT = c_p (T_2 - T_1) = 1.008 \times (400 - 300) = 100.8\, \text{kJ/kg} \\
+    \end{align*}
+\end{example}
+
+\paragraph{Note} Do not associated $c_v$ and $c_p$ with constant volume and constant pressure processes, they are just the specific heat capacities at constant volume and constant pressure, they are properties of the substance that is true for any process.
+Associate them with $\Delta u$ and $\Delta h$.
+
+\begin{example}
+    Consider a piston-cylinder device.
+If we put a weight on top, the gas is compressed, and the piston moves down.
+This boundry work process would change enthalpy, despite not changing the temperature.
+This is becuase:
+    $$
+        dh = du + PdV + VdP = VdP \quad (du = 0, dV = 0)
+    $$
+\end{example}
 \chapter{Heat Transfer}

### 2025-09-16 YEAR2/ECE253_LectureNotes/main.tex
sept 15

@@
 
-\subsection{Hierarchy, Modularity, and Regularity}
+\section{Hierarchy, Modularity, and Regularity}
 
@@
 
-\section{Logic Circuits}
+% ===================== PART A: DIGITAL LOGIC FOUNDATIONS =====================
+\section{Digital Logic Foundations}
+% Representation and arithmetic first, then devices and gates, then algebra/minimization
+% --- Representation ---
 \subsection{Number Systems}
@@
 \end{definition}
+% --- Devices and gates ---
 \subsection{Transistors as Switches}
@@
 We denote the state of the gate as $x \in \{0,1\}$, where 0 is LO and 1 is HI.
-If input end of the swtich is HI, the output end could be modeled by the logic function:
+If input end of the switch is HI, the output end could be modeled by the logic function:
     $$
@@
 \begin{example}[Serial Transistors]
-    Consdier two transistors connected in series, with the input end of the first transistor connected to HI. The output end of the second transistor can be modeled by the following truth table:
+    Consider two transistors connected in series, with the input end of the first transistor connected to HI. The output end of the second transistor can be modeled by the following truth table:
     \begin{table}[h!]
@@
 \begin{example}[XOR Operation]
-    We have two swtiches, when both switches are in the same state (both open or both closed), the output is 0.
+    We have two switches; when both switches are in the same state (both open or both closed), the output is 0.
 When the switches are in different states (one open and one closed), the output is 1.
@@
 \begin{definition}[NAND Gate]
-    A NAND gate outputs 0 only if all inputs are 1. The truth table for a 2-input NAND gate is shown in Table \ref{tab:nand_nor_gates}.
+    A NAND gate outputs 0 only if all inputs are 1. The truth table for a 2-input NAND gate is as expected for the complement of AND.
     \\
@@
 
-\begin{definition}
-
+\begin{definition}[NOR Gate]
+    A NOR gate outputs 1 only if all inputs are 0. The truth table for a 2-input NOR gate is the dual of the OR gate.
+    \\
+    The logic function for a NOR gate with inputs $A$ and $B$ can be expressed as:
+    $$
+        L(A, B) = \overline{A + B}
+    $$
+    % A minimal symbolic sketch could be added similar to OR with an inversion bubble; omitted for brevity.
 \end{definition}
@@
 
+% --- Algebra and minimization ---
 \paragraph{Commonly Used Logic Operators} Below is a table summarizing the commonly used logic operators:
@@
         \hline
-        AND & $\cdot$ or no symbol & Outputs 1 if all inputs are 1 \\
+    AND & $\cdot$ or adjacency & Outputs 1 if all inputs are 1 \\
     OR & $+$ & Outputs 1 if at least one input is 1 \\
-        NOT & $\overline{x}$ & Outputs the inverse of the input \\
-        inverse & $'$ or $\sim$ & Outputs the 2's complement of the input \\
+    NOT & $\overline{x}$ or $x'$ or $\sim x$ & Outputs the logical negation of the input \\
     XOR & $\oplus$ & Outputs 1 if inputs are different \\
@@
 
+\paragraph{Summary of Important Theorems} The following table summarizes the important theorems/laws in Boolean algebra:
+\begin{table}[h!]
+    \centering
+    \caption{Summary of Important Theorems/Laws in Boolean Algebra (original form and dual form)}
+    \label{tab:boolean_algebra_summary}
+    \begin{tabular}{|l|c|c|}
+         \hline
+            \textbf{Law / Theorem} & \textbf{Original Form(s)} & \textbf{Dual Form(s)} \\ \hline
+            Commutative Law & $A + B = B + A$ & $A \cdot B = B \cdot A$ \\ \hline
+            Associative Law & $A + (B + C) = (A + B) + C$ & $A \cdot (B \cdot C) = (A \cdot B) \cdot C$ \\ \hline
+            Distributive Law & $A \cdot (B + C) = A \cdot B + A \cdot C$ & $A + (B \cdot C) = (A + B) \cdot (A + C)$ \\ \hline
+            Identity Law & $A + 0 = A$ & $A \cdot 1 = A$ \\ \hline
+            Null Law & $A + 1 = 1$ & $A \cdot 0 = 0$ \\ \hline
+            Idempotent Law & $A + A = A$ & $A \cdot A = A$ \\ \hline
+            Complement Law & $A + \overline{A} = 1$ & $A \cdot \overline{A} = 0$ \\ \hline
+            Double Negation Law & $\overline{\overline{A}} = A$ & - \\ \hline
+            De Morgan's Theorem & $\overline{A \cdot B} = \overline{A} + \overline{B}$ & $\overline{A + B} = \overline{A} \cdot \overline{B}$ \\ \hline
+            Absorption / Redundancy Theorem & $A + \overline{A}B = A + B$ & $A(\overline{A} + B) = AB$ \\ \hline
+            Combining Theorem & $AB + A\overline{B} = A$ & $(A + B)(A + \overline{B}) = A$ \\ \hline
+            Covering Theorem & $A + AB = A$ & $A(A + B) = A$ \\ \hline  
+    \end{tabular}
+\end{table}
+\newpage
 \begin{shaded}
-    \textbf{Logic Minimization} The goal of logic minimization is to reduce the number of logic gates and inputs in a digital circuit while maintaining its functionality.
+    \paragraph{Logic Minimization} The goal of logic minimization is to reduce the number of logic gates and inputs in a digital circuit while maintaining its functionality.
 This is important because it can lead to cost savings, improved performance, and reduced power consumption.
@@
     \begin{align*}
-        L(s_0, s_1, s_2) &= m_3 + m_4 + m_5 + m_6 + m_7 = \overline{s_0}s_1s_2 + s_2\overline{s_0}\overline{s_2} + s_0\overline{s_1}s_2 + s_0s_1\overline{s_2} + s_0s_1s_2 \\
+        L(s_0, s_1, s_2) &= m_3 + m_4 + m_5 + m_6 + m_7 \\ 
+        &= \overline{s_0}s_1s_2 + s_2\overline{s_0}\overline{s_2} + s_0\overline{s_1}s_2 + s_0s_1\overline{s_2} + s_0s_1s_2 \\
         \intertext{Using the Combining Theorem}
@@
 \end{example}
+
+\begin{example}
+    \textit{Derive a minimal POS expression for $f(x_1, x_2, x_3 = \prod M(0, 2, 4)$}
+    \\
+    We have:
+    \begin{align*}
+        f(x_1, x_2, x_3) &= M_0 M_2 M_4 \\
+        &= (x_1 + x_2 + x_3)(x_1 + \overline{x_2} + x_3)(\overline{x_1} + x_2 + x_3) \\
+    \intertext{Recognizing the combining theorem $(x+y)(x+\overline{y}) = x$}
+        &= (x_1 + x_3)(x_2 + x_3)(\overline{x_1} + x_3) \\
+        &= (x_1 + x_3)(x_2 + \overline{x_1} + x_3) \\
+        &= (x_1 + x_3)(x_2 + x_3)
+    \end{align*}   
+\end{example}
+
+\begin{theorem}[Transporting POS to SOP]
+    A POS circuit can be implemented using only NAND gates.
+This is achieved by applying De Morgan's Theorem and the properties of NAND gates.
+We can use the trick $f = \overline{\overline{f}}$ to convert the POS expression into a form that can be implemented with NAND gates.
+\end{theorem}
+\begin{theorem}[Transporting SOP to POS]
+    An SOP circuit can be implemented using only NOR gates.
+This is achieved by applying De Morgan's Theorem and the properties of NOR gates.
+We can use the trick $f = \overline{\overline{f}}$ to convert the SOP expression into a form that can be implemented with NOR gates.
+\end{theorem}
+
+\begin{example}
+    \textit{Implement the function $f(x_1, x_2, x_3) = \sum m(1, 3, 5, 7)$ using only NAND gates.}
+    \\
+    We have:
+    \begin{align*}
+        f(x_1, x_2, x_3) &= m_1 + m_3 + m_5 + m_7 \\
+        &= \overline{x_1}\overline{x_2}x_3 + \overline{x_1}x_2x_3 + x_1\overline{x_2}x_3 + x_1x_2x_3 \\
+        &= x_3(\overline{x_1}\overline{x_2} + \overline{x_1}x_2 + x_1\overline{x_2} + x_1x_2) \\
+        &= x_3(x_1 + x_2) \quad (\text{Using Combining Theorem}) \\
+        &= \overline{\overline{x_3(x_1 + x_2)}} \quad (\text{Using } f = \overline{\overline{f}}) \\
+        &= \overline{\overline{x_3} + \overline{x_1 + x_2}} \quad (\text{Using De Morgan's Theorem}) \\
+        &= \overline{\overline{x_3} + (\overline{x_1} \cdot \overline{x_2})} \quad (\text{Using De Morgan's Theorem}) 
+    \end{align*}    
+\end{example}
+
+% ===================== PART B: DESIGN METHODS AND HDL =====================
+\section{Combinational Logic Circuits}
+
+\begin{definition}[Combinational Logic Circuit]
+    A combinational logic circuit is a digital circuit that implements a specific logic function using a combination of logic gates.
+The output of a combinational logic circuit depends only on the current inputs and not on any previous inputs or states.
+\end{definition}
+\begin{definition}{Hardware Description Language (HDL)}
+    A hardware description language (HDL) is a specialized programming language used to describe the structure, behavior, and operation of electronic circuits and systems.
+HDLs are used in the design and verification of digital systems, including integrated circuits (ICs) and field-programmable gate arrays (FPGAs).
+The two most commonly used HDLs are VHDL (VHSIC Hardware Description Language) and Verilog.
+\end{definition}
+\subsection{Introduction of Verilog}
+
+\begin{definition}[Module]
+    A module is a self-contained block of hardware that has inputs and outputs and an internal implementation (behavioral or structural). Modules are the unit of hierarchy in Verilog.
+\end{definition}
+
+\begin{example}[Module Block]
+    A module block in Verilog is defined using the `module` keyword, followed by the module name and a list of input and output ports. For example:
+    \begin{verbatim}
+    module basic_logic(input logic a, b,
+                       output logic w, x, y, z);
+        assign w = a & b; // AND gate
+        assign x = a | b; // OR gate
+        assign y = ~a;    // NOT gate
+        assign z = a ^ b; // XOR gate
+    endmodule
+    \end{verbatim}
+
+    \textbf{Keywords} The keywords used in the module block are:
+    \begin{itemize}
+        \item \verb|assign|: Used to define continuous assignments for combinational logic.
+        \item \verb|logic|: A type of variable that can hold binary values (0 or 1).
+    \end{itemize}
+\end{example}
+
+\begin{definition}[Continuous Assignment]
+    A continuous assignment is used to model combinational logic in Verilog.
+It is defined using the \verb|assign| keyword, followed by the output signal, the assignment operator `=`, and the logic expression.
+Continuous assignments are evaluated whenever any of the input signals change.
+That is, the output is considered instantaneously updated when the input changes (ignoring propagation delay).
+\end{definition}
+
+\begin{example}[2-1 Multiplexers (Mux)]
+    \textit{Design a circuit that controls a light $f$ based on either two switches $x$ and $y$.
+The switch that control the light is determined by a control signal $s$.
+If $s = 0$, the light is controlled by switch $x$.
+If $s = 1$, the light is controlled by switch $y$.}
+    \\
+    We have the following truth table:
+
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|c|}
+            \hline
+            $s$ & $x$ & $y$ & $f$ \\
+            \hline
+            0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 0 \\
+            0 & 1 & 0 & 1 \\
+            0 & 1 & 1 & 1 \\
+            1 & 0 & 0 & 0 \\
+            1 & 0 & 1 & 1 \\
+            1 & 1 & 0 & 0 \\
+            1 & 1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Mux}
+        \label{tab:mux_truth_table}
+    \end{table}
+
+    From the truth table, we can derive the following canonical SOP expression:
+    \begin{align*}
+        f(s, x, y) &= m_2 + m_3 + m_5 + m_7 \\
+        &= \overline{s}x\overline{y} + \overline{s}xy + s\overline{x}y + sxy \\
+        &= \overline{s}x(\overline{y} + y) + sy(\overline{x} + x) \quad (\text{Using Distributive Law}) \\
+        &= \overline{s}x \cdot 1 + sy \cdot 1 \quad (\text{Using Complementation}) \\
+        &= \overline{s}x + sy \quad (\text{Using Identity})
+    \end{align*}
+
+    The verilog implementation is as follows:
+    \begin{verbatim}
+    module mux2to1(input logic x, y, s,
+                   output logic f);
+        assign f = (~s & x) | (s & y);
+    endmodule
+    \end{verbatim}
+\end{example}
+
+\paragraph{The diagram of a Mux} is a trapezoid shown below:
+\begin{center}
+    \begin{tikzpicture}[scale=1]
+        % Input nodes
+        \node (x) at (-2,0.5) {$x$};
+        \node (y) at (-2,-0.5) {$y$};
+        \node (s) at (-0.5,1.5) {$s$}; 
+        % Wires to gate
+        \draw (x.east) -- (-1,0.5);
+        \draw (y.east) -- (-1,-0.5);
+        \draw (s.south) -- (-0.5,0.75);
+        % Gate body: trapezoid
+        \draw (-1,1) -- (0,0.5) -- (0,-0.5) -- (-1,-1) -- cycle;
+        % Output
+        \node (f) at (1,0) {$f$};
+
+        \node (0) at (-0.6,0.3) {0};
+        \node (1) at (-0.6,-0.2) {1};
+        \draw (0,0) -- (f.west);
+    \end{tikzpicture}
+\end{center}
+
+\paragraph{4-1 Multiplexer (Mux)} A 4-1 Mux has 4 data inputs ($d_0, d_1, d_2, d_3$).
+The select signal would have two bit ($s_0, s_1$) to select one of the four data inputs to be outputted.
+In general, this is call a multibit signal, a.k.a a bus.
+\begin{definition}
+
+\end{definition}
+
+% ===================== PART C: SEQUENTIAL LOGIC =====================
 \section{Digital Storage Elements}
+% Flip-flops, latches, registers (to be expanded)
 
 \section{Finite State Machines (FSM)}
+% FSM modeling, state diagrams, and Verilog implementation (to be expanded)
 

### 2025-09-16 YEAR2/PHY293_LectureNotes/main.tex
sept 15

@@
 \subsection{Governing Equations of Harmonic Oscillators}
+This subsection collects the baseline ODEs for simple, damped, and driven oscillators to set notation used later.
 
@@
 
+\begin{shaded}
+	\textbf{Note on damping parameter conventions.} Different texts use different symbols and normalizations:
+\begin{itemize}
+    \item Our notes use $\gamma = \tfrac{b}{m}$, so the ODE reads $\ddot{x} + \gamma\dot{x} + \omega_0^2 x = 0$ and the decay envelope is $e^{-\gamma t/2}$.
+    \item Many texts define $\beta = \tfrac{b}{2m}$ and may call it $\gamma$ instead. In that convention the ODE is $\ddot{x} + 2\beta\dot{x} + \omega_0^2 x = 0$ with envelope $e^{-\beta t}$.
+\end{itemize}
+Equivalences: $\beta = \tfrac{\gamma}{2}$ and our underdamped condition $\gamma < 2\omega_0$ corresponds to $\beta < \omega_0$ in the $2\beta$ convention.
+When comparing formulas, check which definition is being used and replace $\gamma \leftrightarrow 2\beta$ accordingly.
+\end{shaded}
+
 \begin{definition}[Driven Harmonic Oscillator]
@@
 \end{definition}
-\subsection{The Wave Equation}
-
-\begin{definition}[The Wave Equation]
-    The wave equation is a second-order linear partial differential equation that describes the propagation of waves, such as sound waves, light waves, and water waves, through a medium.
-In one dimension, it is given by the following PDE:
-    \begin{equation}
-        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
-    \end{equation}
-    where \( u(x,t) \) is the wave function, \( c \) is the speed of the wave in the medium, \( x \) is the spatial coordinate, and \( t \) is time.
-\end{definition}
+% [Moved: The Wave Equation section is presented later for a smoother flow after oscillators]
 
 \subsection{Simple Harmonic Motion}
+We analyze the undamped solution forms, relate constants to initial conditions, and derive period/frequency relations.
 \begin{definition}[Simple Harmonic Motion]
-    You should have leaned the Hooke's law and Newton's second law, which gives us the equation of motion for a simple harmonic oscillator.
-The same with the equation (\ref{eq:hooke}), which can be rewritten as:
+    You should have learned Hooke's law and Newton's second law, which give us the equation of motion for a simple harmonic oscillator.
+The same as equation (\ref{eq:hooke}), which can be rewritten as:
     \begin{equation}
@@
     \end{equation}
-    By setting $\omega^2 = \frac{k}{m}$, ageneral solution can be written as:
+    By setting $\omega^2 = \frac{k}{m}$, a general solution can be written as:
     \begin{equation}
@@
     \end{equation}
-    where \( A \) are the constants determined by the IVP, \( \omega \) is the angular frequency, and \( \phi \) is the phase constant.
-$x_0$ is the equilibrium position where we generally set it to be 0.
-The unknown constant can be determined by knowing $x, \dot{x}$ at specific times.
+    where $A_1$ and $A_2$ are constants determined by the IVP, $\omega$ is the angular frequency, and $\phi$ is the phase constant.
+$x_0$ is the equilibrium position (often set to 0).
+The unknown constants can be determined by knowing $x, \dot{x}$ at specific times.
 \end{definition}
@@
 \subsection{Damped Harmonic Motion}
+We solve the damped ODE, classify regimes (underdamped, critical, overdamped), and connect decay rates with parameters.
 \begin{definition}[Damped Harmonic Motion]
@@
 \subsection{Energy and Quality Factor}
+We study how energy decays under light damping, define the time constant and quality factor Q, and relate them to response.
 \begin{definition}[Energy of a Very Light Damping]
@@
 \subsection{Undamped Forced Oscillations}
+We examine steady-state response to sinusoidal driving without losses and identify resonance behavior.
 \begin{definition}[Undamped Forced Oscillations]
-    Consider a driver porves on the left side of the sping, we have the following differential equation:
+    Consider a driver force acting on the mass-spring system; the ODE becomes:
     \begin{equation}
@@
     \end{equation}
-    where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The general solution to this non-homogeneous differential equation given by:
+    where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The steady-state particular solution is:
     \begin{equation}
@@
     \end{equation}
-    to derive $A(\omega)$ and $\delta$, we can use the equation of motion:
+    To derive $A(\omega)$ and $\delta$, use the equation of motion:
     \begin{align*}
@@
     \end{align*}
-    where \( a = \frac{F_0}{k} \) is the static displacement of the mass when the driving force is constant. We also haev:
+    where \( a = \frac{F_0}{k} \) is the static displacement of the mass when the driving force is constant. We also have:
     \begin{align*}
@@
 \subsection{Damped Forced Oscillations}
+We add damping to the driven case, derive the frequency response amplitude and phase, and study bandwidth and Q.
 \begin{definition}[Damped Forced Oscillations]
@@
 
+\begin{definition}[Power absorbed during forced oscillations] 
+    We can calculate the velocity of the oscillator:
+    \begin{equation}
+        \dot{x}(t) = -\omega A(\omega) \sin(\omega t - \delta) = -v_0 \sin(\omega t - \delta)
+    \end{equation}
+    where \( v_0 = \omega A(\omega) \) is the maximum speed of the oscillator. Energy is lost at the following rate:
+    \begin{equation}
+        P(t) = b v(t)^2 
+    \end{equation}
+    Substituteing the expression for \( v(t) \) into the power equation gives:
+    $$
+        P(t) = b v_0^2 \sin^2(\omega t - \delta)
+    $$
+    The average power \( \langle P \rangle \) over one complete cycle of the driving force is given by:
+   \begin{equation}
+        \bar{P}(\omega) = \frac{b v_0^2}{2} = \frac{\omega^2 F_0^2 \gamma}{2m \left[(\omega^2 - \omega_0^2)^2 + \omega^2 \gamma^2\right]}
+   \end{equation}
+    If the driving frequency \( \omega \) is close to \( \omega_0 \), use $\omega^2 - \omega_0^2 \approx 2\omega_0\Delta \omega$ with $\Delta \omega = \omega - \omega_0$ to get the Lorentzian form:
+    \begin{equation}
+        \bar{P}(\omega) \approx \frac{\omega_0 F_0^2}{2m \gamma \Big[ 1 + \big(\tfrac{2\Delta \omega}{\gamma}\big)^2 \Big]}
+    \end{equation}
+    The maximum average power occurs at resonance:
+    \begin{equation}
+        \bar{P}_{\text{max}} = \frac{\omega_0 F_0^2}{2m \gamma}
+    \end{equation}
+\end{definition}
+
+\paragraph{Full Width Half Height} The graph of \( \bar{P}(\omega) \) versus \( \omega \) has a peak at \( \omega = \omega_0 \) with a maximum value of \( \bar{P}_{\text{max}} \).
+The full width at half maximum (FWHM) is the width of the peak at half of its maximum height.
+The FWHM is given by:
+\begin{equation}
+    \omega_\text{FWHM} = 2\Delta \omega = 2\gamma = \frac{2\omega_0}{Q}
+\end{equation}
+To illustrate the graph, we can plot \( \bar{P}(\omega) \) versus \( \omega \) for different values of the quality factor \( Q \):
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+        \begin{axis}[
+            xlabel={Driving Frequency \( \omega \)},
+            ylabel={Average Power \( \bar{P}(\omega) \)},
+            title={Average Power vs Driving Frequency for Different Quality Factors, Power normalized to \( \bar{P}_{\text{max}}  = 1\)},
+            legend pos=north east,
+            grid=major,
+            width=0.8\textwidth,
+            height=0.5\textwidth,
+            domain=0:3,
+            range=0:1.2,
+            samples=200,
+        ]
+        % Plot for Q = 1
+        \addplot[black] {1/(1 + (2*(x-1))^2)};
+        \addlegendentry{\( Q = 1 \)}
+
+        % Plot for Q = 5
+        \addplot[black, thick, dashed] {1/(1 + (2*(x-1)/0.2)^2)};
+        \addlegendentry{\( Q = 5 \)}
+
+        % Plot for Q = 10
+        \addplot[black, thick] {1/(1 + (2*(x-1)/0.1)^2)};
+        \addlegendentry{\( Q = 10 \)}
+        % add the lenght of the fwhm at P = 0.5
+        \draw[dashed] (axis cs:0.9,0) -- (axis cs:0.9,0.5);
+        \draw[dashed] (axis cs:1.1,0) -- (axis cs:1.1,0.5);
+        \draw[<->] (axis cs:0.9,0.5) -- (axis cs:1.1,0.5) node[midway, above] {\( \Delta \omega = \frac{\omega_0}{Q} \)};
+
+        % vertcal line dotted at omega = 1
+        \draw[dotted, red, thick] (axis cs:1,0) -- (axis cs:1,1);
+        % label the peak
+        \node at (axis cs:1,1.05) {\( \bar{P}_{\text{max}} \)};
+        % label the x axis at omega = 1
+        \node at (axis cs:1,-0.05) {\( \omega_0 \)};
+
+        % label the points
+        \node at (axis cs:0.7,-0.05) {\( \omega_0 - \frac{\gamma}{2} \)};
+        \node at (axis cs:1.3,-0.05) {\( \omega_0 + \frac{\gamma}{2} \)};
+
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Average Power vs Driving Frequency for Different Quality Factors}
+    \label{fig:power_vs_frequency}
+\end{figure}
+
+\subsection{Simple Pendulum}
+Consider the arc length \( s \) of a pendulum bob, we have:
+\begin{equation}
+    s = l\theta \quad m\ddot{s} = -mg\sin(\theta)
+\end{equation}
+
+For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
+\begin{equation}
+    \ddot{\theta} + \frac{g}{l}\theta = 0
+\end{equation}
+And the ODE has solution:
+\begin{equation}
+    \theta(t) = \theta_0 \cos(\omega t + \phi)
+\end{equation}
+
+\paragraph{Energy} For small angle $\theta$, the total mechanical energy \( E \) of the pendulum is given by:
+\begin{equation}
+    E = K + U = \frac{1}{2}ml^2\dot{\theta}^2 + mgl(1 - \cos(\theta)) \approx \frac{1}{2}ml^2\dot{\theta}^2 + \frac{1}{2}mgl\theta^2
+\end{equation}
+This shows that the pendulum behaves like a simple harmonic oscillator with an effective spring constant \( k = \frac{mg}{l} \).
+
+\paragraph{The physical pendulum} For a rigid body swinging about a pivot point, the equation of motion is given by:
+\begin{equation}
+    I\ddot{\theta} = \tau 
+\end{equation}
+For simple rod of length \( L \) and mass \( m \) pivoted at one end, the moment of inertia about the pivot point is \( I = \frac{1}{3}mL^2 \).
+The torque due to gravity when the rod is displaced by an angle \( \theta \) from the vertical is \( \tau = -mg\frac{L}{2}\sin(\theta) \).
+For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
+\begin{equation}
+    \ddot{\theta} + \frac{3g}{2L}\theta = 0
+\end{equation}
+The angular frequency of oscillation for the physical pendulum is:
+\begin{equation}
+    \omega = \sqrt{\frac{3g}{2L}}
+\end{equation}
+
+\paragraph{The LC Circuit} An LC circuit consists of an inductor \( L \) and a capacitor \( C \) connected in series. The charge \( q(t) \) on the capacitor satisfies the differential equation:
+\begin{equation}
+    L\frac{d^2q}{dt^2} + \frac{1}{C}q = 0
+\end{equation}
+This is analogous to the equation of motion for a simple harmonic oscillator, with the angular frequency given by:
+\begin{equation}
+    \omega = \frac{1}{\sqrt{LC}}
+\end{equation}
+
+\paragraph{The RLC Circuit} An RLC circuit consists of a resistor \( R \), inductor \( L \), and capacitor \( C \) connected in series.
+The charge \( q(t) \) on the capacitor satisfies the differential equation:
+\begin{equation}
+    L\frac{d^2q}{dt^2} + R\frac{dq}{dt} + \frac{1}{C}q = 0
+\end{equation}
+This is analogous to the resistor behaving as a damped harmonic oscillator, with the angular frequency given by:
+\begin{equation}
+    \omega = \sqrt{\frac{1}{LC} - \left(\frac{R}{2L}\right)^2}
+\end{equation}
+
+\section{The Wave Equation}
+
+\begin{definition}[The Wave Equation]
+    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water). In one dimension, it is:
+    \begin{equation}
+        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
+    \end{equation}
+    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
+\end{definition}
+
 \chapter{Modern Physics}

### 2025-09-19 YEAR2/AER210_LectureNotes/main.tex
sept19

@@
 
-\subsection{Triple Integrals}
+\subsection{Triple Integrals in Rectangular Coordinates}
 The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
@@
 
-\subsection{Triple Integrals in Rectangular Coordinates}
 \begin{definition}[Triple Integral in Rectangular Region]
@@
 \end{example}
+
+\begin{example}
+    \textit{Change the order of Integration in the following triple integral such that the integration is performed w.r.t. $x$, then $y$, then $z$:}
+    $$
+        \int_{-1}^1 \int_{x^2}^1 \int_0^{1 - y} f(x,y,z) \, dz \, dy \, dx
+    $$
+    The region is limited in the $z$ direction by $0 \le z \le 1 - y$:
+    % use plot 3d to draw z = 1 - y and z = 0
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                    axis lines = middle,
+                    xlabel = $x$,
+                    ylabel = $y$,
+                    zlabel = $z$,
+                    domain=-1:1,
+                    y domain=0:1,
+                    samples=20,
+                    samples y=10,
+                    colormap/cool,
+                    view={120}{30},
+                    width=10cm,
+                    height=8cm,
+                ]
+                % Draw the surface z = 1 - y
+                \addplot3[surf, opacity=0.5] {1 - y};
+                % Draw the plane z = 0
+                \addplot3[surf, opacity=0.5] {0};
+                % Draw the curve y = x^2
+                \addplot3[domain=-1:1, samples=100, thick, blue] ({x}, {x^2}, {0});
+                % Draw the vertical lines at x = -1 and x = 1
+                \addplot3[domain=0:1, samples=100, thick, red] ({-1}, {x}, {0});
+                \addplot3[domain=0:1, samples=100, thick, red] ({1}, {x}, {0});
+            \end{axis}
+        \end{tikzpicture}
+        \caption{The region bounded by $z = 1 - y$, $z = 0$, $y = x^2$, $x = -1$, and $x = 1$.}
+    \end{figure}
+
+    We can also visualize the $R$ that is the reigion projected in the $xy$-plane as follows
+
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}[scale=2]
+            % Draw axes
+            \draw[->] (-1.5,0) -- (1.5,0) node[right] {$x$};
+            \draw[->] (0,-0.5) -- (0,1.5) node[above] {$y$};
+
+            % Draw the curve y = x^2
+            \draw[domain=-1.2:1.2,smooth,variable=\x,blue] plot ({\x},{\x*\x}) node[right] {$y = x^2$};
+            % Draw the line y = 1
+            \draw[blue] (-1,1) -- (1,1) node[right] {$y = 1$};
+
+            %vertical labels for x = pm1
+            \draw[dashed] (-1,0) -- (-1,1);
+            \draw[dashed] (1,0) -- (1,1);
+            \node at (-1,-0.1) {$-1$};
+            \node at (1,-0.1) {$1$};
+
+            % Shade the region
+            \fill[blue!20,opacity=0.5] plot[domain=-1:1,smooth,variable=\x] ({\x},{\x*\x}) -- (1,1) -- (-1,1) -- cycle;
+            % Add labels
+            \node at (0.5,0.5) {$R$};
+        \end{tikzpicture}
+        \caption{The projection of the region onto the $xy$-plane.}
+
+    \end{figure}
+
+    So, we can consider the projection of the triangle on the $zy$-plane as our base region $R'$. Then, we consider how the region changes in the $x$ direction over $R$:
+    $$
+        x^2 = y \implies -\sqrt{y} \le x \le \sqrt{y}
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_V f(x,y,z) \, dV &= \iint_{R'} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dA \\
+        &= \iint_{R'} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dy \, dz
+    \end{align*}
+    Now, we consider how the projection $R'$ change in the $y$ direction over $z$:
+    $$        z = 1 - y \implies 0 \le y \le 1 - z, \quad 0 \le z \le 1
+    $$
+    So we can set up:
+    \begin{align*}
+        \iiint_V f(x,y,z) \, dV &= \int_0^1 \int_0^{1 - z} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dy \, dz
+    \end{align*}
+\end{example}
+\subsubsection{Applications of Triple Integrals}
+\paragraph{Physical Applications} Similar to double integrals, triple integrals can be used to compute physical quantities such as mass, center of mass, and moment of inertia for three-dimensional objects with variable density.
+\begin{definition}[Mass of a Solid]
+    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid. The mass $M$ of the solid can be computed using the following triple integral:
+    \begin{equation}
+        M = \iiint_V \rho(x,y,z) \, dV
+    \end{equation}
+
+\end{definition}
+
+\begin{definition}[Center of Mass of a Solid]
+    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid.
+The coordinates of the center of mass $(\bar{x}, \bar{y}, \bar{z})$ of the solid can be computed using the following formulas:
+    \begin{equation}
+        \bar{x} = \frac{1}{M} \iiint_V x \rho(x,y,z) \, dV, \quad \bar{y} = \frac{1}{M} \iiint_V y \rho(x,y,z) \, dV, \quad \bar{z} = \frac{1}{M} \iiint_V z \rho(x,y,z) \, dV
+    \end{equation}
+    where $M$ is the mass of the solid.
+
+\end{definition}
+
+\begin{definition}[Moment of Inertia of a Solid]
+    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid.
+The moment of inertia of the solid about the $x$-axis ($I_x$), $y$-axis ($I_y$), and $z$-axis ($I_z$) can be computed using the following formulas:
+    \begin{equation}
+        I_x = \iiint_V (y^2 + z^2) \rho(x,y,z) \, dV, \quad I_y = \iiint_V (x^2 + z^2) \rho(x,y,z) \, dV, \quad I_z = \iiint_V (x^2 + y^2) \rho(x,y,z) \, dV
+    \end{equation}
+    where $y^2 + z^2$, $x^2 + z^2$, and $x^2 + y^2$ are the squares of the distances from the respective axes of rotation, and $\rho(x,y,z) \, dV$ represents the mass element of the solid at the point $(x,y,z)$.
+
+    Also for the moment of inertia about the origin ($I_o$):
+    \begin{equation}
+        I_o = \iiint_V (x^2 + y^2 + z^2) \rho(x,y,z) \, dV
+    \end{equation}
+    In general, for an axis defined by a line $ax + by + cz + d = 0$, the moment of inertia about that axis ($I_l$) can be found using the following formula:
+    \begin{equation}
+        I_l = \iiint_V \left( \frac{ax + by + cz + d}{\sqrt{a^2 + b^2 + c^2}} \right)^2 \rho(x,y,z) \, dV
+    \end{equation}
+\end{definition}
+
+\begin{example}
+    \textit{Find the center of mass of a solid of constant density $\rho_0$ that is bounded by the parabolic cylinder $x = y^2$ and the planes $z=0$, $x=z$ and $x=1$.}
+    We visualize the solid as follows:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                    axis lines = middle,
+                    xlabel = $x$,
+                    ylabel = $y$,
+                    zlabel = $z$,
+                    domain=0:1.3,
+                    y domain=-1:1,
+                    samples=20,
+                    samples y=10,
+                    colormap/cool,
+                    view={120}{30},
+                    width=10cm,
+                    height=8cm,
+                ]
+                % Draw the surface z = x
+                \addplot3[surf, opacity=0.5] {x};
+                % Draw the plane z = 0
+                \addplot3[surf, opacity=0.5] {0};
+
+                % Draw the surface cylinder x = y^2 on the plane z = 0, and it extends to z = x
+                % Parabolic cylinder x = y^2 (y in [-1,1], z in [0,1])
+                \addplot3[surf, opacity=0.45, shader=flat, domain=-1:1, y domain=0:1] ({x^2}, {x}, {y});
+                % bottom z = 0
+                \addplot3[surf, opacity=0.25, shader=flat, domain=0:1, y domain=-1:1] ({x}, {y}, {0});
+                % slanted plane x = z (i.e. z = x)
+                \addplot3[surf, opacity=0.35, shader=flat, domain=0:1, y domain=-1:1] ({x}, {y}, {x});
+                % vertical plane x = 1 (closes the solid)
+                \addplot3[surf, opacity=0.30, shader=flat, domain=-1:1, y domain=0:1] ({1}, {x}, {y});
+                % fill the surface such that
+            \end{axis}
+        \end{tikzpicture}
+        \caption{The solid bounded by the surfaces $x = y^2$, $z = 0$, $x = z$, and $x = 1$.}
+    \end{figure}
+    We can find the mass of the solid:
+    \begin{align*}
+        M &= \iiint_V \rho_0 \, dV = \rho_0 \iiint_V 1 \, dV \\
+        &= \rho_0 \int_{-1}^1 \int_{y^2}^1 \int_{0}^x 1 \, dz \, dx \, dy \\
+        &= \rho_0 \cdot \frac{4}{5}
+    \end{align*}
+
+    Next, we can find the coordinates of the center of mass:
+    \begin{align*}
+        \bar{x} &= \frac{1}{M} \iiint_V x \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V x \, dV = 5/7 \\
+        \bar{y} &= \frac{1}{M} \iiint_V y \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V y \, dV = 0 \text{ (by symmetry)} \\
+        \bar{z} &= \frac{1}{M} \iiint_V z \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V z \, dV = 5/14 \\
+    \end{align*}    
+\end{example}
+
+\begin{example}
+    \textit{Find the moment of inertia of a cyliner with radius $a$ and height $h$ about its central axis, assuming the cylinder has a constant density $\rho_0$.}
+    We consider the base region $R$ as the circular base of the cylinder in the $xy$-plane. Then, we consider how the cylinder change in the $z$ direction over $R$:
+    $$
+        0 \le z \le h
+    $$
+    So we can set up:
+    \begin{align*}
+        I_z &= \iiint_V (x^2 + y^2) \rho_0 \, dV = \rho_0 \iint_R \int_0^h (x^2 + y^2) \, dz \, dA \\
+        &= \rho_0 h \iint_R (x^2 + y^2) \, dA
+    \end{align*}
+
+    Then, we can express the base region $R$ in polar coordinates:
+    $$        
+    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
+    $$
+    So we can set up:
+    \begin{align*}
+        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
+        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
+    \end{align*}
+    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
+    $$
+        I_z = \frac{1}{2} M a^2
+    $$
+    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
+\end{example}
+\subsection{Triple Integrals in Cylindrical Coordinates}
+\begin{definition}[Cylindrical Coordinates]
+    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
+    \begin{itemize}
+        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
+        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
+        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
+    \end{itemize}
+    The relationships between cylindrical coordinates and rectangular coordinates are given by:
+    \begin{equation}
+        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
+    \end{equation}
+    and conversely:
+    \begin{equation}
+        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
+    \end{equation}
+\end{definition}
+
+\begin{example}
+
+\end{example}
 \subsection{Triple Integrals in Spherical Coordinates}

### 2025-09-19 YEAR2/CHE260_LectureNotes/main.tex
sept19

@@
 \end{example}
+
+\begin{example}[Control Volume in a Turbine]
+    The length of a fluid inside of control volume given time is:
+    $$
+        dx = V dt
+    $$
+    The volume is given by:
+    $$
+        dV = A dx = A V dt
+    $$
+    The mass of the fluid is:
+    $$
+        \delta m = \rho dV = \rho A V dt
+    $$
+    We can infer that if work is done, then the pressure is non constant. And that:
+    $$
+        \dot{m} = \rho A V = \text{constant}
+    $$
+    So at the inlet we have high $\rho$ and high $P$, and at the outlet we have low $\rho$ and low $P$. Such that $V$ remains relatively constant. We also devucebce specific volume:
+    \begin{equation}
+        v = \frac{1}{\rho}
+    \end{equation}
+    So we have:
+    \begin{equation}
+        \dot{m} = \frac{A V}{v} = \text{constant}
+    \end{equation}
+    At steady state:
+    \begin{equation}
+        \dot{m}_{in} = \dot{m}_{out}
+    \end{equation}
+    For and ideal gas, we have:
+    \begin{equation}
+        Pv = RT
+    \end{equation}
+\end{example}
+\begin{definition}[Fluid Energy per Unit Mass]
+    The total fluid energy per unit mass is given by:
+    \begin{equation}
+        e = h + \frac{V^2}{2} + gz
+    \end{equation}
+    where $h$ is the specific enthalpy, $\frac{V^2}{2}$ is the kinetic energy per unit mass, and $gz$ is the potential energy per unit mass.
+
+    Now, we can write the energy balance at the steady state, for a steady flow device:
+    \begin{equation}
+        q + w = (h_2 - h_1) + \left(\frac{V_2^2}{2} - \frac{V_1^2}{2}\right) + g(z_2 - z_1)
+    \end{equation}
+
+    For turbine, assume $q = 0$, and the KE and PE are small, we have:
+    \begin{equation}
+        w = h_1 - h_2, \quad W = m(h_1 - h_2)
+    \end{equation}
+    For compressor, we have $w_{shaft} = h_2 - h_1 > 0$.
+
+\end{definition}
+
+\begin{example}
+    Consider a compression turbine that has 800$^\circ$C and 600 kPa at the inlet, and 300$^\circ$C and 100 kPa at the outlet.
+Assuming ideal gas behavior, and $c_{p, \text{air}} = 1.005$ kJ/kg K.
+A 10kg flow rate would have:
+    $$
+        w = h_1 - h_2 
+    $$
+    We can compute the enthalpy change:
+    $$
+        \Delta h = \int_{T_1}^{T_2} c_p(T) dT \approx c_p (T_2 - T_1) = 1.005 \times (573 - 1073) = -502.5\, \text{kJ/kg}
+    $$
+    So the power output is:
+    $$
+        W = m \Delta h = 10 \times -502.5 = -5.025\, \text{MW}
+    $$
+\end{example}
+
+\begin{example}[Pump at Constant Temperature]
+    Consider a pump (a up pumping pump) that has 20$^\circ$C and 100 kPa at the inlet, and 20$^\circ$C and 600 kPa at the outlet.
+Assuming incompressible liquid water, and $v = 0.001\, \text{m}^3/\text{kg}$.
+A 10kg flow rate would have:
+    $$
+        w_{shaft} = h_2 - h_1 + g(z_2 - z_1)
+    $$
+    We can compute the enthalpy change:
+    $$
+        \Delta h = v (P_2 - P_1) 
+    $$
+    So the power with considerion of the change in potential energy is:
+    $$
+        W = m \Delta h + mg(z_2 - z_1) = 10 \times 0.001 \times (600 - 100) + 10 \times 9.81 \times (10 - 0) = 98.5\, \text{kW}
+    $$
+\end{example}
 \chapter{Heat Transfer}

### 2025-09-19 YEAR2/ECE253_LectureNotes/main.tex
sept19

@@
 \end{definition}
-
+\subsection{Multiplexers (Mux)}
 \begin{example}[2-1 Multiplexers (Mux)]
@@
 
-\paragraph{The diagram of a Mux} is a trapezoid shown below:
+\paragraph{The Diagram of a Mux} is a trapezoid shown below:
 \begin{center}
@@
 In general, this is call a multibit signal, a.k.a a bus.
-\begin{definition}
-
-\end{definition}
+\begin{definition}[Bus]
+    A bus is a bundle of signals. It is used to transfer data between different components of a digital system.
+\end{definition}
+
+\begin{example}[Verilog for a 2bit 2-1 Mux]
+    The verilog implementation for a 2bit 2-1 Mux is as follows:
+    \begin{verbatim}
+    module mux2to1_2bit(input logic [1:0] x, y, // 2-bit inputs
+                        input logic s,         // select signal
+                        output logic [1:0] f); // 2-bit output
+        assign f[0] = (~s & x[0]) | (s & y[0]); // LSB
+        assign f[1] = (~s & x[1]) | (s & y[1]); // MSB
+        // Alternatively, we can use the following single line:
+        // assign f = s ? y : x; // If s=1, f=y; else f=x
+    endmodule
+    \end{verbatim}
+
+    As we can see, we can initialize the datatype of a bus using the following syntax:
+    \begin{verbatim}
+    logic [n-1:0] bus_name; // n-bit bus
+    \end{verbatim}
+
+\end{example}
+
+\subsection{Adders}
+\begin{definition}[Half Adder]
+    A half adder is a combinational logic circuit that performs the addition of two single-bit binary numbers.
+It has two inputs, typically denoted as $A$ and $B$, and two outputs: the sum ($S$) and the carry ($C$).
+The sum output represents the least significant bit of the addition, while the carry output represents any overflow that occurs when both inputs are 1.
+We start with the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|c|}
+            \hline
+            $A$ & $B$ & $S_1$ & $S_2$ \\
+            \hline
+            0 & 0 & 0 & 0 \\
+            0 & 1 & 1 & 0 \\    
+            1 & 0 & 1 & 0 \\
+            1 & 1 & 0 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Half Adder}
+        \label{tab:half_adder_truth_table}
+    \end{table}
+    From the truth table, we can derive the following expressions using canonical SOP form:
+    \begin{align*}
+        S_0 &= m_1 + m_2 = \overline{A}B + A\overline{B} = A \oplus B \\
+        S_1 &= m_3 = AB
+    \end{align*}
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module half_adder(input logic A, B,
+                      output logic [1:0] S);
+        assign s[1] = a & b;
+        assign s[0] = a ^ b;
+    endmodule
+    \end{verbatim}
+\end{definition}
+
+\begin{definition}[Full Adder]
+    A full adder is a combinational logic circuit that performs the addition of three single-bit binary numbers: two significant bits and a carry-in bit.
+It has three inputs, we denote the signifcant bit as $A_i$ and $B_i$, and the carry-in bit as $C_i$, and two outputs: the sum ($S_i$) and the carry-out ($C_{i + 1}$).
+The sum output represents the least significant bit of the addition, while the carry-out output represents any overflow that occurs when the sum exceeds the value that can be represented by a single bit.
+We start with the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|c|c|}
+            \hline
+            $A_i$ & $B_i$ & $C_i$ & $S_i$ & $C_{i+1}$ \\
+            \hline
+            0 & 0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 1 & 0 \\
+            0 & 1 & 0 & 1 & 0 \\
+            0 & 1 & 1 & 0 & 1 \\
+            1 & 0 & 0 & 1 & 0 \\
+            1 & 0 & 1 & 0 & 1 \\
+            1 & 1 & 0 & 0 & 1 \\
+            1 & 1 & 1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Full Adder}
+        \label{tab:full_adder_truth_table}
+    \end{table}
+    From the truth table, we can derive the following expressions using canonical SOP form:
+    \begin{align*}
+        S_i &= m_1 + m_2 + m_4 + m_7 \\
+            &= \overline{A_i}\overline{B_i}C_i + \overline{A_i}B_i\overline{C_i} + A_i\overline{B_i}\overline{C_i} + A_iB_iC_i \\
+            &= A_i \oplus B_i \oplus C_i \\
+        C_{i+1} &= m_3 + m_5 + m_6 + m_7 \\
+                &= \overline{A_i}B_iC_i + A_i\overline{B_i}C_i + A_iB_i\overline{C_i} + A_iB_iC_i \\
+    \intertext{Using the Distributive Law and A = A + A}
+                &= B_iC_i(\overline{A_i} + A_i) + A_iB_i(\overline{C_i} + C_i) + A_iC_i(\overline{B_i} + B_i) \\
+                &= A_iB_i + B_iC_i + A_iC_i
+    \end{align*}
+    The circuit diagram of a full adder is shown below:
+    \begin{figure}[h!]
+        \begin{circuitikz}
+            \draw
+            % gates 
+            (0,0) node[xor port] (X1) {}
+            (2,0) node[xor port] (X2) {}
+            (1,-2) node[and port] (A1) {}
+            (3,-2) node[and port] (A2) {}
+            (2,-4) node[and port] (A3) {}
+            (4,-4) node[or port] (O1) {} 
+            (6,-4) node[or port] (O2) {};
+
+            % connections
+            \draw (X1.out) -- (X2.in 1);
+            \draw (X2.out) -- ++(1,0) node[right] {$S_i$};
+
+            % inputs
+            \draw (X1.in 1) -- ++(-1,0) node[left] {$A_i$};
+            \draw (X1.in 2) -- ++(-1,0) node[left] {$B_i$};
+            \draw (X2.in 2) -- ++(-0.3,0) node[left] {$C_i$};
+
+            % AND gate connections
+            \draw (X1.in 1) |- (A1.in 1);
+            \draw (X1.in 2) |- (A1.in 2);
+            \draw (X1.in 2) |- (A2.in 1);
+            \draw (X2.in 2) |- (A2.in 2);
+            \draw (X1.in 1) |- (A3.in 1);
+            \draw (X2.in 2) |- (A3.in 2);
+
+            % OR gate connections
+            \draw (A1.out) |- (O1.in 1);
+            \draw (A2.out) |- (O1.in 2);
+            \draw (A3.out) |- (O2.in 1);
+            \draw (O1.out) |- (O2.in 2);
+            \draw (O2.out) -- ++(1,0) node[right] {$C_{i+1}$}
+        ;
+
+        \end{circuitikz}
+        \caption{Circuit Diagram of a Full Adder}
+        \label{fig:full_adder_circuit}
+        \centering 
+    \end{figure}
+
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module full_adder(input logic A_i, B_i, C_i,
+                      output logic S_i, C_i_plus_1)
+        assign S_i = A_i ^ B_i ^ C_i;
+        assign C_i_plus_1 = (A_i & B_i) | (B_i & C_i) | (A_i & C_i);
+    endmodule
+    \end{verbatim}
+\end{definition}
+
+\begin{definition}[Hierarchical Verilog Module]
+    Hierarchical Verilog is a verilog module that instantiates other modules within it. This allows for the creation of complex designs by combining simpler modules.
+\end{definition}
+\begin{example}[3-bit Ripple Carry Adder]
+    A 3-bit ripple carry adder can be implemented using three instances of the full adder module. The Verilog implementation is as follows:
+    \begin{verbatim}
+    module adder3(input logic [2:0] A, B,
+                  input logic C_in,
+                  output logic [2:0] S,
+                  output logic C_out);
+        logic C_1, C_2; // Internal carry signals
+
+        // Instantiate full adders
+        full_adder FA0 (A[0], B[0], C_in, S[0], C_1);
+        full_adder FA1 (A[1], B[1], C_1, S[1], C_2);
+        full_adder FA2 (A[2], B[2], C_2, S[2], C_out);
+    endmodule
+    \end{verbatim}
+\end{example}
 

### 2025-09-19 YEAR2/MAT292_LectureNotes/main.tex
sept19

@@
 
-\begin{example}[Logistic Equation]
+\begin{example}[Logistic Equation] \label{ex:logistic-equation}
     The logistic equation is a first-order nonlinear separable autonomous ODE that models population growth with a carrying capacity. It is given by:
@@
 
+\subsection{Population Dynamics}
+\begin{example}[Simple Model]
+    Consider the ODE:
+    $$
+    P' = rP
+    $$
+    where \( P(t) \) is the population at time \( t \) and \( r \) is the growth rate. This model assumes that the population grows exponentially without any constraints. The solution to this ODE is:
+    $$
+    P(t) = P_0 e^{rt}
+    $$
+
+    \textit{Is this a good model?} Not really, because it assumes unlimited resources and no environmental constraints, which is unrealistic in real-world scenarios.
+\end{example}
+
+\begin{example}[A More Realistic Model]
+    Consider the logistic growth model:
+        $$
+        P' = rP\left(1 - \frac{P}{K}\right)
+        $$
+        where \( K > 0 \) is the carrying capacity of the environment.
+This model accounts for limited resources and environmental constraints.
+When $P = K$, the population growth rate becomes zero, indicating that the population has reached its maximum sustainable size.
+        The solution to this ODE is:
+        $$
+        P(t) = \frac{K}{1 + \left(\frac{K - P_0}{P_0}\right)e^{-rt}}
+        $$
+
+        This model is more realistic as it predicts that the population will grow rapidly when small, but will slow down and stabilize as it approaches the carrying capacity \( K \).
+
+        \textbf{Note} This is the same ODE considered in Example \ref{ex:logistic-equation}.
+
+        % Two figures side-by-side using minipages, scaled to the column width
+        \begin{figure}[h!]
+            \centering
+            \begin{minipage}[b]{0.48\textwidth}
+                \centering
+                \begin{tikzpicture}
+                    \begin{axis}[
+                        axis lines = middle,
+                        xlabel = \( P \),
+                        ylabel = \( P' \),
+                        ymin = -10, ymax = 10,
+                        xmin = -10, xmax = 110,
+                        domain=-1:110,
+                        samples=200,
+                        width=\linewidth,
+                        height=5.5cm,
+                        grid=both,
+                        minor tick num=1,
+                        major grid style={line width=.2pt,draw=gray!50},
+                        minor grid style={line width=.1pt,draw=gray!20},
+                    ]
+                    % Logistic growth curve (example parameters r=0.3, K=100)
+                    \addplot[blue, thick] {0.3*x*(1 - x/100)};
+                    % Fixed points
+                    \node at (axis cs:0,0) [circle, draw, inner sep=1.5pt] {};
+                    \node at (axis cs:100,0) [circle,fill,inner sep=1.5pt]{};
+                    % Labels
+                    \node at (axis cs:50,5) [anchor=south west] {\( P' = rP(1 - P/K) \)};
+                    \node at (axis cs:0,1) [anchor=south west] {\( P^* = 0 \)};
+                    \node at (axis cs:100,1) [anchor=south west] {\( P^* = K \)};
+                    \end{axis}
+                \end{tikzpicture}
+                \vspace{1mm}
+
+                \small\textbf{(a)} Fixed Point Diagram for the Logistic Growth Model
+            \end{minipage}\hfill
+            \begin{minipage}[b]{0.48\textwidth}
+                \centering
+                \begin{tikzpicture}
+                    \begin{axis}[
+                        axis lines = middle,
+                        xlabel = \( t \),
+                        ylabel = \( P \),
+                        ymin = -10, ymax = 150,
+                        xmin = 0, xmax = 50,
+                        domain=0:50,
+                        samples=200,
+                        width=\linewidth,
+                        height=5.5cm,
+                        grid=both,
+                        minor tick num=1,
+                        major grid style={line width=.2pt,draw=gray!50},
+                        minor grid style={line width=.1pt,draw=gray!20},
+                    ]
+                    % Logistic growth solution curves for different initial conditions (K=100, r=0.3)
+                    \addplot[blue, thick] {100/(1 + (100/10 - 1)*exp(-0.3*x))}; % P0 = 10
+                    \addplot[blue, thick] {100/(1 + (100/50 - 1)*exp(-0.3*x))}; % P0 = 50
+                    \addplot[blue, thick] {100/(1 + (100/90 - 1)*exp(-0.3*x))}; % P0 = 90
+
+                    % carrying capacity line
+                    \draw[dashed] (axis cs:0,100) -- (axis cs:50,100);
+                    \node at (axis cs:50.5,100) [anchor=west] {\( P = K \)};
+
+                    \addplot[red, thick] {100/(1 + (100/190 - 1)*exp(-0.3*x))}; % P0 = 10
+                    \addplot[red, thick] {100/(1 + (100/150- 1)*exp(-0.3*x))}; % P0 = 50
+                    \addplot[red, thick] {100/(1 + (100/110 - 1)*exp(-0.3*x))}; % P0 = 90
+
+                    % Labels for initial conditions
+                    \node at (axis cs:25,85) [anchor=south west] {\( P_0 = 10 \)};
+                    \node at (axis cs:25,60) [anchor=south west] {\( P_0 = 50 \)};
+                    \node at (axis cs:25,40) [anchor=south west] {\( P_0 = 90 \)};
+                    \end{axis}
+                \end{tikzpicture}
+                \vspace{1mm}
+
+                \small\textbf{(b)} Sample solution trajectories (different \(P_0\))
+            \end{minipage}
+            \caption{Logistic growth: (a) fixed-point diagram; (b) sample solution trajectories for different initial conditions.}
+            \label{fig:logistic-growth-side-by-side}
+        \end{figure}
+\end{example}
+
+\begin{example}[Logistics with Threshold]
+    Now we introduce a threshold \( T \) below which the population cannot sustain itself:
+        $$
+        P' = rP\left(1 - \frac{P}{K}\right)\left(\frac{P}{T} - 1\right)
+        $$
+        This model introduces an Allee effect, where the population growth rate becomes negative when the population is below the threshold \( T \). The fixed points of this system are:
+        $$
+        P^* = 0, \quad P^* = T, \quad P^* = K
+        $$
+        The stability of these fixed points can be analyzed as follows:
+        \begin{itemize}
+            \item \( P^* = 0 \) is stable (since \( f'(0) < 0 \))
+            \item \( P^* = T \) is unstable (since \( f'(T) > 0 \))
+            \item \( P^* = K \) is stable (since \( f'(K) < 0 \))
+        \end{itemize}
+
+    A phase plot and sample solution trajectories are shown below (the sketch shows guides at \(P=T\) and \(P=K\), and sample trajectories that converge to \(K\) or diverge to \(0\)):
+
+        \begin{figure}[h!]
+            \centering
+            \begin{minipage}[b]{0.48\textwidth}
+                \centering
+                \begin{tikzpicture}
+                    \begin{axis}[
+                        axis lines = middle,
+                        xlabel = \( P \),
+                        ylabel = \( P' \),
+                        ymin = -10, ymax = 10,
+                        xmin = -5, xmax = 110,
+                        domain=-1:110,
+                        samples=200,
+                        width=\linewidth,
+                        height=5.5cm,
+                        grid=both,
+                        minor tick num=1,
+                        major grid style={line width=.2pt,draw=gray!50},
+                        minor grid style={line width=.1pt,draw=gray!20},
+                    ]
+                    % Logistic growth with threshold curve (example parameters r=0.3, K=100, T=40)
+                    \addplot[blue, thick] {0.3*x*(1 - x/100)*(x/40 - 1)};
+                    % Axis baseline and vertical guides at P=T and P=K
+                    \addplot[black] coordinates {(0,0) (110,0)}; % P-axis
+                    \draw[red,dashed]  (axis cs:40,-9) -- (axis cs:40,9) node[black, above right] {\(P=T\)};
+                    \draw[teal,dashed] (axis cs:100,-9) -- (axis cs:100,9) node[black, above right] {\(P=K\)};
+                    % Fixed points markers on P-axis (at P'=0)
+                    \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt,label=below:{\(0\)}] {};
+                    \node at (axis cs:40,0) [circle,draw,fill=white,inner sep=1.5pt,label=below:{\(T\)}] {};
+                    \node at (axis cs:100,0) [circle,fill,inner sep=1.5pt,label=below:{\(K\)}] {};
+                    % Flow direction arrows on P-axis (qualitative)
+                    \draw[->,thick] (axis cs:20,0) -- (axis cs:5,0);    % 0<T: P decreases (left)                    \draw[->,thick] (axis cs:45,0) -- (axis cs:70,0.6);   % T<K: P increases (right)
+                    \draw[->,thick] (axis cs:75,0) -- (axis cs:95,0);   % >K: P decreases (left)
+
+                    \end{axis}
+                \end{tikzpicture}
+                \vspace{1mm}    
+                \small\textbf{(a)} Phase diagram (fixed points and flow)
+            \end{minipage}\hfill
+            \begin{minipage}[b]{0.48\textwidth}
+                \centering
+                \begin{tikzpicture}
+                    \begin{axis}[
+                        axis lines = middle,
+                        xlabel = \( t \),
+                        ylabel = \( P \),
+                        ymin = -5, ymax = 110,
+                        xmin = 0, xmax = 50,    
+                        domain=0:50,
+                        samples=200,
+                        width=\linewidth,
+                        height=5.5cm,
+                        grid=both,
+                        minor tick num=1,   
+                        major grid style={line width=.2pt,draw=gray!50},
+                        minor grid style={line width=.1pt,draw=gray!20},
+                    ]
+                    % carrying capacity line
+                    \draw[dashed] (axis cs:0,100) -- (axis cs:50,100);
+                    \node at (axis cs:50.5,100) [anchor=west] {\( P = K \)};
+                    % threshold line (visual)
+                    \draw[dashed] (axis cs:0,40) -- (axis cs:50,40);
+                    \node at (axis cs:50.5,40) [anchor=west] {\( P = T \)};
+                    % Sample solution trajectories (qualitative sketches)
+                    % Below-threshold -> decline to 0 (four examples)
+                    \addplot[red, thick, opacity=0.9, domain=0:50, samples=200]
+                        {40/(1 + (40/35 - 1)*exp(0.25*x))};
+                    \addplot[red, thick, opacity=0.9, domain=0:50, samples=200]
+                        {40/(1 + (40/25 - 1)*exp(0.25*x))};
+                    \addplot[red, thick, opacity=0.9, domain=0:50, samples=200]
+                        {40/(1 + (40/15 - 1)*exp(0.25*x))};
+                    \addplot[red, thick, opacity=0.9, domain=0:50, samples=200]
+                        {40/(1 + (40/5  - 1)*exp(0.25*x))};
+                    \node[red] at (axis cs:7,26) [anchor=west] {\(P_0<T\) diverges to $0$};
+                    % Above-threshold -> grow to K (four examples)
+                    \addplot[blue, thick, opacity=0.9, domain=0:50, samples=200]
+                        {100/(1 + (100/45 - 1)*exp(-0.3*x))};
+                    \addplot[blue, thick, opacity=0.9, domain=0:50, samples=200]
+                        {100/(1 + (100/60 - 1)*exp(-0.3*x))};
+                    \addplot[blue, thick, opacity=0.9, domain=0:50, samples=200]
+                        {100/(1 + (100/80 - 1)*exp(-0.3*x))};
+                    \addplot[blue, thick, opacity=0.9, domain=0:50, samples=200]
+                        {100/(1 + (100/95 - 1)*exp(-0.3*x))};
+                    \node[blue] at (axis cs:40,82) [anchor=east] {\(P_0>T\) converges to $K$};
+                    \end{axis}
+                \end{tikzpicture}
+                \vspace{1mm}    
+                \small\textbf{(b)} Sample time trajectories: convergence to \(K\) or collapse to \(0\)
+            \end{minipage}
+            \caption{Logistic growth with threshold: (a) phase portrait with horizontal guides at \(P=T\) and \(P=K\); (b) qualitative solution trajectories showing convergence and divergence relative to the threshold.}
+            \label{fig:logistic-growth-threshold-side-by-side}
+        \end{figure}
+\end{example}
+
+\begin{example}[Harvesting Model]
+    Consider a population model with harvesting:
+    $$
+    P' = rP\left(1 - \frac{P}{K}\right) - H
+    $$
+    where \( H \) is the constant harvesting rate. The fixed points of this system are given by solving:
+    $$
+    rP\left(1 - \frac{P}{K}\right) - H = 0
+    $$
+    This leads to a quadratic equation in \( P \):
+    $$
+    P^2 - KP + \frac{KH}{r} = 0
+    $$
+    The solutions to this equation are:
+    $$
+    P^* = \frac{K \pm \sqrt{K^2 - 4\frac{KH}{r}}}{2}
+    $$
+    The stability of these fixed points can be analyzed as follows:
+    \begin{itemize}
+        \item If \( H < \frac{rK}{4} \), there are two real fixed points: one stable and one unstable.
+        \item If \( H = \frac{rK}{4} \), there is one real fixed point (a saddle-node bifurcation point).
+        \item If \( H > \frac{rK}{4} \), there are no real fixed points, indicating that the population will decline to extinction.
+    \end{itemize}
+    \textbf{Fixed Point Diagram} A fixed point diagram is shown below:
+        \begin{figure}[h!]
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                    axis lines = middle,
+                    xlabel = \( y \),
+                    ylabel = \( f(y) \),
+                    ymin = -10, ymax = 10,
+                    xmin = -5, xmax = 105,
+                    domain=-1:105,
+                    samples=200,
+                    width=0.6\linewidth,
+                    height=6cm,
+                    grid=both,
+                    minor tick num=1,
+                    major grid style={line width=.2pt,draw=gray!50},
+                    minor grid style={line width=.1pt,draw=gray!20},
+                    legend pos=north east,
+                ]
+                % Quadratic curve for different H values with legend entries
+                \addplot[blue, thick] {0.3*x*(1 - x/100) - 5};
+                \addlegendentry{\(H < \tfrac{rK}{4}\)}
+                \addplot[red, thick, dashed] {0.3*x*(1 - x/100) - 7.5};
+                \addlegendentry{\(H = \tfrac{rK}{4}\)}
+                \addplot[green, thick, dotted] {0.3*x*(1 - x/100) - 10};
+                \addlegendentry{\(H > \tfrac{rK}{4}\)}
+                % Fixed points using plot marks (uses built-in plot labeling via legend)
+                \addplot[black, only marks, mark=o, mark size=2.5pt] coordinates {(20,0)};
+                \addplot[black, only marks, mark=*, mark size=2.5pt] coordinates {(80,0)};
+                % semi-filled marker for intermediate / degenerate fixed point
+                % half-filled marker: right half black, left half white
+                \node at (axis cs:50,0) [circle, draw, fill=white, inner sep=1.5pt] {};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Fixed point diagram for the harvesting model with different harvesting rates \(H\).}
+        \end{figure}
+
+     \textbf{Bifurcation Diagram} This is a classic example of a saddle-node bifurcation (0 to 2 fixed points). The bifurcation point is at $(k/2, 4h/K)$. The bifurcation diagram is shown below:
+        \begin{figure}[h!]
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                    axis lines = middle,
+                    xlabel = \( r \),
+                    ylabel = \( P^* \),
+                    ymin = -10, ymax = 110,
+                    xmin = -1, xmax = 30,
+                    domain=0:30,
+                    samples=200,
+                    width=0.6\linewidth,
+                    height=6cm,
+                    grid=both,
+                    minor tick num=1,
+                    major grid style={line width=.2pt,draw=gray!50},
+                    minor grid style={line width=.1pt,draw=gray!20},
+                ]
+                % Bifurcation curve (example parameters r=0.3, K=100)
+                \addplot[blue, thick, domain=7.5:30] {50 + 10*sqrt(x-7.5)};
+                \addplot[red, thick, dashed, domain=7.5:30] {50 - 10*sqrt(x-7.5)};
+                \draw[dashed] (axis cs:7.5,-10) -- (axis cs:7.5,110);
+                \node at (axis cs:7.5,110) [anchor=north west] {\( r = \frac{4H}{K} \)};
+                % Mark bifurcation point
+                \node at (axis cs:7.5,50) [circle,fill,inner sep=1.5pt]{};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Bifurcation diagram for the harvesting model showing the saddle-node bifurcation at \(H = \frac{rK}{4} \).}
+        \end{figure}
+
+\end{example}
+
+\begin{theorem}[ID Autonomous ODEs are Monotonic]
+    Consider the ODE:
+    $$
+    y' = f(y)
+    $$
+    where \( f(y) \) is continuous. Then, the solution \( y(t) \) is monotonic on its interval of existence.
+
+\end{theorem}
+\begin{proof}
+    WLOG, suppose \( f(y_0) > 0 \).
+For the sake of contradiction, suppose \( y(t) \) is not monotonic.
+Then, there exists \( t_1 > t_0 \) such that \( y(t_1) < y(t_0) \).
+By the Intermediate Value Theorem, there exists \( t_2 \in (t_0, t_1) \) such that \( y(t_2) = y(t_0) \).
+By the Mean Value Theorem, there exists \( c \in (t_0, t_2) \) such that:
+    $$\frac{y(t_2) - y(t_0)}{t_2 - t_0} = y'(c)$$
+    But \( y(t_2) = y(t_0) \), so the left-hand side is zero.
+Thus, \( y'(c) = 0 \).
+However, since \( f(y) \) is continuous and \( f(y_0) > 0 \), there exists a neighborhood around \( y_0 \) where \( f(y) > 0 \).
+This contradicts the assumption that \( y'(c) = 0 \).
+Therefore, \( y(t) \) must be monotonic.
+\end{proof}
+
+\begin{definition}[System of ODEs]
+    A system of ODEs is a set of coupled first-order ODEs of the form:
+    $$
+    \begin{cases}
+    x_1' = f_1(t, x_1, x_2, \ldots, x_n) \\
+    x_2' = f_2(t, x_1, x_2, \ldots, x_n) \\
+    \vdots \\
+    x_n' = f_n(t, x_1, x_2, \ldots, x_n)
+    \end{cases}
+    $$
+    where \( x_i(t) \) are the state variables and \( f_i(t, x_1, x_2, \ldots, x_n) \) are given functions.
+\end{definition}
+\subsection{Systems of ODEs}
+
+\begin{definition}[Two-Species Interaction Model]
+    Consider a model of two species interacting, such as predator-prey dynamics:
+    $$
+    \begin{cases}
+    x' = x(r_1 - ax -by) \\
+    y' = y(r_2 + cx - dy)
+    \end{cases}
+    $$
+    where \( x(t) \) and \( y(t) \) are the populations of the two species, \( r_1, r_2, a, b, c, d \) are constants representing:
+    \begin{itemize}
+        \item \( r_1, r_2 \): intrinsic growth rates of species \( x \) and \( y \)
+        \item \( a, d \): intrinsic carrying capacity coefficients
+        \item \( b, c \): interspecific coefficients representing the effect of one species on the other
+    \end{itemize}
+\end{definition}
+
+\begin{example}[Lotka-Volterra Model]
+    Now, consider the classic Lotka-Volterra predator-prey model:
+    $$
+    \begin{cases}
+    x' &= ax - bxy \\
+    y' &= -cy + dxy \\
+    (x(0), y(0)) &= (x_0, y_0)
+    \end{cases}
+    $$
+    where \( x(t) \) is the prey population, \( y(t) \) is the predator population, and \( a, b, c, d \) are positive constants representing:
+    \begin{itemize}
+        \item \( a \): growth rate of prey in the absence of predators
+        \item \( b \): predator efficiency (how many prey are eaten per predator per unit time)
+        \item \( c \): natural death rate of predators in the absence of prey
+        \item \( d \): conversion efficiency (how many predators are born per prey eaten)
+    \end{itemize}
+    Consider the fixed points of this system by setting \( x' = 0 \) and \( y' = 0 \). Then we have:
+    $$
+    (x^*, y^*) = (0, 0) \quad \text{and} \quad (x^*, y^*) = \left(\frac{c}{d}, \frac{a}{b}\right)
+    $$
+    The case where \( (x^*, y^*) = (0, 0) \) is trivial (extinction of both species).
+The non-trivial fixed point \( \left(\frac{c}{d}, \frac{a}{b}\right) \) represents a coexistence equilibrium where the rate of the prey being eaten is the rate of the prey being born.
+\end{example}
+
+We can further analyze this by considering Nullclines.
+
+\begin{definition}[Nullclines]
+    Nullclines are curves in the phase plane where the derivative of \textbf{one of the variables} is zero. For the Lotka-Volterra model, the nullclines are given by:
+    \begin{itemize}
+        \item Prey nullcline (\( x' = 0 \)): \( y = \frac{a}{b} \) (horizontal line)
+        \item Predator nullcline (\( y' = 0 \)): \( x = \frac{c}{d} \) (vertical line)
+    \end{itemize}
+    The intersection of these nullclines gives the fixed points of the system.
+
+\end{definition}
+
+\begin{example}[Nullclines of the Lotka-Volterra Model]
+    The x-nullclines ($x' = 0$) are given by:
+    $$x' = ax - bxy = 0 \implies x(a - by) = 0 \implies x = 0 \text{ or } y = \frac{a}{b}$$
+    The y-nullclines ($y' = 0$) are given by:
+    $$y' = -cy + dxy = 0 \implies y(-c + dx) = 0 \implies y = 0 \text{ or } x = \frac{c}{d}$$
+
+    Also, consider the growth of each species in the nullcline regions:
+    \begin{itemize}
+        \item For \( x' = 0 \) (prey nullcline):
+        \begin{itemize}
+            \item If \( x = 0 \), then \( y' < 0 \) (predator population decreases).
+            \item If \( y < \frac{a}{b} \), then \( x' > 0 \) (prey population increases).
+            \item If \( y > \frac{a}{b} \), then \( x' < 0 \) (prey population decreases).
+        \end{itemize}
+        \item For \( y' = 0 \) (predator nullcline):
+        \begin{itemize}
+            \item If \( y = 0 \), then \( x' > 0 \) (prey population increases).
+            \item If \( x < \frac{c}{d} \), then \( y' < 0 \) (predator population decreases).
+            \item If \( x > \frac{c}{d} \), then \( y' > 0 \) (predator population increases).
+        \end{itemize}
+    \end{itemize}
+
+    These directions would ultimately leads to a circular flow around the fixed point \( \left(\frac{c}{d}, \frac{a}{b}\right) \).
+
+    The intersection of these nullclines gives the fixed points of the system. A phase portrait showing the nullclines and the fixed point is illustrated below:
+        \begin{figure}[h!]
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                    axis lines = middle,
+                    view     = {0}{90},
+                    xlabel = \( x \) (Prey),
+                    ylabel = \( y \) (Predator),
+                    ymin = -1, ymax = 6,
+                    xmin = -1, xmax = 6,
+                    domain=-1:6,
+                    samples=200,
+                    width=0.6\linewidth,
+                    height=6cm,
+                    grid=both,
+                    minor tick num=1,
+                    major grid style={line width=.2pt,draw=gray!50},
+                    minor grid style={line width=.1pt,draw=gray!20},
+                ]
+                % Prey nullcline (horizontal line)
+                \addplot[red, thick] {3};
+                % trvial nullclines (axes)
+                \addplot[blue, thick] {0};
+                \node at (axis cs:5,3) [anchor=south west] {\( y = \frac{a}{b} \)};
+                % Predator nullcline (vertical line)
+                \addplot[red, thick] coordinates {(0,0) (0,6)};
+                \addlegendentry{Predator nullcline}
+                % trvial nullclines (axes)
+                \addplot[blue, thick] coordinates {(2,0) (2,6)};
+                \addlegendentry{Prey nullcline}
+
+                \node at (axis cs:2,5) [anchor=north east] {\( x = \frac{c}{d} \)};
+                % Fixed point marker
+                \node at (axis cs:2,3) [circle,fill,inner sep=1.5pt,label=above right:{\(\left(\frac{c}{d}, \frac{a}{b}\right)\)}] {};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Nullclines of the Lotka-Volterra model showing the prey nullcline (horizontal line) and predator nullcline (vertical line).
+The intersection point represents the coexistence equilibrium.}
+        \end{figure}
+
+\end{example}
+
+\begin{example}[Simple Pendulum]
+    Consider a simple pendulum of length \( L \) and mass \( m \) under the influence of gravity \( g \).
+The angle \( \theta(t) \) that the pendulum makes with the vertical satisfies the second-order ODE:
+    $$
+    \theta'' + \frac{g}{L} \sin(\theta) = 0
+    $$
+    To convert this into a system of first-order ODEs, we let:
+    $$\begin{cases}
+    x := \theta  \quad x' = \theta' = y \\
+    y := \theta' \quad y' = \theta'' = -\frac{g}{L} \sin(x)
+    \end{cases}$$
+
+    Consider the fixed points of this system by setting \( x' = 0 \) and \( y' = 0 \). Then we have:
+    $$(x^*, y^*) = (n\pi, 0) \quad n \in \mathbb{Z}$$
+    The case where \( n \) is even (e.g., \( 0, 2, 4, \ldots \)) corresponds to the pendulum hanging straight down (stable equilibrium), while the case where \( n \) is odd (e.g., \( 1, 3, 5, \ldots \)) corresponds to the pendulum being inverted (unstable equilibrium).
+
+    Consider the nullclines:
+    \begin{itemize}
+        \item \( x' = 0 \) (horizontal nullcline): \( y = 0 \)
+        \item \( y' = 0 \) (vertical nullcline): \( \sin(x) = 0 \implies x = n\pi, n \in \mathbb{Z} \) And we have, for $y$:
+
+    \end{itemize}
+\end{example}
+
 \end{document}

### 2025-09-19 YEAR2/PHY293_LectureNotes/main.tex
sept19

@@
 \end{equation}
+And the small angle approximates leads to:
+\begin{equation}
+    \ddot{\theta} + \omega^2 \theta = 0
+\end{equation}
+where \( \omega = \sqrt{\frac{3g}{2L}} \). The solution to this ODE is:
+\begin{equation}
+    \theta(t) = \theta_0 \cos(\omega t + \phi)
+\end{equation}
 The angular frequency of oscillation for the physical pendulum is:
@@
 \end{equation}
+The following are equations that summarize the analogies between mechanical and electrical oscillators:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \textbf{Mechanical Oscillator} & \textbf{Electrical Oscillator} & \textbf{Analogy} \\
+        \hline
+        Mass \( m \) & Inductance \( L \) & Inertia \\
+        Spring constant \( k \) & Inverse capacitance \( \frac{1}{C} \) & Restoring force \\
+        Damping coefficient \( b \) & Resistance \( R \) & Energy dissipation \\
+        Displacement \( x \) & Charge \( q \) & Position \\
+        Velocity \( \dot{x} \) & Current \( I = \frac{dq}{dt} \) & Rate of change of position \\
+        Force \( F \) & Voltage \( V \) & Driving force \\
+        \hline
+    \end{tabular}
+    \caption{Analogies between Mechanical and Electrical Oscillators}
+    \label{tab:oscillator_analogies}
+\end{table}
 
+\subsection{Coupled Oscillators}
+\begin{definition}[Coupled Oscillators]
+    Consider two masses \( m_1 \) and \( m_2 \) on two pendulum connected by springs with spring constant \( k \).  The equations of motion for the two masses are given by:
+
+    \begin{equation}
+        \begin{cases}
+            \ddot{x}_1 &= -\frac{g}{L}x_1 + \frac{k}{m_1} (x_1 - x_2) \quad (a) \\
+            \ddot{x}_2 &= -\frac{g}{L}x_2 - \frac{k}{m_2} (x_1 - x_2) \quad (b)
+        \end{cases}
+    \end{equation}
+    where \( x_1 \) and \( x_2 \) are the displacements of masses \( m_1 \) and \( m_2 \) from their equilibrium positions, respectively. Now if we add and subtract the two equations, we get:
+    $$
+        \begin{cases}
+            \ddot{x}_1 + \ddot{x}_2 &= -\frac{g}{L}(x_1 + x_2) \quad (c) \\
+            \ddot{x}_1 - \ddot{x}_2 &= -\left(\frac{g}{L} + \frac{k}{m_1} + \frac{k}{m_2}\right)(x_1 - x_2) \quad (d)
+        \end{cases}
+    $$
+    Let:
+    $$ 
+    q_1 = x_1 + x_2 \quad \text{and} \quad q_2 = x_1 - x_2
+    $$
+    also, we let:
+    $$
+        \omega_1 = \sqrt{\frac{g}{L}} \quad \text{and} \quad \omega_2 = \sqrt{\frac{g}{L} + \frac{k}{m_1} + \frac{k}{m_2}}
+    $$
+    Then, the equations of motion can be rewritten as:
+    $$
+        \begin{cases}
+            \ddot{q}_1 &= -\omega_1^2 q_1 \\
+            \ddot{q}_2 &= -\omega_2^2 q_2
+        \end{cases}
+    $$
+    Observe that \( q_1 \) and \( q_2 \) are decoupled, and each behaves like a simple harmonic oscillator with angular frequencies \( \omega_1 \) and \( \omega_2 \), respectively.
+The general solutions for \( q_1(t) \) and \( q_2(t) \) are:
+    \begin{equation}
+        \begin{cases}
+            q_1(t) &= A_1 \cos(\omega_1 t + \phi_1) \\
+            q_2(t) &= A_2 \cos(\omega_2 t + \phi_2)
+        \end{cases}
+    \end{equation}
+    where \( A_1, A_2, \phi_1, \) and \( \phi_2 \) are constants determined by the initial conditions.
+Finally, we can expressions for \( x_1(t) \) and \( x_2(t) \) in terms of \( q_1(t) \) and \( q_2(t) \):
+    \begin{equation}
+        \begin{cases}
+            x_1(t) &= \frac{q_1(t) + q_2(t)}{2} = \frac{A_1}{2} \cos(\omega_1 t + \phi_1) + \frac{A_2}{2} \cos(\omega_2 t + \phi_2) \\
+            x_2(t) &= \frac{q_1(t) - q_2(t)}{2} = \frac{A_1}{2} \cos(\omega_1 t + \phi_1) - \frac{A_2}{2} \cos(\omega_2 t + \phi_2)
+        \end{cases}
+    \end{equation}
+    All parts of the system oscillate with the same normal frequency (eigenvalue) in a normal mode (eigenvector).
+\end{definition}
+
+\begin{definition}[Normal Modes]
+    A \textbf{normal mode} is a natural pattern of oscillation in which every part of the system moves in sync with a single frequency.
+
+    For the two-mass pendulum system, there are two normal modes:
+    \begin{itemize}
+        \item \textbf{In-phase mode} (\(\omega_1\)): both masses swing together, reaching maximum displacement on the same side at the same time.  
+        $$
+        x_1 = x_2
+        $$
+
+        \item \textbf{Out-of-phase mode} (\(\omega_2\)): the two masses swing in opposite directions, so when one moves left, the other moves right.  
+        $$
+        x_1 = -x_2
+        $$
+    \end{itemize}
+
+    Any general motion of the system can be written as a combination (superposition) of these two normal modes.
+\end{definition}
+\paragraph{Energy in Coupled Oscillators} The total mechanical energy \( E \) of the coupled oscillator system is given by:
+\begin{equation}
+    E = \frac{1}{2}m_1\dot{x}_1^2 + \frac{1}{2}m_2\dot{x}_2^2 + \frac{1}{2}k(x_1 - x_2)^2 + \frac{1}{2}m_1\frac{g}{L}x_1^2 + \frac{1}{2}m_2\frac{g}{L}x_2^2
+\end{equation}
+Substituting the expressions for \( x_1 = \frac{q_1 + q_2}{2} \) and \( x_2 = \frac{q_1 - q_2}{2} \) into the energy equation, two independent energies terms for $q_1$ and $q_2$ emerge:
+\begin{equation}
+    E = \left[ \frac{1}{4} m \dot{q}_1^2 + \frac{1}{4}\frac{m_1 g}{L} q_1^2 \right] + \left[ \frac{1}{4} m \dot{q}_2^2 + \frac{1}{4}\frac{m_2 g}{L} q_2^2 \right]
+\end{equation}
+This shows that the total energy is the sum of the energies associated with each normal mode, and that no cross term between the modes exists. Thus energy does not flow from one mode to another.
+
+\begin{example}
+    Given two coupled pendulums with $x_0(0) = A$, $x_1(0) = 0$, and both masses at rest initially, we have:
+    \begin{align*}
+        q_0(t) &= C_0 \cos(\omega_0 t) \\
+        q_1(t) &= C_1 \cos(\omega_1 t)
+    \end{align*}
+    where \( C_0 \) and \( C_1 \) are constants determined by the initial conditions. Using the initial conditions. Using the cosine addition formula, which is:
+    \begin{align*}
+        \cos(A) + \cos(B) &= 2\cos\left(\frac{A+B}{2}\right)\cos\left(\frac{A-B}{2}\right) \\
+        \cos(A) - \cos(B) &= -2\sin\left(\frac{A+B}{2}\right)\sin\left(\frac{A-B}{2}\right)
+    \end{align*}
+    we can express the motion of each pendulum as:
+    \begin{align*}
+        x_0(t) &= \frac{q_0 + q_1}{2} = A \cos\left(\frac{\omega_0 + \omega_1}{2} t\right) \cos\left(\frac{\omega_0 - \omega_1}{2} t\right) \\
+        x_1(t) &= \frac{q_0 - q_1}{2} = A \sin\left(\frac{\omega_0 + \omega_1}{2} t\right) \sin\left(\frac{\omega_0 - \omega_1}{2} t\right)
+    \end{align*}
+    Due to the formula for the product of cosines and sines, we see that energy oscillates between the two pendulums with a beat frequency of \( \frac{\omega_1 - \omega_0}{2} \).
+The time for all the energy to transfer from one pendulum to the other is given by:
+    \begin{align*}
+        T &= \frac{2\pi}{\frac{\omega_1 - \omega_0}{2}} = \frac{4\pi}{\omega_1 - \omega_0}
+    \end{align*}
+
+\end{example}
 \section{The Wave Equation}

### 2025-09-20 YEAR2/AER210_LectureNotes/main.tex
sept20

@@
 % set section counter from 15 
-\setcounter{section}{14}
+\setcounter{section}{13}
 \paragraph{Note} the section numbering is based on Stewart's book.
-\section{Double and Triple Integrals}
+
+\section{Partial Derivatives}
+\subsection{Taylor Series for Multivariable Functions}
+\paragraph{Continuing from Calc II} We consider Taylor series of multivariable functions, and we would need to consider partial derivatives.
+
+\begin{shaded}
+    \subsubsection{Review. Taylor Series for Single Variable Functions}
+    Let $f(x)$ be a function that is infinitely differentiable at $x = a$. The Taylor series of $f(x)$ about the point $x = a$ is given by:
+    \begin{equation}
+        f(x) = f(a) + f'(a)(x-a) + \frac{f''(a)}{2!}(x-a)^2 + \frac{f'''(a)}{3!}(x-a)^3 + \cdots = \sum_{n=0}^{\infty} \frac{f^{(n)}(a)}{n!} (x-a)^n
+    \end{equation}
+
+    Also, consider the taylor approximation of $f(x + \Delta x)$ about $x$: 
+    \begin{equation}
+        f(x + \Delta x) = f(x) + f'(x)\Delta x + \frac{f''(x)}{2!}(\Delta x)^2 + \frac{f'''(x)}{3!}(\Delta x)^3 + \cdots = \sum_{n=0}^{\infty} \frac{f^{(n)}(x)}{n!} (\Delta x)^n
+    \end{equation}
+\end{shaded}
+\begin{definition}[Taylor Series for Multivariable Functions]
+    Let $f(x,y)$ be a function that is infinitely differentiable at the point $(a,b)$.
+Consider the increment $\Delta x$ in the $x$-direction and $\Delta y$ in the $y$-direction centred at $(x_0,y_0)$.
+We have the following parametric equations:
+    $$
+        \begin{cases}
+            x = x_0 + \Delta x t \\
+            y = y_0 + \Delta y t
+        \end{cases} \quad t \in [0,1]
+    $$
+    Define a new function $g(t) = f(x_0 + \Delta x t, y_0 + \Delta y t)$, which is a single-variable function in terms of $t$.
+We can then apply the Taylor series for single-variable functions to $g(t)$ about $t = 0$:
+    $$
+        g(t) = g(0) + g'(0)t + \frac{g''(0)}{2!}t^2 + \frac{g'''(0)}{3!}t^3 + \cdots = \sum_{n=0}^{\infty} \frac{g^{(n)}(0)}{n!} t^n
+    $$
+    Using the chain rule, we can compute the taylor series of $f(x,y)$ about the point $(x_0,y_0)$:
+    \begin{align}
+        f(x_0 + \Delta x, y_0 + \Delta y) &= f(x_0,y_0) + \left( \frac{\partial f}{\partial x} \Delta x + \frac{\partial f}{\partial y} \Delta y \right) + \nonumber \\
+        &\quad \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (\Delta x)^2 + 2\frac{\partial^2 f}{\partial x \partial y} \Delta x \Delta y + \frac{\partial^2 f}{\partial y^2} (\Delta y)^2 \right) + \cdots \\
+        &= f(x_0,y_0) + \sum_{n=1}^{\infty} \sum_{k=0}^n \frac{1}{k!(n-k)!} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (\Delta x)^k (\Delta y)^{n-k}
+    \end{align}
+    Or equivalently, we can write:
+    \begin{equation}
+        f(x,y) = f(x_0,y_0) + \sum_{n=1}^{\infty} \sum_{k=0}^n \frac{1}{k!(n-k)!} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (x - x_0)^k (y - y_0)^{n-k}
+    \end{equation}
+    We could also use the fact:
+    $$
+        \frac{1}{k!(n-k)!} = \frac{1}{n!} \binom{n}{k}
+    $$
+    we rewrite the above equations, such that we can use the pascal's triangle to help us remember the coefficients:
+    \begin{equation}
+        f(x,y) = f(x_0,y_0) + \sum_{n=1}^{\infty} \frac{1}{n!} \sum_{k=0}^n \binom{n}{k} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (x - x_0)^k (y - y_0)^{n-k}
+    \end{equation}
+
+\end{definition}
+\section{Multiple Integrals}
 \begin{definition}[Double Integral]
@@
     \end{equation}
-\end{definition}
-
-\begin{example}
+
+    For the change of variables in triple integrals, the volume element $dV$ in cylindrical coordinates is given by:
+    \begin{equation}
+        dV = dA \, dz = r \, dr \, d\theta \, dz
+    \end{equation}
+\end{definition}
+
+\begin{example}
+    Consider the triple integral in the cononical reigion $V$ bounded by the cone $z = \sqrt{x^2 + y^2}$ and the plane $z=2$.
+Changing to cylindrical coordinates, we consider the projection of the cone onto the $xy$-plane as our polar base region $R$.
+Then, we consider how the cone change in the $z$ direction over $R$:
+    $$
+        z = r \implies r \le z \le 2
+    $$
+    So the region $V_{\text{polar}}$ is:
+    $$
+    V_{\text{polar}} = \{(r, \theta, z) \mid 0 \le r \le 2, 0 \le \theta \le 2\pi, r \le z \le 2\}
+    $$
+    So we can set up:
+    $$
+        \iiint_V f(x,y,z) \, dV = \int_0^{2\pi} \int_0^2 \int_r^2 f(r \cos \theta, r \sin \theta, z) \, r \, dz \, dr \, d\theta
+    $$
+
+    Let $f(x,y,z) = x^2 +. y^2 = r^2$. Then, we have:
+    \begin{align*}
+        \iiint_V (x^2 + y^2) \, dV &= \int_0^{2\pi} \int_0^2 \int_r^2 r^2 \cdot r \, dz \, dr \, d\theta \\
+        &= \int_0^{2\pi} \int_0^2 r^3 (2 - r) \, dr \, d\theta = \frac{16\pi}{5}
+    \end{align*}
 
@@
 \subsection{Triple Integrals in Spherical Coordinates}
+\begin{definition}[Spherical Coordinates]
+    The spherical coordinates of a point $P$ in 3-D space are given by the ordered triple $(\rho, \phi, \theta)$, where:
+    \begin{itemize}
+        \item $\rho$ is the distance from the origin to the point $P$,
+        \item $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to point $P$ (also known as the polar angle or colatitude),
+        \item $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to point $P$ onto the $xy$-plane (also known as the azimuthal angle).
+    \end{itemize}
+    The relationships between spherical coordinates and rectangular coordinates are given by:
+    \begin{equation}
+        x = \rho \sin \phi \cos \theta, \quad y = \rho \sin \phi \sin \theta, \quad z = \rho \cos \phi
+    \end{equation}
+    and conversely:
+    \begin{equation}
+        \rho = \sqrt{x^2 + y^2 + z^2}, \quad \phi = \cos^{-1}\left(\frac{z}{\sqrt{x^2 + y^2 + z^2}}\right), \quad \theta = \tan^{-1}\left(\frac{y}{x}\right)
+    \end{equation}
+
+    For the change of variables in triple integrals, the volume element $dV$ in spherical coordinates is given by:
+    \begin{equation}
+        dV = \rho^2 \sin \phi \, d\rho \, d\phi \, d\theta
+    \end{equation}
+\end{definition}
+
+\begin{example}
+    \textit{find the mass of a half sphere of radius $a$ that has a density $ k(2a - \rho)$, where $k$ is a constant and $\rho$ is the distance from the origin.}
+
+    Let $\lambda = k(2a - \rho)$ be the density function. Then, we consider the half sphere $S$ in spherical coordinates:
+    $$       
+    S = \{(\rho, \phi, \theta) \mid 0 \le \rho \le a, 0 \le \phi \le \frac{\pi}{2}, 0 \le \theta \le 2\pi\}
+    $$
+    So we can set up:
+    \begin{align*}
+        M &= \iiint_S \lambda \, dV = \iiint_S k(2a - \rho) \, dV \\
+        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \int_0^a (2a - \rho) \rho^2 \sin \phi \, d\rho \, d\phi \, d\theta \\
+        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \left[ 2a \cdot \frac{\rho^3}{3} - \frac{\rho^4}{4} \right]_{\rho=0}^{\rho=a} \sin \phi \, d\phi \, d\theta \\
+        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \left( \frac{2a^4}{3} - \frac{a^4}{4} \right) \sin \phi \, d\phi \, d\theta \\
+        &= k \cdot \frac{5a^4}{12} \int_0^{2\pi} \left[ -\cos \phi \right]_{\phi=0}^{\phi=\frac{\pi}{2}} \, d\theta = k \cdot \frac{5a^4}{12} \cdot 2\pi \\ 
+        &= \frac{5\pi k a^4}{6}
+    \end{align*}
+
+\end{example}
+
+\subsection{}
 \subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}

### 2025-09-20 YEAR2/CHE260_LectureNotes/main.tex
sept20

@@
 \end{example}
+
+\subsection{Second Law of Thermodynamics and Entropy}
+\paragraph{Motivation} We want to understand the direction of energy transfer and the efficiency of energy conversion processes.
+In an engine or a driven process, we want a measurement for the \textit{quality} of energy, and how much of it is \textit{usable}.
+This is important for understanding how to design efficient systems, and how to analyze thermodynamic processes.
+\begin{definition}[Entropy]
+    Entropy is an extensive property that changes when heat is transferred to or from a system. The entropy change is defined as:
+    \begin{equation}
+        dS = \frac{\delta Q_{rev}}{T}
+    \end{equation}
+    where $dS$ is the change in entropy, $\delta Q_{rev}$ is the reversible heat transfer, and $T$ is the absolute temperature. 
+    Now for the total entropy change of a system, we have:
+    \begin{equation}
+        \Delta S_{system} = S_2 - S_1 = \int_{1}^{2} \frac{\delta Q_{rev}}{T}
+    \end{equation}
+    where $S_1$ and $S_2$ are the initial and final entropies of the system, respectively.
+\end{definition}
+
+\paragraph{Inevitable Entropy Generation} Consider a two systems undergoing heat transfer, we can model the entropy of the entire system as:
+$$
+    \Delta S_{universe} = \frac{Q_{rev}}{T_{hot}} - \frac{Q_{rev}}{T_{cold}}
+$$
+Since $T_{hot} > T_{cold}$, we have $\Delta S_{universe} > 0$.
+This means that for any process that involves heat transfer, the entropy of the universe always increases.
+This is a statement of the second law of thermodynamics.
+
+\begin{theorem}[Second Law of Thermodynamics]
+    The entropy in an isolated system always increases for any real process. This can be expressed as:
+    \begin{equation}
+        \Delta S > 0
+    \end{equation}
+
+    Entropy would generate until it reaches equilibrium, where $\Delta S_{universe} = 0$.
+\end{theorem}
 \chapter{Heat Transfer}

### 2025-09-20 YEAR2/ECE253_LectureNotes/main.tex
sept20

@@
 
+\subsection{Decoders and Encoders}
+\begin{definition}[Decoder]
+    A decoder is a combinational logic circuit that converts binary information from $n$ input lines to a maximum of $2^n$ unique output lines.
+Each output line corresponds to one of the possible combinations of the input lines.
+When a specific combination of input lines is activated (set to 1), the corresponding output line is activated (set to 1), while all other output lines remain inactive (set to 0).
+Decoders are commonly used in applications such as memory address decoding, data multiplexing, and digital display systems.
+
+\end{definition}
+\begin{example}[2-bit 7-Segment Display]
+    A 7-segment display is an electronic display device that consists of seven individual segments (labeled $h_0$ to $h_6$) that can be illuminated in different combinations to represent numerical digits (0-9) and some alphabetic characters.
+The truth table for binary inputs $x_0, x_1, x_2, x_3$ (only $x_0$ and $x_1$ are used in this demonstration) to control the 7-segment display is shown below (note that 0 represents an active segment):
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c||c|c|c|c|c|c|c|}
+            \hline
+            $x_1$ & $x_0$ & $h_0$ & $h_1$ & $h_2$ & $h_3$ & $h_4$ & $h_5$ & $h_6$ \\
+            \hline
+            0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\ % 0
+            0 & 1 & 1 & 0 & 0 & 1 & 1 & 1 & 1 \\ % 1
+            1 & 0 & 0 & 0 & 1 & 0 & 0 & 1 & 0 \\ % 2
+            1 & 1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 \\ % 3
+        \hline
+        \end{tabular}
+        \caption{Truth Table for 7-Segment Display (0-3)}
+        \label{tab:7_segment_display_truth_table}
+    \end{table}
+    So we have:
+    \begin{align*}
+        h_0 &= \overline{x_1}{x_0} \\
+        h_1 &= 0 \\
+        h_2 &= x_1\overline{x_0} \\
+        h_3 &= h_0 = \overline{x_1}{x_0} \\
+        h_4 &= x_0 \\
+        h_5 &= x_0 + x_1 \\
+        h_6 &= \overline{x_1} \\
+    \end{align*}
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module seven_segment(input logic [1:0] x,
+                         output logic [6:0] h);
+        assign h[0] = ~x[1] & x[0];
+        assign h[1] = 1'b0;
+        assign h[2] = x[1] & ~x[0];
+        assign h[3] = ~x[1] & x[0];
+        assign h[4] = x[0];
+        assign h[5] = x[0] | x[1];
+        assign h[6] = ~x[1];
+    endmodule
+    \end{verbatim}
+
+\end{example}
+
+\begin{definition}[If-Else Statement]
+    An if-else statement is a conditional statement that allows for branching in the execution of code based on the evaluation of a boolean expression.
+In Verilog, if-else statements are used to model combinational logic and can be used within always blocks or initial blocks.
+The syntax for an implementation of mux using an if-else statement in Verilog is as follows:
+    \begin{verbatim}
+        module mux(input logic x, y, s,
+                   output logic f);
+            always_comb // Always @(x or y or s) is equivalent.
+            // The sensitivity list gives the signals
+            // that can trigger the always block.
+            begin
+                if (s == 1'b0) // If statements must go inside always block.
+                    f = x; // Keyword 'assign' not used in always block.
+                else
+                    f = y;
+                end
+            end
+        endmodule
+    \end{verbatim}
+    \textbf{Analogy} The always block is like the \verb|useEffect(() => { ... }| hook in React. The sensitivity list is like the dependency array in React.
+\end{definition}
+
+\begin{definition}[Case Statement]
+    A case statement is a control flow statement that allows for multi-way branching based on the value of a single expression.
+In Verilog, case statements are used to model combinational logic and can be used within always blocks.
+The syntax for an implementation of a full 7-segment display using a case statement in Verilog is as follows:
+    \begin{verbatim}
+        module seven_segment_case(input logic [3:0] x,
+                                  output logic [6:0] h);
+            always_comb begin
+                case (x) 
+                    4'b0000: h = 7'b1000000; // 0
+                    4'b0001: h = 7'b1111001; // 1
+                    4'b0010: h = 7'b0100100; // 2
+                    4'b0011: h = 7'b0110000; // 3
+                    4'b0100: h = 7'b0011001; // 4
+                    4'b0101: h = 7'b0010010; // 5
+                    4'b0110: h = 7'b0000010; // 6
+                    4'b0111: h = 7'b1111000; // 7
+                    4'b1000: h = 7'b0000000; // 8
+                    4'b1001: h = 7'b0010000; // 9
+                    default: h = 7'b1111111; // All segments off for invalid input
+                endcase
+            end
+        endmodule
+    \end{verbatim}
+
+\end{definition}
+
 % ===================== PART C: SEQUENTIAL LOGIC =====================

### 2025-09-20 YEAR2/MAT292_LectureNotes/main.tex
sept20

@@
         \end{figure}
-
+        And a figure for the vector field is shown below:
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=0.6\linewidth]{predprey.png}
+            \caption{Vector field of the Lotka-Volterra model showing the circular flow around
+            the fixed point \(\left(\frac{c}{d}, \frac{a}{b}\right)\).}
+        \end{figure}
 \end{example}
@@
     \end{itemize}
+
+    A vector field of the system is shown below:
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=0.6\linewidth]{pendulum.png}
+            \caption{Vector field of the simple pendulum system showing the fixed points at \( (n\pi, 0) \).}
+        \end{figure}
+
+\end{example}
+
+\paragraph{Systems of ODEs as Matrices}
+A system of ODEs can be expressed in matrix form as:
+$$\mathbf{x}' = A\mathbf{x} + \mathbf{b}(t)
+$$where \( \mathbf{x} \) is the state vector, \( A \) is a matrix of coefficients, and \( \mathbf{b}(t) \) is a vector of functions.
+
+\begin{definition}[General System of ODEs]
+    WLOG, for a system of 2 ODEs with $x$ and $y$, we write:
+    $$\begin{cases}
+    x' = f(t, x, y) \\
+    y' = g(t, x, y)
+    \end{cases}$$
+    where \( f \) and \( g \) are given functions:
+    $$
+    f(t, x, y) = p_{11}(t)x + p_{12}(t)y + g_1(t) \\
+    g(t, x, y) = p_{21}(t)x + p_{22}(t)y + g_2(t)
+    $$
+    We can express this in matrix form as:
+    \begin{equation}
+    \begin{bmatrix}x' \\ y'
+    \end{bmatrix} = \begin{bmatrix}
+    p_{11}(t) & p_{12}(t) \\
+    p_{21}(t) & p_{22}(t)
+    \end{bmatrix} \begin{bmatrix}x \\ y
+    \end{bmatrix} + \begin{bmatrix}g_1(t) \\ g_2(t)
+    \end{bmatrix}
+    \end{equation}
+
+    Furthermore, we generalize this to \( n \) dimensions, and the inhomogeneous, non-autonomous system of linears ODEs is given by:
+    \begin{align*}
+    \mathbf{x}'_1 &= \sum^n_{j=1} P_{1j}(t)x_j + g_1(t) \quad i = 1, 2, \ldots, n \\
+     & \vdots \\
+    \mathbf{x}'_n &= \sum^n_{j=1} P_{nj}(t)x_j + g_n(t)
+    \end{align*}
+    or in matrix form:
+    \begin{equation}
+    \mathbf{x}' = A(t)\mathbf{x} + \mathbf{G}(t)
+    \end{equation}
+    where \( A(t) \) is an \( n \times n \) matrix of functions, \( \mathbf{x} \) is the state vector, and \( \mathbf{G}(t) \) is the inhomogeneous term.
+\end{definition}
+
+\begin{theorem}[Existence and Uniqueness for Systems of ODEs]
+    If $p_{ij}(t)$ and $g_i(t)$ are continuous on an open interval $I = (a, b)$ containing $t_0$ for all $i, j = 1, 2, \ldots, n$, then the initial value problem
+    $$\mathbf{x}' = A(t)\mathbf{x} + \mathbf{G}(t), \quad \mathbf{x}(t_0) = \mathbf{x}_0$$
+    has a unique solution on the interval \( I \).
+\end{theorem}
+
+\begin{example}[Linear, homogeneous and autonomous]
+    Consider the simple ODE system:
+    $$
+        \textbf{x}' = A\textbf{x} +\textbf{b}
+    $$
+    We cosidner the fix points by setting \( \textbf{x}' = 0 \):
+    $$A\textbf{x} + \textbf{b} = 0 \implies A\textbf{x} = -\textbf{b}$$
+    If \( A \) is invertible, then the fixed point is given by:
+    $$\textbf{x}^* = -A^{-1}\textbf{b}$$
+    If \( A \) is not invertible, then there are either no fixed points or infinitely many fixed points. Now, we let $\textbf{y} = \textbf{x} + A^{-1}\textbf{b}$, then we have:
+    $$\textbf{y}' =\textbf{x}'$$
+    we have:
+    $$Ay = A\textbf{x} + \textbf{b} = \textbf{x}' = \textbf{y}'$$
+    Thus, the system reduces to:
+    $$\textbf{y}' = A\textbf{y}$$
+    which is a linear, homogeneous, and autonomous system. The fixed point of this system is given by:
+    $$\textbf{y}^* = 0$$
+\end{example}
+
+\begin{example}[Simple ODEs]
+    Consider the system of ODEs:
+    $$
+    \textbf{x}' = \begin{bmatrix} 1 & 0 \\ 0 & -2
+    \end{bmatrix}\textbf{x} 
+    $$
+    and the initial condition:
+    $$
+    \textbf{x}(0) = \begin{bmatrix} x_0 \\ y_0
+    \end{bmatrix}
+    $$
+    Then we can consider the solutions:
+    $$x_1(t) = x_0 e^{t}$$
+    $$x_2(t) = y_0 e^{-2t}$$
+    or we can write the solution in vector form:
+    $$\textbf{x}(t) = x_0 \begin{bmatrix}
+        1 \\ 0
+    \end{bmatrix} e^{t} + y_0 \begin{bmatrix}
+        0 \\ 1
+    \end{bmatrix} e^{-2t}
+    $$
+
+    The fixed point and nullclines are given by:
+    $$\textbf{x}^* = \begin{bmatrix}0 \\ 0
+    \end{bmatrix}$$
+    The nullclines are given by:
+    \begin{itemize}
+        \item \( x' = 0 \): \( x = 0 \) (vertical line)
+        \item \( y' = 0 \): \( y = 0 \) (horizontal line)
+    \end{itemize}
+
+    \textbf{Stability of Nullclines} The $x$-nullcline since it goes to the origin as $t \to -\infty$ and away from the origin as $t \to \infty$, while the $y$-nullcline goes to the origin as $t \to \infty$ and away from the origin as $t \to -\infty$.
+        \begin{figure}[h!]
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                    axis lines = middle,
+                    view     = {0}{90},
+                    xlabel = \( x \),
+                    ylabel = \( y \),
+                    ymin = -3, ymax = 3,
+                    xmin = -3, xmax = 3,
+                    domain=-3:3,
+                    samples=200,
+                    width=0.6\linewidth,
+                    height=6cm,
+                    grid=both,
+                    minor tick num=1,
+                    major grid style={line width=.2pt,draw=gray!50},
+                    minor grid style={line width=.1pt,draw=gray!20},
+                ]
+                % x-nullcline (vertical line)
+                \addplot[blue, thick] coordinates {(0,-3) (0,3)};
+                % y-nullcline (horizontal line)
+                \addplot[red, thick] coordinates {(-3,0) (3,0)};
+                % Fixed point marker
+                \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt,label=above right:{\((0, 0)\)}] {};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Phase portrait of the simple ODE system showing the nullclines and the fixed point at the origin.}
+        \end{figure}
+        \textbf{Note - Stability} The stability here really means that the trajectories approach or move away from the fixed point along the nullclines, this does not mean for perturbation along the nullclines.
+        \textbf{Saddle Node} The fixed point is a saddle node (since one eigenvalue is positive and the other is negative). The trajectories in the phase plane are shown below:
+\end{example}
+
+\begin{example}[Linear Algebra]
+    Consider the system of ODEs:
+    $$
+    \textbf{x}' = A\textbf{x}
+    $$
+    where \( A \) is a constant matrix. Consider:
+    \begin{align*}
+        \textbf{x} &= \vec{v} e^{\lambda t} \\
+        \textbf{x}' &= \lambda \vec{v} e^{\lambda t} \\
+        &= A\vec{v} e^{\lambda t} = A\textbf{x}
+    \end{align*}
+    Thus, we have:
+
+    \begin{equation}
+    A\vec{v} = \lambda \vec{v}
+        \end{equation}
+    which is the eigenvalue problem. 
+\end{example}
+
+\begin{example}[Superposition]
+
+    Assume $x_1(t)$ and $x_2(t)$ are solutions to the system of ODEs:
+    $$\textbf{x}' = A\textbf{x}$$
+    Then, any linear combination of these solutions is also a solution:
+    $$\textbf{x}(t) = c_1 x_1(t) + c_2 x_2(t)$$
+    where \( c_1 \) and \( c_2 \) are constants. This is due to the linearity of the system.
+\end{example}
+
+\begin{definition}[General Eigenproblem for n-dim ODEs System]
+    Combining the above ideas, consider the system:
+    $$\textbf{x}' = A\textbf{x}$$
+    where \( A \in \mathbb{R}^{n \times n} \). We define the \textit{specturm} as:
+    \begin{equation}
+    \sigma(A) = \{(\lambda_i, \vec{v}_i) \in \mathbb{C} \times \mathbb{C}^n : A\vec{v}_i = \lambda_i \vec{v}_i, i = 1, 2, \ldots, n\}
+    \end{equation}
+    where \( \lambda_i \) are the eigenvalues and \( \vec{v}_i \) are the corresponding eigenvectors. The general solution to the system can be expressed as:
+    \begin{equation}
+            \textbf{x}(t) = \sum_{i=1}^{n} c_i \vec{v}_i e^{\lambda_i t}
+    \end{equation}
+    where \( c_i \) are constants determined by the initial conditions.
+\end{definition}
+
+\begin{example}
+    Consider the following system of ODEs:
+    $$
+    \textbf{x}' = \begin{bmatrix} 0 & 1 \\ 1 & 0
+    \end{bmatrix}\textbf{x}
+    $$
+    We find the eigenvalues and eigenvectors by solving the characteristic polynomial:
+    \begin{align*}
+    \det(A - \lambda I) &= \det\begin{bmatrix} -\lambda & 1 \\ 1 & -\lambda
+    \end{bmatrix} \\
+    &= \lambda^2 - 1 = 0 \\
+    \implies \lambda_1 &= 1, \quad \lambda_2 = -1
+    \end{align*}
+    The corresponding eigenvectors are:
+    \begin{align*}
+    \vec{v}_1 &= \begin{bmatrix} 1 \\ 1
+    \end{bmatrix}, \quad \vec{v}_2 = \begin{bmatrix} -1 \\ 1
+    \end{bmatrix}
+    \end{align*}
+    Thus, the general solution to the system is:
+    $$\textbf{x}(t) = c_1 \begin{bmatrix} 1 \\ 1
+    \end{bmatrix} e^{t} + c_2 \begin{bmatrix} -1 \\ 1
+    \end{bmatrix} e^{-t}$$
+    where \( c_1 \) and \( c_2 \) are constants determined by the initial conditions.
+
+    Consider the phase portait.
+If we start on the span of $ [-1, 1]^T$, then we go to the origin as $t \to \infty$.
+If we start on the span of $ [1, 1]^T$, then we go away from the origin as $t \to \infty$.
+Then, the trivial fixed point is a saddle node.
 \end{example}

### 2025-09-20 YEAR2/PHY293_LectureNotes/main.tex
sept20

@@
 \end{example}
+\subsection{Normal Modes}
+\paragraph{Oscilating Masses Connected by Springs} Consider two masses \( m_1 = m_2 = m \) connected by three spring with spring constant \( k \) and fixed to walls on either side.
+The equations of motion for the two masses are given by:
+\begin{equation}
+    \begin{cases}
+        m\ddot{x}_1 &= -k x_1 + k (x_2 - x_1) \quad (a) = m\ddot{x}_1 = -2k x_1 + k x_2 \\
+        m\ddot{x}_2 &= -k x_2 - k (x_2 - x_1) \quad (b) = m\ddot{x}_2 = k x_1 - 2k x_2
+    \end{cases}
+\end{equation}
+To solve these equations, we assume solutions of the form:
+$$
+    \begin{cases}
+        x_1(t) &= A \cos(\omega t + \phi) \\
+        x_2(t) &= B \cos(\omega t + \phi)
+    \end{cases}
+$$
+Substituting these assumed solutions into the equations of motion, we get:
+\begin{equation} \label{eq:coupled_springs}
+    \begin{cases}
+        \frac{A}{B} &= \frac{k}{m\omega^2 - 2k} \\
+        \frac{A}{B} &= \frac{m\omega^2 - 2k}{k}
+    \end{cases}
+\end{equation}
+We can deduce that $A = \pm B$  
+We can set the equation:
+$$
+    (m\omega^2 - 2k)^2 = k^2 \implies m\omega^2 - 2k = \pm k
+$$
+
+So we can derive two normal mode frequencies:
+\begin{equation}
+    \begin{cases}
+        \omega_1 &= \sqrt{\frac{k}{m}} \quad \text{(in-phase mode)} \\
+        \omega_2 &= \sqrt{\frac{3k}{m}} \quad \text{(out-of-phase mode)}
+    \end{cases}
+\end{equation}
+
+\paragraph{Solving as an Eigenproblem} Alternatively, we can create a system of equations from \eqref{eq:coupled_springs}:
+\begin{equation}
+    \begin{bmatrix}
+        \frac{2k}{m} & -\frac{k}{m} \\
+        -\frac{k}{m} & \frac{2k}{m}
+    \end{bmatrix}
+    \begin{bmatrix}
+        A \\ B
+    \end{bmatrix}
+    = \omega^2
+    \begin{bmatrix}
+        A \\ B
+    \end{bmatrix}
+\end{equation}
+To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
+$$
+    \text{det}\left(\begin{bmatrix}
+        \frac{2k}{m} - \omega^2 & -\frac{k}{m} \\
+        -\frac{k}{m} & \frac{2k}{m} - \omega^2
+    \end{bmatrix}\right) = 0
+$$
+Calculating the determinant, we have:
+$$
+    \left(\frac{2k}{m} - \omega^2\right)^2 - \left(-\frac{k}{m}\right)^2 = 0
+$$
+Expanding and simplifying, we get:
+$$
+    \omega^4 - \frac{4k}{m}\omega^2 + \frac{5k^2}{m^2} = 0
+$$
+This gives us the two normal mode frequencies:
+\begin{equation}
+    \begin{cases}
+        \omega_1^2 &= \frac{k}{m} \implies \omega_1 = \sqrt{\frac{k}{m}} \\
+        \omega_2^2 &= \frac{3k}{m} \implies \omega_2 = \sqrt{\frac{3k}{m}}
+    \end{cases}
+\end{equation}
+
+\begin{example}
+    CConsider two equal masses $m$ suspended from identical springs of spring constant $k$.
+The masses are hanged on the celing.
+We can solve this system's frequency via the eigenvalue method.
+The equations of motion are:
+    $$
+        \begin{cases}
+            m\ddot{x}_2 &= -k(x_2 - x_1)\\
+            m\ddot{x}_1 &= -k x_1 + k(x_2 - x_1) = -2k x_1 + k x_2
+        \end{cases}
+    $$
+    Assume solutions of the form:
+    $$        
+    \begin{cases}
+            x_1(t) &= A \cos(\omega t + \phi) \\
+            x_2(t) &= B \cos(\omega t + \phi)
+        \end{cases}
+    $$
+    Substituting these assumed solutions into the equations of motion, we get:
+    $$
+        \begin{cases}
+            -A \omega^2 &= -\frac{k}{m}(-2A + B) \\
+            -B \omega^2 &= -\frac{k}{m}(A - B)
+        \end{cases}
+    $$
+    We can rewrite this as an eigenvalue problem:
+    $$
+        \begin{bmatrix}
+            \frac{2k}{m} & -\frac{k}{m} \\
+            -\frac{k}{m} & \frac{k}{m}
+        \end{bmatrix}
+        \begin{bmatrix}
+            A \\ B
+        \end{bmatrix}
+        = \omega^2
+        \begin{bmatrix}
+            A \\ B
+        \end{bmatrix}
+    $$
+    To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
+    $$
+        \text{det}\left(\begin{bmatrix}
+            \frac{2k}{m} - \omega^2 & -\frac{k}{m} \\
+            -\frac{k}{m} & \frac{k}{m} - \omega^2
+        \end{bmatrix}\right) = 0
+    $$
+    Calculating the determinant, we have:
+    $$
+        \omega^2 = \frac{k}{2m}(3\pm \sqrt{5})
+    $$
+\end{example}
+
 \section{The Wave Equation}

### 2025-09-22 YEAR2/AER210_LectureNotes/main.tex
sept 22

@@
 \section{Partial Derivatives}
+\paragraph{Continuing from Calc II} We consider Taylor series of multivariable functions, and we would need to consider partial derivatives.
+
 \subsection{Taylor Series for Multivariable Functions}
-\paragraph{Continuing from Calc II} We consider Taylor series of multivariable functions, and we would need to consider partial derivatives.
 
@@
     \begin{align}
-        f(x_0 + \Delta x, y_0 + \Delta y) &= f(x_0,y_0) + \left( \frac{\partial f}{\partial x} \Delta x + \frac{\partial f}{\partial y} \Delta y \right) + \nonumber \\
-        &\quad \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (\Delta x)^2 + 2\frac{\partial^2 f}{\partial x \partial y} \Delta x \Delta y + \frac{\partial^2 f}{\partial y^2} (\Delta y)^2 \right) + \cdots \\
+        f(x_0 + \Delta x, y_0 + \Delta y) &= f(x_0,y_0) + \left( \frac{\partial f}{\partial x} \Delta x + \frac{\partial f}{\partial y} \Delta y \right) \nonumber \\
+        & + \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (\Delta x)^2 + 2\frac{\partial^2 f}{\partial x \partial y} \Delta x \Delta y + \frac{\partial^2 f}{\partial y^2} (\Delta y)^2 \right) + \cdots \\
         &= f(x_0,y_0) + \sum_{n=1}^{\infty} \sum_{k=0}^n \frac{1}{k!(n-k)!} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (\Delta x)^k (\Delta y)^{n-k}
@@
 
-\end{definition}
+    WLOG, this can be extended to functions of more than two variables.
+\end{definition}
+
+\begin{example}[Thrid (3$^{\text{rd}}$) order Taylor Polynomial]
+    For some function $f(x,y)$, the third order Taylor polynomial about the point $(x_0,y_0)$ is given by:
+    \begin{align*}
+        f(x,y) &= f(x_0,y_0) + \left( \frac{\partial f}{\partial x} (x - x_0) + \frac{\partial f}{\partial y} (y - y_0) \right) \\
+        &\quad + \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (x - x_0)^2 + 2\frac{\partial^2 f}{\partial x \partial y} (x - x_0)(y - y_0) + \frac{\partial^2 f}{\partial y^2} (y - y_0)^2 \right) \\
+        &\quad + \frac{1}{3!} \left( \frac{\partial^3 f}{\partial x^3} (x - x_0)^3 + 3\frac{\partial^3 f}{\partial x^2 \partial y} (x - x_0)^2 (y - y_0) + 3\frac{\partial^3 f}{\partial x \partial y^2} (x - x_0)(y - y_0)^2 + \frac{\partial^3 f}{\partial y^3} (y - y_0)^3 \right) \\
+        &\quad + O\left( ||(x - x_0, y - y_0)||^4 \right)
+    \end{align*}
+    where $O\left( ||(x - x_0, y - y_0)||^4 \right)$ represents the higher order terms that are of order 4 or higher in the Taylor series expansion.
+\end{example}
+
+\begin{example}[Three Variable Case]
+    For some function $f(x,y,z)$, the second order Taylor polynomial about the point $(x_0,y_0,z_0)$ is given by:
+    \begin{align*}
+        f(x,y,z) &= f(x_0,y_0,z_0) + \left( \frac{\partial f}{\partial x} (x - x_0) + \frac{\partial f}{\partial y} (y - y_0) + \frac{\partial f}{\partial z} (z - z_0) \right) + \\
+        &\quad \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (x - x_0)^2 + 2\frac{\partial^2 f}{\partial x \partial y} (x - x_0)(y - y_0) + 2\frac{\partial^2 f}{\partial x \partial z} (x - x_0)(z - z_0) + \right.
+\\
+        &\quad \left. \frac{\partial^2 f}{\partial y^2} (y - y_0)^2 + 2\frac{\partial^2 f}{\partial y \partial z} (y - y_0)(z - z_0) + \frac{\partial^2 f}{\partial z^2} (z - z_0)^2 \right) + \\
+        &\quad O\left( ||(x - x_0, y - y_0, z - z_0)||^3 \right)
+    \end{align*}
+    where $O\left( ||(x - x_0, y - y_0, z - z_0)||^3 \right)$ represents the higher order terms that are of order 3 or higher in the Taylor series expansion.
+
+\end{example}
+
+\begin{example}
+    \textit{Find the second order Taylor polynomial of $f(x,y) = \sqrt{x^2 + y^3}$ about the point $(1,2)$.}
+
+    First, we compute the necessary partial derivatives at the point $(1,2)$:
+    \begin{align*}
+        f(x,y) &= \sqrt{x^2 + y^3} \implies f(1,2) = \sqrt{1^2 + 2^3} = \sqrt{9} = 3 \\
+        f_x(x,y) &= \frac{x}{\sqrt{x^2 + y^3}} \implies f_x(1,2) = \frac{1}{3} \\
+        f_y(x,y) &= \frac{3y^2}{2\sqrt{x^2 + y^3}} \implies f_y(1,2) = \frac{6}{3} = 2 \\
+        f_{xx}(x,y) &= \frac{y^3}{(x^2 + y^3)^{3/2}} \implies f_{xx}(1,2) = \frac{8}{27} \\
+        f_{yy}(x,y) &= \frac{3x^2}{4(x^2 + y^3)^{3/2}} \implies f_{yy}(1,2) = \frac{3}{27} = \frac{1}{9} \\
+        f_{xy}(x,y) &= -\frac{3xy^2}{2(x^2 + y^3)^{3/2}} \implies f_{xy}(1,2) = -\frac{6}{27} = -\frac{2}{9}
+    \end{align*}
+    Now, we can plug these values into the second order Taylor polynomial formula:
+    \begin{align*}
+        f(x,y) &\approx f(1,2) + f_x(1,2)(x-1) + f_y(1,2)(y-2) \\
+        &\quad + \frac{1}{2} \left( f_{xx}(1,2)(x-1)^2 + 2f_{xy}(1,2)(x-1)(y-2) + f_{yy}(1,2)(y-2)^2 \right) \\
+        &\approx 3 + \frac{1}{3}(x-1) + 2(y-2) + \frac{1}{2} \left( \frac{8}{27}(x-1)^2 - \frac{4}{9}(x-1)(y-2) + \frac{1}{9}(y-2)^2 \right)
+    \end{align*}
+\end{example}
+
+\begin{example}
+    \textit{Find the third-order taylor expansion for $f(x,y) = e^{x-2y}$ about the point $(0,0)$.}
+
+    \textbf{Method 1 (using multivariable taylor series)} We compute the necessary partial derivatives at the point $(0,0)$:
+    \begin{align*}
+        f(x,y) &= e^{x-2y} \implies f(0,0) = e^0 = 1 \\
+        f_x(x,y) &= e^{x-2y} \implies f_x(0,0) = 1 \\
+        f_y(x,y) &= -2e^{x-2y} \implies f_y(0,0) = -2 \\
+        f_{xx}(x,y) &= e^{x-2y} \implies f_{xx}(0,0) = 1 \\
+        f_{yy}(x,y) &= 4e^{x-2y} \implies f_{yy}(0,0) = 4 \\
+        f_{xy}(x,y) &= -2e^{x-2y} \implies f_{xy}(0,0) = -2 \\
+        f_{xxx}(x,y) &= e^{x-2y} \implies f_{xxx}(0,0) = 1 \\
+        f_{yyy}(x,y) &= -8e^{x-2y} \implies f_{yyy}(0,0) = -8 \\
+        f_{xxy}(x,y) &= -2e^{x-2y} \implies f_{xxy}(0,0) = -2 \\
+        f_{xyy}(x,y) &= 4e^{x-2y} \implies f_{xyy}(0,0) = 4
+    \end{align*}
+    Now, we can plug these values into the third order Taylor polynomial formula:
+    \begin{align*}
+        f(x,y) &\approx f(0,0) + f_x(0,0)x + f_y(0,0)y \\
+        &\quad + \frac{1}{2} \left( f_{xx}(0,0)x^2 + 2f_{xy}(0,0)xy + f_{yy}(0,0)y^2 \right) \\
+        &\quad + \frac{1}{6} \left( f_{xxx}(0,0)x^3 + 3f_{xxy}(0,0)x^2y + 3f_{xyy}(0,0)xy^2 + f_{yyy}(0,0)y^3 \right) \\
+        &\approx 1 + x - 2y + \frac{1}{2} \left( x^2 - 4xy + 4y^2 \right) + \frac{1}{6} \left( x^3 - 6x^2y + 12xy^2 - 8y^3 \right)
+    \end{align*}
+    \textbf{Method 2 (using single variable taylor series)} We know that $e^{x - 2y} = e^{x} e^{-2y}$.
+We can find the taylor series of $e^x$ and $e^{-2y}$ about $0$ separately, and then multiply them together.
+Due to margin, this is left as an exercise to the reader.
+\end{example}
 \section{Multiple Integrals}
@@
 \end{example}
-
-\subsection{}
 \subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
+\begin{shaded}
+    \subsubsection{Change of Variables in Single Integrals}
+    Consider the following integral:
+    $$
+        \int_1^3 2x\sqrt{x^2 +1} \, dx
+    $$
+    We can use substitution to solve this integral. Let $u = x^2 + 1$, then $du = 2x \, dx$. Then, we carefully plug in to the integral, and notice how this could extend to multiple integrals:
+    \begin{align*}
+        \int_1^3 2x\sqrt{x^2 +1} \, dx &= \int_{u=2}^{u=10} 2 \sqrt{u} \sqrt{u-1} \cdot \frac{du}{2\sqrt{u-1}} = \int_{u=2}^{u=10} \sqrt{u} \, du \\
+        &= \left[ \frac{2}{3} u^{\frac{3}{2}} \right]_{u=2}^{u=10} = \frac{2}{3} (10^{\frac{3}{2}} - 2^{\frac{3}{2}})
+    \end{align*}
+    The key step is to express $dx$ in terms of $du$:
+    $$        
+    dx = \frac{du}{2x}
+    $$
+    Consider how the area under a curve shifts when we change the variable from $x$ to $u$, we can visualize this as follows:
+    % two minipages side by side one with the graph of y = f(x) from 0 to 3 and the other with the graph of y = f(u) from 2 to 10
+    \end{shaded}
+    \newpage
+
+    \begin{figure}[h!]
+        \centering
+        \begin{minipage}{0.45\textwidth}
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                        axis lines = middle,
+                        xlabel = $x$,
+                        ylabel = $y$,
+                        domain=0:3,
+                        samples=100,
+                        width=6cm,
+                        height=5cm,
+                    ]
+                    % Draw the curve y = f(x) and name the x-axis as a path
+                    \addplot[blue, thick, name path=fcurve] {2*x*sqrt(x^2 + 1)};
+                    \addplot[thick, name path=axisline] {0};
+                    % Draw vertical guide lines (for visual reference)
+                    \draw[black, thick] (axis cs:1,0) -- (axis cs:1,8);
+                    \draw[black, thick] (axis cs:3,0) -- (axis cs:3,18);
+                    % Shade the area under the curve from x=1 to x=3 using fill between curve and axis
+                    \addplot[blue!20, opacity=0.5] fill between[of=fcurve and axisline, soft clip={domain=1:3}];
+                    % Add labels
+                    \node at (axis cs:2,5) {$y = f(x)$};
+                    \node at (axis cs:2,1) {Area};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Area under the curve $y = f(x)$ from $x=1$ to $x=3$.}
+        \end{minipage}
+        \hfill
+        \begin{minipage}{0.45\textwidth}
+            \centering
+            \begin{tikzpicture}
+                \begin{axis}[
+                        axis lines = middle,
+                        xlabel = $u$,
+                        ylabel = $y$,
+                        domain=0:12,
+                        samples=100,
+                        width=6cm,
+                        height=5cm,
+                    ]
+                    % Draw the curve y = f(u) and name the x-axis as a path
+                    \addplot[blue, thick, name path=gcurve] {2*sqrt(x)*sqrt(x - 1)};
+                    \addplot[thick, name path=axisline] {0};
+                    % Shade the area under the curve from u=2 to u=10 using fill between curve and axis
+                    \addplot[blue!20, opacity=0.5] fill between[of=gcurve and axisline, soft clip={domain=2:10}];
+                    % Add labels
+                    \node at (axis cs:6,9) {$y = f(u)$};
+                    \node at (axis cs:6,3) {Area};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Area under the curve $y = f(u)$ from $u=2$ to $u=10$.}
+        \end{minipage}  
+    \end{figure}
+\begin{shaded}
+
+    Although $2x \sqrt{x^2 + 1} = 2\sqrt{u} \sqrt{u-1}$, the area under the curve at any given $x_0$, $u_0 = x_0^2 + 1$ are not necessarily equal: $f(x_0) \neq f(u_0)$.
+This is because the width of each rectangle changes when we change the variable from $x$ to $u$.
+But for small changes, we can approximate the width of each rectangle as follows:
+    $$        
+    \Delta u \approx \frac{du}{dx} \Delta x = 2x \Delta x
+    $$
+    So, the area of each rectangle changes as follows:
+    $$        
+    \text{Area} \approx f(x) \Delta x = f(u) \frac{\Delta u}{2x}
+    $$
+    Taking the limit as $\Delta x \to 0$ (and thus $\Delta u \to 0$), we get:
+    $$        
+    \text{Area} = f(x) \, dx = f(u) \frac{du}{2x}
+    $$
+    This is the key idea behind the change of variables in integrals, and it can be extended to multiple integrals.
+\end{shaded}
+
+\paragraph{Change of Variables in Double Integrals} Consider the following change of variables in double integral $\iint_R f(x,y) \, dA$:
+$$
+\begin{cases}
+    x = g(u,v) \\
+    y = h(u,v)
+\end{cases}
+$$
+Such that we call the corresponding region in the $uv$-plane as $S$.
+Now we assume that there exists a bijective mapping between $(x,y)$ and $(u,v)$, and that the functions $g$ and $h$ have continuous partial derivatives.
+Then, we can express the area element $dA$ in terms of $du$ and $dv$.
+
+We can imagine any uniform small rectangle in the $uv$-plane with sides $\Delta u$ and $\Delta v$. Consider how the four corners of the rectangle map to the $xy$-plane:
+\begin{itemize}
+    \item The corner at $(u,v)$ maps to $(x,y) = (g(u,v), h(u,v))$.
+    \item The corner at $(u + \Delta u, v)$ maps to $(x,y) = (g(u + \Delta u, v), h(u + \Delta u, v))$.
+    \item The corner at $(u, v + \Delta v)$ maps to $(x,y) = (g(u, v + \Delta v), h(u, v + \Delta v))$.
+    \item The corner at $(u + \Delta u, v + \Delta v)$ maps to $(x,y) = (g(u + \Delta u, v + \Delta v), h(u + \Delta u, v + \Delta v))$.
+\end{itemize}
+And we can consider the infinitesimal case as $\Delta u \to 0$ and $\Delta v \to 0$.
+Then, we can approximate the area of the parallelogram (we can prove this is a parallelogram as the dot product of the two sides converge to 0) formed by these four points in the $xy$-plane by the magnitude of the cross product of the two vectors.
+
+First, we consider the first order taylor approximateion of $g$ and $h$ at the above four corners:
+\begin{align*}
+    g(u + \Delta u, v) &\approx g(u,v) + g_u(u,v) \Delta u \\
+    h(u + \Delta u, v) &\approx h(u,v) + h_u(u,v) \Delta u \\
+    g(u, v + \Delta v) &\approx g(u,v) + g_v(u,v) \Delta v \\
+    h(u, v + \Delta v) &\approx h(u,v) + h_v(u,v) \Delta v \\
+    g(u + \Delta u, v + \Delta v) &\approx g(u,v) + g_u(u,v) \Delta u + g_v(u,v) \Delta v \\
+    h(u + \Delta u, v + \Delta v) &\approx h(u,v) + h_u(u,v) \Delta u + h_v(u,v) \Delta v
+\end{align*}
+And then we condsier the vecotr's form of the two sides of the parallelogram:
+\begin{align*}
+    \overrightarrow{A} &= \langle g(u + \Delta u, v) - g(u,v), h(u + \Delta u, v) - h(u,v) \rangle \\
+    &\approx \langle g_u(u,v) \Delta u, h_u(u,v) \Delta u \rangle = \Delta u \langle g_u(u,v), h_u(u,v) \rangle \\
+    \overrightarrow{B} &= \langle g(u, v + \Delta v) - g(u,v), h(u, v + \Delta v) - h(u,v) \rangle \\
+    &\approx \langle g_v(u,v) \Delta v, h_v(u,v) \Delta v \rangle = \Delta v \langle g_v(u,v), h_v(u,v) \rangle
+\end{align*}
+Now, we can compute the area of the parallelogram as follows:
+\begin{align*}
+    \text{Area} &\approx |\overrightarrow{A} \times \overrightarrow{B}| = \left| \begin{vmatrix}
+        \mathbf{i} & \mathbf{j} & \mathbf{k} \\
+        g_u(u,v) \Delta u & h_u(u,v) \Delta u & 0 \\
+        g_v(u,v) \Delta v & h_v(u,v) \Delta v & 0
+    \end{vmatrix} \right| \\
+    &= \left| (0 - 0) \mathbf{i} - (0 - 0) \mathbf{j} + (g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)) \Delta u \Delta v \mathbf{k} \right| \\
+    &= |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \Delta u \Delta v
+\end{align*}
+Taking the limit as $\Delta u \to 0$ and $\Delta v \to 0$, we get:
+$$
+dA = |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \, du \, dv
+$$
+which could be rewritten as the following defintion:
+
+\begin{definition}[Jacobian of a Transformation]
+    Consider the transformation:
+    $$
+    \begin{cases}
+        x = g(u,v) \\
+        y = h(u,v)
+    \end{cases}
+    $$
+    The Jacobian of the transformation is defined as:
+    \begin{equation}
+        J = \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
+            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
+            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
+        \end{vmatrix} = g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)
+        \end{equation}
+    And the area element $dA$ in terms of $du$ and $dv$ is given by:
+    $$
+        dA = |J| \, du \, dv
+    $$
+    Additionally, we if we define the chang of variable the other way around:
+    $$
+    \begin{cases}
+        u = p(x,y) \\
+        v = q(x,y)
+    \end{cases}
+    $$
+    Then, the Jacobian of the transformation is defined as:
+    $$
+        J = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
+            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
+            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
+        \end{vmatrix} = p_x(x,y) q_y(x,y) - p_y(x,y) q_x(x,y)
+    $$
+    And the area element $dA$ in terms of $dx$ and $dy$ is given by:
+    $$        dA = \frac{1}{|J|} \, dx \, dy
+    $$
+
+\end{definition}
 \chapter{Fluid Mechanics}

### 2025-09-22 YEAR2/CHE260_LectureNotes/main.tex
sept 22

@@
 \section{Foundations and Fundamental Concepts}
+\begin{review}[Section overview]
+Key ideas: energy and entropy as fundamental concepts; systems, states, and properties.
+Tips: always sketch system boundary and list knowns/assumptions.
+Units: keep Joule and Watt consistent (W = J/s).
+\end{review}
 \begin{definition}[Energy]
@@
 \paragraph{Efficiency of a Heat Engine} The efficiency of a heat engine is defined as the ratio of the work output to the heat input:
-
-\begin{equation}
-    \eta = \frac{W}{Q_H}
-\end{equation}
+\keyeq{\eta = \frac{W}{Q_H}}
 
@@
 \section{Preview of Thermodynamic Laws and Directionality}
+\begin{review}[Section overview]
+Most-used relations: \noteeq{W = Q_H - Q_C}, sign convention (to system positive), and “heat flows hot→cold.” Pitfall: mixing absolute temperature (K) with Celsius.
+\end{review}
 The first intuitive principle (later formalized) is conservation of energy in transformations between forms:
@@
 In the context of a heat engine, this means that the work output ($W$) is equal to the heat input ($Q_H$) minus the heat rejected ($Q_C$):
-    \begin{equation}
-        W = Q_H - Q_C
-    \end{equation}
+    \keyeq{ W = Q_H - Q_C }
 \end{definition}
@@
     The second law of thermodynamics states that the total entropy of an isolated system can never decrease over time. We consdier entropy as:
-    \begin{equation}
-        S = \frac{Q}{T}
-    \end{equation}
+    \numeq{ S = \frac{Q}{T} }
     In the above example, we can see that entropy increases as heat is transferred from the hot reservoir to the cold reservoir:
 
-    \begin{equation}
-        \Delta S = S_{\text{final}} - S_{\text{initial}} = \frac{Q_C}{T_C} - \frac{Q_H}{T_H} > 0
-    \end{equation}
+    \keyeq{ \Delta S = \frac{Q_C}{T_C} - \frac{Q_H}{T_H} > 0 }
 
@@
 \section{Systems, States, and Properties}
+\begin{review}[Section overview]
+System types: closed (no mass), open/control volume (mass+energy), isolated (no mass or energy).
+Property vs path-dependent: properties (T, P, V, U, m) vs.
+transfers (Q, W).
+Pitfall: use \(\delta W, \delta Q\) for path differentials.
+\end{review}
 \begin{definition}[System]
@@
     (Non-example) Work, Transfer of energy (heat), Mass outside the system (denote $\delta m$)
-    \begin{equation}
-        W = \int dW \quad (\text{depends on path})
-    \end{equation}
+    \numeq{ W = \int \! \delta W \quad (\text{depends on path}) }
     Hence, to remind ourselves, we denote infinitesimal work as $\delta W$ instead of $dW$. Then, we can compute the work done by the force field $\textbf{F}$ along a curve $C$ as:
@@
 \section{Ideal Gas Model and Molecular Basis}
+\begin{review}[Section overview]
+Ideal gas: \noteeq{PV = n R_u T = m R T}.
+Useful processes: isothermal (\noteeq{PV=\text{const}}), adiabatic reversible (\noteeq{PV^\gamma=\text{const}}, \noteeq{TV^{\gamma-1}=\text{const}}).
+Pitfalls: use absolute T (K); check units of R.
+\end{review}
 \begin{definition}[Quasi-Equilibrium Process (Reversible Path)]
@@
 \section{First Law of Thermodynamics (Formal Development)}
+\begin{review}[Section overview]
+Closed system: \noteeq{\Delta U=Q-W}.
+Control volume steady state (single-in/out): \noteeq{\dot W_{shaft}=\dot m(h_1-h_2)+\dot m(\tfrac{V_1^2-V_2^2}{2}+g(z_1-z_2)) - (\dot Q_1-\dot Q_2)}.
+Sign convention: work out of system positive.
+\end{review}
 \begin{shaded}
@@
 \section{Work and Energy Transfer Modes}
+\begin{review}[Section overview]
+Boundary work for quasi-equilibrium: \noteeq{W=\int P\,dV}.
+Useful cases: isobaric \noteeq{W=P\Delta V}, isothermal ideal gas \noteeq{W=nRT\ln(V_2/V_1)}.
+Flow work is included in enthalpy: \noteeq{h=u+Pv}.
+\end{review}
 \begin{definition}[Boundry Work]
@@
 Associate them with $\Delta u$ and $\Delta h$.
-
+\subsection{Control Volume Steady State Energy Equation}
+\begin{review}[Subsection overview]
+Control volume steady state (single-in/out): \\ \noteeq{\dot W_{shaft}=\dot m(h_1-h_2)+\dot m(\tfrac{V_1^2-V_2^2}{2}+g(z_1-z_2)) - (\dot Q_1-\dot Q_2)}. Sign convention: work out of system positive.
+Mass flow rate: \noteeq{\dot m=\rho AV=\frac{AV}{v}}. Ideal gas: \noteeq{Pv=RT}. Fluid energy per unit mass: \noteeq{e=h+\tfrac{V^2}{2}+gz}.
+\end{review}
 \begin{example}
@@
 \subsection{Second Law of Thermodynamics and Entropy}
+\begin{review}[Subsection overview]
+Entropy balance (CV): \noteeq{\dot S_{cv}=\sum \tfrac{\dot Q_j}{T_j}+\sum \dot m_i s_i-\sum \dot m_e s_e+\dot S_{gen}}, with \noteeq{\dot S_{gen}\ge0}.
+Reversible process: \noteeq{dS=\tfrac{\delta Q_{rev}}{T}}.
+\end{review}
 \paragraph{Motivation} We want to understand the direction of energy transfer and the efficiency of energy conversion processes.
@@
     \begin{equation}
-        dS = \frac{\delta Q_{rev}}{T}
-    \end{equation}
-    where $dS$ is the change in entropy, $\delta Q_{rev}$ is the reversible heat transfer, and $T$ is the absolute temperature. 
+        dS = \frac{\delta Q_{\text{rev}}}{T}
+    \end{equation}
+    where $dS$ is the change in entropy, $\delta Q_{\text{rev}}$ is the reversible heat transfer, and $T$ is the absolute temperature. 
     Now for the total entropy change of a system, we have:
     \begin{equation}
-        \Delta S_{system} = S_2 - S_1 = \int_{1}^{2} \frac{\delta Q_{rev}}{T}
+        \Delta S_{\text{system}} = S_2 - S_1 = \int_{1}^{2} \frac{\delta Q_{\text{rev}}}{T}
     \end{equation}
@@
 $$
-    \Delta S_{universe} = \frac{Q_{rev}}{T_{hot}} - \frac{Q_{rev}}{T_{cold}}
-$$
-Since $T_{hot} > T_{cold}$, we have $\Delta S_{universe} > 0$.
+    \Delta S_{\text{universe}} = \frac{Q_{\text{rev}}}{T_{\text{hot}}} - \frac{Q_{\text{rev}}}{T_{\text{cold}}}
+$$
+Since $T_{\text{hot}} > T_{\text{cold}}$, we have $\Delta S_{\text{universe}} > 0$.
 This means that for any process that involves heat transfer, the entropy of the universe always increases.
@@
 \begin{theorem}[Second Law of Thermodynamics]
-    The entropy in an isolated system always increases for any real process. This can be expressed as:
-    \begin{equation}
-        \Delta S > 0
-    \end{equation}
-
-    Entropy would generate until it reaches equilibrium, where $\Delta S_{universe} = 0$.
+    For an isolated system, the entropy never decreases for a real process; it grows until the system reaches equilibrium.
+In a counting formulation this is expressed by the number of accessible microstates \(\Omega\): the equilibrium macrostate is the one with the largest \(\Omega\).
+Using the logarithmic definition below makes these counts add up nicely:
+    \begin{equation}
+        S \equiv \ln \Omega .
+    \end{equation}
+    At equilibrium \(\Omega=\Omega_{\rm eq}\) is (for large systems) overwhelmingly larger than the multiplicities of other macrostates, so the system almost surely occupies the equilibrium macrostate and the entropy is maximal.
 \end{theorem}
+
+\begin{proof}
+    Simple combinatorics gives an instructive example. If we distribute \(N\) indistinguishable energy quanta among \(M\) distinguishable boxes, the number of microstates is
+    \begin{align*}
+        \Omega(\{n_i\}) &= \frac{(N+M-1)!}{\prod_{i=1}^M n_i!\,(M-1)!}
+        \quad\text{and for the total count with only the total \(N\) fixed}\quad \\
+        \Omega_{\rm tot} & =\binom{N+M-1}{N}.
+    \end{align*}
+    Symmetry shows the multiplicity is maximal when the quanta are as equally distributed as possible, i.e.
+\(n_i \approx N/M\) (the most probable macrostate).
+Using Stirling's approximation \(\ln n!\approx n\ln n - n\) one obtains, for small deviations \(\delta n_i\) about the equal distribution,
+    $$
+        \ln\Omega(\{n_i\}) \approx \ln\Omega_{\rm eq} - \frac{1}{2}\sum_{i}\frac{(\delta n_i)^2}{n_{i,\rm eq}} + \cdots.
+    $$
+    The second term shows a Gaussian suppression of multiplicity for deviations: the ratio
+    $$
+        \frac{\Omega(\{n_i\})}{\Omega_{\rm eq}} \approx \exp\!\Big(-\tfrac{1}{2}\sum_i\frac{(\delta n_i)^2}{n_{i,\rm eq}}\Big)
+    $$
+    becomes exponentially small once the deviations are large compared with the typical fluctuation size (~\(\sqrt{N}\)).
+Therefore, as \(N\) grows large, almost all microstates are concentrated in a narrow band around the equal-distribution macrostate and one finds effectively
+    $$
+        \Omega_{\rm eq} \approx \Omega_{\rm tot},
+    $$
+    which implies \(S=\ln\Omega\) is maximal at equilibrium. This concentration of measure is the combinatorial heart of the second law.
+\end{proof}
+
+\begin{definition}[Microstates and Boltzmann Entropy]
+    A microstate is a complete microscopic specification of the system; \(\Omega\) counts how many microstates correspond to a given macroscopic description (a macrostate).
+The entropy defined by the logarithm of that count,
+    \begin{equation}
+        S = \ln \Omega,
+    \end{equation}
+    has the convenient property that independent subsystems add: \(S_{\rm tot}=\ln(\Omega_1\Omega_2)=S_1+S_2\).
+
+    To recover the conventional physical units one multiplies by Boltzmann's constant \(k_B\). The Boltzmann entropy is therefore
+    \begin{equation}
+        S_{\text{Boltzmann}} = k_B \ln \Omega,
+    \end{equation}
+    where \(k_B \approx 1.38\times 10^{-23}\,{\rm J/K}\).
+\end{definition}
+
+\begin{definition}[Maxwell-Boltzmann Distribution]
+    Consider a system with the $n$ particle and energy levels $\epsilon_i$. The system must first satisfy:
+    $$
+        \sum_i n_i = n \quad \text{and} \quad \sum_i n_i \epsilon_i = U
+    $$ 
+    That gives us the number of particles in a particular energy level given a temperature $T$. And this is visualized by the Maxwell-Boltzmann distribution:
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=0.8\textwidth]{MBGraph2.png}
+        \caption{Maxwell-Boltzmann distribution of particles in energy levels}
+    \end{figure}
+\end{definition}
+\begin{definition}[Change of Entropy for an Ideal Gas]
+    The change in entropy of an ideal gas is:
+    \begin{equation}
+        \Delta S = s_2 - s_1 = k_B \ln\left(\frac{\Omega_2}{\Omega_1}\right) \\
+    \end{equation}
+    After some derivation, we have:
+    \begin{equation}
+        \Delta S = mR \left[ \ln\left(\frac{3}{2} \frac{T_2}{T_1}\right) + \ln\left(\frac{V_2}{V_1}\right) \right]
+    \end{equation}
+\end{definition}
 \chapter{Heat Transfer}

### 2025-09-22 YEAR2/ECE253_LectureNotes/main.tex
sept 22

@@
 
-\subsection{Decoders and Encoders}
+\subsection{HEX0 7 Segment Display Decoder}
 \begin{definition}[Decoder]
@@
 
+\subsection{Field Programmable Gate Array (FPGA)}
+\begin{definition}[Gate Array]
+    A gate array is a type of integrated circuit (IC) that consists of a regular array of uncommitted logic gates. These gates can be interconnected to implement various digital functions. 
+\end{definition}
+
+\begin{definition}[Programmable Gate and Lookup Table (LUT)]
+    A programmable gate is a logic gate that can be configured to perform different logic functions based on the input signals.
+A lookup table (LUT) is a memory element that stores the truth table of a logic function.
+The LUT can be programmed to implement any logic function by storing the corresponding output values for each possible combination of input values.
+
+    The implementation of a 2-LUT is via three muxes and 4 memory cells that store the expected output values for each combination of the two input values.
+The three muxes are used to select the appropriate output value based on the input values.
+In general, a $n$-LUT would require $2^n$ memory cells.
+By programming the memory cells with the desired output values, we can configure the LUT to implement any logic function of two variables.
+\end{definition}
+
+\begin{definition}[FPGA Fabric]
+    An FPGA fabric is the underlying architecture of an FPGA that consists of an array of programmable logic blocks (PLBs, i.e.
+LUTs), switch blocks, and connection blocks.
+In addition around the FPGA fabric, there are input/output pads (I/O pads) that provide the interface between the FPGA and the pins in the chip package.
+The FPGA fabric is designed to be highly flexible and reconfigurable, allowing users to implement a wide range of digital circuits and systems.
+
+\end{definition}
+\paragraph{Files} The following files are commonly used in FPGA design:
+\begin{itemize}
+    \item \textbf{.v} or \textbf{.sv} files: These files contain the Verilog or SystemVerilog code that describes the digital circuit or system to be implemented on the FPGA.
+    \item \textbf{.sof} files: These files contain the configuration data that is used to program the FPGA.
+The .sof file is generated by the \textbf{synthesis} and place-and-route tools based on the Verilog or SystemVerilog code.
+    \item \textbf{.qsf} files: These files contain the pin assignments and other configuration settings for the FPGA.
+The .qsf file is used by the synthesis and place-and-route tools to ensure that the design is correctly mapped to the physical resources of the FPGA.
+\end{itemize}
+
+\paragraph{CAD Flow} The following steps are commonly used in the FPGA design process:
+\begin{enumerate}
+    \item \textbf{HDL Coding:} Write the Verilog or SystemVerilog code that describes the digital circuit or system to be implemented on the FPGA.
+    \item \textbf{Synthesis:} Use a synthesis tool to convert the HDL code to logic gates.
+    \item \textbf{Functional Simulation:} Use a simulation tool to verify the functionality of the synthesized design. If corrections are needed, go back to the HDL coding step.
+    \item \textbf{Physical Design (Place-and-Route):} Use a place-and-route tool to map the synthesized design to the physical resources of the FPGA.
+    \item \textbf{Timing Analysis:} In reality, it takes time for signal to propagate through the circuit. In our course, we ignore this.
+    \item \textbf{Generate Bitstream:} Generate the .sof file that contains the configuration data for the FPGA.
+\end{enumerate}
 % ===================== PART C: SEQUENTIAL LOGIC =====================
 \section{Digital Storage Elements}
-% Flip-flops, latches, registers (to be expanded)
+\subsection{Introduction to Sequential Circuits}
+\begin{definition}[Sequential Circuit] 
+    Sequential circuits are digital circuits whose outputs depend not only on the current inputs but also on the history of past inputs.
+In other words, sequential circuits have memory and can store information about previous states.
+This is in contrast to combinational circuits, where the output is solely determined by the current inputs.
+
+    We call sequential circuit a circuit that have \textbf{state}.
+\end{definition}
+\begin{definition}[Cross-Couple RS Latch]
+    A cross-coupled RS latch is a basic memory element that can store one bit of information.
+It consists of two NOR gates connected in a feedback loop.
+The latch has two inputs, labeled R (reset) and S (set), and two outputs, labeled Q and $\overline{Q}$.
+The behavior of the latch is as follows:
+    \begin{itemize}
+        \item When S = 1 and R = 0, the latch is set, and Q = 1 and $\overline{Q} = 0$.
+        \item When S = 0 and R = 1, the latch is reset, and Q = 0 and $\overline{Q} = 1$.
+        \item When S = 0 and R = 0, the latch maintains its previous state (no change).
+        \item When S = 1 and R = 1, this condition is invalid for a NOR-based RS latch, as it would force both outputs to be 0, which violates the requirement that Q and $\overline{Q}$ must always be complements of each other.
+    \end{itemize}
+    The truth table for a cross-coupled RS latch is shown below:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c||c|c|}
+            \hline
+            $S$ & $R$ & $Q$ & $\overline{Q}$ \\
+            \hline
+            0 & 0 & $Q_{prev}$ & $\overline{Q}_{prev}$ \\
+            0 & 1 & 0 & 1 \\
+            1 & 0 & 1 & 0 \\
+            1 & 1 & Invalid & Invalid \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Cross-Coupled RS Latch}
+        \label{tab:rs_latch_truth_table}
+    \end{table}
+\end{definition}
 

### 2025-09-26 YEAR2/AER210_LectureNotes/main.tex
sept 26

@@
     \item The corner at $(u + \Delta u, v)$ maps to $(x,y) = (g(u + \Delta u, v), h(u + \Delta u, v))$.
-    \item The corner at $(u, v + \Delta v)$ maps to $(x,y) = (g(u, v + \Delta v), h(u, v + \Delta v))$.
-    \item The corner at $(u + \Delta u, v + \Delta v)$ maps to $(x,y) = (g(u + \Delta u, v + \Delta v), h(u + \Delta u, v + \Delta v))$.
-\end{itemize}
+    \item The corner at $(u, v + \Delta v)$ maps to $(x,y) = (g(u, v + \Delta v), h(u, v + \Delta v))$.\end{itemize}
 And we can consider the infinitesimal case as $\Delta u \to 0$ and $\Delta v \to 0$.
@@
     h(u, v + \Delta v) &\approx h(u,v) + h_v(u,v) \Delta v \\
-    g(u + \Delta u, v + \Delta v) &\approx g(u,v) + g_u(u,v) \Delta u + g_v(u,v) \Delta v \\
-    h(u + \Delta u, v + \Delta v) &\approx h(u,v) + h_u(u,v) \Delta u + h_v(u,v) \Delta v
 \end{align*}
@@
     \overrightarrow{A} &= \langle g(u + \Delta u, v) - g(u,v), h(u + \Delta u, v) - h(u,v) \rangle \\
-    &\approx \langle g_u(u,v) \Delta u, h_u(u,v) \Delta u \rangle = \Delta u \langle g_u(u,v), h_u(u,v) \rangle \\
+    &\approx \Delta u \langle g_u(u,v), h_u(u,v) \rangle \\
     \overrightarrow{B} &= \langle g(u, v + \Delta v) - g(u,v), h(u, v + \Delta v) - h(u,v) \rangle \\
-    &\approx \langle g_v(u,v) \Delta v, h_v(u,v) \Delta v \rangle = \Delta v \langle g_v(u,v), h_v(u,v) \rangle
+    &\approx \Delta v \langle g_v(u,v), h_v(u,v) \rangle
 \end{align*}
@@
     \begin{equation}
-        J = \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
+        J = \left| \frac{\partial(x,y)}{\partial(u,v)} \right|= \begin{vmatrix}
             \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
@@
     And the area element $dA$ in terms of $dx$ and $dy$ is given by:
-    $$        dA = \frac{1}{|J|} \, dx \, dy
-    $$
-
-\end{definition}
+    $$        
+        dA = \frac{1}{|J|} \, dx \, dy
+    $$
+
+    WLOG, the Jacobian can be defined for any number of variables. The geneal Jacobian for a transformation is defined as:
+    \begin{equation}
+        J = \frac{\partial(x_1, x_2, \ldots, x_n)}{\partial(u_1, u_2, \ldots, u_n)} = \begin{vmatrix}
+            \dfrac{\partial x_1}{\partial u_1} & \dfrac{\partial x_1}{\partial u_2} & \cdots & \dfrac{\partial x_1}{\partial u_n} \\
+            \dfrac{\partial x_2}{\partial u_1} & \dfrac{\partial x_2}{\partial u_2} & \cdots & \dfrac{\partial x_2}{\partial u_n} \\
+            \vdots & \vdots & \ddots & \vdots \\
+            \dfrac{\partial x_n}{\partial u_1} & \dfrac{\partial x_n}{\partial u_2} & \cdots & \dfrac{\partial x_n}{\partial u_n}
+        \end{vmatrix}
+    \end{equation}
+    for the transformation:
+    $$
+    \begin{cases}
+        x_1 = g_1(u_1, u_2, \ldots, u_n) \\
+        x_2 = g_2(u_1, u_2, \ldots, u_n) \\
+        \vdots \\
+        x_n = g_n(u_1, u_2, \ldots, u_n)
+    \end{cases}
+    $$
+\end{definition}
+
+\begin{example}[Change to Polar Coordinates]
+    Consider the change of variables from rectangular coordinates $(x,y)$ to polar coordinates $(r,\theta)$:
+    $$
+    \begin{cases}
+        x = r \cos \theta \\
+        y = r \sin \theta
+    \end{cases}
+    $$
+    We can compute the Jacobian of this transformation as follows:
+    \begin{align*}
+        J &= \frac{\partial(x,y)}{\partial(r,\theta)} = \begin{vmatrix}
+            \dfrac{\partial x}{\partial r} & \dfrac{\partial x}{\partial \theta} \\
+            \dfrac{\partial y}{\partial r} & \dfrac{\partial y}{\partial \theta}
+        \end{vmatrix} = \begin{vmatrix}
+            \cos \theta & -r \sin \theta \\
+            \sin \theta & r \cos \theta
+        \end{vmatrix} \\
+        &= r (\cos^2 \theta + \sin^2 \theta) = r
+    \end{align*}
+    So, the area element $dA$ in terms of $dr$ and $d\theta$ is given by:
+    $$        
+        dA = |J| \, dr \, d\theta = r \, dr \, d\theta
+    $$
+    This is consistent with the formula for the area element in polar coordinates.
+\end{example}
+
+\begin{example}
+    Cosndier the interal:
+    $$
+        \iint_R (x^2 + 2xy) \, dA
+    $$
+    where $R$ the region bounded by the lines $y=2x + 3$ and $y=2x + 1$, $y=5-x$ and $y=2 - x$.
+
+    Motivated by the bounds, we can use the change of variables:
+    $$
+    \begin{cases}
+        u = y - 2x \\
+        v = y + x
+    \end{cases}
+    $$ 
+    So, let $S$ be the corresponding region in the $uv$-plane. We can find the bounds of $S$ by plugging in the lines that bound $R$:
+    \begin{itemize}
+        \item $y = 2x + 3 \implies u = 3$
+        \item $y = 2x + 1 \implies u = 1$
+        \item $y = 5 - x \implies v = 5$
+        \item $y = 2 - x \implies v = 2$
+    \end{itemize}
+
+    Then, we consider the Jacobian. First, we express $x$ and $y$ in terms of $u$ and $v$:
+    $$
+    \begin{cases}
+        u = y - 2x \\
+        v = x + y
+    \end{cases} \implies \begin{cases}
+        x = \frac{1}{3} (v - u) \\
+        y = \frac{1}{3} (2v + u)
+    \end{cases}
+    $$
+    Then, we can compute the Jacobian of this transformation as follows:
+    \begin{align*}
+        J &= \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
+            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
+            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
+        \end{vmatrix} = \begin{vmatrix}
+            -\frac{1}{3} & \frac{1}{3} \\
+            \frac{1}{3} & \frac{2}{3}
+        \end{vmatrix} \\
+        &= -\frac{2}{9} - \frac{1}{9} = -\frac{1}{3}
+    \end{align*}
+    So, we can express the integral in terms of $u$ and $v$:
+    \begin{align*}
+        \iint_R (x^2 + 2xy) \, dA &= \iint_S \left( \left(\frac{1}{3} (v - u)\right)^2 + 2 \cdot \frac{1}{3} (v - u) \cdot \frac{1}{3} (2v + u) \right) |J| \, du \, dv \\
+        &= \iint_S \left( \frac{1}{9} (v - u)^2 + \frac{2}{9} (v - u)(2v + u) \right) \cdot \frac{1}{3} \, du \, dv \\
+        &= \frac{1}{27} \iint_S (5v^2 - u^2) \, du \, dv \\
+        &= \frac{196}{27}
+    \end{align*}
+\end{example}
+
+\begin{example} \label{ex:change_of_variables}
+    \textit{Evaluate the integral}
+    $$
+        \iint_R xy \, dA
+    $$
+    where $R$ is the region bounded by curves $x^2 + y^2 = 4$, $x^2 + y^2 = 9$, $x^2 - y^2 = 1$, and $x^2 - y^2 = 4$.
+
+    Motivated by the bounds, we can use the change of variables:
+    $$
+    \begin{cases}
+        u = x^2 + y^2 \\
+        v = x^2 - y^2
+    \end{cases} \implies \begin{cases}
+        x = \sqrt{\frac{u + v}{2}} \\
+        y = \sqrt{\frac{u - v}{2}}
+    \end{cases}
+    $$
+    The jacobian of this transformation is:
+    \begin{align*}
+        J &= \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
+            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
+            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
+        \end{vmatrix} = \begin{vmatrix}
+            \frac{1}{2\sqrt{2(u+v)}} & \frac{1}{2\sqrt{2(u+v)}} \\
+            \frac{1}{2\sqrt{2(u-v)}} & -\frac{1}{2\sqrt{2(u-v)}}
+        \end{vmatrix} \\
+        &= -\frac{1}{8} \left( \frac{1}{\sqrt{u^2 - v^2}} + \frac{1}{\sqrt{u^2 - v^2}} \right) = -\frac{1}{4\sqrt{u^2 - v^2}}
+    \end{align*}
+    So, we can express the integral in terms of $u$ and $v$:
+    \begin{align*}
+        \iint_R xy \, dA &= \iint_S \sqrt{\frac{u + v}{2}} \cdot \sqrt{\frac{u - v}{2}} \cdot |J| \, du \, dv \\
+        &= \iint_S \frac{\sqrt{u^2 - v^2}}{2} \cdot \frac{1}{4\sqrt{u^2 - v^2}} \, du \, dv = \iint_S \frac{1}{8} \, du \, dv \\
+        &= \int_1^4 \int_4^9 \frac{1}{8} \, dv \, du = \frac{1}{8}(4 - 1)(9 - 4) = \frac{15}{8}
+    \end{align*}
+
+\end{example}
+
+\begin{theorem}[Back Transformation]
+    Consider the Jacobian $J$ of a transformation:
+    $$
+    \begin{cases}
+        x = g(u,v) \\
+        y = h(u,v)
+    \end{cases}
+    $$
+    If $J \neq 0$, then the Jacobian of the inverse transformation:
+    $$
+    \begin{cases}
+        u = p(x,y) \\
+        v = q(x,y)
+    \end{cases}
+    $$
+    is given by:
+    \begin{equation}
+        J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \frac{1}{\frac{\partial(x,y)}{\partial(u,v)}}
+    \end{equation}
+\end{theorem}
+\begin{proof}
+    % a short intuitive proof sketch
+    \textbf{Sketch:} Consider a small rectangle in the $uv$-plane with sides $\Delta u$ and $\Delta v$.
+The area of this rectangle is $\Delta A_{uv} = \Delta u \Delta v$.
+This rectangle maps to a parallelogram in the $xy$-plane with area $\Delta A_{xy} \approx |J| \Delta u \Delta v$.
+Now, consider the inverse transformation.
+A small rectangle in the $xy$-plane with sides $\Delta x$ and $\Delta y$ maps back to a parallelogram in the $uv$-plane with area $\Delta A_{uv} \approx |J^{-1}| \Delta x \Delta y$.
+Since these areas must be consistent under the transformations, we have:
+    $$
+        |J| \Delta u \Delta v = \Delta x \Delta y \quad \text{and} \quad |J^{-1}| \Delta x \Delta y = \Delta u \Delta v
+    $$
+    Dividing the first equation by the second, we get:
+    $$
+        |J| = \frac{1}{|J^{-1}|}
+    $$
+\end{proof}
+
+\begin{example}
+    Consider the transformation and integral in \ref{ex:change_of_variables}. We can verify the result using the back transformation theorem. The Jacobian of the inverse transformation is:
+    \begin{align*}
+        J^{-1} &= \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
+            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
+            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
+        \end{vmatrix} = \begin{vmatrix}
+            2x & 2y \\
+            2x & -2y
+        \end{vmatrix} \\
+        &= -4xy - 4xy = -8xy
+    \end{align*}
+    So, we have:
+    $$        
+        J = \frac{1}{J^{-1}} = -\frac{1}{8xy}
+    $$
+    This is consistent with the result we obtained in \ref{ex:change_of_variables}, and results in a much simpler integral to evaluate.
+\end{example}
 \chapter{Fluid Mechanics}

### 2025-09-26 YEAR2/ECE253_LectureNotes/main.tex
sept 26

@@
 \end{definition}
-\begin{definition}[Cross-Couple RS Latch]
+\begin{definition}[Cross-Couple RS-Latch]
     A cross-coupled RS latch is a basic memory element that can store one bit of information.
@@
         \item When S = 1 and R = 1, this condition is invalid for a NOR-based RS latch, as it would force both outputs to be 0, which violates the requirement that Q and $\overline{Q}$ must always be complements of each other.
+Additoanlly, this would break the logic for subsequent operations.
     \end{itemize}
@@
             1 & 0 & 1 & 0 \\
-            1 & 1 & Invalid & Invalid \\
+            1 & 1 & Invalid (0) & Invalid (0) \\
             \hline
@@
 
+\begin{definition}[Timing Diagram]
+    A timing diagram is a graphical representation of the relationship between signals in a digital circuit over time.
+It shows how the signals change.
+The timing diagram of a RS Latch could be visualized as follows:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}
+            % Time axis
+            \draw[->] (0,0) -- (10,0) node[right] {Time};
+
+            % S signal
+            \draw (0,1) -- (1,1) -- (1,2) -- (3,2) -- (3,1) -- (5,1) -- (5,2) -- (7,2) -- (7,1) -- (10,1);
+            \node at (-0.5,1) {S};
+
+            % R signal
+            \draw (0,-1) -- (2,-1) -- (2,-2) -- (4,-2) -- (4,-1) -- (6,-1) -- (6,-2) -- (8,-2) -- (8,-1) -- (10,-1);
+            \node at (-0.5,-1) {R};
+
+            % Q signal
+            \draw[thick] (0,-3) -- (4,-3) -- (4,-4) -- (6, -4) -- (6.5, -3) -- (7,-4) -- (7.5, -3) -- (8,-4) -- (8.5, -3) -- (9,-4);
+            \node at (-0.5,-3) {Q};
+
+            % Q' signal
+            \draw[thick] (0,-5.3) -- (4,-5.3) -- (4,-4.3) -- (5,-4.3) -- (5,-5.3) -- (6,-5.3);
+            \node at (-0.5,-5.3) {$\overline{Q}$};
+
+            % Time markers
+            \foreach \x in {0,1,...,10} {
+                \draw[dotted] (\x,3) -- (\x,-5.4);
+                \node at (\x,-6) {\x};
+            }
+        \end{tikzpicture}
+        \caption{Timing Diagram of a Cross-Coupled RS Latch}
+        \label{fig:rs_latch_timing_diagram}
+        \centering
+    \end{figure}
+
+\end{definition}
+
+\begin{definition}[Gated RS-Latch]
+    We can put an AND gate along with the R and S signals, and the circuit will only reset to 0 or set to 1 when the clock is HI. The truth table of a gated RS latch is:
+
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c||c|}
+            \hline
+            CLK & $S$ & $R$ & $Q(t+1)$ \\
+            \hline
+            0 & X & X & $Q(t)$ \\
+            1 & 0 & 0 & 0 \\
+            1 & 0 & 1 & 1 \\
+            1 & 1 & 0 & 1 \\
+            1 & 1 & 1 & Invalid \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Gated RS Latch}
+        \label{tab:rs_latch_truth_table}
+    \end{table}
+    \newpage
+    Additionally, we can use purely NAND gates instead of NOR gates and AND gates, using a technique called bubble pushing. A implementation is shown in the below figure.
+    \begin{figure}[h!]
+        \includegraphics[width=0.5\textwidth]{RS-with-NAND.png}
+        \centering
+        \caption{Gated SR latch using NAND}
+    \end{figure}
+\end{definition}
+
+\begin{definition}[Gated D-Latch]
+    We can set $D = S = \overline{R}$, so we would prevent $S = R = 1$.
+So when the clock is HI, $Q$ tracks $D$ (a.k.a.
+Transparent Mode), and when clock is high, it won't store (a.k.a.
+Opaque Mode).
+The truth table is shown as follows:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c||c|}
+            \hline
+            CLK & $D$ & $Q(t+1)$ \\
+            \hline
+            0 & X & $Q(t)$ \\
+            1 & 0 & 0 \\
+            1 & 1 & 1 \\ 
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Gated D Latch}
+        \label{tab:gated_d_latch_truth_table}
+    \end{table}
+\end{definition}
+
+\paragraph{Equivalence with a mux} If we feed back the output of a mux back to one of the input, this is essentially a gated D latch with the SW as CLK and the input is the D.
+
+\begin{definition}[D-Flip Flops]
+    A D flip-flop is a type of flip-flop that captures the value of the D input at a specific portion of the clock cycle (rising edge) and holds that value until the next clock cycle.
+It is done by chaining the output of a D latch to the input of another D-latch.
+While the second D-latch recognizes the CLK, and the first one recognizes $\overline{\text{CLK}}$.
+\end{definition}
 \section{Finite State Machines (FSM)}
@@
 \section{Computer Organization}
-
 \section{Assembly Language}

### 2025-09-26 YEAR2/MAT292_LectureNotes/main.tex
sept 26

@@
 
+\begin{definition}[Computing Eigenvalues via Trace and Determinant]
+    Consider matrix $A \in \mathbb{R}^{2 \times 2}$. We compute the eigenvalues by solving the characteristic polynomial:
+    \begin{align*}
+    0 &= \det(A - \lambda I) \\
+    &= \lambda^2 - \text{tr}(A)\lambda + \det(A) \\
+    \implies \lambda_{1,2} &= \frac{\text{tr}(A) \pm \sqrt{\text{tr}(A)^2 - 4\det(A)}}{2}
+    \end{align*}
+
+    If the roots are real, there are four cases that coud emerge from this (WLOG): 
+    \begin{itemize}
+        \item $\lambda_1 > \lambda_2 > 0$: We see much faster growth in the span assocaited with $\lambda_1$.
+        \item $\lambda_1 < \lambda_2 < 0$: We see much faster decay in the span assocaited with $\lambda_1$.
+        \item $\lambda_1 < 0 < \lambda_2$: We would see the span assocaited with $\lambda_2$ dominate as $t \to \infty$ and the span assocaited with $\lambda_1$ dominate as $t \to -\infty$.
+This is a saddle node.
+        \item $\lambda_2  = 0 < \lambda_1$: Along the span of $v_2$, we would see horizontal line, as we growth only along the span of $v_1$. 
+        \item $\lambda_1 = \lambda_2 > 0$: Along the span of $v_2$, we would see horizontal line, as we decay only along the span of $v_1$.
+        \item $\lambda_1 = \lambda_2 = 0$: The system is trivial, and all points are fixed points.
+        \item $\lambda_1 = \lambda_2 < 0$: It is a inward star that everything goes to the origin.
+        \item $\lambda_1 = \lambda_2 > 0$: It is an outward star that everything goes away from the origin.
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Complex Eigenvalues]
+    Consider the case where the eigenvalues are complex, i.e., \( \lambda = \alpha \pm i\beta \) with \( \beta \neq 0 \). The general solution to the system can be expressed as:
+    $$\textbf{x}(t) = e^{\alpha t} \left( c_1 \text{Re}(\vec{v} e^{i\beta t}) + c_2 \text{Im}(\vec{v} e^{i\beta t}) \right)$$
+    where \( \vec{v} \) is the complex eigenvector corresponding to \( \lambda \), and \( c_1, c_2 \) are constants determined by the initial conditions.
+
+    This would lead to the following definition of a radial and angular component:
+    \begin{equation}
+    v(t) = ve^{\alpha t} \mathrm{cis} (\beta t)
+    \end{equation}
+    where \( r(t) \) represents the radial distance from the origin and \( \theta(t) \) represents the angular position. So, for a general solution, we have:
+    \begin{equation}
+    \textbf{x}(t) = \sum_i c_i v_i e^{\alpha_i t} \mathrm{cis} (\beta_i t)
+    \end{equation}
+    The behavior of the system depends on the sign of \( \alpha \) and $\beta$:
+    \begin{itemize}
+        \item If \( \alpha > 0 \) the system spirals outward (complex source).
+        \begin{itemize}
+            \item If \( \beta > 0 \), the systems sprial counter-clockwise.
+            \item If \( \beta < 0 \), the systems sprial clockwise.
+        \end{itemize}
+        \item If \( \alpha < 0 \) the system spirals inward (complex sink).
+        \begin{itemize}
+            \item If \( \beta > 0 \), the systems sprial counter-clockwise.
+            \item If \( \beta < 0 \), the systems sprial clockwise.
+        \end{itemize}
+        \item If \( \alpha = 0 \) the system is a center (closed orbits).
+        \begin{itemize}
+            \item If \( \beta > 0 \), the systems sprial counter-clockwise.
+            \item If \( \beta < 0 \), the systems sprial clockwise.       
+        \end{itemize}
+    \end{itemize}
+\end{definition}
+
+\paragraph{Complex Eigenvalues Resolves to Real Numbers} Consider eigenvalues $\lambda = \alpha \pm i\beta \in \mathbb{C}$ with $v = \textbf{p} + i\textbf{q}\in \mathbb{C}^n$.
+Then, the general solution $x: \mathbb{R} \to \mathbb{R}^n$ is given by:
+\begin{equation}
+\textbf{x}(t) = c_1 e^{\lambda t} \textbf{v} + c_2 e^{\bar{\lambda} t} \bar{\textbf{v}}
+\end{equation}
+where $c_1, c_2 \in \mathbb{C}$.
+
+We also have:
+$$
+    \bar{\textbf{x}} = \bar{c_1} e^{\bar{\lambda} t} \bar{\textbf{v}} + \bar{c_2} e^{\lambda t} \textbf{v}
+$$
+So, we can say that:
+$$
+c_1 = \bar{c_2}, \quad c_2 = \bar{c_1}
+$$
+Let $c_1 =  \delta + i\gamma$, then $c_2 = \delta - i\gamma$. Thus, we have:
+\begin{equation}
+\textbf{x}(t) = (\delta + i\gamma)e^{(\alpha + i\beta)t}(\textbf{p} + i\textbf{q}) + (\delta - i\gamma)e^{(\alpha - i\beta)t}(\textbf{p} - i\textbf{q}) \\
+\end{equation}
+Seperating the real and imaginary parts, we have:
+\begin{equation}
+\textbf{x}(t) = \delta (ve^{\lambda t} + \bar{v}e^{\bar{\lambda} t}) + i\gamma (ve^{\lambda t} - \bar{v}e^{\bar{\lambda} t})
+\end{equation}
+By the Euler's formula, we have:
+\begin{equation}
+\textbf{x}(t) = 2\delta \mathrm{Re}(ve^{\lambda t}) - 2\gamma \mathrm{Im}(ve^{\lambda t}) \in \mathbb{R}^n
+\end{equation}
+
+\paragraph{Generalized eigenvectors and solution for a defective matrix}
+Consider the matrix
+\[
+A=\begin{bmatrix}2 & 1\\[4pt]0 & 2\end{bmatrix},
+\]
+which has a double eigenvalue \(\lambda=2\) but only one linearly independent eigenvector
+\[
+v=\begin{bmatrix}1\\[4pt]0\end{bmatrix}.
+\]
+
+To complete a Jordan chain of length two we seek a generalized eigenvector \(v_g\) satisfying
+\begin{equation}
+(A-\lambda I)\,v_g = v. \label{eq:gen-eig}
+\end{equation}
+For this \(A\) we have \(A-2I=\begin{bmatrix}0&1\\0&0\end{bmatrix}\). Choosing the convenient solution
+\[
+v_g=\begin{bmatrix}0\\[4pt]1\end{bmatrix}
+\]
+indeed yields \((A-2I)v_g=v\), so \eqref{eq:gen-eig} holds.
+
+The two linearly independent solutions corresponding to this Jordan chain are
+\begin{equation}
+x_1(t)=v\,e^{\lambda t} = \begin{bmatrix}1\\[4pt]0\end{bmatrix} e^{2t}, \label{eq:sol1}
+\end{equation}
+and
+\begin{equation}
+x_2(t)=(v_g + t\,v)\,e^{\lambda t} = \left(\begin{bmatrix}0\\[4pt]1\end{bmatrix} + t\begin{bmatrix}1\\[4pt]0\end{bmatrix}\right)e^{2t}. \label{eq:sol2}
+\end{equation}
+One checks directly that these satisfy \(x_i'(t)=Ax_i(t)\) using \(Av=\lambda v\) and \(Av_g=\lambda v_g + v\).
+
+Therefore the general solution of \(\mathbf{x}'=A\mathbf{x}\) is the linear combination
+\begin{equation}
+\mathbf{x}(t)=C_1\,x_1(t) + C_2\,x_2(t)
+= e^{\lambda t} \begin{bmatrix}
+C_1 + C_2 t \\[4pt] C_2
+\end{bmatrix}
+\end{equation}
+The phase portrait of this system is shown below:
+% insert later
+
+\begin{definition}[Trace-Determinant Plane]
+    We found that we can find the eigenvalues of a \( 2 \times 2 \) matrix \( A \) using:
+    $$
+    \lambda_{1,2} = \frac{\text{tr}(A) \pm \sqrt{\text{tr}(A)^2 - 4\det(A)}}{2}
+    $$
+    where
+    \begin{itemize}
+        \item \( \text{tr}(A) = \lambda_1 + \lambda_2 = a_{11} + a_{22} \) is the trace of \( A \)
+        \item \( \det(A) = \lambda_1 \lambda_2 = a_{11}a_{22} - a_{12}a_{21} \) is the determinant of \( A \)   
+    \end{itemize}
+    And we have the following analysis:
+    \begin{itemize}
+        \item If \( \det(A) < 0 \) and \( \text{tr}(A) > 0 \), then \( \lambda_1 > 0 > \lambda_2 \) (saddle node).
+        \item Consider the discriminant boundry \( \text{tr}(A)^2 - 4\det(A) = 0 \), it is a parabola that upward opening with vertex at the origin.
+Consider region above that, where there is complex eigenvalues:
+        \begin{itemize}
+            \item If \( \text{tr}(A) > 0 \) and \( \det(A) > 0 \), then \( \alpha = \text{tr}(A)/2 > 0 \) (spiral source).
+            \item If \( \text{tr}(A) < 0 \) and \( \det(A) > 0 \), then \( \alpha = \text{tr}(A)/2 < 0 \) (spiral sink).
+            \item If \( \text{tr}(A) = 0 \) and \( \det(A) > 0 \), then \( \lambda_1 = -\lambda_2 = i\beta \) (center).
+        \end{itemize}
+        \item Consider the region below the discriminant boundry, where there is real eigenvalues:
+        \begin{itemize}
+            \item If \( \text{tr}(A) > 0 \) and \( \det(A) > 0 \), then \( \lambda_1 \geq \lambda_2 > 0 \) (nodal source).
+            \item If \( \text{tr}(A) < 0 \) and \( \det(A) > 0 \), then \( 0 > \lambda_1 \geq \lambda_2 \) (nodal sink).
+        \end{itemize}
+        \item \textbf{Degenerate Star Node} Consider the line \( \text{tr}(A)^2 - 4\det(A) = 0 \).
+        \item \textbf{Lines of Nodes} Consider the line \( \det(A) = 0 \).
+    \end{itemize}
+\end{definition}
+
 \end{document}

### 2025-09-26 YEAR2/PHY293_LectureNotes/main.tex
sept 26

@@
 
-\section{The Wave Equation}
+\begin{example}
+    Consider three equal masses \( m \) connected by four identical springs with spring constant \( k \) and fixed to walls on either side. The equations of motion for the three masses are given by:
+    $$
+        \begin{cases}
+            m\ddot{x}_1 &= -2k x_1 + k x_2 \\
+            m\ddot{x}_2 &= -2k x_2 + k x_1 + k x_3 \\
+            m\ddot{x}_3 &= -2k x_3 + k x_2
+        \end{cases}
+    $$
+    Assume solutions of the form:
+    $$        
+    \begin{cases}
+            x_1(t) &= A \cos(\omega t + \phi) \\
+            x_2(t) &= B \cos(\omega t + \phi) \\
+            x_3(t) &= C \cos(\omega t + \phi)
+        \end{cases}
+    $$
+    Substituting these assumed solutions into the equations of motion, we get:
+    $$      
+        \begin{cases}
+            A \omega^2 &= -\frac{k}{m}(-2A + B) \\
+            B \omega^2 &= -\frac{k}{m}(A - 2B + C) \\
+            C \omega^2 &= -\frac{k}{m}(B - 2C)
+        \end{cases}
+    $$
+    We can rewrite this as an eigenvalue problem:
+    $$
+        \begin{bmatrix}
+            -\frac{2k}{m} & \frac{k}{m} & 0 \\
+            \frac{k}{m} & -\frac{2k}{m} & \frac{k}{m} \\
+            0 & \frac{k}{m} & -\frac{2k}{m}
+        \end{bmatrix}
+        \begin{bmatrix}
+            A \\ B \\ C
+        \end{bmatrix}
+        = \omega^2
+        \begin{bmatrix}
+            A \\ B \\ C
+        \end{bmatrix}
+    $$
+    To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
+    $$
+        \text{det}\left(\begin{bmatrix} 
+            -\frac{2k}{m} - \omega^2 & \frac{k}{m} & 0 \\
+            \frac{k}{m} & -\frac{2k}{m} - \omega^2 & \frac{k}{m} \\
+            0 & \frac{k}{m} & -\frac{2k}{m} - \omega^2
+        \end{bmatrix}\right) = 0
+    $$
+    Calculating the determinant, we have:
+    $$        \left(\frac{2k}{m} - \omega^2\right)\left[\left(-\frac{2k}{m} - \omega^2\right)^2 - \left(\frac{k}{m}\right)^2\right] - \left(\frac{k}{m}\right)^2\left(\frac{2k}{m} - \omega^2\right) = 0
+    $$   
+    This gives us the three normal mode frequencies:
+    $$
+        \begin{cases}
+            \omega_1 &= \sqrt{\frac{2k}{m}} \\
+            \omega_2 &= \sqrt{\frac{2k}{m}(2 - \sqrt{2})} \\
+            \omega_3 &= \sqrt{\frac{2k}{m}(2 + \sqrt{2})}
+        \end{cases}
+    $$
+\end{example}
+\paragraph{Intuition} For any perturbation, we can expressed it as a linear combination of the excitations of several normal modes.
+Each normal mode oscillates at its own frequency, and the overall motion is a superposition of these modes.
+\section{Waves}
+
+\subsection{Travelling Waves}
+
+\begin{definition}[Travelling Pulse]
+    A travelling pulse is a disturbance that moves through a medium. For any wave function $f(x,t)$, if it satisfies the property:
+    \begin{equation} \label{eq:travelling_wave}
+        y(x,t) = f(x \pm vt)
+    \end{equation}
+    then it represents a wave travelling in the positive (for $x - vt$) or negative (for $x + vt$) x-direction with speed \( v \). $y$ could represent displacement, pressure, electric field, etc.
+\end{definition}
+
+\begin{definition}[Sinusoidal Wave]
+    A sinusoidal wave is a wave that can be described by a sine or cosine function. The general form of a sinusoidal wave travelling in the positive x-direction is:
+    \begin{equation}
+        y(x,t) = A \cos(kx \pm \omega t + \phi_0)
+    \end{equation}
+    where:
+    \begin{itemize}
+        \item \( A \) is the amplitude (maximum displacement)
+        \item \( k = \frac{2\pi}{\lambda} \) is the wave number, with \( \lambda \) being the wavelength
+        \item \( \omega = 2\pi f \) is the angular frequency, with \( f \) being the frequency
+        \item \( \phi _0 \) is the initial phase.
+    \end{itemize}
+    So we can write the wave function as a form of \eqref{eq:travelling_wave}:
+    \begin{equation}
+        y(x,t) = A \cos\left[k(x \pm vt) + \phi_0\right]
+    \end{equation}
+    The wave speed \( v \) is related to the angular frequency and wave number by:
+    \begin{equation}
+        v = \frac{\omega}{k} = f\lambda
+    \end{equation}
+\end{definition}
+
+\paragraph{Transverse and Longitudinal Waves} There is two (or three) main categories of waves:
+\begin{itemize}
+    \item \textbf{Transverse Waves}: In transverse waves, the oscillations are perpendicular to the direction of wave propagation. \\
+    (Examples) Light waves, water waves, and waves on a string.
+    \item \textbf{Longitudinal Waves}: In longitudinal waves, the oscillations are parallel to the direction of wave propagation. \\
+    (Examples) Sound waves and pressure waves in fluids.
+    \item \textbf{Both}: Some waves can exhibit both transverse and longitudinal characteristics. \\
+    (Examples) Rayleigh surface waves in seismology.
+\end{itemize}
+\begin{definition}[Velocity of a Fixed Particle]
+    For a fixed particle, we have:
+    \begin{equation}
+        v_y = \frac{\partial y}{\partial t} = A \omega \cos(kx \pm \omega t + \phi_0)
+    \end{equation}
+\end{definition}
+
+\begin{example}[Vibrating String]
+    Consider a string in the $xy$-plane, stretched along the x-axis with tension \( \tau \) and linear mass density \( \mu \).
+Now, consider at the end of the string at $x$ makes a small angle \( \theta \) with the x-axis.
+Then the components of the tension are:
+    $$
+    \begin{cases}
+        \tau_x &= \tau \cos(\theta) \approx \tau \\
+        \tau_y &= \tau \sin(\theta) \approx \tau \frac{\partial y}{\partial x}
+    \end{cases}
+    $$
+    Now, consider a small segment of the string between \( x \) and \( x + \delta x \). The vetical force on the other end is given, via linear approximation, by:
+    $$    
+        \tau_y(x + \delta x) \approx \tau_y(x) \left[ \frac{\partial y}{\partial x}_x + \frac{\partial^2 y}{\partial x^2} \delta x \right]
+    $$
+    So the net force is given by:
+    $$
+        F_y = \tau_y(x + \delta x) - \tau_y(x) = \tau \frac{\partial^2 y}{\partial x^2} \delta x
+    $$
+    Becuase $F=ma$, we have:
+    $$
+        \tau \frac{\partial^2 y}{\partial x^2} \delta x = dm \frac{\partial^2 y}{\partial t^2}
+    $$
+    where \( dm = \mu \delta x \) is the mass of the small segment. Thus, we have:
+    $$
+        \frac{\partial^2 y}{\partial t^2} - v^2 \frac{\partial^2 y}{\partial x^2} = 0
+    $$
+    where \( v = \sqrt{\frac{\tau}{\mu}} \) is the wave speed on the string. This is the one-dimensional wave equation.
+
+\end{example}
+
+\subsection{The Wave Equation}
 

### 2025-09-26 YEAR2/AER210_LectureNotes/main.tex
sept 26

@@
     $$        
-        J = \frac{1}{J^{-1}} = -\frac{1}{8xy}
+        J = \frac{1}{J^{-1}} = -\frac{1}{8xy} \implies I = \iint_R xy \, dA = \iint_S \frac{xy}{8xy} \, du \, dv = \iint_S \frac{1}{8} \, du \, dv = \frac{15}{8}
     $$
     This is consistent with the result we obtained in \ref{ex:change_of_variables}, and results in a much simpler integral to evaluate.
+\end{example}
+
+\begin{example}
+    Consider the integral:  
+    $$
+        \iint_R \exp{\left(\frac{x - y}{x + y}\right)} \, dA
+    $$
+    where $R$ is the trapezoid with vertices at $(1,0)$, $(2,0)$, $(0,-2)$ and $(0,-1)$.
+
+    Motivated by the integrabled, we can use the change of variables:
+    $$
+    \begin{cases}
+        u = x + y \\
+        v = x - y
+    \end{cases} 
+    $$
+    Also, we can define our bounds as follow:
+    \begin{itemize}
+        \item $y = x - 2 \implies u = 2$
+        \item $y = x - 1 \implies v = 1$
+        \item $x = 0 \implies u = -v$
+        \item $y = 0 \implies u = v$
+    \end{itemize}
+    We can find the Jacobian of this transformation by its inverse:
+    $$
+    J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
+        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
+        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
+    \end{vmatrix} = \begin{vmatrix}
+        1 & 1 \\
+        1 & -1
+    \end{vmatrix} = -2 \implies J = -\frac{1}{2}
+    $$
+    So, we can write the integral in terms of $u$ and $v$:
+    \begin{align*}
+        \iint_R \exp{\left(\frac{x - y}{x + y}\right)} \, dA &= \iint_S \exp{\left(\frac{v}{u}\right)} \cdot |J| \, du \, dv \\
+        &= \frac{1}{2} \int_1^2 \int_{-v}^{v} \exp{\left(\frac{v}{u}\right)} \, du \, dv \\
+        &= \frac{1}{2} \int_1^2 v(e - e^{-1}) \, dv = \frac{3}{4} (e - e^{-1})
+    \end{align*}
+\end{example}
+
+\begin{example}
+    \textit{Evaluate the integral}
+    $$
+        \iint_R (x^2 - y^2) \exp(xy) \, dA
+    $$
+    where $R$ is the region bounded by the hyperbolas $xy=1$ and $xy=4$, and the lines $y=x$ and $y=x+2$.
+
+    Motivated by the bounds, we can use the change of variables:
+    $$
+    \begin{cases}
+        u = xy \\
+        v = y - x
+    \end{cases}
+    $$
+    So, let $S$ be the corresponding region in the $uv$-plane. We can find the bounds of $S$ by plugging in the lines that bound $R$:   
+    \begin{itemize}
+        \item $xy = 1 \implies u = 1$
+        \item $xy = 4 \implies u = 4$
+        \item $y = x \implies v = 0$
+        \item $y = x + 2 \implies v = 2$
+    \end{itemize}
+    We find the Jacobian of this transformation by its inverse:
+    $$
+    J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
+        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
+        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
+    \end{vmatrix} = \begin{vmatrix}
+        y & x \\
+        -1 & 1
+    \end{vmatrix} = y + x \implies J = \frac{1}{y + x}
+    $$
+    So, we can express the integral in terms of $u$ and $v$:
+    \begin{align*}
+        \iint_R (x^2 - y^2) \exp(xy) \, dA &= \iint_S (x+y)(x-y)e^u \cdot |J| \, du \, dv \\
+        &= \int_0^2 \int_1^4 -v e^u \, du \, dv = \int_0^2 -v(e^4 - e) \, dv = 2(e - e^4)
+    \end{align*}
+\end{example}
+\begin{example}
+    \textit{Find the volume of the region bounded by the hyperbolic cylinders:}
+    $$ 
+    xy = 1, \quad xy = 3, \quad xz = 1, \quad xz = 3, \quad x = 36, \quad y = 25, \quad yz = 49
+    $$
+    Motivated by the bounds, we can use the change of variables:
+    $$
+    \begin{cases}
+        u = xy \\
+        v = xz \\
+        w = yz
+    \end{cases} 
+    $$
+    So, let $S$ be the corresponding region in the $uvw$-space. We can find the bounds of $S$ by plugging in the surfaces that bound the region:
+
+    \begin{align*}
+        1 &\leq u \leq 3 \\
+        1 &\leq v \leq 3 \\
+        25 &\leq w \leq 49
+    \end{align*}
+
 \end{example}

### 2025-09-26 YEAR2/ECE253_LectureNotes/main.tex
sept 26

@@
     \end{table}
+
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module d_latch(input logic D, CLK,
+                   output logic Q);
+        always_latch begin // this is new in SystemVerilog, 
+        // dont write else in always_latch 
+            if (CLK) // Transparent mode
+                Q = D;
+        end
+    endmodule
+    \end{verbatim}
 \end{definition}
@@
 
-\begin{definition}[D-Flip Flops]
+\begin{definition}[D-Flip Flops (Registors)]
     A D flip-flop is a type of flip-flop that captures the value of the D input at a specific portion of the clock cycle (rising edge) and holds that value until the next clock cycle.
@@
 While the second D-latch recognizes the CLK, and the first one recognizes $\overline{\text{CLK}}$.
+
+    The verilog implementation is as follows:
+    \begin{verbatim}
+    module d_ff(input logic D, CLK,
+                output logic Q);
+        always_ff @(posedge CLK) begin
+            Q <= D; // Non-blocking assignment
+        end
+    endmodule
+    \end{verbatim}
+    \verb|postedge| is a keyword used to create a positive edge-triggered FF.
+We use $<= $ instead of $= $ because we want to use non-blocking assignment.
+This is because in a sequential circuit, all the FFs are triggered at the same time, and we want to make sure that the value of Q is updated only after all the FFs have been evaluated.
+\end{definition}
+
+\begin{definition}[8-Bit Register]
+    An 8-bit register is a digital storage device that can hold 8 bits of binary data.
+It is typically implemented using a series of D flip-flops, where each flip-flop stores one bit of data.
+The register has a clock input (CLK) that controls when the data is loaded into the register, and an enable input (EN) that determines whether the register should load new data or retain its current value.
+When EN is high (1), the register loads the data present on the D input into the flip-flops on the rising edge of the clock.
+When EN is low (0), the register retains its current value regardless of changes on the D input.
+
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module register8(input logic [7:0] D,
+                     input logic CLK, EN,
+                     output logic [7:0] Q);
+        always_ff @(posedge CLK) begin
+            if (EN)
+                Q <= D; // Load new data
+            // else retain old data
+        end
+    endmodule
+    \end{verbatim}
+
+\end{definition}
+
+\paragraph{Resets} We have two types of rests to reset the FF to a known state:
+\begin{itemize}
+    \item \textbf{Synchronous reset:} Active only on the clock edge. If reset = 1 at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
+    \item \textbf{Asynchronous reset:} Active immediately when asserted, forcing the flip‑flop to the predefined state (usually 0) regardless of the clock; normal edge‑triggered operation resumes when deasserted.
+\end{itemize}
+\begin{definition}[Synchronous Reset D-FF]
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module d_ff_sync_reset(input logic D, CLK, RST,
+                           output logic Q);
+        always_ff @(posedge CLK) begin
+            if (RST==1'b0)
+                Q <= 1'b0; // Reset to 0
+            else
+                Q <= D; // Normal operation
+        end
+    endmodule
+    \end{verbatim}
+
 \end{definition}

### 2025-09-29 YEAR2/AER210_LectureNotes/main.tex
sept 29

@@
 
-\end{example}
+    We find the Jacobian of this transformation by its inverse:
+    $$
+    J^{-1} = \frac{\partial(u,v,w)}{\partial(x,y,z)} = \begin{vmatrix}
+        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} & \dfrac{\partial u}{\partial z} \\
+        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y} & \dfrac{\partial v}{\partial z} \\
+        \dfrac{\partial w}{\partial x} & \dfrac{\partial w}{\partial y} & \dfrac{\partial w}{\partial z}
+    \end{vmatrix} = \begin{vmatrix}
+        y & x & 0 \\
+        z & 0 & x \\
+        0 & z & y
+    \end{vmatrix} = xyz
+    $$
+    So, we can express the volume of the region in terms of $u$, $v$ and $w$:
+    \begin{align*}
+        V &= \iiint_R 1 \, dV = \iiint_S |J| \, du \, dv \, dw = \iiint_S \frac{1}{xyz} \, du \, dv \, dw \\
+        &= \int_{25}^{49} \int_1^3 \int_1^3 \frac{1}{\sqrt{\frac{uvw}{y^2}}} \, du \, dv \, dw = \int_{25}^{49} \int_1^3 \int_1^3 \frac{y}{\sqrt{uvw}} \, du \, dv \, dw \\
+        &= \int_{25}^{49} \int_1^3 \int_1^3 \frac{5}{\sqrt{uvw}} \, du \, dv \, dw = 5 \int_{25}^{49} \int_1^3 \left[ 2\sqrt{u} \right]_{u=1}^{u=3} \frac{1}{\sqrt{vw}} \, dv \, dw \\
+        &= 10(\sqrt{3} - 1) \int_{25}^{49} \left[ 2\sqrt{v} \right]_{v=1}^{v=3} \frac{1}{\sqrt{w}} \, dw = 20(\sqrt{3} - 1)(\sqrt{3} - 1) \int_{25}^{49} \frac{1}{\sqrt{w}} \, dw \\
+        &= 20(\sqrt{3} - 1)^2 \left[ 2\sqrt{w} \right]_{w=25}^{w=49} = 40(\sqrt{3} - 1)^2 (7 - 5) = 80(\sqrt{3} - 1)^2
+    \end{align*}
+
+\end{example}
+
+\section{Vector Calculus}
+\section{Line Integrals}
+\begin{definition}[Line Integral of a Scalar Field]
+    Consider a scalar field $f(x,y) $ and a smooth curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
+    $$
+        \int_C f(x,y) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*) \Delta s_i
+    $$
+    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.
+\end{definition}
+
+\begin{definition}[Parametrization of a Curve]
+    A smooth curve $C$ in the plane can be parametrized by:
+    \begin{equation}
+        \begin{cases}
+            x = g(t) \\
+            y = h(t)
+        \end{cases} \implies \vec{r(t)} = \langle g(t), h(t) \rangle, \quad a \leq t \leq b
+    \end{equation}
+    where $g$ and $h$ are continuous functions with continuous derivatives on $[a,b]$.
+
+    Under the following asummumption, we can solve the line integral using the parametrization:
+    \begin{itemize}
+        \item The function $f(x,y)$ is continuous on $C$.
+        \item The curve $C$ is smooth and can be parametrized by $\vec{r(t)}$.
+        \begin{itemize}
+            \item $r'(t)$ is continuous on $[a,b]$.
+            \item $r'(t) \neq 0$ for all $t \in [a,b]$. (Except possibly at the endpoints.) 
+        \end{itemize}
+    \end{itemize}
+\end{definition}
+
+\paragraph{Change of variable from $ds$} We can express the line integral in terms of the parameter $t$ as follows:
+\begin{equation}
+    \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
+\end{equation}
+where $||\vec{r'(t)}|| = \sqrt{(g'(t))^2 + (h'(t))^2}$ is the speed of the curve.
+
+\begin{definition}[Center of Mass via Line Integral]
+    Consider a wire in the plane with a density function $\rho(x,y)$ and a smooth curve $C$ parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$.
+Then, the mass of the wire is given by:
+    \begin{equation}
+        m = \int_C \rho(x,y) \, ds = \int_a^b \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
+    \end{equation}
+    And the center of mass $(\bar{x}, \bar{y})$ of the wire is given by:
+    \begin{align}
+        \bar{x} &= \frac{1}{m} \int_C x \rho(x,y) \, ds = \frac{1}{m} \int_a^b g(t) \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt \\
+        \bar{y} &= \frac{1}{m} \int_C y \rho(x,y) \, ds = \frac{1}{m} \int_a^b h(t) \rho(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
+    \end{align}
+\end{definition}
+
+\begin{example}[Center of Mass of a Semi-Circular Wire]
+    Let $a$ be the radius of a semi-circular wire. We want to find the center of mass of the wire. We can parametrize the semi-circular wire as follows:
+    $$
+    \begin{cases}
+        x = a \cos t \\
+        y = a \sin t
+    \end{cases} \quad 0 \leq t \leq \pi
+    $$
+    The density function is constant, $\rho(x,y) = \rho_0$. The mass of the wire is just the length of the wire times the density:
+    $$
+        m = \pi a \rho_0
+    $$
+    The $x$-coordinate of the center of mass is just 0 by symmetry. The $y$-coordinate of the center of mass is given by:
+    \begin{align*}
+        \bar{y} &= \frac{1}{m} \int_C y \rho(x,y) \, ds = \frac{1}{\pi a \rho_0} \int_0^\pi a \sin t \cdot \rho_0 \cdot a \, dt \\
+        &= \frac{a^2}{\pi a} \int_0^\pi \sin t \, dt = \frac{a}{\pi} \left[ -\cos t \right]_0^\pi = \frac{2a}{\pi}
+    \end{align*}
+    So, the center of mass of the semi-circular wire is located at:
+    $$
+        \left( \bar{x}, \bar{y} \right) = \left( 0, \frac{2a}{\pi} \right)
+    $$
+\end{example}
+
+\begin{definition}[Line Integral with Space Curve]
+    Consider a scalar field $f(x,y,z)$ and a smooth space curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
+    $$
+        \int_C f(x,y,z) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta s_i
+    $$
+    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*, z_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.
+
+    A smooth space curve $C$ can be parametrized by:
+    \begin{equation}
+        \begin{cases}
+            x = g(t) \\
+            y = h(t) \\
+            z = k(t)
+        \end{cases} \implies \vec{r(t)} = \langle g(t), h(t), k(t) \rangle, \quad a \leq t \leq b
+    \end{equation}
+    where $g$, $h$ and $k$ are continuous functions with continuous derivatives on $[a,b]$.
+
+    Under the following asummumption, we can solve the line integral using the parametrization:
+    \begin{itemize}
+        \item The function $f(x,y,z)$ is continuous on $C$.
+        \item The curve $C$ is smooth and can be parametrized by $\vec{r(t)}$.
+        \begin{itemize}
+            \item $r'(t)$ is continuous on $[a,b]$.
+            \item $r'(t) \neq 0$ for all $t \in [a,b]$. (Except possibly at the endpoints.) 
+        \end{itemize}
+    \end{itemize}
+
+    We can express the line integral in terms of the parameter $t$ as follows:
+    \begin{equation}
+        \int_C f(x,y,z) \, ds = \int_a^b f(g(t), h(t), k(t)) \, ||\vec{r'(t)}|| \, dt
+    \end{equation}
+    where $||\vec{r'(t)}|| = \sqrt{(g'(t))^2 + (h'(t))^2 + (k'(t))^2}$ is the speed of the curve.
+\end{definition}
+
+\begin{example}[Mass of a Spring]
+    Find the mass of a spring parametrized by:
+    $$
+    \begin{cases}
+        x = 2 \cos t \\
+        y = t \\
+        z = 2 \sin t
+    \end{cases} \quad 0 \leq t \leq 6\pi
+    $$
+    with a density function $\rho(x,y,z) = 2y$.
+
+    The mass of the spring is given by:
+    \begin{align*}
+        m &= \int_C \rho(x,y,z) \, ds = \int_0^{6\pi} 2t \cdot \sqrt{(-2\sin t)^2 + 1^2 + (2\cos t)^2} \, dt \\
+        &= \int_0^{6\pi} 2t \cdot \sqrt{4 + 1} \, dt = \sqrt{5} \int_0^{6\pi} 2t \, dt = \sqrt{5} \left[ t^2 \right]_0^{6\pi} = 36\pi^2 \sqrt{5}
+    \end{align*}    
+\end{example}
+
+\begin{definition}[Line Intergral with Piecewise Smooth Curve]
+    Consider a scalar field $f(x,y)$ and a piecewise smooth curve $C$ composed of $n$ smooth curves $C_1, C_2, \ldots, C_n$.
+Then we define the line integral of $f$ along $C$ as the sum of the line integrals along each smooth curve:
+    \begin{equation}
+        \int_C f(x,y) \, ds = \sum_{i=1}^n \int_{C_i} f(x,y) \, ds
+    \end{equation}
+\end{definition}
+
+\begin{theorem}[Directionality of Scalar Line Integrals] The line integral of a scalar field is independent of the direction of the curve.
+That is, if we reverse the direction of the curve, the value of the line integral remains the same:
+\begin{equation}
+    \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
+\end{equation}
+\end{theorem}
+\begin{proof}
+    Let $C$ be a smooth curve parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$. Then, the line integral of $f$ along $C$ is given by:
+    \begin{equation}
+        \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
+    \end{equation}
+    Now, consider the curve $-C$ obtained by reversing the direction of $C$.
+We can parametrize $-C$ by $\vec{r(-t)} = \langle g(-t), h(-t) \rangle$, $-b \leq t \leq -a$.
+The line integral of $f$ along $-C$ is given by:
+    \begin{align*}
+        \int_{-C} f(x,y) \, ds &= \int_{-b}^{-a} f(g(-t), h(-t)) \, ||\vec{r'(-t)}|| \, dt \\
+        &= \int_a^b f(g(u), h(u)) \, ||\vec{r'(u)}|| \, du \quad (u = -t, du = -dt) \\
+        &= \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
+    \end{align*}
+    Thus, we have shown that:
+    $$
+        \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
+    $$
+\end{proof}
 \chapter{Fluid Mechanics}

### 2025-09-29 YEAR2/CHE260_LectureNotes/main.tex
sept 29

@@
 \end{definition}
+
+\section{The Postulate of Classical Thermodynamics}
+We can now summarize thermal dynamics as a logical system.
+
+\paragraph{Fundamental Concepts} Below are the fundamental properties of thermodynamics:
+\begin{itemize}
+    \item \textbf{Volume} $V$: the space occupied by the system.
+    \item \textbf{Energy} $U$: the internal energy of the system.
+    \item \textbf{Entropy} $S$: the measure of disorder or randomness in the system.
+    \item \textbf{Mass} $m$: the amount of matter in the system.
+\end{itemize}
+
+\paragraph{Postulates} The postulates of classical thermodynamics are:
+\begin{itemize}
+    \item \textbf{State Postulate}: The state of a simple compressible system is completely specified by:
+    \begin{equation}
+        S = S(U, V, m)
+    \end{equation} 
+    \item \textbf{First Law of Thermodynamics}: The change in internal energy of a system is equal to the heat added to the system minus the work done by the system:
+    \begin{equation}
+        \Delta U = Q - W
+    \end{equation}
+    \item \textbf{Second Law of Thermodynamics}: The entropy of an isolated system never decreases; it either increases or remains constant for a reversible process:
+    \begin{equation}
+        \Delta S \geq 0
+    \end{equation}
+    \item \textbf{Third Law of Thermodynamics}: The entropy of a perfect crystal at absolute zero temperature is zero:
+    \begin{equation}
+        S = 0
+    \end{equation}
+    when:
+    \begin{equation}
+        T = 0\, \text{K} \quad \text{or} \quad \frac{\partial U}{\partial S} = 0
+    \end{equation}
+\end{itemize}
+
 \chapter{Heat Transfer}

### 2025-09-29 YEAR2/ECE253_LectureNotes/main.tex
sept 29

@@
 \begin{itemize}
-    \item \textbf{Synchronous reset:} Active only on the clock edge. If reset = 1 at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
+    \item \textbf{Synchronous reset:} Active only on the clock edge.
+If reset = 1 (or resetn = 0) at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
     \item \textbf{Asynchronous reset:} Active immediately when asserted, forcing the flip‑flop to the predefined state (usually 0) regardless of the clock; normal edge‑triggered operation resumes when deasserted.
@@
     \begin{verbatim}
-    module d_ff_sync_reset(input logic D, CLK, RST,
+    module d_ff_sync_reset(input logic D, CLK, RSTn,
                            output logic Q);
         always_ff @(posedge CLK) begin
-            if (RST==1'b0)
+            if (RSTn == 1'b0)
                 Q <= 1'b0; // Reset to 0
@@
 \end{definition}
+
+\begin{definition}[Asynchronous Reset D-FF]
+    The Verilog implementation is as follows:
+    \begin{verbatim}
+    module d_ff_async_reset(input logic D, CLK, RSTn,
+                            output logic Q);
+        always_ff @(posedge CLK or negedge RSTn) begin
+            if (RSTn == 1'b0)
+                Q <= 1'b0; // Reset to 0
+            else
+                Q <= D; // Normal operation
+        end
+    endmodule
+    \end{verbatim}
+\end{definition}
+
+\paragraph{Enable Inputs} We can add an enable input to a D-FF by adding a mux before the D input.
+It determines whether data is loaded on the clock edge or the old data is retained.
+If the enable input is high (1), the data on the D input is loaded into the flip-flop on the rising edge of the clock.
+If the enable input is low (0), there would be a feedback signal going into the mux, so the old data is retained.
+\begin{definition}[Counters]
+    To count binary numbers, we have the follow truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|}
+            \hline
+            CLK Cycle & Binary Count (Q2 Q1 Q0) \\
+            \hline
+            0 & 000 \\
+            1 & 001 \\
+            2 & 010 \\
+            3 & 011 \\
+            4 & 100 \\
+            5 & 101 \\
+            6 & 110 \\
+            7 & 111 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for a 3-bit Counter}
+        \label{tab:3_bit_counter_truth_table}
+    \end{table}
+    From the truth table, we can derive the following rules:
+    \begin{itemize}
+        \item $Q_0$ toggles every clock cycle.
+        $$
+        Q_0(t+1) = Q_0(t) \oplus 1
+        $$
+        \item $Q_1$ toggles when $Q_0$ is 1.
+        $$
+        Q_1(t+1) = Q_1(t) \oplus Q_0(t)
+        $$
+        \item $Q_2$ toggles when $Q_1$ and $Q_0$ are both 1.
+        $$
+        Q_2(t+1) = Q_2(t) \oplus (Q_1(t) Q_0(t))
+        $$
+    \end{itemize}
+    We call this a TQ flip-flop, where T stands for toggle.
+It is done by connecting the Q output back, to a new XOR gate.
+The input of that XOR gate is Q and T.
+And the XOR gate output is connected to the D input of a D-FF.
+The Verilog implementation is as follows:
+    \begin{verbatim}
+    module counter3(input logic CLK, RSTn,
+                    output logic [3:0] Q);
+        always_ff @(posedge CLK or negedge RSTn) begin
+            if (RSTn == 1'b0)
+                Q <= 4'b0000;
+            else begin
+                Q[0] <= Q[0] ^ 1'b1; // Toggle Q0
+                Q[1] <= Q[1] ^ Q[0]; // Toggle Q1
+                Q[2] <= Q[2] ^ (Q[1] & Q[0]); // Toggle Q2
+                Q[3] <= Q[3] ^ (Q[2] & Q[1] & Q[0]); // Toggle Q3
+            end
+        end
+    endmodule
+    \end{verbatim}
+\end{definition}
 \section{Finite State Machines (FSM)}
-% FSM modeling, state diagrams, and Verilog implementation (to be expanded)
 

### 2025-09-29 YEAR2/MAT292_LectureNotes/main.tex
sept 29

@@
 
+\subsection{Wronskian and Fundamental Matrix}
+\begin{definition}[Functionals]
+    A functional is a mapping from a vector space of functions to the real numbers.
+For example, consider the space of continuous functions on an interval \( [a, b] \), denoted by \( C[a, b] \).
+A functional \( J: C[a, b] \to \mathbb{R} \) could be defined as:
+    $$
+    J[f] = \int_a^b f(x) \, dx
+    $$
+    for any function \( f \in C[a, b] \).
+\end{definition}
+
+\paragraph{Motivation} For an $n$th-order differential equation we need $n$ linearly independent solutions to form the general solution.
+These solutions constitute a basis of the solution space, so we must check whether a given set of solutions is linearly independent.
+
+\begin{definition}[Wronskian]
+    Consider \( n \) functions \( x_1(t), x_2(t), \ldots, x_n(t) \) for which \( x: \mathbb{R} \to \mathbb{R}^n \) that are solutions of a $n$th-order linear differential equation on an interval \( I \).
+The Wronskian \( W(t) \) is defined as the determinant of the matrix whose columns are formed by the functions and their derivatives up to order \( n-1 \):
+    \begin{equation}
+    W(t) = \det\begin{bmatrix}
+    x_1(t) & x_2(t) & \cdots & x_n(t) \\
+    x_1'(t) & x_2'(t) & \cdots & x_n'(t) \\
+    \vdots & \vdots & \ddots & \vdots \\
+    x_1^{(n-1)}(t) & x_2^{(n-1)}(t) & \cdots & x_n^{(n-1)}(t)
+    \end{bmatrix} 
+    \end{equation}
+    Iff there exist \( t \) in interval \( I \) such that \( W(t) \neq 0 \), then the functions \( x_1(t), x_2(t), \ldots, x_n(t) \) are linearly independent on \( I \).
+
+\end{definition}
+
+\begin{example}
+    Let $f(x) = \cos (mx)$ and $g(x) = \sin (mx)$ for some constant $m \neq 0$. We compute the Wronskian:
+    \begin{align*}
+    W(f, g)(x) &= \det\begin{bmatrix}
+    \cos(mx) & \sin(mx) \\
+    -m\sin(mx) & m\cos(mx)
+    \end{bmatrix} \\
+    &= m\cos^2(mx) + m\sin^2(mx) \\
+    &= m(\cos^2(mx) + \sin^2(mx)) \\
+    &= m \neq 0
+    \end{align*}
+    Since \( W(f, g)(x) \neq 0 \) for all \( x \), the functions \( \cos(mx) \) and \( \sin(mx) \) are linearly independent on \( \mathbb{R} \).
+\end{example}
+\paragraph{Trading order for dimension} Consider the $n$th-order ODE:
+$$a_n(t)x^{(n)} + a_{n-1}(t)x^{(n-1)} + \ldots + a_1(t)x' + a_0(t)x = g(t)$$
+where \( a_n(t) \neq 0 \) on an interval \( I \). We can convert this into a system of first-order ODEs by defining:
+$$\begin{cases}
+x_1 = x \\
+x_2 = x' \\
+x_3 = x'' \\
+\vdots \\
+x_n = x^{(n-1)}
+\end{cases}$$
+Then, we have solution $x: \mathbb{R} \to \mathbb{R}^n$ given by:
+$$
+x = v_1 e^{\lambda_1 t} + v_2 e^{\lambda_2 t} + \ldots + v_n e^{\lambda_n t}
+$$
+where \( v_i \) are the eigenvectors and \( \lambda_i \) are the eigenvalues of the corresponding matrix \( A \). Then, we can define the Wronskian for this system as:
+\begin{equation}
+W(v_1e^{\lambda_1 t}, v_2 e^{\lambda_2 t}, \ldots, v_n e^{\lambda_n t}) = 
+\begin{vmatrix}
+v_1 e^{\lambda_1 t} & v_2 e^{\lambda_2 t} & \ldots & v_n e^{\lambda_n t} 
+\end{vmatrix} = e^{(\lambda_1 + \lambda_2 + \ldots + \lambda_n)t} \begin{vmatrix}
+v_1 & v_2 & \ldots & v_n 
+\end{vmatrix}
+\end{equation}
+Which the first term is never zero, so the Wronskian is non-zero iff the eigenvectors are linearly independent.
+The basis $\{v_1 e^{\lambda_1 t}, v_2 e^{\lambda_2 t}, \ldots, v_n e^{\lambda_n t}\}$ is a fundamental called the fundamental set of solutions.
+
+\paragraph{General Planer Systems} Consider the system of ODEs:
+$$\textbf{x}' = A\textbf{x}$$
+where \( A \in \mathbb{R}^{2 \times 2} \).
+
 \end{document}

### 2025-09-29 YEAR2/PHY293_LectureNotes/main.tex
sept 29

@@
     $$
-    Becuase $F=ma$, we have:
-    $$
-        \tau \frac{\partial^2 y}{\partial x^2} \delta x = dm \frac{\partial^2 y}{\partial t^2}
-    $$
-    where \( dm = \mu \delta x \) is the mass of the small segment. Thus, we have:
+    Because \( F=ma \), we have:
+    $$
+        \tau \frac{\partial^2 y}{\partial x^2} \delta x = \text{d}m \frac{\partial^2 y}{\partial t^2}
+    $$
+    where \( \text{d}m = \mu \delta x \) is the mass of the small segment. Thus, we have:
     $$
@@
 \end{example}
-
-\subsection{The Wave Equation}
-
 \begin{definition}[The Wave Equation]
-    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water). In one dimension, it is:
+    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water) through a medium. In one dimension, it is given by:
     \begin{equation}
@@
     where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
-\end{definition}
-
+
+    Or in general:
+    \begin{equation}
+        \frac{\partial^2 u}{\partial t^2} - c^2 \nabla^2 u = 0
+    \end{equation}
+    where \( \nabla^2 \) is the Laplacian operator.
+\end{definition}
+
+\begin{definition}[Energy of a String]
+    The kinetic energy of a string segment of mass \( \text{d}m = \mu \text{d}x \) is given by:
+    \begin{equation}
+        \text{d}K = \frac{1}{2} \text{d}m v_y^2 = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2
+    \end{equation}
+    The potential energy stored in the string due to its tension is given by:
+    \begin{equation}
+        \text{d}U = \frac{1}{2} \tau (\delta s - \delta x) 
+    \end{equation}
+    where \( \delta s \) is the actual length of the string segment and \( \delta x \) is the horizontal length. For small displacements, we can approximate:
+    $$
+        \delta s \approx \delta x \left(1 + \frac{1}{2}\left(\frac{\partial y}{\partial x}\right)^2\right)
+    $$
+    Thus, the potential energy becomes:
+    \begin{equation}
+        \text{d}U \approx \frac{1}{2} \tau \delta x \left(\frac{\partial y}{\partial x}\right)^2
+    \end{equation}
+    The total energy \( E \) of the string segment is the sum of its kinetic and potential energies:
+    \begin{equation}
+        \text{d}E = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2 + \frac{1}{2} \tau \text{d}x \left(\frac{\partial y}{\partial x}\right)^2
+    \end{equation}
+    For a sinusoidal wave \( y(x,t) = A \cos(kx - \omega t) \), the change in energy is:
+    \begin{equation}
+        \text{d}E = \frac{1}{2} \mu A^2 \omega^2 \sin^2(kx - \omega t) \text{d}x + \frac{1}{2} \tau A^2 k^2 \sin^2(kx - \omega t) \text{d}x
+    \end{equation}
+    Integrating over one wavelength \( \lambda \), we find the total energy in one wavelength:
+    \begin{equation}
+        E = \int_0^\lambda \text{d}E = \frac{1}{4} \mu A^2 \omega^2 \lambda + \frac{1}{4} \tau A^2 k^2 \lambda
+    \end{equation}
+
+\end{definition}
+
+\begin{definition}[Power Transmitted by a String]
+   The power carried by a transverse wave could be calculated from $P=Fv$. The transverse force on the string is given by:
+    \begin{equation}
+        P(x,t) = A^2 \mu v \omega^2 \sin^2(kx - \omega t + \phi_0)
+    \end{equation}
+    The average power is then given by:
+    \begin{equation}
+        \bar{P} = \frac{1}{2} A^2 \mu v \omega^2 = \frac{1}{2} \mu A^2 \omega^2 \frac{\lambda}{T}
+    \end{equation}
+    And the maximum power is given by:
+    \begin{equation}
+        P_{max} = A^2 \mu v \omega^2
+    \end{equation}
+\end{definition}
+
+\paragraph{Speed of Sound} The speed of sound waves through various meida is as follows:
+\begin{itemize}
+    \item In ideal gas, it is given by:
+    \begin{equation}
+        v = \sqrt{\frac{\gamma RT}{m}} = \sqrt{\frac{\gamma k_B T}{m}}
+    \end{equation}
+    where \( \gamma \) is the adiabatic index, \( R \) is the universal gas constant, \( k_B \) is the Boltzmann constant, \( T \) is the absolute temperature, and \( m \) is the molar mass of the gas.
+$\gamma$ varies based on the type of gas:
+    \begin{itemize}
+        \item Monatomic gases (e.g., helium, neon): \( \gamma = \frac{5}{3} \)
+        \item Diatomic gases (e.g., nitrogen, oxygen): \( \gamma = \frac{7}{5} \)
+        \item Polyatomic gases (e.g., carbon dioxide, methane): \( \gamma \) is typically between 1.3 and 1.4
+    \end{itemize}
+    \item In a liquid, it is given by:
+    \begin{equation}
+        v = \sqrt{\frac{B_{\text{adiabatic}}}{m}} = \sqrt{\frac{B}{\rho}}
+    \end{equation}
+    where \(B_{\text{adiabatic}}\) is the adiabatic bulk modulus, \( B \) is the bulk modulus, \( \rho \) is the density of the liquid and \( m \) is the molar mass of the liquid.
+    \item In a solid, it is given by:
+    \begin{equation}
+        v = \sqrt{\frac{E}{\rho}} 
+    \end{equation}
+    where \( E \) is the Young's modulus and \( \rho \) is the density of the solid. It could also be given by:
+    $$
+        v = \sqrt{\frac{B}{\rho}}
+    $$
+    where \( B \) is the bulk modulus.
+\end{itemize}
+
+\begin{definition}[Impedance]
+    Similar to electrical impedance, we also have acoustic impedance and mechanical, they measure how much a structure recisits motion when subjected to a harmonic force. They are defined as follows:
+    \begin{itemize}
+        \item Electical impedance \( Z_e \):
+        \begin{equation}
+            Z_e = \sqrt{R^2 + \left(\omega L - \frac{1}{\omega C}\right)^2}
+        \end{equation}
+        \item Acoustic impedance \( Z_a \), it is the ratio of acoustic pressure to particle velocity:
+        \begin{equation}
+            Z_a = \frac{p}{v} = \rho v = \sqrt{B\rho}
+        \end{equation}
+        \item Mechanical impedance \( Z_m \):
+        \begin{equation}
+            Z_m = v\mu = \sqrt{\tau\mu}
+        \end{equation}
+    \end{itemize}
+\end{definition}
+
+\paragraph{Wave boundry Conditions} When a wave encounters a boundary between two different media, part of the wave is reflected back into the original medium, and part is transmitted into the new medium.
+This is constrined by the following boundary conditions:
+\begin{itemize}
+    \item The displacement of the wave must be continuous across the boundary:
+    \begin{equation}
+        y_1(x_b, t) = y_2(x_b, t) = y_3(x_b, t)
+    \end{equation}
+    \item The derivative of the displacement with respect to position must also be continuous across the boundary:
+    \begin{equation}
+        \frac{\partial y_1}{\partial x}\bigg|_{x_b} = \frac{\partial y_2}{\partial x}\bigg|_{x_b} = \frac{\partial y_3}{\partial x}\bigg|_{x_b}
+    \end{equation}
+\end{itemize}
+and then from these conditions, we can derive the amplitudes of the reflected and transmitted waves:
+\begin{equation}
+    A_i = A_r + A_t
+\end{equation}
+and 
+\begin{equation}
+    A_i Z_1 = A_r Z_1 + A_t Z_2
+\end{equation}
+From these two equations, we can solve for the reflection and transmission coefficients:
+\begin{equation}
+    R = \frac{A_r}{A_i} = \frac{Z_2 - Z_1}{Z_2 + Z_1}
+\end{equation}
+and
+\begin{equation}
+    T = \frac{A_t}{A_i} = \frac{2Z_2}{Z_2 + Z_1}
+\end{equation}
 \chapter{Modern Physics}

### 2025-10-01 YEAR2/MAT292_LectureNotes/main.tex
30 sept

@@
 where \( A \in \mathbb{R}^{2 \times 2} \).
+We define the fundamental matrix \( \Phi(t) \) as:
+\begin{equation}
+\Phi(t) = \begin{bmatrix}
+x_1(t) & x_2(t) \\ 
+y_1(t) & y_2(t)
+\end{bmatrix}
+\end{equation}
+where \( (x_1(t), y_1(t)) \) and \( (x_2(t), y_2(t)) \) are two linearly independent solutions of the system. The Wronskian of the system is given by:
+\begin{equation}
+W(t) = \det(\Phi(t)) = x_1(t)y_2(t) - x_2(t)y_1(t)
+\end{equation}
+The Wronskian satisfies the differential equation:
+\begin{equation}
+W'(t) = \text{tr}(A) W(t)
+\end{equation}
+with the solution:
+\begin{equation}
+W(t) = W(0) e^{\text{tr}(A)t}
+\end{equation}
+where \( W(0) \) is the Wronskian at \( t = 0 \).
+
+\subsection{Stability of Nonlinear Systems in Local Linear Approximation}
+\begin{definition}[Change of Variables]
+    The change of variable in a higher dimensions, the change of variables related by the Jacobian matrix.
+WLOG, consider change of variables $F: \mathbb{R}^2 \to \mathbb{R}^2$ and system $x' = Ax$.
+Let:
+    $$
+    F(x,y) = \begin{bmatrix}f_1(x,y) \\ f_2(x,y)
+    \end{bmatrix}
+    $$
+
+    Then, the Jacobian matrix \( J_F \) is given by:
+    \begin{equation}
+    J_F = \begin{bmatrix}
+    \frac{\partial f_1}{\partial x} & \frac{\partial f_1}{\partial y} \\
+    \frac{\partial f_2}{\partial x} & \frac{\partial f_2}{\partial y}
+    \end{bmatrix}
+    \end{equation}
+    The change of variables transforms the system \( x' = Ax \) into a new system \( u' = Bu \) where:
+    \begin{equation}
+    B = J_F A J_F^{-1}
+    \end{equation}
+    provided that \( J_F \) is invertible.
+\end{definition}
+
+\begin{shaded}
+\subsubsection*{A reference for simple linear stability analysis.}
+
+\paragraph{Linear Stablity Analysis} Consider the simple linear one-dimensional ODE:
+$$y' = f(y)$$
+where \( f(y) \) is a differentiable function.
+Let \( y^* \) be a fixed point, i.e., \( f(y^*) = 0 \).
+We can perform a linear stability analysis by considering a small perturbation \( \eta(t) = y(t) - y^* \).
+Then, we have:
+\begin{align*}
+\eta' &= y' - (y^*)' \\
+&= f(y) - 0 \\
+&= f(y^* + \eta) \\
+&\approx f(y^*) + f'(y^*)\eta \\
+&= f'(y^*)\eta = A\eta
+\end{align*}
+So, we have the linearized system:
+$$\eta' = A\eta$$
+where \( A = f'(y^*) \). The stability of the fixed point \( y^* \) is determined by the sign of \( A \).
+\end{shaded}
+
+\begin{definition}[Linear Stability Analysis in Higher Dimensions]
+Consider the two-dimensional system:
+$$\begin{cases}
+x' = f(x, y) \\
+y' = g(x, y)
+\end{cases}$$
+where \( f(x, y) \) and \( g(x, y) \) are differentiable functions. Let \( (x^*, y^*) \) be a fixed point. We let perturbations \( u(t) = x(t) - x^* \) and \( v(t) = y(t) - y^* \). Then, we have:
+$$
+\begin{cases}
+    u' &= x' = \cancel{f(x,y)} + \frac{\partial f}{\partial x}(x^*, y^*)u + \frac{\partial f}{\partial y}(x^*, y^*)v + \cancel{O(u^2, v^2, uv)} \\
+    v' &= y' = \cancel{g(x,y)} + \frac{\partial g}{\partial x}(x^*, y^*)u + \frac{\partial g}{\partial y}(x^*, y^*)v + \cancel{O(u^2, v^2, uv)}
+\end{cases}
+$$
+So we can write the linearized system as:
+$$\begin{bmatrix}
+u' \\ v'
+\end{bmatrix} = J \begin{bmatrix}
+u \\ v
+\end{bmatrix}$$
+where \( J \) is the Jacobian matrix evaluated at the fixed point \( (x^*, y^*) \).
+This describes the local behavior of the system near the fixed point.
+Then the stability and behavior is determined by the determinant and trace of the Jacobian matrix \( J \), and hence the eigenvalues and eigenvectors of \( J \).
+\end{definition}
+
+\begin{definition}
+    Consider the nonlinear pendulum equation:
+    $$\theta'' + \frac{g}{L}\sin(\theta) = 0$$
+    where \( \theta \) is the angle of the pendulum, \( g \) is the acceleration due to gravity, and \( L \) is the length of the pendulum.
+We can rewrite this as a system of first-order ODEs by defining:
+    $$\begin{cases}
+    x_1 = \theta \\
+    x_2 = \theta'
+    \end{cases}$$
+    Then, we have:
+    $$\begin{cases}
+    x_1' = x_2 \\
+    x_2' = -\frac{g}{L}\sin(x_1)    
+    \end{cases}$$
+    Now, consider the Jacaobian matrix:
+    $$J = \begin{bmatrix}
+    0 & 1 \\
+    -\frac{g}{L}\cos(x_1) & 0
+    \end{bmatrix}$$
+
+    Consider the fixed point $(n \pi, 0)$ for \( n \in \mathbb{Z} \). We have:
+    $$J = \begin{bmatrix}
+    0 & 1 \\
+    -\frac{g}{L}\cos(n \pi) & 0
+    \end{bmatrix} = \begin{bmatrix}
+    0 & 1 \\
+    -\frac{g}{L}(-1)^n & 0
+    \end{bmatrix}$$
+    So the derminant is:
+    $$\det(J) = -\frac{g}{L}(-1)^n$$
+    and the trace is:
+    $$\text{tr}(J) = 0$$
+    When $n$ is even, we have a positive determinant, so the fixed point is a center. When $n$ is odd, we have a negative determinant, so the fixed point is a saddle node.
+\end{definition}
 

### 2025-10-01 YEAR2/PHY293_LectureNotes/main.tex
30 sept

@@
 \end{equation}
+
+\paragraph{Special Cases} There are two special cases:
+\begin{itemize}
+    \item \textbf{Fixed End}: If the wave encounters a boundary where the medium is fixed (e.g., a string tied to a wall), the reflected wave undergoes a phase change of \( \pi \) (inversion).
+The reflection coefficient is \( R = -1 \) and the transmission coefficient is \( T = 0 \).
+    \item \textbf{Free End}: If the wave encounters a boundary where the medium is free to move (e.g., a string attached to a ring that can slide along a rod), the reflected wave does not undergo a phase change.
+The reflection coefficient is \( R = 1 \) and the transmission coefficient is \( T = 2 \).
+\end{itemize}
+
+\begin{example}[Transmission of 64\% Energy]
+Suppose a wave transmits \(64\%\) of its incident energy across a boundary. Then:
+\begin{equation}
+    T_e = 0.64, 
+    \quad R_e = 1 - T_e = 0.36
+\end{equation}
+so that the amplitude reflection coefficient is
+\begin{equation}
+    R = +\sqrt{R_e} = 0.6 
+\end{equation}
+\textbf{Important} the positive sign indicates the reflected wave is \emph{in phase} with the incident wave.
+
+From the boundary condition in Eq.~\eqref{eq:amp-continuity},
+\begin{equation}
+    T = 1 + R = 1 + 0.6 = 1.6
+\end{equation}
+Thus, the amplitude transmission coefficient is \(T = 1.6\), even though only \(64\%\) of the energy is transmitted.
+This apparent paradox arises because the transmitted energy depends on both the amplitude squared and the ratio of impedances \(\tfrac{Z_2}{Z_1}\).
+\end{example}
+\subsection{Superposition and Standing Waves}
+\begin{definition}[Modulation of two waves]
+    Consider two monochromatic waves:
+    $$
+    \begin{cases}
+        y_1(x,t) &= A \cos(k_1 x - \omega_1 t) \\
+        y_2(x,t) &= A \cos(k_2 x - \omega_2 t)
+    \end{cases}
+    $$
+    Then, the resultant wave $y_1 + y_2$ can be expressed as:
+    \begin{equation}
+        y(x,t) = 2A \cos\left(\frac{(k_1 - k_2)x - (\omega_1 - \omega_2)t}{2}\right) \cos\left(\frac{(k_1 + k_2)x - (\omega_1 + \omega_2)t}{2}\right)
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Coherence]
+    Two waves are said to be coherent if they are monochromatic and maintain a constant phase relationship over time.
+This means that the difference in phase between the two waves does not change as they propagate:
+    \begin{equation}
+        \Delta \phi = \phi_1 - \phi_2 = \text{constant}
+    \end{equation}
+    Coherent waves can produce stable interference patterns, such as constructive and destructive interference.
+\end{definition}
+
+\begin{definition}[Phase and Phase Difference]
+    Phase of the wave is an argument $kx \pm \omega t + \phi_0$ of the wave function. The phase difference between two waves at a given point in space and time is given by:
+    \begin{equation}
+        \Delta \phi = (k_1 x - \omega_1 t + \phi_1) - (k_2 x - \omega_2 t + \phi_2)
+    \end{equation}
+    where \( k_1, k_2 \) are the wave numbers, \( \omega_1, \omega_2 \) are the angular frequencies, and \( \phi_1, \phi_2 \) are the initial phases of the two waves.
+
+    If we consider two waves of the same frequency and wave number, the phase difference simplifies to:
+    \begin{equation}
+        \Delta \phi = \Delta\phi + k\Delta x + \omega \Delta t
+    \end{equation}
+
+\end{definition}
+\begin{definition}[Waves in Two and Three Dimensions]
+    In two or three dimensions, waves can propagate in various directions. The general form of a wave in two dimensions is:
+    \begin{equation}
+        z(x,y,t) = A \cos(k_x x + k_y y - \omega t + \phi_0)
+    \end{equation}
+    where \( k_x \) and \( k_y \) are the wave numbers in the \( x \) and \( y \) directions, respectively, and \( \phi_0 \) is the initial phase. We also define $v$ and $k$ as:
+    \begin{equation}
+        v = \frac{\omega}{k}, \quad k = \sqrt{k_x^2 + k_y^2}
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Spherical Waves]
+    A spherical wave is a wave that propagates outward in all directions from a point source. The general form of a spherical wave is:
+    \begin{equation}
+        y(r,t) = \frac{A}{r} \cos(kr - \omega t + \phi_0)
+    \end{equation}
+    where \( r = \sqrt{x^2 + y^2 + z^2} \) is the radial distance from the source, \( A \) is the amplitude at a reference distance, \( k \) is the wave number, \( \omega \) is the angular frequency, and \( \phi_0 \) is the initial phase.
+The amplitude decreases with distance as \( 1/r \) due to the spreading of the wavefront.
+
+\end{definition}
+
+\begin{definition}[Interference]
+    Consider two coherent in phase sources $S_1$ and $S_2$ separated by a distance \( d \). The path difference \( \Delta r \) between the two waves at a point \( P \) is given by:
+    \begin{equation}
+        \Delta r = r_2 - r_1
+    \end{equation}
+    where \( r_1 \) and \( r_2 \) are the distances from the sources \( S_1 \) and \( S_2 \) to the point \( P \), respectively.
+
+    We have the following two cases for constructive and destructive interference:
+    \begin{itemize}
+        \item \textbf{Constructive Interference}: Occurs when the path difference is an integer multiple of the wavelength:
+        \begin{equation}
+            \Delta r = m \lambda, \quad m = 0, 1, 2, \ldots
+        \end{equation}
+        \item \textbf{Destructive Interference}: Occurs when the path difference is an odd multiple of half the wavelength:
+        \begin{equation}
+            \Delta r = \left(m + \frac{1}{2}\right) \lambda, \quad m = 0, 1, 2, \ldots
+        \end{equation}
+    \end{itemize}
+\end{definition}
+
 \chapter{Modern Physics}

### 2025-10-08 YEAR2/AER210_LectureNotes/main.tex
oct 7

@@
 \subsection{Taylor Series for Multivariable Functions}
+
+\begin{keybox}
+	\textbf{Key Equations:}
+\begin{itemize}
+    \item Multivariable Taylor (about $(x_0,y_0)$): $\displaystyle f(x,y) = f_0 + \sum_i f_i\Delta x_i + \tfrac12\sum_{i,j} f_{ij}\Delta x_i\Delta x_j + \cdots$
+    \item 2D form: $\displaystyle f(x_0{+}\Delta x,y_0{+}\Delta y) = f_0 + f_x\Delta x + f_y\Delta y + \tfrac12(f_{xx}\Delta x^2 + 2f_{xy}\Delta x\Delta y + f_{yy}\Delta y^2)+\cdots$
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Use a line path $(x,y)=(x_0,y_0)+t(\Delta x,\Delta y)$ to reduce to 1D.
+    \item Hessian controls the quadratic term and local shape (min/max/saddle with tests later).
+\end{itemize}
+\end{keybox}
 
@@
 
+\begin{keybox}
+	\textbf{Key Equations:}
+\begin{itemize}
+    \item Rectangle $R=[a,b]\times[c,d]$: $\displaystyle \iint_R f\,dA = \int_a^b\!\int_c^d f(x,y)\,dy\,dx = \int_c^d\!\int_a^b f(x,y)\,dx\,dy$
+    \item Separable $f(x,y)=g(x)h(y)$: product of one-dimensional integrals.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Swap order freely on rectangles; use Fubini/Tonelli conditions.
+    \item Check integrability and symmetry to simplify.
+\end{itemize}
+\end{keybox}
+
  By the point of seeing this note, you should be familiar with the simple case of rectangular, simple cases are provided as examples:
@@
 \subsection{Double Integrals in General Regions}
+
+\begin{keybox}
+	\textbf{Key Ideas:}
+\begin{itemize}
+    \item Type I: $a\le x\le b$, $g_1(x)\le y\le g_2(x)$, integrate $dy$ then $dx$.
+    \item Type II: $c\le y\le d$, $h_1(y)\le x\le h_2(y)$, integrate $dx$ then $dy$.
+\end{itemize}
+	\textbf{Tips/Pitfalls:}
+\begin{itemize}
+    \item Draw the region; switching order often simplifies bounds.
+    \item Split region if boundaries change formula across sub-intervals.
+\end{itemize}
+\end{keybox}
 \paragraph{Types of Regions} When the region $R$ is not rectangular, we can still compute the double integral by expressing the region in terms of inequalities.
@@
 
+\begin{keybox}
+	\textbf{Key Equations/Concepts:}
+\begin{itemize}
+    \item Piecewise descriptions: break $R$ into Type I/II parts.
+    \item Use symmetry and monotonic boundaries to choose efficient order.
+\end{itemize}
+	\textbf{Tip:} Convert to polar if circular arcs or $x^2{+}y^2$ appear.
+\end{keybox}
+
 \begin{theorem}[Change of Variable to Polar Coordinates]
@@
     Consider a region $R$ in the $xy$-plane that can be described in polar coordinates by the inequalities:
-    $$
+    \begin{equation}
         \alpha \le \theta \le \beta, \quad g_1(\theta) \le r \le g_2(\theta)
-    $$
+    \end{equation}
     where $g_1(\theta)$ and $g_2(\theta)$ are continuous functions on the interval $[\alpha, \beta]$.
@@
 \subsection{Applications of Double Integrals}
+
+\begin{keybox}
+	\textbf{Key Formulas:}
+\begin{itemize}
+    \item Mass: $M=\iint_R \rho\,dA$; Center of mass: $\bar x=\frac{1}{M}\iint_R x\rho\,dA$, $\bar y=\frac{1}{M}\iint_R y\rho\,dA$.
+    \item Moments of inertia: $I_x=\iint_R y^2\rho\,dA$, $I_y=\iint_R x^2\rho\,dA$, $I_O=I_x+I_y$ (planar).
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Use symmetry to zero cross-terms and simplify integrals.
+    \item Choose coordinates aligned with geometry.
+\end{itemize}
+\end{keybox}
 \begin{shaded}
@@
 \subsection{Surface Area}
+
+\begin{keybox}
+	\textbf{Key Formula:}
+\begin{itemize}
+    \item Graph $z=f(x,y)$ over $R$: $\displaystyle S=\iint_R \sqrt{1+f_x^2+f_y^2}\,dA$.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item For parametric $\mathbf r(u,v)$ use $\iint\!\lVert \mathbf r_u\times\mathbf r_v\rVert\,du\,dv$.
+    \item Approximate with linearization when small slopes.
+\end{itemize}
+\end{keybox}
 \begin{theorem}[Surface Area]
@@
 \subsection{Triple Integrals in Rectangular Coordinates}
+
+\begin{keybox}
+	\textbf{Key Equations:}
+\begin{itemize}
+    \item $\displaystyle \iiint_E f\,dV = \int\!\int\!\int f(x,y,z)\,dz\,dy\,dx$ over box or layered bounds.
+    \item Mass/CM/ Moments extend by adding $z$ and $dV$.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Choose order by inspecting projection and boundary dependence.
+    \item Slice by planes to keep inner bounds simple.
+\end{itemize}
+\end{keybox}
 The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
@@
 \subsection{Triple Integrals in Cylindrical Coordinates}
+
+\begin{keybox}
+	\textbf{Key Relations:}
+\begin{itemize}
+    \item $(r,\theta,z)$ with $x=r\cos\theta$, $y=r\sin\theta$, $dV=r\,dr\,d\theta\,dz$.
+    \item Use when symmetry around $z$-axis or $x^2{+}y^2$ appears.
+\end{itemize}
+	\textbf{Tip:} Don’t forget the Jacobian factor $r$.
+\end{keybox}
 \begin{definition}[Cylindrical Coordinates]
@@
 \subsection{Triple Integrals in Spherical Coordinates}
+
+\begin{keybox}
+	\textbf{Key Relations:}
+\begin{itemize}
+    \item $(\rho,\phi,\theta)$ with $x=\rho\sin\phi\cos\theta$, $y=\rho\sin\phi\sin\theta$, $z=\rho\cos\phi$.
+    \item $dV=\rho^2\sin\phi\,d\rho\,d\phi\,d\theta$.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Spheres/cones/spherical caps are natural in these coordinates.
+    \item Bound order: typically $\rho$ (inner), then $\phi$, then $\theta$.
+\end{itemize}
+\end{keybox}
 \begin{definition}[Spherical Coordinates]
@@
 \subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
+
+\begin{keybox}
+	\textbf{Key Equations:}
+\begin{itemize}
+    \item $\displaystyle \iint_R f(x,y)\,dA = \iint_{R'} f(x(u,v),y(u,v))\,\left|\det J\right|\,du\,dv$.
+    \item $J=\partial(x,y)/\partial(u,v)$; similarly for 3D.
+\end{itemize}
+	\textbf{Tips/Pitfalls:}
+\begin{itemize}
+    \item Map the region carefully; sketch both $R$ and $R'$.
+    \item Include the Jacobian; watch sign via absolute value.
+\end{itemize}
+\end{keybox}
 \begin{shaded}
@@
 \section{Line Integrals}
+
+\subsection{Line Integrals and Vector Fields}
+
+\begin{keybox}
+	\textbf{Key Equations:}
+\begin{itemize}
+    \item Scalar field: $\displaystyle \int_C f\,ds = \int_a^b f(\mathbf r(t))\,\lVert \mathbf r'(t)\rVert\,dt$.
+    \item Vector field work: $\displaystyle \int_C \mathbf F\cdot d\mathbf r = \int_a^b \mathbf F(\mathbf r(t))\cdot\mathbf r'(t)\,dt$.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Orientation matters; reverse curve flips sign of vector line integral.
+    \item For conservative fields, path independence and potential functions apply.
+\end{itemize}
+\end{keybox}
 \begin{definition}[Line Integral of a Scalar Field]
@@
 \end{proof}
+
+\begin{definition}[Line Integral of a Vector Field]
+    Consider a vector field $\vec{F}(x,y)$ and a smooth curve $C$. Then we define the line integral of $\vec{F}$ along $C$ as the following Riemann sum:
+    \begin{equation}
+        \int_C \vec{F} \cdot d\vec{r} = \lim_{||P|| \to 0} \sum_{i=1}^n \vec{F}(x_i^*, y_i^*) \cdot \Delta \vec{r_i}
+    \end{equation}
+    where $P$ is a partition of $C$ into $n$ subintervals and $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta \vec{r_i}$ is the vector from the start to the end of the $i$-th subinterval.
+We define $d \vec{r}$ as the infinitesimal vector along the curve:
+    \begin{equation}
+        d\vec{r} = \frac{d\vec{r}}{dt} dt = \vec{T} \, ds
+    \end{equation}
+    where $\vec{T}$ is the unit tangent vector to the curve and $ds$ is the infinitesimal arc length along the curve.
+
+    To evaluate the line integral, we can use the parametrization of the curve $C$:
+    \begin{equation}
+        \int_C F \cdot d\vec{r} = \int_a^b \vec{F}(r(t)) \cdot \vec{r'(t)} \, dt
+    \end{equation}
+    where $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$ is the parametrization of the curve $C$.
+\end{definition}
+
+\begin{example}
+    Consider the line integral of $\vec{F}(x, y, z) = yz \, \hat{i} + xz \, \hat{j} + xy \, \hat{k}$ along the curve $C$ parametrized by $ \vec{r}(t) = t^2 \, \hat{i} + t \, \hat{j} + t^3 \, \hat{k}, \quad 0 \leq t \leq 2 $.
+We have that:
+    \begin{align*}
+        \int_C \vec{F} \cdot d\vec{r} &= \int_0^2 \vec{F}(r(t)) \cdot \vec{r'(t)} \, dt \\
+        &= \int_0^2 \langle t \cdot t^3, t^2 \cdot t^3, t^2 \cdot t \rangle \cdot \langle 2t, 1, 3t^2 \rangle \, dt \\
+        &= \int_0^2 \langle t^4, t^5, t^3 \rangle \cdot \langle 2t, 1, 3t^2 \rangle \, dt \\
+        &= \int_0^2 (2t^5 + t^5 + 3t^5) \, dt = \int_0^2 6t^5 \, dt = \left[ t^6 \right]_0^2 = 64
+    \end{align*}
+\end{example}
+
+\subsection{Fundamental Theorem of Line Integrals}
+
+\begin{keybox}
+	\textbf{Key Statement:}
+\begin{itemize}
+    \item If $\mathbf F=\nabla \phi$ and $C$ goes from $A$ to $B$, then $\displaystyle \int_C \mathbf F\cdot d\mathbf r = \phi(B)-\phi(A)$.
+\end{itemize}
+	\textbf{Consequences:}
+\begin{itemize}
+    \item Path independence; integral on closed loops is $0$.
+    \item Test for conservative in simply connected regions: $\nabla\times\mathbf F=\mathbf 0$.
+\end{itemize}
+\end{keybox}
+\begin{shaded}
+Recall, fundametnal theorem of calculus:
+\begin{theorem}[Fundamental Theorem of Calculus]
+    If $f$ is continuous on $[a,b]$ and $F$ is an antiderivative of $f$ on $[a,b]$, then:
+    \begin{equation}
+        \int_a^b f(x) \, dx = F(b) - F(a)
+    \end{equation}
+\end{theorem}
+\end{shaded}
+
+\begin{definition}[Conservative Vector Field]
+    A vector field $\vec{F}$ is said to be conservative if there exists a scalar potential function $\phi$ such that:
+    \begin{equation}
+        \vec{F} = \nabla \phi
+    \end{equation}
+    where $\nabla \phi$ is the gradient of $\phi$.
+
+\end{definition}
+
+\begin{theorem}[Test for Conservative Vector Field]
+    A vector field $\vec{F} = P \, \hat{i} + Q \, \hat{j} + R \, \hat{k}$ is conservative if:
+    \begin{equation}
+        \frac{\partial P}{\partial y} = \frac{\partial Q}{\partial x}, \quad \frac{\partial P}{\partial z} = \frac{\partial R}{\partial x}, \quad \frac{\partial Q}{\partial z} = \frac{\partial R}{\partial y}
+    \end{equation}
+    in a simply connected domain.
+\end{theorem}
+
+\begin{theorem}[Conservative Closed Integral]
+    If \(\vec F\) is a conservative vector field on an open, simply connected region \(D\), then for every closed curve \(C\) in \(D\),
+    \[
+        \oint_C \vec F \cdot d\vec r = 0
+    \]
+\end{theorem}
+
+\begin{theorem}[Fundamental Theorem of Line Integrals]
+    Let $\vec{F}$ be a continuous vector field on an open connected region $D$. If $\vec{F}$ is conservative, then for any smooth curve $C$ in $D$ with endpoints $A$ and $B$, we have:
+    \begin{equation}
+        \int_C \vec{F} \cdot d\vec{r} = \phi(B) - \phi(A)
+    \end{equation}
+    where $\phi$ is the potential function of $\vec{F}$.
+\end{theorem}
+
+\begin{proof}
+Let \(\vec F\) be conservative on the open, connected region \(D\). Then there exists a scalar potential function \(\phi\in C^1(D)\) with \(\nabla\phi=\vec F\).
+
+Let \(C\) be any smooth curve in \(D\) from \(A\) to \(B\). Parametrise \(C\) by a smooth vector function
+\[
+\vec r:[a,b]\to D,\qquad \vec r(t)=(x(t),y(t),z(t)),
+\]
+with \(\vec r(a)=A\) and \(\vec r(b)=B\). Since \(\vec F=\nabla\phi\) and \(\vec r\) is smooth, the integrand \(\nabla\phi(\vec r(t))\cdot\vec r\,'(t)\) is continuous on \([a,b]\). By the chain rule,
+\[
+\frac{d}{dt}\big(\phi(\vec r(t))\big)
+= \nabla\phi(\vec r(t))\cdot\vec r\,'(t).
+\]
+Therefore the line integral of \(\vec F\) along \(C\) becomes
+\[
+\int_C \vec F\cdot d\vec r
+= \int_a^b \nabla\phi(\vec r(t))\cdot\vec r\,'(t)\,dt
+= \int_a^b \frac{d}{dt}\big(\phi(\vec r(t))\big)\,dt.
+\]
+Applying the Fundamental Theorem of Calculus to the last integral yields
+\[
+\int_a^b \frac{d}{dt}\big(\phi(\vec r(t))\big)\,dt
+= \phi(\vec r(b))-\phi(\vec r(a))
+= \phi(B)-\phi(A).
+\]
+If \(C\) is only piecewise smooth the same argument applies on each smooth piece and the contributions telescope to give the same result.
+\end{proof}
+
+\subsection{Green's Theorem}
+
+\begin{keybox}
+	\textbf{Theorem (Plane):}
+\begin{itemize}
+    \item For region $R$ with positively oriented boundary $C$ and $\mathbf F=\langle P,Q\rangle$,
+    $\displaystyle \oint_C P\,dx+Q\,dy=\iint_R \left(\frac{\partial Q}{\partial x}-\frac{\partial P}{\partial y}\right)\,dA$.
+\end{itemize}
+	\textbf{Tips:}
+\begin{itemize}
+    \item Orientation CCW for boundary; decompose region if needed.
+    \item Convert flux/ circulation forms appropriately.
+\end{itemize}
+\end{keybox}
+\begin{definition}[Simple Curve]
+    A simple curve is a curve that does not intersect itself.
+\end{definition}
+\paragraph{Positive Orientation} A curve is said to be positively oriented if it is traversed in a counterclockwise direction.
+Some books denote the direction in the circle of the $\oint$ symbol.
+To convert a negatively oriented curve to a positively oriented curve, we can simply reverse the direction of the curve.
+We also have:
+\begin{equation}
+    \oint_C F \cdot d\vec{r} = -\oint_{-C} F \cdot d\vec{r} \\
+\end{equation}
+where $-C$ is the curve $C$ traversed in the opposite direction.
+\begin{definition}[Closed Integral]
+    A closed integral is a line integral where the curve is closed, i.e., the starting point and the ending point are the same. We denote a closed integral by the symbol \(\oint\). For example:
+    \[
+        \oint_C P \, dx + Q \, dy
+    \]
+\end{definition}
+
+\begin{theorem}[Green's Theorem]
+    Let \(C\) be a positively oriented, piecewise smooth, simple closed curve in the plane, and let \(D\) be the region bounded by \(C\).
+If \(P\) and \(Q\) have continuous partial derivatives on an open region that contains \(D\), then:
+    \[
+    \oint_C (P \, dx + Q \, dy) = \iint_D \left( \frac{\partial Q}{\partial x} - \frac{\partial P}{\partial y} \right) dA
+    \]
+\end{theorem}
+
+\begin{proof}
+    \textbf{Simple Case:}
+    Assume the region \(D\) is a simple region that could be expressed by both a type I and type II region. That is, we can express \(D\) as:
+    \[
+        D = \{(x,y) | a \leq x \leq b, g_1(x) \leq y \leq g_2(x) \} = \{(x,y) | c \leq y \leq d, h_1(y) \leq x \leq h_2(y) \}
+    \]
+    where \(g_1\), \(g_2\), \(h_1\) and \(h_2\) are continuous functions on their respective intervals. WLOG, we can prove for type I region. We can expressed the first part of the integral as:
+    \begin{align*}
+        \oint_C P \, dx &= \int_{C_1} P \, dx + \int_{C_2} P \, dx \\
+        &= \int_a^b P(x, g_1(x)) \, dx + \int_b^a P(x, g_2(x)) \, dx \\
+        &= \int_a^b P(x, g_1(x)) \, dx - \int_a^b P(x, g_2(x)) \, dx \\
+        &= \int_a^b [P(x, g_1(x)) - P(x, g_2(x))] \, dx \\
+        &= -\int_a^b \int_{g_1(x)}^{g_2(x)} \frac{\partial P}{\partial y} \, dy \, dx
+    \end{align*}
+    Then, we can also prove in a similar manner for the second part of the integral by expressing it as a type II region. We obtain:
+    $$
+        \oint_C Q \, dy = \int_c^d \int_{h_1(y)}^{h_2(y)} \frac{\partial Q}{\partial x} \, dx \, dy
+    $$
+    The results of both integrals can be combined, and the theorem follows.
+\end{proof}
+
+\begin{example}
+    \textit{Verify Green's Theorem for the integral:}
+    $$
+        \oint_C y \, dx - x \, dy
+    $$
+    \textit{where \(C\) is the circle \(x^2 + y^2 = 1\) oriented counterclockwise.}
+
+    We can parametrize the curve \(C\) as:
+    $$
+    \begin{cases}
+        x = \cos t \\
+        y = \sin t
+    \end{cases} \quad 0 \leq t \leq 2\pi
+    $$
+    Then, we can evaluate the line integral directly:
+    \begin{align*}
+        \oint_C y \, dx - x \, dy &= \oint_C \langle y, -x \rangle \cdot d\vec{r} = \int_0^{2\pi} \langle \sin t, -\cos t \rangle \cdot \langle -\sin t, \cos t \rangle \, dt \\
+        &= \int_0^{2\pi} \sin t (-\sin t) - \cos t (\cos t) \, dt \\
+        &= \int_0^{2\pi} -\sin^2 t - \cos^2 t \, dt = \int_0^{2\pi} -1 \, dt = -2\pi
+    \end{align*}
+    Now, we can verify Green's Theorem by evaluating the double integral:
+    \begin{align*}
+        \iint_D \left( \frac{\partial Q}{\partial x} - \frac{\partial P}{\partial y} \right) dA &= \iint_D \left( -1 - 1 \right) dA = \iint_D -2 \, dA \\
+        &= -2 \cdot \text{Area}(D) = -2 \cdot \pi = -2\pi
+    \end{align*}
+    Both methods yield the same result, confirming Green's Theorem.
+
+\end{example}
+
+\begin{example}
+    \textit{Evaluate the line integral:}
+    $$
+        \oint_C \left(4 - e^{\sqrt{x}}\right) \, dx + \left(\sin{y} + 3x^2\right) \, dy
+    $$
+    \textit{where \(C\) is the curve in the first quadrant defined by two circles with radius $a$ and $b$ ($b > a$).}
+
+    We can use Green's Theorem to evaluate the line integral. We can reduce to a much simpler double integral:
+    \begin{align*}
+        \iint_D \left( \frac{\partial Q}{\partial x} - \frac{\partial P}{\partial y} \right) dA &= \iint_D \left( 6x - 0 \right) dA = \iint_D 6x \, dA
+    \end{align*}
+    We can express the region \(D\) in polar coordinates as:
+    $$    D = \{(r, \theta) | a \leq r \leq b, 0 \leq \theta \leq \frac{\pi}{2} \} $$
+    So, we can express the double integral in polar coordinates:
+    \begin{align*}
+        \iint_D 6x \, dA &= \int_0^{\frac{\pi}{2}} \int_a^b 6r \cos \theta \cdot r \, dr \, d\theta = \int_0^{\frac{\pi}{2}} 6\cos \theta \int_a^b r^2 \, dr \, d\theta \\
+        &= \int_0^{\frac{\pi}{2}} 6\cos \theta \left[ \frac{r^3}{3} \right]_a^b \, d\theta = 2(b^3 - a^3) \int_0^{\frac{\pi}{2}} \cos \theta \, d\theta \\
+        &= 2(b^3 - a^3)
+    \end{align*}
+\end{example}
+
+\subsection{Surfaces and Their Areas}
+\begin{definition}[3-Dimensional Parametric Surface]
+    A 3-dimensional parametric surface is a surface that can be parametrized by two parameters \(u\) and \(v\):
+    \begin{equation}
+        \begin{cases}
+            x = g(u,v) \\
+            y = h(u,v) \\
+            z = k(u,v)
+        \end{cases} \implies \vec{r(u,v)} = \langle g(u,v), h(u,v), k(u,v) \rangle, \quad (u,v) \in D
+    \end{equation}
+    where \(g\), \(h\) and \(k\) are continuous functions with continuous partial derivatives on the domain \(D\).
+\end{definition}
+
+\begin{example}
+    \textit{Parameterize the upper hemisphere given by \(x^2 + y^2 + z^2 = a^2\).}
+
+    \textbf{Method 1:} We can express \(z\) in terms of \(x\) and \(y\):
+    $$    
+        z = \sqrt{a^2 - x^2 - y^2}
+    $$
+    Then, we let $x = u$ and $y = v$. So, we can parametrize the surface as:
+    $$    
+        \vec{r(u,v)} = \langle u, v, \sqrt{a^2 - u^2 - v^2} \rangle
+    $$
+    where \((u,v) \in D = \{(u,v) | u^2 + v^2 \leq a^2\}\).
+
+    \textbf{Method 2:} We can also use spherical coordinates to parametrize the surface. We have:
+    $$    
+        \begin{cases}
+            x = a \sin \phi \cos \theta \\
+            y = a \sin \phi \sin \theta \\
+            z = a \cos \phi
+        \end{cases} \implies \vec{r(\phi, \theta)} = \langle a \sin \phi \cos \theta, a \sin \phi \sin \theta, a \cos \phi \rangle
+    $$
+    where \(0 \leq \phi \leq \frac{\pi}{2}\) and \(0 \leq \theta \leq 2\pi\).
+\end{example}
+
 \chapter{Fluid Mechanics}

### 2025-10-08 YEAR2/ECE253_LectureNotes/main.tex
oct 7

@@
 \end{definition}
+
+\begin{example}[Clock Reduction]
+    Suppose we want to reduce a 1Mhz clock to a 1Hz clock.
+We can use a 20-bit counter to achieve this.
+Since $2^{20} = 1,048,576$, which is slightly more than 1 million, we can use the most significant bit (MSB) of the counter as the output clock.
+The MSB will toggle every $2^{19}$ clock cycles, which is approximately 1 second when the input clock is 1MHz.
+The Verilog implementation is as follows:
+    \begin{verbatim}
+    module clock_divider(input logic CLK_1MHZ, RSTn,
+                         output logic CLK_1HZ);
+        logic [19:0] counter; // 20-bit counter     
+        always_ff @(posedge CLK_1MHZ or negedge RSTn) begin
+            if (RSTn == 1'b0)
+                counter <= 20'b0;
+            else
+                counter <= counter + 1; // Increment counter
+        end
+        assign CLK_1HZ = counter[19]; // MSB as 1Hz clock
+    endmodule
+    \end{verbatim}
+
+    In a real hardware implementation, we would feed the bit to a d flip-flop, and that would create the rising and falling edge signals.
+\end{example}
+
+\subsection{Optimizations}
+
+\begin{definition}[Karnaugh Map (K-Map)]
+    A Karnaugh map (K-map) is a graphical representation of a truth table that is used to simplify boolean algebra expressions.
+It is a grid-like structure that organizes the minterms of a boolean function in such a way that adjacent cells differ by only one variable.
+By grouping adjacent cells that contain 1s, we can identify common factors and simplify the boolean expression.
+
+    The K-map for a 3-variable function is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{BC} \\
+            \cline{2-5}
+            A & 00 & 01 & 11 & 10 \\
+            \hline
+            0 & m0 & m1 & m3 & m2 \\
+            1 & m4 & m5 & m7 & m6 \\
+        \end{tabular}
+        \caption{3-Variable Karnaugh Map}
+        \label{fig:3_variable_karnaugh_map}
+    \end{figure}
+
+    The K-map for a 4-variable function is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c|c}
+            & \multicolumn{4}{c}{CD} \\
+            \cline{2-5}
+            AB & 00 & 01 & 11 & 10 \\
+            \hline
+            00 & m0 & m1 & m3 & m2 \\
+            01 & m4 & m5 & m7 & m6 \\
+            11 & m12 & m13 & m15 & m14 \\
+            10 & m8 & m9 & m11 & m10 \\
+        \end{tabular}
+        \caption{4-Variable Karnaugh Map}
+        \label{fig:4_variable_karnaugh_map}
+    \end{figure}
+
+    The following rules are used to group the minterms in a K-map:
+    \begin{itemize}
+        \item Groups must contain only 1s.
+        \item Groups must be rectangular and contain $2^n$ cells (where n is a non-negative integer).
+        \item Groups can wrap around the edges of the K-map.
+        \item Each group should be as large as possible.
+        \item Each 1 in the K-map should be covered by at least one group.
+    \end{itemize}
+    By following these rules, we can create groups of 1s in the K-map and simplify the boolean expression accordingly.
+
+\end{definition}
+
+\begin{example}
+    Suppose we have the following truth table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c||c|}
+            \hline
+            A & B & C & F \\
+            \hline
+            0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 1 \\
+            0 & 1 & 0 & 1 \\
+            0 & 1 & 1 & 1 \\
+            1 & 0 & 0 & 0 \\
+            1 & 0 & 1 & 1 \\
+            1 & 1 & 0 & 1 \\
+            1 & 1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table for Example}
+        \label{tab:example_truth_table}
+    \end{table}
+    We can derive the following minterms from the truth table:
+    $$
+    F(A,B,C) = m_1 + m_2 + m_3 + m_5 + m_6 + m_7
+    $$
+    The K-map for the function is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{BC} \\
+            \cline{2-5}
+            A & 00 & 01 & 11 & 10 \\
+            \hline
+            0 & 0 & 1 & 1 & 1 \\
+            1 & 0 & 1 & 1 & 1 \\
+        \end{tabular}
+        \caption{Karnaugh Map for Example}
+        \label{fig:example_karnaugh_map}
+    \end{figure}
+    We can group the minterms as follows:
+    \begin{itemize}
+        \item Group of four covering $m_2, m_3, m_6, m_7$: This group corresponds to $B = 1$ (since B is constant in this group). So this group simplifies to $B$.
+        \item Group of two covering $m_1, m_5$: This group corresponds to $A = 0$ and $C = 1$ (since A and C are constant in this group). So this group simplifies to $\overline{A}C$.
+    \end{itemize}
+    Therefore, the simplified boolean expression is:
+    $$
+    F(A,B,C) = B + \overline{A}C
+    $$
+\end{example}
+
+\begin{definition}[Don't Care Conditions]
+    In some cases, certain input combinations may never occur or their output values are irrelevant to the overall function.
+These input combinations are referred to as "don't care" conditions and are typically represented by an "X" in the truth table.
+When simplifying a boolean expression using a K-map, we can treat these "don't care" conditions as either 0 or 1, whichever leads to a simpler expression.
+This flexibility can help in creating larger groups of 1s in the K-map, leading to a more simplified boolean expression.
+
+\end{definition}
+\begin{example}[Don't Care Conditions]
+    Suppose we have the following truth table with don't care conditions:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c||c|}
+            \hline
+            A & B & C & F \\
+            \hline
+            0 & 0 & 0 & 0 \\
+            0 & 0 & 1 & 1 \\
+            0 & 1 & 0 & X \\
+            0 & 1 & 1 & 1 \\
+            1 & 0 & 0 & 0 \\
+            1 & 0 & 1 & X \\
+            1 & 1 & 0 & 1 \\
+            1 & 1 & 1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{Truth Table with Don't Care Conditions}
+        \label{tab:example_dont_care_truth_table}
+    \end{table}
+    We can derive the following minterms from the truth table:
+    $$
+    F(A,B,C) = m_1 + m_3 + m_6 + m_7
+    $$
+    The K-map for the function is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{BC} \\
+            \cline{2-5}
+            A & 00 & 01 & 11 & 10 \\
+            \hline
+            0 & 0 & X & 1 & X \\
+            1 & 0 & X & 1 & X \\
+        \end{tabular}
+        \caption{Karnaugh Map with Don't Care Conditions}
+        \label{fig:example_dont_care_karnaugh_map}
+    \end{figure}
+    We can group the minterms as follows:
+    \begin{itemize}
+        \item Group of four covering $m_3, m_7$ and the don't care conditions: This group corresponds to $C = 1$ (since C is constant in this group). So this group simplifies to $C$.
+        \item Group of two covering $m_1$: This group corresponds to $A = 0$ and $B = 0$ (since A and B are constant in this group). So this group simplifies to $\overline{A}\overline{B}$.
+    \end{itemize}
+    Therefore, the simplified boolean expression is:
+    $$
+    F(A,B,C) = C + \overline{A}\overline{B}
+    $$
+\end{example}
+
+\paragraph{Gray Code}
+Gray code is a binary numeral system where two successive values differ in only one bit.
+This property makes Gray code useful in minimizing errors in digital systems, particularly in applications like rotary encoders and digital communication.
+The following table shows the 3-bit binary and Gray code representations:
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|}
+        \hline
+        Binary & Gray Code \\
+        \hline
+        000 & 000 \\
+        001 & 001 \\
+        010 & 011 \\
+        011 & 010 \\
+        100 & 110 \\
+        101 & 111 \\
+        110 & 101 \\
+        111 & 100 \\
+        \hline
+    \end{tabular}
+    \caption{3-bit Binary and Gray Code Representations}
+    \label{tab:3_bit_binary_gray_code}
+\end{table}
+\paragraph{Terminologies} For a K-map, we have the following terminologies:
+\begin{itemize}
+    \item \textbf{Implicant:} A product term (AND term) that can be used to cover one or more minterms in the K-map.
+    \item \textbf{Prime Implicant:} An implicant that cannot be combined with any other implicant to form a larger implicant.
+In other words, it is an implicant that covers the maximum number of minterms possible.
+    \item \textbf{Essential Prime Implicant:} A prime implicant that covers at least one minterm that is not covered by any other prime implicant.
+In other words, it is a prime implicant that is necessary to cover all the minterms in the K-map.
+    \item \textbf{Cover:} A set of implicants that together cover all the minterms in the K-map.
+\end{itemize}
+
+\begin{example}[Implicant and Prime Implicant]
+    Consider the following K-map:
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{BC} \\
+            \cline{2-5}
+            A & 00 & 01 & 11 & 10 \\
+            \hline
+            0 & 1 & 1 & 0 & 0 \\
+            1 & 1 & 1 & 1 & 0 \\
+        \end{tabular}
+        \caption{Karnaugh Map for Example}
+        \label{fig:example_karnaugh_map_implicant}
+    \end{figure}
+    We can identify the following implicants:
+    \begin{itemize}
+        \item $A'B'$ covering minterms $m_0, m_1$
+        \item $A'B$ covering minterms $m_2, m_3$
+        \item $AB'$ covering minterms $m_4, m_5$
+        \item $AB$ covering minterms $m_6, m_7$
+        \item $B$ covering minterms $m_2, m_3, m_6, m_7$
+        \item $A$ covering minterms $m_4, m_5, m_6, m_7$
+    \end{itemize}
+    The prime implicants are:
+    \begin{itemize}
+        \item $A'B'$ (cannot be combined with any other implicant)
+        \item $B$ (cannot be combined with any other implicant)
+        \item $A$ (cannot be combined with any other implicant)
+    \end{itemize}
+    The essential prime implicants are:
+    \begin{itemize}
+        \item $A'B'$ (covers minterm $m_0$ which is not covered by any other prime implicant)
+        \item $B$ (covers minterms $m_2, m_3$ which are not covered by any other prime implicant)
+        \item $A$ (covers minterms $m_4, m_5$ which are not covered by any other prime implicant)
+    \end{itemize}
+    Therefore, the cover of the K-map is:
+    $$
+    F(A,B,C) = A'B' + B + A
+    $$
+\end{example}
+
+\begin{example}
+    Let $f(x_1, x_2, x_3, x_4) = \sum m(2, 4, 5, 8, 10, 11, 12, 13, 15)$ be a function represented by the following K-map:
+    % up : x_1 x_2, left: x_3 x_4
+    \begin{figure}[h!]
+        \centering
+        \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{$x_1 x_2$} \\
+            \cline{2-5}
+            $x_3 x_4$ & 00 & 01 & 11 & 10 \\
+            \hline
+            00 & 0 & 1 & 1 & 1 \\
+            01 & 0 & 1 & 1 & 0 \\
+            11 & 0 & 0 & 1 & 1 \\
+            10 & 1 & 0 & 0 & 1 \\   
+        \end{tabular}
+        \caption{Karnaugh Map for Example}
+        \label{fig:example_karnaugh_map_dont_care}
+    \end{figure}
+
+\end{example}
+
+\paragraph{Procedure to fine the minimum cost cover}
+\begin{enumerate}
+    \item Find all prime implicants.
+    \item Identify all essential prime implicants (EPIs) and include them in the cover.
+    \item Remove all minterms covered by the EPIs from the K-map.
+    \item Check and inspect whether it is the minimum cost cover. 
+\end{enumerate}
+
+\begin{definition}[K-Maps with Maxterm]
+    We can also use K-maps to simplify boolean expressions represented by maxterms.
+The procedure is similar to that of minterms, but instead of grouping 1s, we group 0s.
+The following rules are used to group the maxterms in a K-map:
+    \begin{itemize}
+        \item Groups must contain only 0s.
+        \item Groups must be rectangular and contain $2^n$ cells (where n is a non-negative integer).
+        \item Groups can wrap around the edges of the K-map.
+        \item Each group should be as large as possible.
+        \item Each 0 in the K-map should be covered by at least one group.
+    \end{itemize}
+    By following these rules, we can create groups of 0s in the K-map and
+\end{definition}
+
 \section{Finite State Machines (FSM)}
-
+\begin{definition}[State Diagram]
+    A state diagram is a graphical representation of a finite state machine (FSM) that illustrates the states, transitions, inputs, and outputs of the system.
+It consists of nodes (representing states) and directed edges (representing transitions between states).
+Each edge is typically labeled with the input condition that triggers the transition and may also include the output associated with that transition.
+
+    State diagrams are used to visualize the behavior of sequential systems, making it easier to understand how the system responds to different inputs and how it transitions between various states over time.
+\end{definition}
+
+\begin{example}
+    Cosnider a circult that detect the input sequance $w=$`101' using a FSM. The state diagram is shown below:
+    \begin{figure}[h!]
+        \centering
+        \begin{tikzpicture}[node distance=2cm, on grid, auto]
+            % Nodes
+            \node[state, initial] (S0) {S0};
+            \node[state] (S1) [right=of S0] {S1};
+            \node[state] (S2) [right=of S1] {S2};
+            \node[state, accepting] (S3) [right=of S2] {S3};    
+            % Edges
+            \path[->]
+                (S0) edge [loop above] node {0/0} ()
+                     edge node {1/0} (S1)
+                (S1) edge node {0/0} (S2)
+                     edge [loop above] node {1/0} ()
+                (S2) edge node {1/1} (S3)
+                     edge [bend left] node {0/0} (S0)
+                (S3) edge [bend left] node {0/0} (S2)
+                     edge [bend left] node {1/0} (S1);
+        \end{tikzpicture}
+        \caption{State Diagram for FSM that detects `101'}
+        \label{fig:fsm_state_diagram}
+    \end{figure}
+    The FSM has four states: S0, S1, S2, and S3.
+The initial state is S0, and the accepting state is S3.
+The transitions between states are labeled with the input/output pairs.
+For example, from state S0, if the input is 1, the FSM transitions to state S1 and outputs 0.
+The FSM detects the sequence `101' and outputs 1 when it reaches state S3.
+
+    We can also draw its state table as follows:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|cc|c|}
+            \hline
+            Curreent State & $w=0$ & $w=1$ & Output \\
+            \hline
+            S0 & S0 & S1 & 0 \\
+            S1 & S2 & S1 & 0 \\
+            S2 & S0 & S3 & 0 \\
+            S3 & S2 & S1 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{State Table for FSM that detects `101'}
+        \label{tab:fsm_state_table}
+    \end{table}
+
+    Then, we would assign the states. First the circuit has 4 states, so we need at least 2 flip-flops $y_2$, $y_1$ to represent the states. We can assign the states as follows:
+    \begin{itemize}
+        \item S0 = 00, $y_2 y_1$
+        \item S1 = 01, $y_2 \overline{y_1}$
+        \item S2 = 10, $\overline{y_2} y_1$
+        \item S3 = 11, $\overline{y_2} \overline{y_1}$
+    \end{itemize}
+    Then, we can derive the following state transition table:
+
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|cc|c|}
+            \hline
+            Current State ($y_2 y_1$) & $w=0$ & $w=1$ & Output \\
+            \hline
+            00 & 00 & 01 & 0 \\
+            01 & 10 & 01 & 0 \\
+            10 & 00 & 11 & 0 \\
+            11 & 10 & 01 & 1 \\
+            \hline
+        \end{tabular}
+        \caption{State Transition Table for FSM that detects `101'}
+        \label{tab:fsm_state_transition_table}
+    \end{table}
+
+    From the state transition table, we can synthesis the circuit using D flip-flops. Let $Y_2$ and $Y_1$ be the next state of $y_2$ and $y_1$, respectively. We can derive the following equations:
+    \begin{itemize}
+        \item $Y_1 = w$
+        \item $Z = y_2 y_1$
+        \item $Y_2 = \overline{w} y_1 + w \overline{y_2} \overline{y_1}$
+    \end{itemize}
+
+\end{example}
+
+\begin{definition}[Finite State Machine (FSM)]
+    A finite state machine (FSM) is a mathematical model of computation used to design both computer programs and sequential logic circuits.
+It is composed of a finite number of states, transitions between those states, inputs, and outputs.
+FSMs are used to model the behavior of systems that can be in one of a limited number of states at any given time and can transition from one state to another based on input signals.
+
+    There are two main types of FSMs:
+    \begin{itemize}
+        \item \textbf{Moore Machine:} In a Moore machine, the outputs depend only on the current state. The output is associated with each state, and it changes only when the state changes.
+        \item \textbf{Mealy Machine:} In a Mealy machine, the outputs depend on both the current state and the current inputs.
+The output can change immediately in response to changes in the input, even if the state remains the same.
+    \end{itemize}
+\end{definition}
 \chapter{Computer Organization and Assembly Language}

### 2025-10-08 YEAR2/MAT292_LectureNotes/main.tex
oct 7

@@
 
-\begin{definition}
+\begin{example}
     Consider the nonlinear pendulum equation:
@@
     When $n$ is even, we have a positive determinant, so the fixed point is a center. When $n$ is odd, we have a negative determinant, so the fixed point is a saddle node.
-\end{definition}
+\end{example}
+
+\begin{definition}[Polar Coordinates]
+    Consider the transformation from Cartesian coordinates \( (x, y) \) to polar coordinates \( (r, \theta) \) given by:
+    \begin{equation}
+        \begin{cases}
+        x = r \cos(\theta) \\
+        y = r \sin(\theta)
+        \end{cases} \implies x^2 + y^2 = r^2, \quad \tan(\theta) = \frac{y}{x}
+        \end{equation}
+    The inverse transformation is given by:
+    \begin{equation}
+        \begin{cases}
+        r = \sqrt{x^2 + y^2} \\
+        \theta = \tan^{-1}\left(\frac{y}{x}\right)
+        \end{cases}
+    \end{equation}
+
+    A trick we can do is to consider the time derivatives of both sides. We have:
+    \begin{align*}
+     2 r r' &= 2x x' + 2y y'  \\
+     r' &= \frac{x x' + y y'}{r} \\
+     \theta' &= \frac{1}{1 + (y/x)^2} \left(\frac{x y' - y x'}{x^2}\right) \\
+     &= \frac{x y' - y x'}{r^2}
+    \end{align*}
+
+    Then we can transform a system of ODEs in Cartesian coordinates to polar coordinates:
+    \begin{equation}
+        \begin{cases}
+        x' = f(x, y) \\
+        y' = g(x, y)
+        \end{cases} \implies
+        \begin{cases}
+        r' = F(r, \theta) = \frac{x f(x, y) + y g(x, y)}{r} \\
+        \theta' = G(r, \theta) = \frac{x g(x, y) - y f(x, y)}{r^2}
+        \end{cases}
+    \end{equation}
+
+\end{definition}
+
+\paragraph{Relationship to Eigenvectors} Consider the nullclines $\theta' = 0$. Along this nullcline, the direction on the $xy$-plane is purely radial (either inward or outward). This occurs when:
+$$x y' - y x' = 0 \implies \frac{y'}{x'} = \frac{y}{x}$$
+This means that the slope of the trajectory is equal to the slope of the position vector, which corresponds to the direction of an eigenvector of the system.
+Thus, the nullclines in polar coordinates correspond to the directions of the eigenvectors in the Cartesian coordinate system.
+
+\paragraph{Clockwise Parity} For clockwise rotation, we have $\theta' < 0$. For counter-clockwise rotation, we have $\theta' > 0$.
+
+\begin{example}[Limit Cycles and Closed Orbits]
+    Consider the following system of ODEs:
+    $$\begin{cases}
+        x' = -y + x(1 - x^2 - y^2) \\
+        y' = x + y(1 - x^2 - y^2)
+    \end{cases}
+    $$
+    The system is nonlinear. There is a trivial (in fact, it is the only) fixed point at the origin. We can determine the stability of this fixed point by analyzing the Jacobian matrix:
+    \begin{align*}
+        J &= \begin{bmatrix}
+        \frac{\partial f}{\partial x} & \frac{\partial f}{\partial y} \\
+        \frac{\partial g}{\partial x} & \frac{\partial g}{\partial y}
+        \end{bmatrix} \\ 
+        &= \begin{bmatrix}
+        1 - 3x^2 - y^2 & -1 - 2xy \\
+        1 - 2xy & 1 - x^2 - 3y^2
+        \end{bmatrix} \\
+        J(0, 0) &= \begin{bmatrix}
+        1 & -1 \\
+        1 & 1
+        \end{bmatrix} \\
+        \det(J(0, 0)) &= 2 > 0 \\
+        \text{tr}(J(0, 0)) &= 2 > 0
+    \end{align*}
+
+    Then, we conclude that this is a outward spiral (spiral source). We can also convert this system into polar coordinates:
+    \begin{align*}
+        r' &= \left [ r\cos \theta [ -r\sin \theta + r\cos \theta (1 - r^2)] + r\sin \theta [r\cos \theta + r\sin \theta (1 - r^2)] \right ] / r \\
+        &= r(1 - r^2) \\
+        \theta' &= \left [ r\cos \theta [r\cos \theta + r\sin \theta (1 - r^2)] - r\sin \theta [-r\sin \theta + r\cos \theta (1 - r^2)] \right ] / r^2 \\
+        &= 1
+    \end{align*}
+    So, we have a decoupled system with constant angular velocity of 1 (counter-clockwise rotation) and a radial component that is positive when \( r < 1 \) and negative when \( r > 1 \).
+On $r = 1$, we have a have a closed orbit.
+Now, we can solve this system:
+    \begin{align*}
+        \frac{dr}{dt} &= r(1 - r^2) \\
+        \int \frac{1}{r(1 - r^2)} dr &= \int dt \\
+        \ln \left | \frac{r}{\sqrt{1 - r^2}} \right | &= t + C \\
+        \frac{r}{\sqrt{1 - r^2}} &= Ke^t \\
+        r(t) &= \frac{Ke^t}{\sqrt{1 + K^2 e^{2t}}}
+    \end{align*}
+    also, obviously, \( \theta(t) = t + \theta(0) \).
+
+    where \( K = e^C \) is a constant determined by the initial condition. The phase portrait is shown below:
+
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=\textwidth]{limitCycle.jpg}
+        \caption{Phase Portrait of the Nonlinear Pendulum System}
+
+    \end{figure}
+\end{example}
+
+\begin{definition}[Limit Cycle]
+    A limit cycle is a closed trajectory in the phase space of a dynamical system that is isolated.
+This means that there are no other closed trajectories in its immediate vicinity.
+Limit cycles can be stable, unstable, or semi-stable:
+    \begin{itemize}
+        \item A stable limit cycle attracts nearby trajectories, meaning that trajectories starting close to the limit cycle will converge to it as time goes to infinity.
+        \item An unstable limit cycle repels nearby trajectories, meaning that trajectories starting close to the limit cycle will diverge away from it as time goes to infinity.
+        \item A semi-stable limit cycle has both attracting and repelling properties depending on the direction from which nearby trajectories approach it.
+    \end{itemize}
+
+\end{definition}
+
+\begin{example}[Von der Pol Oscillator]
+    Consider the Van der Pol oscillator, which is a classic example of a system that exhibits limit cycle behavior. The system is described by the second-order differential equation:
+    $$\frac{d^2x}{dt^2} - \mu(1 - x^2)\frac{dx}{dt} + x = 0$$
+    where \( \mu \) is a scalar parameter indicating the nonlinearity and the strength of the damping. We can rewrite this as a system of first-order ODEs by defining:
+    $$\begin{cases}
+    x_1 = x \\
+    x_2 = \frac{dx}{dt}
+    \end{cases}$$
+    Then, we have:
+    $$\begin{cases}
+    x_1' = x_2 \\
+    x_2' = \mu(1 - x_1^2)x_2 - x_1
+    \end{cases}$$
+    Consider the fixed point at the origin \( (0, 0) \). We can also consider the nullclines:
+    \begin{itemize}
+        \item \( x_1' = 0 \) implies \( x_2 = 0 \) and \( x_2' = -x_1 \)
+        \item \( x_2' = 0 \) implies \( x_2 = \frac{x_1}{\mu(1 - x_1^2)} \) for \( x_1 \neq \pm 1 \)
+    \end{itemize}
+
+    We now consider the Jacobian matrix:
+    \begin{align*}
+        J &= \begin{bmatrix}
+        \frac{\partial f}{\partial x_1} & \frac{\partial f}{\partial x_2} \\
+        \frac{\partial g}{\partial x_1} & \frac{\partial g}{\partial x_2}
+        \end{bmatrix} \\ 
+        &= \begin{bmatrix}
+        0 & 1 \\
+        -1 - 2\mu x_1 x_2 & \mu(1 - x_1^2)
+        \end{bmatrix} \\
+        J(0, 0) &= \begin{bmatrix}
+        0 & 1 \\
+        -1 & \mu
+        \end{bmatrix} \\
+        \det(J(0, 0)) &= 1 > 0 \\
+        \text{tr}(J(0, 0)) &= \mu
+    \end{align*}
+
+    Consider the line $\tau^2 - 4\Delta = 0$, we have:
+    \begin{itemize}
+        \item If \( \mu < 0 \), then the fixed point is a stable spiral (spiral sink).
+        \item If \( \mu = 0 \), then the fixed point is a center.
+        \item If \( \mu > 0 \), then the fixed point is an unstable spiral (spiral source).
+    \end{itemize}
+    Thus, this is a limit cycle that is stable when \( \mu > 0 \). 
+\end{example}
+
+\begin{theorem}[Pomcor\'e-Bendixson Theorem]
+    Consider a two-dimensional autonomous system:
+    $$\begin{cases}
+        x' = f(x, y) \\
+        y' = g(x, y)
+    \end{cases}$$
+    where \( f \) and \( g \) are continuously differentiable functions.
+If there exists a trajectory that remains in a closed, bounded region \( R \) of the phase plane for all future time and if \( R \) contains only unstable nodes, then all trajectories in \( R \) approach a limit cycle as \( t \to \infty \).
+\end{theorem}
+\paragraph{Intuition} A monotonically increasing function that gets bounded must converge.
+
+\section{Matrices and Linear Algebra}
+
+\subsection{Matrix Exponentials}
+\begin{shaded}
+\subsubsection*{Taylor Series Definition of Exponentials}
+Consider a scalar exponential, then,
+$$e^x = \sum_{n=0}^{\infty} \frac{x^n}{n!}$$
+is the Taylor series expansion of \( e^x \) around \( x = 0 \). This series converges for all real numbers \( x \).
+\end{shaded}
+
+\begin{definition}[Matrix Exponential]
+    Consider a square matrix \( A \in \mathbb{R}^{n \times n} \). The matrix exponential \( e^{At} \) is defined by the Taylor series expansion:
+    \begin{equation}
+    e^{At} = I + At + \frac{(At)^2}{2!} + \frac{(At)^3}{3!} + \ldots = \sum_{k=0}^{\infty} \frac{(At)^k}{k!}
+    \end{equation}
+    where \( I \) is the identity matrix of the same dimension as \( A \), and \( t \) is a scalar parameter (often representing time in applications).
+\end{definition}
+
+\paragraph{Diagonalization} If \( A \) is diagonalizable, i.e., there exists an invertible matrix \( P \) and a diagonal matrix \( D \) such that \( A = PDP^{-1} \), then the matrix exponential can be computed as:
+\begin{equation}
+e^{At} = Pe^{Dt}P^{-1}
+\end{equation}
+where \( e^{Dt} \) is computed by exponentiating each diagonal element of \( D \):
+\begin{equation}
+e^{Dt} = \begin{bmatrix}
+e^{d_1t} & 0 & \cdots & 0 \\
+0 & e^{d_2t} & \cdots & 0 \\
+\vdots & \vdots & \ddots & \vdots \\
+0 & 0 & \cdots & e^{d_nt}
+\end{bmatrix}
+\end{equation}
+this would make it easier to compute the matrix exponential.
+
+\begin{definition}[Properties of Matrix Exponentials]
+    The matrix exponential has several important properties:
+    \begin{itemize}
+        \item \( e^{A \cdot 0} = I \), where \( 0 \) is the zero matrix.
+        \item \( e^{A+B} = e^{A}e^{B} \) if \( A \) and \( B \) commute (i.e., \( AB = BA \)).
+        \item \( e^{A(t+s)} = e^{At}e^{As} \) for any scalars \( t \) and \( s \).
+        \item \( \frac{d}{dt}e^{At} = Ae^{At} = e^{At}A \).
+        \item If \( A \) is invertible, then \( e^{At} \) is also invertible, and its inverse is given by \( (e^{At})^{-1} = e^{-At} \).
+        \item The solution to the system of linear differential equations \( \mathbf{x}'(t) = A\mathbf{x}(t) \) with initial condition \( \mathbf{x}(0) = \mathbf{x}_0 \) is given by:
+        \begin{equation}
+        \mathbf{x}(t) = e^{At}\mathbf{x}_0
+        \end{equation}
+        \item The derivatives of the matrix exponential can be computed using the series expansion:
+        \begin{equation}
+        \frac{d^n}{dt^n}e^{At} = A^n e^{At}
+        \end{equation}
+    \end{itemize}
+\end{definition}
+
+\begin{example}[Schrödinger Equation]
+    Consider the time-dependent Schrödinger equation in quantum mechanics:
+    $$i\hbar \frac{\partial}{\partial t} | \Psi(t) \rangle = \hat{H} | \Psi(t) \rangle$$
+    where \( | \Psi(t) \rangle \) is the state vector (wavefunction) of the quantum system, \( \hat{H} \) is the Hamiltonian operator (which can be represented as a matrix in a finite-dimensional space for which \( \hat{H} = \hat{H}^\dagger \) (Conjugate-Transpose)), \( i \) is the imaginary unit, and \( \hbar \) is the reduced Planck constant.
+We claim that Hermitian matrices have real eigenvalues and orthogonal eigenvectors.
+We first rewrite the soltuion of the Schrödinger equation as:
+    $$
+        | \Psi(t) \rangle = \exp \left(-\frac{i}{\hbar} \hat{H} t \right) | \Psi(0) \rangle
+    $$
+    Let $\langle \phi(n) | = | \phi(n) \rangle^\dagger$, then we call $\langle \phi(n) | \Psi(t) \rangle$ the norm of the wavefunction in the direction of the eigenvector \( | \phi(n) \rangle \).
+We have:
+    \begin{align*}
+        \langle \phi(n) | \Psi(t) \rangle
+        &= \langle \phi(0) | \sum_{k=0}^{\infty} \frac{1}{k!} \left(-\frac{i}{\hbar} \hat{H} t \right)^k | \Psi(0) \rangle \\  
+        &= \langle \phi(0) | \Psi(0) \rangle
+    \end{align*}
+    Then, the solution lies on the same norm as the initial condition, in other words, the solution is in $S^{n-1}$ if the initial condition is in $S^{n-1}$.
+\end{example}
 

### 2025-10-08 YEAR2/PHY293_LectureNotes/main.tex
oct 7

--- 2025-09-26 YEAR2/PHY293_LectureNotes/main.tex
@@
 \subsection{Governing Equations of Harmonic Oscillators}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Simple Harmonic Oscillator: $m\ddot{x} + kx = 0$ or $\ddot{x} + \omega_0^2 x = 0$ where $\omega_0 = \sqrt{k/m}$
+    \item Damped: $m\ddot{x} + b\dot{x} + kx = 0$ or $\ddot{x} + \gamma\dot{x} + \omega_0^2 x = 0$ where $\gamma = b/m$
+    \item Driven: $m\ddot{x} + b\dot{x} + kx = F(t)$
+\end{itemize}
+\textbf{Tip:} Remember to check the damping parameter convention ($\gamma$ vs $2\beta$) when comparing formulas!
+\end{keybox}
+
 This subsection collects the baseline ODEs for simple, damped, and driven oscillators to set notation used later.
@@
     A driven harmonic oscillator is described by the following differential equation, which includes an external driving force $F(t)$:
-    \begin{equation}
+    $$
         m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = F(t)   
-    \end{equation}
+    $$
 \end{definition}
@@
 \subsection{Simple Harmonic Motion}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item General solution: $x(t) = A\cos(\omega t + \phi)$ where $\omega = \sqrt{k/m}$
+    \item Period: $T = 2\pi/\omega = 2\pi\sqrt{m/k}$
+    \item Frequency: $f = 1/T = \omega/(2\pi)$
+    \item Total Energy: $E = \frac{1}{2}kA^2$ (constant for undamped oscillator)
+\end{itemize}
+\textbf{Common Pitfall:} Don't forget to match initial conditions properly when solving for amplitude and phase!
+\end{keybox}
+
 We analyze the undamped solution forms, relate constants to initial conditions, and derive period/frequency relations.
@@
 The same as equation (\ref{eq:hooke}), which can be rewritten as:
-    \begin{equation}
+    $$
         F = m\ddot{x} = -kx
-    \end{equation}
+    $$
     By setting $\omega^2 = \frac{k}{m}$, a general solution can be written as:
-    \begin{equation}
+    $$
         x(t) = x_0 + A_1 \cos(\omega t) + A_2 \sin(\omega t)
-    \end{equation}
+    $$
     where $A_1$ and $A_2$ are constants determined by the IVP, $\omega$ is the angular frequency, and $\phi$ is the phase constant.
@@
     The period \( T \) is the time it takes for one complete cycle of the motion, given by:
-    \begin{equation}
+    $$
         T = 2\pi \sqrt{\frac{m}{k}}
-    \end{equation}
+    $$
     The frequency \( f \) is the number of cycles per unit time, given by:
-    \begin{equation}
+    $$
         f = \frac{1}{T} = \frac{1}{2\pi} \sqrt{\frac{k}{m}}
-    \end{equation}
+    $$
     The angular frequency \( \omega \) is related to the frequency by:
-    \begin{equation}
+    $$
         \omega = 2\pi f = \sqrt{\frac{k}{m}}
-    \end{equation}
+    $$
 
@@
 We can also express the solution in a more compact form using a single cosine function with a phase shift:
-\begin{equation}
+$$
     x(t) = A \cos(\omega t + \phi)
-\end{equation}
+$$
 where
-\begin{subequations}
-    \begin{align}
-        A &= \sqrt{A_1^2 + A_2^2}, \\
-        \phi &= \arctan\!\left(\frac{-A_2}{A_1}\right) 
-              = \arctan\!\left(\frac{-v(0)/\omega}{x(0)}\right).
-    \end{align}
-\end{subequations}
+$$
+        A = \sqrt{A_1^2 + A_2^2}, \qquad
+        \phi = \arctan\!\left(\frac{-A_2}{A_1}\right) 
+              = \arctan\!\left(\frac{-v(0)/\omega}{x(0)}\right)
+$$
 \end{theorem}
@@
     The total mechanical energy \(E\) of a simple harmonic oscillator is the sum of its kinetic energy \(K\) and potential energy \(U\). 
-    \begin{equation}
+    $$
         E = K + U
-    \end{equation}
+    $$
     First we consider the change of potential energy from a position $x_i$ to $x_f$, assuming the path is along the spring or the curve $C$ of the oscillator.
@@
 The change in potential energy can be simply parametized and calculated as follows:
-    \begin{equation}
+    $$
         \Delta U = \int_C F\cdot ds = -\int_{x_i}^{x_f} F\,dx = \int_{x_i}^{x_f} kx\,dx = \left[\frac{1}{2}kx^2\right]_{x_i}^{x_f} = \frac{1}{2}k(x_f^2 - x_i^2)
-    \end{equation}
+    $$
 
     Therefore, the potential energy \(U\) at a position \(x\) (taking the reference point at \(x=0\)) is given by:
-    \begin{equation}
+    $$
         U(x) = \frac{1}{2}kx^2
-    \end{equation}
+    $$
 
     The kinetic energy \(K\) of the oscillator is given by:
-    \begin{equation}
+    $$
         K = \frac{1}{2}m\dot{x}^2
-    \end{equation}
+    $$
     Therefore, the total mechanical energy \(E\) of the simple harmonic oscillator is:
-    \begin{equation}
+    \begin{equation}\label{eq:energy_sho}
         E = K + U = \frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2
@@
 \subsection{Damped Harmonic Motion}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Damped ODE: $\ddot{x} + \gamma\dot{x} + \omega_0^2 x = 0$ where $\gamma = b/m$
+    \item Underdamped ($\gamma < 2\omega_0$): $x(t) = A_0 e^{-\gamma t/2}\cos(\omega_r t + \phi)$ where $\omega_r = \sqrt{\omega_0^2 - \gamma^2/4}$
+    \item Critically damped ($\gamma = 2\omega_0$): $x(t) = (A_1 t + A_2)e^{-\gamma t/2}$
+    \item Overdamped ($\gamma > 2\omega_0$): Exponential decay without oscillation
+\end{itemize}
+\textbf{Concept:} Critical damping returns to equilibrium fastest without oscillating!
+\end{keybox}
+
 We solve the damped ODE, classify regimes (underdamped, critical, overdamped), and connect decay rates with parameters.
@@
 Which could be rewritten as:
-    \begin{equation}
+    $$
         \ddot{x} + \gamma \dot{x} + \omega_0^2 x = 0
-    \end{equation}
+    $$
     where \( \omega_0 = \sqrt{\frac{k}{m}} \) is the natural angular frequency of the undamped oscillator, and \( \gamma = \frac{b}{m} \) is the damping coefficient per unit mass.
@@
     To skip the math, lets assume a solution of the form \( x(t) = e^{i\omega t} \), substituting into the differential equation gives us a formulation for \( \omega \):
-    \begin{align}
-        -\omega^2 - i\gamma\omega + \omega_0^2 &= 0 \nonumber \\
-        \omega &= -i\frac{\gamma}{2} \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} \label{eq:omega_damped}
-    \end{align}
+    \begin{equation}\label{eq:omega_damped}
+        \omega = -i\frac{\gamma}{2} \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}}
+    \end{equation}
     Also, we can characterize the real and imaginary parts of \( \omega \) as:
+    $$
+        \omega_r = \mathrm{Re}(\omega) = \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}}, \qquad
+        \omega_i = \mathrm{Im}(\omega) = -\frac{\gamma}{2}
+    $$
+    The general solution for the damped harmonic oscillator can be written as:
+    $$
+        x(t) = \exp(\omega_i t) \exp(-i \omega_r t) = \exp\left(-\frac{\gamma}{2}t\right) \exp(\mp i \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} t)
+    $$
     \begin{subequations}
-        \begin{equation}
-            \omega_r = \mathrm{Re}(\omega) = \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}}
-        \end{equation}
-        \begin{equation}
-            \omega_i = \mathrm{Im}(\omega) = -\frac{\gamma}{2}
-        \end{equation}
-    \end{subequations}
-    The general solution for the damped harmonic oscillator can be written as:
-    \begin{equation}
-        x(t) = \exp(\omega_i t) \exp(-i \omega_r t) = \exp\left(-\frac{\gamma}{2}t\right) \exp(\mp i \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} t)
-    \end{equation}
-    \begin{subequations}
     \begin{itemize}
         \item \textbf{No Damping} (\( \gamma = 0 \)): The system behaves like a simple harmonic oscillator with angular frequency \( \omega_0 \). Given by:
-        \begin{equation}
+        $$
             z = \exp(-i\omega_0 t)
-        \end{equation}
+        $$
         \item \textbf{Underdamping} (\( 0 < \gamma < 2\omega_0 \)): The system oscillates with a gradually decreasing amplitude.
@@
 Given by:
-        \begin{equation}
+        $$
             z = \exp\left(-\frac{\gamma}{2}t\right) \exp(-i \omega_r t)
-        \end{equation}
+        $$
         The trigonometric form of the solution is:
-        \begin{equation}
+        $$
             x(t) = A_0 \exp\left(-\frac{\gamma}{2}t\right) \cos(\omega_r t + \phi)
-        \end{equation}
+        $$
         where \( A_0 \) and \( \phi \) are constants determined by the initial conditions
@@
         Therefore, the general solution for the critically damped case is:
-        \begin{equation}
+        $$
             x(t) = (A_1 t + A_2) \exp\left(-\frac{\gamma}{2}t\right)
-        \end{equation}
+        $$
         where \( A_1 \) and \( A_2 \) are constants determined by
         \item \textbf{Overdamping} (\( \gamma > 2\omega_0 \)): The system returns to equilibrium without oscillating, but more slowly than in the critically damped case. The solution is given by:
-        \begin{equation}
+        $$
             z = \exp\left(-\frac{\gamma}{2}t\right) \exp\left(\sqrt{\frac{\gamma^2}{4} - \omega_0^2} t\right)
-        \end{equation}
+        $$
         So the general solution is (the solution is via a substitution of $x(t) = e^{-\gamma t/2} f(t)$ into the differential equation, which resolves the ODE to a simple form):
-        \begin{equation}
+        $$
             x(t) = A_1 \exp\left[\left(-\frac{\gamma}{2} + \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right] + A_2 \exp\left[\left(-\frac{\gamma}{2} - \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right]
-        \end{equation}
+        $$
         where \( A_1 \) and \( A_2 \) are constants determined by the initial conditions.
@@
 \subsection{Energy and Quality Factor}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Energy decay: $E(t) = E_0 e^{-\gamma t}$ for light damping
+    \item Time constant: $\tau = 1/\gamma$ (time for energy to drop to $1/e$)
+    \item Quality factor: $Q = \omega_0/\gamma = \omega_0\tau$ (higher $Q$ = less damping)
+    \item Energy loss per cycle: $\frac{\Delta E}{E} \approx -\frac{2\pi}{Q}$
+\end{itemize}
+\textbf{Concept:} $Q$ measures how many oscillations occur before energy decays significantly!
+\end{keybox}
+
 We study how energy decays under light damping, define the time constant and quality factor Q, and relate them to response.
@@
     We can also define the time constant \( \tau \) as the time it takes for the energy to decrease to \( \frac{1}{e} \) of its initial value:
-    \begin{equation}
+    $$
         \tau = \frac{1}{\gamma}
-    \end{equation}
+    $$
     So we have, for very light damping:
-    \begin{equation}
+    \begin{equation}\label{eq:energy_decay}
         E(t) = E_0 e^{-\gamma t} = E_0 e^{-\frac{t}{\tau}}
@@
 However, for a damped harmonic oscillator, \( ma + kx = -b\dot{x} \), leading to:
-    \begin{equation}
+    $$
         \frac{dE}{dt} = -b\dot{x}^2
-    \end{equation}
+    $$
 \end{definition}
@@
     The quality factor \( Q \) is a dimensionless parameter that characterizes the damping of a harmonic oscillator. It is defined as:
-    \begin{subequations}
-    \begin{equation}
+    \begin{equation}\label{eq:Q_factor}
         Q = \frac{\omega}{\gamma} = \omega \tau
@@
     And for very light damping, we can approximate \( \omega \approx \omega_0 \), leading to:
-    \begin{equation}
+    $$
         Q \approx \frac{\omega_0}{\gamma} = \omega_0 \tau
-    \end{equation}
-    \end{subequations}
+    $$
     This allows us to rewrite the equation of a damped harmonic oscillator as:
-    \begin{equation}
+    $$
         \ddot{x} + \frac{\omega_0}{Q} \dot{x} + \omega_0^2 x = 0
-    \end{equation}
+    $$
     and:
-    \begin{equation}
+    $$
         \omega = \omega_0 \sqrt{1 - \frac{1}{4Q^2}}
-    \end{equation}
+    $$
 
@@
 \subsection{Undamped Forced Oscillations}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Driven ODE: $m\ddot{x} + kx = F_0\cos(\omega t)$
+    \item Steady-state solution: $x(t) = A(\omega)\cos(\omega t - \delta)$
+    \item Amplitude: $A(\omega) = \frac{\omega_0^2 a}{|\omega_0^2 - \omega^2|}$ where $a = F_0/k$
+\end{itemize}
+\textbf{Common Pitfall:} At resonance $(\omega \approx \omega_0)$, amplitude becomes infinite without damping!
+\end{keybox}
+
 We examine steady-state response to sinusoidal driving without losses and identify resonance behavior.
@@
     Consider a driver force acting on the mass-spring system; the ODE becomes:
-    \begin{equation}
+    $$
         m\ddot{x} + kx = \eta
-    \end{equation}
+    $$
     where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The steady-state particular solution is:
-    \begin{equation}
+    $$
         x(t) = A(\omega)\cos(\omega t - \delta) 
-    \end{equation}
+    $$
     To derive $A(\omega)$ and $\delta$, use the equation of motion:
@@
     Now, consider the case when $\delta = 0$, we have:
-    \begin{equation}
+    $$
         A(\omega) = \frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
-    \end{equation}
+    $$
     and the case where \( \delta = \pi \), we have:
-    \begin{equation}
+    $$
         A(\omega) = -\frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
-    \end{equation}
+    $$
     Observe, when $\omega \approx \omega_0$, the amplitude \( A(\omega) \) becomes very large, indicating resonance.
@@
 \subsection{Damped Forced Oscillations}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Amplitude: $A(\omega) = \frac{\omega_0^2 a}{\sqrt{\omega^2\gamma^2 + (\omega^2 - \omega_0^2)^2}}$
+    \item Resonance amplitude: $A_{\text{max}} = Qa$ where $Q = \omega_0/\gamma$
+    \item Phase shift: $\tan(\delta) = \frac{\omega\gamma}{\omega_0^2 - \omega^2}$
+    \item FWHM: $\Delta\omega = \gamma = \omega_0/Q$
+\end{itemize}
+\textbf{Concept:} Damping prevents infinite amplitude at resonance and causes phase lag!
+\end{keybox}
+
 We add damping to the driven case, derive the frequency response amplitude and phase, and study bandwidth and Q.
@@
     Consider a damped harmonic oscillator subjected to an external driving force. It is described by the following differential equation:
-    \begin{equation}
+    $$
         m\ddot{x} + b\dot{x} + kx = F_0 \cos(\omega t)
-    \end{equation}
+    $$
     where \( F_0 \) is the amplitude of the driving force, \( \omega \) is the angular frequency of the driving force, \( b \) is the damping coefficient, \( m \) is the mass, and \( k \) is the spring constant.
 The general solution to this non-homogeneous differential equation is given by:
-    \begin{equation}
+    $$
         x(t) = A(\omega)\cos(\omega t - \delta)
-    \end{equation}
+    $$
     and the amplitude \( A(\omega) \) is given by:
-    \begin{equation}
+    \begin{equation}\label{eq:damped_amplitude}
         A(\omega) = \frac{\omega_0^2 a}{\sqrt{\omega^2 \gamma^2 + (\omega^2 - \omega_0^2)^2}}
@@
         \item \textbf{Resonance} (\( \omega \approx \omega_0 \)): At resonance, the amplitude \( A(\omega) \) reaches its maximum value but it does not become infinite due to the presence of damping: 
-        \begin{equation}
+        $$
             A_{\text{max}} = \frac{\omega_0^2 a}{\omega_0 \gamma} = \frac{\omega_0 a}{\gamma} = Q a
-        \end{equation}
+        $$
     \end{enumerate}
     We also have the phase shift \( \delta \) given by:
-    \begin{equation}
+    $$
         \tan(\delta) = \frac{\omega \gamma}{\omega_0^2 - \omega^2}
-    \end{equation}
+    $$
     The phase shift \( \delta \) indicates how much the oscillation of the mass lags behind the driving force. The behavior of \( \delta \) can be summarized as follows:
@@
     We can calculate the velocity of the oscillator:
-    \begin{equation}
+    $$
         \dot{x}(t) = -\omega A(\omega) \sin(\omega t - \delta) = -v_0 \sin(\omega t - \delta)
-    \end{equation}
+    $$
     where \( v_0 = \omega A(\omega) \) is the maximum speed of the oscillator. Energy is lost at the following rate:
-    \begin{equation}
+    $$
         P(t) = b v(t)^2 
-    \end{equation}
+    $$
     Substituteing the expression for \( v(t) \) into the power equation gives:
@@
     The average power \( \langle P \rangle \) over one complete cycle of the driving force is given by:
-   \begin{equation}
+   \begin{equation}\label{eq:avg_power}
         \bar{P}(\omega) = \frac{b v_0^2}{2} = \frac{\omega^2 F_0^2 \gamma}{2m \left[(\omega^2 - \omega_0^2)^2 + \omega^2 \gamma^2\right]}
@@
     If the driving frequency \( \omega \) is close to \( \omega_0 \), use $\omega^2 - \omega_0^2 \approx 2\omega_0\Delta \omega$ with $\Delta \omega = \omega - \omega_0$ to get the Lorentzian form:
-    \begin{equation}
+    $$
         \bar{P}(\omega) \approx \frac{\omega_0 F_0^2}{2m \gamma \Big[ 1 + \big(\tfrac{2\Delta \omega}{\gamma}\big)^2 \Big]}
-    \end{equation}
+    $$
     The maximum average power occurs at resonance:
-    \begin{equation}
+    $$
         \bar{P}_{\text{max}} = \frac{\omega_0 F_0^2}{2m \gamma}
-    \end{equation}
+    $$
 \end{definition}
@@
 The FWHM is given by:
-\begin{equation}
+$$
     \omega_\text{FWHM} = 2\Delta \omega = 2\gamma = \frac{2\omega_0}{Q}
-\end{equation}
+$$
 To illustrate the graph, we can plot \( \bar{P}(\omega) \) versus \( \omega \) for different values of the quality factor \( Q \):
@@
 \subsection{Simple Pendulum}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Small angle approximation: $\ddot{\theta} + \frac{g}{l}\theta = 0$ (valid for $\theta \ll 1$ rad)
+    \item Period: $T = 2\pi\sqrt{l/g}$ (independent of mass!)
+    \item Physical pendulum: $\omega = \sqrt{\frac{mgd}{I}}$ where $d$ is distance to CM
+\end{itemize}
+\textbf{Tip:} LC circuits are analogous to oscillators: $L \leftrightarrow m$, $1/C \leftrightarrow k$, $R \leftrightarrow b$
+\end{keybox}
+
 Consider the arc length \( s \) of a pendulum bob, we have:
-\begin{equation}
+$$
     s = l\theta \quad m\ddot{s} = -mg\sin(\theta)
-\end{equation}
+$$
 
 For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
-\begin{equation}
+$$
     \ddot{\theta} + \frac{g}{l}\theta = 0
-\end{equation}
+$$
 And the ODE has solution:
-\begin{equation}
+$$
     \theta(t) = \theta_0 \cos(\omega t + \phi)
-\end{equation}
+$$
 
 \paragraph{Energy} For small angle $\theta$, the total mechanical energy \( E \) of the pendulum is given by:
-\begin{equation}
+$$
     E = K + U = \frac{1}{2}ml^2\dot{\theta}^2 + mgl(1 - \cos(\theta)) \approx \frac{1}{2}ml^2\dot{\theta}^2 + \frac{1}{2}mgl\theta^2
-\end{equation}
+$$
 This shows that the pendulum behaves like a simple harmonic oscillator with an effective spring constant \( k = \frac{mg}{l} \).
@@
 \paragraph{The physical pendulum} For a rigid body swinging about a pivot point, the equation of motion is given by:
-\begin{equation}
+$$
     I\ddot{\theta} = \tau 
-\end{equation}
+$$
 For simple rod of length \( L \) and mass \( m \) pivoted at one end, the moment of inertia about the pivot point is \( I = \frac{1}{3}mL^2 \).
@@
 For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
-\begin{equation}
+$$
     \ddot{\theta} + \frac{3g}{2L}\theta = 0
-\end{equation}
+$$
 And the small angle approximates leads to:
-\begin{equation}
+$$
     \ddot{\theta} + \omega^2 \theta = 0
-\end{equation}
+$$
 where \( \omega = \sqrt{\frac{3g}{2L}} \). The solution to this ODE is:
-\begin{equation}
+$$
     \theta(t) = \theta_0 \cos(\omega t + \phi)
-\end{equation}
+$$
 The angular frequency of oscillation for the physical pendulum is:
-\begin{equation}
+$$
     \omega = \sqrt{\frac{3g}{2L}}
-\end{equation}
+$$
 
 \paragraph{The LC Circuit} An LC circuit consists of an inductor \( L \) and a capacitor \( C \) connected in series. The charge \( q(t) \) on the capacitor satisfies the differential equation:
-\begin{equation}
+$$
     L\frac{d^2q}{dt^2} + \frac{1}{C}q = 0
-\end{equation}
+$$
 This is analogous to the equation of motion for a simple harmonic oscillator, with the angular frequency given by:
-\begin{equation}
+$$
     \omega = \frac{1}{\sqrt{LC}}
-\end{equation}
+$$
 
@@
 The charge \( q(t) \) on the capacitor satisfies the differential equation:
-\begin{equation}
+$$
     L\frac{d^2q}{dt^2} + R\frac{dq}{dt} + \frac{1}{C}q = 0
-\end{equation}
+$$
 This is analogous to the resistor behaving as a damped harmonic oscillator, with the angular frequency given by:
-\begin{equation}
+$$
     \omega = \sqrt{\frac{1}{LC} - \left(\frac{R}{2L}\right)^2}
-\end{equation}
+$$
 The following are equations that summarize the analogies between mechanical and electrical oscillators:
@@
 \subsection{Coupled Oscillators}
+
+\begin{keybox}
+\textbf{Key Concepts:}
+\begin{itemize}
+    \item Normal coordinates $q_1 = x_1 + x_2$, $q_2 = x_1 - x_2$ decouple the equations
+    \item Each normal mode has its own frequency: $\omega_1$ (in-phase), $\omega_2$ (out-of-phase)
+    \item General motion = superposition of normal modes
+    \item Energy doesn't flow between modes (they're independent)
+\end{itemize}
+\textbf{Tip:} Beat frequency $= (\omega_2 - \omega_1)/2$ describes energy transfer between oscillators!
+\end{keybox}
 \begin{definition}[Coupled Oscillators]
@@
 
-    \begin{equation}
+    $$
         \begin{cases}
@@
         \end{cases}
-    \end{equation}
+    $$
     where \( x_1 \) and \( x_2 \) are the displacements of masses \( m_1 \) and \( m_2 \) from their equilibrium positions, respectively. Now if we add and subtract the two equations, we get:
@@
 The general solutions for \( q_1(t) \) and \( q_2(t) \) are:
-    \begin{equation}
+    $$
         \begin{cases}
@@
         \end{cases}
-    \end{equation}
+    $$
     where \( A_1, A_2, \phi_1, \) and \( \phi_2 \) are constants determined by the initial conditions.
 Finally, we can expressions for \( x_1(t) \) and \( x_2(t) \) in terms of \( q_1(t) \) and \( q_2(t) \):
-    \begin{equation}
+    $$
         \begin{cases}
@@
         \end{cases}
-    \end{equation}
+    $$
     All parts of the system oscillate with the same normal frequency (eigenvalue) in a normal mode (eigenvector).
@@
 \paragraph{Energy in Coupled Oscillators} The total mechanical energy \( E \) of the coupled oscillator system is given by:
-\begin{equation}
+$$
     E = \frac{1}{2}m_1\dot{x}_1^2 + \frac{1}{2}m_2\dot{x}_2^2 + \frac{1}{2}k(x_1 - x_2)^2 + \frac{1}{2}m_1\frac{g}{L}x_1^2 + \frac{1}{2}m_2\frac{g}{L}x_2^2
-\end{equation}
+$$
 Substituting the expressions for \( x_1 = \frac{q_1 + q_2}{2} \) and \( x_2 = \frac{q_1 - q_2}{2} \) into the energy equation, two independent energies terms for $q_1$ and $q_2$ emerge:
-\begin{equation}
+$$
     E = \left[ \frac{1}{4} m \dot{q}_1^2 + \frac{1}{4}\frac{m_1 g}{L} q_1^2 \right] + \left[ \frac{1}{4} m \dot{q}_2^2 + \frac{1}{4}\frac{m_2 g}{L} q_2^2 \right]
-\end{equation}
+$$
 This shows that the total energy is the sum of the energies associated with each normal mode, and that no cross term between the modes exists. Thus energy does not flow from one mode to another.
@@
 \subsection{Normal Modes}
+
+\begin{keybox}
+\textbf{Key Concepts:}
+\begin{itemize}
+    \item Normal mode = pattern where all parts oscillate at same frequency
+    \item Find frequencies by solving eigenvalue problem: det$(K - \omega^2 M) = 0$
+    \item Number of normal modes = number of degrees of freedom
+    \item Eigenvectors give amplitude ratios for each mode
+\end{itemize}
+\textbf{Tip:} For symmetric systems, look for in-phase and out-of-phase modes!
+\end{keybox}
 \paragraph{Oscilating Masses Connected by Springs} Consider two masses \( m_1 = m_2 = m \) connected by three spring with spring constant \( k \) and fixed to walls on either side.
 The equations of motion for the two masses are given by:
-\begin{equation}
+$$
     \begin{cases}
@@
     \end{cases}
-\end{equation}
+$$
 To solve these equations, we assume solutions of the form:
@@
 So we can derive two normal mode frequencies:
-\begin{equation}
+$$
     \begin{cases}
@@
     \end{cases}
-\end{equation}
+$$
 
 \paragraph{Solving as an Eigenproblem} Alternatively, we can create a system of equations from \eqref{eq:coupled_springs}:
-\begin{equation}
+$$
     \begin{bmatrix}
@@
     \end{bmatrix}
-\end{equation}
+$$
 To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
@@
 This gives us the two normal mode frequencies:
-\begin{equation}
+$$
     \begin{cases}
@@
     \end{cases}
-\end{equation}
+$$
 
@@
 
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Wave function: $y(x,t) = A\cos(kx - \omega t + \phi_0)$ (travelling right)
+    \item Wave speed: $v = \omega/k = f\lambda$
+    \item Wave number: $k = 2\pi/\lambda$, Angular frequency: $\omega = 2\pi f$
+    \item String wave speed: $v = \sqrt{\tau/\mu}$ where $\tau$ is tension, $\mu$ is linear density
+\end{itemize}
+\textbf{Common Pitfall:} Don't confuse particle velocity $(\partial y/\partial t)$ with wave velocity $v$!
+\end{keybox}
+
 \begin{definition}[Travelling Pulse]
@@
     A sinusoidal wave is a wave that can be described by a sine or cosine function. The general form of a sinusoidal wave travelling in the positive x-direction is:
-    \begin{equation}
+    $$
         y(x,t) = A \cos(kx \pm \omega t + \phi_0)
-    \end{equation}
+    $$
     where:
@@
     So we can write the wave function as a form of \eqref{eq:travelling_wave}:
-    \begin{equation}
+    $$
         y(x,t) = A \cos\left[k(x \pm vt) + \phi_0\right]
-    \end{equation}
+    $$
     The wave speed \( v \) is related to the angular frequency and wave number by:
-    \begin{equation}
+    \begin{equation}\label{eq:wave_speed}
         v = \frac{\omega}{k} = f\lambda
@@
     For a fixed particle, we have:
-    \begin{equation}
+    $$
         v_y = \frac{\partial y}{\partial t} = A \omega \cos(kx \pm \omega t + \phi_0)
-    \end{equation}
+    $$
 \end{definition}
@@
 \end{example}
-
-\subsection{The Wave Equation}
-
 \begin{definition}[The Wave Equation]
-    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water). In one dimension, it is:
+    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water) through a medium. In one dimension, it is given by:
+    \begin{equation}\label{eq:wave_equation}
+        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
+    \end{equation}
+    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
+
+    Or in general:
+    $$
+        \frac{\partial^2 u}{\partial t^2} - c^2 \nabla^2 u = 0
+    $$
+    where \( \nabla^2 \) is the Laplacian operator.
+\end{definition}
+
+\begin{definition}[Energy of a String]
+    The kinetic energy of a string segment of mass \( \text{d}m = \mu \text{d}x \) is given by:
+    $$
+        \text{d}K = \frac{1}{2} \text{d}m v_y^2 = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2
+    $$
+    The potential energy stored in the string due to its tension is given by:
+    $$
+        \text{d}U = \frac{1}{2} \tau (\delta s - \delta x) 
+    $$
+    where \( \delta s \) is the actual length of the string segment and \( \delta x \) is the horizontal length. For small displacements, we can approximate:
+    $$
+        \delta s \approx \delta x \left(1 + \frac{1}{2}\left(\frac{\partial y}{\partial x}\right)^2\right)
+    $$
+    Thus, the potential energy becomes:
+    $$
+        \text{d}U \approx \frac{1}{2} \tau \delta x \left(\frac{\partial y}{\partial x}\right)^2
+    $$
+    The total energy \( E \) of the string segment is the sum of its kinetic and potential energies:
+    $$
+        \text{d}E = \frac{1}{2} \mu \text{d}x \left(\frac{\partial y}{\partial t}\right)^2 + \frac{1}{2} \tau \text{d}x \left(\frac{\partial y}{\partial x}\right)^2
+    $$
+    For a sinusoidal wave \( y(x,t) = A \cos(kx - \omega t) \), the change in energy is:
+    $$
+        \text{d}E = \frac{1}{2} \mu A^2 \omega^2 \sin^2(kx - \omega t) \text{d}x + \frac{1}{2} \tau A^2 k^2 \sin^2(kx - \omega t) \text{d}x
+    $$
+    Integrating over one wavelength \( \lambda \), we find the total energy in one wavelength:
+    $$
+        E = \int_0^\lambda \text{d}E = \frac{1}{4} \mu A^2 \omega^2 \lambda + \frac{1}{4} \tau A^2 k^2 \lambda
+    $$
+
+\end{definition}
+
+\begin{definition}[Power Transmitted by a String]
+   The power carried by a transverse wave could be calculated from $P=Fv$. The transverse force on the string is given by:
+    $$
+        P(x,t) = A^2 \mu v \omega^2 \sin^2(kx - \omega t + \phi_0)
+    $$
+    The average power is then given by:
+    $$
+        \bar{P} = \frac{1}{2} A^2 \mu v \omega^2 = \frac{1}{2} \mu A^2 \omega^2 \frac{\lambda}{T}
+    $$
+    And the maximum power is given by:
+    $$
+        P_{max} = A^2 \mu v \omega^2
+    $$
+\end{definition}
+
+\paragraph{Speed of Sound} The speed of sound waves through various meida is as follows:
+\begin{itemize}
+    \item In ideal gas, it is given by:
+    $$
+        v = \sqrt{\frac{\gamma RT}{m}} = \sqrt{\frac{\gamma k_B T}{m}}
+    $$
+    where \( \gamma \) is the adiabatic index, \( R \) is the universal gas constant, \( k_B \) is the Boltzmann constant, \( T \) is the absolute temperature, and \( m \) is the molar mass of the gas.
+$\gamma$ varies based on the type of gas:
+    \begin{itemize}
+        \item Monatomic gases (e.g., helium, neon): \( \gamma = \frac{5}{3} \)
+        \item Diatomic gases (e.g., nitrogen, oxygen): \( \gamma = \frac{7}{5} \)
+        \item Polyatomic gases (e.g., carbon dioxide, methane): \( \gamma \) is typically between 1.3 and 1.4
+    \end{itemize}
+    \item In a liquid, it is given by:
+    $$
+        v = \sqrt{\frac{B_{\text{adiabatic}}}{m}} = \sqrt{\frac{B}{\rho}}
+    $$
+    where \(B_{\text{adiabatic}}\) is the adiabatic bulk modulus, \( B \) is the bulk modulus, \( \rho \) is the density of the liquid and \( m \) is the molar mass of the liquid.
+    \item In a solid, it is given by:
+    $$
+        v = \sqrt{\frac{E}{\rho}} 
+    $$
+    where \( E \) is the Young's modulus and \( \rho \) is the density of the solid. It could also be given by:
+    $$
+        v = \sqrt{\frac{B}{\rho}}
+    $$
+    where \( B \) is the bulk modulus.
+\end{itemize}
+
+\begin{definition}[Impedance]
+    Similar to electrical impedance, we also have acoustic impedance and mechanical, they measure how much a structure recisits motion when subjected to a harmonic force. They are defined as follows:
+    \begin{itemize}
+        \item Electical impedance \( Z_e \):
+        $$
+            Z_e = \sqrt{R^2 + \left(\omega L - \frac{1}{\omega C}\right)^2}
+        $$
+        \item Acoustic impedance \( Z_a \), it is the ratio of acoustic pressure to particle velocity:
+        $$
+            Z_a = \frac{p}{v} = \rho v = \sqrt{B\rho}
+        $$
+        \item Mechanical impedance \( Z_m \):
+        $$
+            Z_m = v\mu = \sqrt{\tau\mu}
+        $$
+    \end{itemize}
+\end{definition}
+
+\paragraph{Wave boundry Conditions} When a wave encounters a boundary between two different media, part of the wave is reflected back into the original medium, and part is transmitted into the new medium.
+This is constrined by the following boundary conditions:
+\begin{itemize}
+    \item The displacement of the wave must be continuous across the boundary:
+    $$
+        y_1(x_b, t) = y_2(x_b, t) = y_3(x_b, t)
+    $$
+    \item The derivative of the displacement with respect to position must also be continuous across the boundary:
+    $$
+        \frac{\partial y_1}{\partial x}\bigg|_{x_b} = \frac{\partial y_2}{\partial x}\bigg|_{x_b} = \frac{\partial y_3}{\partial x}\bigg|_{x_b}
+    $$
+\end{itemize}
+and then from these conditions, we can derive the amplitudes of the reflected and transmitted waves:
+\begin{equation}\label{eq:amp-continuity}
+    A_i = A_r + A_t
+\end{equation}
+and 
+$$
+    A_i Z_1 = A_r Z_1 + A_t Z_2
+$$
+From these two equations, we can solve for the reflection and transmission coefficients:
+\begin{equation}\label{eq:reflection_coeff}
+    R = \frac{A_r}{A_i} = \frac{Z_2 - Z_1}{Z_2 + Z_1}
+\end{equation}
+and
+\begin{equation}\label{eq:transmission_coeff}
+    T = \frac{A_t}{A_i} = \frac{2Z_2}{Z_2 + Z_1}
+\end{equation}
+
+\paragraph{Special Cases} There are two special cases:
+\begin{itemize}
+    \item \textbf{Fixed End}: If the wave encounters a boundary where the medium is fixed (e.g., a string tied to a wall), the reflected wave undergoes a phase change of \( \pi \) (inversion).
+The reflection coefficient is \( R = -1 \) and the transmission coefficient is \( T = 0 \).
+    \item \textbf{Free End}: If the wave encounters a boundary where the medium is free to move (e.g., a string attached to a ring that can slide along a rod), the reflected wave does not undergo a phase change.
+The reflection coefficient is \( R = 1 \) and the transmission coefficient is \( T = 2 \).
+\end{itemize}
+
+\begin{example}[Transmission of 64\% Energy]
+Suppose a wave transmits \(64\%\) of its incident energy across a boundary. Then:
+$$
+    T_e = 0.64, 
+    \quad R_e = 1 - T_e = 0.36
+$$
+so that the amplitude reflection coefficient is
+$$
+    R = +\sqrt{R_e} = 0.6 
+$$
+\textbf{Important} the positive sign indicates the reflected wave is \emph{in phase} with the incident wave.
+
+From the boundary condition in Eq.~\eqref{eq:amp-continuity},
+$$
+    T = 1 + R = 1 + 0.6 = 1.6
+$$
+Thus, the amplitude transmission coefficient is \(T = 1.6\), even though only \(64\%\) of the energy is transmitted.
+This apparent paradox arises because the transmitted energy depends on both the amplitude squared and the ratio of impedances \(\tfrac{Z_2}{Z_1}\).
+\end{example}
+\subsection{Superposition and Standing Waves}
+
+\begin{keybox}
+\textbf{Key Equations:}
+\begin{itemize}
+    \item Standing wave: $y(x,t) = 2A\sin(kx)\cos(\omega t)$ (for fixed-fixed ends)
+    \item Allowed wavelengths: $\lambda_n = 2L/n$ where $n = 1,2,3,...$
+    \item Frequencies: $f_n = nv/(2L) = nf_1$ (harmonics)
+    \item Constructive interference: path difference $= m\lambda$
+    \item Destructive interference: path difference $= (m+1/2)\lambda$
+\end{itemize}
+\textbf{Tip:} Nodes are spaced $\lambda/2$ apart; antinodes too!
+\end{keybox}
+\begin{definition}[Modulation of two waves]
+    Consider two monochromatic waves:
+    $$
+    \begin{cases}
+        y_1(x,t) &= A \cos(k_1 x - \omega_1 t) \\
+        y_2(x,t) &= A \cos(k_2 x - \omega_2 t)
+    \end{cases}
+    $$
+    Then, the resultant wave $y_1 + y_2$ can be expressed as:
+    $$
+        y(x,t) = 2A \cos\left(\frac{(k_1 - k_2)x - (\omega_1 - \omega_2)t}{2}\right) \cos\left(\frac{(k_1 + k_2)x - (\omega_1 + \omega_2)t}{2}\right)
+    $$
+\end{definition}
+
+\begin{definition}[Coherence]
+    Two waves are said to be coherent if they are monochromatic and maintain a constant phase relationship over time.
+This means that the difference in phase between the two waves does not change as they propagate:
+    $$
+        \Delta \phi = \phi_1 - \phi_2 = \text{constant}
+    $$
+    Coherent waves can produce stable interference patterns, such as constructive and destructive interference.
+\end{definition}
+
+\begin{definition}[Phase and Phase Difference]
+    Phase of the wave is an argument $kx \pm \omega t + \phi_0$ of the wave function. The phase difference between two waves at a given point in space and time is given by:
+    $$
+        \Delta \phi = (k_1 x - \omega_1 t + \phi_1) - (k_2 x - \omega_2 t + \phi_2)
+    $$
+    where \( k_1, k_2 \) are the wave numbers, \( \omega_1, \omega_2 \) are the angular frequencies, and \( \phi_1, \phi_2 \) are the initial phases of the two waves.
+
+    If we consider two waves of the same frequency and wave number, the phase difference simplifies to:
+    $$
+        \Delta \phi = \Delta\phi + k\Delta x + \omega \Delta t
+    $$
+
+\end{definition}
+\begin{definition}[Waves in Two and Three Dimensions]
+    In two or three dimensions, waves can propagate in various directions. The general form of a wave in two dimensions is:
+    $$
+        z(x,y,t) = A \cos(k_x x + k_y y - \omega t + \phi_0)
+    $$
+    where \( k_x \) and \( k_y \) are the wave numbers in the \( x \) and \( y \) directions, respectively, and \( \phi_0 \) is the initial phase. We also define $v$ and $k$ as:
+    $$
+        v = \frac{\omega}{k}, \quad k = \sqrt{k_x^2 + k_y^2}
+    $$
+\end{definition}
+
+\begin{definition}[Spherical Waves]
+    A spherical wave is a wave that propagates outward in all directions from a point source. The general form of a spherical wave is:
+    $$
+        y(r,t) = \frac{A}{r} \cos(kr - \omega t + \phi_0)
+    $$
+    where \( r = \sqrt{x^2 + y^2 + z^2} \) is the radial distance from the source, \( A \) is the amplitude at a reference distance, \( k \) is the wave number, \( \omega \) is the angular frequency, and \( \phi_0 \) is the initial phase.
+The amplitude decreases with distance as \( 1/r \) due to the spreading of the wavefront.
+
+\end{definition}
+
+\begin{definition}[Interference]
+    Consider two coherent in phase sources $S_1$ and $S_2$ separated by a distance \( d \). The path difference \( \Delta r \) between the two waves at a point \( P \) is given by:
+    $$
+        \Delta r = r_2 - r_1
+    $$
+    where \( r_1 \) and \( r_2 \) are the distances from the sources \( S_1 \) and \( S_2 \) to the point \( P \), respectively.
+
+    We have the following two cases for constructive and destructive interference:
+    \begin{itemize}
+        \item \textbf{Constructive Interference}: Occurs when the path difference is an integer multiple of the wavelength:
+        $$
+            \Delta r = m \lambda, \quad m = 0, 1, 2, \ldots
+        $$
+        \item \textbf{Destructive Interference}: Occurs when the path difference is an odd multiple of half the wavelength:
+        $$
+            \Delta r = \left(m + \frac{1}{2}\right) \lambda, \quad m = 0, 1, 2, \ldots
+        $$
+    \end{itemize}
+\end{definition}
+
+\paragraph{Nodes and Antinodes} In a standing wave, nodes are points where the medium does not move (zero displacement), while antinodes are points where the medium experiences maximum displacement.
+The distance between two consecutive nodes or antinodes is half the wavelength (\( \frac{\lambda}{2} \)).
+Nodes occur at positions where \( kx = n\pi \) (for integer \( n \)), and antinodes occur at positions where \( kx = (n + \frac{1}{2})\pi \).
+\begin{definition}[Nodes]
+    Nodes are points along a standing wave where the wave has zero amplitude. At these points, destructive interference occurs, and the medium remains at rest. Nodes occur at:
+    $$
+        x_n = n \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+    $$
+\end{definition}
+\begin{definition}[Antinodes]
+    Antinodes are points along a standing wave where the wave has maximum amplitude.
+At these points, constructive interference occurs, and the medium oscillates with the greatest displacement.
+Antinodes occur at:
+    $$
+        x_a = \left(n + \frac{1}{2}\right) \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+    $$
+\end{definition}
+
+\paragraph{Antinodal Curves} In two-dimensional standing waves, antinodal curves are the loci of points where the amplitude of the wave is maximum.
+These curves are formed by the constructive interference of waves from multiple sources.
+The shape and spacing of antinodal curves depend on the geometry of the wave sources and the wavelength of the waves.
+An antindal curve could be visuallized as following:
+\begin{figure}[h!]
+    \centering
+    % tikz, one source at (0, -3), another at (0, 3), cetner line at y = 0
+    \resizebox{0.4\textwidth}{!}{
+
+    \begin{tikzpicture}
+        % Draw the sources
+        \filldraw[black] (0, -3) circle (2pt) node[below] {$S_1$};
+        \filldraw[black] (0, 3) circle (2pt) node[above] {$S_2$};   
+        % Draw the center line
+        \draw[thick] (-5, 0) -- (5, 0) node[right] {};
+        % Draw the antinodal curves as circles radiating from the sources
+        \foreach \r in {1, 2, 3, 4} {
+            \draw[blue, thick] (0, -3) circle (\r);
+            \draw[blue, thick] (0, 3) circle (\r);
+        }
+        % Label the antinodal curves
+        \node at (3, -3) [right] {Antinodal Curves};
+        \node at (3, 3) [right] {Antinodal Curves};
+    \end{tikzpicture}
+    }
+    \caption{Antinodal Curves}
+    \label{fig:antinodal_curves}
+\end{figure}
+
+\begin{definition}[Standing Waves]
+    Standing waves are formed by the superposition of two waves of the same frequency and amplitude traveling in opposite directions.
+The solution of the standing waves must satisfy a quantization condition, which leads to discrete frequencies and wavelengths.
+This discretized solution is given by:
+    \begin{equation}\label{eq:standing_wave}
+        y(x,t) = A_n \sin(kx) \cos(\omega t)
+    \end{equation}
+    where \( k = \frac{n\pi}{L} \) for \( n = 1, 2, 3, \ldots \), and \( L \) is the length of the medium (e.g., a string fixed at both ends). The corresponding frequencies are given by: 
+    $$
+        f_n = \frac{n v}{2L}
+    $$
+    In turn, the wavelengths are given by:
+    $$
+        \lambda_n = \frac{2L}{n}
+    $$
+    where \( v \) is the wave speed. The distance between two consecutive nodes or antinodes is given by:
+    $$
+        d = \frac{\lambda}{2}
+    $$
+
+\end{definition}
+
+\paragraph{Systems with no Fixed Ends} For systems with no fixed ends, the general solution is given by:
+$$
+    y(x,t) = A_n \cos(kx) \cos(\omega_n t)
+$$
+where \( k = \frac{n\pi}{L} \) for \( n = 0, 1, 2, \ldots \). 
+the postion of the nodes are:
+$$
+    x_n = n \frac{\lambda}{2}, \quad n = 0, \pm 1, \pm 2, \ldots
+$$
+and the position of the antinodes are:
+$$
+    x_a = \left(n + \frac{1}{2}\right) \frac{\lambda}{2}, \quad n = 0, \pm 1, \pm 2, \ldots
+$$
+
+\paragraph{Systems with no free Ends} For systems with no free ends, the general solution is given by:
+$$
+    y(x,t) = A_n \sin(kx) \cos(\omega_n t)
+$$
+
+The postion of the nodes are:
+$$
+    x_n = \left(n + \frac{1}{2}\right) \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+$$
+and the position of the antinodes are:
+$$
+    x_a = n \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+$$
+
+\paragraph{System with Mixed Ends} For systems with mixed ends (one fixed and one free), the general solution has two cases:
+$$
+    y(x,t) = \begin{cases}
+        A_n \sin(kx) \cos(\omega_n t) & \text{if the fixed end is at } x=0 \\
+        A_n \cos(kx) \cos(\omega_n t) & \text{if the fixed end is at } x=L
+    \end{cases}
+$$
+The postion of the nodes are:
+$$
+    x_n = \left(n + \frac{1}{2}\right) \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+$$
+and the position of the antinodes are:
+$$
+    x_a = n \frac{\lambda}{2}, \quad n = 0, 1, 2, \ldots
+$$
+
+\begin{definition}[Energy of Standing Waves]
+    The total energy stored in a standing wave on a string of length \( L \) is given by:
+    $$
+        E = \frac{1}{4} \mu A^2 \omega^2 L
+    $$
+    where \( \mu \) is the linear mass density of the string, \( A \) is the amplitude of the wave, and \( \omega \) is the angular frequency.
+The energy is equally divided between kinetic and potential energy, with each contributing half of the total energy.
+\end{definition}
+
+\begin{definition}[Average Power of Standing Waves]
+    The average power transmitted by a standing wave on a string is given by:
+    $$
+        \bar{P} = \frac{1}{2} \mu A^2 \omega^2 v
+    $$
+    where \( \mu \) is the linear mass density of the string, \( A \) is the amplitude of the wave, \( \omega \) is the angular frequency, and \( v \) is the wave speed.
+This power represents the rate at which energy is transferred along the string.
+
+\end{definition}
+\subsection{Fourier analysis}
+\begin{definition}[Standing Waves as Normal Modes of a Vibrating String]
+    The superposition principle states that if $y_1(x,t)$ and $y_2(x,t)$ are two solutions to the wave equation, then their sum $y(x,t) = y_1(x,t) + y_2(x,t)$ is also a solution.
+For a string fixed at both ends, the $n^\text{th}$ normal mode of vibration can be expressed as:
     \begin{equation}
-        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
-    \end{equation}
-    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
-\end{definition}
-
-\chapter{Modern Physics}
+        y_n(x,t) = A_n \sin\left(\frac{n\pi x}{L}\right) \cos(\omega_n t)
+    \end{equation}
+
+    The motion of the string will be a superposition of normal modes:
+    \begin{equation}
+        y(x,t) = \sum_{n=1}^{\infty} A_n \sin\left(\frac{n\pi x}{L}\right) \cos(\omega_n t)
+    \end{equation}
+    where \( A_n \) are the amplitudes of the respective modes, and \( \omega_n = \frac{n\pi v}{L} \) are the angular frequencies of the modes.
+\end{definition}
+
+\begin{definition}[The amplitude of normal modes – Fourier analysis]
+    If we look at the initial shape of the string:
+    $$
+        y(x,0) = \sum_n A_n \sin\left(\frac{n\pi x}{L}\right)
+    $$
+    Any initial shape of the string can be expressed as a Fourier sine series. Also, for fixed ends ($f(0)=f(L)=0$), we can write the shape as the following:
+    $$
+        f(x) = \sum_n A_n \sin\left(\frac{n\pi x}{L}\right)
+    $$
+
+    To find the coefficients \( A_n \), we use:
+    \begin{equation}
+        A_n = \frac{2}{L} \int_0^L f(x) \sin\left(\frac{n\pi x}{L}\right) \, dx
+    \end{equation}
+\end{definition}
 \end{document}

### 2025-10-13 YEAR2/AER210_LectureNotes/main.tex
oct 12

@@
 \section{Vector Calculus}
-\section{Line Integrals}
 
@@
 
+\begin{definition}[Surface Area]
+    The surface area of a parametric surface \(\vec{r(u,v)}\) over a domain \(D\) is given by:
+    \begin{equation}
+        S = \iint_D ||\vec{r_u} \times \vec{r_v}|| \, dA
+    \end{equation}
+    where \(\vec{r_u} = \frac{\partial \vec{r}}{\partial u}\) and \(\vec{r_v} = \frac{\partial \vec{r}}{\partial v}\).
+\end{definition}
+
+\begin{example}[Surface Area of a Sphere]
+    Find the surface area of a sphere with radius \(a\).
+
+    We can parametrize the sphere using spherical coordinates:
+    $$    
+        \vec{r}(\phi, \theta) = \langle x, y , z \rangle = \langle a \sin \phi \cos \theta, a \sin \phi \sin \theta, a \cos \phi \rangle
+    $$
+    where over the domain \(D = \{(\phi, \theta) | 0 \leq \phi \leq \pi, 0 \leq \theta \leq 2\pi\}\).
+
+    We can compute the partial derivatives:
+    $$
+        \vec{r_\phi} = \langle a \cos \phi \cos \theta, a \cos \phi \sin \theta, -a \sin \phi \rangle
+    $$
+    $$
+        \vec{r_\theta} = \langle -a \sin \phi \sin \theta, a \sin \phi \cos \theta, 0 \rangle
+    $$
+
+    Then, we can compute the cross product:
+    $$
+        \vec{r_\phi} \times \vec{r_\theta} = 
+        \begin{vmatrix}
+            \hat{i} & \hat{j} & \hat{k} \\ 
+            a\cos\phi\cos\theta & a\cos\phi\sin\theta & -a\sin\phi \\ 
+            -a\sin\phi\sin\theta & a\sin\phi\cos\theta & 0
+        \end{vmatrix}
+        = a^2 \sin^2 \phi \cos \theta \hat{i} + a^2 \sin^2 \phi \sin \theta \hat{j} + a^2 \sin \phi \cos \phi \hat{k}
+    $$
+
+    The magnitude of the cross product is:
+    $$
+        ||\vec{r_\phi} \times \vec{r_\theta}|| = a^2 \sin \phi
+    $$
+
+    Finally, we can compute the surface area:
+    $$
+        S = \iint_D ||\vec{r_\phi} \times \vec{r_\theta}|| \, dA = \int_0^{2\pi} \int_0^\pi a^2 \sin \phi \, d\phi \, d\theta = 2\pi a^2 \left[ -\cos \phi \right]_0^\pi = 4\pi a^2
+    $$
+
+\end{example}
+
+\subsection{Surface Integrals}
+\begin{definition}[Surface Integral over Scalar Field]
+    Consider a scalar field \(f(x,y,z)\) and a smooth parametric surface \(\vec{r(u,v)}\) over a domain \(D\). Then we define the surface integral of \(f\) over the surface as:
+    \begin{equation}
+        \iint_S f(x,y,z) \, dS = \iint_D f(\vec{r(u,v)}) ||\vec{r_u} \times \vec{r_v}|| \, du \, dv
+    \end{equation}
+
+    And if $S$ is piecewise smooth, then we can express the surface integral as the sum of the surface integrals over each smooth piece:
+    \begin{equation}
+        \iint_S f(x,y,z) \, dS = \sum_{i=1}^n \iint_{S_i} f(x,y,z) \, dS
+    \end{equation}
+    where \( \bigcup_{i=1}^n S_i = S \) and each \(S_i\) is a smooth surface.
+\end{definition}
+\begin{example}
+    \textit{Evaluate the $\iint_S \sqrt{x^2 + y^2 + 1} \, dS$ where \(S\) is the surface given by $\vec{r}(u,v) = \langle u\cos v, u\sin v, v \rangle$, $0 \leq u \leq 1$, $0 \leq v \leq 2\pi$.}   
+    We can compute the partial derivatives:
+    $$        \vec{r_u} = \langle \cos v, \sin v, 0 \rangle $$
+    $$        \vec{r_v} = \langle -u \sin v, u \cos v, 1 \rangle $$
+    Then, we can compute the cross product:
+    $$        \vec{r_u} \times \vec{r_v} = 
+        \begin{vmatrix}
+            \hat{i} & \hat{j} & \hat{k} \\ 
+            \cos v & \sin v & 0 \\ 
+            -u \sin v & u \cos v & 1
+        \end{vmatrix}
+        = \sin v \hat{i} - \cos v \hat{j} + u \hat{k}
+    $$
+    The magnitude of the cross product is:
+    $$        ||\vec{r_u} \times \vec{r_v}|| = \sqrt{\sin^2 v + \cos^2 v + u^2} = \sqrt{1 + u^2} $$
+    Finally, we can compute the surface integral:
+    \begin{align*}
+        \iint_S \sqrt{x^2 + y^2 + 1} \, dS &= \iint_D \sqrt{u^2 \cos^2 v + u^2 \sin^2 v + 1} \cdot \sqrt{1 + u^2} \, du \, dv \\
+        &= \int_0^{2\pi} \int_0^1 \sqrt{u^2 + 1} \cdot \sqrt{1 + u^2} \, du \, dv = \int_0^{2\pi} \int_0^1 (1 + u^2) \, du \, dv \\
+        &= \int_0^{2\pi} \left[ u + \frac{u^3}{3} \right]_0^1 \, dv = \int_0^{2\pi} \frac{4}{3} \, dv = \frac{8\pi}{3}
+    \end{align*}
+\end{example}
+\begin{definition}[Orientability]
+    A surface is said to be orientable if it is possible to assign a consistent normal vector at every point on the surface. In other words, a surface is orientable if it has two distinct sides. \\
+    (Examples) Spheres, Torus, Cylinder, Plane \\    
+    (Non-examples) Möbius Strip, Klein Bottle
+\end{definition}
+
+\begin{definition}[Orientation]
+    An orientation of a surface is a choice of a continuous normal vector field on the surface.
+For a given surface, there are two possible orientations: one with the normal vector pointing "outward" and the other with the normal vector pointing "inward".
+The choice of orientation affects the sign of the surface integral.
+For closed surfaces, the outward orientation is typically chosen for obvious reasons.
+\end{definition}
+\begin{definition}[Surface Integrals of Vector Fields]
+    Consider a vector field \(\vec{F}(x,y,z)\) and an orientable smooth surface \(S\) with a chosen orientation. Then we define the surface integral of \(\vec{F}\) over \(S\) as:
+    \begin{equation}
+        \iint_D \vec{F}(\vec{r(u,v)}) \cdot \hat{n} \, dS = \iint_D \vec{F}(\vec{r(u,v)}) \cdot (\vec{r_u} \times \vec{r_v}) \, du \, dv
+    \end{equation}
+    where $\hat{n} = \frac{\vec{r_u} \times \vec{r_v}}{||\vec{r_u} \times \vec{r_v}||}$ is the unit normal vector to the surface and $dS = ||\vec{r_u} \times \vec{r_v}|| \, du \, dv$ is the infinitesimal area element on the surface.
+\end{definition}
+
+\paragraph{Physical Interpretation - Flux Integration} Consider a fluid with density $\rho (x, y, z)$ and velocity filed $\vec{v(x, y, z)}$.
+For each patch of surface $dS$ with unit normal vector $\vec{n}$, the volume of fluid passing through the patch per unit time is given by is one of a prism given by:
+$$
+     \text{Height} = \vec{v} \cdot \vec{n} \, dt, \quad \text{Volume} = (\vec{v} \cdot \vec{n}) \, dS \, dt \, \implies dm = \rho (\vec{v} \cdot \vec{n}) \, dS \, dt
+$$
+The mass flow rate $d\text{Mass}/dt$ is given by:
+$$
+    \frac{\text{Mass}}{dt} = \rho (\vec{v} \cdot \vec{n}) \, dS
+$$
+So we write:
+$$
+    \text{Mass flow rate} = \iint_S \rho \vec{v} \cdot \vec{n}\, dS = \iint_S F \cdot \vec{n} \, dS = \iint_S \vec{F} \cdot d\vec{S}
+$$
+where $\vec{F} = \rho \vec{v}$ is the flux density vector field.
+
+\begin{example}
+    \textit{Calculate the flux of the vector filed $\vec{F}(x,y,z) = x \, \hat{i} + y \, \hat{j} + z \, \hat{k}$ across the closed cylindrical surface \(S\) given by \(x^2 + y^2 = a^2\), \(-1 \leq z \leq 1\), oriented outward.}
+
+    We first parametrize the cylindrical surface as:
+    $$        \vec{r}(\theta, z) = \langle a \cos \theta, a \sin \theta, z \rangle $$
+    over the domain \(D = \{(\theta, z) | 0 \leq \theta \leq 2\pi, -1 \leq z \leq 1\}\).
+
+    Then, we can compute the partial derivatives:
+    $$        \vec{r_\theta} = \langle -a \sin \theta, a \cos \theta, 0 \rangle $$
+    $$        \vec{r_z} = \langle 0, 0, 1 \rangle $$
+    Then, we can compute the cross product:
+    $$        \vec{r_\theta} \times \vec{r_z} = 
+        \begin{vmatrix}
+            \hat{i} & \hat{j} & \hat{k} \\ 
+            -a \sin \theta & a \cos \theta & 0 \\ 
+            0 & 0 & 1
+        \end{vmatrix}
+        = a \cos \theta \hat{i} + a \sin \theta \hat{j} + 0 \hat{k}
+    $$
+    The magnitude of the cross product is:
+    $$        ||\vec{r_\theta} \times \vec{r_z}|| = \sqrt{(a \cos \theta)^2 + (a \sin \theta)^2 + 0^2} = a $$
+    Finally, we can compute the surface integral:
+    \begin{align*}
+        \iint_S \vec{F} \cdot d\vec{S} &= \iint_D \vec{F}(\vec{r}(\theta, z)) \cdot (\vec{r_\theta} \times \vec{r_z}) \, d\theta \, dz \\
+        &= \int_0^{2\pi} \int_{-1}^1 \langle a \cos \theta, a \sin \theta, z \rangle \cdot \langle a \cos \theta, a \sin \theta, 0 \rangle \, dz \, d\theta \\
+        &= \int_0^{2\pi} \int_{-1}^1 a^2 (\cos^2 \theta + \sin^2 \theta) \, dz \, d\theta = 2\int_0^{2\pi} a^2 (1) (1 - 0) \, d\theta = 4\pi a^2
+    \end{align*}
+    And for the top and bottom surfaces, we have:
+    \begin{align*}
+        \iint_{S_{top}} \vec{F} \cdot d\vec{S} &= \iint_D \vec{F}(\vec{r}(r, \theta)) \cdot (\vec{r_r} \times \vec{r_\theta}) \, dr \, d\theta \\
+        &= \int_0^{2\pi} \int_0^a \langle r \cos \theta, r \sin \theta, 1 \rangle \cdot \langle 0, 0, r \rangle \, dr \, d\theta \\
+        &= \int_0^{2\pi} \int_0^a r \, dr \, d\theta = \int_0^{2\pi} \left[ \frac{r^2}{2} \right]_0^a \, d\theta = \int_0^{2\pi} \frac{a^2}{2} \, d\theta = \pi a^2
+    \end{align*}
+    And then, by symmetry, we have:
+    $$        \iint_{S_{bottom}} \vec{F} \cdot d\vec{S} = \pi a^2 $$
+    So, the total flux across the closed surface is:
+    $$        \iint_S \vec{F} \cdot d\vec{S} = 4\pi a^2 + \pi a^2 + \pi a^2 = 6\pi a^2 $$
+\end{example}
+
+\begin{example}
+    Find the flux of $\vec{F}(x,y,z) = \frac{2x \hat{i} + 2y \hat{j}}{x^2 + y^2} + \hat{k}$ across the surface of the paraboloid $\vec{r}(u, v) = \langle u \cos v, u \sin v, u^2 \rangle$, $0 \leq u \leq 1$, $0 \leq v \leq 2\pi$, oriented downward.
+
+    We can compute the partial derivatives:
+    $$        \vec{r_u} = \langle \cos v, \sin v, 2u \rangle $$
+    $$        \vec{r_v} = \langle -u \sin v, u \cos v, 0 \rangle $$
+    Then, we can compute the cross product:
+    $$        \vec{r_u} \times \vec{r_v} = -2u^2 \cos v \hat{i} - 2u^2 \sin v \hat{j} + u \hat{k}
+    $$
+    Notice, since the surface is oriented downward, we take the negative of the cross product:
+    $$        \vec{r_v} \times \vec{r_u} = 2u^2 \cos v \hat{i} + 2u^2 \sin v \hat{j} - u \hat{k}
+    $$
+\end{example}
 \chapter{Fluid Mechanics}
+\section{Introduction to Fluid Mechanics}
+\begin{definition}[Fluid]
+    A fluid is a substance that continuously deforms (flows) under an applied shear stress, no matter how small the stress. Fluids include liquids and gases.
+
+\end{definition}
 \end{document}

### 2025-10-13 YEAR2/ECE253_LectureNotes/main.tex
oct 12

@@
 
+\begin{example}[Alternative Design]
+    Alternatively, we can design the state diagram in which each state represents the last three values. We can set the following states table:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|cc|c|}
+            \hline
+            $y_1 y_2 y_3$ & $w=0$ & $w=1$ & Output \\
+            \hline
+            000 & 000 & 001 & 0 \\
+            001 & 010 & 011 & 0 \\
+            010 & 100 & 101 & 0 \\
+            011 & 110 & 111 & 0 \\
+            100 & 000 & 001 & 0 \\
+            101 & 010 & 011 & 1 \\
+            110 & 100 & 101 & 0 \\
+            111 & 110 & 111 & 0 \\
+            \hline
+        \end{tabular}
+        \caption{State Table for Alternative FSM that detects `101'}
+        \label{tab:alternative_fsm_state_table}
+    \end{table}
+    And then we can draw the following equations:
+    \begin{itemize}
+        \item $Y_3 = w$
+        \item $Y_2 = y_3$
+        \item $Y_1 = y_2$
+    \end{itemize}
+    The output $Z = y_3 \overline{y_2} y_1$.
+    This design is simpler, but it uses more states (8 states instead of 4 states). This is an example of the trade-off between the number of states and the complexity of the logic.
+
+    We call this a shift register, where the input is shifted into the register on each clock cycle. The output is taken from the last bit of the register.
+\end{example}
+
 \begin{definition}[Finite State Machine (FSM)]

### 2025-10-13 YEAR2/MAT292_LectureNotes/main.tex
oct 12

@@
 \end{equation}
-The phase portrait of this system is shown below:
-% insert later
-
+
+\paragraph{Applying the matrix to the generalized eigenvector} We can apply the matrix \( A \) to the generalized eigenvector \( v_g \), then , $Av_g$ would stay in the span of $v$ and $v_g$: In other words, we call this subspace invariant under the transformation $A$.
+
+\begin{example}[Generalized Eigenvectors]
+    Consider the system of ODEs:
+    $$
+    \textbf{x}' = \begin{bmatrix} 4 & 1 \\ -1 & 2
+    \end{bmatrix}\textbf{x}
+    $$
+    We find the eigenvalues and eigenvectors by solving the characteristic polynomial:
+    \begin{align*}
+    \det(A - \lambda I) &= \det\begin{bmatrix} 4 - \lambda & 1 \\ -1 & 2 - \lambda
+    \end{bmatrix} \\
+    &= (\lambda - 3)^2 = 0 \\
+    \implies \lambda_1 &= 3, \quad \lambda_2 = 3
+    \end{align*}
+    The corresponding eigenvector is:
+    \begin{align*}
+    \vec{v}_1 &= \begin{bmatrix} 1 \\ -1
+    \end{bmatrix}
+    \end{align*}
+    Since there is only one linearly independent eigenvector, we need to find a generalized eigenvector \( v_g \) satisfying:
+    $$
+    (A - 3I)v_g = v_1
+    $$
+    Solving this, we find:
+    $$
+    v_g = \begin{bmatrix} 0 \\ 1
+    \end{bmatrix}
+    $$
+    Thus, the two linearly independent solutions to the system are:
+    $$x_1(t) = v_1 e^{3t} = \begin{bmatrix} 1 \\ -1
+    \end{bmatrix} e^{3t}$$
+    $$x_2(t) = (v_g + tv_1)e^{3t} = \left(\begin{bmatrix} 0 \\ 1
+    \end{bmatrix} + t\begin{bmatrix} 1 \\ -1
+    \end{bmatrix}\right)e^{3t}$$
+
+    Therefore, the general solution to the system is:
+    $$\textbf{x}(t) = c_1 x_1(t) + c_2 x_2(t)$$
+    where \( c_1 \) and \( c_2 \) are constants determined by the initial conditions.
+\end{example}
 \begin{definition}[Trace-Determinant Plane]
@@
 \paragraph{Intuition} A monotonically increasing function that gets bounded must converge.
-
-\section{Matrices and Linear Algebra}
-
-\subsection{Matrix Exponentials}
+A similar idea applies to trajectories in a dynamical system: if they are confined to a bounded region and cannot escape, they must eventually settle into a steady state, such as a limit cycle.
+
+\section{Matrix Exponentials}
 \begin{shaded}
@@
 
+\paragraph{What if \( A \) is not diagonalizable?} If \( A \) is not diagonalizable there could be two cases:
+\begin{itemize}
+    \item If we know the fundamental set of solutions \( \{x_1(t), x_2(t), \ldots, x_n(t)\} \), then we can construct the fundamental matrix \( \Phi(t) \) and compute the matrix exponential as:
+    \begin{equation}
+    e^{At} = \Phi(t)\Phi(0)^{-1}
+    \end{equation}
+    where \( \Phi(0) \) is the fundamental matrix evaluated at \( t = 0 \). And $\phi = [x_1, x_2, \ldots, x_n]$.
+    \item Otherwise, we can consider its Jordan Canonical Form.
+\end{itemize}
+
+\paragraph{Jordan Canonical Form} We first define the AM and GM of an eigenvalue:
+\begin{definition}[Geometric and Algebraic Multiplicity]
+    Consider a square matrix \( A \in \mathbb{R}^{n \times n} \) and an eigenvalue \( \lambda \) of \( A \).
+The algebraic multiplicity (AM) of \( \lambda \) is the number of times \( \lambda \) appears as a root of the characteristic polynomial of \( A \).
+The geometric multiplicity (GM) of \( \lambda \) is the dimension of the eigenspace corresponding to \( \lambda \), which is the null space of \( A - \lambda I \).
+Diagonalizability is characterized by the condition that for each eigenvalue, its algebraic multiplicity equals its geometric multiplicity.
+\end{definition}
+
+\begin{definition}[Jordan Block]
+    A Jordan block \( J_k(\lambda) \) of size \( k \) associated with an eigenvalue \( \lambda \) is a \( k \times k \) matrix of the form:
+    \begin{equation}
+    J_k(\lambda) = \begin{bmatrix}
+    \lambda & 1 & 0 & \cdots & 0 \\
+    0 & \lambda & 1 & \cdots & 0 \\
+    0 & 0 & \lambda & \cdots & 0 \\
+    \vdots & \vdots & \vdots & \ddots & 1 \\
+    0 & 0 & 0 & \cdots & \lambda
+    \end{bmatrix}
+    \end{equation}
+    where \( \lambda \) is the eigenvalue and the superdiagonal entries are all ones, while all other entries are zero.
+
+\end{definition}
+
+\begin{definition}[Jordan Canonical Form]
+    A square matrix \( A \in \mathbb{R}^{n \times n} \) is said to be in Jordan canonical form if it is a block diagonal matrix where each block is a Jordan block.
+Specifically, \( A \) can be expressed as:
+    \begin{equation}
+    A = PJP^{-1}
+    \end{equation}
+    where \( P \) is an invertible matrix and \( J \) is the Jordan matrix composed of Jordan blocks along its diagonal.
+The Jordan canonical form provides a way to represent matrices that are not diagonalizable, allowing for the computation of functions of matrices, such as the matrix exponential.
+\end{definition}
+\begin{example}
+    Consider the system of ODEs:
+    $$x' = \begin{bmatrix}
+    1 & -1 \\
+    1 & 3
+    \end{bmatrix} x$$
+    The characteristic polynomial is given by:
+    \begin{align*}
+        p(\lambda) &= \det\begin{bmatrix}
+        1 - \lambda & -1 \\
+        1 & 3 - \lambda
+        \end{bmatrix} \\
+        &= (1 - \lambda)(3 - \lambda) + 1 \\
+        &= \lambda^2 - 4\lambda + 4 \\
+        &= (\lambda - 2)^2
+    \end{align*}
+    So, we have a repeated eigenvalue \( \lambda = 2 \) with algebraic multiplicity 2. To find the eigenvectors, we solve:
+    \begin{align*}
+        (A - 2I)v &= 0 \\
+        \begin{bmatrix}
+        -1 & -1 \\
+        1 & 1
+        \end{bmatrix} \begin{bmatrix}
+        v_1 \\ v_2
+        \end{bmatrix} &= 0
+    \end{align*}
+    This gives us the eigenvector \( v = \begin{bmatrix}1 \\ -1 \end{bmatrix} \). Since the geometric multiplicity is 1, we need to find a generalized eigenvector \( w \) such that:
+    \begin{align*}
+        (A - 2I)w &= v \\
+        \begin{bmatrix}
+        -1 & -1 \\
+        1 & 1
+        \end{bmatrix} \begin{bmatrix}
+        w_1 \\ w_2
+        \end{bmatrix} &= \begin{bmatrix}1 \\ -1
+        \end{bmatrix}
+    \end{align*}
+    Solving this, we can take \( w = \begin{bmatrix}0 \\ -1 \end{bmatrix} \). Now, we can form the Jordan matrix:
+    $$
+    J = \begin{bmatrix}
+    2 & 1 \\
+    0 & 2
+    \end{bmatrix} = 2I + \begin{bmatrix}
+    0 & 1 \\
+    0 & 0
+    \end{bmatrix}
+    $$
+    Obserbe that $N = \begin{bmatrix}
+    0 & 1 \\
+    0 & 0
+    \end{bmatrix}$ is nilpotent since $N^2 = 0$. And that $I$ commutes with everything. So, we can compute the matrix exponential as:
+    \begin{align*}
+        e^{At} &= Pe^{Jt}P^{-1} \\
+        &= P e^{(2I + N)t} P^{-1} \\
+        &= P e^{2tI} e^{Nt} P^{-1} \\
+        &= P e^{2t} \left(I + Nt + \frac{(Nt)^2}{2!} + \ldots \right) P^{-1} \\
+        &= P e^{2t} \left(I + Nt\right) P^{-1} \\
+        &= e^{2t} P \begin{bmatrix}
+        1 & t \\
+        0 & 1
+        \end{bmatrix} P^{-1}
+    \end{align*}
+    where \( P = \begin{bmatrix}1 & 0 \\
+    -1 & -1
+    \end{bmatrix} \) and \( P^{-1} = \begin{bmatrix}
+    1 & 0 \\
+    -1 & -1
+    \end{bmatrix} \). Thus, we have:
+    $$
+        e^{At} = e^{2t} \begin{bmatrix}1 + t & t \\
+        -t & 1 - t
+        \end{bmatrix}
+    $$
+\end{example}
+
+\section{Higher Order ODEs}
+\subsection{Second Order Linear ODEs}
+
+\begin{definition}[General Form of Second Order Linear ODEs]
+    A second-order linear ordinary differential equation (ODE) has the general form:
+    \begin{equation}
+     y'' = f(t, y, y')
+    \end{equation}
+
+\end{definition}
+
+\begin{example} \label{ex:2ndOrderTo1stOrder}
+    We can convert a second-order linear ODE into a system of first-order ODEs. Consider the second-order ODE:
+    $$
+     y'' + p(t)y' + q(t)y = g(t)
+    $$
+    We can define:
+    $$\begin{cases}
+    x_1 = y \\
+    x_2 = y'
+    \end{cases}$$
+    Then, we have the system of first-order ODEs:
+    $$\begin{cases}
+    x_1' = x_2 \\
+    x_2' = -g(t) - p(t)x_2 - q(t)x_1
+    \end{cases}$$
+
+    In matrix form, this can be written as:
+    $$\begin{bmatrix}
+    x_1' \\ x_2'
+    \end{bmatrix} = \begin{bmatrix}
+    0 & 1 \\
+    -q(t) & -p(t)
+    \end{bmatrix} \begin{bmatrix}
+    x_1 \\ x_2
+    \end{bmatrix} + \begin{bmatrix}
+    0 \\ -g(t)
+    \end{bmatrix}$$ 
+    This transformation allows us to apply techniques for first-order systems to analyze and solve second-order ODEs.
+
+\end{example}
+
+\begin{example}
+    Consider the above Example \ref{ex:2ndOrderTo1stOrder} with constant coefficients:
+    $$y'' + p y' + q y = 0
+    $$
+    We can write the corresponding system of first-order ODEs as:
+    $$\begin{bmatrix}
+    x_1' \\ x_2'
+    \end{bmatrix} = \begin{bmatrix}
+    0 & 1 \\
+    -q & -p
+    \end{bmatrix} \begin{bmatrix}
+    x_1 \\ x_2
+    \end{bmatrix}
+    $$
+    And you should be able to solve and analyze the system using the techniques we have learned for first-order systems, especially the trace-determinant plane.
+
+\end{example}
+
+\begin{definition}[Characteristic Equation]
+    The characteristic equation of a second-order linear ODE with constant coefficients:
+    $$y'' + p y' + q y = 0$$
+    is given by the quadratic equation:
+    \begin{equation}
+    r^2 + pr + q = 0
+    \end{equation}
+    where \( r \) represents the roots of the characteristic polynomial.
+The nature of the roots (real and distinct, real and repeated, or complex conjugates) determines the form of the general solution to the ODE.
+This comes from the assuned solution of the form \( y = e^{rt} \).
+\end{definition}
+
+\begin{example}[Cauchy-Euler Equation]
+    Consider the following second-order ODE:
+    $$t^2 y'' - t y' + y = 0$$
+    This is in the form of a Cauchy-Euler equation:
+    $$
+    \sum_m y^{(m)}(t) a_m t^m = 0
+    $$
+    where \( a_2 = 1, a_1 = -1, a_0 = 1 \). We can solve this by assuming a solution of the form \( y = t^r \). Then, we have:
+    \begin{align*}
+        y' &= r t^{r-1} \\
+        y'' &= r(r-1) t^{r-2}
+    \end{align*}
+    Substituting these into the original equation gives:
+    \begin{align*}
+        t^2 (r(r-1) t^{r-2}) - t (r t^{r-1}) + t^r &= 0 \\
+        r(r-1) t^r - r t^r + t^r &= 0 \\
+        (r^2 - 2r + 1) t^r &= 0 \\
+        (r - 1)^2 t^r &= 0
+    \end{align*}
+    Thus, we have a repeated root \( r = 1 \). The general solution for this case is given by:
+    $$y(t) = C_1 t + C_2 t \ln(t)$$
+    where \( C_1 \) and \( C_2 \) are constants determined by initial conditions.
+\end{example}
+
 \end{document}

### 2025-10-13 YEAR2/PHY293_LectureNotes/main.tex
oct 12

@@
     A driven harmonic oscillator is described by the following differential equation, which includes an external driving force $F(t)$:
-    $$
-        m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = F(t)   
-    $$
+    \begin{equation}\label{eq:driven_ho}
+        m \ddot{x} + b \dot{x} + kx = F(t)
+    \end{equation}
 \end{definition}
@@
 The same as equation (\ref{eq:hooke}), which can be rewritten as:
-    $$
+    \begin{equation}\label{eq:newton_hooke}
         F = m\ddot{x} = -kx
-    $$
+    \end{equation}xs
     By setting $\omega^2 = \frac{k}{m}$, a general solution can be written as:
-    $$
+    \begin{equation}\label{eq:sho_general}
         x(t) = x_0 + A_1 \cos(\omega t) + A_2 \sin(\omega t)
-    $$
+    \end{equation}
     where $A_1$ and $A_2$ are constants determined by the IVP, $\omega$ is the angular frequency, and $\phi$ is the phase constant.
@@
     The period \( T \) is the time it takes for one complete cycle of the motion, given by:
-    $$
+    \begin{equation}\label{eq:period}
         T = 2\pi \sqrt{\frac{m}{k}}
-    $$
+    \end{equation}
     The frequency \( f \) is the number of cycles per unit time, given by:
-    $$
+    \begin{equation}\label{eq:frequency}
         f = \frac{1}{T} = \frac{1}{2\pi} \sqrt{\frac{k}{m}}
-    $$
+    \end{equation}
     The angular frequency \( \omega \) is related to the frequency by:
-    $$
+    \begin{equation}\label{eq:omega_rel}
         \omega = 2\pi f = \sqrt{\frac{k}{m}}
-    $$
+    \end{equation}
 
@@
 Which could be rewritten as:
-    $$
+    \begin{equation}\label{eq:damped_ode}
         \ddot{x} + \gamma \dot{x} + \omega_0^2 x = 0
-    $$
+    \end{equation}
     where \( \omega_0 = \sqrt{\frac{k}{m}} \) is the natural angular frequency of the undamped oscillator, and \( \gamma = \frac{b}{m} \) is the damping coefficient per unit mass.
@@
     Consider a driver force acting on the mass-spring system; the ODE becomes:
-    $$
+    \begin{equation}\label{eq:forced_undamped}
         m\ddot{x} + kx = \eta
-    $$
+    \end{equation}
     where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The steady-state particular solution is:
-    $$
+    \begin{equation}\label{eq:forced_ss}
         x(t) = A(\omega)\cos(\omega t - \delta)
-    $$
+    \end{equation}
     To derive $A(\omega)$ and $\delta$, use the equation of motion:
@@
 Consider the arc length \( s \) of a pendulum bob, we have:
-$$
-    s = l\theta \quad m\ddot{s} = -mg\sin(\theta)
-$$
+\begin{equation}\label{eq:pendulum_arc}
+\begin{aligned}
+    s &= l\theta \\
+    m\ddot{s} &= -mg\sin\theta
+\end{aligned}
+\end{equation}
 
 For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
-$$
+\begin{equation}\label{eq:pendulum_small}
     \ddot{\theta} + \frac{g}{l}\theta = 0
-$$
+\end{equation}
 And the ODE has solution:
-$$
-    \theta(t) = \theta_0 \cos(\omega t + \phi)
-$$
+\begin{equation}\label{eq:pendulum_solution}
+    	heta(t) = \theta_0 \cos(\omega t + \phi)
+\end{equation}
 
 \paragraph{Energy} For small angle $\theta$, the total mechanical energy \( E \) of the pendulum is given by:
-$$
-    E = K + U = \frac{1}{2}ml^2\dot{\theta}^2 + mgl(1 - \cos(\theta)) \approx \frac{1}{2}ml^2\dot{\theta}^2 + \frac{1}{2}mgl\theta^2
-$$
+\begin{equation}\label{eq:pendulum_energy}
+    E = K + U = \frac{1}{2}ml^2\dot{\theta}^2 + mgl\big(1 - \cos\theta\big) \approx \frac{1}{2}ml^2\dot{\theta}^2 + \frac{1}{2}mgl\theta^2
+\end{equation}
 This shows that the pendulum behaves like a simple harmonic oscillator with an effective spring constant \( k = \frac{mg}{l} \).
@@
 \paragraph{The physical pendulum} For a rigid body swinging about a pivot point, the equation of motion is given by:
-$$
+\begin{equation}\label{eq:physical_tau}
     I\ddot{\theta} = \tau
-$$
+\end{equation}
 For simple rod of length \( L \) and mass \( m \) pivoted at one end, the moment of inertia about the pivot point is \( I = \frac{1}{3}mL^2 \).
-The torque due to gravity when the rod is displaced by an angle \( \theta \) from the vertical is \( \tau = -mg\frac{L}{2}\sin(\theta) \).
-For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
-$$
+The torque due to gravity when the rod is displaced by an angle \( \theta \) from the vertical is \( \tau = -mg\tfrac{L}{2}\sin\theta \).
+For small angles, we can approximate \( \sin\theta \approx \theta \), leading to:
+\begin{equation}\label{eq:physical_small}
     \ddot{\theta} + \frac{3g}{2L}\theta = 0
-$$
-And the small angle approximates leads to:
-$$
+\end{equation}
+And the small angle approximation leads to:
+\begin{equation}\label{eq:physical_omega_form}
     \ddot{\theta} + \omega^2 \theta = 0
-$$
-where \( \omega = \sqrt{\frac{3g}{2L}} \). The solution to this ODE is:
-$$
-    \theta(t) = \theta_0 \cos(\omega t + \phi)
-$$
+\end{equation}
+where \( \omega = \sqrt{\tfrac{3g}{2L}} \). The solution to this ODE is:
+\begin{equation}\label{eq:physical_solution}
+    	heta(t) = \theta_0 \cos(\omega t + \phi)
+\end{equation}
 The angular frequency of oscillation for the physical pendulum is:
-$$
+\begin{equation}\label{eq:physical_omega}
     \omega = \sqrt{\frac{3g}{2L}}
-$$
+\end{equation}
 
 \paragraph{The LC Circuit} An LC circuit consists of an inductor \( L \) and a capacitor \( C \) connected in series. The charge \( q(t) \) on the capacitor satisfies the differential equation:
-$$
+\begin{equation}\label{eq:LC_ode}
     L\frac{d^2q}{dt^2} + \frac{1}{C}q = 0
-$$
+\end{equation}
 This is analogous to the equation of motion for a simple harmonic oscillator, with the angular frequency given by:
-$$
+\begin{equation}\label{eq:LC_omega}
     \omega = \frac{1}{\sqrt{LC}}
-$$
+\end{equation}
 
@@
 The charge \( q(t) \) on the capacitor satisfies the differential equation:
-$$
+\begin{equation}\label{eq:RLC_ode}
     L\frac{d^2q}{dt^2} + R\frac{dq}{dt} + \frac{1}{C}q = 0
-$$
+\end{equation}
 This is analogous to the resistor behaving as a damped harmonic oscillator, with the angular frequency given by:
-$$
+\begin{equation}\label{eq:RLC_omega}
     \omega = \sqrt{\frac{1}{LC} - \left(\frac{R}{2L}\right)^2}
-$$
+\end{equation}
 The following are equations that summarize the analogies between mechanical and electrical oscillators:
@@
 
-    $$
-        \begin{cases}
-            \ddot{x}_1 &= -\frac{g}{L}x_1 + \frac{k}{m_1} (x_1 - x_2) \quad (a) \\
-            \ddot{x}_2 &= -\frac{g}{L}x_2 - \frac{k}{m_2} (x_1 - x_2) \quad (b)
-        \end{cases}
-    $$
+    \begin{equation}\label{eq:coupled_pendulum}
+        \begin{cases}
+            \ddot{x}_1 = -\dfrac{g}{L}x_1 + \dfrac{k}{m_1} (x_1 - x_2) \\
+            \ddot{x}_2 = -\dfrac{g}{L}x_2 - \dfrac{k}{m_2} (x_1 - x_2)
+        \end{cases}
+    \end{equation}
     where \( x_1 \) and \( x_2 \) are the displacements of masses \( m_1 \) and \( m_2 \) from their equilibrium positions, respectively. Now if we add and subtract the two equations, we get:
-    $$
-        \begin{cases}
-            \ddot{x}_1 + \ddot{x}_2 &= -\frac{g}{L}(x_1 + x_2) \quad (c) \\
-            \ddot{x}_1 - \ddot{x}_2 &= -\left(\frac{g}{L} + \frac{k}{m_1} + \frac{k}{m_2}\right)(x_1 - x_2) \quad (d)
-        \end{cases}
-    $$
+    \begin{equation}\label{eq:sum_diff}
+        \begin{cases}
+            \ddot{x}_1 + \ddot{x}_2 = -\dfrac{g}{L}(x_1 + x_2) \\
+            \ddot{x}_1 - \ddot{x}_2 = -\left(\dfrac{g}{L} + \dfrac{k}{m_1} + \dfrac{k}{m_2}\right)(x_1 - x_2)
+        \end{cases}
+    \end{equation}
     Let:
-    $$ 
-    q_1 = x_1 + x_2 \quad \text{and} \quad q_2 = x_1 - x_2
-    $$
+    \begin{equation}\label{eq:q_defs}
+    q_1 = x_1 + x_2, \quad q_2 = x_1 - x_2
+    \end{equation}
     also, we let:
-    $$
-        \omega_1 = \sqrt{\frac{g}{L}} \quad \text{and} \quad \omega_2 = \sqrt{\frac{g}{L} + \frac{k}{m_1} + \frac{k}{m_2}}
-    $$
+    \begin{equation}\label{eq:omega12}
+        \omega_1 = \sqrt{\frac{g}{L}}, \quad \omega_2 = \sqrt{\frac{g}{L} + \frac{k}{m_1} + \frac{k}{m_2}}
+    \end{equation}
     Then, the equations of motion can be rewritten as:
-    $$
-        \begin{cases}
-            \ddot{q}_1 &= -\omega_1^2 q_1 \\
-            \ddot{q}_2 &= -\omega_2^2 q_2
-        \end{cases}
-    $$
+    \begin{equation}\label{eq:q_odes}
+        \begin{cases}
+            \ddot{q}_1 = -\omega_1^2 q_1 \\
+            \ddot{q}_2 = -\omega_2^2 q_2
+        \end{cases}
+    \end{equation}
     Observe that \( q_1 \) and \( q_2 \) are decoupled, and each behaves like a simple harmonic oscillator with angular frequencies \( \omega_1 \) and \( \omega_2 \), respectively.
 The general solutions for \( q_1(t) \) and \( q_2(t) \) are:
-    $$
-        \begin{cases}
-            q_1(t) &= A_1 \cos(\omega_1 t + \phi_1) \\
-            q_2(t) &= A_2 \cos(\omega_2 t + \phi_2)
-        \end{cases}
-    $$
+    \begin{equation}\label{eq:q_solutions}
+        \begin{cases}
+            q_1(t) = A_1 \cos(\omega_1 t + \phi_1) \\
+            q_2(t) = A_2 \cos(\omega_2 t + \phi_2)
+        \end{cases}
+    \end{equation}
     where \( A_1, A_2, \phi_1, \) and \( \phi_2 \) are constants determined by the initial conditions.
 Finally, we can expressions for \( x_1(t) \) and \( x_2(t) \) in terms of \( q_1(t) \) and \( q_2(t) \):
-    $$
-        \begin{cases}
-            x_1(t) &= \frac{q_1(t) + q_2(t)}{2} = \frac{A_1}{2} \cos(\omega_1 t + \phi_1) + \frac{A_2}{2} \cos(\omega_2 t + \phi_2) \\
-            x_2(t) &= \frac{q_1(t) - q_2(t)}{2} = \frac{A_1}{2} \cos(\omega_1 t + \phi_1) - \frac{A_2}{2} \cos(\omega_2 t + \phi_2)
-        \end{cases}
-    $$
+    \begin{equation}\label{eq:x_from_q}
+        \begin{cases}
+            x_1(t) = \dfrac{q_1(t) + q_2(t)}{2} = \dfrac{A_1}{2} \cos(\omega_1 t + \phi_1) + \dfrac{A_2}{2} \cos(\omega_2 t + \phi_2) \\
+            x_2(t) = \dfrac{q_1(t) - q_2(t)}{2} = \dfrac{A_1}{2} \cos(\omega_1 t + \phi_1) - \dfrac{A_2}{2} \cos(\omega_2 t + \phi_2)
+        \end{cases}
+    \end{equation}
     All parts of the system oscillate with the same normal frequency (eigenvalue) in a normal mode (eigenvector).
@@
         \item \textbf{In-phase mode} (\(\omega_1\)): both masses swing together, reaching maximum displacement on the same side at the same time.  
-        $$
+    \begin{equation}\label{eq:in_phase}
     x_1 = x_2
-        $$
+    \end{equation}
 
         \item \textbf{Out-of-phase mode} (\(\omega_2\)): the two masses swing in opposite directions, so when one moves left, the other moves right.  
-        $$
+    \begin{equation}\label{eq:out_phase}
     x_1 = -x_2
-        $$
+    \end{equation}
     \end{itemize}
@@
 \paragraph{Energy in Coupled Oscillators} The total mechanical energy \( E \) of the coupled oscillator system is given by:
-$$
+\begin{equation}\label{eq:coupled_energy}
     E = \frac{1}{2}m_1\dot{x}_1^2 + \frac{1}{2}m_2\dot{x}_2^2 + \frac{1}{2}k(x_1 - x_2)^2 + \frac{1}{2}m_1\frac{g}{L}x_1^2 + \frac{1}{2}m_2\frac{g}{L}x_2^2
-$$
+\end{equation}
 Substituting the expressions for \( x_1 = \frac{q_1 + q_2}{2} \) and \( x_2 = \frac{q_1 - q_2}{2} \) into the energy equation, two independent energies terms for $q_1$ and $q_2$ emerge:
-$$
+\begin{equation}\label{eq:mode_energy}
     E = \left[ \frac{1}{4} m \dot{q}_1^2 + \frac{1}{4}\frac{m_1 g}{L} q_1^2 \right] + \left[ \frac{1}{4} m \dot{q}_2^2 + \frac{1}{4}\frac{m_2 g}{L} q_2^2 \right]
-$$
+\end{equation}
 This shows that the total energy is the sum of the energies associated with each normal mode, and that no cross term between the modes exists. Thus energy does not flow from one mode to another.
@@
 
+\paragraph{Energy Reflection and Transmission} The energy reflection and transmission coefficients are given by:
+\begin{equation}\label{eq:energy_coeff}
+    R_e = R^2 = \left(\frac{Z_2 - Z_1}{Z_2 + Z_1}\right)^2, \quad T_e = \frac{P_t}{P_i} = \frac{Z_1}{Z_2} T^2 = \frac{4Z_1 Z_2}{(Z_2 + Z_1)^2}
+\end{equation}
+Note that \( R_e + T_e = 1 \), which means that the total energy is conserved across the boundary.
+
 \begin{example}[Transmission of 64\% Energy]
@@
 This apparent paradox arises because the transmitted energy depends on both the amplitude squared and the ratio of impedances \(\tfrac{Z_2}{Z_1}\).
+\end{example}
+\begin{example}[Tricky Case: Impedance Mismatch with Energy Conservation]
+Let \(Z_2 = 3Z_1\). Then:
+\[
+R = \frac{Z_2 - Z_1}{Z_2 + Z_1} = \frac{1}{2}, 
+\quad T = \frac{2Z_2}{Z_2 + Z_1} = 1.5
+\]
+Amplitude transmission seems large, but power depends on both amplitude and impedance.
+
+Energy (power) transmission and reflection coefficients are:
+\[
+R_e = R^2 = 0.25, \qquad 
+T_e = \frac{Z_1}{Z_2} T^2 = \frac{1}{3}(1.5)^2 = 0.75
+\]
+so
+\[
+R_e + T_e = 1.00
+\]
+confirming energy conservation.
+
+\textbf{Interpretation:}
+The transmitted amplitude is higher because the second medium has greater impedance (\(Z_2 > Z_1\)), which means each unit of amplitude carries less energy.
+Therefore, amplitude ratios alone can be misleading — energy flow must always be evaluated using both amplitude and impedance.
 \end{example}
@@
 \end{definition}
+\subsection{Dispersion of Waves}
+\begin{definition}[Wave Number]
+    The wave number \( k \) is defined as:
+    \begin{equation}
+        k = \frac{2\pi}{\lambda}
+    \end{equation}
+    where \( \lambda \) is the wavelength of the wave. The wave number represents the spatial frequency of the wave, indicating how many wavelengths fit into a unit distance.
+\end{definition}
+\begin{definition}[Dispersion Relation]
+    The dispersion relation describes how the angular frequency \( \omega \) of a wave depends on its wave number \( k \). It is given by:
+    \begin{equation}
+        \omega = \omega(k)
+    \end{equation}
+    The specific form of the dispersion relation depends on the type of wave and the properties of the medium through which it propagates.
+
+    \textbf{Non-dispersive waves}: If \( \omega \) is directly proportional to \( k \) (i.e., \( \omega = vk \)), the wave is non-dispersive, meaning all frequency components travel at the same phase velocity \( v \).
+\end{definition}
+
+\begin{definition}[Phase and Group Velocity]
+    The phase velocity \( v_p \) is the rate at which the phase of the wave propagates in space. It is given by:
+    \begin{equation}
+        v_p = \frac{\omega}{k}  = \frac{c}{n}
+    \end{equation}
+    where \( c \) is the speed of light in a vacuum and \( n \) is the refractive index of the medium.
+
+    The group velocity \( v_g \) is the rate at which the envelope of a wave packet (or group of waves) propagates. It is given by:
+    \begin{equation}
+        v_g = \frac{d\omega}{dk} = v - k \frac{dv}{dk} \frac{\lambda}{k} = v - \lambda \frac{dv}{d\lambda}
+    \end{equation}
+    where \( v \) is the average of the phase velocities of the individual waves in the packet. $\lambda$ is average wavelength of the wave packet.
+
+    We have the following three cases:
+    \begin{itemize}
+        \item If \( v_p = v_g \), the wave is non-dispersive.
+        \item If \( v_p > v_g \), the wave is said to be normally dispersive.
+        \item If \( v_p < v_g \), the wave is said to be anomalously dispersive.
+    \end{itemize}
+\end{definition}
+\begin{definition}[Refractive Index]
+    The refractive index \( n \) of a medium is defined as the ratio of the speed of light in a vacuum \( c \) to the phase velocity \( v_p \) of light in the medium:
+    \begin{equation}
+        n = \frac{c}{v_p}
+    \end{equation}
+    The refractive index indicates how much the light slows down when it enters the medium compared to its speed in a vacuum.
+A higher refractive index means that light travels more slowly in that medium.
+\end{definition}
+\begin{example}
+    The yellow light from sodium has two components with wavelengths of 589.00 nm and 589.59 nm.
+The refractive index $n$ of a particular glass at these wavelengths has values of 1.6351 and 1.6350, respectively.
+    \textit{\begin{enumerate}
+        \item What are the phase velocities of the light at these wavelengths?
+        \item What is the velocity of the narrow pulse of sodium transmitted through the glass?
+    \end{enumerate}}
+    Also, we are given the following data:
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=0.5\textwidth]{nlambda.png}
+        \caption{$n-\lambda$ data}
+        \label{fig:refractive_index_data}
+    \end{figure}
+    \begin{enumerate}
+        \item The phase velocity \( v_p \) is given by:
+        $$
+            v_p = \frac{c}{n}
+        $$
+        For the first wavelength (589.00 nm):
+        $$
+            v_{p1} = \frac{3 \times 10^8 \,\text{m/s}}{1.6351} \approx 1.834 \times 10^8 \,\text{m/s}
+        $$
+        For the second wavelength (589.59 nm):
+        $$
+            v_{p2} = \frac{3 \times 10^8 \,\text{m/s}}{1.6350} \approx 1.834 \times 10^8 \,\text{m/s}
+        $$
+        \item To find the group velocity \( v_g \), we first need to calculate the average phase velocity \( v \) and average wavelength \( \lambda \):
+        $$            v = \frac{v_{p1} + v_{p2}}{2} \approx 1.834 \times 10^8 \,\text{m/s} $$
+        $$            \lambda = \frac{589.00 \,\text{nm} + 589.59 \,\text{nm}}{2} = 589.295 \,\text{nm} = 5.89295 \times 10^{-7} \,\text{m} $$
+        We are given the $n-\lambda$ data, so we have $\frac{dn}{d\lambda} = -0.0001/0.59 \text{nm} = -1.6949 \times 10^{-7} \text{nm}^{-1} = -1.6949 \times 10^{-16} \text{m}^{-1}$.
+And by chain rule, we have:
+        $$            \frac{dv}{d\lambda} = \frac{dv}{dn} \frac{dn}{d\lambda} = -\frac{c}{n^2} \frac{dn}{d\lambda} = -\frac{v}{n} \frac{dn}{d\lambda} $$
+        Thus, we can calculate the group velocity \( v_g \):
+        $$
+            v_g = v - \lambda \frac{dv}{d\lambda} = v - \lambda \frac{v}{n} \frac{dn}{d\lambda} = 0.5742c \approx 1.722 \times 10^8 \,\text{m/s}   
+        $$
+
+    \end{enumerate}
+\end{example}
+
+\subsection{Wave Pockets}
+\begin{definition}[Wave Packet]
+    A wave packet is a localized disturbance that results from the superposition of multiple waves with different wavelengths and frequencies. It can be represented mathematically as:
+    \begin{equation}
+        y(x,t) = \int_{-\infty}^{\infty} A(k) e^{i(kx - \omega t)} \, dk
+    \end{equation}
+    where \( A(k) \) is the amplitude distribution function in wave number space, \( k \) is the wave number, and \( \omega \) is the angular frequency.
+The wave packet travels with the group velocity \( v_g = \frac{d\omega}{dk} \).
+\end{definition}
+
+\begin{definition}[Width of the Wave Packet]
+    The spatial width \( \Delta x \) of a wave packet is inversely related to the spread in wave numbers \( \Delta k \) of the constituent waves. This relationship is given by:
+    \begin{equation}
+        \Delta x \Delta k \approx 1
+    \end{equation}
+    where \( \Delta x \) is the uncertainty in position and \( \Delta k \) is the uncertainty in wave number.
+A narrower wave packet (smaller \( \Delta x \)) corresponds to a broader range of wave numbers (larger \( \Delta k \)), and vice versa.
+
+\end{definition}
+
 \end{document}

### 2025-10-20 YEAR2/AER210_LectureNotes/main.tex
oct 20

@@
 \end{example}
+\subsection{Divergence and Curl}
+
+\begin{definition}[Divergence]
+    The divergence of a vector field \(\vec{F} = P \, \hat{i} + Q \, \hat{j} + R \, \hat{k}\) is a scalar function defined as:
+    \begin{equation}
+        \nabla \cdot \vec{F} = \frac{\partial P}{\partial x} + \frac{\partial Q}{\partial y} + \frac{\partial R}{\partial z}
+    \end{equation}
+    It measures the rate at which "density" exits a point in the field.
+\end{definition}
+
+\begin{definition}[Curl]
+    The curl of a vector field \(\vec{F} = P \, \hat{i} + Q \, \hat{j} + R \, \hat{k}\) is a vector function defined as:
+    \begin{equation}
+        \nabla \times \vec{F} = \begin{vmatrix}
+            \hat{i} & \hat{j} & \hat{k} \\
+            \frac{\partial}{\partial x} & \frac{\partial}{\partial y} & \frac{\partial}{\partial z} \\
+            P & Q & R
+        \end{vmatrix}
+    \end{equation}
+    It measures the rotation of the field around a point.
+\end{definition}
+
+\begin{theorem}[Properties of Divergence and Curl]
+    Let \(\vec{F}\) and \(\vec{G}\) be vector fields, and \(f\) be a scalar function. Then:
+    \begin{itemize}
+        \item \(\nabla \cdot (f \vec{F}) = f (\nabla \cdot \vec{F}) + \vec{F} \cdot (\nabla f)\)
+        \item \(\nabla \times (f \vec{F}) = f (\nabla \times \vec{F}) + (\nabla f) \times \vec{F}\)
+        \item \(\nabla \cdot (\nabla \times \vec{F}) = 0\)
+        \item \(\nabla \times (\nabla f) = 0\)
+    \end{itemize}
+\end{theorem}
+
+\begin{definition}[Consdervative Vector Field]
+    A vector field \(\vec{F}\) is said to be conservative if its curl is zero:
+    \begin{equation}
+        \nabla \times \vec{F} = 0
+    \end{equation}
+
+\end{definition}
+
+\begin{definition}[Incompressible Vector Field]
+    A vector field \(\vec{F}\) is said to be incompressible if its divergence is zero:
+    \begin{equation}
+        \nabla \cdot \vec{F} = 0
+    \end{equation}
+
+\end{definition}
+
+\subsection{Divergence Theorem}
+\begin{theorem}
+    Let \(E\) be a solid region in \(\mathbb{R}^3\) with a smooth boundary surface \(S\) oriented outward.
+If \(\vec{F}\) is a vector field with continuous partial derivatives on an open region containing \(E\), then:
+    \begin{equation}
+        \iint_S \vec{F} \cdot \vec{n} \, dS = \iiint_E (\nabla \cdot \vec{F}) \, dV
+    \end{equation}
+    and $\vec{n}$ is the unit normal vector to the surface \(S\).
+\end{theorem}
+
+\begin{definition}[Flux]
+    The flux of a vector field \(\vec{F}\) across a surface \(S\) is defined as the surface integral of the normal component of \(\vec{F}\) over \(S\):
+    \begin{equation}
+        \text{Flux} = \iint_S \vec{F} \cdot \vec{n} \, dS
+    \end{equation}
+\end{definition}
+\begin{example}[Flux]
+\textit{Compute the flux of the vector field \(\vec{F}(x,y,z) = x \, \hat{i} + y \, \hat{j} + z \, \hat{k}\) across the surface of the unit sphere \(x^2 + y^2 + z^2 = 1\).}
+
+We can use the Divergence Theorem to compute the flux. Notice the following equality holds under the Divergence Theorem:
+\[\iint_S \vec{F} \cdot \vec{n} \, dS = \iiint_E (\nabla \cdot \vec{F}) \, dV\]
+where \(E\) is the solid region enclosed by the surface \(S\).
+
+First, we compute the divergence of \(\vec{F}\):
+\[
+\nabla \cdot \vec{F} = \frac{\partial x}{\partial x} + \frac{\partial y}{\partial y} + \frac{\partial z}{\partial z} = 1 + 1 + 1 = 3
+\]
+Next, we compute the volume integral over the unit sphere:
+\[
+\iiint_E (\nabla \cdot \vec{F}) \, dV = \iiint_E 3 \, dV = 3 \cdot \text{Volume of unit sphere} = 3 \cdot \frac{4}{3} \pi = 4\pi
+\]
+Thus, the flux of \(\vec{F}\) across the surface of the unit sphere is \(4\pi\).
+
+\end{example}
+\subsection{Stokes Theorem}
+\begin{theorem}[Stokes' Theorem]
+    Let \(S\) be an oriented smooth surface bounded by a simple, closed, piecewise-smooth curve \(C\) with positive orientation.
+If \(\vec{F}\) is a vector field with continuous partial derivatives on an open region containing \(S\), then:
+    \begin{equation}
+        \iint_S (\nabla \times \vec{F}) \cdot \vec{n} \, dS = \oint_C \vec{F} \cdot d\vec{r}
+    \end{equation}
+    where \(\vec{n}\) is the unit normal vector to the surface \(S\).
+
+    In other words, Stokes' theorem is a generalization of Green's theorem to three dimensions.
+\end{theorem}
+
+\begin{example}
+    Let $S$ be the part of the paraboloid $z = 9 - x^2 - y^2$ that lies above the $xy$-plane.
+Let $C$ be the trace of $S$ in the $xy$-plane.
+Let $\vec{F}(x,y,z) = \langle 3z, 4x, 2y \rangle$.
+\textit{Verify Stokes' Theorem by computing both $\iint_S (\nabla \times \vec{F}) \cdot \vec{n} \, dS$ and $\oint_C \vec{F} \cdot d\vec{r}$.}
+
+First, we compute the curl of \(\vec{F}\):
+\[\nabla \times \vec{F} = \begin{vmatrix}
+\hat{i} & \hat{j} & \hat{k} \\
+\frac{\partial}{\partial x} & \frac{\partial}{\partial y} & \frac{\partial}{\partial z} \\
+3z & 4x & 2y
+\end{vmatrix} = \langle 0 - 0, 0 - 3, 4 - 0 \rangle = \langle 0, -3, 4 \rangle\]    
+Next, we parametrize the surface \(S\) using polar coordinates:
+\[\vec{r}(r, \theta) = \langle r \cos \theta, r \sin \theta, 9 - r^2 \rangle\]
+where \(0 \leq r \leq 3\) and \(0 \leq \theta \leq 2\pi\).
+
+We compute the partial derivatives:
+\[\vec{r_r} = \langle \cos \theta, \sin \theta, -2r \rangle\]
+\[\vec{r_\theta} = \langle -r \sin \theta, r \cos \theta, 0 \rangle\]
+Then, we compute the cross product:
+\[\vec{r_r} \times \vec{r_\theta} = \begin{vmatrix}
+\hat{i} & \hat{j} & \hat{k} \\
+\cos \theta & \sin \theta & -2r \\
+-r \sin \theta & r \cos \theta & 0 
+\end{vmatrix} = 2r^2 \cos \theta \hat{i} + 2r^2 \sin \theta \hat{j} + r \hat{k}\]
+The magnitude of the cross product is:
+\[||\vec{r_r} \times \vec{r_\theta}|| = \sqrt{(2r^2 \cos \theta)^2 + (2r^2 \sin \theta)^2 + r^2} = \sqrt{4r^4 + r^2} = r \sqrt{4r^2 + 1}\]
+Now, we can compute the surface integral:
+\[\iint_S (\nabla \times \vec{F}) \cdot \vec{n} \, dS = \iint_D (\nabla \times \vec{F}) \cdot (\vec{r_r} \times \vec{r_\theta}) \, dr \, d\theta\]
+\begin{align*}
+&= \int_0^{2\pi} \int_0^3 \langle 0, -3, 4 \rangle \cdot \langle 2r^2 \cos \theta, 2r^2 \sin \theta, r \rangle \, dr \, d\theta \\
+&= \int_0^{2\pi} \int_0^3 (-6r^2 \sin \theta + 4r) \, dr \, d\theta \\
+&= \int_0^{2\pi} \left[ -2r^3 \sin \theta + 2r^2 \right]_0^3 \, d\theta \\
+&= \int_0^{2\pi} (-54 \sin \theta + 18) \, d\theta \\
+&= \left[ 54 \cos \theta + 18 \theta \right]_0^{2\pi} = 36\pi   
+\end{align*}
+
+Next, we compute the line integral around the curve \(C\). The curve \(C\) is the circle \(x^2 + y^2 = 9\) in the \(xy\)-plane. We can parametrize \(C\) as:
+\[\vec{r}(t) = \langle 3 \cos t, 3 \sin t, 0 \rangle\]
+where \(0 \leq t \leq 2\pi\).
+We compute \(d\vec{r}\):
+\[d\vec{r} = \langle -3 \sin t, 3 \cos t, 0 \rangle \, dt\]
+Now, we can compute the line integral:
+\[\oint_C \vec{F} \cdot d\vec{r} = \int_0^{2\pi} \vec{F}(\vec{r(t)}) \cdot d\vec{r}\]
+\begin{align*}
+&= \int_0^{2\pi} \langle 0, 12 \cos t, 6 \sin t \rangle \cdot \langle -3 \sin t, 3 \cos t, 0 \rangle \, dt \\   
+&= \int_0^{2\pi} (0 + 36 \cos^2 t + 0) \, dt \\
+&= 36 \int_0^{2\pi} \cos^2 t \, dt \\
+&= 36 \cdot \pi = 36\pi
+\end{align*}        
+Both the surface integral and the line integral yield the same result, confirming Stokes' Theorem.
+\end{example}
+
 \chapter{Fluid Mechanics}
@@
 \end{definition}
+\subsection{Basic Concepts in Fluid Mechanics}
+\begin{definition}[Continuum Hypothesis]
+    The continuum hypothesis assumes that fluids can be treated as continuous media, meaning that their properties (such as density, pressure, and velocity) vary smoothly and continuously throughout the fluid, without abrupt changes or discontinuities.
+This allows us to use calculus and differential equations to model fluid behavior.
+\end{definition}
+\begin{definition}[Stress Field and Stress Tensor]
+    A \textbf{stress field} is a vector field that describes the distribution of internal forces (stresses) within a fluid at every point in space.
+
+    The \textbf{stress tensor} is a mathematical representation of the stress field, which encapsulates the normal and shear stresses acting on infinitesimal fluid elements in all directions.
+It is typically represented as a 3x3 matrix as follows:
+    \[\mathbf{T} = \begin{bmatrix}
+    \sigma_{xx} & \tau_{xy} & \tau_{xz} \\
+    \tau_{yx} & \sigma_{yy} & \tau_{yz} \\
+    \tau_{zx} & \tau_{zy} & \sigma_{zz}
+    \end{bmatrix}\]
+    where \(\sigma_{ii}\) are the normal stresses and \(\tau_{ij}\) are the shear stresses.
+
+\end{definition}
+
+\subsection{Viscosity}
+\begin{definition}[Viscosity]
+    Viscosity is a measure of a fluid's resistance to deformation or flow.
+It quantifies the internal friction within the fluid, which arises from the interactions between its molecules.
+A fluid with high viscosity (like honey) resists motion and flows slowly, while a fluid with low viscosity (like water) flows easily.
+\end{definition}
+
+\begin{definition}[Newtonian Fluid]
+    A Newtonian fluid is a type of fluid whose viscosity remains constant regardless of the applied shear rate.
+In other words, the relationship between shear stress and shear rate is linear.
+Mathematically, this can be expressed as:
+    \begin{equation}
+        \tau_\text{newtonian}= \mu \frac{du}{dy}
+    \end{equation}
+    where \(\tau_\text{newtonian}\) is the shear stress, \(\mu\) is the dynamic viscosity (a constant for Newtonian fluids), and \(\frac{du}{dy}\) is the shear rate (the rate of change of velocity with respect to distance in the direction perpendicular to the flow).
+
+    A non-Newtonian fluid, on the other hand, exhibits a non-linear relationship between shear stress and shear rate, meaning that its viscosity can change with the applied shear rate.
+Examples of non-Newtonian fluids include ketchup, blood, and certain polymers.
+Non-newtonian fluids can be further classified into various types:
+    \begin{itemize}
+        \item \textbf{Shear-thinning fluids (pseudoplastic)} Viscosity decreases with increasing shear rate (e.g., paint, blood, ketchup).
+        \item \textbf{Shear-thickening fluids (dilatant)} Viscosity increases with increasing shear rate (e.g., cornstarch in water).
+        \item \textbf{Bingham plastics} Behave as a solid until a certain yield stress is exceeded, after which they flow like a viscous fluid (e.g., toothpaste, mud).
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Temperature Dependency of Viscosity]
+    The viscosity of fluids is generally temperature-dependent. The behavours differs between liquids and gases:
+    \begin{itemize}
+        \item \textbf{Liquids} As temperature increases, the viscosity of liquids typically decreases.
+This is because higher temperatures provide more thermal energy to the molecules, allowing them to overcome intermolecular forces more easily, resulting in reduced resistance to flow.
+        \item \textbf{Gases} As temperature increases, the viscosity of gases typically increases.
+This is because higher temperatures lead to more frequent and energetic molecular collisions, which enhances the momentum transfer between layers of the gas, resulting in increased resistance to flow.
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Kinematic Viscosity]
+    Kinematic viscosity is a measure of a fluid's resistance to flow under the influence of gravity. It is defined as the ratio of dynamic viscosity to fluid density:
+    \begin{equation}
+        \nu = \frac{\mu}{\rho}
+    \end{equation}
+    where \(\nu\) is the kinematic viscosity, \(\mu\) is the dynamic viscosity, and \(\rho\) is the fluid density.
+\end{definition}
+
+\subsection{Compressibility of Fluids}
+\begin{subequations}
+\begin{definition}[Compressibility]
+    Compressibility is a measure of the change in volume of a fluid in response to a change in pressure. It is quantified by the \textbf{bulk modulus} \(E_v\), which is defined as:
+    \begin{equation}
+        E_v = -V \frac{dP}{dV}
+    \end{equation}
+    where \(V\) is the volume of the fluid, \(P\) is the pressure, and \(\frac{dP}{dV}\) is the rate of change of pressure with respect to volume.
+A fluid with high compressibility can be easily compressed, while a fluid with low compressibility is more resistant to compression.
+
+    The negative sign indicates that an increase in pressure typically results in a decrease in volume.
+\end{definition}
+
+\begin{definition}[Bulk Modulus]
+    The bulk modulus $E_v$ also can be written in terms of density rather than volume:
+    \begin{equation}
+        E_v = \rho \frac{dP}{d\rho}
+    \end{equation}
+\end{definition}
+
+In other words, the larger the bulk modulus, the less compressible the fluid is.
+For water, the bulk modulus is approximately \(2.2 \times 10^9 \, \text{Pa}\) at $20^\circ$C, indicating that it is relatively incompressible.
+
+\end{subequations}
+
+\begin{example}[Compressibility of Water]
+    \textit{Calculate the pressure needed to compress water by 0.05\% at $20^\circ$C.}
+
+    Given that the bulk modulus of water at $20^\circ$C is approximately \(E_v = 2.2 \times 10^9 \, \text{Pa}\), we can use the definition of bulk modulus to find the required pressure change:
+    \[E_v = -V \frac{dP}{dV} \implies dP = -E_v \frac{dV}{V}\]
+    Since we want to compress water by 0.05\%, we have:
+    \[\frac{dV}{V} = -0.0005\]
+    Substituting this into the equation for \(dP\):
+    \[dP = - (2.2 \times 10^9 \,\text{Pa}) \times (-0.0005) = 1.1 \times 10^6 \, \text{Pa}\]
+    Therefore, a pressure increase of approximately 1.1 MPa is needed to compress water by 0.05\% at $20^\circ$C.
+\end{example}
+In fact, we often take water as a nearly incompressible fluid in fluid mechanics problems due to its high bulk modulus.
+
+\begin{definition}[Incompressible Fluid]
+    An incompressible fluid is a fluid whose density remains constant regardless of changes in pressure or temperature.
+In other words, the volume of an incompressible fluid does not change significantly when subjected to external forces.
+Mathematically, this can be expressed as:
+    $$
+        \frac{d\rho}{dt} = 0
+    $$
+    where \(\rho\) is the density of the fluid and \(t\) is time. In some cases we would assume incompressibility if the density changes are negligible.
+
+\end{definition}
+
+\section{Dimensional Analysis and Similitude}
+\begin{definition}[Dimensional Analysis]
+    Dimensional analysis is a mathematical technique used to analyze the relationships between physical quantities by examining their dimensions (units of measurement).
+It helps to identify the fundamental variables that govern a physical phenomenon and to derive dimensionless groups that can be used to simplify complex problems.
+\end{definition}
+
+\section{Hydrostatics}
+\begin{example}
+    A cylindrical tank of height \(h\) contains a rubber membrane at its base.
+When the tank is filled with water to a height \(H\), the membrane deflects by an amount \(d_c\) due to the hydrostatic pressure.
+Consider now a similar setup using a smaller-diameter cylinder but the \emph{same} membrane and the same water height \(H\).
+Let the corresponding deflection be \(d_s\).
+Empirically, it is found that
+    \[
+    d_c = d_s.
+    \] 
+    This section would like to explain why this is the case.
+\end{example}
+
+\begin{definition}[Hydrostatic]
+    Hydrostatics is the branch of fluid mechanics that studies fluids at rest or in rigid body motion.
+It focuses on the behavior of fluids under the influence of gravity and other external forces, and it examines the pressure distribution, buoyancy, and stability of fluids in various configurations.
+
+    We cosndier two main forces acting on a fluid at rest:
+    \begin{itemize}
+        \item \textbf{Body Forces}: These are forces that act on the entire volume of the fluid, such as gravitational forces. The body force per unit volume is given by:
+        \[\vec{f_b} = \rho \vec{g}\]
+        where \(\rho\) is the fluid density and \(\vec{g}\) is the acceleration due to gravity.
+        \item \textbf{Surface Forces}: These are forces that act on the surface of the fluid, such as pressure forces. The surface force per unit area is given by:
+        \[\vec{f_s} = -P \vec{n}\]
+        where \(P\) is the pressure and \(\vec{n}\) is the unit normal vector to the surface.
+    \end{itemize}
+    We, however, DO NOT consider shear forces since the fluid is at rest.
+\end{definition}
+
+\subsection{Pressure in a Fluid at Rest}
+\begin{definition}[Pressure]
+    Pressure is the result of molecular collisions on a real or imaginary surface.
+\end{definition}
+
+\paragraph{Sign Convention} A negative force indicates a compressive force (acts on the molecules by the surface), while a positive force indicates a tensile force (acts on the surface by the molecules).
+
+\begin{theorem}[Surface Pressure]
+    Consider a surface that is has complicated geometry.
+We pressure acting on the surface is the inward acting normal force per unit area.
+Thus, we ignore shear forces and only consider normal forces.
+Mathematically, we write:
+    \begin{equation}
+        P = -\frac{dF_n}{dA}
+    \end{equation}
+\end{theorem}
+
+\begin{theorem}[Pascal's Law]
+    The pressure at a point in a hydrostatic fluid is the same in all directions. Mathematically, we write:
+    \begin{equation}
+        P = P(x,y,z)
+    \end{equation}
+\end{theorem}
+
+\begin{proof}
+    Consider a small fluid element in the shape of a tetrahedron with one face oriented in the \(xy\)-plane, one face in the \(yz\)-plane, and one face in the \(xz\)-plane.
+Let the area of the face in the \(xy\)-plane be \(dA\), and let the areas of the other two faces be \(dA_x\) and \(dA_y\).
+The height of the tetrahedron is \(dz\).
+
+    The forces acting on the tetrahedron are:
+    \begin{itemize}
+        \item The pressure force on the face in the \(xy\)-plane: \(P \cdot dA\)
+        \item The pressure force on the face in the \(yz\)-plane: \(P_x \cdot dA_x\)
+        \item The pressure force on the face in the \(xz\)-plane: \(P_y \cdot dA_y\)
+    \end{itemize}
+
+    Since the fluid is at rest, the net force on the tetrahedron must be zero. Therefore, we have:
+    \[
+    P \cdot dA = P_x \cdot dA_x + P_y \cdot dA_y
+    \]
+    Dividing both sides by \(dA\), we get:
+    \[
+    P = P_x \frac{dA_x}{dA} + P_y \frac{dA_y}{dA}
+    \]
+    As we let the dimensions of the tetrahedron approach zero, we find that:
+    \[
+    P = P_x = P_y
+    \]
+    Thus, we conclude that the pressure at a point in a hydrostatic fluid is the same in all directions.
+\end{proof}
+
+\subsection{Basic Eqaution for Preessure Fiekd}
+\begin{theorem}[Hydrostatic Pressure Equation]
+    In a fluid at rest, the pressure gradient is related to the body forces acting on the fluid. Mathematically, we write:
+    \begin{equation}
+        \nabla P = \rho \vec{g}
+    \end{equation}
+    where \(\nabla P\) is the pressure gradient, \(\rho\) is the fluid density, and \(\vec{g}\) is the acceleration due to gravity. 
+\end{theorem}
 \end{document}

### 2025-10-20 YEAR2/CHE260_LectureNotes/main.tex
oct 20

@@
 \chapter{Heat Transfer}
-
+\section{Modes of Heat Transfer}
+\subsection{Conduction}
+\begin{definition}[Conduction]
+    Conduction is the transfer of heat through a solid material without the movement of the material itself.
+It occurs due to the vibration and collision of atoms and molecules within the material.
+The rate of heat transfer by conduction is given by Fourier's law:
+    \begin{equation}
+        \dot{Q} = -k A \frac{dT}{dx}
+    \end{equation}
+    where $\dot{Q}$ is the heat transfer rate, $k$ is the thermal conductivity of the material, $A$ is the cross-sectional area through which heat is being transferred, and $\frac{dT}{dx}$ is the temperature gradient in the direction of heat transfer.
+\end{definition}
+\begin{definition}[Heat Flux]
+    Heat flux is the rate of heat transfer per unit area. It is given by:
+    \begin{equation}
+        q'' = \frac{\dot{Q}}{A} = -k \frac{dT}{dx}
+    \end{equation}
+    where $q''$ is the heat flux, $\dot{Q}$ is the heat transfer rate, $A$ is the cross-sectional area, $k$ is the thermal conductivity, and $\frac{dT}{dx}$ is the temperature gradient.
+
+\end{definition}
+
+\begin{definition}[Fourier's Law]
+    Fourier's law states that the rate of heat transfer through a material is proportional to the negative gradient of temperature and the area through which the heat is being transferred.
+Mathematically, it is expressed as:
+    \begin{equation}
+        \dot{Q} = -k A \frac{dT}{dx}
+    \end{equation}
+    where $\dot{Q}$ is the heat transfer rate, $k$ is the thermal conductivity of the material, $A$ is the cross-sectional area, and $\frac{dT}{dx}$ is the temperature gradient.
+
+    In other words, heat flows from regions of higher temperature to regions of lower temperature, and the rate of heat transfer is influenced by the material's thermal conductivity and the geometry of the system.
+\end{definition}
+\begin{example}
+   \textit{ A cone frustum is conducting heat at staady-state with no generation.
+If the frustum is insulated, and the there is heat flux onto the smaller face, the heat flux in would be \underline{\hspace{2cm}} the heat flux out:}
+    \begin{itemize}
+        \item Equal to
+        \item Greater than
+        \item Less than
+    \end{itemize}
+
+    \textbf{Answer:} Greater than. We can set up the energy balance:
+    \begin{align*}
+        \dot{Q}_\text{in} - \dot{Q}_\text{out} + \dot{Q}_\text{gen} &= 0 \\
+        \dot{q}_{in} A_{in} - \dot{q}_\text{out} A_\text{out} + 0 &= 0 \\
+        \dot{q}_\text{in} A_\text{in} &= \dot{q}_\text{out} A_\text{out} \\
+        \dot{q}_\text{out} &= \dot{q}_\text{in} \frac{A_\text{in}}{A_\text{out}} \\
+        &< \dot{q}_\text{in} \quad \text{since } A_\text{out} > A_\text{in}
+    \end{align*}
+\end{example}
+
+\begin{definition}[Common Thermal Conductors]
+    Common thermal conductors and their thermal conductivities ($k$) at room temperature are:
+    \begin{itemize}
+        \item Copper: $k \approx 400\, \text{W/m·K}$
+        \item Aluminum: $k \approx 237\, \text{W/m·K}$
+        \item Steel: $k \approx 50\, \text{W/m·K}$
+        \item Glass: $k \approx 1\, \text{W/m·K}$
+        \item Wood: $k \approx 0.1 - 0.2\, \text{W/m·K}$
+        \item Diamond: $k \approx 2200\, \text{W/m·K}$
+    \end{itemize}
+\end{definition}
+
+Diamond conducts well because of its crystalline structure, but why does metal conduct well?
+Metals have free electrons that can move easily through the lattice structure, allowing them to transfer thermal energy quickly.
+The free electrons collide with atoms and other electrons, facilitating the transfer of heat.
+This is why metals generally have high thermal conductivity compared to non-metals, which rely primarily on lattice vibrations (phonons) for heat transfer.
+\begin{definition}[Electron Conduction]
+    In metals, heat conduction occurs primarily through the movement of free electrons.
+These electrons can move freely throughout the metal lattice and transverse quickly through different energy states.
+When a temperature gradient is present, the higher-energy electrons from the hotter region move towards the cooler region, transferring their kinetic energy to the lattice atoms through collisions.
+This process results in efficient heat conduction in metals.
+\end{definition}
 \end{document}

### 2025-10-20 YEAR2/ECE253_LectureNotes/main.tex
oct 20

@@
 \section{Computer Organization}
-\section{Assembly Language}
+\section{RISC-V Assembly Language}
+\begin{definition}[Arithmetic Logical Unit]
+    An Arithmetic Logical Unit (ALU) is a digital circuit that performs arithmetic and logical operations on binary numbers.
+It is a fundamental building block of the central processing unit (CPU) in a computer.
+The ALU takes two binary inputs, performs the specified operation, and produces a binary output.
+
+    The ALU can perform a variety of operations, including:
+    \begin{itemize}
+        \item Arithmetic operations: addition, subtraction, multiplication, division
+        \item Logical operations: AND, OR, NOT, XOR
+        \item Bitwise operations: shift left, shift right
+        \item Comparison operations: equal to, less than, greater than
+    \end{itemize}
+    The specific operations supported by an ALU depend on its design and the architecture of the computer it is used in.
+\end{definition}
+
+\begin{definition}[Instructions]
+    Instructions are the basic commands that a CPU can execute.
+They are typically represented in assembly language, which is a low-level programming language that is specific to a particular computer architecture.
+Each instruction consists of an operation code (opcode) and operands, which specify the data to be operated on.
+
+    In RISC-V architecture, instructions are 32 bits long and are divided into different formats based on the type of operation being performed.
+The most common instruction formats are R-type, I-type, S-type, B-type, U-type, and J-type.
+Each format has a specific layout for the opcode and operands.
+
+    RISC-V instructions can perform a variety of operations, including arithmetic operations (e.g., addition, subtraction), logical operations (e.g., AND, OR), data movement (e.g., load, store), and control flow (e.g., branch, jump).
+
+\end{definition}
+
+\begin{definition}[Registers]
+    Registers are small, high-speed storage locations within a CPU that hold data temporarily during the execution of instructions.
+They are used to store operands, intermediate results, and memory addresses.
+Registers are faster to access than main memory, which allows for quicker data manipulation and processing.
+
+    In RISC-V architecture, there are 32 general-purpose registers (x0 to x31), each 32 bits wide in the RV32I variant.
+The register x0 is hardwired to zero and cannot be modified.
+The other registers can be used for various purposes, such as holding function arguments, return values, and temporary data.
+\end{definition}
+
+\begin{definition}[Intermediates]
+    Intermediates are temporary values that are used during the execution of instructions.
+They are typically stored in registers and are used to hold the results of operations before they are written back to memory or used as inputs to other operations.
+\end{definition}
+
+\begin{definition}[Operahands and Instruction Operations]
+    The following is a table of RISC-V instructions and the operations they perform:
+
+    The registers are \verb|rX| and \verb|rY|, and \verb|D| is a 12 bit value.
+
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{|c|c|c|}
+            \hline
+            Instruction & Operation & Description \\
+            \hline
+            MV rX, rY & rX = rY & Move value from rY to rX \\
+            MV rX, #D & rX = D & Load sign extended immediate D into rX \\
+            \hline
+            ADD rX, rY & rX = rX + rY & Add rX and rY, store result in rX \\
+            ADD rX, #D & rX = rX + D & Add immediate D to rX, store result in rX \\
+            \hline 
+            SUB rX, rY & rX = rX - rY & Subtract rY from rX, store result in rX \\
+            SUB rX, #D & rX = rX - D & Subtract immediate D from rX, store result in rX \\
+            \hline
+            MUL rX, rY & rX = rX * rY & Multiply rX and rY, store result in rX \\
+            \hline
+            DIV rX, rY & rX = rX / rY & Divide rX by rY, store quotient in rX \\
+            \hline
+        \end{tabular}
+        \caption{RISC-V Arithmetic Instructions}
+        \label{tab:riscv_arithmetic_instructions}
+    \end{table}
+\end{definition}
+\begin{definition}[Instruction Encoding]
+    Encodings are 16-bit or 32-bit binary representations of instructions that the CPU can understand and execute.
+Each instruction is represented by a unique binary code that specifies the operation to be performed and the operands involved.
+The encoding of an instruction typically includes the following components:
+    \begin{itemize}
+        \item \textbf{Opcode:} A unique binary code that identifies the operation to be performed (e.g., addition, subtraction, load, store).
+        \item \textbf
+\end{definition}
 \end{document}

### 2025-10-20 YEAR2/MAT292_LectureNotes/main.tex
oct 20

@@
 \end{example}
-
+\subsection{Second Order Homogeneous ODE}
 \begin{definition}[Characteristic Equation]
@@
 
+\subsection{Second Order Inhomogeneous ODE}
+\begin{theorem}[General Solution of Inhomogeneous ODE]
+    The general solution of a second-order inhomogeneous ODE:
+    $$y'' + p y' + q y = \sum_i g_i(t)$$
+    is given by:
+    \begin{equation}
+    y(t) = c_1 y_1(t) + c_2 y_2(t) + \sum_i y_{p_i}(t)
+    \end{equation}
+    where \(c_1, c_2\) are constants determined by initial conditions, \( y_h(t) = \{ y_1(t) , y_2(t) \} \) is the general solution of the corresponding homogeneous equation \( y'' + p y' + q y = 0 \), and \( y_{p_i}(t) \) are particular solutions corresponding to each \( g_i(t) \).
+Each \( y_{p_i}(t) \) satisfies:
+    \begin{enumerate}
+        \item $\text{Null}(L) = \{y_h\}$
+        \item $L(y_{p_i}) = g_i(t)$
+    \end{enumerate}
+\end{theorem}
+\begin{theorem}[Uniqueness of Particular Solutions]
+    If \( y_{p_1}(t) \) and \( y_{p_2}(t) \) are two particular solutions of the inhomogeneous ODE:
+    $$y'' + p y' + q y = g(t)$$
+    then their difference \( y_{p_1}(t) - y_{p_2}(t) \) is a solution of the corresponding homogeneous equation:
+    $$y'' + p y' + q y = 0$$
+
+    And since the homongenous solution is unique up to a linear combination of basis solutions, we have:
+    $$y_{p_1}(t) - y_{p_2}(t) = c_1 y_1(t) + c_2 y_2(t)$$
+    where \( c_1 \) and \( c_2 \) are constants, and \( y_1(t) \) and \( y_2(t) \) are the fundamental solutions of the corresponding homogeneous equation.
+\end{theorem}
+\begin{proof}
+    We have:
+    \begin{align*}
+        L(y_{p_1}) &= g(t) \\
+        L(y_{p_2}) &= g(t)
+    \end{align*}
+    Subtracting these two equations gives:
+    \begin{align*}
+        L(y_{p_1}) - L(y_{p_2}) &= g(t) - g(t) \\
+        L(y_{p_1} - y_{p_2}) &= 0
+    \end{align*}
+    Thus, \( y_{p_1}(t) - y_{p_2}(t) \) is indeed a solution of the corresponding homogeneous equation.
+\end{proof}
+
+\begin{definition}[Wronskian Derivation of $y_p$]
+    If $y_1(t)$ and $y_2(t)$ are two linearly independent solutions of the corresponding homogeneous equation (a.k.a.
+the fundamental set of homogeneous solutions), then we can assume a particular solution of the form:
+    $$y_p(t) = -y_1(t) \int \frac{y_2(t) g(t)}{W(t)} dt + y_2(t) \int \frac{y_1(t) g(t)}{W(t)} dt$$
+    where \( W(t) = y_1(t)y_2'(t) - y_2(t)y_1'(t) \) is the Wronskian determinant of \( y_1 \) and \( y_2 \).
+\end{definition}
+
+\begin{example}
+    Consider the following second-order inhomogeneous ODE:
+    $$y'' - 3y'- 4y = 3e^{2t} + \sin(t)$$
+    The corresponding homogeneous equation is:
+    $$y'' - 3y' - 4y = 0$$
+    The characteristic equation is:
+    $$r^2 - 3r - 4 = 0$$
+    Solving this gives the roots \( r_1 = 4 \) and \( r_2 = -1 \). Thus, the general solution of the homogeneous equation is:
+    $$y_h(t) = C_1 e^{4t} + C_2 e^{-t}$$
+    where \( C_1 \) and \( C_2 \) are constants determined by initial conditions.
+
+    Next, we find particular solutions for each term on the right-hand side of the inhomogeneous equation.
+For the term \( 3e^{2t} \), we can assume a particular solution of the form $y_{p_1}(t) = A e^{2t}$.
+Substituting this into the left-hand side of the ODE gives:
+    \begin{align*}
+        y_{p_1}'' - 3y_{p_1}' - 4y_{p_1} &= 4A e^{2t} - 6A e^{2t} - 4A e^{2t} \\
+        &= -6A e^{2t}
+    \end{align*}
+    Setting this equal to \( 3e^{2t} \) gives:
+    $$-6A e^{2t} = 3e^{2t} \implies A = -\frac{1}{2}$$
+    Thus, we have:
+    $$y_{p_1}(t) = -\frac{1}{2} e^{2t}$$
+    For the term \( \sin(t) \), we can assume a particular solution of the form \( y_{p_2}(t) = B \cos(t) + C \sin(t) \). Substituting this into the left-hand side of the ODE gives:
+    \begin{align*}
+        y_{p_2}'' - 3y_{p_2}' - 4y_{p_2} &= -B \cos(t) - C \sin(t) + 3B \sin(t) - 3C \cos(t) - 4B \cos(t) - 4C \sin(t) \\
+        &= (-5B - 3C) \cos(t) + (3B - 5C) \sin(t)
+    \end{align*}
+    Setting this equal to \( \sin(t) \) gives the system of equations:
+    \begin{align*}
+        -5B - 3C &= 0 \\
+        3B - 5C &= 1
+    \end{align*}
+    Solving this system gives \( B = -\frac{3}{34} \) and \( C = \frac{5}{34} \). Thus, we have:
+    $$y_{p_2}(t) = -\frac{3}{34} \cos(t) + \frac{5}{34} \sin(t)$$
+    Finally, the general solution of the inhomogeneous ODE is:
+    $$y(t) = C_1 e^{4t} + C_2 e^{-t} - \frac{1}{2} e^{2t} - \frac{3}{34} \cos(t) + \frac{5}{34} \sin(t)$$
+\end{example}
+
+\paragraph{Operator Theory} We call the differential operator \( L = D^2 + pD + qI \) where \( D \) is the differentiation operator and \( I \) is the identity operator.
+Then, we would call $L: C^2(\mathbb{R}) \to C(\mathbb{R})$ a transformation from the space of twice differentiable functions to the space of continuous functions.
+We can consider the factorization of this operator:
+$$L = (D - \lambda_1 I)(D - \lambda_2 I)$$
+where \( \lambda_1 \) and \( \lambda_2 \) are the roots of the characteristic equation.
+
+Then, we can consider the eigenvalues and eigenfunctions of this operator:
+\begin{definition}[Eigenvalues and Eigenfunctions of Differential Operators]
+    An eigenvalue \( \lambda \) of a differential operator \( L \) is a scalar such that there exists a non-trivial function \( y(t) \) (called an eigenfunction) satisfying the equation:
+    \begin{equation}
+    L[y(t)] = \lambda y(t)
+    \end{equation}
+    In the context of second-order linear ODEs, if we consider the operator \( L = D^2 + pD + qI \), then the eigenvalue problem becomes:
+    \begin{equation}
+    y'' + py' + qy = \lambda y
+    \end{equation}
+    The solutions to this equation can provide insights into the behavior of the system described by the differential equation, including stability and oscillatory properties.
+\end{definition}
+
+\begin{example}[Eigenfunctions of $D$]
+    Consider the differential operator \( D \) defined by \( D[y(t)] = y'(t) \). We want to find the eigenvalues and eigenfunctions of this operator. The eigenvalue equation is given by:
+    $$D[y(t)] = \lambda y(t)$$
+    which translates to the first-order ODE, solving this gives:
+    $$ f_\lambda = e^{\lambda t} $$
+    where \( \lambda \) is any complex number. Thus, the eigenfunctions of the operator \( D \) exists in a dense, uncoutable spectrum.
+
+    If we impose boundary conditions, for example, \( y(0) = 1 = y(2\pi) \), then we have:
+    $$ e^{\lambda 2\pi} = 1 \implies \lambda = ik, k \in \mathbb{Z} $$
+    Thus, the eigenvalues are discrete and countable, and the corresponding eigenfunctions are:
+    $$ f_k(t) = e^{ikt}, k \in \mathbb{Z} $$
+    which form a basis for the space of periodic functions with period \( 2\pi \).
+\end{example}
+
+\begin{definition}[Fourier Basis]
+    The set of functions \( \{ e^{ikt} \mid k \in \mathbb{Z} \} \) forms an orthogonal basis for the space of square-integrable functions on the interval \( [0, 2\pi] \).
+This basis is known as the Fourier basis.
+Any function \( f(t) \) in this space can be expressed as a Fourier series:
+    \begin{equation}
+    f(t) = \sum_{k=-\infty}^{\infty} c_k e^{ikt}
+    \end{equation}
+    where the coefficients \( c_k \) are given by:
+    \begin{equation}
+    c_k = \frac{1}{2\pi} \int_0^{2\pi} f(t) e^{-ikt} dt
+    \end{equation}
+    The Fourier basis is widely used in various fields, including signal processing, heat transfer, and quantum mechanics, due to its ability to represent periodic functions and analyze frequency components.
+
+\end{definition}
+
+\begin{definition}[Generalized Eigenfunction]
+    A generalized eigenfunction of a differential operator \( L \) associated with an eigenvalue \( \lambda \) is a function \( y(t) \) that satisfies the equation:
+    \begin{equation}
+    (L - \lambda I)^m[y(t)] = 0
+    \end{equation}
+    for some positive integer \( m \).
+Here, \( I \) is the identity operator, and \( m \) is the order of the generalized eigenfunction.
+Generalized eigenfunctions arise in cases where the operator \( L \) has repeated eigenvalues, and they extend the concept of eigenfunctions to include functions that may not satisfy the standard eigenvalue equation but still provide valuable information about the structure of the solution space.
+
+\end{definition}
+
+\begin{definition}[Spectrum]
+    The spectrum of a differential operator \( L \) is the set of all eigenvalues \( \lambda \) for which there exists a non-trivial solution \( y(t) \) (eigenfunction):
+    \begin{equation}
+        \mathrm{spec}(L) = \{ \lambda \in \mathbb{C} \mid L[y(t)] = \lambda y(t) \text{ has a non-trivial solution } y(t) \}
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Variation of Parameters]
+    Variation of parameters is a method used to find a particular solution to a non-homogeneous linear differential equation. For a second-order ODE of the form:
+    $$y'' + p(t)y' + q(t)y = g(t)$$
+    where \( y_1(t) \) and \( y_2(t) \) are two linearly independent solutions of the corresponding homogeneous equation, a particular solution \( y_p(t) \) can be expressed as:
+    \begin{equation}
+    y_p(t) = u_1(t)y_1(t) + u_2(t)y_2(t)
+    \end{equation}
+    where \( u_1(t) \) and \( u_2(t) \) are functions to be determined. These functions satisfy the system of equations:
+    \begin{align*}
+        u_1'(t)y_1(t) + u_2'(t)y_2(t) &= 0 \\
+        u_1'(t)y_1'(t) + u_2'(t)y_2'(t) &= g(t)
+    \end{align*}
+    Solving this system allows us to find \( u_1(t) \) and \( u_2(t) \), and thus the particular solution \( y_p(t) \).
+
+\end{definition}
 \end{document}

### 2025-10-20 YEAR2/PHY293_LectureNotes/main.tex
oct 20

--- 2025-10-08 YEAR2/PHY293_LectureNotes/main.tex
@@
 \end{definition}
+
+\chapter{Modern Physics}
+\section{Quantum Mechanics}
+\begin{definition}[Quantization]
+    Quantization is the constraint that energy and other physical properties can only take on discrete values (\textit{quanta}) rather than a continuous range.
+\end{definition}
+\begin{definition}[Probabilistic Framework]
+    Predictions in quantum mechanics are inherently probabilistic. The exact outcome of a measurement cannot be determined, only the probability of different outcomes.
+\end{definition}
+
+\begin{definition}[Black Body]
+    A black body is an idealized physical object that absorbs all incident electromagnetic radiation, regardless of frequency or angle of incidence.
+It also emits radiation in a characteristic spectrum that depends solely on its temperature.
+\end{definition}
+\begin{definition}[Classical View of Black Body Radiation]
+    According to classical physics, Reyleigh-Jeans law predicts that the energy density of black body radiation increases indefinitely with frequency, leading to the "ultraviolet catastrophe." The classical formula is given by:
+    $$    u(\nu, T) = \frac{8\pi \nu^2 k_B T}{c^3} $$
+    where \( u(\nu, T) \) is the energy density per unit frequency, \( \nu \) is the frequency, \( k_B \) is the Boltzmann constant, \( T \) is the absolute temperature, and \( c \) is the speed of light.
+\end{definition}
+\begin{definition}[Quantum View of Black Body Radiation]
+    Max Planck resolved the ultraviolet catastrophe by proposing that electromagnetic energy is quantized.
+He introduced the concept that energy is emitted or absorbed in discrete packets called \textbf{quanta}, with energy given by:
+    \begin{equation}
+        E_n = n h f \quad n = 0, 1, 2, \ldots
+    \end{equation}
+    where \( h \) is Planck's constant and \( f \) is the frequency of the radiation. This leads to Planck's law for black body radiation. The resulting law for the black body spectrum is:
+    \begin{equation}
+        u(\nu, T) = \frac{8\pi h \nu^3}{c^3} \frac{1}{e^{\frac{h\nu}{k_B T}} - 1}
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Photoelectric Effect]
+    The photoelectric effect is the phenomenon where electrons are emitted from a material (usually a metal) when it is exposed to light of sufficient frequency.
+The key observations of the photoelectric effect are:
+    \begin{itemize}
+        \item Electrons are emitted only if the incident light has a frequency above a certain threshold frequency \( f_0 \).
+        \item The kinetic energy of the emitted electrons depends on the frequency of the incident light, not its intensity.
+        \item The number of emitted electrons is proportional to the intensity of the incident light, provided the frequency is above the threshold.
+    \end{itemize}
+    The energy of the incident photons is given by:
+    $$
+        E = h f
+    $$
+    where \( h \) is Planck's constant and \( f \) is the frequency of the light. The maximum kinetic energy \( K_{max} \) of the emitted electrons is given by:
+    $$
+        K_{max} = h f - \phi
+    $$
+    where \( \phi \) is the work function of the material, representing the minimum energy required to liberate an electron from the surface.
+
+\end{definition}
+
+\begin{definition}[Campton Effect]
+    The Compton effect is the increase in wavelength (or decrease in energy) of X-rays or gamma rays when they are scattered by electrons.
+This phenomenon provides evidence for the particle-like behavior of light.
+The change in wavelength \( \Delta \lambda \) of the scattered photon is given by the Compton formula:
+    \begin{equation}
+        \Delta \lambda = \lambda' - \lambda = \frac{h}{m_e c} (1 - \cos \theta)
+    \end{equation}
+    where:
+    \begin{itemize}
+        \item \( \lambda \) is the initial wavelength of the photon,
+        \item \( \lambda' \) is the wavelength after scattering,
+        \item \( h \) is Planck's constant,
+        \item \( m_e \) is the rest mass of the electron,
+        \item \( c \) is the speed of light,
+        \item \( \theta \) is the angle at which the photon is scattered.
+    \end{itemize}
+
+\end{definition}
 \end{document}

### 2025-11-03 YEAR2/ECE253_LectureNotes/main.tex
nov 3

@@
 \end{definition}
+
 \chapter{Computer Organization and Assembly Language}
@@
 \section{Computer Organization}
-\section{RISC-V Assembly Language}
+\subsection{Basic Computational Units}
 \begin{definition}[Arithmetic Logical Unit]
@@
                 \hline
-            MV rX, rY & rX = rY & Move value from rY to rX \\
-            MV rX, #D & rX = D & Load sign extended immediate D into rX \\
-            \hline
-            ADD rX, rY & rX = rX + rY & Add rX and rY, store result in rX \\
-            ADD rX, #D & rX = rX + D & Add immediate D to rX, store result in rX \\
-            \hline 
-            SUB rX, rY & rX = rX - rY & Subtract rY from rX, store result in rX \\
-            SUB rX, #D & rX = rX - D & Subtract immediate D from rX, store result in rX \\
-            \hline
-            MUL rX, rY & rX = rX * rY & Multiply rX and rY, store result in rX \\
-            \hline
-            DIV rX, rY & rX = rX / rY & Divide rX by rY, store quotient in rX \\
+                \verb|MV rX, rY| & rX = rY & Move value from rY to rX \\
+                \verb|MV rX, #D| & rX = D & Load sign extended immediate D into rX \\
+                \hline
+                \verb|ADD rX, rY| & rX = rX + rY & Add rX and rY, store result in rX \\
+                \verb|ADD rX, #D| & rX = rX + D & Add immediate D to rX, store result in rX \\
+                \hline
+                \verb|SUB rX, rY| & rX = rX - rY & Subtract rY from rX, store result in rX \\
+                \verb|SUB rX, #D| & rX = rX - D & Subtract immediate D from rX, store result in rX \\
+                \hline
+                \verb|MUL rX, rY| & rX = rX * rY & Multiply rX and rY, store result in rX \\
+                \hline
+                \verb|DIV rX, rY| & rX = rX / rY & Divide rX by rY, store quotient in rX \\
                 \hline
@@
 \end{definition}
-\begin{definition}[Instruction Encoding]
-    Encodings are 16-bit or 32-bit binary representations of instructions that the CPU can understand and execute.
-Each instruction is represented by a unique binary code that specifies the operation to be performed and the operands involved.
-The encoding of an instruction typically includes the following components:
-    \begin{itemize}
-        \item \textbf{Opcode:} A unique binary code that identifies the operation to be performed (e.g., addition, subtraction, load, store).
-        \item \textbf
+\begin{definition}[Opcodes]
+    An opcode (operation code) is a portion of a machine language instruction that specifies the operation to be performed.
+It is typically represented as a binary or hexadecimal value and is used by the CPU to determine which operation to execute.
+
+    In RISC-V architecture, opcodes are 7 bits long and are located in the least significant bits of the instruction.
+The opcode determines the type of instruction (e.g., arithmetic, logical, load/store, branch) and the specific operation to be performed within that type.
+
+    For example, the opcode for the ADD instruction is 0000000, while the opcode for the SUB instruction is 0100000. The CPU uses these opcodes to identify the operation and execute it accordingly.
+\end{definition}
+
+\begin{definition}[Instrctuion Cycles]
+    An instruction cycle is the process by which a CPU fetches, decodes, and executes an instruction. It consists of several stages, including:
+    \begin{itemize}
+        \item \textbf{Fetch:} The CPU retrieves the instruction from memory.
+        \item \textbf{Decode:} The CPU decodes the instruction to determine the operation to be performed and the operands involved.
+        \item \textbf{Execute:} The CPU performs the operation specified by the instruction.
+        \item \textbf{Write Back:} The CPU writes the result of the operation back to memory or a register.
+    \end{itemize}
+    The instruction cycle is repeated for each instruction in a program until the program is complete.
+\end{definition}
+
+\subsection{ALU Control and Datapath}
+\begin{definition}[Datapath]
+    A datapath is a collection of functional units, registers, and buses that work together to perform data processing operations within a computer system.
+It is responsible for the flow of data between different components of the system, such as the CPU, memory, and input/output devices.
+
+    The datapath typically includes components such as:
+    \begin{itemize}
+        \item \textbf{Registers:} Temporary storage locations for data.
+        \item \textbf{ALU (Arithmetic Logic Unit):} Performs arithmetic and logical operations on data.
+        \item \textbf{Multiplexers:} Selects between multiple data sources.
+        \item \textbf{Buses:} Pathways for data transfer between components.
+    \end{itemize}
+    The datapath works in conjunction with the control unit, which generates control signals to coordinate the operation of the datapath components based on the instructions being executed.
+\end{definition}
+\begin{definition}[Control Path]
+    A control path is a collection of control signals and logic that manage the operation of the datapath in a computer system.
+It is responsible for generating the necessary control signals to coordinate the activities of the various components within the datapath, such as registers, ALU, multiplexers, and buses.
+
+    The control path typically includes components such as:
+    \begin{itemize}
+        \item \textbf{Control Unit:} Generates control signals based on the current instruction being executed.
+        \item \textbf{Instruction Decoder:} Decodes the instruction to determine the required operations and generates corresponding control signals.
+        \item \textbf{Timing Signals:} Synchronizes the operation of the datapath components with the system clock.
+    \end{itemize}
+    The control path works in conjunction with the datapath to ensure that data is processed correctly and efficiently according to the instructions being executed.
+
+    The control signal would cotnrol the following:
+    \begin{itemize}
+        \item Which operation the ALU would perform.
+        \item Which registers to read from and write to.
+        \item When to read from or write to memory.
+        \item How to route data through the datapath using multiplexers.
+    \end{itemize}
+\end{definition}
+\section{Memory Hierarchy}
+\begin{definition}[Memory]
+    The CPU uses a memory address to specify the location of the data it wants to access.
+For example, a 32-bit system specifies a memory address using 32 bits, which amounts to a total addressable memory of around 4GB ($2^32$ bytes).
+\end{definition}
+
+\begin{definition}[Bit, Byte, Word]
+    A \textbf{bit} is the smallest unit of data in a computer and can have a value of either 0 or 1.
+
+    A \textbf{byte} is a group of 8 bits and is the standard unit of data used to represent a character in most computer systems.
+
+    A \textbf{word} is a group of bytes that is used to represent a larger unit of data, such as an integer or a floating-point number.
+The size of a word can vary depending on the architecture of the computer, but it is typically 4 bytes (32 bits).
+\end{definition}
+
+\begin{definition}[Memory Array]
+    A memory array is a collection of memory cells that are organized in a grid-like structure.
+On a high-level, we think of it as a shape-2 matrix, where each row represents a memory address and each column represents a bit within that address.
+
+    Each cell in the array can store a single bit of data (0 or 1). The memory array is typically organized into words, where each word consists of a fixed number of bits.
+
+    For example, take a 4-byte word, then the memory array would have 4 columns, and $n-2$ rows, where $n$ is the total number of bits in the memory. The address format would be, for example:
+    $$
+    \overbrace{A_n \dots A_2}^{\text{which word}} \underbrace{A_1 A_0}_{\text{which byte in the word}}
+    $$
+\end{definition}
+
+\section{Assembly Language}
+\begin{definition}[Assembly Language]
+    Assembly language is a low-level programming language that is specific to a particular computer architecture.
+It provides a human-readable representation of machine code instructions, allowing programmers to write code that can be directly executed by the CPU.
+
+    Assembly language uses mnemonics to represent machine instructions, making it easier for humans to understand and write code.
+Each assembly instruction corresponds to a specific machine code instruction, and the assembly code is typically translated into machine code using an assembler.
+\end{definition}
+
+\paragraph{List of Assembly Languages} The following are some common assembly languages:
+\begin{itemize}
+    \item x86 Assembly: Used in Intel and AMD processors. It is a CISC (\textbf{C}omplex \textbf{I}nstruction \textbf{S}et \textbf{C}omputer) architecture.
+    \item ARM Assembly: Used in ARM processors, commonly found in mobile devices. It is a RISC (\textbf{R}educed \textbf{I}nstruction \textbf{S}et \textbf{C}omputer) architecture.
+    \item RISC-V Assembly: An open-source instruction set architecture (ISA) that is gaining popularity in academia and industry.
+As in its name, it is a RISC architecture.
+In this course, we will sue the RISC-V 32I flavor.
+\end{itemize}
+\paragraph{RISC vs CISC} RISC architectures have a smaller set of instructions, which are designed to be executed quickly and efficiently.
+CISC architectures have a larger set of instructions, which can perform more complex operations in a single instruction.
+This can lead to more efficient code in some cases, but it can also make the architecture more complex and harder to optimize.
+
+\begin{definition}[Instruction Set Architecture (ISA)]
+    An \textbf{I}nstruction \textbf{S}et \textbf{A}rchitecture (ISA) is a specification that defines the set of instructions that a computer's CPU can execute.
+It serves as the interface between software and hardware, providing a standardized way for programmers to write code that can be executed on a specific architecture.
+
+    The ISA includes details such as:
+    \begin{itemize}
+        \item The format and encoding of instructions
+        \item The types of operations that can be performed (e.g., arithmetic, logical, data movement)
+        \item The number and types of registers available
+        \item The addressing modes used to access memory
+        \item The conventions for function calls and returns
+    \end{itemize}
+    Different ISAs are designed for different types of processors and applications, and they can vary significantly in terms of complexity and capabilities.
+
+    Notably, the instruction set does not define the underlying hardware implementation, which can vary even for processors that implement the same ISA.
+\end{definition}
+
+\subsection{Types of Instructions}
+\begin{definition}[Operand]
+    An operand is a value or data that an instruction operates on.
+In assembly language, operands can be registers, memory addresses, or immediate values (constants).
+The operands specify the source and destination of the data for the operation being performed by the instruction.
+\end{definition}
+\begin{definition}[Arithmetic Instructions]
+    Arithmetic instructions are a type of assembly language instruction that perform mathematical operations on numerical data.
+These instructions typically operate on values stored in registers or memory locations and can include operations such as addition, subtraction, multiplication, and division.
+
+    Lets start with high level code:
+    \begin{verbatim}
+    int a = b + c; // addition
+    int a = b - c; // subtraction
+    \end{verbatim}
+
+    The equivalent assembly code using RISC-V instructions would be:
+    \begin{verbatim}
+    ADD rA, rB, rC  # rA = rB + rC
+    SUB rA, rB, rC  # rA = rB - rC
+    \end{verbatim}  
+    the \verb|rA| occupies the destination operand, while \verb|rB| and \verb|rC| are the source operands.
+
+\end{definition}
+
+\begin{definition}[Register Set]
+    A register set is a collection of registers within a CPU that are used for temporary storage of data during the execution of instructions.
+Registers are small, high-speed storage locations that can hold data such as operands, intermediate results, and memory addresses.
+
+    In RISC-V architecture, there are 32 general-purpose registers (x0 to x31), each 32 bits wide in the RV32I variant.
+The register x0 is hardwired to zero and cannot be modified.
+The other registers can be used for various purposes, such as holding function arguments, return values, and temporary data.
 \end{definition}

### 2025-11-03 YEAR2/MAT292_LectureNotes/main.tex
nov 3

@@
 
+\begin{definition}[Cauchy-Euler Equation and Indicial Equation] \label{def:cauchyEuler}
+    A Cauchy-Euler equation is a type of second-order linear ODE of the form:
+    \begin{equation}
+    t^2 y'' + a t y' + b y = 0
+    \end{equation}
+    where \( a \) and \( b \) are constants.
+To solve this equation, we assume a solution of the form \( y = t^r \), where \( r \) is a constant to be determined.
+Substituting this assumed solution into the Cauchy-Euler equation leads to the indicial equation:
+    \begin{equation}
+    r(r - 1) + a r + b = 0
+    \end{equation}
+    The roots of the indicial equation determine the general solution of the Cauchy-Euler equation, which can take different forms depending on whether the roots are real and distinct, real and repeated, or complex conjugates.
+
+    To generalize, a Cauchy-Euler equation can be expressed as: 
+    \begin{equation}
+    \sum_m y^{(m)}(t) a_m t^m = 0
+    \end{equation}
+    where \( a_m \) are constants.
+
+    And the indicial equation in this generalized form is given by:
+    \begin{equation}
+    \sum_m a_m r (r - 1) (r - 2) \ldots (r - m + 1) = 0
+    \end{equation}
+\end{definition}
+
 \begin{example}[Cauchy-Euler Equation] 
@@
 
-\begin{definition}[Wronskian Derivation of $y_p$]
+\begin{definition}[Wronskian Derivation of $y_p$] \label{def:wronskian_method}
     If $y_1(t)$ and $y_2(t)$ are two linearly independent solutions of the corresponding homogeneous equation (a.k.a.
@@
 
-\begin{definition}[Variation of Parameters]
+\begin{theorem}[Variation of Parameters]
     Variation of parameters is a method used to find a particular solution to a non-homogeneous linear differential equation. For a second-order ODE of the form:
@@
     where \( u_1(t) \) and \( u_2(t) \) are functions to be determined. These functions satisfy the system of equations:
-    \begin{align*}
-        u_1'(t)y_1(t) + u_2'(t)y_2(t) &= 0 \\
-        u_1'(t)y_1'(t) + u_2'(t)y_2'(t) &= g(t)
-    \end{align*}
+    \begin{equation}
+        \begin{cases}
+        u_1'(t)y_1(t) + u_2'(t)y_2(t) = 0 \\
+        u_1'(t)y_1'(t) + u_2'(t)y
+        \end{cases}
+    \end{equation}
     Solving this system allows us to find \( u_1(t) \) and \( u_2(t) \), and thus the particular solution \( y_p(t) \).
 
-\end{definition}
+    Or, we can also plug in the following formula directly:
+    \begin{equation}
+    \begin{cases}
+        u_1(t) = -\int \frac{y_2(t) g(t)}{W(t)} dt \\
+        u_2(t) = \int \frac{y_1(t) g(t)}{W(t)} dt
+    \end{cases}
+    \end{equation}
+    where \( W(t) = y_1(t)y_2'(t) - y_2(t)y_1'(t) \) is the Wronskian determinant of \( y_1 \) and \( y_2 \). However, practically, we only use this when $g$ is a sum of functions that are:
+    \begin{itemize}
+        \item Trigonometric functions
+        \item Exponential functions
+        \item Polynomials
+    \end{itemize}
+\end{theorem}
+\begin{proof}
+    We start with Defintion \ref{def:wronskian_method} and assume a particular solution of the form:
+    $$y_p(t) = -y_1(t) \int \frac{y_2(t) g(t)}{W(t)} dt + y_2(t) \int \frac{y_1(t) g(t)}{W(t)} dt$$
+    where \( W(t) = y_1(t)y_2'(t) - y_2(t)y_1'(t) \) is the Wronskian determinant of \( y_1 \) and \( y_2 \).
+To verify that this is indeed a particular solution, we need to show that it satisfies the original inhomogeneous ODE:
+    $$y'' + p(t)y' + q(t)y = g(t)$$
+    We can compute the first and second derivatives of \( y_p(t) \) using the product rule and the chain rule.
+After substituting these derivatives back into the left-hand side of the ODE, we can simplify the expression.
+The key step involves using the properties of the Wronskian determinant and the fact that \( y_1(t) \) and \( y_2(t) \) are solutions of the corresponding homogeneous equation.
+After simplification, we find that the left-hand side reduces to \( g(t) \), confirming that \( y_p(t) \) is indeed a particular solution of the inhomogeneous ODE.
+\end{proof}
+
+\begin{example}
+    Conisder the following second-order inhomogeneous ODE:
+    $$y'' + y = \tan(t)$$
+    The corresponding homogeneous equation is:
+    $$y'' + y = 0$$
+    The characteristic equation is:
+    $$r^2 + 1 = 0$$
+    Solving this gives the roots \( r_1 = i \) and \( r_2 = -i \). Thus, the fundamental set of solutions to the homogeneous equation is:
+    $$y_1(t) = \cos(t), \quad y_2(t) = \sin(t)$$
+    The Wronskian determinant \( W(t) \) is given by:
+    $$W(t) = y_1(t)y_2'(t) - y_2(t)y_1'(t) = \cos(t)\cos(t) - \sin(t)(-\sin(t)) = \cos^2(t) + \sin^2(t) = 1$$
+    Using the variation of parameters formula, we can find a particular solution \( y_p(t) \):
+    \begin{align*}
+        u_1(t) &= -\int \frac{y_2(t) g(t)}{W(t)} dt = -\int \sin(t) \tan(t) dt = -\int \sin^2(t)/\cos(t) dt \\
+        &= -\int (1 - \cos^2(t))/\cos(t) dt = -\int \sec(t) dt + \int \cos(t) dt \\
+        &= -\ln|\sec(t) + \tan(t)| + \sin(t) + C_1
+    \end{align*}
+    \begin{align*}
+        u_2(t) &= \int \frac{y_1(t) g(t)}{W(t)} dt = \int \cos(t) \tan(t) dt = \int \sin(t) dt = -\cos(t) + C_2
+    \end{align*}
+    Thus, a particular solution is given by:
+    \begin{align*}
+        y_p(t) &= u_1(t)y_1(t) + u_2(t)y_2(t) \\
+        &= \left[ -\ln|\sec(t) + \tan(t)| + \sin(t)\right] \cos(t) + \left[-\cos(t)\right] \sin(t) \\
+        &= -\cos(t) \ln|\sec(t) + \tan(t)|
+    \end{align*}
+    Finally, the general solution of the inhomogeneous ODE is:
+    $$y(t) = C_1 \cos(t) + C_2 \sin(t) - \cos(t) \ln|\sec(t) + \tan(t)|$$
+\end{example}
+
+\paragraph{Finding the fundamental set of solutions} We can find the fundamental set of solutions using:
+\begin{itemize}
+    \item Cauchy-Euler Equation method if applicable (see Definition \ref{def:cauchyEuler})
+    \item Reduction of Order if one solution is known:
+    \begin{equation}
+    y_2(t) = y_1(t) \int \frac{e^{-\int p(t) dt}}{(y_1(t))^2} dt
+    \end{equation}
+    where \( y_1(t) \) is a known solution of the homogeneous equation.
+    \item Variation of Parameters
+\end{itemize}
+
+\section{Numerical Methods for ODEs}
+\begin{definition}[Euler's Method]
+    Euler's method is a numerical technique used to approximate the solution of ordinary differential equations (ODEs) with a given initial value.
+It is a first-order method that uses a simple iterative approach to estimate the value of the solution at discrete points.
+Given an initial value problem of the form:
+    $$y' = f(t, y), \quad y(t_0) = y_0$$
+    Euler's method approximates the solution at successive time steps \( t_n = t_0 + nh \) (where \( h \) is the step size) using the formula:
+    \begin{equation}
+    y_{n+1} = y_n + h f(t_n, y_n)
+    \end{equation}
+    for \( n = 0, 1, 2, \ldots \).
+The method provides a straightforward way to compute approximate solutions, but it may require small step sizes for accurate results, especially for stiff equations or problems with rapid changes.
+
+    The local truncation error of Euler's method is \( O(h^2) \), and the global truncation error is \( O(h) \).
+The formula, given the upperbound \( M \) on the second derivative of the true solution, is:
+    \begin{equation}
+    |y(t_n) - y_n| \leq \frac{hM}{2} (e^{L(t_n - t_0)} - 1)
+    \end{equation}
+    where \( L \) is the Lipschitz constant for \( f \) with respect to \( y \).
+\end{definition}
+
+\begin{definition}[Improved Euler's Method (Heun's Method)]
+    Improved Euler's method, also known as Heun's method, is a numerical technique used to approximate the solution of ordinary differential equations (ODEs) with a given initial value.
+It is a second-order method that improves upon the basic Euler's method by using an average of slopes to estimate the next value.
+Given an initial value problem of the form:
+    $$y' = f(t, y), \quad y(t_0) = y_0$$
+    Improved Euler's method approximates the solution at successive time steps \( t_n = t_0 + nh \) (where \( h \) is the step size) using the following steps:
+    \begin{subequations}
+    \begin{align}
+        k_1 &= f(t_n, y_n) \\
+        k_2 &= f(t_n + h, y_n + h k_1) \\
+        y_{n+1} &= y_n + \frac{h}{2} (k_1 + k_2)
+    \end{align}
+    \end{subequations}
+
+    The local truncation error of Improved Euler's method is \( O(h^3) \), and the global truncation error is \( O(h^2) \).
+The formula, given the upperbound \( M \) on the second derivative of the true solution, is:
+    \begin{equation}
+    |y(t_n) - y_n| \leq \frac{h^2 M}{2} (e^{L(t_n - t_0)} - 1)
+    \end{equation}
+    where \( L \) is the Lipschitz constant for \( f \) with respect to \( y \).
+\end{definition}
+
+\begin{definition}[Runge-Kutta Method]
+    The Runge-Kutta method is a family of iterative methods used to approximate the solutions of ordinary differential equations (ODEs).
+The most commonly used version is the fourth-order Runge-Kutta method (RK4), which provides a good balance between accuracy and computational efficiency.
+Given an initial value problem of the form:
+    $$y' = f(t, y), \quad y(t_0) = y_0$$
+    the RK4 method approximates the solution at successive time steps \( t_n = t_0 + nh \) (where \( h \) is the step size) using the following steps:
+    1. Compute the slopes:
+    \begin{align*}
+        k_1 &= f(t_n, y_n) \\
+        k_2 &= f\left(t_n + \frac{h}{2}, y_n + \frac{h}{2} k_1\right) \\
+        k_3 &= f\left(t_n + \frac{h}{2}, y_n + \frac{h}{2} k_2\right) \\
+        k_4 &= f(t_n + h, y_n + h k_3)
+    \end{align*}
+    2. Update the solution using a weighted average of these slopes:
+    $$y_{n+1} = y_n + \frac{h}{6} (k_1 + 2k_2 + 2k_3 + k_4)$$
+    for \( n = 0, 1, 2, \ldots \). The RK4 method is widely used due to its high accuracy and stability for a variety of ODE problems.
+
+    The local truncation error of the RK4 method is \( O(h^5) \), and the global truncation error is \( O(h^4) \).
+The formula, given the upperbound \( M \) on the fifth derivative of the true solution, is:
+    \begin{equation}
+    |y(t_n) - y_n| \leq \frac{h^4 M}{90} (e^{L(t_n - t_0)} - 1)
+    \end{equation}
+    where \( L = \max_{t,y} \left| \frac{\partial f}{\partial y} \right| \) is the Lipschitz constant for \( f \) with respect to \( y \).
+\end{definition}
+
+\begin{definition}[Order of a Method]
+    The order of a numerical method for solving ordinary differential equations (ODEs) refers to the rate at which the local truncation error decreases as the step size \( h \) approaches zero.
+Specifically, a method is said to be of order \( p \) if the local truncation error \( \tau(h) \) satisfies:
+    \begin{equation}
+    \tau(h) = O(h^{p+1})
+    \end{equation}
+    This means that as the step size \( h \) is halved, the local truncation error decreases by a factor of approximately \( 2^{p+1} \).
+Higher-order methods generally provide more accurate results for a given step size, but they may also require more computational effort per step.
+For example, Euler's method is a first-order method, while the fourth-order Runge-Kutta method (RK4) is a fourth-order method.
+\end{definition}
+
+\begin{example}[Estimating the Order of a Method]
+    Given the following approximations of the solution to an ODE at a specific point using different step sizes:
+    \begin{align*}
+        y(h) &= 1.2 \\
+        y(h/2) &= 1.1 \\
+        y(h/4) &= 1.05
+    \end{align*}
+    We can estimate the order of the numerical method used. Let \( E(h) \) be the error associated with step size \( h \). Assuming the error behaves as:
+    $$
+    E(h) = C h^p
+    $$
+    for some constant \( C \) and order \( p \), we can set up the following equations based on the given approximations:
+    \begin{align*}
+        E(h) &= |y(h) - y_{\text{exact}}| \\
+        E(h/2) &= |y(h/2) - y_{\text{exact}}| \\
+        E(h/4) &= |y(h/4) - y_{\text{exact}}|
+    \end{align*}
+    By taking the ratios of these errors, we can solve for \( p \):
+    \begin{align*}
+        \frac{E(h)}{E(h/2)} &= \frac{C h^p}{C (h/2)^p} = 2^p \\
+        \frac{E(h/2)}{E(h/4)} &= \frac{C (h/2)^p}{C (h/4)^p} = 2^p
+    \end{align*}
+    Taking the logarithm of both sides gives:
+    \begin{align*}
+        p &= \log_2\left(\frac{E(h)}{E(h/2)}\right) \\
+        p &= \log_2\left(\frac{E(h/2)}{E(h/4)}\right)
+    \end{align*}
+    By substituting the actual error values (which can be computed if the exact solution is known), we can estimate the order \( p \) of the numerical method used.
+In this case, we find that \( p \approx 1 \), indicating that the method is first-order accurate.
+\end{example}
 \end{document}

### 2025-11-03 YEAR2/PHY293_LectureNotes/main.tex
nov 3

@@
 
+\subsection{Quantization of Light}
+
+\begin{definition}[De Broglie Wavelength]
+    The de Broglie wavelength is the wavelength associated with a particle and is given by:
+    \begin{equation}
+        \lambda = \frac{h}{p}
+    \end{equation}
+    where \( h \) is Planck's constant and \( p \) is the momentum of the particle.
+This concept suggests that all matter exhibits wave-like properties, with the wavelength inversely proportional to the momentum.
+
+\end{definition}
+
+\begin{definition}[Energy of a Wave]
+    The energy of a photon (quantum of light) is given by:
+    \begin{equation}
+        E = h f = \frac{h c}{\lambda}
+    \end{equation}
+    where \( E \) is the energy of the photon, \( h \) is Planck's constant, and \( f \) is the frequency of the electromagnetic wave.
+
+    And the momentum of a photon is given by:
+    \begin{equation}
+        p = \frac{E}{c} = \frac{h f}{c}
+    \end{equation}
+\end{definition}
 \begin{definition}[Photoelectric Effect]
@@
 \end{definition}
+
+\begin{example}[Photoelectric Experiment]
+    Consider the folling $V-f$ graph of a photoelectric experiment:
+    % work fuction at f = 1, slope around 4
+    \begin{figure}[h!]
+        \centering
+        % tikz graph
+        \begin{tikzpicture}
+            \begin{axis}[
+                axis lines = left,
+                xlabel = {$f$ (frequency)},
+                ylabel = {$V$ (stopping potential)},
+                xmin=0, xmax=5,
+                ymin=-1, ymax=5,
+                xtick={0,1,2,3,4,5},
+                ytick={0,1,2,3,4,5},
+                grid=major,
+                width=10cm,
+                height=8cm,
+            ]
+            \addplot[
+                domain=0:5,
+                samples=100,
+                color=blue,
+            ]
+            {4*x - 4};
+            \addlegendentry{$V = \frac{h}{e} f - \frac{\phi}{e}$}
+            \end{axis}
+        \end{tikzpicture}
+        \caption{Photoelectric Experiment $V-f$ Graph}
+        \label{fig:photoelectric_experiment}
+    \end{figure}
+    From the graph, we can determine the work function \( \phi \) and Planck's constant \( h \):
+
+    First, we find the planks constant from the slope of the line:
+    $$
+        \text{slope} = \frac{h}{e} = 4 \,\text{V·s}
+    $$
+    Thus, we have:
+    $$
+        h = 4e \,\text{V·s} = 4 \times 1.602 \times 10^{-19} \,\text{C·V·s} = 6.408 \times 10^{-19} \,\text{J·s}
+    $$
+    which is approximately equal to the known value of Planck's constant \( h \approx 6.626 \times 10^{-34} \,\text{J·s} \).
+We can find the work function from the x-intercept, and covnert to energy using the experimental value of \( h \):
+    $$
+        f_0 = 1 \,\text{Hz} \implies \phi = h f_0 = 6.408 \times 10^{-19} \,\text{J}
+    $$
+    converting to electron volts:
+    $$
+        \phi = \frac{6.408 \times 10^{-19} \,\text{J}}{1.602 \times 10^{-19} \,\text{J/eV}} \approx 4 \,\text{eV}
+    $$
+
+\end{example}
+\subsection{Quantization of Atomic Energy Level}
+% the 1 electron atom model, rydburg formula, hydrogen spectrum, and the 13.6/n^2 eV
+
+\begin{definition}[Bohr Model of the Hydrogen Atom]
+    The Bohr model describes the hydrogen atom as a system where the electron orbits the nucleus in discrete energy levels. The key postulates of the Bohr model are:
+    \begin{itemize}
+        \item The electron moves in circular orbits around the nucleus under the influence of the Coulomb force.
+        \item Only certain discrete orbits are allowed, corresponding to quantized angular momentum:
+        \begin{equation}
+            L = n \hbar = n \frac{h}{2\pi}, \quad n = 1, 2, 3, \ldots
+        \end{equation}
+        \item The electron does not radiate energy while in a stable orbit.
+        \item Energy is emitted or absorbed when the electron transitions between these quantized orbits, with the energy difference given by:
+        \begin{equation}
+            \Delta E = E_f - E_i = h f
+        \end{equation}
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Energy Levels of the Hydrogen Atom]
+    The energy levels of the hydrogen atom are quantized and given by the formula:
+    \begin{equation}
+        E_n = -\frac{13.6 \,\text{eV}}{n^2}, \quad n = 1, 2, 3, \ldots
+    \end{equation}
+    where \( E_n \) is the energy of the electron in the \( n^\text{th} \) energy level.
+The negative sign indicates that the electron is bound to the nucleus.
+The energy difference between two levels \( n_i \) and \( n_f \) corresponds to the emission or absorption of a photon with frequency \( f \):
+    \begin{equation}
+        h f = E_{n_f} - E_{n_i} = -13.6 \,\text{eV} \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right)
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Energy Spectrum of Hydrogen-Like Atoms]
+    The energy spectrum of hydrogen-like atoms (atoms with a single electron, such as He\(^+\), Li\(^{2+}\), etc.) is given by:
+    \begin{equation}
+        E_n = -\frac{Z^2 \cdot 13.6 \,\text{eV}}{n^2}, \quad n = 1, 2, 3, \ldots
+    \end{equation}
+    where \( Z \) is the atomic number of the nucleus (number of protons).
+The energy difference between two levels \( n_i \) and \( n_f \) corresponds to the emission or absorption of a photon with frequency \( f \):
+    \begin{equation}
+        h f = E_{n_f} - E_{n_i} = -Z^2 \cdot 13.6 \,\text{eV} \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right)
+    \end{equation}
+\end{definition}
+
+\paragraph{Hydrogen-Like Atoms} Atoms with one valence electron, such as He\(^+\), Li\(^{2+}\), Be\(^{3+}\), etc. Or, alkaline metals like Na, K, Rb, Cs, Fr.
+
+\begin{definition}[Characteristic X-Rays and Moseley's Law] When high-energy electrons strike a metal target, they can eject inner-shell electrons, leading to characteristic X-ray emission as outer-shell electrons fill the vacancies.
+Moseley's law relates the frequency of the emitted X-rays to the atomic number \( Z \) of the target element:
+\begin{equation}
+    \sqrt{f} = K (Z - b)
+\end{equation}
+where \( f \) is the frequency of the emitted X-ray, \( K \) is a constant, and \( b \) is a screening constant that accounts for the shielding effect of the inner electrons.
+\end{definition}
+
+\subsection{De Broglie's Wavelength}
+\begin{definition}[De Broglie Hypothesis]
+    The de Broglie hypothesis posits that all matter exhibits wave-like properties, with the wavelength \( \lambda \) of a particle being inversely proportional to its momentum \( p \):
+    \begin{equation}
+        \lambda = \frac{h}{p}
+    \end{equation}
+    where \( h \) is Planck's constant.
+This concept extends wave-particle duality to all particles, suggesting that particles such as electrons, protons, and even larger objects have associated wavelengths.
+\end{definition}
+
+\begin{definition}[Wave Nature of Electrons]
+    The electron behave as wave satisfying the de Broglie relation. For a stable robit, the elevtron wave must fit exactly into the circumference of the orbit:
+    \begin{equation}
+        2 \pi r = n \lambda, \quad n = 1, 2, 3, \ldots
+    \end{equation}
+    where \( r \) is the radius of the orbit and \( \lambda \) is the de Broglie wavelength of the electron. This condition leads to the quantization of angular momentum in Bohr's model of the atom:
+    \begin{equation}
+        L = rp = n \hbar = n \frac{h}{2\pi}, \quad n = 1, 2, 3, \ldots
+    \end{equation}
+
+\end{definition}
+
+\begin{definition}[De Broglie Wavelength of an Electron]
+    \textit{Calculate the de Broglie wavelength of an electron moving with a velocity of \( 2 \times 10^6 \) m/s.
+The mass of the electron is \( 9.11 \times 10^{-31} \) kg and Planck's constant is \( 6.626 \times 10^{-34} \) J·s.
+Ignore relativistic effects.}
+
+    To calculate the de Broglie wavelength \( \lambda \) of the electron, we use the de Broglie relation:
+   $$
+        \lambda = \frac{h}{p}
+    $$
+    where \( p \) is the momentum of the electron, given by:
+    $$
+        p = mv
+    $$
+    where \( m \) is the mass of the electron and \( v \) is its velocity.
+
+    First, we calculate the momentum \( p \):
+    $$        p = (9.11 \times 10^{-31} \, \text{kg})(2 \times 10^6 \, \text{m/s}) = 1.822 \times 10^{-24} \, \text{kg·m/s} $$
+
+    Now, we can calculate the de Broglie wavelength \( \lambda \):
+    $$        \lambda = \frac{6.626 \times 10^{-34} \, \text{J·s}}{1.822 \times 10^{-24} \, \text{kg·m/s}} \approx 3.64 \times 10^{-10} \, \text{m} $$
+    Therefore, the de Broglie wavelength of the electron is approximately \( 3.64 \times 10^{-10} \) meters.
+\end{definition}
+
+\begin{example}[De Broglie Wavelength of a Baseball]
+    \textit{Jeff Hoffman was pitching a baseball with a speed of \( 30 \) m/s, sending Blue Jays to the World Series.
+The mass of the baseball is \( 0.1 \) kg.
+Calculate the de Broglie wavelength of the baseball.}
+
+    Using the same method, we have:
+    $$
+    \lambda = \frac{6.63 \times 10^{-34} \, \text{J·s}}{(0.1 \, \text{kg})(30 \, \text{m/s})} = 2.21 \times 10^{-34} \, \text{m}
+    $$
+    Therefore, the de Broglie wavelength of the baseball is approximately \( 2.21 \times 10^{-34} \) meters, which is extremely small and not observable in everyday life.
+\end{example}
+
+\begin{example}[Momentum is Wavelength]
+    \textit{The mass of the proton is 1836 times greater than the mass of the electron.
+In order for a proton to have the same momentum as an electron, how must the de Broglie wavelength of the proton compare to that of the electron?}
+    According to the de Broglie relation:
+    $$        \lambda = \frac{h}{p}
+    $$
+    For the proton and electron to have the same momentum \( p \), we have:
+    $$        \lambda_p = \frac{h}{p} \quad \text{and} \quad \lambda_e = \frac{h}{p} $$
+    Therefore,:
+    $$        \lambda_p = \lambda_e $$
+    Thus, the de Broglie wavelength of the proton must be equal to that of the electron in order for them to have the same momentum.
+\end{example}
+
+\begin{example}[Davisson-Germer Experiment]
+    In the Davisson-Germer experiment, electrons were accelerated through a potential difference of \( 54 \) V before striking a nickel crystal.
+Calculate the de Broglie wavelength of the electrons after acceleration.
+
+    The interplaner spacing of the nickel crystal is \( d = 0.091 \) nm, and the first-order diffraction maximum was observed at an angle of \( \theta = 50^\circ \).
+
+    First, we calculate the kinetic energy \( K \) of the electrons after being accelerated through the potential difference \( V \):
+    $$        
+    K = eV = (1.602 \times 10^{-19} \, \text{C})(54 \, \text{V}) = 8.65 \times 10^{-18} \, \text{J} 
+    $$
+    Next, we calculate the momentum \( p \) of the electrons using the kinetic energy:
+    $$        K = \frac{p^2}{2m} \implies p = \sqrt{2mK} $$
+    where \( m \) is the mass of the electron (\(9.11 \times 10^{-31} \, \text{kg}\)). Thus,
+    $$        p = \sqrt{2(9.11 \times 10^{-31} \, \text{kg})(8.65 \times 10^{-18} \,\text{J})} \approx 1.26 \times 10^{-24} \text{kg·m/s} $$
+    Now, we can calculate the de Broglie wavelength \( \lambda \) of the electrons:
+    $$        \lambda = \frac{h}{p} = \frac{6.626 \times 10^{-34} \, \text{J·s}}{1.26 \times 10^{-24} \, \text{kg·m/s}} \approx 1.67 \times 10^{-10} \, \text{m} $$
+    Finally, we can verify this wavelength using Bragg's law for diffraction:
+    $$        n\lambda = 2d \sin \theta $$
+    For the first-order maximum (\( n = 1 \)), \( \theta = \frac{180 - 50}{2} = 65^\circ \):
+    $$        \lambda = 2(0.091 \times 10^{-9} \, \text{m}) \sin(65^\circ) \approx 1.67 \times 10^{-10} \, \text{m} $$
+    This matches our earlier calculation, confirming the de Broglie wavelength of the electrons is approximately \( 1.67 \times 10^{-10} \) meters.
+\end{example}
 \end{document}

### 2025-11-05 YEAR2/AER210_LectureNotes/main.tex
nov 5

@@
 \end{theorem}
+
+\begin{definition}[Hydrodatic Force on a Submerged Surface]
+    The hydrostatic force \(F\) acting on a submerged surface \(S\) is given by:
+    \begin{equation}
+        F = \iint_S P \, dA
+    \end{equation}
+    where \(P\) is the pressure at each point on the surface and \(dA\) is the infinitesimal area element of the surface.
+
+    The point of application of the hydrostatic force, known as the center of pressure, can be determined by finding the centroid of the pressure distribution on the surface.
+\end{definition}
+
+\begin{definition}[Pressure Prism]
+    A pressure prism is a graphical representation of the pressure distribution on a submerged surface.
+It is typically used to visualize the variation of pressure with depth in a fluid.
+The height of the prism at any point on the surface corresponds to the pressure at that point, and the area of the prism represents the total hydrostatic force acting on the surface.
+\end{definition}
+
+\begin{example}[Pressure Prsim]
+    Cosndier a rectangular dam with a width \(b\) and a height \(h\).
+The dam is submerged in water, and we want to calculate the hydrostatic force acting on the dam.
+The pressure at a depth \(z\) below the water surface is given by:
+    \[P(z) = \rho g z\]
+    where \(\rho\) is the density of water and \(g\) is the acceleration due to gravity. The average pressure on the dam can be calculated as:
+    \[P_{avg} = \frac{\rho g h}{2}\]
+    The hydrostatic force acting on the dam is then given by:
+    \[F = P_{avg} \cdot A = \frac{\rho g h}{2} \cdot (b \cdot h) = \frac{1}{2} \rho g b h^2\]
+    where \(A = b \cdot h\) is the area of the dam.
+To visulzied this, we can draw a pressure prism in the \textit{prssure-area space} where the base of the prism is the area of the dam and the height of the prism is the average pressure acting on the dam.
+Then, we can condlue that the magnitude of the hydrostatic force is equal to the volume of the pressure prism.
+
+    The centroid of the pressure prism can be found to determine the point of application of the hydrostatic force.
+For a rectangular dam, the centroid is located at a depth of \(\frac{h}{3}\) from the base of the dam.
+
+    \textbf{Important} This method is \textbf{only valid for planar surfaces}.
+\end{example}
+
+\begin{definition}[Determining Centroid]
+ Recall, the centroid of a shape with multiple geometries is calculated by: 
+\begin{equation}
+\bar{y} = \frac{\sum_i V_i  y_i}{\sum_i V_i}
+\end{equation}
+\end{definition}
+
+\paragraph{Note} Effect of atmospheric pressure: In most hydrostatic problems, the effect of atmospheric pressure is neglected because it acts uniformly on all surfaces and does not contribute to the net force on submerged surfaces.
+However, if the surface is exposed to the atmosphere, the atmospheric pressure must be considered in addition to the hydrostatic pressure.
+
 \end{document}

### 2025-11-05 YEAR2/ECE253_LectureNotes/main.tex
nov 5

@@
 \end{definition}
+
+\begin{definition}[\texttt{lw} and \texttt{la} Instructions]
+    The \verb|lw| (load word) instruction is used to load a 32-bit word from memory into a register. The syntax for the \verb|lw| instruction is as follows:
+    \begin{verbatim}
+    lw rX, offset(rY)
+    \end{verbatim}
+    This instruction loads the word located at the memory address calculated by adding the value in register \verb|rY| and the signed immediate \verb|offset| into register \verb|rX|.
+
+    The \verb|la| (load address) instruction is used to load the address of a label or variable into a register. The syntax for the \verb|la| instruction is as follows:
+    \begin{verbatim}
+    la rX, LABEL
+    \end{verbatim}
+    This instruction loads the address of the label or variable \verb|LABEL| into register \verb|rX|.
+\end{definition}
+
+\begin{example}[RISC-V Assembly Code - Rolling Sum]
+    Consider the following C code that uses that performs a rolling sum using pointers of a 4 words array:
+    \begin{verbatim}
+    void rolling_sum(int *arr, int n) {
+        int sum = 0;
+        for (int i = 0; i < n; i++) {
+            sum += arr[i];
+        }
+        return sum;
+    }
+
+    int main() {
+        int LIST[4] = {1, 2, 3, 4};
+        int result = rolling_sum(LIST, 4);
+        return 0;
+    }
+    \end{verbatim}
+    The equivalent RISC-V assembly code would be:
+    \begin{verbatim}
+    .data
+    LIST: .word 1, 2, 3, 4  # Example array
+
+    .global _start # label is visible to other files
+    .text # where the instructions are located
+
+    _start:
+        la s1, LIST   # Load address of LIST into s1
+        lw s2, 0(s1)  # Load first element of LIST into s2
+        lw s3, 4(s1)  # Load second element of LIST into s3
+        add s2 , s2, s3 # sum = arr[0] + arr[1]
+        lw s3, 8(s1)  # Load third element of LIST into s3
+        add s2 , s2, s3 # sum += arr[2]
+        lw s3, 12(s1)  # Load fourth element of LIST into s3
+        add s2 , s2, s3 # sum += arr[3]
+        # sum is now in s2
+    END: ebreak # we use ebreak to end the program, to transfer control to the simulator
+    \end{verbatim}
+\end{example}
+
+\begin{definition}[Logic Instructions]
+    Logic instructions are a type of assembly language instruction that perform bitwise operations on binary data.
+These instructions typically operate on values stored in registers or memory locations and can include operations such as AND, OR, NOT, and XOR.
+
+    For example, consider the following high-level code:
+    \begin{verbatim}
+    int a = b & c; // bitwise AND
+    int a = b | c; // bitwise OR
+    int a = b ^ c; // bitwise XOR
+    int a = ~b;    // bitwise NOT
+    \end{verbatim}
+
+    The equivalent assembly code using RISC-V instructions would be:
+    \begin{verbatim}
+    AND rA, rB, rC  # rA = rB & rC
+    OR  rA, rB, rC  # rA = rB | rC
+    XOR rA, rB, rC  # rA = rB ^ rC
+    NOT rA, rB      # rA = ~rB; Pseudo-instruction,
+    # use XORI rA, rB, -1, since -1 = 0xFFFFFFFF
+    \end{verbatim}
+    where \verb|rA| occupies the destination operand, while \verb|rB| and \verb|rC| are the source operands.
+
+    We can also use the immediate values as operands:
+    \begin{verbatim}
+    ANDI rA, rB, #D  # rA = rB & D
+    ORI  rA, rB, #D  # rA = rB | D
+    XORI rA, rB, #D  # rA = rB ^ D
+    \end{verbatim}
+\end{definition}
+
+\begin{definition}[Shift Instructions]
+    Shift instructions are a type of assembly language instruction that perform bitwise shift operations on binary data.
+These instructions typically operate on values stored in registers and can include operations such as logical shift left (LSL), logical shift right (LSR), and arithmetic shift right (ASR).
+
+    For example, consider the following high-level code:
+    \begin{verbatim}
+    int a = b << n; // logical shift left
+    int a = b >> n; // logical shift right
+    int a = b >> n; // arithmetic shift right
+    \end{verbatim}
+
+    The equivalent assembly code using RISC-V instructions would be:
+    \begin{verbatim}
+    SLL rA, rB, rC  # rA = rB << rC (logical shift left)
+    SRL rA, rB, rC  # rA = rB >> rC (logical shift right)
+    SRA rA, rB, rC  # rA = rB >> rC (arithmetic shift right)
+    \end{verbatim}
+    where \verb|rA| occupies the destination operand, while \verb|rB| is the source operand and \verb|rC| specifies the number of bits to shift.
+
+    The arithmetic shift right (ASR) preserves the sign bit (the most significant bit) when shifting right, which is important for signed integers.
+In contrast, the logical shift right (LSR) fills the vacated bits with zeros, regardless of the sign of the number.
+
+    To multiply or divide by powers of two, we can use shift instructions:
+    \begin{verbatim}
+    SLA rA, rB, #D  # rA = rB * 2^D (multiply by 2^D)
+    SRA rA, rB, #D  # rA = rB / 2^D (divide by 2^D, if it is divisable)
+    \end{verbatim}
+
+    We can also use immediate values as operands:
+    \begin{verbatim}
+    SLLI rA, rB, #D  # rA = rB << D (logical shift left)
+    SRLI rA, rB, #D  # rA = rB >> D (logical shift right)
+    SRAI rA, rB, #D  # rA = rB >> D (arithmetic shift right)
+    \end{verbatim}
+\end{definition}
+
+\begin{definition}[Accessing a byte and half word]
+    In RISC-V, we can access individual bytes and half-words (16 bits) using specific load and store instructions.
+
+    To load a byte from memory into a register, we use the \verb|lb| (load byte) instruction:
+    \begin{verbatim}
+    lb rX, offset(rY)
+    \end{verbatim}
+    This instruction loads the byte located at the memory address calculated by adding the value in register \verb|rY| and the signed immediate \verb|offset| into register \verb|rX|.
+The loaded byte is sign-extended to fill the 32-bit register.
+
+    To load a half-word from memory into a register, we use the \verb|lh| (load half-word) instruction:
+    \begin{verbatim}
+    lh rX, offset(rY)
+    \end{verbatim}
+    This instruction loads the half-word located at the memory address calculated by adding the value in register \verb|rY| and the signed immediate \verb|offset| into register \verb|rX|.
+The loaded half-word is sign-extended to fill the 32-bit register.
+
+    Similarly, to store a byte from a register into memory, we use the \verb|sb| (store byte) instruction:
+    \begin{verbatim}
+    sb rX, offset(rY)
+    \end{verbatim}
+    This instruction stores the least significant byte of register \verb|rX| into the memory address calculated by adding the value in register \verb|rY| and the signed immediate \verb|offset|.
+
+    To store a half-word from a register into memory, we use the \verb|sh| (store half-word) instruction:
+    \begin{verbatim}
+    sh rX, offset(rY)
+    \end{verbatim}
+    This instruction stores the least significant half-word of register \verb|rX| into the memory address calculated by adding the value in register \verb|rY| and the signed immediate \verb|offset|.
+
+\end{definition}
 \end{document}

### 2025-11-05 YEAR2/MAT292_LectureNotes/main.tex
nov 5

@@
 
+\begin{definition}[General Planar System of ODEs]
+    Consider the general planar system of ODEs:
+    $$
+    X' = P(t)X + G(t)
+    $$
+    where \( X = \begin{bmatrix} x \\ y \end{bmatrix} \), \( P(t) = \begin{bmatrix} p_{11}(t) & p_{12}(t) \\ p_{21}(t) & p_{22}(t) \end{bmatrix} \), and \( G(t) = \begin{bmatrix} g_1(t) \\ g_2(t) \end{bmatrix} \).
+The system is called:
+    \begin{itemize}
+        \item \textbf{Homogeneous} if \( G(t) = 0 \) for all \( t \).
+        \item \textbf{Non-homogeneous} if \( G(t) \neq 0 \) for some \( t \).
+        \item \textbf{Autonomous} if \( P(t) \) and \( G(t) \) are constant matrices/vectors.
+        \item \textbf{Non-autonomous} if \( P(t) \) and/or \( G(t) \) are functions of \( t \).
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Method to Solve General Planar System of ODEs]
+    To solve the general planar system of ODEs, we can translate the system into a higher-dimensional homogeneous system by translating the fixed point to the origin.
+Let \( X^* \) be the fixed point of the system, then we define a new variable \( Y = X - X^* \).
+The system becomes:
+    $$Y' = P(t)Y
+$$
+    We can then solve this homogeneous system using methods for linear systems of ODEs.
+\end{definition}
+
+\begin{example}
+    Consider the system of ODEs:
+    $$
+    \textbf{x}' = \begin{bmatrix} 2 & 3 \\ 1 & 4
+    \end{bmatrix}\textbf{x} + \begin{bmatrix} 5 \\ 6
+    \end{bmatrix}
+    $$
+    We find the fixed point by solving:
+    $$
+    A\textbf{x}^* + \textbf{b} = 0 \implies \begin{bmatrix} 2 & 3 \\ 1 & 4
+    \end{bmatrix}\textbf{x}^* = -\begin{bmatrix} 5 \\ 6
+    \end{bmatrix}
+    $$
+    Solving this, we find:
+    $$
+    \textbf{x}^* = \begin{bmatrix} -3 \\ 2
+    \end{bmatrix}
+    $$
+    We define a new variable \( \textbf{y} = \textbf{x} - \textbf{x}^* \), then the system becomes:
+    $$
+    \textbf{y}' = \begin{bmatrix} 2 & 3 \\ 1 & 4
+    \end{bmatrix}\textbf{y}
+    $$
+    We can then solve this homogeneous system using methods for linear systems of ODEs.
+\end{example}
+
 \subsection{Wronskian and Fundamental Matrix}
@@
 
+We want to formalize the two methods of finding particular solutions: undetermined coefficients and variation of parameters.
+
+\begin{theorem}[Method of Undetermined Coefficients]
+    The method of undetermined coefficients is a technique used to find a particular solution to a non-homogeneous linear differential equation with constant coefficients.
+For a second-order ODE of the form:
+    $$y'' + p y' + q y = g(t)$$
+    where \( g(t) \) is a linear combination of functions such as polynomials, exponentials, sines, and cosines, we can assume a particular solution \( y_p(t) \) of the same form as \( g(t) \), but with undetermined coefficients.
+The steps to apply this method are as follows:
+    \begin{enumerate}
+        \item Identify the form of \( g(t) \).
+        \item Assume a particular solution \( y_p(t) \) with undetermined coefficients based on the form of \( g(t) \).
+        \item Substitute \( y_p(t) \) into the left-hand side of the differential equation.
+        \item Solve for the undetermined coefficients by equating the resulting expression to \( g(t) \).
+    \end{enumerate}
+    This method is effective when \( g(t) \) is composed of functions whose derivatives are similar in form to the original function.
+\end{theorem}
+
+\begin{example}
+    Consider the ODE:
+    $$y'' - 3y' + 2y = e^{2t} + \sin(t)$$
+    The corresponding homogeneous equation is:
+    $$y'' - 3y' + 2y = 0$$
+    The characteristic equation is:
+    $$r^2 - 3r + 2 = 0$$
+    Solving this gives the roots \( r_1 = 2 \) and \( r_2 = 1 \). Thus, the general solution of the homogeneous equation is:
+    $$y_h(t) = C_1 e^{2t} + C_2 e^{t}$$
+    where \( C_1 \) and \( C_2 \) are constants determined by initial conditions.
+
+    Next, we find a particular solution for each term on the right-hand side of the inhomogeneous equation.
+For the term \( e^{2t} \), since \( e^{2t} \) is already a solution of the homogeneous equation, we assume a particular solution of the form \( y_{p_1}(t) = At e^{2t} \).
+Substituting this into the left-hand side of the ODE gives:
+    \begin{align*}
+        y_{p_1}'' - 3y_{p_1}' + 2y_{p_1} &= (4At + 4A)e^{2t} - 3(2At + A)e^{2t} + 2At e^{2t} \\
+        &= (4At + 4A - 6At - 3A + 2At)e^{2t} \\
+        &= (0At + A)e^{2t} \\
+        &= Ae^{2t}
+    \end{align*}
+
+    Setting this equal to \( e^{2t} \) gives:
+    $$Ae^{2t} = e^{2t} \implies A = 1$$
+    Thus, we have:
+    $$y_{p_1}(t) = t e^{2t}$$
+
+    For the term \( \sin(t) \), we can assume a particular solution of the form \( y_{p_2}(t) = B \cos(t) + C \sin(t) \). Substituting this into the left-hand side of the ODE gives:
+    \begin{align*}
+        y_{p_2}'' - 3y_{p_2}' + 2y_{p_2} &= -B \cos(t) - C \sin(t) + 3B \sin(t) - 3C \cos(t) + 2B \cos(t) + 2C \sin(t) \\
+        &= (B - 3C) \cos(t) + (3B - C) \sin(t)
+    \end{align*}
+    Setting this equal to \( \sin(t) \) gives the system of equations:
+    \begin{align*}
+        B - 3C &= 0 \\
+        3B - C &= 1
+    \end{align*}
+    Solving this system gives \( B = \frac{3}{10} \) and \( C = \frac{1}{10} \). Thus, we have:
+    $$y_{p_2}(t) = \frac{3}{10} \cos(t) + \frac{1}{10} \sin(t)$$
+    Finally, the general solution of the inhomogeneous ODE is:
+    $$y(t) = C_1 e^{2t} + C_2 e^{t} + t e^{2t} + \frac{3}{10} \cos(t) + \frac{1}{10} \sin(t)$$
+\end{example}
+
 \begin{theorem}[Variation of Parameters]

### 2025-11-05 YEAR2/PHY293_LectureNotes/main.tex
nov 5

@@
 \end{example}
+\subsection{Matter Waves}
+\begin{definition}[Matter Wave Function]
+    In quantum mechanics, the matter wave function \( \psi(x,t) \) describes the quantum state of a particle. The probability density
+
+    \begin{equation}
+        D(x,t) = |\psi(x,t)|^2
+    \end{equation}
+    gives the likelihood of finding the particle at position \( x \) and time \( t \). The wave function must satisfy the normalization condition:
+
+    \begin{equation}
+        \int_{-\infty}^{\infty} |\psi(x,t)|^2 \, dx = 1
+    \end{equation}
+
+    The wave function can be expressed as a superposition of plane waves:
+\end{definition}
+
+\begin{definition}[Fourier Integtal]
+    The foruer transformer of a function $f(x)$ is given by:
+    $$
+        A(k) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{\infty} f(x) e^{-ikx} \, dx
+    $$
+    The inverse fourier transform is given by:
+    $$       
+        f(x) = \int _{-\infty}^{\infty} A(k) e^{ikx} \, dk
+    $$
+\end{definition}
+\begin{definition}[Postion and Momentum Space]
+    In quantum mechanics, the wave function can be represented in both position space and momentum space.
+In terms of fourier transformation, $f(x)$ is the wave function in position space, while $A(k)$ is the wave function in momentum space, where \( k \) is related to the momentum \( p \) by \( p = \hbar k \).
+The probability density in momentum space is given by:
+\end{definition}
+
+\begin{definition}[Localization Trade-off]
+    In fourier transform, there is a trade-off between the localization of a function in position space and its localization in momentum (wave-number) space.
+A function that is highly localized in position space will have a broad distribution in momentum space, and vice versa.
+This is a manifestation of the Heisenberg uncertainty principle.
+That is, if the position is well-defined, we would see a broad distribution of momenta, and if the momentum is well-defined, we would see a broad distribution of positions.
+\end{definition}
+
+\begin{theorem}[Fourier Uncertainty Relation]
+    The Fourier uncertainty relation states that the product of the standard deviations of a function in position space and its Fourier transform in momentum space is bounded by a constant:
+    \begin{equation}
+        \Delta x \Delta k \geq \frac{1}{2}
+    \end{equation}
+    where \( \Delta x \) is the standard deviation in position space and \( \Delta k \) is the standard deviation in wave-number space:
+    $$
+        \Delta x = \sqrt{\langle x^2 \rangle - \langle x \rangle^2}, \quad \Delta k = \sqrt{\langle k^2 \rangle - \langle k \rangle^2}
+    $$  
+    Also, we have, in the time space:
+    \begin{equation}
+        \Delta t \Delta \omega \geq \frac{1}{2}
+    \end{equation}
+    where \( \Delta t \) is the standard deviation in time and \( \Delta \omega \) is the standard deviation in angular frequency space.
+\end{theorem}
+
+\begin{theorem}[Heisenberg Uncertainty Principle]
+    The Heisenberg uncertainty principle states that it is impossible to simultaneously know both the exact position and exact momentum of a particle.
+The product of the uncertainties in position \( \Delta x \) and momentum \( \Delta p \) is bounded by a constant:
+    \begin{equation}
+        \Delta x \Delta p \geq \frac{\hbar}{2}
+    \end{equation}
+    where \( \hbar = \frac{h}{2\pi} \) is the reduced Planck's constant. This principle reflects the fundamental limit on the precision with which certain pairs of physical properties can be known.
+
+\end{theorem}
+\begin{proof}
+    (Direct Proof) Starting from the Fourier uncertainty relation:
+    $$
+        \Delta x \Delta k \geq \frac{1}{2}
+    $$
+    We know that momentum \( p \) is related to wave-number \( k \) by:
+    $$
+        p = \hbar k
+    $$
+    Therefore, the uncertainty in momentum \( \Delta p \) can be expressed in terms of the uncertainty in wave-number \( \Delta k \):
+    $$
+        \Delta p = \hbar \Delta k
+    $$
+    Substituting this into the Fourier uncertainty relation gives:
+    $$
+        \Delta x \left( \frac{\Delta p}{\hbar} \right) \geq \frac{1}{2}
+    $$
+    Multiplying both sides by \( \hbar \) yields:
+    $$
+        \Delta x \Delta p \geq \frac{\hbar}{2}
+    $$
+    This completes the proof of the Heisenberg uncertainty principle.
+\end{proof}
+
+\begin{definition}[Time-Energy Uncertainty Principle]
+    The time-energy uncertainty principle states that there is a fundamental limit to the precision with which energy and time can be simultaneously known.
+The product of the uncertainties in wave angular frequency \( \Delta \omega \) and time \( \Delta t \) is bounded by a constant:
+    \begin{equation}
+        \Delta \omega \Delta t \geq \frac{\hbar}{2}
+    \end{equation}
+    This principle implies that if a system exists in a state for a very short time, there is a corresponding uncertainty in its energy.
+\end{definition}
+\begin{proof}
+    (Sketch) Starting from the Fourier uncertainty relation in the time domain, we use the same reasoning as in the position-momentum case. 
+\end{proof}
+\begin{definition}[Hilbert Space]
+    A Hilbert space is a complete inner product space that provides the mathematical framework for quantum mechanics. it is the space cotnaining all physically allowed wave function that are:
+    \begin{itemize}
+        \item Complex, continuous, and normalizable.
+        \item Can be added together and multiplied by complex scalars (i.e., it is a vector space).
+        \item Equipped with an inner product that allows for the calculation of probabilities and expectation values.
+    \end{itemize}
+\end{definition}
+\begin{theorem}[From fourier to Hilbert Space]
+    Every wave function $\psi(x,t)$ can be represented as a vector in a Hilbert space.
+Where each of the basis vectors corresponds to a specific state of the system, such as a position eigenstate or momentum eigenstate.
+The inner product in this space allows us to calculate probabilities and expectation values of observables.
+Of coruse, it must satisfy the normalization condition:
+    $$        \int_{-\infty}^{\infty} |\psi(x,t)|^2 \, dx = 1
+    $$
+    This idea will be explored further in the next chapters.
+\end{theorem}
+
 \end{document}

### 2025-11-20 YEAR2/AER210_LectureNotes/main.tex
LTNS

@@
 
-\subsection{Basic Eqaution for Preessure Fiekd}
+\subsection{Basic Eqaution for Preessure Field}
 \begin{theorem}[Hydrostatic Pressure Equation]
@@
 However, if the surface is exposed to the atmosphere, the atmospheric pressure must be considered in addition to the hydrostatic pressure.
-
+\subsection{Buoyancy}
+\begin{definition}[Buoyant Force]
+    The buoyant force \(F_b\) acting on a submerged or partially submerged object is given by Archimedes' principle:
+    \begin{equation}
+        F_b = \rho_f V_d g
+    \end{equation}
+    where \(\rho_f\) is the density of the fluid, \(V_d\) is the volume of fluid displaced by the object, and \(g\) is the acceleration due to gravity.
+\end{definition}    
+\begin{definition}[Stability of Submerged Bodies]
+    The stability of a submerged body is determined by the relative positions of its center of gravity (CG) and center of buoyancy (CB).
+A submerged body is considered stable if, when tilted, it returns to its original position.
+This occurs when the CB creates a righting moment that opposes the tilt.
+If the CG is below the CB, the body is stable; if the CG is above the CB, the body is unstable and will tip over.
+\end{definition}
+\begin{definition}[Stability of Floating Bodies]
+    The stability of a floating body is determined by the relative positions of its center of gravity (CG) and center of buoyancy (CB).
+A floating body is considered stable if, when tilted, it returns to its original position.
+This occurs when the metacenter (M), the point where the line of action of the buoyant force intersects the vertical line through the CG when the body is tilted, is above the CG.
+If M is below CG, the body is unstable and will tip over.
+\end{definition}
+\section{Fluids in Rigid Body Motion}
+\begin{example}[Fluid in Linear Rigid Body Motion]
+    Consider a container partially filled with liquid. A particle is moving on a straight path with constant acceleration. Determine the shape of the liquid surface inside the container.
+
+    Using the following equation for pressure in a fluid in rigid body motion:
+    \[
+    -\nabla P + \rho (\vec{g} - \vec{a}) = 0
+    \]
+    where \(\vec{a}\) is the acceleration of the particle, we can analyze the situation. The acceleration vector \(\vec{a}\) can be expressed as:
+    \[\vec{a} = a_y \hat{j} + a_z \hat{k}\]
+    where \(a_y\) and \(a_z\) are the components of acceleration in the \(y\) and \(z\) directions, respectively. Then, we can apply to the vector equation:
+    \[-\nabla P + \rho ( - a_y \hat{j} + (g - a_z) \hat{k}) = 0\]
+    This gives us the following system of equations:
+    \[\frac{\partial P}{\partial x} = 0\]
+    \[\frac{\partial P}{\partial y} = -\rho a_y\]
+    \[\frac{\partial P}{\partial z} = -\rho (g + a_z)\]
+    Integrating these equations, we find:
+    \[P = -\rho a_y y - \rho (g + a_z) z + C\]
+    where \(C\) is a constant of integration.
+
+    To determine the shape of the liquid surface, we set the pressure \(P\) equal to the atmospheric pressure \(P_{atm}\) at the free surface:
+    \[P_{atm} = -\rho a_y y - \rho (g + a_z) z + C\]
+    Rearranging this equation, we find the equation of the liquid surface:
+    \[z = \frac{C - P_{atm}}{\rho (g + a_z)} - \frac{a_y}{g + a_z} y = C_1 - \frac{a_y}{g + a_z} y\]
+    where \(C_1 = \frac{C - P_{atm}}{\rho (g + a_z)}\) is a new constant.
+
+    This equation describes a plane inclined at an angle determined by the ratio of the acceleration components.
+Thus, the liquid surface inside the container will be a plane inclined in the direction of the acceleration.
+
+\end{example}
 \end{document}

### 2025-11-20 YEAR2/MAT292_LectureNotes/main.tex
LTNS

@@
 \end{example}
+
+\section{Laplace Transform}
+\begin{center}
+    ``Laplace Transform kills derivatives''
+\end{center}
+\begin{definition}[Integral Transform]
+    An integral transform \( F :  C^0(\mathbb{C}) \to C^0(\mathbb{C}) \) is a functional such that it transforms integrable functions \( f(t) \) into other functions \( F[f(t)](s) \) via an integral of the form:
+    \begin{equation}
+    F[f(t)](s) = \int_\alpha^{\beta} K(t,s) f(t) dt
+    \end{equation}
+    where \( K(t,s) \) is called the kernel of the transform.
+\end{definition}
+\begin{definition}[Laplace Transform]
+    The Laplace transform is an integral transform used to convert a function of time \( f(t) \) into a function of a complex variable \( s \).
+It uses the kernel \( K(t,s) = e^{-st} \) and is defined as:
+    \begin{equation}
+    \mathcal{L}\{f(t)\} = F(s) = \int_0^{\infty} e^{-st} f(t) dt
+    \end{equation}
+    where \( s \in \mathbb{C} \). The Laplace transform is widely used in engineering, physics, and applied mathematics to solve differential equations, analyze systems, and perform control theory.
+
+    The Laplace transform turns a differential operator into an algebraic operator:
+    \begin{equation}
+    \mathcal{L}\{f^{(n)}(t)\} = s^n F(s) - s^{n-1} f(0) - s^{n-2} f'(0) - \ldots - f^{(n-1)}(0)
+    \end{equation}
+
+    By transforming it from the $t$-domain to the $s$-domain, we can solve ODEs and then inverse transform it back to the $t$-domain.
+\end{definition}
+\subsection{Analysis of Laplace Transform}
+\begin{definition}[Poles on the Complex Plane]
+    In the context of the Laplace transform, poles are specific values of the complex variable \( s \) at which the Laplace transform \( F(s) \) becomes unbounded or undefined.
+These poles correspond to the singularities of the transformed function and play a crucial role in determining the behavior of the original function \( f(t) \) in the time domain.
+
+    The location of poles in the complex plane provides insights into the stability and dynamics of systems described by differential equations.
+For instance, poles with positive real parts indicate instability, while those with negative real parts suggest stability.
+The imaginary parts of the poles are associated with oscillatory behavior in the time domain.
+\end{definition}
+\begin{example}[Lapalce Transform of the Unit]
+    The Laplace transform of 1 is given by:
+    \begin{equation*}
+    \mathcal{L}\{1\} = \int_0^{\infty} e^{-st} \cdot 1 \, dt = \int_0^{\infty} e^{-st} dt
+    \end{equation*}
+    Evaluating this integral, we have:
+    \begin{equation*}
+    \mathcal{L}\{1\} = \left[ -\frac{1}{s} e^{-st} \right]_0^{\infty} = \frac{1}{s}
+    \end{equation*}
+    where \( s = 0 \) is a pole in the complex plane.
+\end{example}
+\begin{example}[Laplace Transform of $\sin{(\alpha t)}$]
+    To find the Laplace transform of the function \( f(t) = \sin(\alpha t) \), we start with the definition of the Laplace transform:
+    \begin{equation*}
+    \mathcal{L}\{\sin(\alpha t)\} = \int_0^{\infty} e^{-st} \sin(\alpha t) dt
+    \end{equation*}
+    We can evaluate this integral using integration by parts or by recognizing it as a standard Laplace transform. The result is:
+    \begin{equation*}
+    \mathcal{L}\{\sin(\alpha t)\} = \frac{\alpha}{s^2 + \alpha^2}
+    \end{equation*}
+    where \( s > 0 \) to ensure convergence of the integral.
+
+    The poles of this transform are located at \( s = \pm i\alpha \) in the complex plane.
+\end{example}
+
+\begin{example}
+    Consider the following function:
+    $$f(t) = e^{at} \sin(bt)$$
+    To find the Laplace transform of \( f(t) \), we start with the definition of the Laplace transform:
+    \begin{equation*}
+    \mathcal{L}\{e^{at} \sin(bt)\} = \int_0^{\infty} e^{-st} e^{at} \sin(bt) dt
+    \end{equation*}
+    Simplifying the integrand gives:
+    \begin{equation*}
+    \mathcal{L}\{e^{at} \sin(bt)\} = \int_0^{\infty} e^{-(s - a)t} \sin(bt) dt
+    \end{equation*}
+    We can evaluate this integral using the standard Laplace transform result for \( \sin(bt) \):
+    \begin{equation*}
+    \mathcal{L}\{e^{at} \sin(bt)\} = \frac{b}{(s - a)^2 + b^2}
+    \end{equation*}
+    where \( s > a \) to ensure convergence of the integral.
+
+    The poles of this transform are located at \( s = a \pm ib \) in the complex plane.
+\end{example}
+
+\begin{definition}[Exponential Order] \label{def:exp_order}
+    A function \( f: \mathbb{R} \to \mathbb{R} \) is said to be of exponential order \( \alpha \) if there exist positive constants \( M \) and \( T \) such that:
+    \begin{equation}
+    |f(t)| \leq M e^{\alpha t} \quad \text{for all } t > T
+    \end{equation}
+    This condition ensures that the function does not grow faster than an exponential function as \( t \) approaches infinity.
+Functions of exponential order are important in the context of the Laplace transform, as they guarantee the existence of the transform for sufficiently large values of the complex variable \( s \).
+\end{definition}
+
+\begin{theorem}[Existence of the Laplace Transform]
+    The Laplace transform \( \mathcal{L}\{f(t)\} \) of a function \( f(t) \) exists for all complex numbers \( s \) with $\text{Re}(s) > \alpha$, for some $\alpha$ if the following conditions are satisfied:
+    \begin{enumerate}
+        \item The function \( f(t) \) is piecewise continuous on the interval \( [0, A] \) for any finite \( A > 0 \); and
+        \item The function \( f(t) \) is of exponential order \( \alpha \); that is, there exist positive constants \( M \) and \( T \) such that Definition \ref{def:exp_order} holds.
+    \end{enumerate}
+    In other words, the function has to be \textbf{initially} well-behaved and not grow too fast.
+\end{theorem}
+\begin{example}
+    Consider the following t-domain function:
+    $$f(t) = e^\alpha t$$
+    We can check its Laplace transform:
+    \begin{equation*}
+    \mathcal{L}\{e^{\alpha t}\} = \int_0^{\infty} e^{-st} e^{\alpha t} dt = \int_0^{\infty} e^{-(s - \alpha)t} dt
+    \end{equation*}
+    Evaluating this integral, we have:
+    \begin{equation*}
+    \mathcal{L}\{e^{\alpha t}\} = \left[ -\frac{1}{s - \alpha} e^{-(s - \alpha)t} \right]_0^{\infty} = \frac{1}{s - \alpha}
+    \end{equation*}
+    where \( \text{Re}(s) > \alpha \) to ensure convergence of the integral.
+Thus, the Laplace transform of \( f(t) = e^{\alpha t} \) exists for all complex numbers \( s \) with \( \text{Re}(s) > \alpha \).
+\end{example}
+
+\paragraph{Properties of the Laplace Transform} Following are some important properties of the Laplace transform:
+\begin{itemize}
+    \item Linearity: \( \mathcal{L}\{af(t) + bg(t)\} = a\mathcal{L}\{f(t)\} + b\mathcal{L}\{g(t)\} \)
+    \item First Derivative: \( \mathcal{L}\{f'(t)\} = sF(s) - f(0) \)
+    \item Second Derivative: \( \mathcal{L}\{f''(t)\} = s^2F(s) - sf(0) - f'(0) \)
+    \item Time Shifting: \( \mathcal{L}\{f(t - a)u(t - a)\} = e^{-as}F(s) \)
+    \item Frequency Shifting: \( \mathcal{L}\{e^{at}f(t)\} = F(s - a) \)
+    \item Convolution: \( \mathcal{L}\{f(t) * g(t)\} = F(s)G(s) \)  
+\end{itemize}
+We delve deeper into the linearity property:
+\begin{theorem}[Linearity of the Laplace Transform]
+    The Laplace transform is a linear operator.
+This means that for any two functions \( f(t) \) and \( g(t) \) that are piecewise continuous on \( [0, \infty) \) and of exponential order, and for any constants \( a \) and \( b \), the following property holds:
+    \begin{equation}
+    \mathcal{L}\{af(t) + bg(t)\} = a\mathcal{L}\{f(t)\} + b\mathcal{L}\{g(t)\}
+    \end{equation}
+    where \( \mathcal{L}\{f(t)\} = F(s) \) and \( \mathcal{L}\{g(t)\} = G(s) \).
+\end{theorem}
+\begin{proof}
+    Suppose that the Laplace transforms of \( f(t) \) and \( g(t) \) exist for \( \text{Re}(s_1), \text{Re}(s_2) > \alpha_1, \alpha_2 \) respectively. Then:
+    \begin{align*}
+        \mathcal{L}\{af(t) + bg(t)\} &= \int_0^{\infty} e^{-st} (af(t) + bg(t)) dt \\
+        &= a \int_0^{\infty} e^{-st} f(t) dt + b \int_0^{\infty} e^{-st} g(t) dt \\
+        &= a\mathcal{L}\{f(t)\} + b\mathcal{L}\{g(t)\} \\
+        &= aF(s) + bG(s)
+    \end{align*}
+    Thus, the Laplace transform is linear, and there exist \( \text{Re}(s) > \max(\alpha_1, \alpha_2) \) such that the property holds.
+
+    Note, the RHS of the equation only holds when both transforms exist.
+\end{proof}
+
+\begin{example}
+    Consider the function:
+    $$f_1(t) = e^{a_1 t}, \quad f_2(t) = e^{a_2 t}$$
+    Then, for:
+    $$
+    f_{(\alpha, \beta)} = \alpha f_1(t) + \beta f_2(t) = \alpha e^{a_1 t} + \beta e^{a_2 t}
+    $$
+    The Laplace transform exists for \( \text{Re}(s) > \max(a_1, a_2) \).
+\end{example}
+
+\subsection{Solution to ODEs using Laplace Transform}
+\subsubsection{Transforming ODEs to a s-domain Algebraic Equation}
+\begin{definition}[Laplace Transform of the Derivative Operator]
+    The derivative operator \( D \) defined by \( D[f(t)] = f'(t) \) can be transformed using the Laplace transform. The Laplace transform of the derivative operator is given by:
+    \begin{equation*}
+    \mathcal{L}\{D[f(t)]\} = \int_0^{\infty} e^{-st} D[f(t)] \, dt
+    \end{equation*}
+    By integrating by parts, we can show that:
+    \begin{equation}
+    \mathcal{L}\{D[f(t)]\} = sF(s) - f(0)
+    \end{equation} 
+    where \( F(s) = \mathcal{L}\{f(t)\} \).
+This shows that the Laplace transform converts the differentiation operation in the time domain into an algebraic operation in the frequency domain.
+We can extend this to $n$-th derivatives as follows:
+    \begin{equation}
+    \mathcal{L}\{D^n[f(t)]\} = s^n F(s) - s^{n-1} f(0) - s^{n-2} f'(0) - \ldots - f^{(n-1)}(0)
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Symetrical Differentiation Property]
+    We scaling by $t$ in the time domain corresponds to differentiating with respect to $s$ in the frequency domain. Specifically, the Laplace transform of \( t f(t) \) is given by:
+    \begin{equation}
+    \mathcal{L}\{t f(t)\} = -\frac{dF(s)}{ds}
+    \end{equation}
+    where \( F(s) = \mathcal{L}\{f(t)\} \).
+
+\end{definition}
+\begin{example}
+    Consider the Laplace transform:
+    $$
+    \mathcal{L}\{t f(t)\}
+    $$
+    We can demonstrate this property by starting with the definition of the Laplace transform:
+    \begin{equation*}
+    \mathcal{L}\{t f(t)\} = \int_0^{\infty} e^{-st} t f(t) dt
+    \end{equation*}
+    We can differentiate \( F(s) = \mathcal{L}\{f(t)\} \) with respect to \( s \):
+    \begin{equation*}
+    \frac{dF(s)}{ds} = \frac{d}{ds} \int_0^{\infty} e^{-st} f(t) dt = \int_0^{\infty} \frac{d}{ds} (e^{-st}) f(t) dt = \int_0^{\infty} -t e^{-st} f(t) dt
+    \end{equation*}
+    Thus, we have:
+    \begin{equation*}
+    \mathcal{L}\{t f(t)\} = -\frac{dF(s)}{ds}
+    \end{equation*}    
+\end{example}
+
+\begin{definition}[Laplace Transform of the Integral Operator]
+    The integral operator \( I \) defined by \( I[f(t)] = \int_0^t f(\tau) d\tau \) can be transformed using the Laplace transform. The Laplace transform of the integral operator is given by:
+    \begin{equation*}
+    \mathcal{L}\{I[f(t)]\} = \int_0^{\infty} e^{-st} I[f(t)] \, dt
+    \end{equation*}
+    By changing the order of integration, we can show that:
+    \begin{equation}
+    \mathcal{L}\{I[f(t)]\} = \frac{F(s)}{s}
+    \end{equation} 
+    where \( F(s) = \mathcal{L}\{f(t)\} \). This shows that the Laplace transform converts the integration operation in the time domain into an algebraic operation in the frequency domain.
+\end{definition}
+
+\begin{example}[Transforming and ODEs to Algebraic Equation]
+    Consider the following ODE:
+    $$y'' + by' + cy = g(t)$$
+    with initial conditions \( y(0) = y_0 \) and \( y'(0) = y_1 \). Taking the Laplace transform of both sides gives:
+    $$\mathcal{L}\{y''\} + b\mathcal{L}\{y'\} + c\mathcal{L}\{y\} = \mathcal{L}\{g(t)\}$$
+    Using the properties of the Laplace transform, we have:
+    $$(s^2Y(s) - sy_0 - y_1) + b(sY(s) - y_0) + cY(s) = G(s)$$
+    where \( Y(s) = \mathcal{L}\{y(t)\} \) and \( G(s) = \mathcal{L}\{g(t)\} \). Rearranging terms gives:
+    $$Y(s)(s^2 + bs + c) = G(s) + sy_0 + y_1 + by_0$$
+    Thus, we can solve for \( Y(s) \):
+    $$Y(s) = \frac{G(s) + sy_0 + y_1 + by_0}{s^2 + bs + c}$$
+    Finally, we can find \( y(t) \) by taking the inverse Laplace transform of \( Y(s) \), which would be disccussed later:
+    $$y(t) = \mathcal{L}^{-1}\{Y(s)\}$$
+\end{example}
+\paragraph{Common Laplace Transforms} See Appendix  \ref{apx:common_laplace_transforms} for a list of common Laplace transforms.
+\begin{example}
+    Condider the following IVP:
+    $$y'' + ky = 0, \quad y(0) = 0, \quad y'(0) = \sqrt{k}$$
+    Taking the Laplace transform of both sides gives:
+    $$\mathcal{L}\{y''\} + k\mathcal{L}\{y\} = 0$$
+    Using the properties of the Laplace transform, we have:
+    $$(s^2Y(s) - sy(0) - y'(0)) + kY(s) = 0$$
+    Substituting the initial conditions:
+    $$(s^2Y(s) - \sqrt{k}) + kY(s) = 0$$
+    Rearranging terms gives:
+    $$Y(s)(s^2 + k) = \sqrt{k}$$
+    Thus, we can solve for \( Y(s) \):
+    $$Y(s) = \frac{\sqrt{k}}{s^2 + k} = \frac{\sqrt{k}}{s^2 + (\sqrt{k})^2}$$
+    Finally, we can find \( y(t) \) by taking the inverse Laplace transform of \( Y(s) \). Using the common Laplace transform table, we find:
+    $$y(t) = \mathcal{L}^{-1}\{Y(s)\} = \sin(\sqrt{k}t)$$
+\end{example}
+
+\begin{example}[Inverse Laplace Transform using Common Transforms and Partial Fraction Decomposition]
+    Consider the following equation in the s-domain:
+    $$0 = s^2 Y(s) - 2s- 1 + kY(s)$$
+    Rearranging terms gives:
+    $$Y(s)(s^2 + k) = 2s + 1$$
+    Thus, we can solve for \( Y(s) \):
+    $$Y(s) = \frac{2s + 1}{s^2 + k}$$
+    To find \( y(t) \), we need to take the inverse Laplace transform of \( Y(s) \). We can split \( Y(s) \) into two separate fractions:   
+    $$Y(s) = 2\frac{s}{s^2 + k} + \frac{1}{s^2 + k}$$
+    Using the common Laplace transform table, we find:
+    $$y(t) = \mathcal{L}^{-1}\{Y(s)\} = 2\cos(\sqrt{k}t) + \frac{1}{\sqrt{k}}\sin(\sqrt{k}t)$$
+
+\end{example}
+
+\begin{definition}[Inverse Laplace Transform]
+    The inverse Laplace transform is a mathematical operation that allows us to recover the original time-domain function \( f(t) \) from its Laplace transform \( F(s) \).
+It is denoted by \( \mathcal{L}^{-1}\{F(s)\} \) and is defined as:
+    \begin{equation}
+    f(t) = \mathcal{L}^{-1}[F(s)] = \frac{1}{2\pi i} \lim_{T \to \infty} \int_{\gamma - iT}^{\gamma + iT} e^{st} F(s) ds
+    \end{equation}
+    where \( \gamma \) is a real number such that all singularities of \( F(s) \) are to the left of the line \( \text{Re}(s) = \gamma \) in the complex plane.
+The inverse Laplace transform is widely used in engineering, physics, and applied mathematics to solve differential equations and analyze systems by transforming back from the frequency domain to the time domain.
+
+    In this course, we will primarily use common Laplace transform pairs and properties (from table in Appendix \ref{apx:common_laplace_transforms}) to find the inverse Laplace transform rather than evaluating the integral directly.
+\end{definition}
+
+\paragraph{Partial Fraction Decomposition} When finding the inverse Laplace transform, we may need to use partial fraction decomposition to break down complex rational functions into simpler fractions that can be easily inverted using common Laplace transform pairs.
+In our course, almost every function can be decomposed using partial fractions.
+\begin{theorem}[Linear Fraction]
+    Given the following function in the s-domain:
+    $$F(s) = \frac{a}{s - s_0}$$
+    We can find its inverse Laplace transform using the common Laplace transform table:
+    \begin{equation}
+    f(t) = \mathcal{L}^{-1}\{F(s)\} = a e^{s_0 t}
+    \end{equation}
+\end{theorem}
+
+\begin{theorem}[Repeated Linear Fraction]
+    Given the following function in the s-domain:
+    $$F(s) = \frac{a}{(s - s_0)^n}$$
+    We can find its inverse Laplace transform using the common Laplace transform table:
+    \begin{equation}
+    f(t) = \mathcal{L}^{-1}\{F(s)\} = \frac{a t^{n-1} e^{s_0 t}}{(n-1)!}
+    \end{equation}
+
+    This comes from the fact that multiplying by $e^{s_0 t}$ in the time domain corresponds to shifting by $s_0$ in the s-domain.
+So we can use the $\mathcal{L}[t^n] = \frac{n!}{s^{n+1}}$ property and shift it.
+\end{theorem}
+
+\begin{example}
+    Consider the following IVP:
+    $$y'' - 2y' + y = 0, \quad y(0) = 1, \quad y'(0) = 0$$
+    Taking the Laplace transform of both sides gives:
+    $$\mathcal{L}\{y''\} - 2\mathcal{L}\{y'\} + \mathcal{L}\{y\} = 0$$
+    Using the properties of the Laplace transform, we have:
+    $$(s^2Y(s) - sy(0) - y'(0)) - 2(sY(s) - y(0)) + Y(s) = 0$$
+    Substituting the initial conditions:
+    $$(s^2Y(s) - s) - 2(sY(s) - 1) + Y(s) = 0$$
+    Rearranging terms gives:
+    $$Y(s)(s^2 - 2s + 1) = s - 2$$
+    Thus, we can solve for \( Y(s) \):
+    $$Y(s) = \frac{s - 2}{(s - 1)^2}$$
+    To find \( y(t) \), we need to take the inverse Laplace transform of \( Y(s) \). We can split \( Y(s) \) into two separate fractions using partial fraction decomposition:   
+    $$Y(s) = \frac{s - 1}{(s - 1)^2} - \frac{1}{(s - 1)^2} = \frac{1}{s - 1} - \frac{1}{(s - 1)^2}$$
+    Using the common Laplace transform table and the repeated linear fraction theorem, we find:
+    $$y(t) = \mathcal{L}^{-1}\{Y(s)\} = e^{t} - t e^{t} = (1 - t)e^{t}$$
+\end{example}
+
+\begin{definition}[Transfer Function]
+    In control theory and signal processing, a transfer function is a mathematical representation that describes the relationship between the input and output of a linear time-invariant (LTI) system in the frequency domain.
+It is typically denoted by \( H(s) \) or \( G(s) \) and is defined as the ratio of the Laplace transform of the output \( Y(s) \) to the Laplace transform of the input \( X(s) \):
+    \begin{equation}
+    H(s) = \frac{Y(s)}{X(s)}
+    \end{equation}
+    where \( s \) is a complex variable.
+The transfer function encapsulates the system's dynamics, including its poles and zeros, which determine its stability and frequency response characteristics.
+Transfer functions are widely used in engineering disciplines to analyze and design control systems, filters, and other dynamic systems.
+\end{definition}
+
+\newpage
+\appendix 
+\section{Common Laplace Transforms} \label{apx:}
+\input{common_laplace_transforms.tex}
 \end{document}

### 2025-11-20 YEAR2/PHY293_LectureNotes/main.tex
LTNS

@@
     \end{equation}
+    , where $|\psi(x,t)|^2$ is:
+    $$
+        |\psi(x,t)|^2 = \psi^*(x,t) \psi(x,t)
+    $$
+    Here, \( \psi^*(x,t) \) is the complex conjugate of
 
@@
 
+\section{The Schrödinger Equation in One-Dimensional Space}
+\paragraph{Standing Waves and Stationary States}
+\begin{itemize}
+    \item A particle confined in space behaves like a standing wave.
+    \item For a stationary state:
+    \[
+        \Psi(x,t) = \psi(x)e^{-iEt/\hbar},
+    \]
+    giving a time-independent probability density $|\Psi|^2 = |\psi|^2$.
+    \item Superpositions of stationary states produce time-dependent densities.
+\end{itemize}
+\begin{definition}[Time Dependent Schrödinger Equation]
+    The time-dependent Schrödinger equation describes how the quantum state of a physical system changes over time. It is given by:
+    \begin{equation}
+        i \hbar \frac{\partial \Psi(x,t)}{\partial t} = -\frac{\hbar^2}{2m} \frac{\partial^2 \Psi(x,t)}{\partial x^2} + V(x) \Psi(x,t)
+    \end{equation}
+    where \( \Psi(x,t) \) is the wave function of the particle, \( m \) is the mass of the particle, \( V(x) \) is the potential energy as a function of position, and \( \hbar \) is the reduced Planck's constant.
+
+\end{definition}
+\begin{example}[Particle in a Box]
+    Consider a particle confined in a one-dimensional box of length \( L \) with infinitely high potential walls at \( x = 0 \) and \( x = L \). The potential \( V(x) \) is given by:
+    $$
+        V(x) = \begin{cases}
+            0 & \text{for } 0 < x < L \\
+            \infty & \text{otherwise}
+        \end{cases}
+    $$
+    The time-independent Schrödinger equation inside the box is:
+    $$
+        -\frac{\hbar^2}{2m} \frac{d^2 \psi(x)}{dx^2} = E \psi(x)
+    $$
+    The general solution to this equation is:
+    $$
+        \psi(x) = A \sin(kx) + B \cos(kx)
+    $$
+    where \( k = \sqrt{\frac{2mE}{\hbar^2}} \).
+
+    Applying the boundary conditions \( \psi(0) = 0 \) and \( \psi(L) = 0 \), we find that \( B = 0 \) and \( kL = n\pi \) for \( n = 1, 2, 3, \ldots \). Thus, the allowed wave numbers are:
+    $$
+        k_n = \frac{n\pi}{L}
+    $$
+    The corresponding energy levels are quantized:
+    $$
+        E_n = \frac{\hbar^2 k_n^2}{2m} = \frac{n^2 \pi^2 \hbar^2}{2mL^2}, \quad n = 1, 2, 3, \ldots
+    $$
+    The normalized wave functions are:
+    $$
+        \psi_n(x) = \sqrt{\frac{2}{L}} \sin\left(\frac{n\pi x}{L}\right)
+    $$
+    Therefore, the stationary states of the particle in the box are given by:
+    $$        
+    \Psi_n(x,t) = \sqrt{\frac{2}{L}} \sin\left(\frac{n\pi x}{L}\right) e^{-iE_n t/\hbar}
+    $$
+\end{example}
+
+\begin{definition}[Energy Quantization]
+    \begin{itemize}
+    \item Only discrete energies $E_n$ yield normalizable solutions satisfying boundary conditions.
+    \item Examples:
+    \[
+        \begin{aligned}
+        U(x) &= 0 &\text{(infinite square well)}\\
+        U(x) &= \tfrac{1}{2}kx^2 &\text{(harmonic oscillator)}\\
+        U(r) &= -\frac{ke^2}{r} &\text{(hydrogen atom)}
+        \end{aligned}
+    \]
+\end{itemize}
+
+\end{definition}
+\subsection{Eigenstates and Observables}
+\begin{definition}[Ovservables and Operators]
+    In quantum mechanics, physical observables such as position, momentum, and energy are represented by operators that act on the wave function.
+An operator is a mathematical entity that transforms one function into another.
+For example:
+    \begin{itemize}
+        \item Position operator: \( \hat{x} = x \)
+        \item Momentum operator: \( \hat{p} = -i\hbar \frac{d}{dx} \)
+        \item Hamiltonian (energy) operator: \( \hat{H} = -\frac{\hbar^2}{2m} \frac{d^2}{dx^2} + V(x) \)
+    \end{itemize}
+\end{definition}
+\paragraph{From Schrödinger Equation to Observables}
+The time-independent Schrödinger equation (TISE) governs quantum systems:
+\begin{equation}
+-\frac{\hbar^2}{2m}\frac{d^2\psi(x)}{dx^2} + U(x)\psi(x) = E\psi(x)
+\end{equation}
+Notice, that it is an \textbf{eigenvalue equation}.
+
+\paragraph{Eigenvalue Equation}
+When an operator $\hat{A}$ acts on a special function called an \textbf{eigenfunction} $\psi(x)$, it returns the function scaled by a measurable number called an \textbf{eigenvalue} $\lambda$:
+\begin{equation}
+\hat{A}\psi(x) = \lambda\psi(x)
+\end{equation}
+
+The TISE is precisely this structure, where $\hat{H}$ (the Hamiltonian) is the operator and $E$ is the eigenvalue:
+\begin{equation}
+    \hat{H}\psi(x) = E\psi(x)
+\end{equation}
+
+\paragraph{Common Operators} Below is a table of common quantum operators:
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+\textbf{Observable} & \textbf{Operator} & \textbf{Eigenvalue} \\
+\hline
+Position & $\hat{x} = x$ & $x_0$ \\
+Momentum & $\hat{p} = -i\hbar\frac{d}{dx}$ & $p_0$ \\
+Potential Energy & $\hat{V} = V(x)$ & $V(x_0)$ \\
+Kinetic Energy & $\hat{K} = -\frac{\hbar^2}{2m}\frac{d^2}{dx^2}$ & $K$ \\
+Energy (Hamiltonian) & $\hat{H} = -\frac{\hbar^2}{2m}\frac{d^2}{dx^2} + V(x)$ & $E$ \\
+\hline
+\end{tabular}
+\end{center}
+The choice of $V(x)$ depends on the physical system being modeled:
+\begin{itemize}
+    \item Infinite square well: $V(x) = 0$ inside the well, $V(x) = \infty$ outside.
+    \item 
+    \item Harmonic oscillator: $V(x) = \frac{1}{2}kx^2$.
+    \item Hydrogen atom: $V(r) = -\frac{ke^2}{r}$.
+\end{itemize}
+\begin{definition}[Expected Value]
+In quantum mechanics, the expectation value of an observable represented by operator $\hat{A}$ is:
+\begin{equation}
+    \langle A \rangle = \int_{-\infty}^{\infty} \psi^*(x) \, \hat{A}\psi(x) \, dx
+\end{equation}
+where $\psi(x)$ is the wave function of the system, and $\psi^*(x)$ is its complex conjugate.
+The expectation value represents the average outcome of measurements of the observable $A$ on a large number of identically prepared systems.
+\end{definition}
+\begin{example}
+    A particle of mass \( m \) is confined in a one-dimensional infinite potential well of width \( L \). Its wave function is given by:
+    $$
+        \psi_n(x) = Ae^{x/L} \quad \text{for } 0 < x < L
+    $$
+    and \( \psi_n(x) = 0 \) elsewhere. Determine $\langle x \rangle$ for the particle and $\langle H \rangle$.
+
+    First, we need to normalize the wave function to find the constant \( A \):
+    $$        \int_0^L |\psi_n(x)|^2 \, dx = 1 $$
+    This gives:
+    $$        |A|^2 \int_0^L e^{2x/L} \, dx = 1 $$
+    Evaluating the integral:
+    $$        |A|^2 \left[ \frac{L}{2} e^{2x/L} \right]_0^L = 1 $$
+    $$        |A|^2 \left( \frac{L}{2} (e^2 - 1) \right) = 1 $$
+    Thus, we find:
+    $$        |A|^2 = \frac{2}{L(e^2 - 1)} \implies A = \sqrt{\frac{2}{L(e^2 - 1)}} $$
+    Now, we can calculate the expectation value of position \( \langle x \rangle \):
+    $$        \langle x \rangle = \int_0^L \psi_n^*(x) \, x \, \psi_n(x) \, dx $$
+    $$        = |A|^2 \int_0^L x e^{2x/L} \, dx $$
+    Evaluating the integral using integration by parts:
+    $$        \int x e^{2x/L} \, dx = \left(\frac{L}{2} x e^{2x/L} - \frac{L^2}{4} e^{2x/L}\right) + C $$
+    Thus,
+    $$        \langle x \rangle = |A|^2 \left[ \frac{L}{2} x e^{2x/L} - \frac{L^2}{4} e^{2x/L} \right]_0^L $$
+    $$        = |A|^2 \left( \frac{L^2}{2} e^2 - \frac{L^2}{4} e^2 + \frac{L^2}{4} \right) $$
+    $$        = |A|^2 \frac{L^2}{4} (e^2 + 1) $$
+    Substituting \( |A|^2 \):
+    $$        \langle x \rangle = \frac{2}{L(e^2 - 1)} \cdot \frac{L^2}{4} (e^2 + 1) = \frac{L(e^2 + 1)}{2(e^2 - 1)} $$
+    Next, we calculate the expectation value of the Hamiltonian \( \langle H \rangle \):
+    $$        \langle H \rangle = \int_0^L \psi_n^*(x) \, \hat{H} \, \psi_n(x) \, dx $$
+    where the Hamiltonian operator \( \hat{H} \) for a particle in an infinite potential well is:
+    $$        \hat{H} = -\frac{\hbar^2}{2m} \frac{d^2}{dx^2} $$
+    Calculating the second derivative of \( \psi_n(x) \):
+    $$        \frac{d^2 \psi_n(x)}{dx^2} = \frac{A}{L^2} e^{x/L} $$
+    Thus,
+    $$        \hat{H} \psi_n(x) = -\frac{\hbar^2}{2m} \cdot \frac{A}{L^2} e^{x/L} = -\frac{\hbar^2 A}{2m L^2} e^{x/L} $$
+    Now, we can compute \( \langle H \rangle \):
+    $$        \langle H \rangle = -\frac{\hbar^2 A^2}{2m L^2} \int_0^L e^{2x/L} \, dx $$
+    Evaluating the integral:
+    $$        \int_0^L e^{2x/L} \, dx = \left[ \frac{L}{2} e^{2x/L} \right]_0^L = \frac{L}{2} (e^2 - 1) $$
+    Thus,
+    $$        \langle H \rangle = -\frac{\hbar^2 A^2}{2m L^2} \cdot \frac{L}{2} (e^2 - 1) = -\frac{\hbar^2 A^2 (e^2 - 1)}{4m L} $$
+    Substituting \( A^2 \):
+    $$        \langle H \rangle = -\frac{\hbar^2}{4m L} \cdot \frac{2}{L(e^2 - 1)} (e^2 - 1) = -\frac{\hbar^2}{2m L^2} $$
+    Therefore, the expectation values are:
+    $$        \langle x \rangle = \frac{L(e^2 + 1)}{2(e^2 - 1)}, \quad \langle H \rangle = -\frac{\hbar^2}{2m L^2} $$
+\end{example}
 \end{document}

### 2025-11-25 YEAR2/AER210_LectureNotes/main.tex
nov 25

@@
 \end{example}
+
+\section{Flowing Fluids}
+\begin{definition}[Flowing Fluid]
+    A flowing fluid is a fluid that is in motion, characterized by the movement of its particles from one location to another.
+This motion can be caused by various factors, such as pressure differences, gravitational forces, or external forces applied to the fluid.
+Flowing fluids can exhibit different types of flow patterns, including laminar flow (smooth and orderly) and turbulent flow (chaotic and irregular).
+\end{definition}
+
+\paragraph{Eulerian and Lagrangian Descriptions} There are two main approaches to describe the motion of flowing fluids:
+\begin{definition}[Eulerian Description]
+    The Eulerian description focuses on specific locations in the fluid and examines how the fluid properties (such as velocity, pressure, and density) change over time at those fixed points.
+In this approach, the observer is stationary, and the fluid flows past them.
+We can represent the fluid properties as functions of space and time, such as:
+    \[ \vec{v} = \vec{v}(x, y, z, t) \]
+    where \(\vec{v}\) is the velocity field of the fluid.
+\end{definition}
+
+\begin{definition}[Lagrangian Description]
+    The Lagrangian description, on the other hand, follows individual fluid particles as they move through space and time.
+In this approach, the observer moves with the fluid particles, tracking their trajectories and changes in properties.
+The fluid properties are expressed as functions of the initial position of the particles and time, such as:
+    \[ \vec{r} = \vec{r}(\vec{r_0}, t) \]
+    where \(\vec{r}\) is the position of a fluid particle at time \(t\) and \(\vec{r_0}\) is its initial position.
+\end{definition}
+
+\paragraph{Flow visualization} To visualize fluid flow, we often use:
+\begin{itemize}
+    \item \textbf{Streamlines}: Lines that are tangent to the velocity vector of the fluid at every point, representing the path that a fluid particle would follow.
+    \item \textbf{Pathlines}: The actual trajectory that a fluid particle follows over time.
+    \item \textbf{Streaklines}: The locus of all fluid particles that have passed through a specific point in space over time.
+\end{itemize}
+
+\begin{definition}[Steady and Unsteady Flow]
+    A flow is said to be steady if the fluid properties at any given point do not change over time.
+Mathematically, this means that the partial derivative of any fluid property with respect to time is zero:
+    \[
+    \frac{\partial \phi}{\partial t} = 0
+    \]
+    where \(\phi\) represents any fluid property (such as velocity, pressure, or density).
+
+    Conversely, a flow is unsteady if the fluid properties at a given point do change over time, meaning:
+    \[
+    \frac{\partial \phi}{\partial t} \neq 0
+    \]    
+\end{definition}
+
+\begin{definition}[Turbulent and Laminar Flow]
+    Laminar flow is characterized by smooth, orderly fluid motion in which fluid particles move in parallel layers with minimal mixing between them.
+In laminar flow, the velocity of the fluid is relatively low, and the flow is dominated by viscous forces.
+
+    Turbulent flow, on the other hand, is characterized by chaotic, irregular fluid motion with significant mixing between fluid layers.
+In turbulent flow, the velocity of the fluid is relatively high, and inertial forces dominate over viscous forces.
+This results in eddies, vortices, and fluctuations in velocity and pressure throughout the fluid.
+\end{definition}
+
+\paragraph{Viscous versus Inviscid (Nonviscous) Regions of Flow} In fluid dynamics, the distinction between viscous and inviscid (nonviscous) regions of flow is important for understanding the behavior of fluids under different conditions:
+
+\subsection{Mass and Energy Conservation}
+\begin{definition}[Volume Flow Rate]
+    The volume flow rate \(Q\) is defined as the volume of fluid passing through a given cross-sectional area per unit time. Mathematically, it is expressed as:
+    \begin{equation}
+        Q = \iint_A \vec{v} \cdot \vec{n} \, dA
+    \end{equation}
+    where \(\vec{v}\) is the velocity vector of the fluid, \(\vec{n}\) is the unit normal vector to the cross-sectional area \(A\), and \(dA\) is the infinitesimal area element.
+\end{definition}
+
+\begin{definition}[Mass Flow Rate]
+    The mass flow rate \(\dot{m}\) is defined as the mass of fluid passing through a given cross-sectional area per unit time. Mathematically, it is expressed as:
+    \begin{equation}
+        \dot{m} = \iint_A \rho \vec{v} \cdot \vec{n} \, dA
+    \end{equation}
+    where \(\rho\) is the fluid density, \(\vec{v}\) is the velocity vector of the fluid, \(\vec{n}\) is the unit normal vector to the cross-sectional area \(A\), and \(dA\) is the infinitesimal area element.
+\end{definition}
+
+\paragraph{Conservation of Mass along a Streamtube} For an incompressible fluid flowing through a streamtube, the volume flow rate must be constant along the length of the streamtube.
+This means that the volume flow rate at any cross-section of the streamtube is equal to the volume flow rate at any other cross-section.
+Mathematically, this can be expressed as:
+\[
+Q_1 = Q_2
+\]
+where \(Q_1\) and \(Q_2\) are the volume flow rates
+
 \end{document}

### 2025-11-25 YEAR2/MAT292_LectureNotes/main.tex
nov 25

@@
 
-\newpage
+\subsection{Convolutions}
+
+\begin{definition}[Convolution]
+    Convolution is a mathematical operation that combines two functions to produce a third function, expressing how the shape of one is modified by the other.
+In the context of Laplace transforms, the convolution of two functions \( f(t) \) and \( g(t) \) is defined as:
+    \begin{equation}
+    (f * g)(t) = \int_0^t f(\tau) g(t - \tau) d\tau
+    \end{equation}
+    where \( \tau \) is a dummy variable of integration. The convolution operation is commutative, meaning that \( f * g = g * f \).
+
+    In the context of Laplace transforms, the convolution theorem states that the Laplace transform of the convolution of two functions is equal to the product of their individual Laplace transforms:
+    \begin{equation}
+    \mathcal{L}\{f * g\}(s) = F(s) G(s)
+    \end{equation}
+    where \( F(s) = \mathcal{L}\{f(t)\} \) and \( G(s) = \mathcal{L}\{g(t)\} \).
+\end{definition}
+
+\begin{example}
+    Consider the following IVP:
+    $$ a y'' + b y' + c y = g(t), \quad y(0)=0,\quad y'(0)=0 $$
+
+    We can represent this system using the transfer function \(H(s)\):
+    $$ H(s) = \frac{Y(s)}{G(s)} = \frac{1}{as^2 + bs + c}. $$
+
+    Thus,
+    $$ 
+    Y(s) =  H(s)\big[(a s + b) y_0 + a y_1\big]  \;+\;H(s)G(s).
+    $$
+
+    Taking the inverse Laplace transform and applying the convolution theorem:
+    $$
+    y(t) = 
+    \mathcal{L}^{-1}\!\left\{
+       {H(s)\big[(a s + b) y_0 + a y_1\big]}
+    \right\}
+    \;+\;
+    \mathcal{L}^{-1}\{H(s) G(s)\}
+    $$
+    $$
+    y(t) =
+     \underbrace{\mathcal{L}^{-1}\{H(s)[(a s + b) y_0 + a y_1]\}}_{\text{free response}}
+    \;+\;
+    \underbrace{(h * g)(t).}_{\text{forced response}}
+    $$
+
+\end{example}
+
+\begin{definition}[Forced Response]
+    In the context of differential equations and system dynamics, the forced response refers to the part of the solution that arises due to an external input or forcing function acting on the system.
+It represents how the system responds to external stimuli over time.
+
+    Mathematically, for a linear time-invariant (LTI) system described by a differential equation, the total response \( y(t) \) can be expressed as the sum of the free response (or homogeneous solution) and the forced response (or particular solution):
+    \begin{equation}
+    y(t) = y_{\text{free}}(t) + y_{\text{forced}}(t)
+    \end{equation}
+    where \( y_{\text{free}}(t) \) is determined by the system's initial conditions and natural dynamics, while \( y_{\text{forced}}(t) \) is determined by the external input \( g(t) \).
+
+    The forced response can often be computed using techniques such as the method of undetermined coefficients, variation of parameters, or convolution with the system's impulse response.
+\end{definition}
+
+\begin{definition}[Free Response]
+    In the context of differential equations and system dynamics, the free response refers to the part of the solution that arises solely from the system's initial conditions, without any external input or forcing function.
+It represents how the system behaves naturally over time based on its inherent properties.
+
+    Mathematically, for a linear time-invariant (LTI) system described by a differential equation, the total response \( y(t) \) can be expressed as the sum of the free response (or homogeneous solution) and the forced response (or particular solution):
+    \begin{equation}
+    y(t) = y_{\text{free}}(t) + y_{\text{forced}}(t)
+    \end{equation}
+    where \( y_{\text{free}}(t) \) is determined by the system's initial conditions and natural dynamics, while \( y_{\text{forced}}(t) \) is determined by any external input \( g(t) \).
+
+    The free response is typically found by solving the homogeneous version of the differential equation, which involves setting the external input to zero.
+\end{definition}
+
+\begin{example}
+    % use e i t = cis t for convulution example
+    Consider the following IVP:
+    $$ y '' + 2y ' + 5y = t, \quad y(0) = 0, \quad y'(0) = 0 $$
+    Taking the Laplace transform of both sides gives:
+    $$ \mathcal{L}\{y''\} + 2\mathcal{L}\{y'\} + 5\mathcal{L}\{y\} = \mathcal{L}\{t\} $$
+    Using the properties of the Laplace transform, we have:
+    $$(s^2Y(s) - sy(0) - y'(0)) + 2(sY(s) - y(0)) + 5Y(s) = \frac{1}{s^2} = G(s)$$
+    So for the transfer function:
+    $$ H(s) = \frac{Y(s)}{G(s)} = \frac{1}{s^2 + 2s + 5} $$
+    Thus,
+    $$ 
+    Y(s) =  H(s)\big[(s) y_0 + y_1\big]  \;+\;H(s)G(s).
+    $$
+    Taking the inverse Laplace transform and applying the convolution theorem:  
+    $$
+    y(t) = 
+    \mathcal{L}^{-1}\!\left\{
+       {H(s)\big[(s) y_0 + y_1\big]} 
+    \right\}
+    \;+\;
+    \mathcal{L}^{-1}\{H(s) G(s)\}
+    $$
+    For the forced response, we repsonse, we can evaulte the convultion. 
+    \begin{align*}
+    y(t) &= (h * g)(t) \\
+    &= \int_0^t h(\tau) g(t - \tau) d\tau \\
+    &= \int_0^t e^{-\tau} \sin(2\tau) (t - \tau) d\tau \\
+    \end{align*}
+        We use the euler's formula to simplify the integral:
+
+    \begin{align*}
+        y(t)
+        &= \mathrm{Re}\!\left\{ 
+        \frac{1}{2i}\!\left[
+        \int_0^t (t-\tau)e^{(-1+i2)\tau} d\tau
+        -
+        \int_0^t (t-\tau)e^{(-1-i2)\tau} d\tau
+        \right]\right\} \\[4pt]
+        &= -\mathrm{Im}\!\left\{ 
+        \int_0^t (t-\tau)e^{(-1+i2)\tau} d\tau +
+        \int_0^t (t-\tau)e^{(-1-i2)\tau} d\tau
+        \right\} \\[4pt]
+        &= \frac{1}{5}
+        - \frac{e^{-t}}{25} 
+        \Big[(5-2t)\cos(2t) + (2+t)\sin(2t)\Big].
+    \end{align*}
+
+\end{example}
+
+\section{Fourier Transform}
+\begin{definition}[Fourier Transform]
+    The Fourier transform is an integral transform that converts a time-domain function \( f(t) \) into a frequency-domain function \( F(\omega) \). It is defined as:
+    \begin{equation}
+    \mathcal{F}\{f(t)\} = F(\omega) = \int_{-\infty}^{\infty} f(t) e^{-i\omega t} dt
+    \end{equation}
+    where \( \omega \) is the angular frequency in radians per second.
+The Fourier transform is widely used in signal processing, physics, and engineering to analyze the frequency content of signals and systems.
+
+    The inverse Fourier transform allows us to recover the original time-domain function from its frequency-domain representation:
+    \begin{equation}
+    f(t) = \mathcal{F}^{-1}\{F(\omega)\} = \frac{1}{2\pi} \int_{-\infty}^{\infty} F(\omega) e^{i\omega t} d\omega
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Orthonormal Family of Functions]
+    Let $\mathrm{PC}[a,b]$ be the vector space of piecewise continuous functions on the interval $[a,b] \subset \mathbb{R}$.
+A family of functions $\{ \phi_n(t) \}_{n=0}^{\infty} \subset \mathrm{PC}[a,b]$ is said to be orthonormal on $[a,b]$ if it satisfies the following two conditions:
+    \begin{enumerate}
+        \item \textbf{Orthogonality}: For any two distinct functions $\phi_m(t)$ and $\phi_n(t)$ in the family (where $m \neq n$), the inner product is zero:
+        \begin{equation}
+        \langle \phi_m, \phi_n \rangle = \int_a^b \phi_m(t) \phi_n(t) dt = 0
+        \end{equation}
+        \item \textbf{Normalization}: Each function $\phi_n(t)$ has a unit norm, meaning that the inner product of the function with itself is equal to one:
+        \begin{equation}
+        \langle \phi_n, \phi_n \rangle = \int_a^b \phi_n(t)^2 dt = 1
+        \end{equation}
+    \end{enumerate}
+    If both conditions are satisfied, the family $\{ \phi_n(t) \}_{n=0}^{\infty}$ is called an orthonormal family of functions on the interval $[a,b]$.
+For example, the basis of wave function $\psi_n$ of a partiple in a infinite potential well is an orthonormal family on the interval $[0,L]$.
+\end{definition}
+
+\begin{example}
+    Cosndier the following set on the interval \( [-L, L] \):
+    $$ \left\{ \sqrt{\frac{2}{L}} \frac{1}{2} , \sqrt{\frac{1}{L}} \cos{\left(\frac{n \pi t}{L}\right)}, \sqrt{\frac{1}{L}} \sin{\left(\frac{n \pi t}{L}\right)} \right\}_{n=1}^{\infty} $$
+    This set forms an orthonormal family on the interval \( [-L, L] \). To verify this, we need to check both orthogonality and normalization conditions.
+
+    \begin{enumerate}
+        \item \textbf{Orthogonality}:
+        For any two distinct functions \( \phi_m(t) \) and \( \phi_n(t) \) in the set, we need to show that their inner product is zero:
+        \begin{equation*}
+        \langle \phi_m, \phi_n \rangle = \int_{-L}^L \phi_m(t) \phi_n(t) dt = 0
+        \end{equation*}
+        This can be verified using trigonometric identities and properties of integrals.
+First, we note that the integral of the product of sine and cosine functions over a full period is zero.
+Specifically, for \( m \neq n \):
+
+        Additionally, the constant function is orthogonal to both sine and cosine functions over the interval       
+        \item \textbf{Normalization}:
+        For each function \( \phi_n(t) \) in the set, we need to show that its inner product with itself is equal to one:
+        \begin{equation*}
+        \langle \phi_n, \phi_n \rangle = \int_{-L}^L \phi_n(t)^2 dt = 1
+        \end{equation*}
+        This can also be verified through direct integration.
+    \end{enumerate}
+
+\end{example}
+
+\begin{theorem}[Fourier Theorem]
+    Suppose $f$ is periodic with period $2L$ and; $f$, $f'$ are piecewise continuous on $[-L,L]$. Then $f$ can be represented as a Fourier series:
+    \begin{equation}
+    f(t) = \frac{a_0}{2} + \sum_{n=1}^{\infty} \left[ a_n \cos\left(\frac{n \pi t}{L}\right) + b_n \sin\left(\frac{n \pi t}{L}\right) \right]
+    \end{equation}
+    where the Fourier coefficients \( a_n \) and \( b_n \) are given by:
+    \begin{subequations}
+
+    \begin{equation}
+    a_n = \langle f(t), \sqrt{\frac{1}{L}} \cos\left(\frac{n \pi t}{L}\right) \rangle = \frac{1}{L} \int_{-L}^{L} f(t) \cos\left(\frac{n \pi t}{L}\right) dt
+    \end{equation}
+    \begin{equation}
+    b_n = \langle f(t), \sqrt{\frac{1}{L}} \sin\left(\frac{n \pi t}{L}\right) \rangle = \frac{1}{L} \int_{-L}^{L} f(t) \sin\left(\frac{n \pi t}{L}\right) dt
+    \end{equation}
+        \end{subequations}
+\end{theorem}
+\section{Partial Differential Equations}
+\begin{definition}[Partial Differential Equation]
+    A partial differential equation (PDE) is a mathematical equation that involves multiple independent variables, their partial derivatives, and an unknown function.
+PDEs are used to describe a wide range of physical phenomena, including heat conduction, wave propagation, fluid dynamics, and quantum mechanics.
+
+    A general form of a PDE can be expressed as:
+    \begin{equation}
+    F\left(x_1, x_2, \ldots, x_n, u, \frac{\partial u}{\partial x_1}, \frac{\partial u}{\partial x_2}, \ldots, \frac{\partial u}{\partial x_n}, \frac{\partial^2 u}{\partial x_1^2}, \ldots\right) = 0
+    \end{equation}
+    where \( u = u(x_1, x_2, \ldots, x_n) \) is the unknown function of the independent variables \( x_1, x_2, \ldots, x_n \), and \( F \) is a given function that relates these variables and their partial derivatives.
+
+    PDEs can be classified based on their order (the highest order of partial derivatives present), linearity (linear or nonlinear), and the number of independent variables involved.
+Solving PDEs often requires specialized techniques such as separation of variables, method of characteristics, or numerical methods.
+\end{definition}
+
+\begin{definition}[Boundry Value Problem]
+    A boundary value problem (BVP) is a type of differential equation problem characterized by a PDE and specified initial time condition and boundary conditions on the spatial domain.
+In a BVP, the solution is sought for a function \( u(x, t) \) that satisfies the PDE within a defined spatial domain and adheres to given conditions at the boundaries of that domain.
+
+    A general form of a BVP can be expressed as:
+    \begin{equation}
+    \begin{cases}
+    \frac{\partial u}{\partial t} = \mathcal{L}u + f(x, t), & x \in \Omega, \; t > 0 \\
+    u(x, 0) = u_0(x), & x \in \Omega \\
+    B[u] = g(x, t), & x \in \partial\Omega, \; t > 0
+    \end{cases}
+    \end{equation}
+    where \( \mathcal{L} \) is a differential operator, \( f(x, t) \) is a source term, \( u_0(x) \) is the initial condition, \( B[u] \) represents the boundary conditions applied on the boundary \( \partial\Omega \) of the spatial domain \( \Omega \), and \( g(x, t) \) is a function defining the boundary conditions.
+    BVPs are commonly encountered in various fields of science and engineering, such as heat conduction, fluid dynamics, and electromagnetics, where the behavior of a system is influenced by both its internal dynamics and external constraints.
+\end{definition}
+
+\paragraph{Using Fourier Transform} We can transform PDEs into a algebraic domain using the Fourier transform, similar to how we used the Laplace transform for ODEs.
+This allows us to solve the PDE in the frequency domain and then transform back to the spatial domain using the inverse Fourier transform.
+\paragraph{Seperabilty Assumption} Many PDEs can be solved using the method of separation of variables, which assumes that the solution can be expressed as a product of functions, each depending on a single independent variable.
+This method reduces the PDE into a set of ODEs that can be solved individually.
+We assume that the solution can be written as:
+$$ u(x,t) = X(x)T(t) $$
+Substituting this into the PDE and separating variables allows us to derive ODEs for \( X(x) \) and \( T(t) \).
+
+\begin{example}
+    GIven the following PDE:
+    $$\begin{cases}
+         u_t = \alpha^2 u_{xx}, & 0 < x < L, \; t > 0 \\
+         u(0,t) = 0 = u(L,t), & t \geq 0 \\
+         u(x,0) = f(x), & 0 \leq x \leq L
+    \end{cases}$$
+
+    Assume a separable solution of the form:
+    $$ u(x,t) = X(x)T(t) $$
+    Substituting into the PDE gives:
+    $$ X(x) T'(t) = \alpha^2 X''(x) T(t) $$
+    Dividing both sides by \( \alpha^2 X(x) T(t) \) gives:
+    $$ \frac{T'(t)}{\alpha^2 T(t)} = \frac{X''(x)}{X(x)} = -\lambda $$
+    This leads to two ODEs:
+    \begin{align*}
+        T'(t) + \alpha^2 \lambda T(t) &= 0 \\
+        X''(x) + \lambda X(x) &= 0
+    \end{align*}
+
+    For the spatial part, we have the boundary conditions:
+    \begin{align*}
+        X(0) &= 0 \\
+        X(L) &= 0
+    \end{align*}
+    The general solution for \( X(x) \) is:
+    $$ X(x) = A \sin(\sqrt{\lambda} x) + B \cos(\sqrt{\lambda} x) $$
+    Applying the boundary conditions leads to:
+    \begin{align*}
+        B &= 0 \\
+        \sqrt{\lambda} L &= n \pi, \quad n = 1, 2, 3, \ldots
+    \end{align*}
+    Thus, the eigenvalues are:
+    $$ \lambda_n = \left( \frac{n \pi}{L} \right)^2 $$
+    and the corresponding eigenfunctions are:
+    $$ X_n(x) = A_n \sin\left( \frac{n \pi x}{L} \right) $$
+
+    For the temporal part, the solution is:
+    $$ T_n(t) = C_n e^{-\alpha^2 \lambda_n t} = C_n \exp\left( -\alpha^2 \left( \frac{n \pi}{L} \right)^2 t \right) $$
+
+    So the fundamental solutions are:
+    $$ u_n(x,t) = A_n C_n \sin\left( \frac{n \pi x}{L} \right) \exp\left( -\alpha^2 \left( \frac{n \pi}{L} \right)^2 t \right) $$
+
+    The general solution is a superposition of all fundamental solutions:
+    $$ u(x,t) = \sum_{n=1}^{\infty} a_n \sin\left( \frac{n \pi x}{L} \right) \exp\left( -\alpha^2 \left( \frac{n \pi}{L} \right)^2 t \right) $$ 
+    where \( a_n = A_n C_n \).
+
+    To determine the coefficients \( a_n \), we use the initial condition:
+    $$ u(x,0) = f(x) = \sum_{n=1}^{\infty} a_n \sin\left( \frac{n \pi x}{L} \right) $$
+    This is a Fourier sine series expansion of \( f(x) \) on the interval \( [0, L] \). The coefficients \( a_n \) are given by:
+    $$ a_n = \frac{2}{L} \int_0^L f(x) \sin\left( \frac{n \pi x}{L} \right) dx $$
+\end{example}
 \appendix 

### 2025-11-25 YEAR2/PHY293_LectureNotes/main.tex
nov 25

@@
 \end{example}
+
+\subsection{Wave functions}
+\begin{definition}[Properties of wave functions]
+    A valid wave function \( \psi(x,t) \) must satisfy the following properties:
+    \begin{itemize}
+        \item \textbf{Normalization}: The total probability of finding the particle in all space must be 1:
+        \begin{equation}
+            \int_{-\infty}^{\infty} |\psi(x,t)|^2 \, dx = 1
+        \end{equation}
+        \item \textbf{Continuity}: The wave function and its first derivative must be continuous everywhere, except at points where the potential is infinite.
+        \item \textbf{Single-valuedness}: The wave function must have a unique value at each point in space and time.
+        \item \textbf{Square Integrability}: The integral of the square of the wave function over all space must be finite:
+        \begin{equation}
+            \int_{-\infty}^{\infty} |\psi(x,t)|^2 \, dx < \infty
+        \end{equation}
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Energy well]
+    An energy well is a potential energy profile that confines a particle within a specific region of space.
+It is characterized by a lower potential energy inside the well compared to the surrounding areas.
+Energy wells can be classified into different types based on their shape and depth, such as infinite square wells, finite square wells, and harmonic oscillator potentials.
+The quantization of energy levels arises from the boundary conditions imposed by the well, leading to discrete allowed energy states for the confined particle.
+\end{definition}
+
+\begin{definition}[Strödinger Equation - Time Independent]
+    The time-independent Schrödinger equation (TISE) describes the stationary states of a quantum system. It is given by:
+    \begin{equation}
+        -\frac{\hbar^2}{2m} \frac{d^2 \psi(x)}{dx^2} + V(x) \psi(x) = E \psi(x)
+    \end{equation}
+    where \( \psi(x) \) is the spatial part of the wave function, \( m \) is the mass of the particle, \( V(x) \) is the potential energy as a function of position, and \( E \) is the total energy of the particle.
+The TISE is derived from the time-dependent Schrödinger equation by separating variables and assuming a solution of the form \( \Psi(x,t) = \psi(x)e^{-iEt/\hbar} \).
+\end{definition}
+
+\begin{definition}[Particle in a One-Dimensional Infinite Potential Well]
+    A particle confined in a one-dimensional infinite potential well (also known as a "particle in a box") has quantized energy levels and corresponding wave functions.
+The potential \( V(x) \) is defined as:
+    \begin{equation}
+        V(x) = \begin{cases}
+            0 & \text{for } 0 < x < L \\
+            \infty & \text{otherwise}
+        \end{cases}
+    \end{equation}
+    The allowed energy levels are given by:
+    \begin{equation}
+        E_n = \frac{n^2 \pi^2 \hbar^2}{2mL^2}, \quad n = 1, 2, 3, \ldots
+    \end{equation}
+    The corresponding normalized wave functions are:
+    \begin{equation}
+        \psi_n(x) = \sqrt{\frac{2}{L}} \sin\left(\frac{n\pi x}{L}\right), \quad 0 < x < L
+    \end{equation}
+    and \( \psi_n(x) = 0 \) elsewhere.
+
+    The probability density of finding the particle are the square of the constnat in the wave function:
+    \begin{equation}
+        |\psi_n(x)|^2 = \frac{2}{L} c_n^2 \sin^2\left(\frac{n\pi x}{L}\right)
+    \end{equation}
+\end{definition}
+
+\begin{definition}[Particle in a Finite Potential Well]
+    A particle confined in a one-dimensional finite potential well has quantized energy levels and corresponding wave functions. The potential \( V(x) \) is defined as:
+    \begin{equation}
+        V(x) = \begin{cases}
+            0 & \text{for } 0 < x < L \\
+            V_0 & \text{otherwise}
+        \end{cases}
+    \end{equation}
+    where \( V_0 \) is a finite positive constant representing the potential outside the well.
+The allowed energy levels are determined by solving the time-independent Schrödinger equation with appropriate boundary conditions.
+The wave functions inside and outside the well must be continuous and differentiable at the boundaries \( x = 0 \) and \( x = L \).
+The energy levels are quantized, but unlike the infinite well, they depend on the depth \( V_0 \) of the well.
+
+\end{definition}
+
+\begin{definition}[Quantum Tunneling]
+    Quantum tunneling is a phenomenon where a particle has a non-zero probability of crossing a potential barrier, even if its energy is less than the height of the barrier.
+This occurs due to the wave-like nature of particles in quantum mechanics, allowing the wave function to extend into and beyond the barrier region.
+The probability of tunneling depends on factors such as the width and height of the barrier, as well as the energy of the particle.
+Quantum tunneling has important applications in various fields, including nuclear fusion, semiconductor physics, and scanning tunneling microscopy.
+
+    This is a result of the wave function penetrating into the classically forbidden region, leading to a finite probability of finding the particle on the other side of the barrier.
+\end{definition}
+
+\begin{definition}[Timed Case]
+    For each wave funtion, we can add the time dependent part:
+    \begin{equation}    
+        \Psi_n(x,t) = \psi_n(x) e^{-iE_n t/\hbar}
+    \end{equation}
+    where \( \Psi_n(x,t) \) is the full time-dependent wave function, \( \psi_n(x) \) is the spatial part of the wave function, \( E_n \) is the energy of the state, and \( \hbar \) is the reduced Planck's constant.
+\end{definition}
+
+\section{Relativity}
+\begin{definition}[Postulates of Special Relativity]
+    \begin{itemize}
+        \item The laws of physics are the same in all inertial frames of reference.
+        \item The speed of light in a vacuum is constant and independent of the motion of the light source or observer.
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Lorentz Transformation]
+    The Lorentz transformation relates the space and time coordinates of events as observed in two inertial frames of reference moving at a constant velocity relative to each other.
+If one frame \( S' \) is moving at a velocity \( v \) relative to another frame \( S \), the transformations are given by:
+    \begin{equation}
+        x' = \gamma (x - vt)
+    \end{equation}
+    \begin{equation}
+        t' = \gamma \left( t - \frac{vx}{c^2} \right)
+    \end{equation}
+    where \( \gamma = \frac{1}{\sqrt{1 - \frac{v^2}{c^2}}} \) is the Lorentz factor, and \( c \) is the speed of light in a vacuum.
+
+    These are observable when the relative velocity \( v \) approaches the speed of light \( c \), leading to significant effects such as time dilation and length contraction.
+    \begin{itemize}
+        \item \textbf{Faster Object Sees Slower Clocks}: A moving clock runs slower when observed from a stationary frame.
+For a fast moving object, the symmetry of time dilation means each observer sees the other's clock running slow.
+It breaks when we goes back to the original frame.
+A fast travelling would be younger than the one who stayed.
+        \item \textbf{Faster Object Sees Shorter Lengths}: An object moving at a high speed appears shorter in the direction of motion when observed from a stationary frame.
+This effect is known as length contraction.
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Simultaneity]
+    Simultaneity is the concept that two events occurring at the same time in one frame of reference may not occur at the same time in another frame of reference moving relative to the first.
+This is a direct consequence of the relativity of time in special relativity.
+The lack of absolute simultaneity means that observers in different inertial frames may disagree on the timing of events, leading to different interpretations of cause and effect.
+
+    Imagine two car crashes occuring at the same time in the Eath Frame, one is London one in New York.
+An observer on a plane flying from London to New York would see the crash in London happen first, while an observer on a plane flying from New York to London would see the crash in New York happen first.
+This illustrates how simultaneity is relative and depends on the observer's frame of reference.
+\end{definition}
+
 \end{document}
